---
title: "Yuba River Redd Survey QC"
author: "Maddee Rubenson"
date: "08/30/2022"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Yuba River Redd Surveys

## Description of Monitoring Data

**Timeframe:** 

2011-09-19 through 2021-04-20

**Completeness of Record throughout timeframe:** 
Data was collected primarily during October - December. The number of samples per month is summarized below. 

- January: 75
- February: 34
- March: 40
- April: 15
- May: 0
- June: 0
- July: 0 
- August: 0
- September: 174
- October: 759
- November: 639
- December: 188

**Sampling Location:**

Throughout the Yuba River, identified by unique `redd_id` and `latitude`/`longitude`

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, include = F, echo = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/yuba-river/data-raw/yuba_Redd data request Aug2022.xlsx",
  bucket = gcs_get_global_bucket(),
  saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/yuba-river/yuba_Redd_data_request_Aug2022.xlsx",
  overwrite = TRUE
)

gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/yuba-river/data-raw/2017Redd data request Aug2022.xlsx",
  bucket = gcs_get_global_bucket(),
  saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/yuba-river/2017Redd_data_request_Aug2022.xlsx",
  overwrite = TRUE
)
```


```{r, eval=TRUE}
sheets <- readxl::excel_sheets(here::here('data-raw', 'qc-markdowns', 'adult-holding-redd-and-carcass-surveys', 'yuba-river/yuba_Redd_data_request_Aug2022.xlsx'))
raw_yuba_redd <- readxl::read_excel(here::here('data-raw', 'qc-markdowns', 'adult-holding-redd-and-carcass-surveys', 'yuba-river/yuba_Redd_data_request_Aug2022.xlsx'))
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r include=FALSE}
# read in data to clean 

sheets %>%
  purrr::map(function(sheet){ # iterate through each sheet name
  assign(x = paste0('redd_',sheet),
         value = readxl::read_xlsx(path = here::here('data-raw', 'qc-markdowns', 'adult-holding-redd-and-carcass-surveys', 'yuba-river/yuba_Redd_data_request_Aug2022.xlsx'), sheet = sheet) %>% janitor::clean_names(),
         envir = .GlobalEnv)
})

redd_2017 <- readxl::read_xlsx(here::here("data-raw","qc-markdowns", "adult-holding-redd-and-carcass-surveys", "yuba-river", "2017Redd_data_request_Aug2022.xlsx")) %>% janitor::clean_names()
```

## Data transformations

```{r}

clean_redd_2008 <- redd_2008 %>% 
  mutate(across(where(is.character), str_remove_all, pattern = fixed(" "))) %>%
  mutate(latitude = as.numeric(latitude),
         longitude = as.numeric(longitude),
         latitude = latitude/10000,
         longitude = longitude/10000 * -1,
         year = 2008) 

clean_redd_2009 <- redd_2009 %>% 
  mutate(year = 2009) 

clean_redd_2010 <- redd_2010 %>% 
  select(-c(pl:x80_percent_depth, imposition_potential:percent_silt_clay_1, morphological_unit)) %>%
  rename(redd_id = redd_i_d) %>%
  mutate(year = 2010) 

clean_redd_2011 <- redd_2011 %>% 
  rename(redd_id = redd_i_d) %>%
  select(-c(imposition_potential, distance_to_cover, cover_type, 
           comments:tail_spill_width_2, nose_velocity_0_15m)) %>%
  mutate(year = 2011) %>%
  filter(year(date) < 2012) # removes a few dates that are formatted incorrectly 

clean_redd_2012 <- redd_2012 %>% 
  rename(redd_id = redd_i_d) %>%
  select(-c(crew_a_or, num_of_fis:tail_spi_2, adcp_veloc:comments2)) %>%
  mutate(year = 2012) 

clean_redd_2013 <- redd_2013 %>% 
  rename(redd_id = redd_i_d) %>%
  select(-c(imposition_potential:tail_spill_width_2)) %>%
  mutate(year = 2013) 

clean_redd_2014 <- redd_2014 %>% 
  rename(redd_id = redd_i_d) %>%
  select(-c(imposition_potential:tail_spill_width_2)) %>%
  mutate(year = 2014)

clean_redd_2015 <- redd_2015 %>% 
  rename(redd_id = redd_i_d,
         depth_m = depth) %>%
  mutate(year = 2015) %>%
  select(names(clean_redd_2014)) 

clean_redd_2016 <- redd_2016 %>% 
  rename(redd_id = redd_i_d,
         depth_m = depth) %>%
  mutate(year = 2016) %>%
  select(names(clean_redd_2014)) 

clean_redd_2017 <- redd_2017 %>% 
  rename(redd_id = redd_i_d,
         depth_m = depth) %>%
  mutate(year = 2017) %>%
  select(names(clean_redd_2014))
  

clean_redd_2018 <- redd_2018 %>% 
  rename(redd_id = redd_i_d,
         depth_m = depth) %>%
    mutate(year = 2018) %>%
  select(names(clean_redd_2014)) 

clean_redd_2019 <- redd_2019 %>% 
  rename(redd_id = redd_i_d,
         depth_m = depth) %>%
    mutate(year = 2019) %>%
  select(names(clean_redd_2014)) 

clean_redd_2020 <- redd_2020 %>% 
  rename(redd_id = redd_i_d,
         depth_m = depth) %>%
    mutate(year = 2020) %>%
  select(names(clean_redd_2014)) 

all_yuba_redd <- bind_rows(clean_redd_2008, 
                           clean_redd_2009,
                           clean_redd_2010, 
                           clean_redd_2011, 
                           clean_redd_2012, 
                           clean_redd_2013, 
                           clean_redd_2014, 
                           clean_redd_2015, 
                           clean_redd_2016, 
                           clean_redd_2017,
                           clean_redd_2018, 
                           clean_redd_2019, 
                           clean_redd_2020) %>%
  # take average velocity 
  mutate(mean_velocity = rowMeans(.[grep("^vel", names(.))], na.rm = TRUE)) %>% 
  select(-c(velocity_20_depth:velocity_80_percent_depth)) %>% 
  rename(velocity = mean_velocity) %>% 
  mutate_if(is.numeric, funs(ifelse(is.nan(.), NA, .))) %>%
  filter(velocity < 10) # single outlier value 

knitr::kable(head(all_yuba_redd, 15), digits = 2) 
```


## Explore Numeric Variables: {.tabset}

### Variable: 

**Plotting latitude/longitude over Period of Record**

```{r}
length(unique(all_yuba_redd$latitude))
length(unique(all_yuba_redd$longitude))

length(all_yuba_redd[is.na(all_yuba_redd$latitude), ]) # 9 NAs
length(all_yuba_redd[is.na(all_yuba_redd$longitude), ]) # 9 NAs
```

**Numeric Summary of year over Period of Record**

```{r}
table(all_yuba_redd$year)
```

**Numeric Summary of depth_m**
```{r}
ggplot(data = all_yuba_redd) + 
  geom_point(aes(x = date, y = depth_m))

```


**Numeric Summary of velocity**
```{r}
ggplot(data = all_yuba_redd) + 
  geom_point(aes(x = date, y = velocity))

```

**Numeric Summary of num_of_fish_on_redd**
```{r}
ggplot(data = all_yuba_redd) + 
  geom_jitter(aes(x = date, y = num_of_fish_on_redd))

```
**NA and Unknown Values**

Provide a stat on NA or unknown values

## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
all_yuba_redd[is.na(all_yuba_redd$species), ] # no NAs for species
all_yuba_redd[is.na(all_yuba_redd$redd_id), ] # no NAs for redd_id
all_yuba_redd[is.na(all_yuba_redd$date), ] # no NAs for date
```


### Variable: `[name]`
```{r}
table(all_yuba_redd$species)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

**Create lookup rda for [variable] encoding:** 
```{r}
save(all_yuba_redd, file = here::here('data/redd-carcass-holding/yuba_redd.rda'))
```

**NA and Unknown Values**

## Summary of identified issues


## Save cleaned data back to google cloud 

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_yuba_redd,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/yuba-river/data/2yuba_redd.csv",
           predefinedAcl = "bucketLevel")
```
