---
title: "Yuba River Carcass QC"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)

google_cloud_access <- FALSE

```

# Yuba River Carcass Data

## Description of Monitoring Data

**Timeframe:**

Min date: `2009-09-08` Max date: `2022-01-26`

**Video Season:**

**Completeness of Record throughout timeframe:**

**Sampling Location:**

Sampling location is based on river mile.

**Data Contact:**

Jessica Nichols [Escapement Protocols](/data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/yuba-river/Escapement%20Protocol%202011%20-%20internal%20review%20draft_09-16-2011.docx)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# Open object from google cloud storage

if(google_cloud_access == TRUE) {
  gcs_auth(json_file = Sys.getenv('GCS_AUTH_FILE'))
  gcs_global_bucket(bucket = Sys.getenv('GCS_DEFAULT_BUCKET'))
  
  gcs_list_objects()
  # git data and save as xlsx
  gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/yuba-river/data-raw/Carcass_datarequest07122022.xlsx",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/yuba-river/raw_carcass_yuba_river.xlsx",
                 overwrite = TRUE)
}
```

Read in data from google cloud, glimpse raw data and domain description sheet:
- there are two tabs in the excel file: `Carcass`, which contains biological sampling (one individual per row), and `Chops`, which contains bulk chop counts that were not biologically sampled (one row is one date, and count column is total chops that day).

```{r}
# read in data to clean 
excel_sheets("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/yuba-river/raw_carcass_yuba_river.xlsx") 
raw_carcass_individual <- read_excel("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/yuba-river/raw_carcass_yuba_river.xlsx", 
                               sheet = "Carcass") %>% glimpse()
raw_carcass_counts <- read_excel("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/yuba-river/raw_carcass_yuba_river.xlsx", 
                                 sheet = "Chop") |> glimpse()

```

## Data transformations

### `Carcass` table
- modified `survey_id` to `date` and converted to date format
- added `individual_chops` column (`1` for all rows)

```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns

clean_yuba_carcass <- raw_carcass_individual %>% janitor::clean_names() %>% 
  rename(river_mile = rm) %>%
  mutate(date = as_date(survey_id),
         individual_chops = 1) %>% 
  select(-survey_id) %>%
  glimpse()

summary(clean_yuba_carcass)
```
### `Chop` table
- modified `survey_id` to `date` and converted to date format
- renamed `total_chops` column to `bulk_chops` column to distinguish from `individual_chops` column in `Carcass` table
```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns

clean_yuba_chops <- raw_carcass_counts %>% janitor::clean_names() %>%
  mutate(date = as_date(survey_id)) %>% 
  select(-survey_id) %>%
  rename(bulk_chops = total_chops) |> 
  glimpse()

summary(clean_yuba_chops)
```
### Combined
- Each row in the combined dataset represents a sampling event of either
  * biological sampling of an individual carcass (count stored in `individual_chops` column). This is from the `Carcass` table.
  * bulk sampling of chops on a given date (count stored in `bulk_chops` column). This is from the `Chop` table.
- To get a `total_count` value for a given `date`, sum `individual_chops` and `bulk_chops`.
- converted `adipose_clip` column to a logical variable
- modified values for `sex` and `spawned` columns to be more explicit
- renamed `fresh` to `carcass_condition` to standardize across streams and modified values to be more explicit

```{r}
clean_yuba_combined <- 
  bind_rows(clean_yuba_carcass, clean_yuba_chops) |> 
         mutate(adipose_clip = case_when(adipose_clip == "Y" ~ TRUE,
                                  adipose_clip == "N" ~ FALSE,
                                  adipose_clip %in% c(NA, "Unk") ~ NA),
         spawned = case_when(spawned == "Y" ~ "spawned",
                             spawned == "N" ~ "unspawned",
                             is.na(spawned) ~ NA_character_),
         sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         sex == "U" ~ "unknown",
                         TRUE ~ sex),
         carcass_condition = case_when(fresh == "Y" ~ "fresh",
                                       fresh == "N" ~ "non fresh",
                                       is.na(fresh) ~ NA_character_)) |>
  select(-fresh) |> 
  glimpse()
```

## Explore Numeric Variables: {.tabset}

```{r}
clean_yuba_combined |> 
  select_if(is.numeric) |>
  colnames()
```

### `river_mile`

**Plotting river_mile over Period of Record**

```{r}
summary(clean_yuba_combined$river_mile)

ggplot() + 
  geom_histogram(data = clean_yuba_combined, aes(river_mile), binwidth = 0.75, fill = "#74A089") + 
  theme_minimal()

```

### `fork_length`

```{r}

summary(clean_yuba_combined$fork_length)

# histogram
ggplot() + 
  geom_histogram(data = clean_yuba_combined, aes(fork_length), 
                 fill = "#74A089") + 
  theme_minimal()

# boxplot
ggplot() + 
  geom_boxplot(data = clean_yuba_combined, aes(y = fork_length, x = spawned)) +
  geom_jitter(data = clean_yuba_combined, aes(y = fork_length, x = spawned), alpha = 0.5) + 
  theme_minimal()
```

### `individual_chops` over time

```{r}
clean_yuba_combined |> 
  group_by(date) |> 
  summarise(count = sum(individual_chops, na.rm = T)) |> 
  ggplot(aes(x = date, y = count)) +
  geom_col(fill = "#74A089") + 
  theme_minimal()
```

### `bulk_chops` over time

```{r}
clean_yuba_combined |> 
  group_by(date) |> 
  summarise(count = sum(bulk_chops, na.rm = T)) |> 
  ggplot(aes(x = date, y = count)) +
  geom_col(fill = "#74A089") + 
  theme_minimal()
```


### total chops (`individual_chops` and `bulk_chops`) by year

```{r}
clean_yuba_combined |> 
  mutate(year = year(date)) |>
  group_by(year) |>
  summarise(count = sum(individual_chops, bulk_chops, na.rm = T)) |> 
  ggplot(aes(x = year, y = count)) +
  geom_col(fill = "#74A089") + 
  theme_minimal()
```

## Explore Categorical variables: {.tabset}

```{r}
clean_yuba_combined |> 
  select(!is.numeric) |> 
  colnames()
```

### `date`

```{r}
summary(clean_yuba_combined$date)
```

### `carcass_condition`

```{r}
table(clean_yuba_combined$carcass_condition, useNA = "ifany")
```

### `adipose_clip`

```{r}
table(clean_yuba_combined$adipose_clip, useNA = "ifany")

```

### `sex`

```{r}
table(clean_yuba_combined$sex, useNA = "ifany")

```

### `spawned`

```{r}
table(clean_yuba_combined$spawned, useNA = "ifany")

```

## Summary of identified issues

-   The data are in good shape and standardized across years. 
-   Important to be aware that bulk chop counts and individual chop counts are stored in separate variables; this can be changed easily. 

## Save cleaned data back to google cloud

```{r, include = F}
yuba_carcass <- clean_yuba_combined

# gcs_list_objects()
f <- function(input, output) write_csv(input, file = output)

gcs_upload(yuba_carcass,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/yuba-river/data/yuba_carcass.csv")
```
