---
title: "Feather - Check CAMP tables and variables"
output: html_document
date: "2022-09-13"
---

```{r}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r}
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
mdb.get(feather_camp, tables = T)
```

# Project Description 

Only one record: contains metadata information and is not needed as a separate data table

```{r}
project_description <- mdb.get(feather_camp, tables = "ProjectDescription")
```

# Site

I think this should be included as metadata, not a separate data table. Site will
be joined with other tables.

## variables to exclude
- beginDate
- endDate
- srcXCoord
- srcYCoord
- srcCoordSystem
- llSource
- srcDatum
- srcProjection
- coordSystem
- projection
- georefQAstatus
- relative
- usgsStationNumber
- dischargeURL
- comments
- activeID
- dataRecorder
- dataRecorderAgencyID,
- creationTime
- updateTime

```{r}
site <- mdb.get(feather_camp, tables = "Site")
```

# Subsite 

I think this should be joined with the Site table and included as metadata

## variables to exclude
- projectDescriptionID
- srcXCoord
- srcYCoord
- srcCoordSystem
- llSource
- srcDatum
- srcProjection
- coordSystem
- projection
- georefQAstatus
- comments
- activeID
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime

```{r}
subsite <- mdb.get(feather_camp, tables = "SubSite")
```

# TrapVisit

## variables to exclude
- projectDescriptionID (all one project)
- crew (internal only)
- sampleGearID (metadata)
- timeSampleStarted
- timeSampleEnded
- batchDate
- counterOld,
- counterOldTime
- counterAtEndTime
- counterNew
- coneDepthAtStart
- coneDepthAtEnd
- includeCatchComments
- coneDebrisIntake1
- coneDebrisIntake2
- dataCollectedID
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

```{r}
trap_visit <- mdb.get(feather_camp, tables = "TrapVisit")
# These are all NA
filter(trap_visit, !is.na(timeSampleStarted))
filter(trap_visit, !is.na(timeSampleEnded))
filter(trap_visit, !is.na(batchDate))
filter(trap_visit, !is.na(counterOld))
filter(trap_visit, !is.na(counterOldTime))
filter(trap_visit, !is.na(counterAtEndTime))
filter(trap_visit, !is.na(counterNew))
filter(trap_visit, !is.na(rpmSecondsAtStart))
# only two records where rpm is not 60 seconds
filter(trap_visit, rpmSecondsAtStart != 60)
filter(trap_visit, !is.na(coneDepthAtStart))
filter(trap_visit, !is.na(coneDepthAtEnd))
filter(trap_visit, !is.na(includeCatchID))
unique(trap_visit$includeCatchID)
filter(trap_visit, includeCatchID == 2)
filter(trap_visit, !is.na(includeCatchComments))
filter(trap_visit, includeCatchComments != "")
filter(trap_visit, !is.na(ConeDebrisIntake1))
filter(trap_visit, !is.na(ConeDebrisIntake2))
filter(trap_visit, !is.na(debrisVolumeCatID))
filter(trap_visit, !is.na(debrisVolume))
filter(trap_visit, !is.na(debrisType))
```

# CatchRaw

## variables to exclude
- projectDescriptionID
- auxiliaryRunID
- auxiliaryRunMethodID
- stdLength 
- subsampleMethodID
- subsampleNumerator
- subsampleDenominator
- fishHealth
- gillRating
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

```{r}
catch_raw <- mdb.get(feather_camp, tables = "CatchRaw")

filter(catch_raw, !is.na(auxiliaryRunID))
filter(catch_raw, !is.na(auxiliaryRunMethodID))
filter(catch_raw, !is.na(totalLength))
filter(catch_raw, !is.na(stdLength))
filter(catch_raw, !is.na(weight))
filter(catch_raw, !is.na(randomID))
filter(catch_raw, !is.na(subsampleMethodID))
filter(catch_raw, subsampleMethodID != 1)
filter(catch_raw, !is.na(subsampleNumerator))
filter(catch_raw, !is.na(actualCountID))
filter(catch_raw, actualCountID != 1)
filter(catch_raw, !is.na(fishHealth))
filter(catch_raw, !is.na(gillRating))
```

# MarkExisting

## variables to exclude
- projectDescriptionID
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

```{r}
mark <-  mdb.get(feather_camp, tables = "MarkExisting")
filter(mark, !is.na(markCode)) # rarely used but include to future proof
```

# Specimen

This table is not used. Exclude.

```{r}
specimen <-  mdb.get(feather_camp, tables = "Specimen")
```

# PostHandleMort

This table is not used. Exclude.

```{r}
post_handle_mort <- mdb.get(feather_camp, "PostHandleMort")
```

# PostHandleMortMark

This table is not used. Exclude.

```{r}
post_handle_mort_mark <- mdb.get(feather_camp, "PostHandleMortMark")
```

# EnvDataRaw

## variables to exclude
- measureTime does not exist
- projectDescriptionID
- subSiteID does not exist
- waterDepth
- waterDepthUnitID
- waterDepthSampleGearID
- waterVelReadingStart
- waterVelReadingEnd
- waterVelTime
- waterVelTimeUnitID
- waterVelConvertFactor
- airTemp
- airTempUnitID
- airTempSampleGearID
- dissolvedOxygen
- dissolvedOxygenUnitID
- dissolvedOxygenSampleGearID
- conductivity
- conductivityUnitID
- conductivitySampleGearID
- barometer
- barometerUnitID
- barometerSampleGearID
- cloudCover
- dataCollectedID
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

```{r}
environmental <- mdb.get(feather_camp, tables = "EnvDataRaw")
filter(environmental, !is.na(measureTime))
filter(environmental, !is.na(subSiteID))
filter(environmental, !is.na(discharge))
filter(environmental, !is.na(waterDepth))
filter(environmental, !is.na(waterVel))
filter(environmental, !is.na(waterVelReadingStart))
filter(environmental, !is.na(waterVelReadingEnd))
filter(environmental, !is.na(waterVelTime))
filter(environmental, !is.na(waterVelConvertFactor))
filter(environmental, !is.na(waterVelUnitID))
filter(environmental, !is.na(airTemp))
filter(environmental, !is.na(waterTemp))
filter(environmental, !is.na(lightPenetration))
filter(environmental, !is.na(turbidity))
filter(environmental, !is.na(dissolvedOxygen))
filter(environmental, !is.na(conductivity))
filter(environmental, !is.na(barometer))
filter(environmental, !is.na(cloudCover))
filter(environmental, !is.na(weather))
filter(environmental, !is.na(dataCollectedID))
filter(environmental, dataCollectedID != 1)
```

# EnvDataRawXTargetSite

If we join EnvDataRaw to the TrapVisit table do we need this table?

```{r}
env_raw_target <- mdb.get(feather_camp, "EnvDataRawXTargetSite")
```

# EnvDataDaily

This table is not used. Exclude.

```{r}
env_daily <- mdb.get(feather_camp, "EnvDataDaily")
```

# EnvDataDailyXTargetSite

This table is not used. Exclude.

```{r}
env_daily_target <- mdb.get(feather_camp, "EnvDataDailyXTargetSite")
```

# Release

## variables to exclude
- projectDescriptionID
- checkedTime
- nMarkExamined
- nMarkNotRetained
- propMarked
- dataCollectedID
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

```{r}
release <- mdb.get(feather_camp, tables = "Release")
filter(release, !is.na(releaseSiteID))
filter(release, !is.na(releasePurposeID))
filter(release, !is.na(markedTaxonID))
filter(release, !is.na(markedRunID))
filter(release, !is.na(markedLifeStageID))
filter(release, !is.na(markedFishOriginID))
filter(release, !is.na(sourceOfFishSiteID))
filter(release, !is.na(releaseSubSiteID))
filter(release, !is.na(checkedTime))
filter(release, !is.na(nMortWhileHandling))
filter(release, !is.na(nMortAtCheck))
filter(release, !is.na(nMarkExamined))
filter(release, !is.na(nMarkNotRetained))
filter(release, !is.na(propMarked))
filter(release, !is.na(releaseTime))
filter(release, !is.na(releaseLightConditionID))
filter(release, !is.na(includeTestID))
filter(release, includeTestID != 1)
filter(release, !is.na(includeTestComments))
```

# ReleaseXTargetSite

## variables to exclude
- projectDescriptionID
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime

```{r}
release_target <- mdb.get(feather_camp, tables = "ReleaseXTargetSite")
```

# MarkApplied

## variables to exclude
- projectDescriptionID
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime

```{r}
mark_applied <- mdb.get(feather_camp, tables = "MarkApplied")
filter(mark_applied, !is.na(appliedMarkTypeID))
filter(mark_applied, !is.na(appliedMarkColorID))
filter(mark_applied, !is.na(appliedMarkPositionID))
filter(mark_applied, !is.na(appliedMarkCode))
```

# ReleaseFish

## variables to exclude
- projectDescriptionID
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

```{r}
# no helpful information in this table but we want forklengths
release_fish <- mdb.get(feather_camp, tables = "ReleaseFish")
```
