---
title: "Feather CAMP Data Pull"
output: html_document
date: "2022-09-02"
---

This document queries the CAMP database provided by Kassie Hickey to extract
RST data for upper Feather traps. 

If this file doesn't run, make sure operating system is being set correctly. 
Depending on the operating system this script will run either the query_feather_camp 
'4pc' or '4mac'

*Important to filter releaseID == 0 or releasedID == 255 to exclude recaptures*

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(googleCloudStorageR)
library(Hmisc)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

# Pull in Access db from Google Cloud

```{r}
gcs_get_object(object_name = "rst/CAMP/feather_river/CAMP.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"),
               overwrite = TRUE)
```

# Query data tables

```{r}
# runs the mac or pc version of the query tables file depending on operating system
operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

if(operating_system == "pc") {
  source(here::here("data-raw", "qc-markdowns", "rst", "feather-river", "query_feather_camp_4pc.R"))
  } else{
    source(here::here("data-raw", "qc-markdowns", "rst", "feather-river", "query_feather_camp_4mac.R"))
  }
```

# Site

Joins site and subsite table

```{r}
site_subsite <- full_join(site_lu, subsite_lu) %>% glimpse()
```

# Trap

```{r}
detach(package:Hmisc)
# all NA
# filter(trap_visit, !is.na(timeSampleStarted))
# filter(trap_visit, !is.na(timeSampleEnded))
# filter(trap_visit, !is.na(coneDepthAtStart))
# filter(trap_visit, !is.na(coneDepthAtEnd))
# filter(trap_visit, !is.na(debrisVolume))
# 
# # not NA
# filter(trap_visit, !is.na(inThalwegID))
# filter(trap_visit, !is.na(sampleGearID))

trap_visit_format <- trap_visit %>% 
  select(projectDescriptionID, trapVisitID, trapPositionID, visitTime, visitTime2, visitTypeID, fishProcessedID, inThalwegID, trapFunctioningID, counterAtStart, counterAtEnd, rpmRevolutionsAtStart, rpmSecondsAtStart, rpmRevolutionsAtEnd, rpmSecondsAtEnd, halfConeID, includeCatchID, debrisVolumeCatID, debrisVolume, debrisVolumeUnits, comments) %>% 
  left_join(site_subsite, by = c("trapPositionID" = "subSiteID")) %>% 
  select(-trapPositionID, -siteID) %>% 
  left_join(visit_type_lu, by = c("visitTypeID" = "visitTypeID")) %>% 
  select(-visitTypeID) %>% 
  left_join(processed_lu, by = c("fishProcessedID" = "fishProcessedID")) %>% 
  select(-fishProcessedID) %>% 
  left_join(no_yes_lu, by = c("inThalwegID" = "noYesID")) %>% 
  rename(inThalweg = noYes) %>% 
  select(-inThalwegID) %>% 
  left_join(trap_function_lu, by = c("trapFunctioningID" = "trapFunctioningID")) %>% 
  select(-trapFunctioningID) %>% 
  left_join(no_yes_lu, by = c("halfConeID" = "noYesID")) %>% 
  rename(halfCone = noYes) %>% 
  select(-halfConeID) %>% 
  left_join(no_yes_lu, by = c("includeCatchID" = "noYesID")) %>% 
  rename(includeCatch = noYes) %>% 
  select(-includeCatchID) %>% 
  left_join(debris_volume_lu, by = c("debrisVolumeCatID" = "debrisVolumeCatID")) %>% 
  select(-debrisVolumeCatID) %>% glimpse()

# Save data to google cloud
f <- function(input, output) write_csv(input, file = output)
gcs_upload(trap_visit_format,
           object_function = f,
           type = "csv",
           name = "rst/feather-river/data-raw/feather_trap_camp.csv",
           predefinedAcl = "bucketLevel")
```

# Catch (including recaptures)

```{r}
mark_existing_w_joins <- mark %>% 
  left_join(mark_type_lu, by = c("markTypeID" = "markTypeID")) %>% 
  select(-markTypeID) %>% 
  left_join(color_lu, by = c("markColorID" = "colorID")) %>% 
  rename(markColor = color) %>% 
  select(-markColorID) %>% 
  left_join(body_part_lu, by = c("markPositionID" = "bodyPartID")) %>% 
  rename(markPosition = bodyPart) %>% 
  select(-markPositionID) %>% 
  select(projectDescriptionID, catchRawID, markCode, markType, markColor, markPosition) %>% 
  mutate(markCode = as.character(markCode)) %>% 
  pivot_longer(cols = starts_with("mark"),
               names_to = "mark",
               values_to = "value") %>% 
  group_by(catchRawID, projectDescriptionID, mark) %>% 
  mutate(number = 1:n(),
         mark = paste0(mark,"_",number)) %>% 
  select(-number) %>% 
  pivot_wider(id_cols = c("catchRawID", "projectDescriptionID"), names_from = "mark", values_from = "value")

selected_catch <- catch_raw %>% 
  select(projectDescriptionID, catchRawID, trapVisitID, taxonID, atCaptureRunID, atCaptureRunMethodID, finalRunID, finalRunMethodID, fishOriginID, lifeStageID, forkLength, totalLength, weight, n, randomID, actualCountID, releaseID, mortID, comments) %>% 
  left_join(trap_visit_format %>% 
              select(projectDescriptionID, trapVisitID, visitTime, visitTime2, visitType, siteName, subSiteName), by = c("trapVisitID" = "trapVisitID", "projectDescriptionID" = "projectDescriptionID")) %>% 
  left_join(mark_existing_w_joins, by = c("catchRawID" = "catchRawID", "projectDescriptionID" = "projectDescriptionID")) %>% 
  mutate(taxonID = as.character(taxonID)) %>% 
  left_join(taxon_lu %>% 
              mutate(taxonID = as.character(taxonID)), 
            by = c("taxonID" = "taxonID")) %>% 
  select(-taxonID) %>% 
  left_join(run_lu, by = c("atCaptureRunID" = "runID")) %>% 
  rename(atCaptureRun = run) %>% 
  select(-atCaptureRunID) %>% 
  left_join(run_method_lu, by = c("atCaptureRunMethodID" = "runMethodID")) %>% 
  rename(atCaptureRunMethod = runMethod) %>% 
  select(-atCaptureRunMethodID) %>% 
  left_join(run_lu, by = c("finalRunID" = "runID")) %>% 
  rename(finalRun = run) %>% 
  select(-finalRunID) %>% 
  left_join(run_method_lu, by = c("finalRunMethodID" = "runMethodID")) %>% 
  rename(finalRunMethod = runMethod) %>% 
  select(-finalRunMethodID) %>% 
  left_join(origin_lu, by = c("fishOriginID" = "fishOriginID")) %>% 
  select(-fishOriginID) %>% 
  left_join(lifestage_lu, by = c("lifeStageID" = "lifeStageID")) %>% 
  select(-lifeStageID, -lifeStageCAMPID) %>% 
  left_join(no_yes_lu, by = c("randomID" = "noYesID")) %>% 
  rename(random = noYes) %>% 
  select(-randomID) %>% 
  left_join(no_yes_lu, by = c("actualCountID" = "noYesID")) %>% 
  rename(actualCount = noYes) %>% 
  select(-actualCountID) %>% 
  left_join(no_yes_lu, by = c("mortID" = "noYesID")) %>% 
  rename(mort = noYes) %>% 
  select(-mortID) %>%  glimpse()

# Save data to google cloud
f <- function(input, output) write_csv(input, file = output)
gcs_upload(selected_catch,
           object_function = f,
           type = "csv",
           name = "rst/feather-river/data-raw/feather_catch_camp.csv",
           predefinedAcl = "bucketLevel")
```

# Recaptures

```{r}
trap_visit_recaptures <- trap_visit %>%
  select(trapVisitID, visitTime, visitTime2, visitTypeID, halfConeID) %>%
  left_join(visit_type_lu, by = c("visitTypeID" = "visitTypeID")) %>% 
  mutate(halfCone = ifelse(halfConeID == 1, T, F)) %>% 
  select(-visitTypeID, -halfConeID)

# add atcapturerun
selected_recaptured <- catch_raw %>% 
  select(catchRawID, taxonID, atCaptureRunID, atCaptureRunMethodID, finalRunID, finalRunMethodID, fishOriginID, lifeStageID, forkLength, weight, n, trapVisitID, releaseID, mortID) %>% 
  mutate(taxonID = as.numeric(taxonID)) %>% 
  left_join(taxon_lu, by = c("taxonID" = "taxonID")) %>%
  # filter(commonName == "Chinook salmon") %>% 
  left_join(run_lu, by = c("finalRunID" = "runID")) %>%
  rename(final_run = run) %>% 
  left_join(run_lu, by = c("atCaptureRunID" = "runID")) %>%
  left_join(run_method_lu, by = c("finalRunMethodID" = "runMethodID")) %>%
  rename(final_run_method = runMethod) %>% 
  left_join(run_method_lu, by = c("atCaptureRunMethodID" = "runMethodID")) %>%
  left_join(lifestage_lu %>% 
              select(-lifeStageCAMPID), by = c("lifeStageID" = "lifeStageID")) %>%
  left_join(origin_lu, by = c("fishOriginID" = "fishOriginID")) %>%
  left_join(no_yes_lu, by = c("mortID" = "noYesID")) %>% 
  rename(dead = noYes) %>% 
  select(-lifeStageID, -finalRunMethodID, -finalRunID, -atCaptureRunMethodID, -atCaptureRunID, -taxonID, -fishOriginID, - mortID) %>%
  left_join(trap_visit_recaptures) %>%
  # filter out marked and recapture fish by selecting for only fish where release trial does not exist
  filter(releaseID != 0 & releaseID != 255) %>% 
  glimpse()

# Save data to google cloud
f <- function(input, output) write_csv(input, file = output)
gcs_upload(selected_recaptured,
           object_function = f,
           type = "csv",
           name = "rst/feather-river/data-raw/feather_recaptures_camp.csv",
           predefinedAcl = "bucketLevel")
```

# Releases

```{r}
selected_release <- release %>% 
  select(projectDescriptionID, releaseID, releasePurposeID, markedTaxonID, markedRunID, markedLifeStageID, markedFishOriginID, sourceOfFishSiteID, releaseSiteID, releaseSubSiteID, nMortWhileHandling, nMortAtCheck, nReleased, releaseTime, releaseLightConditionID,
         testDays, includeTestID, includeTestComments, comments) %>% 
  left_join(release_purpose_lu, by = c("releasePurposeID" = "releasePursposeID")) %>% 
  select(-releasePurposeID) %>% 
  left_join(taxon_lu, by = c("markedTaxonID" = "taxonID")) %>% 
  rename(markedTaxon = commonName) %>% 
  select(-markedTaxonID) %>% 
  left_join(run_lu, by = c("markedRunID" = "runID")) %>% 
  rename(markedRun = run) %>% 
  select(-markedRunID) %>% 
  left_join(lifestage_lu, by = c("markedLifeStageID" = "lifeStageID")) %>% 
  rename(markedLifeStage = lifeStage) %>% 
  select(-markedLifeStageID) %>% 
  left_join(origin_lu, by = c("markedFishOriginID" = "fishOriginID")) %>% 
  rename(markedFishOrigin = fishOrigin) %>% 
  select(-markedFishOrigin) %>% 
  left_join(site_subsite %>% 
              select(siteID, siteName), by = c("sourceOfFishSiteID" = "siteID")) %>% 
  rename(sourceOfFishSite = siteName) %>% 
  select(-sourceOfFishSiteID) %>% 
  left_join(site_subsite %>% 
              select(siteID, siteName), by = c("releaseSiteID" = "siteID")) %>% 
  rename(releaseSite = siteName) %>% 
  select(-releaseSiteID) %>% 
  left_join(site_subsite %>% 
              select(subSiteID, subSiteName), by = c("releaseSubSiteID" = "subSiteID")) %>% 
  rename(releaseSubSite = subSiteName) %>% 
  select(-releaseSubSiteID) %>% 
  left_join(light_condition_lu, by = c("releaseLightConditionID" = "lightConditionID")) %>% 
  rename(releaseLightCondition = lightCondition) %>% 
  select(-releaseLightConditionID) %>%  
  left_join(no_yes_lu, by = c("includeTestID" = "noYesID")) %>% 
  rename(includeTest = noYes) %>% 
  select(-includeTestID) %>% glimpse()

# Save data to google cloud
f <- function(input, output) write_csv(input, file = output)
gcs_upload(selected_release,
           object_function = f,
           type = "csv",
           name = "rst/feather-river/data-raw/feather_releases_camp.csv",
           predefinedAcl = "bucketLevel")
```

# Environmental

```{r}
# all NA
# filter(environmental, !is.na(waterVelReadingStart))
# filter(environmental, !is.na(waterVelReadingEnd))
# filter(environmental, !is.na(waterVelTime))
# filter(environmental, !is.na(airTemp))
# filter(environmental, !is.na(cloudCover))
# filter(environmental, !is.na(waterDepth))
# filter(environmental, !is.na(dissolvedOxygen))
# filter(environmental, !is.na(conductivity))
# filter(environmental, !is.na(barometer))
# 
# # contains data
# filter(environmental, !is.na(discharge))
# filter(environmental, !is.na(dischargeUnitID))
# filter(environmental, !is.na(dischargeSampleGearID))
# filter(environmental, !is.na(waterVel))
# filter(environmental, !is.na(waterTemp))
# filter(environmental, !is.na(lightPenetration))
# filter(environmental, !is.na(turbidity))
# filter(environmental, !is.na(weather))

# select relevant variables from envDataRaw table

discharge <- environmental %>% 
  select(trapVisitID, envDataRawID, discharge, dischargeUnitID, dischargeSampleGearID) %>% 
  rename(value = discharge,
         unitID = dischargeUnitID,
         sampleGearID = dischargeSampleGearID) %>% 
  mutate(parameter = "discharge")

# depth <- environmental %>% 
#   select(trapVisitID, envDataRawID, waterDepth, waterDepthUnitID, waterDepthSampleGearID) %>% 
#   rename(value = waterDepth,
#          unitID = waterDepthUnitID,
#          sampleGearID = waterDepthSampleGearID) %>% 
#   mutate(parameter = "water depth")

velocity <- environmental %>% 
  select(trapVisitID, envDataRawID, waterVel, waterVelUnitID, waterVelSampleGearID) %>% 
  rename(value = waterVel,
         unitID = waterVelUnitID,
         sampleGearID = waterVelSampleGearID) %>% 
  mutate(parameter = "velocity")

temperature <- environmental %>% 
  select(trapVisitID, envDataRawID, waterTemp, waterTempUnitID, waterTempSampleGearID) %>% 
  rename(value = waterTemp,
         unitID = waterTempUnitID,
         sampleGearID = waterTempSampleGearID) %>% 
  mutate(parameter = "temperature")

light <- environmental %>% 
  select(trapVisitID, envDataRawID,  lightPenetration, lightPenetrationUnitID, lightPenetrationSampleGearID) %>% 
  rename(value = lightPenetration,
         unitID = lightPenetrationUnitID,
         sampleGearID = lightPenetrationSampleGearID) %>% 
  mutate(parameter = "light penetration")

turbidity <- environmental %>% 
  select(trapVisitID, envDataRawID,  turbidity, turbidityUnitID, turbiditySampleGearID) %>% 
  rename(value = turbidity,
         unitID = turbidityUnitID,
         sampleGearID = turbiditySampleGearID) %>% 
  mutate(parameter = "turbidity")

# oxygen <- environmental %>% 
#   select(trapVisitID, envDataRawID,  dissolvedOxygen, dissolvedOxygenUnitID, dissolvedOxygenSampleGearID) %>% 
#   rename(value = dissolvedOxygen,
#          unitID = dissolvedOxygenUnitID,
#          sampleGearID = dissolvedOxygenSampleGearID) %>% 
#   mutate(parameter = "dissolved oxygen")
# 
# conductivity <- environmental %>% 
#   select(trapVisitID, envDataRawID,  conductivity, conductivityUnitID, conductivitySampleGearID) %>% 
#   rename(value = conductivity,
#          unitID = conductivityUnitID,
#          sampleGearID = conductivitySampleGearID) %>% 
#   mutate(parameter = "conductivity")
# 
# barometer <- environmental %>% 
#   select(trapVisitID, envDataRawID,  barometer, barometerUnitID, barometerSampleGearID) %>% 
#   rename(value = barometer,
#          unitID = barometerUnitID,
#          sampleGearID = barometerSampleGearID) %>% 
#   mutate(parameter = "barometer")

trap_visit_subsite <- trap_visit %>%
  select(trapVisitID, visitTime, visitTime2, trapPositionID) %>%
  left_join(site_subsite, by = c("trapPositionID" = "subSiteID")) %>% 
  select(-trapPositionID, -siteID) %>% 
  rename(site = siteName,
         subsite = subSiteName) %>% 
  mutate(site = tolower(site),
         subsite = tolower(subsite))

environmental_format <- bind_rows(discharge,
                                  velocity,
                                  temperature,
                                  light,
                                  turbidity) %>% 
  left_join(trap_visit_subsite) %>% 
  left_join(unit_lu) %>% 
  left_join(sample_gear_lu) %>% 
  select(-c(unitID, sampleGearID, visitTime2)) %>% 
  rename(date = visitTime,
         sample_gear = sampleGear) %>% 
  mutate(unit = case_when(is.na(unit) & !is.na(value) ~ "not recorded", 
                          is.na(unit) & is.na(value) ~ NA_character_,
                          T ~ unit),
         sample_gear = case_when(is.na(sample_gear) & !is.na(value) ~ "not recorded", 
                                 is.na(sample_gear) & is.na(value) ~ NA_character_,
                                 T ~ sample_gear)) %>%  glimpse()

# Save data to google cloud
f <- function(input, output) write_csv(input, file = output)
gcs_upload(environmental_format,
           object_function = f,
           type = "csv",
           name = "rst/feather-river/data-raw/feather_environmental_camp.csv",
           predefinedAcl = "bucketLevel")
```
