---
title: "Lower Sacramento (Tisdale) RST data QC"
author: "Ashley Vizek"
date: "11/10/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(knitr)
```

## Description of Monitoring Data

**Timeframe:** 

07/07/2010 through 05/28/2020

**Completeness of Record throughout timeframe:** 

Water year 2012 has incomplete data.

**Sampling Location:**

Tisdale

**Data Contact:** 

[Drew Huneycutt](mailto::andrew.huneycutt@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()

gcs_get_object(object_name = "rst/lower-sac-river/data/tisdale/tisdale_rst_raw_clean.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "tisdale_rst_raw_clean.csv",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet.

```{r}
# read in data to clean 
raw_clean <- read_csv("tisdale_rst_raw_clean.csv", show_col_types = F)
glimpse(raw_clean)
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- raw_clean %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(raw_clean),
                          description = c("Date of sampling",
                                          "Position of trap. Options are river left (RL) or river right (RR).",
                                          "Describes if fish were caught and processed",
                                          "Species of fish caught. This data only includes Chinook.",
                                          "Fork length of sampled fish in mm.",
                                          "Weight of sampled fish in grams",
                                          "Lifestage of sampled fish. Options include parr, silvery parr, fry, button-up fry, yolk sac fry, smolt, juvenile",
                                          "Run determined when fish was captured.",
                                          "Adjusted run.",
                                          "Describes if fish was dead when collected.",
                                          "Describes if sample was random.",
                                          "Describes if analyses was conducted on the data. All cases are Yes.",
                                          "Rearing origin of fish. Options are Natural or Hatchery.",
                                          "Identifier for the location fish were released. All are the same.",
                                          "Type of mark observed. Options are None, Pigment, Elastomer.",
                                          "Position of mark observed. Options are Adipose fin, Whole body, Nose, Pelvic fin right, Pelvic fin left.",
                                          "Color of mark observed. Options are Blue, Red, Brown, Yellow, Green, White, Orange.",
                                          "Qualitative comments about sampling.",
                                          "Qualitative comments about the fish caught.",
                                          "Site name. All are Tisdale Weir RST.",
                                          "Count of fish caught."
),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
```

## Data transformations

This work was done in other rmd (1_tisdale_initial_clean.Rmd)

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables (this way we know we do not miss any)
raw_clean %>%
  select_if(is.numeric) %>%
  colnames()
```

### Variable: `fork_length_mm`

**Plotting `fork_length_mm` over Period of Record**

There are some outliers - fork length greater than 300.

```{r}
filter(raw_clean, fork_length_mm < 300) %>%
ggplot(aes(x = fork_length_mm)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

**Numeric Summary of `fork_length_mm` over Period of Record**

```{r}
raw_clean %>%
  group_by(year(date)) %>%
  summarize(mean = mean(fork_length_mm, na.rm = T),
            median = median(fork_length_mm, na.rm = T),
            min = min(fork_length_mm, na.rm = T),
            max = max(fork_length_mm, na.rm = T),
            na = length(which(is.na(fork_length_mm))))
```

Looks like the outlier should be 40 rather than 4038.

```{r}
filter(raw_clean, fork_length_mm > 300)

raw_clean %>%
  group_by(life_stage) %>%
  summarize(mean = mean(fork_length_mm, na.rm = T),
            median = median(fork_length_mm, na.rm = T))
```


**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(raw_clean$fork_length_mm))/nrow(raw_clean), 3) * 100` % of values in the `fork_length_mm` column are NA.

### Variable: `weight`

**Plotting `weight` over Period of Record**

```{r}
filter(raw_clean, weight < 20) %>%
ggplot(aes(x = weight)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

**Numeric Summary of `weight` over Period of Record**

```{r}
raw_clean %>%
  group_by(year(date)) %>%
  summarize(mean = mean(weight, na.rm = T),
            median = median(weight, na.rm = T),
            min = min(weight, na.rm = T),
            max = max(weight, na.rm = T),
            na = length(which(is.na(weight))))
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(raw_clean$weight))/nrow(raw_clean), 3) * 100` % of values in the `weight` column are NA.

### Variable: `count`

**Plotting `count` over Period of Record**

```{r}
raw_clean %>%
  group_by(date, at_capture_run) %>%
  summarize(count = sum(count)) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
ggplot(aes(y = count, x = fake_date, color = at_capture_run)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line() +
  facet_wrap(~wy, scales = "free_y") +
  xlab("") +
  theme(legend.position="bottom")
```

**Numeric Summary of `count` over Period of Record**

```{r}
knitr::kable(raw_clean %>%
  group_by(date, at_capture_run) %>%
  summarize(count = sum(count)) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(mean = round(mean(count, na.rm = T),2),
            median = round(median(count, na.rm = T),2),
            min = round(min(count, na.rm = T),2),
            max = round(max(count, na.rm = T),2),
            na = length(which(is.na(count)))))
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(raw_clean$count))/nrow(raw_clean), 3) * 100` % of values in the `count` column are NA.

### Variable: `release_id`

**Plotting `release_id` over Period of Record**

`release_id` is always the same value. Make this a character instead of numeric.
  
## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

Need to change data to the correct format. 

I don't think we need the variables random and analyses but will leave them in.

```{r}
raw_clean %>%
  select_if(is.character) %>%
  colnames()
```

### Variable: `trap_position`

 `RL` is river left and `RR` is river right
 
```{r}
unique(raw_clean$trap_position) 

filter(raw_clean, trap_position == "N/A")

group_by(raw_clean, date) %>%
  tally() %>%
  filter(n > 2)
```

Some years do not have `trap_location` variable filled in but it is labeled as "N / A" so change that to NA. 

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$trap_position))/nrow(raw_clean), 3) * 100` % of values in the `trap_position` column are NA.

### Variable: `fish_processed`

Two instances where `fish_processed` is `No fish were caught` when count is greater than 0. I think these are mistakes and should be changes to `Processed fished`

```{r}
unique(raw_clean$fish_processed) 
filter(raw_clean, fish_processed == "No fish were caught", count > 0) %>%
  select(date, fish_processed, fork_length_mm, at_capture_run, count)
filter(raw_clean, fish_processed == "N/A; not a sampling visit") %>%
  select(date, fish_processed, fork_length_mm, at_capture_run, count)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$fish_processed))/nrow(raw_clean), 3) * 100` % of values in the `fish_processed` column are NA.

### Variable: `species`

Data only includes `Chinook salmon`

```{r}
unique(raw_clean$species)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$species))/nrow(raw_clean), 3) * 100` % of values in the `species` column are NA.

### Variable: `life_stage`

Parr is the most abundant life stage.

```{r}
unique(raw_clean$life_stage)
```

```{r}
total <- raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(total = sum(count))
raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, life_stage) %>%
  summarize(count = sum(count)) %>%
  left_join(total) %>%
  mutate(proportion_total = count/total) %>%
  ggplot(aes(y = life_stage, x = proportion_total)) +
  geom_col() +
  facet_wrap(~wy)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$life_stage))/nrow(raw_clean), 3) * 100` % of values in the `life_stage` column are NA.

### Variable: `at_capture_run`

10 instances where `at_capture_run` is `Not applicable (n/a)`. Change these to be NA.

Fall is the most abundant run.

```{r}
unique(raw_clean$at_capture_run)

filter(raw_clean, at_capture_run == "Not applicable (n/a)") %>%
  select(date, fish_processed, fork_length_mm, at_capture_run, count)
```

```{r}
total <- raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(total = sum(count))
raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, at_capture_run) %>%
  summarize(count = sum(count)) %>%
  left_join(total) %>%
  mutate(proportion_total = count/total) %>%
  ggplot(aes(y = at_capture_run, x = proportion_total)) +
  geom_col() +
  facet_wrap(~wy)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$at_capture_run))/nrow(raw_clean), 3) * 100` % of values in the `at_capture_run` column are NA.


### Variable: `final_run`

Assuming that `final_run` is the qa/qc version of `at_capture_run`. For the most part it doesn't make much of a difference - there are only 13 instances when the run is adjusted in `final_run`. I think we should use `final_run` as the `run` variable.

```{r}
unique(raw_clean$final_run)

filter(raw_clean, !is.na(final_run), final_run != at_capture_run, final_run != "Not recorded") %>%
  select(date, fish_processed, fork_length_mm, at_capture_run, count)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$final_run))/nrow(raw_clean), 3) * 100` % of values in the `final_run` column are NA.

### Variable: `mortality`

Transform to T/F. Typically alive.

```{r}
unique(raw_clean$mortality)
```

```{r}
total <- raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(total = sum(count))
raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, mortality) %>%
  summarize(count = sum(count)) %>%
  left_join(total) %>%
  mutate(proportion_total = count/total) %>%
  ggplot(aes(y = mortality, x = proportion_total)) +
  geom_col() +
  facet_wrap(~wy)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$mortality))/nrow(raw_clean), 3) * 100` % of values in the `mortality` column are NA.

### Variable: `random`

Only two values where `random` is `Not recorded`. Most all cases are `Yes`. I'm not sure how useful this information is. I will keep it in the dataset though.

Transform to T/F.

```{r}
unique(raw_clean$random)

filter(raw_clean, random == "Not recorded") %>%
  select(date, fish_processed, fork_length_mm, at_capture_run, count, random)
```

```{r}
total <- raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(total = sum(count))
raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, random) %>%
  summarize(count = sum(count)) %>%
  left_join(total) %>%
  mutate(proportion_total = count/total)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$random))/nrow(raw_clean), 3) * 100` % of values in the `random` column are NA.

### Variable: `analyses`

All cases are `Yes`. This doesn't provide any information. Remove this variable from dataset.

```{r}
unique(raw_clean$analyses)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$analyses))/nrow(raw_clean), 3) * 100` % of values in the `analyses` column are NA.

### Variable: `rearing`

Natural and Hatchery are the options for this variable. Most all are Natural but there are some Hatchery.

```{r}
unique(raw_clean$rearing)

filter(raw_clean, rearing == "Natural", mark_type != "No marks or tags") %>%
  select(date, fork_length_mm, at_capture_run, count, rearing)
```

```{r}
total <- raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(total = sum(count))
raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, rearing) %>%
  summarize(count = sum(count)) %>%
  left_join(total) %>%
  mutate(proportion_total = count/total) %>%
  ggplot(aes(y = rearing, x = proportion_total)) +
  geom_col() +
  facet_wrap(~wy)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$rearing))/nrow(raw_clean), 3) * 100` % of values in the `rearing`` column are NA.

### Variable: `mark_type`

Simplified the names of categories:

- No marks or tags = None
- Pigment / dye = Pigment
- Elastomer = Elastomer

Data is for fish with mostly no marks or tags.

```{r}
unique(raw_clean$mark_type)
# curious whether there are cases when natural fish have marks.
filter(raw_clean, rearing == "Natural", mark_type != "No marks or tags") %>%
  select(date, fork_length_mm, at_capture_run, count, rearing, mark_type)
```

```{r}
total <- raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(total = sum(count))
raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, mark_type) %>%
  summarize(count = sum(count)) %>%
  left_join(total) %>%
  mutate(proportion_total = count/total) %>%
  ggplot(aes(y = mark_type, x = proportion_total)) +
  geom_col() +
  facet_wrap(~wy)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$mark_type))/nrow(raw_clean), 3) * 100` % of values in the `mark_type`` column are NA.

### Variable: `mark_position`

Left category names as is but changed Not applicable (n/a) to NA.

- Adipose fin
- Whole body
- Nose
- Pelvic fin, right
- Pelvic fin, left

Most all marks are on the adipose fin.

```{r}
unique(raw_clean$mark_position)
# curious whether this variable is filled out when mark_type is none
filter(raw_clean, mark_type == "No marks or tags", mark_position != "Not applicable (n/a)")
```

```{r}
total <- raw_clean %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(total = sum(count))
filter(raw_clean, mark_type != "No marks or tags") %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, mark_position) %>%
  summarize(count = sum(count)) %>%
  left_join(total) %>%
  mutate(proportion_total = count/total) %>%
  ggplot(aes(y = mark_position, x = proportion_total)) +
  geom_col() +
  facet_wrap(~wy)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$mark_position))/nrow(raw_clean), 3) * 100` % of values in the `mark_positon` column are NA.

### Variable: `mark_color`

Left category names as is but changed Not applicable (n/a) to NA.

- Blue
- Red
- Brown
- Yellow
- Green
- White
- Orange

Most all data is NA. Only 35 cases when a color is recorded.

```{r}
unique(raw_clean$mark_color)
# curious whether this variable is filled out when mark_type is none
filter(raw_clean, mark_type == "No marks or tags", mark_color != "Not applicable (n/a)")

filter(raw_clean, mark_color != "Not applicable (n/a)") %>%
  select(date, fork_length_mm, at_capture_run, count, rearing, mark_type, mark_color)
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$mark_color))/nrow(raw_clean), 3) * 100` % of values in the `mark_color` column are NA.


### Variable: `trap_visit_comment`

There are some comments entered. I am not going to standardize these at this point.

```{r}
filter(raw_clean, !is.na(trap_visit_comment)) %>%
  group_by(year(date)) %>%
  tally()

filter(raw_clean, !is.na(trap_visit_comment)) %>%
  select(trap_visit_comment) %>%
  glimpse()
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$trap_visit_comment))/nrow(raw_clean), 3) * 100` % of values in the `trap_visit_comment` column are NA.

### Variable: `catch_comment`

There are no comments entered. Remove this variable.

```{r}
filter(raw_clean, !is.na(catch_comment)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$catch_comment))/nrow(raw_clean), 3) * 100` % of values in the `catch_comment` column are NA.

### Variable: `location`

Location is Tisdale Weir RST

```{r}
unique(raw_clean$location)
filter(raw_clean, is.na(location))
```

**NA and Unknown Values**

`r round(sum(is.na(raw_clean$location))/nrow(raw_clean), 3) * 100` % of values in the `location` column are NA.


## Make any changes identified

```{r make_changes}
rst_changes <- raw_clean %>%
  select(-analyses, -catch_comment) %>%
  # change release_id to categorical
  mutate(release_id = as.character(release_id),
         date = as.Date(date),
         trap_position = ifelse(trap_position == "N/A", NA_character_, trap_position),
         fork_length_mm = ifelse(fork_length_mm == 4038, 40, fork_length_mm),
         fish_processed = case_when(fish_processed == "No fish were caught" & count > 0 ~ "Processed fish",
                                    T ~ fish_processed),
         at_capture_run = ifelse(at_capture_run == "Not applicable (n/a)", NA_character_, at_capture_run),
         run = case_when(!is.na(final_run) & final_run != "Not recorded" ~ final_run,
                         T ~ at_capture_run),
         mortality = ifelse(mortality == "Yes",T,F),
         random = case_when(random == "Yes" ~ T,
                            random == "No" ~ F,
                            T ~ NA),
         mark_type = case_when(mark_type == "No marks or tags" ~ "None",
                               mark_type == "Pigment / dye" ~ "Pigment",
                               T ~ mark_type),
         mark_position = ifelse(mark_position == "Not applicable (n/a)", NA_character_, mark_position))
```

### Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
# f <- function(input, output) write_csv(input, file = output)
# 
# gcs_upload(rst_changes,
#            object_function = f,
#            type = "csv",
#            name = "rst/lower-sac-river/data/tisdale/rst_clean.csv")
```
