---
title: "Mill Creek Adult Upstream Passage Estimate QC"
author: "Inigo Peng"
date: "10/19/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
```

# Mill Creek Adult Upstream Passage Estimate Data 2007 to 2023

**Description of Monitoring Data**

Adult spring run daily passage estimate is based on data recorded at Ward Dam via video monitoring.

**Timeframe:** 

2007 to 2023

**Completeness of Record throughout timeframe:**

* All NA values for count are converted to `0`
* Flow and temperature are only collected for early years (`2007-2012`) and are mostly NA

**Sampling Location:** 

* Ward Dam   

**Data Contact:** [Ryan Revnak](mailto:Ryan.Revnak@wildlife.ca.gov) and [Doug Killam](mailto:Doug.Killam@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# get filenames for mill creek
mill_creek_filenames <- gcs_list_objects() |> 
  filter(str_detect(name, "adult-upstream-passage-monitoring/mill-creek/data-raw/MCVS")) |> 
  pull(name)

# define function to download all files
get_files_from_cloud <- function(filename) {
  print(substr(filename, 55, nchar(filename)))
  gcs_get_object(object_name = filename,
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = paste0("data-raw/qc-markdowns/adult-upstream-passage-monitoring/mill-creek/", substr(filename, 55, nchar(filename))),
                 overwrite = TRUE)
  message(paste0("downloaded ", filename))
}

lapply(mill_creek_filenames, get_files_from_cloud)

# get the one that says COPY of
# 2020-2021 passage data
gcs_get_object(object_name = "adult-upstream-passage-monitoring/mill-creek/data-raw/Copy of MCVS 2021-22 spring FINAL 8-17-22.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-upstream-passage-monitoring/mill-creek/mill_creek_passage_counts_2021.xlsx",
               overwrite = TRUE)
```
## Read in and clean raw count data for each year

```{r, message = FALSE, warning = FALSE}
save_to_prefix <- "data-raw/qc-markdowns/adult-upstream-passage-monitoring/mill-creek/"
```

This requires a different approach for each year because the file formats change.

### 2006-2007
```{r, message = FALSE, warning = FALSE}
mill_2006_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2006-2007 as of 2-14-2008.xls"), 
                                  sheet = "Daily Totals",
                                  skip = 3) |> 
  glimpse()

mill_2006 <- mill_2006_raw |> 
  janitor::clean_names() |> 
  select(date, avg_water_temp, avg_flow_in_cfs, salmon_up, salmon_down) |> 
  pivot_longer(salmon_up:salmon_down,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "salmon_up", "up", "down")) |> 
  glimpse()
```
### 2007-2008
```{r, message = FALSE, warning = FALSE}
mill_2007_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2007-2008 as of 7-24-08.xls"), 
                                  sheet = "Daily Totals",
                                  skip = 3) |> 
  glimpse()

mill_2007 <- mill_2007_raw |> 
  janitor::clean_names() |> 
  select(date, avg_water_temp = water_temp, number_up, number_down) |> 
  pivot_longer(number_up:number_down,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "number_up", "up", "down")) |> 
  glimpse()
```

### 2008-2009
```{r, message = FALSE, warning = FALSE}
mill_2008_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2008-2009 as of 8-10-10.xls"), 
                                  sheet = "Summary Totals ",
                                  skip = 4) |> 
  glimpse()

mill_2008 <- mill_2008_raw |> 
  janitor::clean_names() |> 
  select(date, avg_water_temp = water_temp_avg, avg_flow_in_cfs = flow_avg, salmon_number_up, salmon_number_down) |> 
  pivot_longer(salmon_number_up:salmon_number_down,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "salmon_number_up", "up", "down")) |> 
  glimpse()
```
### 2009-2010
```{r, message = FALSE, warning = FALSE}
mill_2009_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2009-2010 as of  9-21-10.xls"), 
                                  sheet = "SUMMARY",
                                  skip = 4) |> 
  glimpse()

mill_2009 <- mill_2009_raw |> 
  janitor::clean_names() |> 
  select(date, avg_water_temp = water_temp_avg, avg_flow_in_cfs = flow_avg, salmon_number_up, salmon_number_down) |> 
  pivot_longer(salmon_number_up:salmon_number_down,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "salmon_number_up", "up", "down")) |> 
  glimpse()
```

### 2010-2011
- no `flow` or `temperature`, but some columns `flow_mch_max` and `mch_low_water_tmp_avg`.

```{r, message = FALSE, warning = FALSE}
mill_2010_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2010-2011 as of 10-16-12.xls"), 
                                  sheet = "SUMMARY",
                                  skip = 4) |> 
  glimpse()

mill_2010 <- mill_2010_raw |> 
  janitor::clean_names() |> 
  select(date, salmon_number_up, salmon_number_down) |> 
  pivot_longer(salmon_number_up:salmon_number_down,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "salmon_number_up", "up", "down")) |> 
  glimpse()
```

### 2011-2012
- seems they started collecting more flow and temp variables, adding a descriptor `mch` to them and collecting `min` and `max`.

```{r, message = FALSE, warning = FALSE}
mill_2011_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2011-2012 as of 9-22-13.xls"), 
                                  sheet = "SUMMARY",
                                  skip = 4,
                                  col_types = c("date", rep("text", 23))) |> 
  glimpse()

mill_2011 <- mill_2011_raw |> 
  janitor::clean_names() |> 
  select(date, avg_water_temp = average_water_temp_mch, avg_flow_in_cfs = avg_flow_mch_low, number_salmon_up, number_salmon_down) |> 
  mutate(across(avg_water_temp:number_salmon_down, as.numeric)) |> 
  pivot_longer(number_salmon_up:number_salmon_down,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "number_salmon_up", "up", "down")) |> 
  glimpse()
```
### 2012-2013
- this moves to a new format for data collection
- There is a tab `Post R MCVS spring 2013` which is summed by day and has `original.counts` vs `consensus.counts` and `adjusted.counts`
- we are using the tab `DATA ENTRY` which has raw counts with `passage_direction`, `time`, and `view_adjust`.
- no covariates

```{r, message = FALSE, warning = FALSE}
mill_2012_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2012-2013 as of 9-22-13 .xls"), 
                                  sheet = "DATA ENTRY",
                                  skip = 4) |> 
  glimpse()

mill_2012 <- mill_2012_raw |> 
  janitor::clean_names() |>
  mutate(up_salmon = as.numeric(up_salmon),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_salmon,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "number_salmon_up", "up", "down"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

### 2013-2014
```{r, message = FALSE, warning = FALSE}
mill_2013_raw <- readxl::read_xls(paste0(save_to_prefix, "MCVS 2013-2014 as of 11-24-2014.xls"), 
                                  sheet = "DATA ENTRY",
                                  skip = 4) |> 
  glimpse()

mill_2013 <- mill_2013_raw |> 
  janitor::clean_names() |>
  mutate(up_salmon = as.numeric(up_salmon),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_salmon,
               names_to = "passage_direction", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(passage_direction == "number_salmon_up", "up", "down"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

### 2014-2015
- includes `jack` now

```{r, message = FALSE, warning = FALSE}
mill_2014_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2014-2015 as of  7-3-2015.xlsx"), 
                                  sheet = "DATA ENTRY",
                                  skip = 3) |> 
  glimpse()

mill_2014 <- mill_2014_raw |> 
  janitor::clean_names() |>
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```
### 2015-2016

```{r, message = FALSE, warning = FALSE}
mill_2015_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2015-2016 as of  4-25-2017.xlsx"), 
                                  sheet = "DATA ENTRY",
                                  skip = 5) |> 
  glimpse()

mill_2015 <- mill_2015_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

### 2016-2017

```{r, message = FALSE, warning = FALSE}
mill_2016_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2016-2017 as of  9-25-18.xlsx"), 
                                  sheet = "DATA ENTRY 16-17",
                                  skip = 5) |> 
  glimpse()

mill_2016 <- mill_2016_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```

### 2017-2018

```{r, message = FALSE, warning = FALSE}
mill_2017_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2017-2018 as of  9-14-18.xlsx"), 
                                  sheet = "MILL-17-18 ",
                                  skip = 5) |> 
  glimpse()

mill_2017 <- mill_2017_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

### 2018-2019

```{r, message = FALSE, warning = FALSE}
mill_2018_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2018-2019 as of 3-19-20.xlsx"), 
                                  sheet = "MILL-18-19",
                                  skip = 5) |> 
  glimpse()

mill_2018 <- mill_2018_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks_less_24, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks_less_24, dn_jacks_less_24, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks_less_24,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

### 2019-2020

```{r, message = FALSE, warning = FALSE}
mill_2019_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2019-20 as of  9-9-21.xlsx"), 
                                  sheet = "MILL-19-20",
                                  skip = 5) |> 
  glimpse()

mill_2019 <- mill_2019_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks_less_24, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks_less_24, dn_jacks_less_24, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks_less_24,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

### 2020-2021

```{r, message = FALSE, warning = FALSE}
mill_2020_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2020-21 as of  3-3-23.xlsx"), 
                                  sheet = "MILL-20-21",
                                  skip = 3) |> 
  glimpse()

mill_2020 <- mill_2020_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks_less_24, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks_less_24, dn_jacks_less_24, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks_less_24,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

### 2021-2022

```{r, message = FALSE, warning = FALSE}
mill_2021_raw <- readxl::read_xlsx("data-raw/qc-markdowns/adult-upstream-passage-monitoring/mill-creek/mill_creek_passage_counts_2021.xlsx", 
                                         sheet = "MILL-21-22", 
                                         skip = 5)

mill_2021 <- mill_2021_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks_less_24, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks_less_24, dn_jacks_less_24, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks_less_24,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()

```

### 2022-2023

```{r, message = FALSE, warning = FALSE}
mill_2022_raw <- readxl::read_xlsx(paste0(save_to_prefix, "MCVS 2022-23 as of 8-28-23.xlsx"), 
                                         sheet = "MILL-22-23", 
                                         skip = 5) |> 
  glimpse()

mill_2022 <- mill_2022_raw |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks_less_24, as.numeric),
         start_time = format(start_time, "%H:%M:%S")) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks_less_24, dn_jacks_less_24, viewing_adjust) |> 
  pivot_longer(up_salmon:dn_jacks_less_24,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "number_salmon_up", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE),
         viewing_adjust = as.character(viewing_adjust)) |> 
  glimpse()
```

## Data transformations

```{r, message=FALSE, warning=FALSE}
all_years_raw <- bind_rows(mill_2006, mill_2007, mill_2008, 
                           mill_2009, mill_2010, mill_2011, 
                           mill_2012, mill_2013, mill_2014, 
                           mill_2015, mill_2016, mill_2017,
                           mill_2018, mill_2019, mill_2020,
                           mill_2021, mill_2022)

all_mill_data <- all_years_raw |> 
  select(-count_type) |> 
  mutate(viewing_adjust = case_when(viewing_adjust == 0 ~ "normal",
                                    viewing_adjust == 1 ~ "turbid",
                                    viewing_adjust == 2 ~ "weir down",
                                    viewing_adjust == 3 ~ "equipment fail",
                                    viewing_adjust == 4 ~ "DIDSON/Aris",
                                    is.na(viewing_adjust) ~ "VAKI",
                                    TRUE ~ NA),
         count = ifelse(is.na(count), 0, count)) |> 
  relocate(date, start_time, count, passage_direction, jack, viewing_adjust, avg_water_temp, avg_flow_in_cfs) |> 
  filter(year(date) > 2000) |> 
  filter(!is.na(date)) |> 
  mutate(date = as.Date(date)) |> 
  rename(flow = avg_flow_in_cfs, temperature = avg_water_temp) |> 
  glimpse()
```


## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data. 

```{r data_dictionary}
percent_na <- all_mill_data  |>
  summarise_all(list(name = ~sum(is.na(.))/length(.)))  |>
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(all_mill_data),
                          description = c("Date of sampling",
                                          "Time of obseration",
                                          "Counts of spring run Chinook",
                                          "Direction of fish passage",
                                          "Whether or not the fish was less than 24 inches (jack size)",
                                          "Reason camera needed to be adjusted", "Average water temperature", "Flow in CFS"),
                          
                          percent_na = round(percent_na$value*100)
                          
)
knitr::kable(data_dictionary)
```


## Explore `date`

Check for outlier and NA values

```{r}
summary(all_mill_data$date)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_mill_data$date))/nrow(all_mill_data), 3)*100` % of values in the `date` column are NA.

## Explore Numerical Values

### Variable:`count` 
```{r}
all_mill_data  |>
  mutate(year = as.factor(year(date)))  |> 
  # glimpse()
  ggplot(aes(x=date, y = count))+
  geom_line()+
  facet_wrap(~year, scales = "free")+
  theme_minimal()+
  labs(title = "Daily Counts From 2007 - 2023")
```


```{r}
all_mill_data  |> 
  filter(date != is.na(date))  |>
  mutate(year = as.factor(year(date)))  |>
  group_by(year)  |> 
  summarise(total = sum(count, na.rm  = TRUE))  |>
  ggplot(aes(x = year, y = total, group = 1))+
  geom_line()+
  geom_point(aes(x=year, y = total))+
  theme_minimal()+
  labs(title = "Total Annual Passage Estimate from 2007 - 2023",
       y = "Total count")
```

**Numeric Summary of count From 2007 to 2023**
```{r}
summary(all_mill_data$count)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_mill_data$count))/nrow(all_mill_data), 3)*100` % of values in the `count` column are NA.

### Variable:`flow`

Flow in cfs

```{r fig.height=8, fig.width=8}
all_mill_data  |> 
  filter(date != is.na(date))  |>
  group_by(date)  |>
  mutate(avg_flow = mean(flow, na.rm = T))  |>
  ungroup()  |> 
  mutate(year = as.factor(year(date)),
         fake_year = 1900,
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date))))  |> 
  ggplot(aes(x = fake_date, y = avg_flow, color = year)) + 
  scale_color_brewer(palette = "Dark2")+
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Water Flow (colored by year)",
       y = "Average Daily Flow", 
       x = "Date")  
```
```{r fig.height=8, fig.width=8}
all_mill_data  |> 
  mutate(year = as.factor(year(date)))  |> 
  ggplot(aes(x=flow, fill = year))+
  scale_fill_brewer(palette = "Dark2")+
  geom_histogram()+
  theme_minimal()+
  labs(title = "Distribution of Flow")
```

**Numeric Summary of flow From 2007 to 2023**
```{r}
summary(all_mill_data$flow)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_mill_data$flow))/nrow(all_mill_data), 3)*100` % of values in the `flow` column are NA.

### Variable:`temperature`

Temperature in F, convert to C below 

```{r fig.height=8, fig.width=8}
all_mill_data  |> 
  filter(date != is.na(date))  |>
  group_by(date)  |>
  mutate(avg_temp = mean(temperature, na.rm = T))  |>
  ungroup()  |> 
  mutate(year = as.factor(year(date)),
         fake_year = 1900,
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date))))  |> 
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  scale_color_brewer(palette = "Dark2")+
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Water Temperature (colored by year)",
       y = "Average Daily Temperature", 
       x = "Date")  
```

```{r fig.height=8, fig.width=8}
all_mill_data  |> 
  filter(date != is.na(date))  |>
  mutate(year = as.factor(year(date)))  |> 
  ggplot(aes(x=temperature, fill = year))+
  scale_fill_brewer(palette = "Dark2")+
  geom_histogram(bins = 10)+
  theme_minimal()+
  labs(title = "Distribution of Temperature")
```

**Numeric Summary of temperature From 2007 to 2023**
```{r}
summary(all_mill_data$temperature)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_mill_data$temperature))/nrow(all_mill_data), 3)*100` % of values in the `temperature` column are NA.

### Notes and Issues

* count drops significantly in 2023
* We are treating these as counts (in standard format, rename passage estimates to counts) - are these passage estimates for the earlier data?
* Temperature in F, convert to C below 

```{r}
all_mill_data <- all_mill_data  |>
  mutate(temperature = (temperature - 32) * (5/9))
```


### Add cleaned data back onto google cloud

```{r}
mill_upstream_counts <- all_mill_data  |> glimpse()
```

```{r, eval = FALSE}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(mill_upstream_counts,
           object_function = f,
           type = "csv",
           name = "adult-upstream-passage-monitoring/mill-creek/data/mill_upstream_counts.csv",
           predefinedAcl = "bucketLevel")
```