---
title: "Battle Creek Adult Holding Survey QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(knitr)
```

# Battle Creek Adult Holding Survey 

## Description of Monitoring Data

These data were acquired via snorkel and kayak surveys on Battle Creek from 2001 to 2019 and describe the locations of live adult Chinook Salmon seen in Battle Creek. Generally, spring-run Chinook Salmon are in holding position prior to August 15 and in spawning position after August 15.  						


**Timeframe:** 2001 - 2019

**Survey Season:** may - november

**Completeness of Record throughout timeframe:** Data for every year in timeframe, for some years there are only a few sample dates. 

**Sampling Location:** Battle Creek 

**Data Contact:** [Natasha Wingerter](mailto:natasha_wingerter@fws.gov); [RJ Bottaro](mailto:rj_bottaro@fws.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```

Read in data from google cloud, glimpse raw data sheet: 

```{r}
sheets <- excel_sheets("raw_adult_spawn_hold_carcass.xlsx")
sheets 
raw_holding_data <-read_excel("raw_adult_spawn_hold_carcass.xlsx", sheet = "Live Holding Spawning") %>% glimpse()
```

## Data transformations

```{r}
cleaner_holding_data <- raw_holding_data %>% 
  janitor::clean_names() %>%
  rename("date" = sample_date,
         "river_mile" = rivermile,
         "count" = quantity) %>%
  mutate(date = as.Date(date),
         jacks = as.numeric(jacks)) %>%
  select(-project, -year) %>% # All projects = snorkel, year can be pulled from date
  glimpse
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- cleaner_holding_data %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_holding_data),
                          description = c("GPS X point",
                                          "GPS Y point",
                                          "Date of sample",
                                          "Reach number (1-7)",
                                          "River mile number",
                                          "Total number of salmon observed",
                                          "Number of jacks seen",
                                          "General comments from survey crew"
                                          ),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_holding_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `longitude`, `latitude`

**Numeric Summary of lat long over Period of Record**

```{r}
summary(cleaner_holding_data$latitude)
summary(cleaner_holding_data$longitude)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_holding_data$latitude))/nrow(cleaner_holding_data), 3) * 100` % of values in the `latitude` column are NA.
* `r round(sum(is.na(cleaner_holding_data$longitude))/nrow(cleaner_holding_data), 3) * 100` % of values in the `longitude` column are NA.

### Variable: `river_mile`

**Plotting river mile over Period of Record**

```{r}
cleaner_holding_data %>% 
  ggplot(aes(x = river_mile, y = year(date))) +
  geom_point(alpha = .75, size = 1.5, color = "blue") + 
  labs(y = "Year", x = "River Mile") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

It looks like river miles 0 - 6 and 12 - 17 most commonly have holding chinooks. In most recent years there are less points and many of the holding chinooks are before mile 5. 

```{r}
cleaner_holding_data %>% 
  ggplot(aes(x = river_mile)) +
  geom_histogram(alpha = .75) + 
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Numeric Summary of river mile over Period of Record**

```{r}
summary(cleaner_holding_data$river_mile)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_holding_data$river_mile))/nrow(cleaner_holding_data), 3) * 100` % of values in the `river_mile` column are NA.

### Variable: `count`

**Plotting  Counts over Period of Record**
```{r}
cleaner_holding_data %>% 
  mutate(year = as.factor(year(date)),
         fake_date = as.Date(paste0("1990", "-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = count)) + 
  geom_col() + 
  facet_wrap(~year(date), scales = "free_y") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 23),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Passage Count", 
       x = "Date")  

```


```{r}
cleaner_holding_data %>% 
  group_by(date) %>%
  summarise(daily_count = sum(count)) %>%
  filter(daily_count < 100) %>%
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = year, y = daily_count)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Daily Count Sumarized by Year") + 
  theme(text = element_text(size = 23),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))   + 
  scale_y_continuous(limits = c(0, 101))
```

```{r}
cleaner_holding_data  %>%
  mutate(year = as.factor(year(date))) %>%
  group_by(year = year(date)) %>%
  summarise(total_catch = sum(count, na.rm = T)) %>%
  ungroup() %>%
  ggplot(aes(x = year, y = total_catch)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Yearly Fish Count",
       y = "Total fish count") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of Passage Counts over Period of Record**

```{r}
# Table with summary statistics 
summary(cleaner_holding_data$count)

# Daily numeric summary of passage data
cleaner_holding_data %>% group_by(date) %>%
  summarise(count = sum(count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_holding_data$count))/nrow(cleaner_holding_data), 3) * 100` % of values in the `count` column are NA. 

### Variable: `jacks`

**Plotting distribution of jacks**

Number of jacks seen				


```{r}
cleaner_holding_data %>% 
  ggplot(aes(x = jacks)) +
  geom_histogram(bins = 4) +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

Most of the jacks are 0

**Numeric Summary of jacks over Period of Record**

```{r}
summary(cleaner_holding_data$jacks)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_holding_data$jacks))/nrow(cleaner_holding_data), 3) * 100` % of values in the `jacks` column are NA. 

## Explore Categorical variables: {.tabset}


```{r}
cleaner_holding_data %>% select_if(is.character) %>% colnames()
```

### Variable: `reach`

The stream section

```{r}
table(cleaner_holding_data$reach) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_holding_data$reach))/nrow(cleaner_holding_data), 3) * 100` % of values in the `reach` column are NA. 

### Variable: `notes`

```{r}
unique(cleaner_holding_data$notes)[1:5]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_holding_data$notes))/nrow(cleaner_holding_data), 3) * 100` % of values in the `notes` column are NA. 

## Summary of identified issues

* Looks like a few surveys done every year

## Next steps

### Columns to remove

- `notes` may not be needed.
- if `latitude` and `longitude` are filled in then may not need `reach` or `river_mile`

## Save cleaned data back to google cloud 

```{r}
battle_holding <- cleaner_holding_data %>% select(date, reach, river_mile, count, jacks, latitude, longitude, notes) %>%glimpse()
```

```{r, eval=FALSE}
gcs_list_objects()
f <- function(input, output) write_csv(input, file = output)
gcs_upload(battle_holding,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_holding.csv")
```