---
title: "mill-creek-redd-survey-qc-checklist"
author: "Inigo Peng"
date: "10/19/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
library(knitr)
```
# Mill Creek Redd Survey Data 

**Description of Monitoring Data**

* Redd surveys Buckhorn Gulch to Upper Dam are conducted by air (helicopter).  All other survey sections completed on foot

**Timeframe:** 

1997 to 2023

**Completeness of Record throughout timeframe:**

* Missing some elevation data
* Buckhorn Gulch To Upper Dam missing significant count data
* Canyon Camp to Sooner Place missing redd count data from 1997 - 2006
* McCarthy Place to Savercool Place and Savercool Place to Black Rock reaches combined 1997 to 2006

**Sampling Location:** 

* Above Hwy 36
* Hwy 36 to Little Hole-in-Ground
* Litte Hole-in-Ground to Hole-in-Ground
* Hole-in-Ground to Ishi Trail Head   
* Ishi Trail Head to Big Bend   
* Big Bend to Canyon Camp  
* Canyon Camp to Sooner Place
* Sooner Place to McCarthy Place
* McCarthy Place to Savercool Place
* Savercool Place  to Black Rock
* Black Rock to below Ranch House
* Below Ranch House to above Avery   
* Above Avery to Pape Place
* Pape Place to Buckhorn Gulch 
* Buckhorn Gulch to Upper Dam (aerial survey)

**Data Contact:** [Matt Johnson](mailto:Matt.Johnson@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# original data
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/mill-creek/data-raw/Mill Creek SRCS Redd Counts by Section 1997-2020 Reformatted.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data-raw", "qc-markdowns","adult-holding-redd-and-carcass-surveys", "mill-creek",
                                       "mill_creek_redd_raw.xlsx"))
               # Overwrite = TRUE)

# updated data (from 4-30-2024)
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/mill-creek/data-raw/Mill Creek spring-run Redd Counts 2007-2023.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/mill-creek/mill_creek_snorkel_2007-2023_raw.xlsx")
```

### read in original data (1997-2020)
```{r}
raw_data <- readxl::read_excel(here::here("data-raw", "qc-markdowns","adult-holding-redd-and-carcass-surveys", "mill-creek",
                                       "mill_creek_redd_raw.xlsx"),
                              sheet = '2007-2020 SR redds') |> 
  glimpse()
```

### read in updated data (2007-2023)
```{r}
raw_data_2023 <- readxl::read_excel(here::here("data-raw", "qc-markdowns","adult-holding-redd-and-carcass-surveys", "mill-creek",
                                       "mill_creek_snorkel_2007-2023_raw.xlsx"),
                              sheet = '2007-2023 SR redds',
                              skip = 1) |> 
  glimpse()
```

## Data Transformations

```{r}
# get 1997-2006
cleaner_data_early <- raw_data |> 
  select(location = Section, starting_elevation = `Starting Elevevation`,
         seq(3, 21, 2)) |> 
  filter(!location %in% c("Total Redds Counted:", NA)) |> 
  pivot_longer(3:12, 
               names_to = "year",
               values_to = "redd_count") |> 
  mutate(year = readr::parse_number(year),
         surveyed = ifelse(redd_count == "n/s", FALSE, TRUE),
         redd_count = ifelse(surveyed & is.na(redd_count), 0, as.numeric(redd_count)),
         starting_elevation = as.numeric(str_remove(starting_elevation, "'"))) |> 
  glimpse()

cleaner_data_raw <- raw_data_2023 |> 
  rename(location = `...1`) |> 
  select(-c(seq(2, 36, 2))) |>
  mutate_at(c(2:18), suppressWarnings(as.numeric)) |> 
  filter(!location %in% c("Section", "Total Redds Observed:", NA)) |>
  pivot_longer(2:18,
  names_to = "year",
  values_to = "redd_count") |> 
  mutate(year = readr::parse_number(year)) |> 
  glimpse()

cleaner_data <- bind_rows(cleaner_data_early, cleaner_data_raw) |> 
  mutate(location = str_replace_all(location, " Place", ""),
         location = str_replace_all(location, "HIG", "Hole-in-Ground"),
         location = case_when(location == "Savercool  to Black Rock" ~ "Savercool to Black Rock",
         TRUE ~ location)) |> 
  glimpse()
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- cleaner_data |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_data),
                          description = c("Survey reach",
                                          "Survey year",
                                          "Number of redds observed",
                                          "Elevation at sample start point",
                                          "Whether or not a reach was surveyed for a given year"),
                          data_type = c("character",
                                        "year",
                                        "numeric",
                                        "numeric",
                                        "logical"),
                          encoding = c(NA,
                                       NA,
                                       NA,
                                       NA, NA),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
```

## Explore Categorical Variables

```{r}
table(cleaner_data$location)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$location))/nrow(cleaner_data), 3)*100` % of values in the `location` column are NA.

## Explore Numeric Variables

### Variable `redd_count`
```{r fig.height=7, fig.width=10}
#Find the most distinctive colours for visual
colourCount = length(unique(cleaner_data$location))
getPalette = colorRampPalette(brewer.pal(12, "Paired"))
cleaner_data |>
  mutate(date =lubridate::ymd(year, truncated = 2L), .keep = "unused") |> 
  mutate(year = as.factor(year(date))) |> 
  ggplot(aes(x = year, y = redd_count, fill = location))+
  scale_fill_manual(values = getPalette(colourCount))+
  geom_col()+
  theme_minimal()+
  labs(title = "Adult Redd Count By Year")+
  ylim(0, 800)+
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5))
```
```{r}
cleaner_data |> 
  ggplot(aes(x = redd_count, y = location))+
  geom_boxplot() +
  theme_minimal()+
  labs(title = "Redd Count By Reach")
```

**Numeric Annual Summary of redd_count From 1997 to 2020**
```{r}
cleaner_data |>
  group_by(year) |>
  summarise(count = sum(redd_count, na.rm = T)) |>
  pull(count) |>
  summary()
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$redd_count))/nrow(cleaner_data), 3)*100` % of values in the `redd_count` column are NA.

### Variable `starting_elevation`

```{r fig.height=7, fig.width=10}
cleaner_data |>
  filter(!is.na(starting_elevation)) |> 
  ggplot(aes(y = starting_elevation, x = location))+
  geom_col() +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, size = 7))+
  labs(title = "Starting Elevation By Reach")
```

**Numeric Summary of starting_elevation From 1997 to 2020**
```{r}
cleaner_data |>
  summarise(starting_elevation) |>
  pull(starting_elevation) |>
  summary()
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$starting_elevation))/nrow(cleaner_data), 3)*100` % of values in the `starting_elevation` column are NA.

## Next steps

- Acquire starting elevation for all reaches

### Add cleaned data back onto google cloud
```{r}
mill_redd_survey <- cleaner_data |> glimpse()
```
```{r, include = F}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(mill_redd_survey,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/mill-creek/data/mill_redd_survey.csv")
```
