---
title: "Create standard format for Holding data"
output: html_document
---
```{r, include = F}
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(knitr)
library(hms)


Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")
```

Data were checked and cleaned in scripts available [here]()
Cleaned data were saved on the jpe-dev-bucket on google cloud.

```{r, data_pull, include = F, echo = F}
# # Data pull ---------------------------------------------------------------

# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# define files to pull in (in some cases there is just one catch file while others
# have sample and environmental data)
files_battle <- tibble(path = rep("adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_holding"),
                       name = c(""),
                       save = rep("data/redd-carcass-holding/battle_holding"))
files_butte <- tibble(path = rep("adult-holding-redd-and-carcass-surveys/butte-creek/data/butte_",17),
                       name = c("holding_2001",
                                "holding_2002", "holding_2003", "holding_2004", "holding_2005", "holding_2006",
                                "holding_2007", "holding_2008", "holding_2009", "holding_2010", "holding_2011",
                                "holding_2012", "holding_2013", "holding_2014", "holding_2015", "holding_2016",
                                "holding_2017"),
                      save = "data/redd-carcass-holding/butte_")
files_clear <- tibble(path = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_holding",
                      name = c(""),
                      save = rep("data/redd-carcass-holding/clear_holding"))
files_deer <- tibble(path = rep("adult-holding-redd-and-carcass-surveys/deer-creek/data/deer_",2),
                     name = c("adult_holding_1986_to_1996",
                              "adult_holding_1997_to_2020"),
                     save = "data/redd-carcass-holding/deer_")
# function to save file to disk
get_data <- function(path, name, save) {
  gcs_get_object(object_name = paste0(path, name, ".csv"),
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = paste0(save, name, ".csv"),
                 overwrite = TRUE)
}
# apply function to each set of files
pmap(files_battle, get_data)
pmap(files_butte, get_data)
pmap(files_clear, get_data)
pmap(files_deer, get_data)


# load in data
# catch data
battle_holding <- read_csv("data/redd-carcass-holding/battle_holding.csv")
butte_holding_01 <- read_csv("data/redd-carcass-holding/butte_holding_2001.csv")
butte_holding_02 <- read_csv("data/redd-carcass-holding/butte_holding_2002.csv")
butte_holding_03 <- read_csv("data/redd-carcass-holding/butte_holding_2003.csv")
butte_holding_04 <- read_csv("data/redd-carcass-holding/butte_holding_2004.csv")
butte_holding_05 <- read_csv("data/redd-carcass-holding/butte_holding_2005.csv")
butte_holding_06 <- read_csv("data/redd-carcass-holding/butte_holding_2006.csv")
butte_holding_07 <- read_csv("data/redd-carcass-holding/butte_holding_2007.csv")
butte_holding_08 <- read_csv("data/redd-carcass-holding/butte_holding_2008.csv")
butte_holding_09 <- read_csv("data/redd-carcass-holding/butte_holding_2009.csv")
butte_holding_10 <- read_csv("data/redd-carcass-holding/butte_holding_2010.csv")
butte_holding_11 <- read_csv("data/redd-carcass-holding/butte_holding_2011.csv")
butte_holding_12 <- read_csv("data/redd-carcass-holding/butte_holding_2012.csv")
butte_holding_13 <- read_csv("data/redd-carcass-holding/butte_holding_2013.csv")
butte_holding_14 <- read_csv("data/redd-carcass-holding/butte_holding_2014.csv")
butte_holding_15 <- read_csv("data/redd-carcass-holding/butte_holding_2015.csv")
butte_holding_16 <- read_csv("data/redd-carcass-holding/butte_holding_2016.csv")
butte_holding_17 <- read_csv("data/redd-carcass-holding/butte_holding_2017.csv")
clear_holding <- read_csv("data/redd-carcass-holding/clear_holding.csv")
deer_holding_1997_2020 <- read_csv("data/redd-carcass-holding/deer_adult_holding_1997_to_2020.csv")
deer_holding_1986_1996 <- read_csv("data/redd-carcass-holding/deer_adult_holding_1986_to_1996.csv")

```

# Data formatting {.tabset}

## Battle Creek

```{r, battle_holding_data_dictionary}
battle_data_dictionary <- read_rds("data/redd-carcass-holding/battle_holding_data_dictionary.rds")
kable(battle_data_dictionary)
```
```{r, battle_clean}
battle_holding_clean <- battle_holding %>%
  select(-notes) %>%
  mutate(watershed = "Battle Creek",
         year = lubridate::year(date))
  

```

### Varibles removed
- notes

## Clear Creek

```{r, clear_holding_data_dictionary}
clear_data_dictionary <- read_rds("data/redd-carcass-holding/clear_holding_data_dictionary.rds")
kable(clear_data_dictionary)
```
```{r, clear_clean}
clear_holding_clean <- clear_holding %>%
  select(-c(comments, picket_weir_location_rm, picket_weir_relate)) %>%
  rename(jacks = jack_count) %>%
  mutate(watershed = "Clear Creek",
         year = lubridate::year(date))
```

### Variables removed
- comments, picket_weir_location_rm, picket_weir_relate

## Deer Creek

```{r, deer_holding_data_dictionary}
deer_data_dictionary <- read_rds("data/redd-carcass-holding/deer_holding_data_dictionary.rds")
kable(deer_data_dictionary)
```

```{r, deer_clean}
deer_holding_clean <- bind_rows(deer_holding_1986_1996,
                                deer_holding_1997_2020) %>%
  rename(reach = location) %>%
  mutate(watershed = "Deer Creek")
```

### Variables removed
- none

## Butte Creek

```{r}
# percent_na <- bind_rows(butte_holding_01 %>% select(-comments),
#                                  butte_holding_02,
#                                  butte_holding_03,
#                                  butte_holding_04,
#                                  butte_holding_05,
#                                  butte_holding_06,
#                                  butte_holding_07, 
#                                  butte_holding_08,
#                                  butte_holding_09,
#                                  butte_holding_10,
#                                  butte_holding_11,
#                                  butte_holding_12,
#                                  butte_holding_13,
#                                  butte_holding_14,
#                                  butte_holding_15,
#                                  butte_holding_16,
#                                  butte_holding_17) %>% 
#   summarise_all(list(name = ~sum(is.na(.)/length(.)))) %>% 
#                   pivot_longer(cols = everything())

# butte_data_dictionary <- tibble(variables = colnames(butte_holding_02), description = c("Date of Sampling", "Reach within watershed", "Name of survey person", "Fish count", "Comments why count is NA"), percent_na = round(percent_na$value*100))
# saveRDS(butte_data_dictionary, file = "butte_holding_data_dictionary.rds")
```

```{r, butte_holding_data_dictionary}
butte_data_dictionary <- read_rds("data/redd-carcass-holding/butte_holding_data_dictionary.rds")
kable(butte_data_dictionary)
```
```{r, butte_clean}
butte_holding_clean <- bind_rows(butte_holding_01 %>% select(-comments),
                                 butte_holding_02,
                                 butte_holding_03,
                                 butte_holding_04,
                                 butte_holding_05,
                                 butte_holding_06,
                                 butte_holding_07, 
                                 butte_holding_08,
                                 butte_holding_09,
                                 butte_holding_10,
                                 butte_holding_11,
                                 butte_holding_12,
                                 butte_holding_13,
                                 butte_holding_14,
                                 butte_holding_15,
                                 butte_holding_16,
                                 butte_holding_17) %>%
  group_by(date, reach) %>%
  summarize(avg_count = mean(fish_count, na.rm = T)) %>%
  rename(count = avg_count) %>%
  mutate(watershed = "Butte Creek",
         year = lubridate::year(date)) %>% 
  glimpse()
```

### Variables removed
- comments, personnel removed

# Standard format {.tabset}
```{r, combined}
combined_holding <- bind_rows(battle_holding_clean,
                              butte_holding_clean,
                              clear_holding_clean,
                              deer_holding_clean)
```

## Data dictionary

```{r, data_dictionary_prep, include = F}
colnames(combined_holding)
holding_catch_data_dictionary <- tibble(variable_name = colnames(combined_holding), description = c("Date survey was conducted", "Reach number within watershed", "River mile number", "Number of fish observed", "Number of jacks observed", "Latitude measurement of adult holding", "Longitude measurement of adult holding", "Watershed name", "Reason survey was conducted", "Year survey was conducted"), encoding = c("YYYY-MM-DD", NA,NA,NA,NA,NA,NA,NA,"august index, spawning","YYYY" ), collected = c("All but Deer", "All","Battle and Clear", "All", "Battle and Clear", "Battle and Clear", "Battle and Clear", "All", "Clear", "All"))
```
```{r, data_dictionary}
kable(holding_catch_data_dictionary)
```

## Exploratory plots

Total count per year for each waterbody

```{r}
combined_holding %>% 
  group_by(year, watershed) %>%
  ggplot(aes(x = year, y = count, fill = watershed))+
  geom_col()
```

Average count per month. This chart shows only Battle, Butte and Clear as Deer does not have monthly count.

```{r}
combined_holding %>% 
  mutate(month = lubridate::month(date)) %>% 
  group_by(month, watershed) %>% 
  summarize(average_count = mean(count, na.rm = T)) %>%
  ggplot(aes(x = month, y = average_count, fill = watershed)) +
  geom_col()
```

Average count of jacks per month. Jacks is only collected at Battle and Clear.
```{r}
combined_holding %>% 
  mutate(month = lubridate::month(date)) %>% 
  group_by(month, watershed) %>% 
  summarize(average_jacks = mean(jacks, na.rm = T)) %>% 
  ggplot(aes(x = month, y = average_jacks, fill = watershed)) + 
  geom_col()
```

# QA/QC {.tabset}

# Count

Monthly count is not recorded for Deer.

Figure below shows the percentage count by month for each waterbody except Deer.
```{r, count_qaqc}

total_count <- combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(watershed) %>% 
  summarize(total = sum(count, na.rm = T))

combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(watershed, month) %>% 
  summarize(count = sum(count, na.rm = T)) %>% 
  left_join(total_count) %>% 
  mutate(percent = (count/total)*100) %>% 
  ggplot(aes(x= month, y = percent, fill = watershed)) +
  geom_col() +
  facet_wrap(~watershed, scales = "free_y")
```

*  `r round(sum(is.na(combined_holding$count))/nrow(combined_holding), 3)*100` % of values in the `count` column are NA.

## Jacks
Jacks are not recorded for Deer and Butte

Figure below whos the percentage count by month for Battle and Clear.

```{r, jack_qa_qc}
total_jacks <- combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(watershed) %>% 
  summarize(total = sum(jacks, na.rm = T))

combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(watershed, month) %>% 
  summarize(count = sum(jacks, na.rm = T)) %>% 
  left_join(total_jacks) %>% 
  mutate(percent = (count/total)*100) %>% 
  ggplot(aes(x= month, y = percent, fill = watershed)) +
  geom_col() +
  facet_wrap(~watershed, scales = "free_y")
```

*  `r round(sum(is.na(combined_holding$jacks))/nrow(combined_holding), 3)*100` % of values in the `jacks` column are NA.

# Save data
```{r, save_data}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(combined_holding,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_holding.csv",
           predefinedAcl = "bucketLevel")

```

