---
title: "rst_eda"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)

combined_rst <- read_rds("data/rst/combined_rst.rds")

combined_trap <- read_rds("data/rst/combined_trap.rds")
```

```{r}
rst_prep <- combined_rst %>%
  mutate(fake_year = ifelse(month(date) %in% 10:12, 1999, 2000),
         fake_date = ymd(paste(fake_year, month(date), day(date))))


```

# Daily catch by watershed

notes

- consider adding color for run (spring run higher alpha)
- filter out hatchery?


```{r}
# data prep
daily_catch_dat <- rst_prep %>%
  filter(species == "chinook" | is.na(species)) %>%
  group_by(watershed, date, site, fake_date) %>%
  summarize(count = sum(count))

# plot function
daily_catch_bar_chart <- function(dat, watershed_filter) {
  ggplot(filter(dat, watershed == watershed_filter), aes(x = fake_date, y = count, fill = site)) +
    geom_col(alpha = 0.75) +
    facet_wrap(~year(date), scales = "free_y") +
    scale_x_date(date_breaks = "3 months", date_labels = "%b") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    xlab("")
}

# plots for each watershed
daily_catch_bar_chart(daily_catch_dat, "Battle Creek")
daily_catch_bar_chart(daily_catch_dat, "Butte Creek")
daily_catch_bar_chart(daily_catch_dat, "Clear Creek")
daily_catch_bar_chart(daily_catch_dat, "Deer Creek")
daily_catch_bar_chart(daily_catch_dat, "Feather River")
# There are 5 cases where the site is NA for Tisdale
daily_catch_bar_chart(filter(daily_catch_dat, site != "Tisdale NA"), "Lower Sac")
daily_catch_bar_chart(daily_catch_dat, "Mill Creek")
daily_catch_bar_chart(daily_catch_dat, "Yuba River")
```

# Daily number of traps with non zero catch by watershed

```{r}
site_number <- daily_catch_dat %>%
  filter(count > 0) %>%
  group_by(watershed, date, fake_date) %>%
  tally() %>%
  rename(number_sites = n) 

site_number_table <- site_number %>%
  group_by(watershed, year(date)) %>%
  summarize(mean_number_sites = mean(number_sites))

ggplot(site_number_table, aes(x = `year(date)`, y = mean_number_sites)) +
  geom_col() +
  facet_wrap(~watershed) +
  xlab("")
```

# Mean weekly fork length by year and watershed

TODO: figure out knights landing min max fork length

notes

- consider adding color for run (spring run higher alpha)
- filter out hatchery?

```{r}
# Summarize the mean fork length by week
weekly_fork_length <- rst_prep %>%
  filter(species == "chinook") %>%
  mutate(week = week(date),
         year = year(date)) %>%
  group_by(watershed, week, year, site) %>%
  summarize(avg_fork_length = mean(fork_length, na.rm = T),
            fake_date = min(date),
            fake_year = ifelse(month(date) %in% 10:12, 1999, 2000),
            fake_date = ymd(paste(fake_year, month(fake_date), day(fake_date)))) %>%
  glimpse

weekly_fork_length_bar_chart <- function(dat, watershed_filter) {
  ggplot(filter(dat, watershed == watershed_filter), aes(x = fake_date, y = avg_fork_length, fill = site)) +
    geom_col(alpha = 0.7) +
    facet_wrap(~year) +
    scale_x_date(date_breaks = "3 months", date_labels = "%b") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    xlab("") +
    ylab("Mean weekly fork length (mm)")
}

weekly_fork_length_bar_chart(weekly_fork_length, watershed_filter = "Battle Creek")
weekly_fork_length_bar_chart(weekly_fork_length, watershed_filter = "Butte Creek")
weekly_fork_length_bar_chart(filter(weekly_fork_length, site != "Tisdale NA"), "Lower Sac")
```

# Hours fished boxplot for each year and watershed

```{r}
# Method to calculated hours fished #
# We can only calculate when we have the start and stop times
# 1) Add 24 to start_time and substract end_time 
# 2) Create datetime variable and subtract

# Used method 2. 
# Assumption - for data without start date assume that prior day is the start date
# and current day is end date

#TODO lots of errors with this code
# there are some cases where the lag doesn't apply - for instance trap date in dec and then next in march. assumed it stopped in this time
# these will be removed from the summary plots
trap_prep <- combined_trap %>%
  mutate(date_start_date = case_when(is.na(start_date) ~ date, 
                                     T ~ start_date),
         wy = ifelse(month(date_start_date) %in% 10:12, year(date_start_date) + 1, year(date_start_date))) %>%
  group_by(wy, watershed, site, gear_id, cone_id) %>%
  arrange(start_date, start_time) %>%
  mutate(start_date = case_when(is.na(start_date) ~ lag(date),
                                T ~ start_date),
         start_date = as.Date(start_date),
         start_time = case_when(is.na(start_time) ~ lag(time),
                                T ~ start_time),
         end_date = case_when(is.na(end_date) ~ date,
                              T ~ end_date),
         end_time = case_when(is.na(end_time) ~ time,
                              T ~ end_time),
         end_date = as.Date(end_date),
         start_datetime = case_when(!is.na(start_date) & !is.na(start_time) ~ ymd_hms(paste(start_date, start_time)),
                                                                                      T ~ NA_real_),
         end_datetime = case_when(!is.na(end_date) & !is.na(end_time) ~ ymd_hms(paste(end_date, end_time)),
                                                                                T ~ NA_real_),
         hrs_fished = case_when(is.na(hrs_fished) ~ difftime(end_datetime, start_datetime, units = "hours"),
                                T ~ hrs_fished),
         hrs_fished = as.numeric(hrs_fished),
         wy = as.factor(wy))


```


```{r}

ggplot(filter(trap_prep, watershed == "Battle Creek", hrs_fished < 1000), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()

# Butte looks weird, a lot with hours less than 24 - TODO why
# Butte is grouped by site and gear_id, in some cases gear_id is NA which messes up the start/date hrs etc.
# Remove gear_id NA for this summary
ggplot(filter(trap_prep, watershed == "Butte Creek", hrs_fished < 100, !is.na(gear_id)), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()
ck <- filter(trap_prep, watershed == "Butte Creek", wy == 1998, gear_id == "rotary screw trap 2")

ggplot(filter(trap_prep, watershed == "Clear Creek", hrs_fished < 100), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()

# Can't calculate deer because they do not collect time. TODO double check this
ggplot(filter(trap_prep, watershed == "Deer Creek", hrs_fished < 1000), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()

# Lot of variance for Feather. TODO Look into this more.
ggplot(filter(trap_prep, watershed == "Feather River", hrs_fished < 250), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()

# Lot of variance for Lower Sac. TODO look into this more.
ggplot(filter(trap_prep, watershed == "Lower Sac", hrs_fished < 100), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()

# Can't calculate mill because they do not collect time. TODO double check this
ggplot(filter(trap_prep, watershed == "Mill Creek", hrs_fished < 100), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()

ggplot(filter(trap_prep, watershed == "Yuba River", hrs_fished < 100), aes(x = wy, y = hrs_fished)) +
  geom_boxplot()
```
