---
title: "checking date format"
output: html_document
date: "2022-08-09"
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(lubridate)
library(googleCloudStorageR)
library(tidyverse)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

Checking date format between standard_catch and standard_flow

# QC Checks {.tabset}

## Standard flow

Date is formatted as a date without hms.

```{r}
flow_raw <- read_csv(here::here("data","standard-format-data", "standard_flow.csv")) %>% glimpse
```


## Standard catch

Date is formatted as POSIXct

```{r}
catch_raw <- read_csv(here::here("data","standard-format-data", "standard_catch.csv")) %>% glimpse
```

## Standard trap

Date is formatted as POSIXct
```{r}
trap_raw <- read_csv(here::here("data","standard-format-data", "standard_trap.csv")) %>% glimpse
```
## Standard environmental

Date formatted as date.

```{r}
environment_raw <- read_csv(here::here("data","standard-format-data", "standard_environmental.csv")) %>% glimpse
```
## Standard recapture

Date formatted as POSIXct.

```{r}
recapture_raw <- read_csv(here::here("data","standard-format-data", "standard_recapture.csv")) %>% glimpse
```
## Standard release

Date formatted as POSIXct.

```{r}
release_raw <- read_csv(here::here("data","standard-format-data", "standard_release.csv")) %>% glimpse
```

# Fix

Identified fix is to format standard_catch date as date rather than POSIXct. This
will be done in the rst_catch_standard_format.Rmd. This is the Rmd that generates
the standard format date so best to keep all fixes in one place.

Check to make sure fix worked.
```{r}
gcs_get_object(object_name = "standard-format-data/standard_rst_catch.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard-format-data/standard_catch.csv",
               overwrite = TRUE)

catch_raw <- read_csv(here::here("data","standard-format-data", "standard_catch.csv")) %>% glimpse
```


Additionally check standard_trap. Decided to format trap_start_date and trap_stop_date
as date rather than POSIXct. 

Checked standard recapture. Format recaptured_date to date rather than POSIXct.

Checked standard release. Format release_date to date rather than POSIXct.