---
title: "Estimating number of fish needed for efficiency trials"
output: html_document
date: '2022-07-25'
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(lubridate)
library(googleCloudStorageR)
library(waterYearType)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# water year type
sac_wy <- CDECRetrieve::water_year_indices %>% 
  filter(location == "Sacramento Valley") %>% 
  select(wy, yr_type) %>%
  add_row(wy = 2020, yr_type = "Dry") %>%
  add_row(wy = 2021, yr_type = "Dry")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_get_object(object_name = "standard-format-data/standard_recapture.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard-format-data/standard_recapture.csv",
               overwrite = TRUE)

gcs_get_object(object_name = "standard-format-data/standard_release.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard-format-data/standard_release.csv",
               overwrite = TRUE)

gcs_get_object(object_name = "standard-format-data/standard_rst_catch.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard-format-data/standard_rst_catch.csv",
               overwrite = TRUE)
# 
# gcs_get_object(object_name = "standard-format-data/standard_rst_catch_lad.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "data/standard-format-data/standard_rst_catch_lad.csv",
#                overwrite = TRUE)

release_raw <- read_csv(here::here("data","standard-format-data", "standard_release.csv"))
recapture_raw <- read_csv(here::here("data","standard-format-data", "standard_recapture.csv"))

# Load in standard RST data (with river model run)
catch_raw <- read_csv(here::here("data", "standard-format-data", "standard_rst_catch.csv")) %>%
  mutate(stream = case_when(site == "knights landing" ~ "knights landing",
                            site == "tisdale" ~ "tisdale",
                            T ~ stream))
```

# Number of fish needed per trial

```{r, include = F, echo = F}
number_released_flow <- release_raw %>% 
  select(stream, site, release_id, release_date, number_released, flow_at_release)

number_recaptured <- recapture_raw %>% 
  mutate(number_recaptured = ifelse(is.na(number_recaptured), 0, number_recaptured)) %>% 
  group_by(stream, release_id) %>% 
  summarize(number_recaptured = sum(number_recaptured))

# Use average efficiency by stream for those with historical estimates
efficiency_stream <- left_join(number_released_flow, number_recaptured) %>% 
  mutate(efficiency = number_recaptured/number_released) %>% 
  # filter(efficiency > 0, efficiency < 1) %>% 
  group_by(stream) %>% 
  summarize(mean_efficiency = mean(efficiency, na.rm = T)) 

# Use global mean for those without historical estimates
# efficiency_global <- left_join(number_released_flow, number_recaptured) %>% 
#   mutate(efficiency = number_recaptured/number_released) %>% 
#   summarize(mean_efficiency = mean(efficiency, na.rm = T),
#             median_efficiency = median(efficiency, na.rm = T)) %>% 
#   mutate(stream = "global")

efficiency_global_avg <- efficiency_stream %>% 
  summarize(mean_efficiency = mean(mean_efficiency)) %>% 
  mutate(stream = "global")

historic_efficiency <- bind_rows(efficiency_global_avg, efficiency_stream) %>% 
  mutate(mean_efficiency = round(mean_efficiency, 3))
```


```{r, echo = F}
# Number of fish needed per trial
number_released <- function(efficiency) {
  cv_z <- 0.2
  p <- efficiency
  N <- ((1-p)/((cv_z^2)*p))
  N <- round(N, 0)
  N
}

efficiency <- historic_efficiency$mean_efficiency
streams <- historic_efficiency$stream

number_released_table <- tibble(stream = streams,
                                historic_efficiency = efficiency,
                                release_per_trial = c(485, 343, 327, 438, 6225))

kable(number_released_table)
```

# Weekly historic fish

```{r}
# Summarize catch daily ---------------------------------------------------

daily_catch <- catch_raw %>% 
  group_by(stream, date, site) %>% 
  # sum fish across fork lengths and runs
  summarize(count = sum(count)) %>% 
  ungroup() %>% 
  group_by(stream, date) %>%
  # handling multiple sites: take max catch on a given day
  summarize(count = max(count))

# Summarize catch weekly  --------------------------------------------

weekly_catch <- daily_catch %>% 
  mutate(week = week(date),
         year = year(date),
         month = month(date),
         wy = ifelse(week %in% 40:53, year + 1, year)) %>% 
  group_by(stream, week, wy, month) %>% 
  # sum count across days in week to weekly counts
  summarize(count = sum(count)) 

weekly_catch_summary <- weekly_catch %>% 
  group_by(stream, month) %>% 
  summarize(catch_10 = quantile(count, 0.10),
            catch_90 = quantile(count, 0.90))

kable(weekly_catch_summary)

```


# Peak outmigration window

| stream | start timing	| end timing | 
| ----------- | ----------- | ----------- | 
| Battle Creek | mid december | beginning february |
| Butte Creek | mid december | beginning february |
| Clear Creek | mid november | beginning january |
| Deer Creek | mid march | beginning may |
| Feather River | beginning december | mid february |
| Mill Creek | mid march | beginning may | 
| Sacramento - Knights Landing | mid february | beginning april |
| Sacramento - Tisdale | mid january | beginning march |
| Yuba River | mid november | beginning january |

# Total estimate of fish needed by month


```{r, echo = F}

weeks_month <- tibble(stream = c(rep("battle creek",3),
                                 rep("butte creek",3),
                                 rep("clear creek",3),
                                 rep("deer creek",3),
                                 rep("feather river",3),
                                 rep("mill creek",3),
                                 rep("knights landing",3),
                                 rep("tisdale",3),
                                 rep("yuba river", 3)),
                      month = c("december", "january", "february",
                                "december", "january", "february",
                                "november", "december", "january",
                                "march", "april","may",
                                "december", "january", "february",
                                "march", "april","may",
                                "february","march","april",
                                "january", "february", "march",
                                "november", "december", "january"),
                      weeks = c(2,4,1,
                                2,4,1,
                                2,4,1,
                                2,4,1,
                                3,4,2,
                                2,4,1,
                                2,4,1,
                                2,4,1,
                                2,4,1)
                      )
release_window <- number_released_table %>% 
  add_row(stream = c("butte creek",
                     "deer creek",
                     "mill creek",
                     "yuba river",
                     "knights landing",
                     "tisdale"),
          historic_efficiency = c(rep(0.049,4),
                              rep(0.004, 2)),
          release_per_trial = c(rep(485, 4),
                            rep(6225, 2))) %>% 
  filter(!stream %in% c("global", "sacramento river")) %>% 
  left_join(weeks_month) %>% 
  mutate(fish_needed = weeks * release_per_trial)

release_window_historic <- release_window %>% 
  mutate(month = case_when(month == "november" ~ 11,
                           month == "december" ~ 12,
                           month == "january" ~ 1,
                           month == "february" ~ 2,
                           month == "march" ~ 3,
                           month == "april" ~ 4,
                           month == "may" ~ 5)) %>% 
  left_join(weekly_catch_summary) %>% 
  mutate(total_fish_needed = ifelse(stream %in% c("feather river", "clear creek"), fish_needed*2, fish_needed),
         hatchery_max = total_fish_needed - catch_10,
         hatchery_min = total_fish_needed - catch_90,
         hatchery_max = ifelse(hatchery_max < 0, 0, hatchery_max),
         hatchery_min = ifelse(hatchery_min < 0, 0, hatchery_min))

total_hatchery_counts <- release_window_historic %>% 
  group_by(stream) %>% 
  summarize(total_fish_needed = sum(total_fish_needed),
            hatchery_max = sum(hatchery_max),
            hatchery_min = sum(hatchery_min))
# original table without historic catch
total_counts <- release_window %>% 
  select(stream, month, fish_needed) %>% 
  pivot_wider(id_cols = stream, names_from = "month", values_from = "fish_needed", values_fill = 0) %>% 
  mutate(total = november + december + january + february + march + april + may)

total_counts <- total_counts[,c(1,5,2,3,4,6,7,8,9)]

kable(total_counts)

kable(total_hatchery_counts)
```




```{r}
efficiency_dat <- left_join(number_released_flow, number_recaptured) %>% 
    left_join(standard_flow %>% 
                rename(release_date = date)) %>% 
  mutate(efficiency = number_recaptured/number_released)

effic_release_plot <- function(stream_name) {
  efficiency_dat %>% 
  filter(stream == stream_name, efficiency < 1) %>% 
    ggplot(aes(x = number_released, y = efficiency, color = flow_cfs)) +
    geom_point() +
    labs(x = "number fish released",
         y = "efficiency estimate") +
    theme_minimal()
}

flow_efficiency_plot <- function(stream_name) {
  efficiency_dat %>% 
  filter(stream == stream_name, efficiency < 1, number_released < 5000) %>% 
    ggplot(aes(x = flow_cfs, y = efficiency, color = number_released)) +
    geom_point() +
    labs(x = "flow (cfs)",
         y = "efficiency estimate") +
    theme_minimal()
}

flow_efficiency_plot("sacramento river")
effic_release_plot("sacramento river")
effic_release_plot("feather river")
effic_release_plot("battle creek")
effic_release_plot("clear creek")
```

```{r}
# release_group_effic <- function(stream_name) {
#   left_join(number_released_flow, number_recaptured) %>% 
#   mutate(efficiency = number_recaptured/number_released,
#          release_group = case_when(number_released <= 100 ~ "0-100",
#                                    number_released > 100 & number_released <= 200 ~ "101-200",
#                                    number_released > 200 & number_released <= 300 ~ "201-300",
#                                    number_released > 300 & number_released <= 400 ~ "301-400",
#                                    number_released > 400 & number_released <= 500 ~ "401-500",
#                                    number_released > 500 & number_released <= 600 ~ "501-600",
#                                    number_released > 600 & number_released <= 700 ~ "601-700",
#                                    number_released > 700 & number_released <= 800 ~ "701-800",
#                                    number_released > 800 & number_released <= 900 ~ "801-900",
#                                    number_released > 900 & number_released <= 1000 ~ "901-1000",
#                                    number_released > 1000 ~ ">1000"),
#          release_group = factor(release_group, levels = c("0-100", "101-200", "201-300", "301-400", "401-500", "501-600", "601-700","701-800","801-900","901-1000",">1000"))) %>% 
#   #filter(stream == stream_name) %>% 
#     filter(efficiency < 1) %>% 
#     ggplot(aes(x = release_group, y = efficiency)) +
#     geom_boxplot() +
#     labs(x = "number fish released",
#          y = "efficiency estimate") +
#     theme(axis.text.x = element_text(angle = 0.45)) +
#     theme_minimal() +
#     facet_wrap(~stream, scales = "free")
# }
# under_1000_effic("sacramento river")
# under_1000_effic("feather river")
```

```{r}
n_yrs <- number_released_flow %>% 
  mutate(year = year(release_date)) %>% 
  group_by(stream) %>% 
  distinct(year) %>% 
  tally()
n_obs <- number_recaptured %>% 
  group_by(stream) %>% 
  distinct(release_id) %>% 
  tally()

n_recaptured_sac <- number_recaptured %>% 
  filter(stream == "sacramento river")
n_released_sac <- number_released_flow %>% 
  filter(stream == "sacramento river") %>% 
  summarize(mean = mean(number_released),
            median = median(number_released),
            min = min(number_released),
            max = max(number_released))

sac_flow <- filter(standard_flow, stream == "sacramento river", site == "knights landing") 

ck <- sac_flow %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  filter(month %in% 2:4) %>% 
  group_by(year) %>% 
  summarize(min = min(flow_cfs),
            mean = mean(flow_cfs),
            max = max(flow_cfs))

sac_efficiency <- filter(efficiency_dat, stream == "sacramento river")

filter(sac_efficiency, efficiency == 0) %>% tally()
filter(sac_efficiency, efficiency == 0) %>% summarize(mean_released = mean(number_released),
                                                      median_released = median(number_released))
filter(sac_efficiency, efficiency > 0) %>% summarize(mean_released = mean(number_released),
                                                      median_released = median(number_released))

sac_compare <- sac_efficiency %>% 
  mutate(efficiency_zero = ifelse(efficiency == 0, "0% efficiency", "greater than 0% efficiency"))

sac_compare %>% 
  ggplot(aes(x = efficiency_zero, y = number_released)) +
  geom_boxplot() +
  labs(x ="", 
       y = "number of fish released") +
  theme_minimal()

ggsave("knights_efficiency_boxplot.png", width = 8, height = 4)

t.test(number_released~efficiency_zero, data= sac_compare)
```