---
title: "Yuba River Upstream Passage QC"
author: "Erin Cain"
date: "11/8/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Yuba River Upstream Passage Monitoring Video Data

## Description of Monitoring Data

Yuba river upstream passage data provided by Mike Healey in a xlsx doc. No associated description or metadata. 
					

**Timeframe:** 2004 - 2019

**Video Season:** Appears to be January through October - we see most fish March - June

**Completeness of Record throughout timeframe:** Data from each year, records on hours sampled each date

**Sampling Location:** Yuba River North and Sourth Fish Ladder

**Data Contact:** [Mike Healey](mailto:Mike.Healey@wildlife.ca.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
getwd() #to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "adult-upstream-passage-monitoring/yuba-river/data-raw/MASTER_Vaki_ALL PASSAGE and OPS_BY2004-BY2019.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_yuba_river_passage_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data: 
```{r}
# read in data to clean 
sheets <- readxl::excel_sheets("raw_yuba_river_passage_data.xlsx")
sheets
raw_passage_data <- read_excel("raw_yuba_river_passage_data.xlsx", sheet = "All Passage") %>% glimpse()
raw_passage_operations <- read_excel("raw_yuba_river_passage_data.xlsx", sheet = "All Ops") %>% glimpse()
```

## Data transformations

```{r}
cleaner_passage_data <- raw_passage_data %>% 
  left_join(raw_passage_operations, c("Date", "Ladder")) %>% # Join passage data, no data on hours monitored in first 2 months
  janitor::clean_names() %>% 
  rename(speed_m_per_s = speed_m_s,
         catagory = categorie,
         passage_direction = direction_of_passage) %>%
  mutate(count = 1, 
         date = as.Date(date), 
         time = hms::as_hms(time)) %>%
  select(-percent) %>% # remove because just hours/24
  glimpse()
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_passage_data %>% select_if(is.numeric) %>% colnames()
```
### Variable: `count`

**Plotting Passage Counts Moving Up over Period of Record**

```{r}
cleaner_passage_data %>% filter(year(date) > 2010) %>% # show only last 10 years to make graphs more legible 
  group_by(date) %>%
  mutate(total_count = sum(count)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = total_count, fill = passage_direction)) + 
  geom_col() + 
  facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1901-01-01"), as.Date("1901-10-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 23),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Count of Passage All Runs", 
       x = "Date")  
```

Spring Run Chinook appear to be seen moving upstream April through September.

```{r}
# Boxplots of daily counts by year
cleaner_passage_data %>% group_by(date, passage_direction) %>%
  mutate(daily_count_upstream = sum(count)) %>%
  mutate(year = as.factor(year(date))) %>% 
  ungroup() %>%
  ggplot(aes(x = year, y = daily_count_upstream, color = passage_direction)) + 
  geom_boxplot() + 
  theme_minimal() +
  theme(text = element_text(size = 23),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Count of Upstream Passage Sumarized") 
```

```{r}
cleaner_passage_data  %>%
  mutate(year = as.factor(year(date))) %>%
  group_by(year, passage_direction) %>%
  mutate(total_count = sum(count)) %>%
  ggplot(aes(x = year, y = total_count, fill = passage_direction)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Yearly Upstream Fish Counts",
       y = "Total fish count") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of Passage Counts Moving Up over Period of Record**

```{r}
# daily numeric summary 
cleaner_passage_data %>% group_by(date) %>%
  summarise(count = sum(count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$count))/nrow(cleaner_passage_data), 3) * 100` % of values in the `count` column are NA. Expected given I created the column assuming each row is one fish. 

### Variable: `depth_mm`

**Plotting depth_mm over Period of Record**

```{r}
cleaner_passage_data %>%
  ggplot(aes(x = depth_mm)) + 
  geom_histogram(breaks=seq(25, 275, by=2)) + 
  scale_x_continuous(breaks=seq(25, 275, by=25)) +
  theme_minimal() +
  labs(title = "depth distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of depth_mm over Period of Record**

```{r}
summary(cleaner_passage_data$depth_mm)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$depth_mm))/nrow(cleaner_passage_data), 3) * 100` % of values in the `depth_mm` column are NA. 

### Variable: `length_cm`

Length of fish in cm							

**Plotting length_cm**
  
```{r}
cleaner_passage_data %>%
  ggplot(aes(x = length_cm)) + 
  geom_histogram(breaks=seq(0, 175, by=2)) + 
  scale_x_continuous(breaks=seq(0, 175, by=25)) +
  theme_minimal() +
  labs(title = "length distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of length_cm over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_passage_data$length_cm)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_passage_data$length_cm))/nrow(cleaner_passage_data), 3) * 100` % of values in the `length_cm` column are NA. 

Quite a few 0 values and NA values. Need to figure out if 0 values can also be considered NA. 

### Variable: `speed_m_per_s`

Speed of fish in meter per second 

**Plotting speed_m_per_s over Period of Record**

```{r}
cleaner_passage_data %>%
  ggplot(aes(x = speed_m_per_s)) + 
  geom_histogram(breaks=seq(-3, 4, by=.25)) + 
  scale_x_continuous(breaks=seq(-3, 5, by=1)) +
  theme_minimal() +
  labs(title = "speed distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of speed_m_per_s over Period of Record**

```{r}
summary(cleaner_passage_data$speed_m_per_s)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$speed_m_per_s))/nrow(cleaner_passage_data), 3) * 100` % of values in the `speed_m_per_s` column are NA. 

### Variable: `position_in_frame`

Fish position in video frame. Not sure how to map numbers to positions. 

**Plotting position_in_frame over Period of Record**

```{r}
cleaner_passage_data %>%
  ggplot(aes(x = position_in_frame)) + 
  geom_histogram(breaks=seq(0, 50, by=1)) + 
  scale_x_continuous(breaks=seq(0, 50, by=5)) +
  theme_minimal() +
  labs(title = "Position in frame distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of position_in_frame over Period of Record**

```{r}
summary(cleaner_passage_data$position_in_frame)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$position_in_frame))/nrow(cleaner_passage_data), 3) * 100` % of values in the `position_in_frame` column are NA. 

### Variable: `hours`

Number of hours in day that were monitored  

**Plotting hours over Period of Record**

```{r}
cleaner_passage_data %>%
  ggplot(aes(x = hours)) + 
  geom_histogram(breaks=seq(0, 25, by=1)) + 
  scale_x_continuous(breaks=seq(0, 25, by=1)) +
  theme_minimal() +
  labs(title = "Position in frame distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of hours over Period of Record**

```{r}
summary(cleaner_passage_data$hours)
```

Often monitored all 24 hours. Some days where hours are 0. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$hours))/nrow(cleaner_passage_data), 3) * 100` % of values in the `hours` column are NA. 


## Explore Categorical variables: {.tabset}

```{r}
cleaner_passage_data %>% select_if(is.character) %>% colnames()
```


### Variable: `catagory`
```{r}
table(cleaner_passage_data$catagory)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
catagories_to_keep <- str_detect(cleaner_passage_data$catagory, c("chinook", "chn")) 
cleaner_passage_data <- filter(cleaner_passage_data, str_detect(tolower(cleaner_passage_data$catagory), c("chinook", "chn")))
cleaner_passage_data$catagory <- tolower(cleaner_passage_data$catagory)

# Create a column for adipose 
cleaner_passage_data$adipose <- case_when(cleaner_passage_data$catagory %in% 
                                            c("chinook ad-unidentified", "chinook ad undetermined", "chinook ad-undetermi", 
                                               "chinook ad-undetermined", "chinook ad clip unde", "chinook ad underterm", 
                                               "chinook ad undetermi") ~ "unknown", 
                                          cleaner_passage_data$catagory %in% 
                                            c("chinook ad-clip", "chinook ad clip", "chinook salmon ad cl", 
                                               "chinook salmon ad c", "chinook salomon ad c") ~ "clipped")
table(cleaner_passage_data$catagory)
# TODO figure out what the whole -p/+s and +p/+s means 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$catagory))/nrow(cleaner_passage_data), 3) * 100` % of values in the `category` column are NA. 

### Variable: `passage_direction`

States if the fish is moving upstream (Up) or downstream (Down)

```{r}
table(cleaner_passage_data$passage_direction) 
cleaner_passage_data$passage_direction <- tolower(cleaner_passage_data$passage_direction)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$passage_direction))/nrow(cleaner_passage_data), 3) * 100` % of values in the `passage_direction` column are NA. 

### Variable: `ladder`

Refers to which fish ladder observation occured on. 

```{r}
table(cleaner_passage_data$ladder) 
cleaner_passage_data$ladder <- tolower(cleaner_passage_data$ladder)
```
 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_passage_data$ladder))/nrow(cleaner_passage_data), 3) * 100` % of values in the `category` column are NA. 

## Summary of identified issues

* The categories are very confusing, need to get definitions for the different category types (TODO ask about this)
* not sure if length is fork length or a different measure of length (TODO ask about this)

## Save cleaned data back to google cloud 
```{r}
yuba_upstream_passage <- cleaner_passage_data %>% glimpse()
```


```{r, eval=FALSE}
# Write to google cloud 
# Name file [watershed]_[data type].csv
f <- function(input, output) write_csv(input, file = output)

gcs_upload(yuba_upstream_passage,
           object_function = f,
           type = "csv",
           name = "adult-upstream-passage-monitoring/yuba-river/data/yuba_upstream_passage")
```



