---
title: "feather-river-adult-holding-redd-survey-qc-checklist-2010"
author: "Inigo Peng"
date: "10/6/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(leaflet)
```
# Feather River Redd Survey Data  

## Description of Monitoring Data

**Timeframe:** 2010


**Completeness of Record throughout timeframe: **\

* Longitude and latitude data are not available for 2009, 2010, 2011, 2012, 2019, 2020. NA values will be filled in for these data sets in final cleaned data set.

**Sampling Location:** Feather River


**Data Contact:** [Chris Cook](Chris.Cook@water.ca.gov)


Additional Info:  
1. Latitude and longitude are in NAD 1983 UTM Zone 10N\
2. The substrate is observed visually and an estimate of the percentage of 5 size classes:   
* fines <1cm  
* small 1-5cm  
* medium 6-15cm  
* large 16-30cm  
* boulder >30cm

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/redd_survey/2010_Chinook_Redd_Survey_Data.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "2010_Chinook_Redd_Survey_Data_raw.xlsx",
               Overwrite = TRUE)


```

Read in data from google cloud, glimpse raw data:
```{r}
raw_data_2010 = readxl::read_excel("2010_Chinook_Redd_Survey_Data_raw.xlsx")
glimpse(raw_data_2010)
```

## Data Transformation

```{r}
raw_data_2010$'redd width (ft)' = raw_data_2010$'redd width (ft)'/3.281
raw_data_2010$'redd length (ft)' = raw_data_2010$'redd length (ft)'/3.281
cleaner_data_2010 <- raw_data_2010 %>% 
  select(-c('Survey Wk', 'File #', '# of redds')) %>% 
  rename('salmon_counted'= '# salmon',
         'latitude' = 'Latitude',
         'longitude' = 'Longitude',
         'depth_m' = 'Depth (m)',
         'velocity_m/s' = 'Velocity (m/s)',
         'pot_depth_m' = 'Pot Depth (m)',
         'percent_fine_substrate' = '% fines(<1 cm)',
         'percent_small_substrate' = '% small (1-5 cm)',
         'percent_medium_substrate'= '% med (5-15 cm)',
         'percent_large_substrate' = '% large (15-30 cm)',
         'percent_boulder' = '% boulder (>30 cm)',
         'redd_width_m' = 'redd width (ft)',
         'redd_length_m' = 'redd length (ft)',
         ) %>% 
  mutate(latitude = as.numeric(latitude),
         longitude = as.numeric(longitude)) %>% 
  filter(salmon_counted > 0) %>% 
  glimpse()
```
```{r}
cleaner_data_2010 <- cleaner_data_2010 %>% 
  set_names(tolower(colnames(cleaner_data_2010))) %>% 
  mutate(date = as.Date(date)) %>% 
  glimpse()
```

## Explore Categorical Variables

```{r}
cleaner_data_2010 %>% 
  select_if(is.character) %>% colnames()
```

### Variable:`location`

```{r}
table(cleaner_data_2010$location)
```
Locations names are changed to be consistent with the rest of the Feather River redd survey files:
```{r}
cleaner_data_2010 <- cleaner_data_2010 %>% 
  mutate(location = tolower(location), 
         location = if_else(location == "g95 east top", "g95 east side channel top", location),
         location = if_else(location == "middle auditorium", "mid auditorium", location),
         location = if_else(location == "moes ditch", "moe's ditch", location),        
         location = if_else(location == "top hour", "upper hour", location),
         location = if_else(location == "top of moes ditch", "upper moe's ditch", location),
         location = if_else(location == "top of steep", "upper steep", location),
         location = if_else(location == "top moes side channel", "upper moe's side channel", location),
         location = if_else(location == "upper mcfarlin", "upper mcfarland", location),
         location = if_else(location == "vance east top", "upper vance east", location),
         location = if_else(location == "g95 bottom", "g95 main bottom", location)
         )
table(cleaner_data_2010$location)
```

* `r round(sum(is.na(cleaner_data_2010$location))/nrow(cleaner_data_2010), 3)*100` % of values in the `location` column are NA.

## Variable:`Type`

Description:  
Area - polygon mapped with Trimble GPS unit
Point - points mapped with Trimble GPS unit
Questionable redds - polygon mapped with Trimble GPS unit where the substrate was disturbed but did not have the proper characteristics to be called a redd - it was no longer recorded after 2011

```{r}
table(cleaner_data_2010$type)
```
```{r}
cleaner_data_2010 <- cleaner_data_2010 %>% 
  mutate(type = tolower(type),
         type = if_else(type == 'a', 'Area', type),
         type = if_else(type == 'p', 'Point', type),
         type = if_else(type == 'q', 'Questionable Redds', type))
glimpse(cleaner_data_2010$type)
```



## Expore Numeric Variables

```{r}
cleaner_data_2010 %>% 
  select_if(is.numeric) %>% colnames()
```
### Variable:`salmon_counted`  
#### Plotting salmon counted in 2010
```{r}
cleaner_data_2010 %>% 
  ggplot(aes(x = date, y = salmon_counted)) + 
  geom_col() +
  facet_wrap(~year(date), scales = "free") +
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Daily Count of Salmon Counted in 2010")
```

**Numeric Daily Summary of salmon_counted Over 2010**
```{r}
cleaner_data_2010 %>%
  group_by(date) %>%
  summarise(count = sum(salmon_counted, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

```{r fig.height = 8, fig.width = 8}
cleaner_data_2010  %>%
  ggplot(aes(y = location, x = salmon_counted))+
  geom_boxplot() +
  theme_minimal() +
  theme(text = element_text(size = 10))+
  scale_y_discrete()+
  theme(axis.text.y = element_text(size = 8,vjust = 0.1, hjust=0.2))+
  labs(title = "Salmon Count By Locations")
```

**Numeric summary of salmon_counted by location in 2010**
```{r}
cleaner_data_2010 %>%
  group_by(location) %>% 
  summarise(count = sum(salmon_counted, na.rm = T)) %>% 
  pull(count) %>%
  summary()
```

**NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$salmon_counted))/nrow(cleaner_data_2010), 5)*100` % of values in the `salmon_counted` column are NA. 


### Variable:`redd_width_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>%
  group_by(location) %>%
  summarise(mean_redd_width = mean(redd_width_m, na.rm = TRUE)) %>%
  ggplot(aes(y = location, x = mean_redd_width)) +
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size = 8)) +
  labs(title = "Mean Redd Width By Location")
```
```{r}
cleaner_data_2010 %>%
  ggplot(aes(x = redd_width_m)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "white") +
  scale_x_continuous(breaks = round(seq(min(cleaner_data_2010$redd_width_m, na.rm = TRUE), max(cleaner_data_2010$redd_width_m, na.rm = TRUE), by = 1),0))+
  labs(title = "Count of Redd Width")
```

**Numeric Summary of redd_width_m Over 2010**
```{r}
summary(cleaner_data_2010$redd_width_m)
```
**NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'redd_width_m'))/nrow(cleaner_data_2010), 3)*100` % of values in the `redd_width_m` column are NA.


### Variable: `redd_length_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>%
  group_by(location) %>%
  summarise(mean_redd_length = mean(redd_length_m, na.rm = TRUE)) %>%
  ggplot(aes(y = location, x = mean_redd_length)) +
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size = 8)) +
  labs(title = "Mean Redd Length By Location")
```
```{r}
cleaner_data_2010 %>%
  ggplot(aes(x = redd_length_m)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "white") +
  scale_x_continuous(breaks = round(seq(min(cleaner_data_2010$redd_length_m, na.rm = TRUE), max(cleaner_data_2010$redd_length_m, na.rm = TRUE), by = 1),0))+
  labs(title = "Count of Redd Length")
```

**Numeric Summary of redd_length_m Over 2010**
```{r}
summary(cleaner_data_2010$redd_length_m)
```
**NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'redd_length_m'))/nrow(cleaner_data_2010), 3)*100` % of values in the `redd_length_m` column are NA.

### Location Physical Attributes
### Variable:`percent_fine_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>%
  group_by(location) %>% 
  summarise(mean_fine_substrate = mean(percent_fine_substrate, na.rm = TRUE)) %>%
  ggplot(aes(y = location, x = mean_fine_substrate)) +
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size = 8)) +
  labs(title = "Average Fine Substrate By Location")
```

**Numeric Summary of percent_fine_substrate Over 2010**
```{r}
summary(cleaner_data_2010$percent_fine_substrate)
```
**NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'percent_fine_substrate'))/nrow(cleaner_data_2010), 3)*100` % of values in the `percent_fine_substrate` column are NA.

### Variable:`percent_small_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>%
  group_by(location) %>% 
  summarise(mean_small_substrate = mean(percent_small_substrate, na.rm = TRUE)) %>%
  ggplot(aes(y = location, x = mean_small_substrate)) +
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size = 8)) +
  labs(title = "Average Percentage of Small Substrate By Location")
```

**Numeric Summary of percent_small_substrate Over 2010**
```{r}
summary(cleaner_data_2010$percent_small_substrate)
```

**NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'percent_small_substrate'))/nrow(cleaner_data_2010), 3)*100` % of values in the `percent_small_substrate` column are NA.

### Variable:`percent_medium_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>%
  group_by(location) %>% 
  summarise(mean_medium_substrate = mean(percent_medium_substrate, na.rm = TRUE)) %>%
  ggplot(aes(y = location, x = mean_medium_substrate)) +
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size = 8)) +
  labs(title = "Average Percentage of Medium Substrate By Location")
```

**Numeric Summary of percent_medium_substrate Over 2010**
```{r}
summary(cleaner_data_2010$percent_medium_substrate)
```
**NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'percent_medium_substrate'))/nrow(cleaner_data_2010), 3)*100` % of values in the `percent_medium_substrate` column are NA.

### Variable:`percent_large_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>%
  group_by(location) %>% 
  summarise(mean_large_substrate = mean(percent_large_substrate, na.rm = TRUE)) %>%
  ggplot(aes(y = location, x = mean_large_substrate)) +
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size = 8)) +
  labs(title = "Average Percentage of Large Substrate By Location")
```

**Numeric Summary of percent_large_substrate Over 2010**
```{r}
summary(cleaner_data_2010$percent_large_substrate)
```
**NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'percent_large_substrate'))/nrow(cleaner_data_2010), 3)*100` % of values in the `percent_large_substrate` column are NA.

### Variable:`percent_boulder`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>%
  group_by(location) %>% 
  summarise(mean_boulder = mean(percent_boulder, na.rm = TRUE)) %>%
  ggplot(aes(y = location, x = mean_boulder)) +
  geom_col() +
  theme_minimal() +
  theme(text = element_text(size = 8)) +
  labs(title = "Average Percentage of Boulder By Location")
```

**Numeric Summary of percent_boulder Over 2010**
```{r}
summary(cleaner_data_2010$percent_boulder)
```

**NA and Unknown Values**
NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'percent_boulder'))/nrow(cleaner_data_2010), 3)*100` % of values in the `percent_large_substrate` column are NA.

### Variable: `depth_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>% 
  group_by(location) %>% 
  summarise(mean_depth_m = mean(depth_m, na.rm = TRUE)) %>%
  ggplot(aes(x = mean_depth_m, y = location)) + 
  geom_col() + 
  theme_minimal() + 
  theme(text = element_text(size = 8))+
  labs(title = "Average Depth By Location")
```
**Numeric Summary of depth_m Over 2010**
```{r}
summary(cleaner_data_2010$depth_m)
```

**NA and Unknown Values**
NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'depth_m'))/nrow(cleaner_data_2010), 3)*100` % of values in the `depth_m` column are NA.


### Variable: `pot_depth_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>% 
   group_by(location) %>% 
  summarise(mean_pot_depth_m = mean(pot_depth_m, na.rm = TRUE)) %>%
  ggplot(aes(x = mean_pot_depth_m, y = location)) + 
  geom_col() + 
  theme_minimal() + 
  theme(text = element_text(size = 8))+
  labs(title = "Average Pot Depth By Location")
```
**Numeric Summary of pot_depth_m Over 2010**
```{r}
summary(cleaner_data_2010$pot_depth_m)
```

**NA and Unknown Values**
NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'pot_depth_m'))/nrow(cleaner_data_2010), 3)*100` % of values in the `pot_depth_m` column are NA.

### Variable: `velocity_m/s`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2010 %>% 
  group_by(location) %>% 
  summarise(`mean_velocity_m/s` = mean(`velocity_m/s`, na.rm = TRUE)) %>%
  ggplot(aes(x = `mean_velocity_m/s`, y = location)) + 
  geom_col() + 
  theme_minimal() + 
  theme(text = element_text(size = 8))+
  labs(title = "Average Velocity By Location")
```
**Numeric Summary of velocity_m/s Over 2010**
```{r}
summary(cleaner_data_2010$`velocity_m/s`)
```

**NA and Unknown Values**
NA and Unknown Values**
* `r round(sum(is.na(cleaner_data_2010$'velocity_m/s'))/nrow(cleaner_data_2010), 3)*100` % of values in the `velocity_m/s` column are NA.

```{r}
feather_redd_survey_2010 <- cleaner_data_2010 %>% glimpse()
```

### Add cleaned data back onto google cloud
```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_redd_survey_2010,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/2010_Chinook_Redd_Survey_Data.csv")
```