---
title: "feather-river-redd-survey-qc-checklist-2019"
author: "Inigo Peng"
date: "10/6/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(leaflet)
library (RColorBrewer)
```
# Feather River Redd Survey Data  

## Description of Monitoring Data

**Timeframe:** 2019


**Completeness of Record throughout timeframe:**  

* Longitude and latitude data are not available for 2009, 2010, 2011, 2012, 2019, 2020. NA values will be filled in for these data sets in final cleaned data set.


**Sampling Location:** Various sampling locations on Feather River.


**Data Contact:** [Chris Cook](mailto::Chris.Cook@water.ca.gov)


Additional Info:  
1. Latitude and longitude are in NAD 1983 UTM Zone 10N\
2. The substrate is observed visually and an estimate of the percentage of 5 size classes:   

* fines <1cm  
* small 1-5cm  
* medium 6-15cm  
* large 16-30cm  
* boulder >30cm

## Access Cloud Data
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/redd_survey/2019_Chinook_Redd_Survey_Data_raw.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "2019_Chinook_Redd_Survey_Data_raw.xlsx",
               overwrite = TRUE)


```

Read in data from google cloud, glimpse raw data:
```{r}
raw_data_2019 = readxl::read_excel("2019_Chinook_Redd_Survey_Data_raw.xlsx",
                                   col_types = c("date","text","text","numeric","text","numeric", "numeric", "numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric"
                                                 ,"numeric","numeric","numeric","numeric"))
glimpse(raw_data_2019)
```

## Data Transformations

```{r}
cleaner_data_2019 <- raw_data_2019 %>% 
  select(-c('Survey Wk', 'File #')) %>% 
  rename('redd_count' = '# redds',
         'salmon_count'= '# salmon',
         'latitude' = 'Latitude mN',
         'longitude' = 'Longitude mE',
         'depth_m' = 'Depth (m)',
         'pot_depth_m' = 'Pot Depth (m)',
         'velocity_m_per_s' = 'Velocity (m/s)',
         'percent_fine_substrate' = '% fines',
         'percent_small_substrate' = '% small',
         'percent_medium_substrate'= '% med',
         'percent_large_substrate' = '% large',
         'percent_boulder' = '% boulder',
         'redd_width_m' = 'redd width (m)',
         'redd_length_m' = 'redd length (m)',
         ) %>% 
  mutate(Date = as.Date(Date))
cleaner_data_2019 <- cleaner_data_2019 %>% 
  set_names(tolower(colnames(cleaner_data_2019))) %>% 
  glimpse()
```
## Explore `date`
```{r}
cleaner_data_2019 %>%
  ggplot(aes(x = date)) +
  geom_histogram(binwidth = 7, position = 'stack', color = "black") +
  labs(title = "Value Counts For Survey Season Dates")+
  theme(legend.text = element_text(size = 8))
```

**Numeric summary of date in 2019**
```{r}
summary(cleaner_data_2019$date)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'date'))/nrow(cleaner_data_2019), 3)*100` % of values in the `date` column are NA.


## Explore Categorical Variables

```{r}
cleaner_data_2019 %>% 
  select_if(is.character) %>% colnames()
```

### Variable:`location`

```{r}
table(cleaner_data_2019$location)
```
Locations names are changed to be consistent with the rest of the Feather River redd survey files:
```{r}
cleaner_data_2019 <- cleaner_data_2019 %>% 
  mutate(location = str_to_title(location),
         location = if_else(location == "G95 East Bottom", "G95 East Side Channel Bottom", location),
         location = if_else(location == "G95 East Top", "G95 East Side Channel Top", location),
         location = if_else(location == "G95 West", "G95 West Side Channel", location),
         location = if_else(location == "Matthews", "Mathews", location),
         location = if_else(location == "Moe's Side Channel", "Moes Side Channel", location),
         location = if_else(location == "Middle Auditorium", "Mid Auditorium", location),
         location = if_else(location == "Upper Matthews", "Upper Mathews", location),
         location = if_else(location == "Upper Mcfarland", "Upper McFarland", location)
         )
table(cleaner_data_2019$location)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$location))/nrow(cleaner_data_2019), 3)*100` % of values in the `location` column are NA.

## Variable:`type`

# Description:  

* Area - polygon mapped with Trimble GPS unit  

* Point - points mapped with Trimble GPS unit  

* Questionable redds - polygon mapped with Trimble GPS unit where the substrate was disturbed but did not have the proper characteristics to be called a redd - it was no longer recorded after 2011 

```{r}
table(cleaner_data_2019$type)
```
```{r}
cleaner_data_2019 <- cleaner_data_2019 %>% 
  mutate(type = tolower(type),
         type = if_else(type == 'p', 'Point', type))
table(cleaner_data_2019$type)
```


## Explore Numeric Variables

```{r}
cleaner_data_2019 %>% 
  select_if(is.numeric) %>% colnames()
```
### Variable:`salmon_count`  

```{r}
cleaner_data_2019 %>% 
  ggplot(aes(x = date, y = salmon_count)) + 
  geom_col() +
  facet_wrap(~year(date), scales = "free") +
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Daily Salmon Count in 2019")
```


**Numeric Daily Summary of salmon_count Over 2019**
```{r}
cleaner_data_2019 %>%
  group_by(date) %>%
  summarise(count = sum(salmon_count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

```{r fig.height = 8, fig.width = 8}
#Find the most distinctive colours for visual
colourCount = length(unique(cleaner_data_2019$location))
getPalette = colorRampPalette(brewer.pal(12, "Paired"))

cleaner_data_2019  %>%
  filter(!is.na(location)) %>% 
  ggplot(aes(x = salmon_count, fill = location))+
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram() +
  theme_minimal() +
  theme(text = element_text(size = 7))+
  theme(axis.text.x = element_text(size = 10,vjust = 0.5, hjust=0.1))+
  labs(title = "Daily Salmon Count Distribution",
       x = 'Daily Salmon Count')+
  guides(fill = guide_legend(nrow = 21),
         shape = guide_legend(orride.aes = list(size =0.5)),
         color = guide_legend(orride.aes = list(size = 0.5)))
```
**Numeric summary of salmon_count by location in 2019**
```{r}
cleaner_data_2019 %>%
  group_by(location) %>% 
  summarise(count = sum(salmon_count, na.rm = T)) %>% 
  pull(count) %>%
  summary()
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$salmon_count))/nrow(cleaner_data_2019), 5)*100` % of values in the `salmon_count` column are NA. 

### Variable:`redd_count` 

```{r}
cleaner_data_2019 %>% 
  ggplot(aes(x = date, y = redd_count)) + 
  geom_col() +
  facet_wrap(~year(date), scales = "free") +
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Daily Count of Redds in 2019")
```

```{r fig.height = 8, fig.width = 8}
cleaner_data_2019  %>%
  filter(!is.na(location)) %>% 
  ggplot(aes(x = redd_count, fill = location))+
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram() +
  theme_minimal() +
  theme(text = element_text(size = 7))+
  theme(axis.text.x = element_text(size = 10,vjust = 0.5, hjust=0.1))+
  labs(title = "Daily Redd Count Distribution",
       x = 'Daily Redd Count')+
  guides(fill = guide_legend(nrow = 22),
         shape = guide_legend(orride.aes = list(size =0.5)),
         color = guide_legend(orride.aes = list(size = 0.5)))
```

**Numeric Daily Summary of redd_count Over 2019**
```{r}
cleaner_data_2019 %>%
  group_by(date) %>%
  summarise(count = sum(redd_count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data_2019$redd_count))/nrow(cleaner_data_2019), 5)*100` % of values in the `redd_count` column are NA.

### Variable:`redd_width_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = redd_width_m)) +
  geom_histogram(binwidth = 0.3, color = "black", fill = "white") +
  scale_x_continuous(breaks = round(seq(min(cleaner_data_2019$redd_width_m, na.rm = TRUE), max(cleaner_data_2019$redd_width_m, na.rm = TRUE), by = 0.5),0))+
  labs(title = "Redd Width Distribution")
```

**Numeric Summary of redd_width_m Over 2019**
```{r}
summary(cleaner_data_2019$redd_width_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'redd_width_m'))/nrow(cleaner_data_2019), 3)*100` % of values in the `redd_width_m` column are NA.


### Variable: `redd_length_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = redd_length_m)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "white") +
  scale_x_continuous(breaks = round(seq(min(cleaner_data_2019$redd_length_m, na.rm = TRUE), max(cleaner_data_2019$redd_length_m, na.rm = TRUE), by = 1),0))+
  labs(title = "Redd Length Distribution")
```

**Numeric Summary of redd_length_m Over 2019**
```{r}
summary(cleaner_data_2019$redd_length_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'redd_length_m'))/nrow(cleaner_data_2019), 3)*100` % of values in the `redd_length_m` column are NA.

### Physical Attributes

### Variable:`percent_fine_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = percent_fine_substrate, fill = location)) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 5, position = 'stack', color = "black") +
  labs(title = "Percent Fine Substrate Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```

**Numeric Summary of percent_fine_substrate Over 2019**
```{r}
summary(cleaner_data_2019$percent_fine_substrate)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'percent_fine_substrate'))/nrow(cleaner_data_2019), 3)*100` % of values in the `percent_fine_substrate` column are NA.

### Variable:`percent_small_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = percent_small_substrate, fill = location)) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 10, position = 'stack', color = "black") +
  labs(title = "Percent Small Substrate Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```

**Numeric Summary of percent_small_substrate Over 2019**
```{r}
summary(cleaner_data_2019$percent_small_substrate)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'percent_small_substrate'))/nrow(cleaner_data_2019), 3)*100` % of values in the `percent_small_substrate` column are NA.

### Variable:`percent_medium_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = percent_medium_substrate, fill = location)) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 10, position = 'stack', color = "black") +
  labs(title = "Percent Medium Substrate Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```

**Numeric Summary of percent_medium_substrate Over 2019**
```{r}
summary(cleaner_data_2019$percent_medium_substrate)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'percent_medium_substrate'))/nrow(cleaner_data_2019), 3)*100` % of values in the `percent_medium_substrate` column are NA.

### Variable:`percent_large_substrate`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = percent_large_substrate, fill = location)) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 10, position = 'stack', color = "black") +
  labs(title = "Percent Large Substrate Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```

**Numeric Summary of percent_large_substrate Over 2019**
```{r}
summary(cleaner_data_2019$percent_large_substrate)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'percent_large_substrate'))/nrow(cleaner_data_2019), 3)*100` % of values in the `percent_large_substrate` column are NA.

### Variable:`percent_boulder`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = percent_boulder, fill = location)) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 4, position = 'stack', color = "black") +
  labs(title = "Percent Boulder Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```

**Numeric Summary of percent_boulder Over 2019**
```{r}
summary(cleaner_data_2019$percent_boulder)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'percent_boulder'))/nrow(cleaner_data_2019), 3)*100` % of values in the `percent_large_substrate` column are NA.

### Summary of Mean Percent Substrate In Each Location

```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>% 
  group_by(location) %>% 
  summarise(mean_percent_fine_substrate = mean(percent_fine_substrate, na.rm = TRUE),
            mean_percent_small_substrate = mean(percent_small_substrate, na.rm = TRUE),
            mean_percent_medium_substrate = mean(percent_medium_substrate, na.rm = TRUE),
            mean_percent_large_substrate = mean(percent_large_substrate, na.rm = TRUE),
            mean_percent_boulder = mean(percent_boulder, na.rm = TRUE),
            ) %>% 
  pivot_longer(
    cols = starts_with("mean"),
    names_to = "substrate_type",
    values_to = "percent",
    values_drop_na = TRUE
  ) %>%
  ggplot(aes(fill = substrate_type,
             y = location,
             x = percent))+
  geom_bar(position = 'stack', stat = 'identity', color = 'black')+
  labs(title = "Mean Percent Substrate by Location")
```

### Variable: `depth_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = depth_m, fill = location, )) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 0.1, position = 'stack', color = "black") +
  labs(title = "Depth Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```
**Numeric Summary of depth_m Over 2019**
```{r}
summary(cleaner_data_2019$depth_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'depth_m'))/nrow(cleaner_data_2019), 3)*100` % of values in the `depth_m` column are NA.


### Variable: `pot_depth_m`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = pot_depth_m, fill = location, )) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 0.2, position = 'stack', color = "black") +
  labs(title = "Pot Depth Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```

**Numeric Summary of pot_depth_m Over 2019**
```{r}
summary(cleaner_data_2019$pot_depth_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'pot_depth_m'))/nrow(cleaner_data_2019), 3)*100` % of values in the `pot_depth_m` column are NA.

### Variable: `velocity_m_per_s`
```{r fig.height = 8, fig.width = 8}
cleaner_data_2019 %>%
  ggplot(aes(x = velocity_m_per_s, fill = location, )) +
  scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 0.1, position = 'stack', color = "black") +
  labs(title = "Velocity Distribution")+
  theme(legend.text = element_text(size = 8)) +
  guides(fill = guide_legend(nrow = 20))
```

**Numeric Summary of velocity_m_per_s Over 2019**
```{r}
summary(cleaner_data_2019$`velocity_m_per_s`)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data_2019$'velocity_m_per_s'))/nrow(cleaner_data_2019), 3)*100` % of values in the `velocity_m_per_s` column are NA.


### Add cleaned data back onto google cloud
```{r}
feather_redd_survey_2019 <- cleaner_data_2019 %>% glimpse()
```

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_redd_survey_2019,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_redd_2019.csv")
```
