---
title: "Feather River Seine Data 1997 - 2001 QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River Seine and Snorkel Data 

## Description of Monitoring Data

Feather River Seine data 1997 - 2021. This dataset contains less attributes than the more recent data and sampling occurs relatively infrequently. 

**Timeframe:** 1997 - 2001

**Seine Season:** First survey can be done as early as mid December and then sampling can continue until late August. 

**Completeness of Record throughout timeframe:** Sampled all 5 years in date range. Sampling does not occur at regular time intervals throughout the years sampled. 

**Sampling Location:** Feather River

**Data Contact:** [Casey Campos](mailto:Casey.Campos@water.ca.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/JPE_SR_all_seine_individuals1997-2001.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_seine_1997-2001.xlsx",
               overwrite = TRUE)

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r, warning=FALSE}
# read in data to clean 
raw_seine_1997 <- read_xlsx("raw_seine_1997-2001.xlsx") %>% 
  glimpse()

```

## Data transformations

```{r}
cleaner_seine_data <- raw_seine_1997 %>%
  janitor::clean_names() %>% 
  filter(species_code %in% c("CHN", "CHNF", "CHNFT", "CHNI",
                        "CHNL", "CHNS", "CHNST", "CHNT","CHNW")) %>%
  rename("cover" = hu_ccover, 
         "substrate" = hu_csubstrate, 
         "HUC_stream_feature" = hu_cunit,
         "lifestage" = salmon_life_stage_code,
         "run" = race,
         "depth_1" = depth1,
         "depth_2" = depth2,
         "depth_1_dist" = depth1dist,
         "depth_2_dist" = depth2dist,
         "gear" = gear_size_code,
         "condition" = condition_code, 
         "weather" = weather_code) %>% 
  mutate(time = hms::as_hms(time),
         seine_id = as.character(seine_id), 
         substrate = as.character(substrate),
         count = 1) %>% # each row is a individual fish catch
  select(-species_code, -individ_auto_id, -mark, -recorder, -gear_type, -crew) %>% 
  glimpse()

``` 


## Explore Numeric Variables: {.tabset}

```{r}
cleaner_seine_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `water_temp`

**Plotting water_temp over Period of Record**

Daily average water temperature measures appear to be lower in Dec - March and tehn increase April - September. They appear to typically range from 44 - 72. 
```{r}
cleaner_seine_data %>% 
  group_by(date = as.Date(date)) %>%
  mutate(avg_temp = mean(water_temp)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  geom_point(alpha = .25, size = 2) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Daily Water Temperature (colored by year)",
       y = "Average daily temp", 
       x = "Date")  
```

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = water_temp)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Temperature distribution (farenheit)") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of water_temp over Period of Record**

```{r}
summary(cleaner_seine_data$water_temp)
```

### Variable: `sechi`

**Plotting sechi over Period of Record**

```{r}
cleaner_seine_data %>% 
  group_by(date = as.Date(date)) %>%
  mutate(avg_sechi = mean(sechi)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_sechi, color = year)) + 
  geom_point(alpha = .75, size = 2) + 
  geom_line(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Daily Average Sechi (colored by year)",
       y = "Sechi", 
       x = "Date")  
```

Most secchi measurments are 0. 

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = sechi)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Sechi distribution") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of sechi over Period of Record**

```{r}
summary(cleaner_seine_data$sechi)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$sechi))/nrow(cleaner_seine_data), 3) * 100` % of values in the `sechi` column are NA. 

### Variable: `flow`

**Plotting flow over Period of Record**

```{r}
cleaner_seine_data %>% 
  group_by(date = as.Date(date)) %>%
  mutate(avg_flow = mean(flow)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_flow, color = year)) + 
  geom_point(alpha = .75, size = 2) + 
  geom_line(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Daily Average Flow (colored by year)",
       y = "Flow (CFS)", 
       x = "Date")  
```

Flow measurements are not taken frequently. It looks like about ~ 15 - 30 days out of the year. 

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = flow)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Flow distribution (cfs)") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of flow over Period of Record**

```{r}
summary(cleaner_seine_data$flow)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$flow))/nrow(cleaner_seine_data), 3) * 100` % of values in the `flow` column are NA. 

### Variable: `distance_out`

**Plotting distribution of `distance_out`**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = distance_out)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Distance Seine Out") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of seine distance out over Period of Record**

```{r}
summary(cleaner_seine_data$distance_out)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$distance_out))/nrow(cleaner_seine_data), 3) * 100` % of values in the `distance_out` column are NA. 


### Variable: `count`

**Plotting Count over Period of Record**

```{r}
cleaner_seine_data %>% 
  group_by(date) %>%
  summarize(count = sum(count)) %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = count)) + 
  geom_col() + 
  facet_wrap(~year(date), scales = "free_y") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Seine Count", 
       x = "Date")  
```

```{r}
cleaner_seine_data %>% group_by(date) %>%
  mutate(daily_count_upstream = sum(count)) %>%
  mutate(year = as.factor(year(date))) %>% 
  ungroup() %>%
  ggplot(aes(x = year, y = daily_count_upstream)) + 
  geom_boxplot() + 
  theme_minimal() +
  theme(text = element_text(size = 23)) + 
  labs(title = "Daily Count of Seine Catch Sumarized by Year") 
```

There is no data on runs (or fork length) so we cannot differentiate spring from other runs. 

**Numeric Summary of Count over Period of Record**

```{r}
# daily numeric summary 
cleaner_seine_data %>% group_by(date) %>%
  summarise(daily_count = sum(count, na.rm = T)) %>%
  pull(daily_count) %>%
  summary()
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$count))/nrow(cleaner_seine_data), 3) * 100` % of values in the `count` column are NA. However, there are clearly gaps in data. 


### Variable: `start_width`, `end_width`

**Plotting distribution of start and end width**

```{r}
cleaner_seine_data %>%  
  ggplot() + 
  geom_histogram(aes(x = start_width), alpha = .5) +
  geom_histogram(aes(x = end_width), fill = "blue", alpha = .5) +
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Distance Seine Out",
       x = "Start (gray) and end (blue) width") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
Start width has a much more spread out distribution. End width is mostly clustered between 4 and 10

**Numeric Summary of seine start and end width out over Period of Record**

```{r}
summary(cleaner_seine_data$start_width)
summary(cleaner_seine_data$end_width)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$start_width))/nrow(cleaner_seine_data), 3) * 100` % of values in the `start_width` column are NA.
* `r round(sum(is.na(cleaner_seine_data$end_width))/nrow(cleaner_seine_data), 3) * 100` % of values in the `end_width` column are NA.

### Variable: `length`

**Plotting distribution of `length`**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = length)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Length beach seine") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of length over Period of Record**

```{r}
summary(cleaner_seine_data$length)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$length))/nrow(cleaner_seine_data), 3) * 100` % of values in the `length` column are NA.


### Variable: `width`

**Plotting distribution of `width`**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = width)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of width over Period of Record**

```{r}
summary(cleaner_seine_data$width)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$width))/nrow(cleaner_seine_data), 3) * 100` % of values in the `width` column are NA.


### Variable: `depth_1`, `depth_2`

**Plotting distribution of `depth_1`, `depth_2`**

```{r}
cleaner_seine_data %>%  
  ggplot() + 
  geom_histogram(aes(x = depth_1), alpha = .5) + 
  geom_histogram(aes(x = depth_2), fill = "blue", alpha = .5) +
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Distance Seine Out") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
Depth 1 goes up to about 1.7, depth 2 goes all the way up to 7
Most values still fall before 2.

**Numeric Summary of depth over Period of Record**

```{r}
summary(cleaner_seine_data$depth_1)
summary(cleaner_seine_data$depth_2)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$depth_1))/nrow(cleaner_seine_data), 3) * 100` % of values in the `depth_1` column are NA.

* `r round(sum(is.na(cleaner_seine_data$depth_2))/nrow(cleaner_seine_data), 3) * 100` % of values in the `depth_2` column are NA.


### Variable: `depth_1_dist`, `depth_2_dist`

TODO figure out what these mean

**Plotting distribution of depth 1 and 2 dist**

```{r}
cleaner_seine_data %>%  
  ggplot() + 
  geom_histogram(aes(x = depth_1_dist), alpha = .5) + 
  geom_histogram(aes(x = depth_2_dist), fill = "blue", alpha = .5) +
  scale_x_continuous() +
  theme_minimal() +
  labs(x = "Depth 1 dist (gray) and Depth 2 dist (blue)") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of depth dist over Period of Record**

```{r}
summary(cleaner_seine_data$depth_1_dist)
summary(cleaner_seine_data$depth_2_dist)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$depth_1_dist))/nrow(cleaner_seine_data), 3) * 100` % of values in the `depth_1_dist` column are NA.
* `r round(sum(is.na(cleaner_seine_data$depth_2_dist))/nrow(cleaner_seine_data), 3) * 100` % of values in the `depth_2_dist` column are NA.


### Variable: `sample_area`

**Plotting distribution of `sample_area`**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = sample_area)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of sample area over Period of Record**

```{r}
summary(cleaner_seine_data$sample_area)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$sample_area))/nrow(cleaner_seine_data), 3) * 100` % of values in the `sample_area` column are NA.



### Variable: `sample_shape`

**Plotting distribution of `sample_shape`**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = sample_shape)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

Sample shape values appear to be either 0, 1, or 2. No info in access database describing what these mean. TODO ask Casey. 

**Numeric Summary of seine distance out over Period of Record**

```{r}
summary(cleaner_seine_data$sample_shape)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$sample_shape))/nrow(cleaner_seine_data), 3) * 100` % of values in the `sample_shape` column are NA.


## Explore Categorical variables: {.tabset}

```{r}
cleaner_seine_data %>% select_if(is.character) %>% colnames()
```


### Variable: `seine_id`

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$seine_id))/nrow(cleaner_seine_data), 3) * 100` % of values in the `seine_id` column are NA.

### Variable: `lifestage`

```{r}
table(cleaner_seine_data$lifestage)
```
```{r}
cleaner_seine_data$lifestage <- case_when(
  cleaner_seine_data$lifestage == 1 ~ "yolk-sac fry",
  cleaner_seine_data$lifestage == 2 ~ "fry",
  cleaner_seine_data$lifestage == 4 ~ "silvery parr",
  cleaner_seine_data$lifestage == "P" | 
    cleaner_seine_data$lifestage == 3 ~ "parr",
  cleaner_seine_data$lifestage == "S" | 
    cleaner_seine_data$lifestage == 5 ~ "smolt",
  cleaner_seine_data$lifestage == "X" ~ "between parr and smolt",)

table(cleaner_seine_data$lifestage)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$lifestage))/nrow(cleaner_seine_data), 3) * 100` % of values in the `lifestage` column are NA.

### Variable: `run`


```{r}
table(cleaner_seine_data$run)
```
```{r}
cleaner_seine_data$run <- case_when(cleaner_seine_data$run == "F" ~ "fall",
                                    cleaner_seine_data$run == "LF" ~ "late fall",
                                    cleaner_seine_data$run == "S" ~ "spring", 
                                    cleaner_seine_data$run == "W" ~ "winter")
table(cleaner_seine_data$run)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$seine_id))/nrow(cleaner_seine_data), 3) * 100` % of values in the `run` column are NA.


### Variable: `location`

```{r}
table(cleaner_seine_data$location)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
format_site_name <- function(string) {
  clean <- str_replace_all(string, "1/2", "half") %>%
    str_replace_all("1/4", "quarter") %>%
    str_replace_all("d/s", "downstream") %>%
    str_replace_all("u/s", "upstream") %>%
    str_replace_all("S.C.", "SC") %>%
    str_replace_all("'", "") %>%
    str_replace_all("G-95", "G95") %>% 
    str_replace_all("Isl", "Island") %>%
    str_replace_all("[^[:alnum:]]", " ") %>% 
    trimws() %>% 
    stringr::str_squish() %>%
    stringr::str_to_title()
}

cleaner_seine_data$location <- format_site_name(cleaner_seine_data$location)
table(cleaner_seine_data$location)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$location))/nrow(cleaner_seine_data), 3) * 100` % of values in the `location` column are NA.

### Variable: `station_code`


```{r}
table(cleaner_seine_data$station_code)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$seine_id))/nrow(cleaner_seine_data), 3) * 100` % of values in the `station_code` column are NA.

### Variable: `gear`

| GearNum |	GearCode |	Gear description |
| -------| ------- | ------------------- | 
| 1 |	SEIN25 |	25 foot beach seine with bag |
| 2 |	SEIN39 |	39 foot beach seine with bag |
| 3 |	SEIN50 |	50 foot beach seine with bag |
| 4 |	SEIN100 |	100 foot beach seine with bag |
| 5 |	EFSEIN |	backpack shock into any seine |

```{r}
table(cleaner_seine_data$gear)
```
```{r}
cleaner_seine_data$gear <- case_when(
  cleaner_seine_data$gear == "1" ~ "SEIN25", 
  cleaner_seine_data$gear == "2" ~ "SEIN39"
)

table(cleaner_seine_data$gear)
```



**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$gear))/nrow(cleaner_seine_data), 3) * 100` % of values in the `gear` column are NA.

### Variable: `condition`

```{r}
table(cleaner_seine_data$condition)
```

* Good - The haul went smoothly without getting hung up on anything or net rolling.
* Fair - The haul went relatively smoothly, but you had to stop once or twice to get it unsnagged or take big rocks out of the net.
* Poor - The haul didn’t go well; you had to stop numerous times while the poles were still a good distance from the shore to free the seine from snags or big rocks.

All other values (0, 4) corresond to NA

```{r}
cleaner_seine_data$condition <- case_when(cleaner_seine_data$condition == "1" ~ "good",
                                               cleaner_seine_data$condition == "2" ~ "fair",
                                               cleaner_seine_data$condition == "3" ~ "poor")


table(cleaner_seine_data$condition)
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$condition))/nrow(cleaner_seine_data), 3) * 100` % of values in the `condition` column are NA.


### Variable: `weather`

```{r}
table(cleaner_seine_data$weather)
```

```{r}
cleaner_seine_data$weather <- case_when(
  cleaner_seine_data$weather == "CLD" ~ "overcast",
  cleaner_seine_data$weather == "CLR" ~ "sunny",
  cleaner_seine_data$weather == "RAN" ~ "precipitation")


table(cleaner_seine_data$weather)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$weather))/nrow(cleaner_seine_data), 3) * 100` % of values in the `weather` column are NA.

### Variable: `substrate`

| SubstrateCode |	Substrate |
| -------------- | ---------------------- |
| 1 |	Fine - small gravel (0-50mm) (0-2in.) |
| 2 |	Small - medium gravel (50-150mm) (2-6in.) |
| 3 |	Medium - large cobble (150-300mm) (6-12in.) |
| 4 |	Boulder (>300mm) (>12in.) |
| 5 |	Pavement (Boat Ramp) |

```{r}
table(cleaner_seine_data$substrate)
```

```{r}
cleaner_seine_data$substrate <- case_when(
  cleaner_seine_data$substrate == "1" ~ "fine",
  cleaner_seine_data$substrate == "2" ~ "small",
  cleaner_seine_data$substrate == "3" ~ "medium",
  cleaner_seine_data$substrate == "4" ~ "boulder",
  cleaner_seine_data$substrate == "5" ~ "pavement")

table(cleaner_seine_data$substrate)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$substrate))/nrow(cleaner_seine_data), 3) * 100` % of values in the `substrate` column are NA.

### Variable: `cover`

```{r}
table(cleaner_seine_data$cover)
```

**Create lookup rda for cover encoding:** 
```{r}
feather_seine_cover <- c("A", "B", "C", "D", "E", "BD", "DE" )
names(feather_seine_cover) <- c("no apparent cover",
"small instream objects",
"large instream objects",
"overhead objects",
"submerged aquatic vegetation",
"small instream objects and overhead objects",
"overhead objects and submerged aquatic vegetation")

tibble(code = feather_seine_cover, 
       definitions = names(feather_seine_cover))

# write_rds(feather_seine_cover, "../../../data/feather_seine_cover.rds")
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$cover))/nrow(cleaner_seine_data), 3) * 100` % of values in the `cover` column are NA.

### Variable: `HUC_strean_feature`

| UnitCode |	Unit |
| --------- | ------ |
| G |	Glide |
| GM |	Glide Edgewater |
| M |	Riffle Margin Eddy |
| P |	Pool |
| R |	Riffle |
| RM |	Riffle Edgewater |
| W |	Backwater |

```{r}
table(cleaner_seine_data$HUC_stream_feature)
```

```{r}
cleaner_seine_data$HUC_stream_feature <- case_when(
  cleaner_seine_data$HUC_stream_feature == "G" ~ "glide",
  cleaner_seine_data$HUC_stream_feature == "GM" ~ "glide edgewater",
  cleaner_seine_data$HUC_stream_feature == "M" ~ "riffle margin eddy",
  cleaner_seine_data$HUC_stream_feature == "P" ~ "pool",
  cleaner_seine_data$HUC_stream_feature == "R" ~ "riffle",
  cleaner_seine_data$HUC_stream_feature == "RM" ~ "riffle edgewater",
  cleaner_seine_data$HUC_stream_feature == "W" ~ "backwater")

table(cleaner_seine_data$HUC_stream_feature)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$HUC_stream_feature))/nrow(cleaner_seine_data), 3) * 100` % of values in the `HUC_stream_feature` column are NA.

### Variable: `comments`


```{r}
unique(cleaner_seine_data$comments)[1:5]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$comments))/nrow(cleaner_seine_data), 3) * 100` % of values in the `comments` column are NA.

## Summary of identified issues

* Sampling does not occur that frequently each year
* Figure out depth measures and sample shape values (TODO contact Casey)

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```

