---
title: "Feather River Snorkel QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River Snorkel Data 

## Description of Monitoring Data

Feather River Snorkel Data from 2004 - 2020. This data was provided to flowwest in an access database  "Snorkel_Revided.mdb. We queried it to have all year and important variables. 

**Timeframe:** 2004 - 2020 

**Seine Season:** 

**Completeness of Record throughout timeframe:** 

No environmental data collected pre 2010.


No data for: 

* 2006
* 2008
* 2009
* 2014


**Sampling Location:** Feather River

**Data Contact:** [Casey Campos](mailto:Casey.Campos@water.ca.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/feather_snorkel_and_catch.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_snorkel.xlsx",
               overwrite = TRUE)

```

Read in data from google cloud, glimpse raw data: 
```{r, warning=FALSE}
# read in data to clean 
raw_snorkel <- read_xlsx("raw_snorkel.xlsx") %>% glimpse
```

## Data transformations



```{r}
cleaner_snorkel_data <- raw_snorkel %>% 
  janitor::clean_names() %>%
  rename(start_time = snorkel_start_ttime, 
         end_time = snorkel_end_time, 
         lwd_number = lwd
        ) %>%
  mutate(date = as.Date(date), 
         start_time = hms::as_hms(start_time), 
         end_time = hms::as_hms(end_time),
         time_of_temperature = hms::as_hms(time_of_temperature),
         survey_id = as.character(survey_id), 
         obs_id = as.character(obs_id),
         substrate = as.character(substrate), 
         overhead_cover = as.character(overhead_cover),
         lwd_number = as.character(lwd_number)) %>% # I think this is some sort of tag number so I am changing to character
  select(-observer, -sid) %>% # remove sid because it is the same as survey_id 
  filter(!is.na(date), # survey ID 1 and 106 do not specify date 
         species %in% c("CHN", "NONE", "CHNU", "CHNT", "CHNC", "CHNF", "CHNS")) %>% # filter species to relevant values (none is helpful to show they snorkeled and did not see anything)
  glimpse()

```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_snorkel_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `flow`

**Plotting flow over Period of Record**

```{r}
cleaner_snorkel_data %>% 
  ggplot(aes(x = date, y = flow)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  facet_wrap(~year(date), scales = "free_x") + 
  theme_minimal()
```

Very inconsistent and sparse flow measures. 

```{r}
cleaner_snorkel_data %>% 
  ggplot(aes(x = flow)) +
  geom_histogram() + 
  theme_minimal()
```

Flow is between 0 - 8000. 0 values must be errors or NA.  

**Numeric Summary of flow over Period of Record**

```{r}
summary(cleaner_snorkel_data$flow)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$flow))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `flow` column are NA. 


### Variable: `turbidity`

**Plotting turbidity over Period of Record**

```{r}
cleaner_snorkel_data %>% 
  ggplot(aes(x = date, y = turbidity)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  facet_wrap(~year(date), scales = "free") + 
  theme_minimal()
```

We only have non zero turbidity values for 2011, 2012, 2017, 2018, and 2019

```{r}
cleaner_snorkel_data %>%
  ggplot(aes(x = turbidity)) + 
  geom_histogram()
```

Most turbidity measures are 0. All fall between 0 and 4.5. 

**Numeric Summary of turbidity over Period of Record**

```{r}
summary(cleaner_snorkel_data$turbidity)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$turbidity))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `turbidity` column are NA.

### Variable: `temperature`

**Plotting temperature over Period of Record**

```{r}
cleaner_snorkel_data %>% 
  ggplot(aes(x = date, y = temperature)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  facet_wrap(~year(date), scales = "free") + 
  theme_minimal()
```

No temp measures before 2010, consistency of temperature measures varies throughout the years. 

```{r}
cleaner_snorkel_data %>% 
  ggplot(aes(x = temperature)) +
  geom_histogram() +
  theme_minimal()
```

Interesting temperature distribution. I wonder if some measures are in C and some in F? 0 values are likely also a mistake or NA.

**Numeric Summary of temperature over Period of Record**

```{r}
summary(cleaner_snorkel_data$temperature)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$temperature))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `temperature` column are NA.



### Variable: `count`

**Plotting count over Period of Record**

```{r, include=FALSE}
sac_indices <- waterYearType::water_year_indices %>% 
    filter(location == "Sacramento Valley") %>% 
    transmute(water_year = WY, year_type = as.character(Yr_type))
```

```{r}
cleaner_snorkel_data %>% 
  group_by(date) %>%
  summarise(total_daily_catch = sum(count, na.rm = T)) %>%
  # filter(year(date) > 2014, year(date) < 2021) %>%
  mutate(water_year = ifelse(month(date) %in% 10:12, year(date) + 1, year(date))) %>% 
  left_join(sac_indices) %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = total_daily_catch, fill = year_type)) + 
  geom_col() + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1901-03-01"), as.Date("1901-10-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Total Daily Raw Fish Count",
       y = "Total daily catch",
       x = "Date")+ 
  facet_wrap(~water_year, scales = "free") + 
  scale_fill_manual(values = wesanderson::wes_palette("Moonrise2", 5, type = "continuous"))
```


It looks like snorkel surveys do not happen very often but when they do happen a lot of fish can be caught in a day. 
```{r}
cleaner_snorkel_data  %>%
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = year, y = count)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Fish Counted each Year",
       y = "Total fish count") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of count over Period of Record**

```{r}
summary(cleaner_snorkel_data$count)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$count))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `count` column are NA.



### Variable: `est_size`

TODO figure out what size is describing. Is it length? or weight? 

**Plotting est_size over Period of Record**

```{r}
cleaner_snorkel_data %>% 
  ggplot(aes(x = est_size)) + 
  geom_histogram() + 
  theme_minimal()
  
```

Estimated sizes fall between 0 and 1000. Again we need to figure out what type of size variable they are estimating 

**Numeric Summary of est_size over Period of Record**

```{r}
summary(cleaner_snorkel_data$est_size)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$est_size))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `est_size` column are NA.



### Variable: `water_depth_m`

Depth of water, units: m 

**Plotting water depth over Period of Record**

```{r}
cleaner_snorkel_data %>% 
  ggplot(aes(x = water_depth_m)) + 
  geom_histogram() + 
  theme_minimal()
```
Depth of water is between 0 and 3 meters. 

**Numeric Summary of depth of water over Period of Record**

```{r}
summary(cleaner_snorkel_data$water_depth_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$water_depth_m))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `water_depth_m` column are NA.




## Explore Categorical variables: {.tabset}


```{r}
cleaner_snorkel_data %>% select_if(is.character) %>% colnames()
```


### Variable: `survey_id`

Each row does not correspond to a unique survey. 

```{r}
length(unique(cleaner_snorkel_data$survey_id)) == length(cleaner_snorkel_data)
```

There are `r length(unique(cleaner_snorkel_data$survey_id))` unique survey ids

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$survey_id))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `survey_id` column are NA.

### Variable: `weather_code`

No lookup table for weather codes provided. However, they are fairly easy to interperate. Would be best if we could get weather codes standardized throughout and condense these down. 

```{r}
table(cleaner_snorkel_data$weather_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_snorkel_data$weather_code <- case_when(cleaner_snorkel_data$weather_code %in% c("CLD", "CLDY") ~ "cloudy", 
                                               cleaner_snorkel_data$weather_code %in% c("CLR (Hot)", "CLR/Hot", "Hot and CLR", "CLR Hot") ~ "clear and hot",
                                               cleaner_snorkel_data$weather_code %in% c("RAIN", "RAN", "CLD/RAIN", "LT RAIN", "CLD, Wind, Light Sprinkles") ~
                                                 "precipitation",
                                               cleaner_snorkel_data$weather_code %in% c("CLR 95", "CLR") ~ "clear",
                                               cleaner_snorkel_data$weather_code %in% c("PT. CLDY", "CLR/CLD") ~ "partly cloudy",
                                               cleaner_snorkel_data$weather_code %in% c("sun", "SUN") ~ "sunny",
                                               cleaner_snorkel_data$weather_code == c("CLR WINDY") ~ "clear and windy",
                                               cleaner_snorkel_data$weather_code == c("WND") ~ "windy",
                                               cleaner_snorkel_data$weather_code == c("LT CLD/HAZE") ~ "hazy")

table(cleaner_snorkel_data$weather_code) 
```

I would still like to simplify these down more but this is better for now. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$weather_code))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `weather_code` column are NA.

### Variable: `section_name`

A lot of differnt section names. TODO clean these up 

```{r}
table(cleaner_snorkel_data$section_name) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$section_name))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `section_name` column are NA.

### Variable: `units_covered`

A lot of units described here. TODO clean up

```{r}
table(cleaner_snorkel_data$units_covered) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$units_covered))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `units_covered` column are NA.

### Variable: `survey_comments` `observation_comments`
```{r}
unique(cleaner_snorkel_data$survey_comments)[1:5]
```

```{r}
unique(cleaner_snorkel_data$observation_comments)[1:5]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$survey_comments))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `survey_comments` column are NA.

* `r round(sum(is.na(cleaner_snorkel_data$observation_comments))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `observation_comments` column are NA.

### Variable: `obs_id`

Each row does not correspond to a unique observation.  
```{r}
length(unique(cleaner_snorkel_data$obs_id)) == length(cleaner_snorkel_data)
```

There are `r length(unique(cleaner_snorkel_data$obs_id))` unique observation ids

* `r round(sum(is.na(cleaner_snorkel_data$obs_id))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `obs_id` column are NA.

### Variable: `unit`

TODO figure out what unit is refering to
```{r}
table(cleaner_snorkel_data$unit) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$unit))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `unit` column are NA.

### Variable: `species`

Species lookup table:

| SpeciesCode	| Species | 
| -------- | ------------- |
| CHN |	Chinook salmon |
| CHNC |	Chinook Salmon - Clipped |
| CHNT |	Chinook salmon - Tagged |
| CHNU |	Chinook salmon - Unknown |
| NONE | Z. nada | 

Species lookup table does not give a definition for CHNF, CHNS. My guess is that these refer to Chinook salmon Spring and Chinook salmon fall but I would want to confirm this. 

```{r}
table(cleaner_snorkel_data$species) 
```

TODO create clipped/adipose column from this column

TODO create a run column from this column 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$species))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `species` column are NA.

### Variable: `size_class`

| Size Class | Size Range |
| ---------- | ---------- |
| I |	0-50 |
| II |	51-75 |
| III |	76-100 |
| IV |	101-150 |
| V |	151-300 |
| VI |	301-499 |
| VII |	500+ |

```{r}
table(cleaner_snorkel_data$size_class) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$size_class))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `size_class` column are NA.

### Variable: `substrate`

| SubstrateCode |	Substrate |
| ------------- | --------- |
| 1 |	Organic Fines, Mud (0.05 mm) |
| 2 |	Sand (0.05 to 2 mm) |
| 3 |	Small Gravel (2 to 50 mm) |
| 4 |	Large Gravel (50 to 150 mm) |
| 5 |	Cobble (150 to 300 mm) |
| 6 |	Boulder (> 300 mm) |
| 0 | ? |

It appears that if more than one substrate is present they list all by creating multi-didget numbers. Not sure what 0 stands for? TODO 

```{r}
table(cleaner_snorkel_data$substrate) 
```

What is the best way to show multiple substrates at different sites? In the seine data they have many substrate columns but I do not think I like that better. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$substrate))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `substrate` column are NA.

### Variable: `instream_cover`

| ICoverCode |	ICover |
| ---------- | ------- |
| A |	No apparent cover |
| B |	Small instream objects/small-medium woody debris |
| C |	Large instream objects/large woody debris |
| D |	Overhead objects |
| E |	Submerged aquatic veg/filamentous algae |
| F |	Undercut bank |

Instream cover appears to use a similar method as substrate where they add letters on if multiple covers are present. 

```{r}
table(cleaner_snorkel_data$instream_cover) 
```

What is the best way to display/simplify these?

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$instream_cover))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `instream_cover` column are NA.


### Variable: `overhead_cover`

| OCoverCode |	OCover |
| --------- | -------- |
| 0 |	No Apparent Cover |
| 1 |	Overhanging veg/obj (< 0.5 m above surface) |
| 2 |	Overhanging veg/obj (0.5 to 2 m above surface |
| 3 |	Surface turbulence, bubble curtain |

Overhead cover appears to use a similar method as substrate and instream cover where they add letters on if multiple covers are present. 

```{r}
table(cleaner_snorkel_data$overhead_cover) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$overhead_cover))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `overhead_cover` column are NA.



### Variable: `hydrology_code`

Looks like these are already defined for the most part, GM = Glide Edgewater so we can change that. 

```{r}
table(cleaner_snorkel_data$hydrology_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_snorkel_data$hydrology_code <- ifelse(cleaner_snorkel_data$hydrology_code == "GM", "Glide Edgewater", cleaner_snorkel_data$hydrology_code)

table(cleaner_snorkel_data$hydrology_code) 
```

* `r round(sum(is.na(cleaner_snorkel_data$hydrology_code))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `hydrology_code` column are NA.



### Variable: `lwd_number`

I am not sure what this number is refering to, TODO

```{r}
table(cleaner_snorkel_data$lwd_number) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$lwd_number))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `lwd_number` column are NA.


## Summary of identified issues

* Snorkel surveys are not done every year in the timeframe 
* Environmental data very sparse, especially in early years
* Still a few unknown columns that we need to define and understand 
* Figure out the best way to display and standardize substrate and cover information

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```

