---
title: "yuba-river-rst-qc-checklist"
author: "Inigo Peng"
date: "10/11/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(leaflet)
library(gridExtra)
```
# Yuba River RST Data  

## Description of Monitoring Data:

Robyn Bilski sent the Yuba RST data in an Access Database via email. The data is limited to 2000 through 2008. Initial query in the Access Database was required before uploading it to Googlecloud. Further information regarding the data can be found in the Draft CAMP screw trap database dictionary.doc and Rotary Screws Traps Report 2007-2008. 

**Timeframe:** 2000-2009


**Completeness of Record throughout timeframe:**

* There is no longitude or latitude data  for RST locations
* Fish life stage data was only tracked from 2007-2009
* Most of the salmon counted were fall run salmon

**Sampling Location:** Yuba River


**Data Contact:** [Robyn Bilski](Robyn.Bilski@Wildlife.ca.gov)


## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "rst/yuba-river/data-raw/yuba-river-rst-data.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "yuba_rst_raw.xlsx",
               overwrite = TRUE)


```

Read in data from google cloud, glimpse raw data:
```{r}
raw_data = readxl::read_excel('yuba_rst_raw.xlsx', col_types = c("text","date","date","text","numeric","numeric","numeric",
                                                             "text","numeric","numeric","text","numeric","numeric","text",
                                                             "text","text","text","numeric","numeric","text","numeric",'text','numeric','numeric'))
glimpse(raw_data)
```

## Data Transformation

```{r}
cleaner_data <- raw_data %>% 
  select(-c('Sample_Entry_SampleRowID','Catch_Entry_SampleRowID', 'CatchRowID', 'Northing', 'Easting')) %>% #Northing, Easting are NA, 
  rename('date'= SampleDate,
         'time' = SampleTime,
         'method' = MethodCode,
         'temperature' = WaterTemperature,
         'turbidity' = Turbidity,
         'velocity' = WaterVelocity,
         'trap_status' = TrapStatus,
         'trap_revolutions' = TrapRevolutions,
         'trap_revolutions2' = TrapRevolutions2,
         'rpms_before' = RPMsBefore,
         'rpms_after' = RPMsAfter,
         'organism_code' = OrganismCode,
         'fork_length' = ForkLength,
         'lifestage' = StageCode,
         'comments' = Comments
         ) %>%
  mutate(time = hms::as_hms(time))
  # filter(organism_code == 'CHN', rm.na = TRUE)
cleaner_data <- cleaner_data %>% 
  set_names(tolower(colnames(cleaner_data))) %>% 
  glimpse()
```

## Explore Categorical Variables
```{r}
cleaner_data %>% select_if(is.character) %>% colnames()
```
### Variable: `organism_code`

**Description: organism_code indicate the alphanumeric code for a species of fish or other organism present in the catch.**

```{r}
table(cleaner_data$organism_code)
```

Filtering data by Chinooks only.
```{r}
cleaner_data <- cleaner_data %>% 
  filter(organism_code == "CHN")

table(cleaner_data$organism_code)
```


**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'organism_code'))/nrow(cleaner_data), 3)*100` % of values in the `organism_code` column are NA.

### Variable: `method`

**Description: method is an alphanumeric code defining the type of sampling action taken.**

* FSR - Fish Screen Diversion Trap

* RSTR - Rotary Screen Trap

```{r}
table(cleaner_data$method) 
```
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(method = case_when(
    method == "FSR" ~ "fish screen diversion trap",
    method == "RSTR" ~ "rotary screw trap"
  ))
table(cleaner_data$method)
```
**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'method'))/nrow(cleaner_data), 3)*100` % of values in the `method` column are NA.


### Variable: `trap_status`

**Description: trap_status indicate if trap was checked for fish or just set or pulled.**

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(trap_status = set_names(tolower(trap_status)))
table(cleaner_data$trap_status) 
```
**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'trap_status'))/nrow(cleaner_data), 3)*100` % of values in the `trap_status` column are NA.


### Variable: `lifestage`

**Description: lifestage describes the life stage of an organism. It is a combination of 'run' and 'lifestage' data. The following operations creates a new column based on run.** 

* FR - Fall Run salmon
* FRY - Fry
* lfr - Late Fall Run salmon
* n/p - Not Provided
* out - Outlier Salmon
* PAR - Parr 
* SMT - Smolt
* SPR - Silvery Parr
* SR - Spring Run salmon
* UNK - Unknown
* YSF - Yolk-sac Fry

```{r}
table(cleaner_data$lifestage) 
```

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(lifestage = set_names(tolower(lifestage))) %>% 
  mutate(run = case_when(
           lifestage == "fr" ~ "fall",
           lifestage == "lfr" ~ "late fall",
           lifestage == "unk" ~ "unknown",
           lifestage == "sr" ~ "spring",
           lifestage == "out" ~ "outlier salmon"),
         lifestage = case_when(
           lifestage == 'n/p'~ 'not provided',
           lifestage == 'par'~ 'parr',
           lifestage == 'smt'~ 'smolt',
           lifestage == 'spr'~ 'silvery parr',
           lifestage == 'ysf'~ 'yolk-sac fry'
         )
    ) %>% 
  glimpse()

```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'lifestage'))/nrow(cleaner_data), 3)*100` % of values in the `lifestage` column are NA.

* `r round(sum(cleaner_data$'lifestage' == "not provided", na.rm = T)/nrow(cleaner_data), 3)*100` % of value in the `lifestage` column are `not provided`.

* `r round(sum(is.na(cleaner_data$'run'))/nrow(cleaner_data), 3)*100` % of values in the `run` column are NA.

* `r round(sum(cleaner_data$'run' == "unknown", na.rm = T)/nrow(cleaner_data), 3)*100` % of value in the `run` column are `unknown`.

### Variable: `location`
**Description: location of Rotary Screw Traps**

```{r}
table(cleaner_data$location)
```
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(location = case_when(
    location == '5 foot RST at Halwood' ~ '5 foot RST at Hallwood',
    location == 'RST # 2 at Hallwood' ~  'RST 2 at Hallwood',
    location == 'RST 1 at Hallwood on Yuba River' ~ 'RST 1 at Hallwood on Yuba River',
    TRUE ~ as.character(location))
  )
table(cleaner_data$location)
```

## Explore Numeric Variables

```{r}
cleaner_data %>% 
  select_if(is.numeric) %>% colnames()
```

### Variable:`temperature`

**Description: water temperature measured with an ACME Industries infrared Cramomatic 444 automatic sensor. Units: degrees Celsius.** 

```{r}
cleaner_data %>% 
  group_by(date) %>%
  mutate(avg_temp = mean(temperature, na.rm = T)) %>%
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Water Temperature (colored by year)",
       y = "Average daily temp", 
       x = "Date")  
```

There are a series of temperature points that is above 50 degrees celsius. It is unlikely that the water reached such high temperature.

```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = temperature, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Water Temperature summarized by year",
       x = "Water Temperature (C)") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of temperature From 1999 to 2009**

```{r}
summary(cleaner_data$temperature)
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'temperature'))/nrow(cleaner_data), 3)*100` % of values in the `temperature` column are NA.


```{r}
summary(cleaner_data$temperature)
```


### Variable:`turbidity`

**Description: water turbidity measured with an ACME Industries Turbido2000 meter. Units: NTU.**

```{r}
cleaner_data %>% 
  group_by(location) %>% 
  summarise(mean_turbidity = mean(turbidity, na.rm = TRUE)) %>% 
  ggplot(aes(y = location,
             x = mean_turbidity))+
  theme_minimal()+
  geom_bar(position = 'stack', stat = 'identity') +
  labs(title = "Mean Turbidity by Location",
       x = "Turbidity (NTU)")
```

```{r}
cleaner_data %>% 
  group_by(date) %>%
  mutate(avg_turb = mean(turbidity, na.rm = T)) %>%
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = avg_turb, color = year)) + 
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Water Turbidity (colored by year)",
       y = "Average daily turbdity", 
       x = "Date")  
```
```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = turbidity, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Turbdity summarized by year",
       x = "Turbidity (NTU)") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of turbidity 1999 to 2009**

```{r}
summary(cleaner_data$turbidity)
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'turbidity'))/nrow(cleaner_data), 3)*100` % of values in the `turbidity` column are NA.

### Variable:`velocity`

**Description: Water velocity measured with ACME Industries' heavy duty Torpedo 3000 velorecroder. Unit: feet per second. The following operations convert the data to m per second for standardization with the rest of the JPE data.** 

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(velocity = velocity/3.281)

cleaner_data %>% 
  group_by(location) %>% 
  summarise(mean_velocity = mean(velocity, na.rm = TRUE)) %>% 
  ggplot(aes(y = location,
             x = mean_velocity))+
  geom_bar(position = 'stack', stat = 'identity') +
  labs(title = "Mean Water Velocity by Location",
       x = 'Water Velocity (m/s)')
```

```{r}
cleaner_data %>% 
  group_by(date) %>%
  mutate(avg_velocity = mean(velocity, na.rm = T)) %>%
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = avg_velocity, color = year)) + 
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Water Velocity (colored by year)",
       y = "Average daily velocity", 
       x = "Date")  
```
```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = velocity, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Water velocity summarized by year",
       x = "Water Velocity") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of velocity 1999 to 2009**

```{r}
summary(cleaner_data$velocity)
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'velocity'))/nrow(cleaner_data), 3)*100` % of values in the `velocity` column are NA.

### Variable:`trap_revolutions`

**Description: Number of trap revolutions corresponding to a particular catch.**

```{r}
#Filter out significant high outliers (revolution > 10,000 )
revolution_plot_1 <- cleaner_data %>% 
  ggplot(aes(x=trap_revolutions))+
  geom_histogram(binwidth = 1000)+
  theme_minimal()+
  labs(title = "Trap Revolutions Distribution Without Filtering")+
  theme(plot.title = element_text(size = 10))
```

```{r}
#Filter out significant significant higher values (277 points at revolution between 112178 rpms and 890793 rpms )
revolution_plot_2 <- cleaner_data %>% 
  filter(trap_revolutions < 100000) %>% 
  ggplot(aes(x=trap_revolutions))+
  geom_histogram(binwidth = 1000)+
  theme_minimal()+
  labs(title = "Trap Revolutions Distribution After Filtering")+
  theme(plot.title = element_text(size = 10))
```
```{r}
grid.arrange(revolution_plot_1, revolution_plot_2, ncol=2)
```

The trap_revolutions have a 277 points at revolution between 112178 rpm and 890793 rpm. These points are significantly higher than the rest of the trap_revolutions which are below 10000 rpm.

**Numeric Summary of trap_revolutions 1999 to 2009**
```{r}
summary(cleaner_data$trap_revolutions)
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'trap_revolutions'))/nrow(cleaner_data), 3)*100` % of values in the `trap_revolutions` column are NA.

### Variable:`trap_revolutions2`

**Description: Number of trap revolutions corresponding to a particular catch, results from second counter.**
```{r}
revolution2_plot_1 <- cleaner_data %>% 
  ggplot(aes(x=trap_revolutions2))+
  geom_histogram(binwidth = 1000)+
  theme_minimal()+
  labs(title = "Trap Revolutions2 Distribution Before Filtering")+
  theme(plot.title = element_text(size = 10))
```

```{r}
#Filter out significant significant higher values (220 points at revolution between 112178 rpms and 890793 rpms )
revolution2_plot_2 <- cleaner_data %>% 
  filter(trap_revolutions2 < 100000) %>%
  ggplot(aes(x=trap_revolutions2))+
  geom_histogram(binwidth = 1000)+
  theme_minimal()+
  labs(title = "Trap Revolutions2 Distribution After Filtering")+
  theme(plot.title = element_text(size = 10))
```
```{r}
grid.arrange(revolution2_plot_1, revolution2_plot_2, ncol=2)
```

The trap_revolutions have a 220 points at revolution between 112178 rpm and 890793 rpm. These points are significantly higher than the rest of the trap_revolutions which are below 10000 rpm.

**Numeric Summary of trap_revolutions2 1999 to 2009**
```{r}
summary(cleaner_data$trap_revolutions2)
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'trap_revolutions2'))/nrow(cleaner_data), 3)*100` % of values in the `trap_revolutions2` column are NA.

### Variable:`rpms_before`

**Description: Revolutions per minute before the trap was sampled.**

```{r}
cleaner_data %>% 
  filter(!is.na(location)) %>% 
  ggplot(aes(x = rpms_before, y = location)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "RPMs Before summarized by locations",
       x = "rpms_before") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of rpms_before 1999 to 2009**
```{r}
summary(cleaner_data$rpms_before)
```
**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'rpms_before'))/nrow(cleaner_data), 3)*100` % of values in the `rpms_before` column are NA.


### Variable:`rpms_after`

**Description: Revolutions per minute after the trap was sampled**


```{r}
cleaner_data %>% 
  filter(!is.na(location)) %>% 
  filter(rpms_after < 50) %>% #filter out 149 points that had significantly higher value-66 points at 50 rpm, 83 points at 1877 rpm 
  ggplot(aes(x = rpms_after, y = location)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "RPMs After Summarized By Locations",
       x = "rpms_after") + 
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```



**Numeric Summary of rpms_after 1999 to 2009**
```{r}
summary(cleaner_data$rpms_after)
```
**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'rpms_after'))/nrow(cleaner_data), 3)*100` % of values in the `rpms_after` column are NA.


### Variable:`fork_length`

Transforming fork_length = 0 to NA as there shouldn't be any actual 0 fork_length values

```{r}
cleaner_data$fork_length[cleaner_data$fork_length == 0] <- NA
```

```{r}
cleaner_data %>%
  filter(fork_length < 200) %>%  #filter out outlier bigger than 110 for more clear distribution graph
  ggplot(aes(x=fork_length))+
  geom_histogram(breaks = seq(0,110, by =2))+
  labs(title = "Fork Length Distribution")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1 ))
```
```{r}
cleaner_data %>% 
  filter(fork_length < 200) %>% #filtered one outlier (>600) from 2001
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = fork_length, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Fork length summarized by year",
       x = "Fork Length") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
```{r}
cleaner_data %>% 
  filter(fork_length < 200) %>% 
  ggplot(aes(x = fork_length, y = lifestage)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Fork length summarized by lifestage") + 
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of fork_length from 1999 to 2009**
```{r}
summary(cleaner_data$fork_length)
```
**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'fork_length'))/nrow(cleaner_data), 3)*100` % of values in the `fork_length` column are NA.

### Variable:`weight`

**Description:** Unit grams

```{r}
#Filter out 216 points with weight bigger than 12 for a more clear distribution graph
cleaner_data %>% 
  filter(weight<12) %>% 
  ggplot(aes(x=weight))+
  geom_histogram(breaks = seq(0,11, by =0.2))+
  scale_x_continuous(breaks=seq(0, 12, by=1)) +
  labs(title = "Weight Distribution")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust =1 ))
```
```{r}
cleaner_data %>% 
  filter(weight< 60) %>% 
  ggplot(aes(x = weight, y = lifestage)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Weight summarized by lifestage") + 
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Daily Summary of weight from 1999 to 2009**
```{r}
summary(cleaner_data$weight)
```
**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'weight'))/nrow(cleaner_data), 3)*100` % of values in the `weight` column are NA.

### Variable:`count`

```{r}
cleaner_data %>% 
  mutate(water_year = ifelse(month(date) %in% 10:12, year(date) + 1, year(date))) %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  group_by(date) %>%
  mutate(total_daily_catch = sum(count)) %>%
  ungroup() %>%
  ggplot(aes(x = fake_date, y = total_daily_catch)) + 
  geom_col() + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1900-10-01"), as.Date("1901-06-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Total Daily Raw Passage 2000 - 2010",
       y = "Total daily catch",
       x = "Date")+ 
  facet_wrap(~water_year, scales = "free")
```

```{r}
cleaner_data  %>%
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = year, y = count)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Fish Counted Each Year by Run",
       y = "Total fish count") + 
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap(~run, scales = "free_y")
```
```{r}
cleaner_data  %>%
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = year, y = count)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Fish Counted Each Year by Life Stage",
       y = "Total fish count") + 
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap(~lifestage, scales = "free_y")
```


**Numeric Daily Summary of count from 1999 to 2009**
```{r}
summary(cleaner_data$count)
```
**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$'count'))/nrow(cleaner_data), 3)*100` % of values in the `count` column are NA.


### Identified issues with this dataset

* Fish screen diversion trap is only for some years. Maybe could be filtered out when standardizing the data.

* There is significant high temperature days (in the 60s) in the early 2000s. On the same day, different trap locations had temperature difference of up to 45 degrees. 

* TrapRevolutions and TrapRevolutions2 both have about 200 points that are significantly higher (750,000 revolutions vs 30,000 revolutions) 

* Depending on the importance of trap revolution and rpms data, we might need more metadata to describe the variable.

* A few outliers for both weight and forklength measured. 

* 2005 had the highest fish count followed by 2009 and 2008.However some of these values are extremely big such as max count of 85000/day.

### Add cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv

yuba_rst <- cleaner_data %>% glimpse()
write_csv(yuba_rst, "yuba_rst.csv")
```
```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(yuba_rst,
           object_function = f,
           type = "csv",
           name = "rst/yuba-river/data/yuba_rst.csv")
```

