---
title: "Knights Landing Pre 2002"
output: html_document
date: "2022-11-07"
---

```{r, include = F}
library(tidyverse)
library(foreign)
library(googleCloudStorageR)
library(readxl)
library(lubridate)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

# Assumptions

- Assuming mark or clip as "A" means adipose-clipped
- Assuming mark as "B" or "BB" means bismark brown and is a recaptured fish

# 1996

TODO - Need to figure out what codes for lifestage, race, traptype

```{r, include = F}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KLSUM96.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/sum96.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/MARK96.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/mark96.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/RSTFL96.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/rst96.DBF",
                 overwrite = TRUE)
```

- sum table: date, trapid, time, effrt_time, rpm_averag, pt_vel
- mark table: date, trapid, traptype, dye_type, n0_dyed, recap_type n0_recap
- rst table: date, time, trapid, traptype, fl, ww, lifestage, rage, clip

```{r}
path <- "data-raw/qc-markdowns/rst/lower-sac/knights_landing/"

sum96 <- read.dbf(paste0(path, "sum96.DBF")) %>% glimpse()
mark96 <- read.dbf(paste0(path, "mark96.DBF")) %>% glimpse()
rst96 <- read.dbf(paste0(path, "rst96.DBF")) %>% glimpse()
#1996: adfl, sum, mark

rst96 %>% group_by(LIFESTAGE) %>% tally()
rst96 %>% group_by(RACE) %>% tally()
```

```{r}
trap96_clean <- sum96 %>% 
  janitor::clean_names() %>% 
  select(date, trapid, time, effrt_time, rpm_averag, pt_vel) %>% 
  rename(effort = effrt_time,
         avg_rpm = rpm_averag,
         velocity = pt_vel)

efficiency96_clean <- mark96 %>% 
  janitor::clean_names() %>% 
  select(date, n0_dyed, no_recap) %>% 
  group_by(date) %>% 
  summarise(number_marked = sum(n0_dyed, na.rm = T),
            number_recaptured = sum(no_recap, na.rm = T))

rst96_clean <- rst96 %>% 
  janitor::clean_names() %>% 
  mutate(lifestage = as.factor(lifestage),
         adipose_clip = ifelse(clip == "A",T, F)) %>% 
  select(date, time, trapid, fl, ww, lifestage, race, adipose_clip)
```

# 1997

What does A and B mark mean? A for adipose clip? B for bismark brown?

```{r, include = F}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/97KLSUMT.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/sum97.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KLFL97T.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/rst97.DBF",
                 overwrite = TRUE)
```

```{r}
sum97 <- read.dbf(paste0(path, "sum97.DBF")) %>% glimpse()
rst97 <- read.dbf(paste0(path, "rst97.DBF")) %>% glimpse()

```

```{r}
trap97_clean <- sum97 %>% 
  janitor::clean_names() %>% 
  select(date, trapid, time, effrt_time, rpm_averag, pt_vel) %>% 
  rename(effort = effrt_time,
         avg_rpm = rpm_averag,
         velocity = pt_vel)

rst97_clean <- rst97 %>% 
  janitor::clean_names() %>% 
  mutate(lifestage = as.factor(lifestage),
         race = as.factor(race),
         adipose_clip = ifelse(mark == "A", T, F),
         mark = case_when(mark == "A" | mark == 0 ~ F, 
                          mark == "B" | mark == "BB" ~ T)
         ) %>% 
  select(date, time, trapid, fl, ww, lifestage, race, mark, adipose_clip)

rst97_clean %>% group_by(mark) %>% tally()
```

# 1998

What does A and B mark mean? A for adipose clip? B for bismark brown?

```{r, include = F}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KLSUM98.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/sum98.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KLFL98.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/rst98.DBF",
                 overwrite = TRUE)
```

```{r}
sum98 <- read.dbf(paste0(path, "sum98.DBF")) %>% glimpse()
rst98 <- read.dbf(paste0(path, "rst98.DBF")) %>% glimpse()

```

```{r}
trap98_clean <- sum98 %>% 
  janitor::clean_names() %>% 
  select(date, trapid, time, effrt_time, rpm_averag, pt_vel) %>% 
  rename(effort = effrt_time,
         avg_rpm = rpm_averag,
         velocity = pt_vel)

rst98_clean <- rst98 %>% 
  janitor::clean_names() %>% 
  mutate(lifestage = as.factor(lifestage),
         race = as.factor(race),
         adipose_clip = ifelse(mark == "A", T, F),
         mark = case_when(mark == "A" | mark == 0 ~ F, 
                          mark == "B" | mark == "BB" ~ T)) %>% 
  select(date, time, trapid, fl, ww, lifestage, race, mark, mort, adipose_clip)

rst98_clean %>% group_by(mark) %>% tally()
```

# 1999

If we knew what mark was then we could pull out the recaptured fish.

```{r, include = F}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KLSUM99.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/sum99.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KLFL99.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/rst99.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/STAIN99.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/stain99.DBF",
                 overwrite = TRUE)
```

```{r}
sum99 <- read.dbf(paste0(path, "sum99.DBF")) %>% glimpse()
rst99 <- read.dbf(paste0(path, "rst99.DBF")) %>% glimpse()
stain99 <- read.dbf(paste0(path, "stain99.DBF")) %>% glimpse()
```

```{r}
trap99_clean <- sum99 %>% 
  janitor::clean_names() %>% 
  select(date, trapid, time, effrt_time, rpm_averag, pt_vel) %>% 
  rename(effort = effrt_time,
         avg_rpm = rpm_averag,
         velocity = pt_vel)

rst99_clean <- rst99 %>% 
  janitor::clean_names() %>% 
   mutate(lifestage = as.factor(lifestage),
         race = as.factor(race),
         adipose_clip = ifelse(mark == "A", T, F),
         mark = case_when(mark == "A" | mark == 0 ~ F, 
                          mark == "B" | mark == "BB" ~ T)
         ) %>% 
  select(date, time, trapid, fl, ww, lifestage, race, mark, mort, adipose_clip)

mark99_clean <- stain99 %>% 
  janitor::clean_names() %>% 
  select(date_relea, time_relea, cs_release) %>% 
  filter(!is.na(date_relea)) %>% 
  rename(date_released = date_relea,
         time_released = time_relea,
         number_released = cs_release)

recap99_clean <- filter(rst99_clean, mark %in% c("B", "BB"))

rst99_clean %>% group_by(mark) %>% tally()
```


# 2000

```{r, include = F}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/SUM00.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/sum00.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/CSFL00.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/rst00.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/csstaind00.dbf",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/stain00.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/EFFORT00.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/effort00.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/CSBISBRN00.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/bisbrn00.DBF",
                 overwrite = TRUE)
```

```{r}
sum00 <- read.dbf(paste0(path, "sum00.DBF")) %>% glimpse()
rst00 <- read.dbf(paste0(path, "rst00.DBF")) %>% glimpse()
stain00 <- read.dbf(paste0(path, "stain00.DBF")) %>% glimpse()
bisbrn00 <- read.dbf(paste0(path, "bisbrn00.DBF")) %>% glimpse()
# decided we don't need this because info is in the sum table
effort00 <- read.dbf(paste0(path, "effort00.DBF")) %>% glimpse() 
```

```{r}
trap00_clean <- sum00 %>% 
  janitor::clean_names() %>% 
  select(date, trapid, time, effrt_time, rpm_averag, pt_vel) %>% 
  rename(effort = effrt_time,
         avg_rpm = rpm_averag,
         velocity = pt_vel)

rst00_clean <- rst00 %>% 
  janitor::clean_names() %>% 
   mutate(lifestage = as.factor(lifestage),
         race = as.factor(race),
         time = as.factor(time),
         adipose_clip = ifelse(mark == "A", T, F),
         mark = case_when(mark == "A" | mark == 0 ~ F, 
                          mark == "B" | mark == "BB" ~ T)) %>% 
  select(date, time, trapid, fl, ww, lifestage, race, mark, mort, adipose_clip)

mark00_clean <- stain00 %>% 
  janitor::clean_names() %>% 
  select(date, no_cs_rel) %>% 
  filter(!is.na(date)) %>% 
  rename(date_released = date,
         number_released = no_cs_rel)

recap00_clean <- bisbrn00 %>% 
  janitor::clean_names() %>% 
  select(date, time, trapid, fl, ww, lifestage, race) %>% 
  mutate(time = as.factor(time),
         lifestage = as.factor(lifestage),
         race = as.factor(race))

rst00_clean %>% group_by(mark) %>% tally()
```

# 2001

```{r, include = F}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KL01SUM.mdb",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/sum01.mdb",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/KL01CSFL.DBF",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/rst01.DBF",
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/2001_RECAP.XLS",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = "data-raw/qc-markdowns/rst/lower-sac/knights_landing/efficiency01.xls",
                 overwrite = TRUE)
```

```{r}
operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

if(operating_system == "pc") {
  library(RODBC)
  sum01 <- sqlFetch(paste0(path, "sum01.mdb"), "KL01SUM1")
  } else{
    library(Hmisc)
    sum01 <- mdb.get(paste0(path, "sum01.mdb"), tables = "KL01SUM1") %>% glimpse()
  }

rst01 <- read.dbf(paste0(path, "rst01.DBF")) %>% glimpse()
efficiency01 <- read_excel(paste0(path, "efficiency01.xls"), sheet = 2) %>% glimpse()

rst01 %>% group_by(RACE) %>% tally()
```

```{r}
trap01_clean <- sum01 %>% 
  janitor::clean_names() %>% 
  select(date, trap, efft) %>% 
  rename(trapid = trap,
         effort = efft)

rst01_clean <- rst01 %>% 
  janitor::clean_names() %>% 
  select(date, trapid, fl, ww, stage, race, mort) %>% 
  rename(lifestage = stage) %>% 
  mutate(lifestage = as.factor(lifestage),
         race = as.factor(race))

efficiency01_clean <- efficiency01 %>% 
  janitor::clean_names() %>% 
  select(week, number_marked, number_recovered) %>% 
  filter(week != "Total") %>% 
  mutate(year = 2001) %>% 
  rename(number_recaptured = number_recovered)
```

# Combine 1996-2001

trap: date, time, trapid, effort, avg_rpm, velocity
rst: date, trapid, fl, ww, lifestage, race, mort, clip, mark, id_num
efficiency: date, week, year, number_marked, number_recaptured
marked: date_released, time_released, number_released
recaptured: date, time, trapid, fl, ww, lifestage, race

## trap
```{r}
trap_combined <- bind_rows(trap96_clean,
                           trap97_clean,
                           trap98_clean,
                           trap99_clean,
                           trap00_clean,
                           trap01_clean) %>% glimpse()
```
### time
```{r}
trap_combined %>% group_by(year = year(date), time) %>% tally()
```

### trapid

We will need to figure out how to assign these trapids to the current 8.3 and 
8.4 sites for Knights Landing

```{r}
unique(trap_combined$trapid)
trap_combined %>% group_by(year = year(date), trapid) %>% tally()
```

### effort
```{r}
trap_combined %>% 
  ggplot(aes(x = as.factor(year(date)), y = effort)) +
  geom_boxplot()
```

### avg_rpm
```{r}
trap_combined %>% 
  ggplot(aes(x = as.factor(year(date)), y = avg_rpm)) +
  geom_boxplot()
```

### velocity
```{r}
trap_combined %>% 
  ggplot(aes(x = as.factor(year(date)), y = velocity)) +
  geom_boxplot()
```
## rst 
```{r}
rst_combined <- bind_rows(rst96_clean,
                          rst97_clean,
                          rst98_clean,
                          rst99_clean,
                          rst00_clean,
                          rst01_clean) %>% glimpse()
```
### time
```{r}
rst_combined %>% group_by(year = year(date), time) %>% tally()
```

### trapid
```{r}
rst_combined %>% group_by(year = year(date), trapid) %>% tally()
```

### fl
```{r}
rst_combined %>% 
  ggplot(aes(x = as.factor(year(date)), y = fl)) +
  geom_boxplot()

rst_combined %>% 
  filter(fl < 250) %>% 
  ggplot(aes(x = as.factor(year(date)), y = fl)) +
  geom_boxplot()
```
### ww
```{r}
rst_combined %>% 
  ggplot(aes(x = as.factor(year(date)), y = ww)) +
  geom_boxplot()

rst_combined %>% 
  filter(ww < 20) %>% 
  ggplot(aes(x = as.factor(year(date)), y = ww)) +
  geom_boxplot()
```

### lifestage

We need to assign labels to these codes

```{r}
rst_combined %>% 
  group_by(year = year(date), lifestage) %>% 
  tally() %>% 
  ggplot(aes(x = year, y = n, fill = lifestage)) +
  geom_col()
```

### race

We need to assign labels to these codes

```{r}
rst_combined %>% 
  group_by(year = year(date), race) %>% 
  tally() %>% 
  ggplot(aes(x = year, y = n, fill = race)) +
  geom_col()
```

### adipose_clip
```{r}
rst_combined %>% 
  group_by(year = year(date), adipose_clip) %>% 
  tally() %>% 
  ggplot(aes(x = year, y = n, fill = adipose_clip)) +
  geom_col()
```

### mark
```{r}
unique(rst_combined$mark)

rst_combined %>% 
  group_by(year = year(date), mark) %>% 
  tally() %>% 
  ggplot(aes(x = year, y = n, fill = mark)) +
  geom_col()
```

### mort
```{r}
unique(rst_combined$mort)

rst_combined %>% 
  group_by(year = year(date), mort) %>% 
  tally() %>% 
  ggplot(aes(x = year, y = n, fill = mort)) +
  geom_col()
```

## efficiency 

Ton of fish were marked in week 8 and 19 in 2001

```{r}
efficiency_combined <- bind_rows(efficiency96_clean,
                                 efficiency01_clean) %>% glimpse()
```

```{r}
efficiency_combined %>% 
  mutate(year = ifelse(is.na(year), year(date), year),
         week = ifelse(is.na(week), week(date), week)) %>% 
  pivot_longer(cols = c(number_marked, number_recaptured), names_to = "type", values_to = "number") %>% 
  ggplot(aes(x = week, y = number, fill = type)) +
  geom_col() +
  facet_wrap(~year)
```


## marked

- Ton of fish released prior to week 10 in 2000
- Few releases in 1998

```{r}
marked_combined <- bind_rows(mark99_clean,
                             mark00_clean) %>% glimpse()
```

```{r}
marked_combined %>% 
  mutate(year = year(date_released),
         week = week(date_released)) %>% 
  ggplot(aes(x = week, y = number_released)) +
  geom_col() +
  facet_wrap(~year)
```
  

## recap
```{r}
recap_combined <- bind_rows(recap99_clean,
                            recap00_clean) %>% glimpse()
```

### time
```{r}
recap_combined %>% group_by(year = year(date), time) %>% tally()
```

### trapid
```{r}
recap_combined %>% group_by(year = year(date), trapid) %>% tally()
```

### fl
```{r}
recap_combined %>% 
  ggplot(aes(x = as.factor(year(date)), y = fl)) +
  geom_boxplot()
```
### ww
```{r}
recap_combined %>% 
  ggplot(aes(x = as.factor(year(date)), y = ww)) +
  geom_boxplot()
```

### lifestage

We need to assign labels to these codes

```{r}
recap_combined %>% 
  group_by(year = year(date), lifestage) %>% 
  tally() %>% 
  ggplot(aes(x = as.factor(year), y = n, fill = lifestage)) +
  geom_col()
```

### race

We need to assign labels to these codes

```{r}
recap_combined %>% 
  group_by(year = year(date), race) %>% 
  tally() %>% 
  ggplot(aes(x = as.factor(year), y = n, fill = race)) +
  geom_col()
```


### mort
```{r}
unique(recap_combined$mort)

recap_combined %>% 
  group_by(year = year(date), mort) %>% 
  tally() %>% 
  ggplot(aes(x = as.factor(year), y = n, fill = mort)) +
  geom_col()
```
# edits

Assume that trapid is 1 or 2 is 8.3, trapid is 3 or 4 is 8.4. According to catch
data trapid is either 2 or 3 in earlier years. Need to make consistent with later
years.

## trap

```{r}
trap_clean <- trap_combined %>% 
  mutate(trapid = case_when(trapid %in% c(1,2) ~ 8.3,
                            trapid %in% c(3,4) ~ 8.4))

unique(trap_clean$trapid)
```

## rst

reformat trapid

```{r}
rst_clean <- rst_combined %>% 
  mutate(mort = ifelse(mort == 1, T, F),
         trapid = case_when(trapid %in% c(1,2) ~ 8.3,
                            trapid %in% c(3,4) ~ 8.4))

unique(rst_clean$trapid)
```

## efficiency

```{r}
# add 1999 and 2000 to efficiency
number_recaptured <- recap_combined %>% 
  group_by(week = week(date),
           year = year(date)) %>% 
  tally() %>% 
  rename(number_recaptured = n)

number_marked <- marked_combined %>% 
  mutate(week = week(date_released),
           year = year(date_released))

eff <- left_join(number_marked, number_recaptured)

efficiency_clean <- left_join(number_marked, number_recaptured) %>% 
              select(date_released, week, year, number_released, number_recaptured) %>% 
              mutate(number_recaptured = as.numeric(ifelse(is.na(number_recaptured), 0, number_recaptured)),
                     number_released = as.numeric(number_released)) %>% 
  bind_rows(efficiency_combined %>% 
  mutate(year = ifelse(is.na(year), year(date), year),
         week = as.numeric(ifelse(is.na(week), week(date), week))) %>% 
  rename(date_released = date,
         number_released = number_marked))
```

# save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(trap_clean,
           object_function = f,
           type = "csv",
           name = "rst/lower-sac-river/data/knights-landing/pre-2002-trap.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(rst_clean,
           object_function = f,
           type = "csv",
           name = "rst/lower-sac-river/data/knights-landing/pre-2002-rst.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(efficiency_clean,
           object_function = f,
           type = "csv",
           name = "rst/lower-sac-river/data/knights-landing/pre-2002-efficiency.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(marked_combined,
           object_function = f,
           type = "csv",
           name = "rst/lower-sac-river/data/knights-landing/pre-2002-mark.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(recap_combined,
           object_function = f,
           type = "csv",
           name = "rst/lower-sac-river/data/knights-landing/pre-2002-recap.csv",
           predefinedAcl = "bucketLevel")
```
