---
title: "Deer Creek RST Data QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Deer Creek RST Data

## Description of Monitoring Data
RST data was shared with us in an access database. I queried it to include both fish catch and trap conditions information. 

**Timeframe:** 1992 - 2010 

**Completeness of Record throughout timeframe:** 

* Very little environmental data collected in beginning of timeframe
* Very little trap efficiency data collected in beginning of timeframe 


**Sampling Location:** 

4 sites on Deer Creek: Deer Creek Canyon Mouth, Deer Creek Canyon Mouth Trap, Deer Creek Canyon Mouth Diversion, Deer Creek RST 

**Data Contact:** [Matt Johnson](Matt.Johnson@wildlife.ca.gov)

Comments from Matt: Let me know if you need any help with interpreting what is in that database.  It is not mine, I inherited it from a retired CDFW biologist, but I did work on those RST’s as a Scientific Aide, and entered data into the database.  So I at least have some working knowledge of the forms/tables in there.  

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
gcs_get_object(object_name = 
                 "rst/deer-creek/data-raw/mill_deer_rst_trap_catch_and_conditions.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_rst_deer_and_mill.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data: 
```{r}
raw_rst_data <- read_excel("raw_rst_deer_and_mill.xlsx", col_types =c("date", "text", "numeric", "text", 
                                                                      "numeric", "numeric", "numeric", 
                                                                      "numeric", "numeric", "numeric", 
                                                                      "numeric", "numeric", "numeric", 
                                                                      "text")) %>% glimpse()
unique(raw_rst_data$Location)
```

## Data transformations

Filter to just include Deer Creek Locations and chinook, clean names, and edit column types. 

```{r}
cleaner_rst_data <- raw_rst_data%>% 
  filter(Location %in% c("Deer Creek Canyon Mouth", "Deer Creek Canyon Mouth Diversion", 
                         "Deer Creek Canyon Mouth Trap", "Deer Creek RSTR"),
         Species == "CHISAL") %>%
  janitor::clean_names() %>% 
  rename(fork_length = length) %>% # confirmed with matt that this is fork length
  mutate(date = as.Date(date),
         trap_condition_code = as.character(trap_condition_code), 
         weather = as.character(weather)) %>%
  select(-species) %>%
  glimpse()
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_rst_data %>% select_if(is.numeric) %>% colnames()
```


### Variable: `count`

**Plotting fish counts over period of record**
```{r, include=FALSE}
sac_indices <- waterYearType::water_year_indices %>% 
    filter(location == "Sacramento Valley") %>% 
    transmute(water_year = WY, year_type = as.character(Yr_type))
```

```{r}
cleaner_rst_data %>% 
  group_by(date) %>%
  summarise(total_daily_catch = sum(count, na.rm = T)) %>%
  mutate(water_year = ifelse(month(date) %in% 10:12, year(date) + 1, year(date))) %>% 
  left_join(sac_indices) %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  filter(water_year < 2021) %>%
  ungroup() %>%
  ggplot(aes(x = fake_date, y = total_daily_catch, fill = year_type)) + 
  geom_col() + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1900-10-01"), as.Date("1901-06-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Total Daily Raw Catch",
       y = "Total daily raw catch",
       x = "Date")+ 
  facet_wrap(~water_year, scales = "free") +
  scale_color_manual(palette)
```

  
Early years looks like sampling happened infrequently. 

```{r}
cleaner_rst_data  %>%
  mutate(year = as.factor(year(date))) %>%
  group_by(year) %>%
  summarise(total_yearly_catch = sum(count, na.rm = T)) %>%
  ggplot(aes(x = year, y = total_yearly_catch)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Fish Counted each Year by run",
       y = "Total fish raw catch") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```



**Numeric Summary of counts over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$count)


daily_count <- cleaner_rst_data %>% group_by(date) %>% summarise(daily_count = sum(count, na.rm = T))
summary(daily_count$daily_count)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$count))/nrow(cleaner_rst_data), 3) * 100` % of values in the `count` column are NA.


### Variable: `fork_length`

fork_length of the fish captured. Units: mm 				

**Plotting fork fork_length**
  
```{r}
cleaner_rst_data %>% 
  ggplot(aes(x = fork_length)) + 
  geom_histogram(breaks=seq(0, 200, by=2)) + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  theme_minimal() +
  labs(title = "Fork Length distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of fork length over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$fork_length)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$fork_length))/nrow(cleaner_rst_data), 3) * 100` % of values in the `fork_length` column are NA. 

### Variable: `weight`

Weight of the fish captured. Units: grams			

**Plotting weight**
  
```{r}
cleaner_rst_data %>% 
  ggplot(aes(x = weight)) + 
  geom_histogram() + 
  theme_minimal() +
  labs(title = "Weight distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of weight over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$weight)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$weight))/nrow(cleaner_rst_data), 3) * 100` % of values in the `weight` column are NA. 

### Variable: `flow`

Flow, cfs

**Plotting flow over Period of Record**
  
```{r}
# Make whatever plot is appropriate 
# maybe 2 plots is appropriate
cleaner_rst_data %>% 
  group_by(date) %>%
  mutate(avg_flow = mean(flow)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_flow, color = year)) + 
  geom_point(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Flow Measures (colored by year)",
       x = "Date", 
       y = "Average Daily Flow")  
```
```{r}
cleaner_rst_data %>% 
  ggplot(aes(x = flow)) + 
  geom_histogram() + 
  theme_minimal() +
  labs(title = "Flow distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_data %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  ggplot(aes(x = fake_date, y = flow)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("Date") +
  facet_wrap(~wy, scales = "free_y") + 
  theme_minimal()
```


Notes: 

* No measures pre 1998
* Even after 1998 a lot of data gaps

**Numeric Summary of flow over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$flow)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$flow))/nrow(cleaner_rst_data), 3) * 100` % of values in the `flow` column are NA.



### Variable: `time_for_10_revolutions`
		

**Plotting time it takes RST to do 10 revolutions**
  
```{r}
cleaner_rst_data %>% 
  ggplot(aes(x = time_for_10_revolutions)) + 
  geom_histogram() + 
  # scale_x_continuous(breaks=seq(0, 200, by=25)) +
  theme_minimal() +
  labs(title = "Time it takes to do 10 revolutions distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
A lot of spread in this time. 0 seems like a non functional trap or could be NA coded as 0 (although there are already NA values). Very high values seem like outliers. 

**Numeric Summary of time_for_10_revolutions over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$time_for_10_revolutions)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$time_for_10_revolutions))/nrow(cleaner_rst_data), 3) * 100` % of values in the `time_for_10_revolutions` column are NA. 


### Variable: `tubs_of_debris`

**Plotting Debris**
  
```{r}
cleaner_rst_data %>% 
  ggplot(aes(x = tubs_of_debris)) + 
  geom_histogram() + 
  theme_minimal() +
  labs(title = "Tub of debris distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of tub of debris over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$tubs_of_debris)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$tubs_of_debris))/nrow(cleaner_rst_data), 3) * 100` % of values in the `tubs_of_debris` column are NA. 



### Variable: `water_temperature`

Water Tempearture in degrees F

**Plotting water_temperature over Period of Record**
  
```{r}
# Make whatever plot is appropriate 
# maybe 2 plots is appropriate
cleaner_rst_data %>% 
  group_by(date) %>%
  mutate(avg_temp = mean(water_temperature)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  geom_point(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Temp Measures F (colored by year)",
       x = "Date", 
       y = "Average Daily Temp")  
```
```{r}
cleaner_rst_data %>% 
  ggplot(aes(x = water_temperature)) + 
  geom_histogram() + 
  theme_minimal() +
  labs(title = "Temperature distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_data %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  ggplot(aes(x = fake_date, y = water_temperature)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("Date") +
  facet_wrap(~wy, scales = "free_y") + 
  theme_minimal()
```

Gaps in temperature measurements. 

**Numeric Summary of water_temperature over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$water_temperature)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$water_temperature))/nrow(cleaner_rst_data), 3) * 100` % of values in the `water_temperature` column are NA.



### Variable: `turbidity`

turbidity, NTU's 

**Plotting turbidity over Period of Record**
  
```{r}
# Make whatever plot is appropriate 
# maybe 2 plots is appropriate
cleaner_rst_data %>% 
  group_by(date) %>%
  mutate(avg_turbidity = mean(turbidity)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_turbidity, color = year)) + 
  geom_point(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily turbidity Measures (colored by year)",
       x = "Date", 
       y = "Average Daily turbidity NTUs")  
```
```{r}
cleaner_rst_data %>% 
  ggplot(aes(x = turbidity)) + 
  geom_histogram() + 
  theme_minimal() +
  labs(title = "Turbidity distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_data %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  ggplot(aes(x = fake_date, y = turbidity)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("Date") +
  facet_wrap(~wy, scales = "free_y") + 
  theme_minimal()
```

Notes: 

* No measures pre 1999, 2007 or 2008
* One outlier greater than 150 

**Numeric Summary of turbidity over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_data$turbidity)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_data$turbidity))/nrow(cleaner_rst_data), 3) * 100` % of values in the `turbidity` column are NA.

## Explore Categorical variables: {.tabset}


```{r}
cleaner_rst_data %>% select_if(is.character) %>% colnames()
```


### Variable: `location`
```{r}
table(cleaner_rst_data$location) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_data$location))/nrow(cleaner_rst_data), 3) * 100` % of values in the `location` column are NA.

### Variable: `trap_condition_code`

Code describing trap condition.  

Trap condition codes are referenced by the following numerals only:  1,2,3 and 4.  Any other numerals in that field are data entry errors.  1= normal, 2 = partial blockage, 3 = total blockage, and 4 = cone stopped.

```{r}
table(cleaner_rst_data$trap_condition_code) 
```

Add in definitions in place of codes: 

```{r}
cleaner_rst_data$trap_condition_code <- case_when(
  cleaner_rst_data$trap_condition_code == 1 ~ "normal", 
  cleaner_rst_data$trap_condition_code == 2 ~ "partial blockage", 
  cleaner_rst_data$trap_condition_code == 3 ~ "total blockage",
  cleaner_rst_data$trap_condition_code == 4 ~ "cone stopped")

table(cleaner_rst_data$trap_condition_code) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_data$trap_condition_code))/nrow(cleaner_rst_data), 3) * 100` % of values in the `trap_condition_code` column are NA.

### Variable: `weather`

Code describing weather condition. 

Weather codes are referenced by the following numerals only:  1,2,3,4,5 and 6.  Any other numerals in that field are data entry errors.  1= sunny, 2= partly cloudy, 3= cloudy, 4= rain, 5= snow, and 6= fog.

```{r}
table(cleaner_rst_data$weather) 
```

Add in definitions in place of codes: 

```{r}
cleaner_rst_data$weather <- case_when(
  cleaner_rst_data$weather == 1 ~ "sunny", 
  cleaner_rst_data$weather == 2 ~ "partly cloudy", 
  cleaner_rst_data$weather == 3 ~ "cloudy",
  cleaner_rst_data$weather == 4 ~ "rain",
  cleaner_rst_data$weather == 5 ~ "snow", 
  cleaner_rst_data$weather == 6 ~ "fog")

table(cleaner_rst_data$weather) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_data$weather))/nrow(cleaner_rst_data), 3) * 100` % of values in the `weather` column are NA.

### Variable: `comments`

First 5 unique comments displayed to see what info they contain

```{r}
unique(cleaner_rst_data$comments)[1:5] 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_data$comments))/nrow(cleaner_rst_data), 3) * 100` % of values in the `comments` column are NA.

## Summary of identified issues

* Lots of environmental variables are not collected before 1997
* Lots of sampling gaps

## Save cleaned data back to google cloud 

```{r}
deer_rst <- cleaner_rst_data %>% glimpse()
```

```{r, eval = FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(deer_rst,
           object_function = f,
           type = "csv",
           name = "rst/deer-creek/data/deer_rst.csv")
```