---
title: "tisdale efficiency data"
output: html_document
date: "2022-09-02"
---

```{r, include = F}
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(knitr)
library(hms)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

# Lower Sacramento (Tisdale) Rotary Screw Trap Data - Efficiency Trials

## Description of Monitoring Data

**Timeframe:** 

Efficiency trials conducted:
2012, 2015-2019, 2021-2022

**Completeness of Record throughout timeframe:** 

No trials conducted before 2012; no trials in 2013, 2014 and 2020

**Sampling Location:**

Tisdale

**Data Contact:** 

[Drew Huneycutt](mailto::andrew.huneycutt@wildlife.ca.gov)

Drew provided the data in an excel spreadsheet that was slightly modified to make
reading the file easier. The original is saved on the google cloud here: 
"rst/lower-sac-river/data-raw/tisdale/tisdale_trap_efficiency.xlsx"

Tisdale efficiency trials may have been originally recorded on paper and then
transferred to this spreadsheet.

No information about fork length of released or recaptured fish or days after release
were provided. Information about the fork length of recaptured fish may be included in
CAMP.

TODO ask Drew if there he has fork lengths of released fish.

## Data pull
```{r}
# # Data pull ---------------------------------------------------------------
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```
### Data from excel
```{r}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/tisdale/tisdale_trap_efficiency_format.csv",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = here::here("data-raw", "qc-markdowns", "rst", "tisdale_trap_efficiency.csv"),
                 overwrite = TRUE)
```
```{r}
tisdale_efficiency_raw <- read_csv(here::here("data-raw", "qc-markdowns", "rst", "tisdale_trap_efficiency.csv"))
```

### Data from CAMP
```{r}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/tisdale/tisdale_releases_camp.csv",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = here::here("data-raw", "qc-markdowns", "rst", "tisdale_releases_camp.csv"),
                 overwrite = TRUE)

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/tisdale/tisdale_recaptures_camp.csv",
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = here::here("data-raw", "qc-markdowns", "rst", "tisdale_recaptures_camp.csv"),
                 overwrite = TRUE)
```
```{r}
tisdale_releases_raw <- read_csv(here::here("data-raw", "qc-markdowns", "rst", "tisdale_releases_camp.csv"))
tisdale_recaptures_raw <- read_csv(here::here("data-raw", "qc-markdowns", "rst", "tisdale_recaptures_camp.csv"))
```

## Cleaning and formatting

Data from CAMP has more information for recaptures so use that instead of the
excel data where available.

```{r}
# excel data
tisdale_efficiency_clean <- tisdale_efficiency_raw %>% 
  rename(date_released = `Release Date`,
         number_released = `Number Released`,
         number_recaptured = `Number Recaptured`) %>% 
  mutate(date_released = as.Date(date_released, "%m/%d/%y"),
         stream = "sacramento river",
         site = "tisdale")

# CAMP data
tisdale_releases_camp <- tisdale_releases_raw %>% 
  select(stream, site, releaseID, nReleased, releaseTime, markedRun, markedLifeStage, sourceOfFishSite, includeTest, markedFishOrigin, releaseLightCondition, releaseSite, releaseSubSite) %>% 
  filter(!releaseID %in% c(0,255)) %>% 
  rename(release_id = releaseID,
         number_released = nReleased,
         date_released = releaseTime,
         run_released = markedRun,
         origin_released = markedFishOrigin,
         lifestage_released = markedLifeStage,
         light_condition = releaseLightCondition,
         source_released = sourceOfFishSite,
         site_released = releaseSite,
         subsite_released = releaseSubSite,
         include = includeTest) %>% 
  mutate(time_released = as_hms(date_released),
         date_released = as.Date(date_released)) %>% 
  # there are duplicates in the camp data for some reason
  group_by(date_released, time_released, release_id) %>% 
  distinct()

tisdale_recaptures_camp <- tisdale_recaptures_raw %>% 
  mutate(subsite = tolower(subSiteName)) %>% 
  select(stream, site, subsite, catchRawID, forkLength, weight, n, releaseID, run, final_run, fishOrigin, visitTime, lifeStage) %>% 
  mutate(run = ifelse(!is.na(final_run), final_run, run)) %>% 
  select(-final_run) %>% 
  rename(release_id = releaseID,
         fork_length = forkLength,
         number_recaptured = n,
         origin_recaptured = fishOrigin,
         date_recaptured = visitTime,
         run_recaptured = run,
         lifestage_recaptured = lifeStage) %>% 
  mutate(time_recaptured = as_hms(date_recaptured),
         date_recaptured = as.Date(date_recaptured),
         release_id = as.character(release_id))

# checking full data
tisdale_recaptures_summary <- tisdale_recaptures_camp %>% 
              group_by(release_id) %>% 
              summarize(number_recaptured = sum(number_recaptured),
                        median_fork_length = median(fork_length, na.rm = T))
tisdale_efficiency_camp <- tisdale_releases_camp %>% 
  mutate(release_id = as.character(release_id)) %>% 
  left_join(tisdale_recaptures_summary)

# combining excel and CAMP data
tisdale_releases <- tisdale_efficiency_clean %>% 
  select(date_released, number_released, stream, site) %>% 
  full_join(tisdale_releases_camp %>% 
              select(release_id, date_released, time_released, number_released, run_released, origin_released, light_condition, site_released, subsite_released, stream, site, lifestage_released, source_released, include)) %>% 
  mutate(release_id = ifelse(is.na(release_id), paste0("TIS", row_number()), release_id))

# check that no duplicates
tisdale_releases %>% 
  group_by(date_released) %>% 
  tally() %>% 
  filter(n>1)

tisdale_excel_recaptures <- tisdale_efficiency_clean %>% 
  full_join(tisdale_releases %>% 
              select(stream, site, date_released, number_released, release_id)) %>% 
  filter(grepl("TIS", release_id)) %>% 
  mutate(date_recaptured = date_released + 1) %>% 
  ungroup() %>% 
  select(-date_released,-number_released)

tisdale_recaptures <- tisdale_recaptures_camp %>% 
  select(release_id, number_recaptured, fork_length, weight, run_recaptured, origin_recaptured, date_recaptured, time_recaptured, lifestage_recaptured, stream, site, subsite) %>% 
  bind_rows(tisdale_excel_recaptures)
  
```

## Exploratory plots
```{r}
tisdale_efficiency_clean %>% 
  mutate(efficiency = number_recaptured/number_released) %>% 
  ggplot(aes(x = date_released, y = efficiency)) +
  geom_point()

```

```{r}
tisdale_efficiency_clean %>% 
  mutate(efficiency = number_recaptured/number_released,
         wy = ifelse(month(date_released) %in% 10:12, year(date_released) + 1, year(date_released)),
         wy = as.factor(wy)) %>% 
  ggplot(aes(x = wy, y = efficiency, fill = stream)) +
  geom_boxplot()
```

```{r}
tisdale_efficiency_clean %>% 
  mutate(efficiency = number_recaptured/number_released,
         wy = ifelse(month(date_released) %in% 10:12, year(date_released) + 1, year(date_released)),
         wy = as.factor(wy)) %>% 
  group_by(wy) %>% 
  summarise(n = length(unique(date_released)),
            mean_efficiency = mean(efficiency)) %>%  
  ggplot(aes(x = wy, y = mean_efficiency, fill = n)) +
  geom_col()
```

## Save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(tisdale_releases,
           object_function = f,
           type = "csv",
           name = "rst/lower-sac-river/data/tisdale/tisdale_releases.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(tisdale_recaptures,
           object_function = f,
           type = "csv",
           name = "rst/lower-sac-river/data/tisdale/tisdale_recaptures.csv",
           predefinedAcl = "bucketLevel")
```
