---
title: "Battle and Clear Creek data from EDI"
author: "ashley"
date: "1/17/2024"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(EDIutils)

color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#8C5899")
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

This script pulls data from EDI for Battle and Clear creeks. Previously data had been provided in another format. FlowWest then worked with USFWS (Natasha Wingerter and Mike Schraml) to prepare an EDI data package. We believe the data on EDI are more complete and robust than the data previously provided. This script is replacing the Battle and Clear RST QC markdowns.

```{r, include =F }
# Pull in data from EDI
res <- read_data_entity_names(packageId = "edi.1509.1")
# catch
raw <- read_data_entity(packageId = "edi.1509.1", entityId = res$entityId[1])
catch <- readr::read_csv(file = raw)
#trap
raw <- read_data_entity(packageId = "edi.1509.1", entityId = res$entityId[2])
trap <- readr::read_csv(file = raw)
#recapture
raw <- read_data_entity(packageId = "edi.1509.1", entityId = res$entityId[3])
recapture <- readr::read_csv(file = raw)
#release
raw <- read_data_entity(packageId = "edi.1509.1", entityId = res$entityId[4])
release <- readr::read_csv(file = raw)
```

# Catch

```{r}
# TODO what is subsample? should i include? 
# TODO what is powerhous battle?
# TODO should the NA counts be 0s? These likely have been estimated and have values for r_catch. Should we use r_catch?
# TODO there is no adipose_clipped or hatchery variable in the data

catch_format <- catch |> 
  select(sample_date, station_code, count, fork_length, weight, fws_run, interp, common_name, life_stage) |> 
  rename(date = sample_date,
         site = station_code,
         run = fws_run,
         interpolated = interp,
         species = common_name,
         lifestage = life_stage) |> 
  mutate(stream = case_when(grepl("clear", site) ~ "clear creek",
                            grepl("battle", site) ~ "battle creek"),
         site = case_when(site == "upper battle creek" ~ "ubc",
                          site == "lower battle creek" ~ "lbc",
                          site == "lower clear creek" ~ "lcc", 
                          site == "upper clear creek" ~ "ucc"),
         subsite = site,
         species = tolower(species),
         lifestage = case_when(lifestage == "not provided" ~ "not recorded",
                               lifestage == "obvious fry" ~ "fry",
                               T ~ lifestage),
         run = case_when(is.na(run) & count > 0 ~ "not recorded",
                         T ~ run))
  
```

## QA/QC {.tabset}

### run

```{r, run_qaqc}
unique(catch_format$run)

total_count <- catch_format %>%
  mutate(year = year(date)) %>%
  group_by(site, year) %>%
  summarize(total = sum(count, na.rm = T))

catch_format %>%
  mutate(year = year(date)) %>%
  group_by(site, year, run) %>%
  summarize(count = sum(count, na.rm = T)) %>%
  left_join(total_count) %>%
  mutate(percent = (count/total)*100) %>%
  ggplot(aes(x = year, y = percent, fill = run)) +
  geom_col() +
  facet_wrap(~site, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### fork_length

Figure below shows the percent of total fish counted that have a fork length measurement.

```{r, forklength_qaqc}
catch_format %>%
  filter(!is.na(fork_length)) %>%
  mutate(year = year(date)) %>%
  group_by(site, year) %>%
  summarize(count = sum(count, na.rm = T)) %>%
  left_join(total_count) %>%
  mutate(percent = (count/total)*100) %>%
  ggplot(aes(x = year, y = percent)) +
  geom_col() +
  facet_wrap(~site)
  
```

### lifestage

Figure below shows the percentage by lifestage for each tributary and year.

```{r, lifestage_qaqc}
catch_format %>%
  mutate(year = year(date)) %>%
  group_by(site, year, lifestage) %>%
  summarize(count = sum(count, na.rm = T)) %>%
  left_join(total_count) %>%
  mutate(percent = (count/total)*100) %>%
  ggplot(aes(x = year, y = percent, fill = lifestage)) +
  geom_col() +
  facet_wrap(~site, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
### count

```{r, count_qaqc}
filter(catch_format, is.na(count)) %>% glimpse()
```
### interpolated

```{r, interpolated_qaqc}
catch_format %>%
  filter(interpolated == T) %>%
  mutate(year = year(date)) %>%
  group_by(site, year) %>%
  summarize(count = sum(count, na.rm = T)) %>%
  left_join(total_count) %>%
  mutate(percent = (count/total)*100) %>%
  ggplot(aes(x = year, y = percent)) +
  geom_col() +
  facet_wrap(~site)
```

### adipose_clipped

```{r, origin_qaqc}
# catch_format %>%
#   mutate(year = year(date)) %>%
#   group_by(site, year, adipose_clipped) %>%
#   summarize(count = sum(count, na.rm = T)) %>%
#   left_join(total_count) %>%
#   mutate(percent = (count/total)*100) %>%
#   ggplot(aes(x = year, y = percent, fill = adipose_clipped)) +
#   geom_col() +
#   facet_wrap(~site, scales = "free_y") +
#   theme_minimal() +
#   theme(legend.position = "bottom")
```

### weight

Figure below shows the percent of total fish counted that have a weight measurement.

```{r, weight_qaqc}
catch_format %>%
  filter(!is.na(weight)) %>%
  mutate(year = year(date)) %>%
  group_by(site, year) %>%
  summarize(count = sum(count, na.rm = T)) %>%
  left_join(total_count) %>%
  mutate(percent = (count/total)*100) %>%
  ggplot(aes(x = year, y = percent)) +
  geom_col() +
  facet_wrap(~site)
```
## Save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(catch_format,
           object_function = f,
           type = "csv",
           name = "rst/battle_clear_catch_edi.csv",
           predefinedAcl = "bucketLevel")
```


# Trap

```{r}
# TODO the QC issues noted in code below have persisted so might be worth contacting them so they can fix on their end
trap_format <- trap |> 
  select(sample_date, sample_time, trap_start_date, trap_start_time, station_code, depth_adjust, thalweg, cone, end_counter, debris_tubs, gear_condition, avg_time_per_rev, sample_id) |> 
  rename(trap_visit_id = sample_id,
         trap_stop_date = sample_date,
         trap_stop_time = sample_time,
         counter_end = end_counter,
         is_half_cone_configuration = cone,
         in_thalweg = thalweg,
         site = station_code) %>%
  mutate(stream = case_when(grepl("clear", site) ~ "clear creek",
                            grepl("battle", site) ~ "battle creek"),
         site = case_when(site == "upper battle creek" ~ "ubc",
                          site == "lower battle creek" ~ "lbc",
                          site == "lower clear creek" ~ "lcc", 
                          site == "upper clear creek" ~ "ucc"),
         subsite = site,
         # debris in gallons (1 tub = 10 gallons)
         debris_volume = debris_tubs*10,
         is_half_cone_configuration = ifelse(is_half_cone_configuration == 0.5, T, F),
         # convert average time per revolution to revolutions per minute
         rpms_start = 60/avg_time_per_rev,
         # fix typos in trap start date
         trap_start_date = case_when(
                                trap_visit_id == "207_20" ~ as_date("2020-07-24"),
                                trap_visit_id == "049_20" ~ as_date("2020-02-17"),
                                trap_visit_id == "085_19" ~ as_date("2020-03-25"),
                                trap_visit_id == "274_18" ~ as_date("2018-09-20"),
                                trap_visit_id == "006_16" ~ as_date("2016-01-05"),
                                T ~ trap_start_date),
         trap_stop_date = case_when(trap_visit_id == "050_20" & trap_start_date == "2020-02-20" ~ as_date("2020-02-20"),
                                    T ~ trap_stop_date),
  # fill in missing start dates with the stop date from previous day
         trap_start_date = case_when(is.na(trap_start_date) ~ lag(trap_stop_date),
                                T ~ trap_start_date),
         trap_start_time = case_when(is.na(trap_start_time) ~ lag(trap_stop_time), 
                                T ~ trap_start_time)) %>% 
  select(-debris_tubs, -avg_time_per_rev) %>%
  filter(!is.na(trap_start_date)) %>%
  data.frame() %>%
  distinct()

```

## QA/QC {.tabset}

### trap_start_date, trap_sample_date

```{r}
trap_format %>% 
  ggplot(aes(x = trap_stop_date, y = site, color = site)) +
  geom_point()
```

### trap_start_time, trap_sample_time

Does not appear that traps are checked at a specific time

```{r}
trap_format %>% 
  ggplot(aes(x = trap_stop_time, y = site, color = site)) +
  geom_point()
```

### is_half_cone_configuration

Battle and Clear collect data on cone setting. Battle is almost always at full cone
and Clear is at half cone more frequently.

```{r}
trap_format %>%
  filter(!is.na(is_half_cone_configuration)) %>%
  group_by(stream, is_half_cone_configuration) %>%
  tally() %>%
  knitr::kable()
```


### thalweg

Battle and Clear collect information of trap in thalweg and it is almost always TRUE.

```{r}
trap_format %>%
  filter(!is.na(in_thalweg)) %>%
  group_by(stream, in_thalweg) %>%
  tally() %>%
  knitr::kable()
```

### location, site, subsite

```{r}
trap_format %>%
  group_by(stream, site, subsite) %>%
  tally()
```

### debris_volume

```{r}
trap_format %>%
  ggplot(aes(x = trap_stop_date, y = debris_volume, color = stream)) +
  geom_point()

trap_format %>%
  ggplot(aes(x = stream, y = debris_volume)) +
  geom_boxplot()
```

### rpms_start, rpms_end

```{r}
trap_format %>%
  ggplot(aes(x = trap_stop_date, y = rpms_start, color = stream)) +
  geom_point()
```

## Save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(trap_format,
           object_function = f,
           type = "csv",
           name = "rst/battle_clear_trap_edi.csv",
           predefinedAcl = "bucketLevel")
```




# Environmental

```{r}
environmental_format <- trap |> 
  select(sample_date, station_code, velocity, turbidity,  river_left_depth, river_center_depth, river_right_depth) |> 
  pivot_longer(cols = c(velocity, turbidity, river_left_depth, river_center_depth, river_right_depth,), names_to = "parameter", values_to = "value") |> 
  mutate(text = NA_character_) |> 
  rbind(trap |> 
          select(station_code, sample_date, weather, habitat, diel) |> 
          pivot_longer(cols = c(weather, habitat, diel), names_to = "parameter", values_to = "text") |> 
        mutate(value = NA_integer_)) |> 
  rename(date = sample_date,
         site = station_code) |> 
   mutate(stream = case_when(grepl("clear", site) ~ "clear creek",
                            grepl("battle", site) ~ "battle creek"),
         site = case_when(site == "upper battle creek" ~ "ubc",
                          site == "lower battle creek" ~ "lbc",
                          site == "lower clear creek" ~ "lcc", 
                          site == "upper clear creek" ~ "ucc"),
         subsite = site) |> 
  filter(!(is.na(value) & is.na(text)))
```

## QA/QC {.tabset}

### date

```{r}
environmental_format %>% 
  ggplot() + 
  geom_point(aes(x = date, y = site, color = stream), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) + 
  theme(legend.position = "none", 
        text = element_text(size = 18)) 
```

### velocity

```{r}
environmental_format %>% 
  filter(parameter == "velocity") %>% 
  ggplot() + 
  geom_boxplot(aes(x = value, y = stream, color = stream), alpha = .5, binwidth = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

Zoom in on 0 - 10 velocity range:

```{r}
environmental_format %>% 
  filter(parameter == "velocity") %>% 
  filter(value < 10) %>%
  ggplot() + 
  geom_histogram(aes(x = value, fill = stream), alpha = .5, binwidth = .3, position="identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### turbidity

All watersheds provide turbidity data

```{r}
environmental_format%>% 
  filter(parameter == "turbidity") %>% 
  ggplot() + 
  geom_boxplot(aes(x = value, y = stream, color = stream), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

Zoom in on 0 - 25 turbidity range:

```{r}
environmental_format%>% 
  filter(parameter == "turbidity") %>% 
  filter(value < 50) %>% # 1,647 values above 50 (274,049 rows total)
  ggplot() + 
  geom_histogram(aes(x = value, fill = stream), alpha = .5, binwidth = .3, position="identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### weather

```{r}
unique(environmental_format$weather)
environmental_format %>% 
  filter(parameter == "weather") %>% 
  ggplot() + 
  geom_bar(aes(x = text, fill = stream), position = "dodge") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

### habitat

```{r}
environmental_format %>% 
  filter(parameter == "habitat") %>% 
  ggplot() + 
  geom_bar(aes(x = text, fill = stream), position = "dodge") +
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))


```

## Save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(environmental_format,
           object_function = f,
           type = "csv",
           name = "rst/battle_clear_environmental_edi.csv",
           predefinedAcl = "bucketLevel")
```

# Release

```{r}
release_format <- release |> 
  select(-hatchery_origin) |> 
  rename(site_released = release_site,
         night_release = day_or_night_release,
         temperature_at_release = release_temp,
         turbidity_at_release = release_turbidity,
         flow_at_release = release_flow,
         run_released = fws_run) |> 
  mutate(stream = case_when(grepl("clear", site) ~ "clear creek",
                            grepl("battle", site) ~ "battle creek"),
         site = case_when(site == "upper battle creek" ~ "ubc",
                          site == "lower battle creek" ~ "lbc",
                          site == "lower clear creek" ~ "lcc", 
                          site == "upper clear creek" ~ "ucc"),
         night_release = ifelse(night_release == "night", T, F),
         origin_released = ifelse(is.na(origin_released), "not recorded", origin_released),
         run_released = ifelse(is.na(run_released), "not recorded", run_released))
         

```

## QA/QC {.tabset}

### release_id

Release id are unique identifiers for an efficiency trial on a specific tributary. There are 1809 unique release IDs

```{r}
release_format %>% 
  group_by(site) %>% 
  tally()
```
### release_date

```{r}
release_format %>% ggplot() + 
  geom_point(aes(x = date_released, y = stream, color = stream), alpha = .1) + 
  scale_color_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### release_time

```{r}
release_format %>% ggplot() + 
  geom_point(aes(x = time_released, y = stream, color = stream), alpha = .1) + 
  scale_color_manual(values = color_pal) +
  theme_minimal() + 
  scale_x_time() +
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### number_released

```{r}
release_format %>% ggplot() + 
  geom_histogram(aes(x = number_released, fill = stream), alpha = .5) + 
  scale_fill_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### median_fork_length_released

```{r}
release_format %>% 
  ggplot(aes(x = median_fork_length_released, fill = stream)) + 
  geom_histogram(breaks=seq(0, 200, by=2), alpha = .5) + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  scale_fill_manual(values = color_pal) +
  theme_minimal() +
  labs(title = "median fork length released distribution") + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
release_format %>% 
  mutate(year = as.factor(lubridate::year(date_released))) %>%
  ggplot(aes(x = median_fork_length_released, y = year, fill = stream)) + 
  geom_boxplot() + 
  scale_fill_manual(values = color_pal) +
  theme_minimal() +
  labs(title = "median fork length released by year") + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### night_release

Percent night release

```{r}
total_count <- release_format %>%
  mutate(year = year(date_released)) %>%
  group_by(site, year) %>%
  summarize(total = length(release_id))

release_format %>%
  filter(night_release == T) %>%
  mutate(year = year(date_released)) %>%
  group_by(site, year) %>%
  summarize(count = length(release_id)) %>%
  left_join(total_count) %>%
  mutate(percent = (count/total)*100) %>%
  ggplot(aes(x = year, y = percent)) +
  geom_col() +
  facet_wrap(~site)
```

### days_held_post_mark

Only battle and clear provide the days held post mark (pre release), most fish are held one day at most before release

```{r}
release_format %>% 
  ggplot() + 
  geom_histogram(aes(x = days_held_post_mark, fill = stream), alpha = .5, binwidth = 1) +
  scale_fill_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### flow_at_release

```{r}
release_format %>% 
  ggplot() + 
  geom_histogram(aes(x = flow_at_release, fill = stream), alpha = .5, binwidth = 25) +
  scale_fill_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### temperature_at_release

```{r}
release_format %>% 
  ggplot() + 
  geom_histogram(aes(x = temperature_at_release, fill = stream), alpha = .5, binwidth = 2.5) +
  scale_fill_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### turbidity_at_release

```{r}
release_format %>% 
  ggplot() + 
  geom_histogram(aes(x = turbidity_at_release, fill = stream), alpha = .5, binwidth = 2) +
  scale_fill_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
## Save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(release_format,
           object_function = f,
           type = "csv",
           name = "rst/battle_clear_release_edi.csv",
           predefinedAcl = "bucketLevel")
```

# Recapture

```{r}
# TODO this error persists and might be good to communciate with them: number_recaptured == 1180
recapture_format <- recapture |> 
  select(-c(release_site, fws_run, hatchery_origin)) |> 
  mutate(stream = case_when(grepl("clear", site) ~ "clear creek",
                            grepl("battle", site) ~ "battle creek"),
         site = case_when(site == "upper battle creek" ~ "ubc",
                          site == "lower battle creek" ~ "lbc",
                          site == "lower clear creek" ~ "lcc", 
                          site == "upper clear creek" ~ "ucc"),
         subsite = site,
         number_recaptured = ifelse(number_recaptured == 1180, 11, number_recaptured))

```

## QA/QC {.tabset}

### release_id

```{r}
recapture_format%>% 
  group_by(site) %>% 
  tally()
```

### recaptured_date

```{r}
recapture_format %>% ggplot() + 
  geom_point(aes(x = date_recaptured, y = stream, color = stream), alpha = .1) + 
  scale_color_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

### number_recaptured

```{r}
recapture_format %>% 
  group_by(release_id, stream) %>% 
  summarise(number_recaptured = sum(number_recaptured, na.rm = T)) %>% 
  ungroup() %>%
  ggplot() + 
  geom_histogram(aes(x = number_recaptured, fill = stream), alpha = .5, binwidth = 10, position = "identity") +
  scale_fill_manual(values = color_pal) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
recapture_format %>% 
  filter(number_recaptured > 250)
```

Manually update battle value of 1180 - We used the totals column in the raw data to update this outlier data point. The totals column reported 11 when the catch-by-day columns had an error of 1180. Fixed above

### median_fork_length_recaptured

Individual fork lengths for recaptured fish are not provided by Battle Creek or Clear Creek. Battle and Clear Creek only show median fork length per efficiency trial.

```{r}
recapture_format %>% 
  ggplot(aes(x = median_fork_length_recaptured, fill = stream)) + 
  geom_histogram(breaks=seq(0, 200, by=2),  alpha = .5, position = "identity") + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  scale_fill_manual(values = color_pal) +
  theme_minimal() +
  labs(title = "median fork length recaptured distribution") + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
recapture_format %>% 
  mutate(year = as.factor(lubridate::year(date_recaptured))) %>%
  ggplot(aes(x = median_fork_length_recaptured, y = year, fill = stream)) + 
  geom_boxplot() + 
  scale_fill_manual(values = color_pal) +
  theme_minimal() +
  labs(title = "median fork length recaptured by year") + 
  theme(legend.position = "bottom", 
        text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
## Save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(recapture_format,
           object_function = f,
           type = "csv",
           name = "rst/battle_clear_recapture_edi.csv",
           predefinedAcl = "bucketLevel")
```