---
title: ""
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
library(pheatmap)
library(RColorBrewer)
library(kableExtra)
library(ComplexHeatmap)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
tisdale_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "tisdale", "CAMP.mdb")
knights_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "knights_landing", "CAMP.mdb")
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
```

```{r, include = F}
project_description <- mdb.get(tisdale_camp, tables = "ProjectDescription") %>%
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(knights_camp, tables = "ProjectDescription")) %>% 
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ProjectDescription")) %>%  
  select( projectName, projectDescriptionID) 

project_id <- project_description %>% 
  distinct() %>% 
  filter(projectDescriptionID != 0) %>% 
  mutate(projectName = case_when(projectName == "Butte Creek Parrot-Phelan Diversion Dam\n-\nPhelan Diversion \nDam" ~ "Butte",
                                 projectName == "Lower Feather River RST" ~ "L Feather", 
                                 projectName == "Sacramento River Knights Landing RST" ~ "Knights",
                                 projectName == "Sacramento River Tisdale RST" ~ "Tisdale",
                                 projectName == "Feather River RST Program" ~ "U Feather")) %>% 
  filter(projectName != "Butte")
```

```{r, include = F}
make_percent_complete <- function(data) {
  data %>% 
  group_by(projectName) %>% 
  summarise_all(list(~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na") %>% 
  mutate(percent_nonNA = (1-prop_na) * 100)
    
}

make_heatmap_data <- function(percent_complete_data) {
  percent_complete_data %>% 
  select(-prop_na) %>% 
  pivot_wider(id_cols = "projectName", names_from = "field", values_from = "percent_nonNA")
}

make_heatmap <- function(df_num) {
  ComplexHeatmap::pheatmap(t(df_num), cluster_rows = F, cluster_cols = F,
                         display_numbers = T, 
                         number_format = "%.0f",
                         color = colorRampPalette(brewer.pal(n = 7, name =
  "Purples"))(100), 
                         border_color = "black", number_color = "black",
                         fontsize_number = 8, row_names_side = c("left"),
                         angle_col = c("0"), heatmap_legend_param = list(title ="Percent Complete"))
}

catch_raw <- mdb.get(tisdale_camp, tables = "CatchRaw") %>% 
  mutate(qcDoneTime = as.Date(qcDoneTime)) %>% 
  bind_rows(mdb.get(knights_camp, tables = "CatchRaw") %>% 
              mutate(qcDoneTime = as.Date(qcDoneTime))) %>% 
  bind_rows(mdb.get(feather_camp, tables = "CatchRaw") %>% 
              mutate(taxonID = as.numeric(taxonID),
                     qcDoneTime = as.Date(qcDoneTime))) %>% 
  full_join(project_id)
  

percent_complete_data <- catch_raw %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:36])
rownames(df_num) <- heatmap_data$projectName
```

### CatchRaw

```{r, echo = F}
make_heatmap(df_num)
```

#### CatchRaw - Useful fields

These fields contain useful information describing fish caught in trap. This information can be joined to other tables using the projectDescriptionID, catchRawID, trapVisitID, releaseID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "catchRawID",
                         "trapVisitID",
                         "taxonID",
                         "atCaptureRunID",
                         "atCaptureRunMethodID",
                         "finalRunID",
                         "finalRunMethodID",
                         "fishOriginID",
                         "lifeStageID",
                         "forkLength",
                         "totalLength",
                         "weight",
                         "n",
                         "randomID",
                         "actualCountID",
                         "releaseID",
                         "mortID",
                         "comments"),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for record in table (fish caught in trap)",
                              "unique identifier for trap visit, useful for joining with TrapVisit table",
                              "taxon of species, use lookup table for common name of species",
                              "run designation when fish captured in field",
                              "method used to designate run when captured in field",
                              "final run designation made later, not all programs fill this in the same way else percent complete would be the same for atCapture and final",
                              "method used to designate final run of fish",
                              "describes the origin or production of fish",
                              "describes lifestage of fish",
                              "fork length of fish",
                              "total length of fish, although not all use this field it is retained because in some cases fork length is not available and total length is",
                              "weight of fish, although not all use this field it is retained because it is an important metric to describe fish condition",
                              "count",
                              "describes if fish sampled randomly",
                              "describes if count is actual or estimate",
                              "unique identifier of release group, useful for joining with Release table or ReleaseFish table",
                              "describes if fish is dead or not",
                              "qualitative comments"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### CatchRaw - Are all fields needed?

All fields are not needed. Some fields are not being used. There are 7 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("auxiliaryRunID",
                            "auxiliaryRunMethodID",
                            "stdLength",
                            "subsampleMethodID",
                            "subsampleNumberator",
                            "subsampleDenominator",
                            "fishHealth",
                            "gillRating",
                            "dataSheetNumber",
                            "dataRecorder",
                            "dataRecorderAgencyID",
                            "creationTime",
                            "updateTime",
                            "qcDone",
                            "qcDoneTime",
                            "qcComments"),
                Rationale = c("not used",
                              "not used",
                              "not used",
                              "subSampleMethod is always 1 for all meaning that all fish are counted or estimated not using subsampling and this field is not needed",
                              "not needed because all fish counted or estimated not using subsampling",
                              "not needed because all fish counted or estimated not using subsampling",
                              "not used by majority",
                              "not used",
                              "not used",
                              "used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used by majority",
                              "not used",
                              "not used by majority"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```
