---
title: "CAMP MarkApplied"
author: "Ashley Vizek"
date: "9/16/2022"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

# How do monitoring programs use the CAMP MarkApplied table?

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
tisdale_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "tisdale", "CAMP.mdb")
knights_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "knights_landing", "CAMP.mdb")
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
```

```{r, include = F}
heatmap <- function(dat, table) {
ggplot(dat, aes(x = projectName, y = field, fill = percent_nonNA)) +
  geom_tile() +
  scale_fill_distiller(palette = "BuPu", direction = 1) +
  theme(axis.text.x = element_text(size = 8)) +
  labs(x = "",
       y = "CAMP field",
       title = table,
       fill = "Percent Complete") +
  theme(legend.position = "bottom")
}
```

```{r, include = F}
project_description <- mdb.get(tisdale_camp, tables = "ProjectDescription") %>%
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(knights_camp, tables = "ProjectDescription")) %>% 
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ProjectDescription")) %>%  
  select( projectName, projectDescriptionID) 

project_id <- project_description %>% 
  distinct() %>% 
  filter(projectDescriptionID != 0) %>% 
  mutate(projectName = case_when(projectName == "Butte Creek Parrot-Phelan Diversion Dam\n-\nPhelan Diversion \nDam" ~ "Butte",
                                 projectName == "Lower Feather River RST" ~ "Lower Feather", 
                                 projectName == "Sacramento River Knights Landing RST" ~ "Knights Landing",
                                 projectName == "Sacramento River Tisdale RST" ~ "Tisdale",
                                 projectName == "Feather River RST Program" ~ "Upper Feather")) %>% 
  filter(projectName != "Butte")
```

```{r}
mark_applied <- mdb.get(tisdale_camp, tables = "MarkApplied") %>% 
  bind_rows(mdb.get(knights_camp, tables = "MarkApplied")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "MarkApplied")) %>% 
  full_join(project_id)

mark_applied_summary <- mark_applied %>% 
  group_by(projectName) %>% 
  summarise_all(list(~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na") %>% 
  mutate(percent_nonNA = (1-prop_na) * 100)

heatmap(mark_applied_summary, "MarkApplied")
```

Proposed list of fields to keep and exclude. Select the field for more
detail.

| Keep                  | Exclude              |
|-----------------------|----------------------|
| projectDescriptionID  | dataRecorder         |
| releaseID             | dataRecorderAgencyID |
| appliedMarkTypeID     | creationTime         |
| appliedMarkColorID    | updateTime           |
| appliedMarkPositionID | qcDone               |
| markAppliedID         | qcDoneTime           |
|                       | comments             |
|                       | appliedMarkCode      |

## Keep

### projectDescriptionID

The projectDescriptionID is a unique identifier for your monitoring program.

#### Use

When combining data with other monitoring programs and making joins with other tables.
Joins should include projectDescriptionID so that the record identifier is unique to the
monitoring program.

### releaseID

The releaseID is a unique identifier for an individual efficiency trial and represents
one record in the CAMP release table.

#### Use 

To join the MarkApplied table to the Release table. When combining data from
multiple monitoring programs, releaseID and projectDescriptionID should be used
to make the join. 

### markAppliedID

### appliedMarkTypeID

### appliedMarkColorID

### appliedMarkPositionID

### Exclude

### dataRecorder, dataRecorderAgencyID, creationTime, updateTime, qcDone, qcDoneTime

### appliedMarkCode

### comments
