---
title: "CAMP comparisons"
output: html_document
date: "2022-09-14"
---

This document compares the proportion of NA within each CAMP data table across
monitoring programs.

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
tisdale_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "tisdale", "CAMP.mdb")
knights_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "knights_landing", "CAMP.mdb")
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
```

```{r, include = F}
heatmap <- function(dat, table) {
ggplot(dat, aes(x = projectName, y = field, fill = prop_na)) +
  geom_tile() +
  scale_fill_viridis_c(option = "D") +
  theme(axis.text.x = element_text(size = 8)) +
  labs(x = "",
       y = "CAMP field",
       title = table,
       fill = "Proportion NA")
}
```

# CAMP data tables {.tabset}

## ProjectDescription 

This table contains project level metadata and will be captured in metadata documents. 

```{r}
project_description <- mdb.get(tisdale_camp, tables = "ProjectDescription") %>% 
  bind_rows(mdb.get(knights_camp, tables = "ProjectDescription")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ProjectDescription")) %>%  glimpse

project_id <- select(project_description, projectName, projectDescriptionID) %>% 
  distinct() %>% 
  filter(projectDescriptionID != 0) %>% 
  mutate(projectName = case_when(projectName == "Butte Creek Parrot-Phelan Diversion Dam\n-\nPhelan Diversion \nDam" ~ "Butte",
                                 projectName == "Lower Feather River RST" ~ "Lower Feather", 
                                 projectName == "Sacramento River Knights Landing RST" ~ "Knights Landing",
                                 projectName == "Sacramento River Tisdale RST" ~ "Tisdale",
                                 projectName == "Feather River RST Program" ~ "Upper Feather"))
```

## Site

Site location information will be captured in the metadata. All tables will include
site name.

```{r}
site <- mdb.get(tisdale_camp, tables = "Site") %>% 
  bind_rows(mdb.get(knights_camp, tables = "Site")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "Site")) %>%  glimpse()
```

## SubSite

Subsite information will be captured in metadata. All tables will include subsite
name. 

```{r}
subsite <- mdb.get(tisdale_camp, tables = "SubSite") %>% 
  bind_rows(mdb.get(knights_camp, tables = "SubSite")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "SubSite")) %>% 
  full_join(project_id)

subsite_summary <- subsite %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(subsite_summary, "SubSite")
```

## TrapVisit

### variables to exclude
- crew (internal only)
- sampleGearID (metadata)
- timeSampleStarted
- timeSampleEnded
- batchDate
- counterOld
- counterOldTime
- counterAtEndTime
- counterNew
- coneDepthAtStart
- coneDepthAtEnd
- includeCatchComments
- coneDebrisIntake1
- coneDebrisIntake2
- debrisType
- dataCollectedID
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

### variables to include
- projectDescriptionID
- trapVisitID (used to join with CatchRaw)
- trapPositionID (join with site/subsite)
- visitTime
- visitTime2
- visitTypeID
- fishProcessedID (luFishProcessed)
- inThalwegID (luNoYes)
- trapFunctioningID (luTrapFunctioning)
- counterAtStart
- counterAtEnd
- rpmRevolutionsAtStart
- rpmSecondsAtStart
- rpmRevolutionsAtEnd
- rpmSecondsAtEnd
- halfConeID (luNoYes)
- includeCatchID (luNoYes)
- debrisVolumeCatID (luVolumeCat)
- debrisVolume
- debrisVolumeUnits 
- comments

```{r}
trap_visit <- mdb.get(tisdale_camp, tables = "TrapVisit") %>% 
  bind_rows(mdb.get(knights_camp, tables = "TrapVisit")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "TrapVisit")) %>% 
  full_join(project_id)

trap_visit_summary <- trap_visit %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(trap_visit_summary, "TrapVisit")
```

## CatchRaw

### variables to exclude
- auxiliaryRunID
- auxiliaryRunMethodID
- stdLength 
- subsampleMethodID
- subsampleNumerator
- subsampleDenominator
- fishHealth
- gillRating
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

### variables to include
- projectDescriptionID
- catchRawID
- trapVisitID (used to join with TrapVisit table)
- taxonID (luTaxon)
- atCaptureRunID (luRun)
- atCaptureRunMethodID (luRunMethod)
- finalRunID (luRun)
- finalRunMethodID (luRunMethod)
- fishOriginID (luFishOrigin)
- lifeStageID (luLifeStage)
- forkLength
- totalLength
- weight
- n
- randomID (luNoYes)
- actualCountID (luNoYes)
- releaseID (used to join with Release table)
- mortID (luNoYes)
- comments

```{r}
catch_raw <- mdb.get(tisdale_camp, tables = "CatchRaw") %>% 
  bind_rows(mdb.get(knights_camp, tables = "CatchRaw")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "CatchRaw") %>% 
              mutate(taxonID = as.numeric(taxonID))) %>% 
  full_join(project_id)

catch_raw_summary <- catch_raw %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(catch_raw_summary, "CatchRaw")
```

## MarkExisting

### variables to exclude
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

### variables to include
- projectDescriptionID
- catchRawID (used to join with CatchRaw table)
- markTypeID (luMarkType)
- markColorID (luColor)
- markPositionID (luBodyPart)
- markExistingID
- markCode

```{r}
mark_existing <- mdb.get(tisdale_camp, tables = "MarkExisting") %>% 
  bind_rows(mdb.get(knights_camp, tables = "MarkExisting")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "MarkExisting") %>% 
              mutate(markCode = as.character(markCode))) %>% 
  full_join(project_id)

mark_existing_summary <- mark_existing %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(mark_existing_summary, "MarkExisting")
```

## Specimen

Do not plan to include this table. For internal use only.

```{r}
specimen <- mdb.get(tisdale_camp, tables = "Specimen") %>% 
  bind_rows(mdb.get(knights_camp, tables = "Specimen")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "Specimen") %>% 
            mutate(physicalSpecimenCode = as.character(physicalSpecimenCode),
                   comments = as.character(comments),
                   dataRecorder = as.character(dataRecorder),
                   creationTime = as.Date(creationTime),
                   updateTime = as.Date(updateTime))) %>% 
  full_join(project_id)

specimen_summary <- specimen %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(specimen_summary, "Specimen")
```

## PostHandleMort

No data.

```{r}
post_handle_mort <- mdb.get(tisdale_camp, "PostHandleMort") %>% 
  bind_rows(mdb.get(knights_camp, "PostHandleMort")) %>% 
  bind_rows(mdb.get(feather_camp, "PostHandleMort")) %>% 
  full_join(project_id)

post_handle_mort_summary <- post_handle_mort %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(post_handle_mort_summary, "PostHandleMort")
```

## PostHandleMortMark

No data.

```{r}
post_handle_mort_mark <- mdb.get(tisdale_camp, "PostHandleMortMark") %>% 
  bind_rows(mdb.get(knights_camp, "PostHandleMortMark")) %>% 
  bind_rows(mdb.get(feather_camp, "PostHandleMortMark")) %>% 
  full_join(project_id)

post_handle_mort_mark_summary <- post_handle_mort_mark %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(post_handle_mort_mark_summary, "PostHandleMortMark")
```

## EnvDataRaw

### variables to exclude
- measureTime does not exist
- subSiteID does not exist
- waterVelReadingStart
- waterVelReadingEnd
- waterVelTime
- waterVelTimeUnitID
- waterVelConvertFactor
- airTemp
- airTempUnitID
- airTempSampleGearID
- cloudCover
- dataCollectedID
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

### variables to include
- projectDescriptionID
- trapVisitID (used to join with TrapVisit table)
- envDataRawID
- discharge
- dischargeUnitID (luUnit)
- dischargeSampleGearID (luSampleGear)
- waterDepth
- waterDepthUnitID (luUnit)
- waterDepthSampleGearID (luSampleGear)
- waterVel
- waterVelUnitID (luUnit)
- waterVelSampleGearID (luSampleGear)
- waterTemp
- waterTempUnitID (luUnit)
- waterTempSampleGearID (luSampleGear)
- lightPenetration
- lightPenetrationUnitID (luUnit)
- lightPenetrationSampleGearID (luSampleGear)
- turbidity
- turbidityUnitID (luUnit)
- turbiditySampleGearID (luSampleGear)
- dissolvedOxygen
- dissolvedOxygenUnitID (luUnit)
- dissolvedOxygenSampleGearID (luSampleGear)
- conductivity
- conductivityUnitID (luUnit)
- conductivitySampleGearID (luSampleGear)
- barometer
- barometerUnitID (luUnit)
- barometerSampleGearID (luSampleGear)
- cloudCover
- weather
- comments

```{r}
env_data_raw <- mdb.get(tisdale_camp, tables = "EnvDataRaw") %>% 
  bind_rows(mdb.get(knights_camp, tables = "EnvDataRaw")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "EnvDataRaw")) %>% 
  full_join(project_id)

env_data_raw_summary <- env_data_raw %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(env_data_raw_summary, "EnvDataRaw")
```

## EnvDataRawXTargetSite

We do not need this table because join can be made to TrapVisit table.

```{r}
env_data_raw_target <- mdb.get(tisdale_camp, tables = "EnvDataRawXTargetSite") %>% 
  bind_rows(mdb.get(knights_camp, tables = "EnvDataRawXTargetSite")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "EnvDataRawXTargetSite"))  %>% 
  full_join(project_id)

env_data_raw_target_summary <- env_data_raw_target %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(env_data_raw_target_summary, "EnvDataRawXTargetSite")
```
```{r, include = F}
# join envdataraw with trapvisit

e1 <- env_data_raw %>% 
  left_join(env_data_raw_target)
# join envdataraw with envdatarawtargetsite

e2 <- env_data_raw %>% 
  left_join(trap_visit)

```

## EnvDataDaily

No data.

```{r}
env_data_daily <- mdb.get(tisdale_camp, tables = "EnvDataDaily") %>% 
  bind_rows(mdb.get(knights_camp, tables = "EnvDataDaily")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "EnvDataDaily")) %>% 
  full_join(project_id)

env_data_daily_summary <- env_data_daily %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(env_data_daily_summary, "EnvDataDaily")
```

## EnvDataDailyXTargetSite

No data.

```{r}
env_data_daily_target <- mdb.get(tisdale_camp, tables = "EnvDataDailyXTargetSite") %>% 
  bind_rows(mdb.get(knights_camp, tables = "EnvDataDailyXTargetSite")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "EnvDataDailyXTargetSite")) %>% 
  full_join(project_id)

env_data_daily_target_summary <- env_data_daily_target %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(env_data_daily_target_summary, "EnvDataDailyXTargetSite")
```

## Release

### variables to exclude
- checkedTime
- nMarkExamined
- nMarkNotRetained
- propMarked
- crew
- dataCollectedID
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

### variables to include
- projectDescriptionID
- releaseID (used to join with CatchRaw table)
- releasePurposeID (luReleasePurpose)
- markedTaxonID (luTaxon)
- markedRunID (luRun)
- markedLifeStageID (luLifeStage)
- markedFishOriginID (luFishOrigin)
- sourceOfFishSiteID (join with site)
- releaseSiteID (join with site)
- releaseSubSiteID (join with site)
- nMortWhileHandling
- nMortAtCheck
- nReleased
- releaseTime
- releaseLightConditionID (luLightCondition)
- testDays
- includeTestID (luNoYes)
- includeTestComments
- comments

```{r}
release <- mdb.get(tisdale_camp, tables = "Release") %>% 
  bind_rows(mdb.get(knights_camp, tables = "Release")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "Release")) %>% 
  full_join(project_id)

release_summary <- release %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(release_summary, "Release")
```

## ReleaseXTargetSite

```{r}
release_target <- mdb.get(tisdale_camp, tables = "ReleaseXTargetSite") %>% 
  bind_rows(mdb.get(knights_camp, tables = "ReleaseXTargetSite")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ReleaseXTargetSite")) %>% 
  full_join(project_id)

release_target_summary <- release_target %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(release_target_summary, "ReleaseXTargetSite")
```

## MarkApplied

### variables to exclude
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime

### variables to include
- projectDescriptionID
- releaseID (used to join with Release table)
- appliedMarkTypeID (luMarkType)
- appliedMarkColorID (luColor)
- appliedMarkPositionID (luBodyPart)
- markAppliedID
- appliedMarkCode
- comments

```{r}
mark_applied <- mdb.get(tisdale_camp, tables = "MarkApplied") %>% 
  bind_rows(mdb.get(knights_camp, tables = "MarkApplied")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "MarkApplied")) %>% 
  full_join(project_id)

mark_applied_summary <- mark_applied %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(mark_applied_summary, "MarkApplied")
```

## ReleaseFish

### variables to exclude
- dataSheetNumber
- dataRecorder
- dataRecorderAgencyID
- creationTime
- updateTime
- qcDone
- qcDoneTime
- qcComments

### variables to include
- projectDescriptionID
- releaseFishID
- releaseID (used to join with Release table)
- nMarked
- forkLength
- weight
- timeMarked
- comments

```{r}
release_fish <- mdb.get(tisdale_camp, tables = "ReleaseFish") %>% 
  bind_rows(mdb.get(knights_camp, tables = "ReleaseFish")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ReleaseFish")) %>% 
  full_join(project_id)

release_fish_summary <- release_fish %>% 
  group_by(projectName) %>% 
  summarise_all(list(na = ~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na")

heatmap(release_fish_summary, "ReleaseFish")
```

# CAMP Lookup Tables

Using Tisdale lookup tables for an example.

```{r}
visit_type_lu <- mdb.get(tisdale_camp, "luVisitType") %>% 
  select(-activeID) %>%glimpse

run_lu <- mdb.get(tisdale_camp, "luRun") %>% 
  select(-activeID) %>% glimpse

run_method_lu <- mdb.get(tisdale_camp, "luRunMethod") %>% 
  select(-activeID) %>% glimpse

lifestage_lu <- mdb.get(tisdale_camp, "luLifeStage") %>% 
  select(-activeID) %>% glimpse

taxon_lu <-  mdb.get(tisdale_camp, "luTaxon") %>% 
  mutate(taxonID = as.numeric(taxonID)) %>% 
  select(taxonID, commonName) %>% glimpse

origin_lu <-  mdb.get(tisdale_camp, "luFishOrigin") %>% 
  select(-activeID) %>% glimpse

debris_volume_lu <- mdb.get(tisdale_camp,  "luDebrisVolumeCat") %>% 
  select(-activeID) %>% 
  glimpse

trap_function_lu <- mdb.get(tisdale_camp, "luTrapFunctioning") %>%
  select(-activeID) %>% 
  glimpse

processed_lu <- mdb.get(tisdale_camp, "luFishProcessed") %>%
  select(-activeID) %>% 
  glimpse

subsite_lu <- mdb.get(tisdale_camp, "SubSite") %>% 
  select(subSiteName, subSiteID, siteID) %>% 
  filter(subSiteName != "N/A")

site_lu <- mdb.get(tisdale_camp, "Site") %>% 
  select(siteName, siteID)

sample_gear_lu <- mdb.get(tisdale_camp, "luSampleGear") %>% 
  select(-activeID) %>% glimpse()

no_yes_lu <- mdb.get(tisdale_camp, "luNoYes")

# lookups for existing mark
color_lu <- mdb.get(tisdale_camp, "luColor") %>%
  select(-activeID) %>% glimpse()

mark_type_lu <- mdb.get(tisdale_camp, "luMarkType") %>%
  select(-activeID) %>% glimpse()

body_part_lu <- mdb.get(tisdale_camp, "luBodyPart") %>%
  select(-activeID) %>% glimpse()

# lookups for releases
release_purpose_lu <- mdb.get(tisdale_camp, "luReleasePurpose") %>% 
  select(-activeID) %>% glimpse()

light_condition_lu <- mdb.get(tisdale_camp, "luLightCondition") %>% 
  select(-activeID) %>% glimpse()

unit_lu <- mdb.get(tisdale_camp, "luUnit") %>% 
  select(-activeID) %>% glimpse()

```
# CAMP Views for Data Sharing

## Site and SubSite

Not shared but linked with other tables.

```{r}
site_subsite <- full_join(site %>% select(siteName, siteID), 
                          subsite %>%  select(subSiteName, subSiteID, siteID))
```

## Trap

```{r}
trap_table <- trap_visit %>% 
  select(projectDescriptionID, trapVisitID, trapPositionID, visitTime, visitTime2, visitTypeID, fishProcessedID, inThalwegID, trapFunctioningID, counterAtStart, counterAtEnd, rpmRevolutionsAtStart, rpmSecondsAtStart, rpmRevolutionsAtEnd, rpmSecondsAtEnd, halfConeID, includeCatchID, debrisVolumeCatID, debrisVolume, debrisVolumeUnits, comments) %>% 
  left_join(site_subsite, by = c("trapPositionID" = "subSiteID")) %>% 
  select(-trapPositionID, -siteID) %>% 
  left_join(visit_type_lu, by = c("visitTypeID" = "visitTypeID")) %>% 
  select(-visitTypeID) %>% 
  left_join(processed_lu, by = c("fishProcessedID" = "fishProcessedID")) %>% 
  select(-fishProcessedID) %>% 
  left_join(no_yes_lu, by = c("inThalwegID" = "noYesID")) %>% 
  rename(inThalweg = noYes) %>% 
  select(-inThalwegID) %>% 
  left_join(trap_function_lu, by = c("trapFunctioningID" = "trapFunctioningID")) %>% 
  select(-trapFunctioningID) %>% 
  left_join(no_yes_lu, by = c("halfConeID" = "noYesID")) %>% 
  rename(halfCone = noYes) %>% 
  select(-halfConeID) %>% 
  left_join(no_yes_lu, by = c("includeCatchID" = "noYesID")) %>% 
  rename(includeCatch = noYes) %>% 
  select(-includeCatchID) %>% 
  left_join(debris_volume_lu, by = c("debrisVolumeCatID" = "debrisVolumeCatID")) %>% 
  select(-debrisVolumeCatID)
```

## Catch

### TrapVisit
- projectDescriptionID (used to join with CatchRaw)
- trapVisitID (used to join with CatchRaw)
- visitTime
- visitTime2
- visitType (luVisitType)
- siteName (join with site/subsite)
- subSiteName (join with site/subsite)

### MarkExisting
- projectDescriptionID (used to join with CatchRaw table)
- catchRawID (used to join with CatchRaw table)
- markTypeID (luMarkType)
- markColorID (luColor)
- markPositionID (luBodyPart)
- markExistingID
- markCode

In some cases one catchRawID can have multiple markExisitingIDs. This creates duplicates in the catchRaw table when joining. To avoid duplicates the markExisting table needs to spread to include columns for each mark.

```{r}
mark_existing_w_joins <- mark_existing %>% 
  left_join(mark_type_lu, by = c("markTypeID" = "markTypeID")) %>% 
  select(-markTypeID) %>% 
  left_join(color_lu, by = c("markColorID" = "colorID")) %>% 
  rename(markColor = color) %>% 
  select(-markColorID) %>% 
  left_join(body_part_lu, by = c("markPositionID" = "bodyPartID")) %>% 
  rename(markPosition = bodyPart) %>% 
  select(-markPositionID) %>% 
  select(projectDescriptionID, catchRawID, markCode, markType, markColor, markPosition) %>% 
  pivot_longer(cols = starts_with("mark"),
               names_to = "mark",
               values_to = "value") %>% 
  group_by(catchRawID, projectDescriptionID, mark) %>% 
  mutate(number = 1:n(),
         mark = paste0(mark,"_",number)) %>% 
  select(-number) %>% 
  pivot_wider(id_cols = c("catchRawID", "projectDescriptionID"), names_from = "mark", values_from = "value")

catch_table <- catch_raw %>% 
  select(projectDescriptionID, catchRawID, trapVisitID, taxonID, atCaptureRunID, atCaptureRunMethodID, finalRunID, finalRunMethodID, fishOriginID, lifeStageID, forkLength, totalLength, weight, n, randomID, actualCountID, releaseID, mortID, comments) %>% 
  left_join(filter(trap_table, fishProcessed %in% c("Processed fish", "Processed fish")) %>% 
              select(projectDescriptionID, trapVisitID, visitTime, visitTime2, visitType, siteName, subSiteName), by = c("trapVisitID" = "trapVisitID", "projectDescriptionID" = "projectDescriptionID")) %>% 
  left_join(mark_existing_w_joins, by = c("catchRawID" = "catchRawID", "projectDescriptionID" = "projectDescriptionID"))
```

## Release

### Release Summary
- projectDescriptionID
- releaseID (used to join with CatchRaw table)
- releasePurposeID (luReleasePurpose)
- markedTaxonID (luTaxon)
- markedRunID (luRun)
- markedLifeStageID (luLifeStage)
- markedFishOriginID (luFishOrigin)
- sourceOfFishSiteID (join with site)
- releaseSiteID (join with site)
- releaseSubSiteID (join with site)
- nMortWhileHandling
- nMortAtCheck
- nReleased
- releaseTime
- releaseLightConditionID (luLightCondition)
- testDays
- includeTestID (luNoYes)
- includeTestComments
- comments

### MarkApplied
- projectDescriptionID
- releaseID (used to join with Release table)
- appliedMarkTypeID (luMarkType)
- appliedMarkColorID (luColor)
- appliedMarkPositionID (luBodyPart)
- markAppliedID
- appliedMarkCode
- comments

```{r}
release_summary_table <- release %>% 
  select(projectDescriptionID, releaseID, releasePurposeID, markedTaxonID, markedRunID, markedLifeStageID, markedFishOriginID, sourceOfFishSiteID, releaseSiteID, releaseSubSiteID, nMortWhileHandling, nMortAtCheck, nReleased, releaseTime, releaseLightConditionID,
         testDays, includeTestID, includeTestComments, comments) %>% 
  left_join(release_purpose_lu, by = c("releasePurposeID" = "releasePursposeID")) %>% 
  select(-releasePurposeID) %>% 
  left_join(taxon_lu, by = c("markedTaxonID" = "taxonID")) %>% 
  rename(markedTaxon = commonName) %>% 
  select(-markedTaxonID) %>% 
  left_join(run_lu, by = c("markedRunID" = "runID")) %>% 
  rename(markedRun = run) %>% 
  select(-markedRunID) %>% 
  left_join(lifestage_lu, by = c("markedLifeStageID" = "lifeStageID")) %>% 
  rename(markedLifeStage = lifeStage) %>% 
  select(-markedLifeStageID) %>% 
  left_join(origin_lu, by = c("markedFishOriginID" = "fishOriginID")) %>% 
  rename(markedFishOrigin = fishOrigin) %>% 
  select(-markedFishOrigin) %>% 
  left_join(site_subsite %>% 
              select(siteID, siteName), by = c("sourceOfFishSiteID" = "siteID")) %>% 
  rename(sourceOfFishSite = siteName) %>% 
  select(-sourceOfFishSiteID) %>% 
  left_join(site_subsite %>% 
              select(siteID, siteName), by = c("releaseSiteID" = "siteID")) %>% 
  rename(releaseSite = siteName) %>% 
  select(-releaseSiteID) %>% 
  left_join(site_subsite %>% 
              select(subSiteID, subSiteName), by = c("releaseSubSiteID" = "subSiteID")) %>% 
  rename(releaseSubSite = subSiteName) %>% 
  select(-releaseSubSiteID) %>% 
  left_join(light_condition_lu, by = c("releaseLightConditionID" = "lightConditionID")) %>% 
  rename(releaseLightCondition = lightCondition) %>% 
  select(-releaseLightConditionID) %>%  
  left_join(no_yes_lu, by = c("includeTestID" = "noYesID")) %>% 
  rename(includeTest = noYes) %>% 
  select(-includeTestID)
```

## Released Fish
- projectDescriptionID
- releaseFishID
- releaseID (used to join with Release table)
- nMarked
- forkLength
- weight
- timeMarked
- comments

Option would be to add rows of recaptured fish to this table

```{r}
release_fish_table <- release_fish %>% 
  select(projectDescriptionID, releaseFishID, releaseID, nMarked, forkLength, weight, timeMarked, comments)
```

## Environmental

### TrapVisit
- projectDescriptionID
- trapVisitID
- visitTime
- visitTime2
- visitType
- siteName
- subSiteName

### EnvDataRaw
- projectDescriptionID
- trapVisitID (used to join with TrapVisit table)
- envDataRawID
- discharge
- dischargeUnitID (luUnit)
- dischargeSampleGearID (luSampleGear)
- waterDepth
- waterDepthUnitID (luUnit)
- waterDepthSampleGearID (luSampleGear)
- waterVel
- waterVelUnitID (luUnit)
- waterVelSampleGearID (luSampleGear)
- waterTemp
- waterTempUnitID (luUnit)
- waterTempSampleGearID (luSampleGear)
- lightPenetration
- lightPenetrationUnitID (luUnit)
- lightPenetrationSampleGearID (luSampleGear)
- turbidity
- turbidityUnitID (luUnit)
- turbiditySampleGearID (luSampleGear)
- dissolvedOxygen
- dissolvedOxygenUnitID (luUnit)
- dissolvedOxygenSampleGearID (luSampleGear)
- conductivity
- conductivityUnitID (luUnit)
- conductivitySampleGearID (luSampleGear)
- barometer
- barometerUnitID (luUnit)
- barometerSampleGearID (luSampleGear)
- comments

```{r}
environmental_table <- env_data_raw %>% 
  select(projectDescriptionID, trapVisitID, envDataRawID, discharge, dischargeUnitID, dischargeSampleGearID, waterDepth, waterDepthUnitID, waterDepthSampleGearID, waterVel, waterVelUnitID, waterVelSampleGearID, waterTemp, waterTempUnitID, waterTempSampleGearID, lightPenetration, lightPenetrationUnitID, lightPenetrationSampleGearID, turbidity, turbidityUnitID, turbiditySampleGearID, dissolvedOxygen, dissolvedOxygenUnitID, dissolvedOxygenSampleGearID, conductivity, conductivityUnitID, conductivitySampleGearID, barometer, barometerUnitID, barometerSampleGearID, comments) %>% 
  pivot_longer(c(discharge, waterDepth, waterVel, waterTemp, lightPenetration, turbidity, dissolvedOxygen, conductivity, barometer), names_to = "parameter", values_to = "value") %>% 
  mutate(unitID = case_when(parameter == "discharge" ~ dischargeUnitID,
                          parameter == "waterDepth" ~ waterDepthUnitID,
                          parameter == "waterVel" ~ waterVelUnitID,
                          parameter == "waterTemp" ~ waterTempUnitID,
                          parameter == "lightPenetration" ~ lightPenetrationUnitID,
                          parameter == "turbidity" ~ turbidityUnitID,
                          parameter == "dissolvedOxygen" ~ dissolvedOxygenUnitID,
                          parameter == "conductivity" ~ conductivityUnitID,
                          parameter == "barometer" ~ barometerUnitID),
         sampleGearID = case_when(parameter == "discharge" ~ dischargeSampleGearID,
                          parameter == "waterDepth" ~ waterDepthSampleGearID,
                          parameter == "waterVel" ~ waterVelSampleGearID,
                          parameter == "waterTemp" ~ waterTempSampleGearID,
                          parameter == "lightPenetration" ~ lightPenetrationSampleGearID,
                          parameter == "turbidity" ~ turbiditySampleGearID,
                          parameter == "dissolvedOxygen" ~ dissolvedOxygenSampleGearID,
                          parameter == "conductivity" ~ conductivitySampleGearID,
                          parameter == "barometer" ~ barometerSampleGearID)) %>% 
  select(-c(dischargeUnitID, dischargeSampleGearID, waterDepthUnitID, waterDepthSampleGearID, waterVelUnitID, waterVelSampleGearID, waterTempUnitID, waterTempSampleGearID, lightPenetrationUnitID, lightPenetrationSampleGearID, turbidityUnitID, turbiditySampleGearID, dissolvedOxygenUnitID, dissolvedOxygenSampleGearID, conductivityUnitID, conductivitySampleGearID, barometerUnitID, barometerSampleGearID)) %>% 
  left_join(unit_lu, by = c("unitID" = "unitID")) %>% 
  left_join(sample_gear_lu, by = c("sampleGearID" = "sampleGearID")) %>% 
  select(-unitID, -sampleGearID) %>% 
  left_join(trap_table %>% 
              select(projectDescriptionID, trapVisitID, visitTime, visitTime2, visitType, siteName, subSiteName), by = c("trapVisitID" = "trapVisitID", "projectDescriptionID" = "projectDescriptionID"))
  
```
