---
title: ""
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
library(pheatmap)
library(RColorBrewer)
library(kableExtra)
library(ComplexHeatmap)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
tisdale_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "tisdale", "CAMP.mdb")
knights_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "knights_landing", "CAMP.mdb")
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
```

```{r, include = F}
project_description <- mdb.get(tisdale_camp, tables = "ProjectDescription") %>%
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(knights_camp, tables = "ProjectDescription")) %>% 
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ProjectDescription")) %>%  
  select( projectName, projectDescriptionID) 

project_id <- project_description %>% 
  distinct() %>% 
  filter(projectDescriptionID != 0) %>% 
  mutate(projectName = case_when(projectName == "Butte Creek Parrot-Phelan Diversion Dam\n-\nPhelan Diversion \nDam" ~ "Butte",
                                 projectName == "Lower Feather River RST" ~ "L Feather", 
                                 projectName == "Sacramento River Knights Landing RST" ~ "Knights",
                                 projectName == "Sacramento River Tisdale RST" ~ "Tisdale",
                                 projectName == "Feather River RST Program" ~ "U Feather")) %>% 
  filter(projectName != "Butte")
```

```{r, include = F}
make_percent_complete <- function(data) {
  data %>% 
  group_by(projectName) %>% 
  summarise_all(list(~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na") %>% 
  mutate(percent_nonNA = (1-prop_na) * 100)
    
}

make_heatmap_data <- function(percent_complete_data) {
  percent_complete_data %>% 
  select(-prop_na) %>% 
  pivot_wider(id_cols = "projectName", names_from = "field", values_from = "percent_nonNA")
}

make_heatmap <- function(df_num) {
  ComplexHeatmap::pheatmap(t(df_num), cluster_rows = F, cluster_cols = F,
                         display_numbers = T, 
                         number_format = "%.0f",
                         color = colorRampPalette(brewer.pal(n = 7, name =
  "Purples"))(100), 
                         border_color = "black", number_color = "black",
                         fontsize_number = 8, row_names_side = c("left"),
                         angle_col = c("0"), heatmap_legend_param = list(title ="Percent Complete"))
}

mark_existing <- mdb.get(tisdale_camp, tables = "MarkExisting") %>% 
  bind_rows(mdb.get(knights_camp, tables = "MarkExisting")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "MarkExisting") %>% 
              mutate(markCode = as.character(markCode))) %>% 
  full_join(project_id) 
  

percent_complete_data <- mark_existing %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:15])
rownames(df_num) <- heatmap_data$projectName
```

### MarkExisting

```{r, echo = F}
make_heatmap(df_num)
```

#### MarkExisting - Useful fields

These fields contain useful information to identify the release group that corresponds with an existing mark. This information can be joined to other tables using the projectDescriptionID and catchRawID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "catchRawID",
                         "markTypeID",
                         "markColorID",
                         "markPositionID",
                         "markCode",
                         "markExistingID"),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for fish caught in trap, useful for joining with CatchRaw table",
                              "type of existing mark, useful for identification of release group",
                              "color of existing mark, useful for identification of release group",
                              "position of existing mark, useful for identification of release group",
                              "code of existing mark (e.g., PIT, CWT), useful to connect with other studies and releases",
                              "unique identifier for record in table"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### MarkExisting - Are all fields needed?

All fields are not needed. There are 7 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("dataRecorder",
                         "dataRecorderAgencyID",
                         "creationTime",
                         "updateTime",
                         "qcDone",
                         "qcDoneTime",
                         "qcComments"),
                Rationale = c("used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used",
                              "not used",
                              "not used by majority"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```
