---
title: ""
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
library(pheatmap)
library(RColorBrewer)
library(kableExtra)
library(ComplexHeatmap)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(fig.width=7, fig.height=6) 
```

```{r, include = F}
tisdale_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "tisdale", "CAMP.mdb")
knights_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "knights_landing", "CAMP.mdb")
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
```

```{r, include = F}
project_description <- mdb.get(tisdale_camp, tables = "ProjectDescription") %>%
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(knights_camp, tables = "ProjectDescription")) %>% 
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ProjectDescription")) %>%  
  select( projectName, projectDescriptionID) 

project_id <- project_description %>% 
  distinct() %>% 
  filter(projectDescriptionID != 0) %>% 
  mutate(projectName = case_when(projectName == "Butte Creek Parrot-Phelan Diversion Dam\n-\nPhelan Diversion \nDam" ~ "Butte",
                                 projectName == "Lower Feather River RST" ~ "L Feather", 
                                 projectName == "Sacramento River Knights Landing RST" ~ "Knights",
                                 projectName == "Sacramento River Tisdale RST" ~ "Tisdale",
                                 projectName == "Feather River RST Program" ~ "U Feather")) %>% 
  filter(projectName != "Butte")
```

```{r, include = F}
make_percent_complete <- function(data) {
  data %>% 
  group_by(projectName) %>% 
  summarise_all(list(~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na") %>% 
  mutate(percent_nonNA = (1-prop_na) * 100)
    
}

make_heatmap_data <- function(percent_complete_data) {
  percent_complete_data %>% 
  select(-prop_na) %>% 
  pivot_wider(id_cols = "projectName", names_from = "field", values_from = "percent_nonNA")
}

make_heatmap <- function(df_num) {
  ComplexHeatmap::pheatmap(t(df_num), cluster_rows = F, cluster_cols = F,
                         display_numbers = T, 
                         number_format = "%.0f",
                         color = colorRampPalette(brewer.pal(n = 7, name =
  "Purples"))(100), 
                         border_color = "black", number_color = "black",
                         fontsize_number = 8, row_names_side = c("left"),
                         angle_col = c("0"), heatmap_legend_param = list(title ="Percent Complete"))
}

trap_visit <- mdb.get(tisdale_camp, tables = "TrapVisit") %>% 
  bind_rows(mdb.get(knights_camp, tables = "TrapVisit")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "TrapVisit")) %>% 
  full_join(project_id)
  

percent_complete_data <- trap_visit %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:46])
rownames(df_num) <- heatmap_data$projectName
```

### TrapVisit

```{r, echo = F}
make_heatmap(df_num)
```

#### TrapVisit - Useful fields

These fields contain useful information describing trap visit. This information can be joined to other tables using the projectDescriptionID, trapVisitID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "trapVisitID",
                         "trapPositionID",
                         "visitTime",
                         "visitTime2",
                         "visitTypeID",
                         "fishProcessedID",
                         "inThalwegID",
                         "trapFunctioningID",
                         "counterAtStart",
                         "counterAtEnd",
                         "rpmRevolutionsAtStart",
                         "rpmSecondsAtStart",
                         "rpmRevolutionsAtEnd",
                         "rpmSecondsAtEnd",
                         "halfConeID",
                         "includeCatchID",
                         "debrisVolumeCatID",
                         "debrisVolume",
                         "debrisVolumeUnits",
                         "comments"
                     ),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for record in table (trap visit), useful for joining with CatchRaw table",
                              "description of trap position, joins with subSite table",
                              "sample time",
                              "trap start time",
                              "describes type of visit, useful in correct handling of visitTime and visitTime2",
                              "describes if fish were processed",
                              "describes if trap fishing in thalweg, although not used by all is important to describe sampling effort",
                              "describes if trap functioning",
                              "revolutions on counter at start of trap visit",
                              "revolutions on counter at end of trap visit",
                              "revolutions per minute at start of trap visit",
                              "number of seconds used to measure rpmRevolutionsAtStart, almost always 60",
                              "revolutions per minute at end of trap visit",
                              "number of seconds used to measure rpmRevolutionsAtEnd, almost always 60",
                              "describes if trap fished at half cone, useful for sampling effort",
                              "describes if this record should be included in analysis or if is low quality and should be excluded",
                              "categorical assessment of debris in trap",
                              "volume of debris in trap",
                              "units used to measure volume of debris in trap",
                              "qualitative comments"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### TrapVisit - Are all fields needed?

All fields are not needed. Some fields are not being used. There are 7 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("crew",
                            "sampleGearID",
                            "timeSampleStarted",
                            "timeSampleEnded",
                            "batchDate",
                            "counterOld",
                            "counterOldTime",
                            "counterAtEndTime",
                            "counterNew",
                            "coneDepthAtStart",
                            "coneDepthAtEnd",
                            "includeCatchComments",
                            "coneDebrisIntake1",
                            "coneDebrisIntake2",
                            "debrisType",
                            "dataCollectedID",
                            "dataSheetNumber",
                            "dataRecorder",
                            "dataRecorderAgencyID",
                            "creationTime",
                            "updateTime",
                            "qcDone",
                            "qcDoneTime",
                            "qcComments"),
                Rationale = c("used internally for QC",
                              "this is metadata",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "used internally for QC",
                              "not used",
                              "not used",
                              "not used by majority",
                              "always 1 meaning data was successfull collected, field not being used",
                              "not used",
                              "used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used by majority",
                              "not used",
                              "not used by majority"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```
