---
title: ""
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
library(pheatmap)
library(RColorBrewer)
library(kableExtra)
library(ComplexHeatmap)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
tisdale_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "tisdale", "CAMP.mdb")
knights_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "knights_landing", "CAMP.mdb")
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
```

```{r, include = F}
project_description <- mdb.get(tisdale_camp, tables = "ProjectDescription") %>%
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(knights_camp, tables = "ProjectDescription")) %>% 
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ProjectDescription")) %>%  
  select( projectName, projectDescriptionID) 

project_id <- project_description %>% 
  distinct() %>% 
  filter(projectDescriptionID != 0) %>% 
  mutate(projectName = case_when(projectName == "Butte Creek Parrot-Phelan Diversion Dam\n-\nPhelan Diversion \nDam" ~ "Butte",
                                 projectName == "Lower Feather River RST" ~ "L Feather", 
                                 projectName == "Sacramento River Knights Landing RST" ~ "Knights",
                                 projectName == "Sacramento River Tisdale RST" ~ "Tisdale",
                                 projectName == "Feather River RST Program" ~ "U Feather")) %>% 
  filter(projectName != "Butte")
```

```{r, include = F}
make_percent_complete <- function(data) {
  data %>% 
  group_by(projectName) %>% 
  summarise_all(list(~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na") %>% 
  mutate(percent_nonNA = (1-prop_na) * 100)
    
}

make_heatmap_data <- function(percent_complete_data) {
  percent_complete_data %>% 
  select(-prop_na) %>% 
  pivot_wider(id_cols = "projectName", names_from = "field", values_from = "percent_nonNA")
}

make_heatmap <- function(df_num) {
  ComplexHeatmap::pheatmap(t(df_num), cluster_rows = F, cluster_cols = F,
                         display_numbers = T, 
                         number_format = "%.0f",
                         color = colorRampPalette(brewer.pal(n = 7, name =
  "Purples"))(100), 
                         border_color = "black", number_color = "black",
                         fontsize_number = 8, row_names_side = c("left"),
                         angle_col = c("0"), heatmap_legend_param = list(title ="Percent Complete"))
}

mark_applied <- mdb.get(tisdale_camp, tables = "MarkApplied") %>% 
  bind_rows(mdb.get(knights_camp, tables = "MarkApplied")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "MarkApplied")) %>% 
  full_join(project_id)

percent_complete_data <- mark_applied %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:16])
rownames(df_num) <- heatmap_data$projectName
```

### MarkApplied

```{r, echo = F}
make_heatmap(df_num)
```

#### MarkApplied - Useful fields

These fields contain useful information to identify the mark applied associated with a release group. This information can be joined to other tables using the projectDescriptionID and releaseID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "releaseID",
                         "appliedMarkTypeID",
                         "appliedMarkColorID",
                         "appliedMarkPositionID",
                         "markAppliedID"),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for release group",
                              "type of mark applied useful for identification during recapture",
                              "color of mark applied useful for identification during recapture",
                              "position of mark applied useful for identification during recapture",
                              "unique identifier for record in table"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### MarkApplied - Are all fields needed?

All fields are not needed. There are 6 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("dataRecorder",
                         "dataRecorderAgencyID",
                         "creationTime",
                         "updateTime",
                         "qcDone",
                         "qcDoneTime",
                         "comments",
                         "appliedMarkCode"),
                Rationale = c("used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used",
                              "not used by majority",
                              "not used"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

