---
title: ""
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = F}
library(tidyverse)
library(knitr)
library(lubridate)
library(Hmisc)
library(pheatmap)
library(RColorBrewer)
library(kableExtra)
library(ComplexHeatmap)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
tisdale_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "tisdale", "CAMP.mdb")
knights_camp <- here::here("data-raw", "qc-markdowns", "rst", "lower-sac", "knights_landing", "CAMP.mdb")
feather_camp <- (here::here("data-raw", "qc-markdowns", "rst", "feather-river", "CAMP.mdb"))
```

```{r, include = F}
project_description <- mdb.get(tisdale_camp, tables = "ProjectDescription") %>%
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(knights_camp, tables = "ProjectDescription")) %>% 
  select(projectName, projectDescriptionID) %>% 
  bind_rows(mdb.get(feather_camp, tables = "ProjectDescription")) %>%  
  select( projectName, projectDescriptionID) 

project_id <- project_description %>% 
  distinct() %>% 
  filter(projectDescriptionID != 0) %>% 
  mutate(projectName = case_when(projectName == "Butte Creek Parrot-Phelan Diversion Dam\n-\nPhelan Diversion \nDam" ~ "Butte",
                                 projectName == "Lower Feather River RST" ~ "L Feather", 
                                 projectName == "Sacramento River Knights Landing RST" ~ "Knights",
                                 projectName == "Sacramento River Tisdale RST" ~ "Tisdale",
                                 projectName == "Feather River RST Program" ~ "U Feather")) %>% 
  filter(projectName != "Butte")
```

```{r, include = F}
make_percent_complete <- function(data) {
  data %>% 
  group_by(projectName) %>% 
  summarise_all(list(~sum(is.na(.))/length(.))) %>% 
  pivot_longer(!projectName, names_to = "field", values_to = "prop_na") %>% 
  mutate(percent_nonNA = (1-prop_na) * 100)
    
}

make_heatmap_data <- function(percent_complete_data) {
  percent_complete_data %>% 
  select(-prop_na) %>% 
  pivot_wider(id_cols = "projectName", names_from = "field", values_from = "percent_nonNA")
}

make_heatmap <- function(df_num) {
  ComplexHeatmap::pheatmap(t(df_num), cluster_rows = F, cluster_cols = F,
                         display_numbers = T, 
                         number_format = "%.0f",
                         color = colorRampPalette(brewer.pal(n = 7, name =
  "Purples"))(100), 
                         border_color = "black", number_color = "black",
                         fontsize_number = 8, row_names_side = c("left"),
                         angle_col = c("0"), heatmap_legend_param = list(title ="Percent Complete"))
}
```

### MarkApplied

```{r, echo = F}
mark_applied <- mdb.get(tisdale_camp, tables = "MarkApplied") %>% 
  bind_rows(mdb.get(knights_camp, tables = "MarkApplied")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "MarkApplied")) %>% 
  full_join(project_id)

percent_complete_data <- mark_applied %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:16])
rownames(df_num) <- heatmap_data$projectName

make_heatmap(df_num)

```

#### MarkApplied - Useful fields

These fields contain useful information to identify the mark applied associated with a release group. This information can be joined to other tables using the projectDescriptionID and releaseID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "releaseID",
                         "appliedMarkTypeID",
                         "appliedMarkColorID",
                         "appliedMarkPositionID",
                         "markAppliedID"),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for release group",
                              "type of mark applied useful for identification during recapture",
                              "color of mark applied useful for identification during recapture",
                              "position of mark applied useful for identification during recapture",
                              "unique identifier for record in table"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### MarkApplied - Are all fields needed?

All fields are not needed. There are 6 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("dataRecorder",
                         "dataRecorderAgencyID",
                         "creationTime",
                         "updateTime",
                         "qcDone",
                         "qcDoneTime",
                         "comments",
                         "appliedMarkCode"),
                Rationale = c("used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used",
                              "not used by majority",
                              "not used"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

### CatchRaw

```{r, echo = F}
catch_raw <- mdb.get(tisdale_camp, tables = "CatchRaw") %>% 
  mutate(qcDoneTime = as.Date(qcDoneTime)) %>% 
  bind_rows(mdb.get(knights_camp, tables = "CatchRaw") %>% 
              mutate(qcDoneTime = as.Date(qcDoneTime))) %>% 
  bind_rows(mdb.get(feather_camp, tables = "CatchRaw") %>% 
              mutate(taxonID = as.numeric(taxonID),
                     qcDoneTime = as.Date(qcDoneTime))) %>% 
  full_join(project_id)
  

percent_complete_data <- catch_raw %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:36])
rownames(df_num) <- heatmap_data$projectName

make_heatmap(df_num)
```

#### CatchRaw - Useful fields

These fields contain useful information describing fish caught in trap. This information can be joined to other tables using the projectDescriptionID, catchRawID, trapVisitID, releaseID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "catchRawID",
                         "trapVisitID",
                         "taxonID",
                         "atCaptureRunID",
                         "atCaptureRunMethodID",
                         "finalRunID",
                         "finalRunMethodID",
                         "fishOriginID",
                         "lifeStageID",
                         "forkLength",
                         "totalLength",
                         "weight",
                         "n",
                         "randomID",
                         "actualCountID",
                         "releaseID",
                         "mortID",
                         "comments"),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for record in table (fish caught in trap)",
                              "unique identifier for trap visit, useful for joining with TrapVisit table",
                              "taxon of species, use lookup table for common name of species",
                              "run designation when fish captured in field",
                              "method used to designate run when captured in field",
                              "final run designation made later, not all programs fill this in the same way else percent complete would be the same for atCapture and final",
                              "method used to designate final run of fish",
                              "describes the origin or production of fish",
                              "describes lifestage of fish",
                              "fork length of fish",
                              "total length of fish, although not all use this field it is retained because in some cases fork length is not available and total length is",
                              "weight of fish, although not all use this field it is retained because it is an important metric to describe fish condition",
                              "count",
                              "describes if fish sampled randomly",
                              "describes if count is actual or estimate",
                              "unique identifier of release group, useful for joining with Release table or ReleaseFish table",
                              "describes if fish is dead or not",
                              "qualitative comments"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### CatchRaw - Are all fields needed?

All fields are not needed. Some fields are not being used. There are 7 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("auxiliaryRunID",
                            "auxiliaryRunMethodID",
                            "stdLength",
                            "subsampleMethodID",
                            "subsampleNumberator",
                            "subsampleDenominator",
                            "fishHealth",
                            "gillRating",
                            "dataSheetNumber",
                            "dataRecorder",
                            "dataRecorderAgencyID",
                            "creationTime",
                            "updateTime",
                            "qcDone",
                            "qcDoneTime",
                            "qcComments"),
                Rationale = c("not used",
                              "not used",
                              "not used",
                              "subSampleMethod is always 1 for all meaning that all fish are counted or estimated not using subsampling and this field is not needed",
                              "not needed because all fish counted or estimated not using subsampling",
                              "not needed because all fish counted or estimated not using subsampling",
                              "not used by majority",
                              "not used",
                              "not used",
                              "used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used by majority",
                              "not used",
                              "not used by majority"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

### TrapVisit

```{r, echo = F}
trap_visit <- mdb.get(tisdale_camp, tables = "TrapVisit") %>% 
  bind_rows(mdb.get(knights_camp, tables = "TrapVisit")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "TrapVisit")) %>% 
  full_join(project_id)
  

percent_complete_data <- trap_visit %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:46])
rownames(df_num) <- heatmap_data$projectName

make_heatmap(df_num)
```

#### TrapVisit - Useful fields

These fields contain useful information describing trap visit. This information can be joined to other tables using the projectDescriptionID, trapVisitID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "trapVisitID",
                         "trapPositionID",
                         "visitTime",
                         "visitTime2",
                         "visitTypeID",
                         "fishProcessedID",
                         "inThalwegID",
                         "trapFunctioningID",
                         "counterAtStart",
                         "counterAtEnd",
                         "rpmRevolutionsAtStart",
                         "rpmSecondsAtStart",
                         "rpmRevolutionsAtEnd",
                         "rpmSecondsAtEnd",
                         "halfConeID",
                         "includeCatchID",
                         "debrisVolumeCatID",
                         "debrisVolume",
                         "debrisVolumeUnits",
                         "comments"
                     ),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for record in table (trap visit), useful for joining with CatchRaw table",
                              "description of trap position, joins with subSite table",
                              "sample time",
                              "trap start time",
                              "describes type of visit, useful in correct handling of visitTime and visitTime2",
                              "describes if fish were processed",
                              "describes if trap fishing in thalweg, although not used by all is important to describe sampling effort",
                              "describes if trap functioning",
                              "revolutions on counter at start of trap visit",
                              "revolutions on counter at end of trap visit",
                              "revolutions per minute at start of trap visit",
                              "number of seconds used to measure rpmRevolutionsAtStart, almost always 60",
                              "revolutions per minute at end of trap visit",
                              "number of seconds used to measure rpmRevolutionsAtEnd, almost always 60",
                              "describes if trap fished at half cone, useful for sampling effort",
                              "describes if this record should be included in analysis or if is low quality and should be excluded",
                              "categorical assessment of debris in trap",
                              "volume of debris in trap",
                              "units used to measure volume of debris in trap",
                              "qualitative comments"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### TrapVisit - Are all fields needed?

All fields are not needed. Some fields are not being used. There are 7 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("crew",
                            "sampleGearID",
                            "timeSampleStarted",
                            "timeSampleEnded",
                            "batchDate",
                            "counterOld",
                            "counterOldTime",
                            "counterAtEndTime",
                            "counterNew",
                            "coneDepthAtStart",
                            "coneDepthAtEnd",
                            "includeCatchComments",
                            "coneDebrisIntake1",
                            "coneDebrisIntake2",
                            "debrisType",
                            "dataCollectedID",
                            "dataSheetNumber",
                            "dataRecorder",
                            "dataRecorderAgencyID",
                            "creationTime",
                            "updateTime",
                            "qcDone",
                            "qcDoneTime",
                            "qcComments"),
                Rationale = c("used internally for QC",
                              "this is metadata",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "not used",
                              "used internally for QC",
                              "not used",
                              "not used",
                              "not used by majority",
                              "always 1 meaning data was successfull collected, field not being used",
                              "not used",
                              "used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used by majority",
                              "not used",
                              "not used by majority"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

### MarkExisting

```{r, echo = F}
mark_existing <- mdb.get(tisdale_camp, tables = "MarkExisting") %>% 
  bind_rows(mdb.get(knights_camp, tables = "MarkExisting")) %>% 
  bind_rows(mdb.get(feather_camp, tables = "MarkExisting") %>% 
              mutate(markCode = as.character(markCode))) %>% 
  full_join(project_id) 
  

percent_complete_data <- mark_existing %>% 
  make_percent_complete()

heatmap_data <- percent_complete_data %>% 
  make_heatmap_data()

df_num <- as.matrix(heatmap_data[,2:15])
rownames(df_num) <- heatmap_data$projectName

make_heatmap(df_num)
```

#### MarkExisting - Useful fields

These fields contain useful information to identify the release group that corresponds with an existing mark. This information can be joined to other tables using the projectDescriptionID and catchRawID.

```{r, echo = F}
table <- tibble(Keep = c("projectDescriptionID",
                         "catchRawID",
                         "markTypeID",
                         "markColorID",
                         "markPositionID",
                         "markCode",
                         "markExistingID"),
                Rationale = c("unique identifier for monitoring program",
                              "unique identifier for fish caught in trap, useful for joining with CatchRaw table",
                              "type of existing mark, useful for identification of release group",
                              "color of existing mark, useful for identification of release group",
                              "position of existing mark, useful for identification of release group",
                              "code of existing mark (e.g., PIT, CWT), useful to connect with other studies and releases",
                              "unique identifier for record in table"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```

#### MarkExisting - Are all fields needed?

All fields are not needed. There are 7 fields that are primarily for internal use
when performing QC. When sharing data it is assumed that these QC checks have already
been performed and this information is no longer needed.

```{r, echo = F}
table <- tibble(Exclude = c("dataRecorder",
                         "dataRecorderAgencyID",
                         "creationTime",
                         "updateTime",
                         "qcDone",
                         "qcDoneTime",
                         "qcComments"),
                Rationale = c("used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "used internally for QC",
                              "not used",
                              "not used",
                              "not used",
                              "not used by majority"))
kbl(table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14)
```
