---
title: "yuba-river-rst-qc-checklist"
author: "Inigo Peng"
date: "10/11/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(leaflet)
```
# Yuba River RST Data  

## Description of Monitoring Data

**Timeframe:** 2000-2008


**Completeness of Record throughout timeframe: **\

TODO


**Sampling Location:** Yuba River


**Data Contact:** [Robyn Bilski](Robyn.Bilski@Wildlife.ca.gov)


## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "rst/yuba-river/data-raw/yuba-river-rst-data.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "yuba-rst.xlsx",
               overwrite = TRUE)


```

Read in data from google cloud, glimpse raw data:
```{r}
raw_data = readxl::read_excel('yuba-rst.xlsx', col_types = c("text",
                                                             "date",
                                                             "date",
                                                             "text",
                                                             "numeric",
                                                             "numeric",
                                                             "numeric",
                                                             "text",
                                                             "numeric",
                                                             "numeric",
                                                             "numeric",
                                                             "numeric",
                                                             "numeric",
                                                             "text",
                                                             "text",
                                                             "text",
                                                             "text",
                                                             "numeric",
                                                             "numeric",
                                                             "text",
                                                             "numeric",
                                                             'text',
                                                             'numeric',
                                                             'numeric'))
glimpse(raw_data)
```

## Data Transformation

```{r}
cleaner_data <- raw_data %>% 
  select(-c('Catch_Entry_SampleRowID', 'CatchRowID')) %>% 
  rename('id' = Sample_Entry_SampleRowID,
         'date'= SampleDate,
         'time' = SampleTime,
         'method' = MethodCode,
         'water_temp' = WaterTemperature,
         'turbidity' = Turbidity,
         'water_velocity' = WaterVelocity,
         'trap_status' = TrapStatus,
         'trap_revolutions' = TrapRevolutions,
         'trap_revolutions2' = TrapRevolutions2,
         'rpms_before' = RPMsBefore,
         'rpms_after' = RPMsAfter,
         'organism_code' = OrganismCode,
         'fork_length' = ForkLength,
         'stage_code' = StageCode
         ) %>%
  mutate(time = hms::as_hms(time)) %>%
  filter(organism_code == 'CHN', rm.na = TRUE) %>% 
  glimpse()
```
```{r}
cleaner_data <- cleaner_data %>% 
  set_names(tolower(colnames(cleaner_data))) %>% 
  glimpse()
```

## Explore Numeric Variables

```{r}
cleaner_data %>% 
  select_if(is.numeric) %>% colnames()
```

### Variable:`water_temp`
#TODO: figure out what's happenign with the high temperature
#note: 2000 - fst had lower temperature (15s) while rst had high temp (60s) on the same day

```{r}
cleaner_data %>% 
  group_by(date) %>%
  mutate(avg_temp = mean(water_temp, na.rm = T)) %>%
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Water Temperature (colored by year)",
       y = "Average daily temp", 
       x = "Date")  
```
```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = water_temp, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Water Temperature summarized by year",
       x = "Water Temperature C") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


```{r}
summary(cleaner_data$water_temp)
```


### Variable:`turbidity`
```{r}
cleaner_data %>% 
  group_by(date) %>%
  mutate(avg_temp = mean(turbidity, na.rm = T)) %>%
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Water Turbidity (colored by year)",
       y = "Average daily turbdity", 
       x = "Date")  
```
```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = turbidity, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Turbdity summarized by year",
       x = "Turbidity") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


