---
title: "Battle Creek RST QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
palette <- wesanderson::wes_palette(name = "Moonrise2")
```

# Battle Creek Rotary Screw Trap Data

## Description of Monitoring Data

These data were collected by the U.S. Fish and Wildlife Service, Red Bluff Fish and Wildlife Office, Battle Creek Monitoring Program. These data represent environmental conditions for Battle Creek RST. 									

**Timeframe:** 2003 - 2021

**Screw Trap Season:** September - June

**Completeness of Record throughout timeframe:** Sample Year tab on excel sheet describes start and end date for trap each year. Sampled every year from 1998 - 2019, some years not fished on weekends or during high flow events. Proxy dates are used when environmental conditions were not measured (see WeekSub column).

**Sampling Location:** Upper Battle Creek (UBC)

**Data Contact:** [Mike Schraml](mailto:mike_schraml@fws.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "rst/battle-creek/data-raw/UBC Spring.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_battle_rst_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
sheets <- excel_sheets("raw_battle_rst_data.xlsx")
sheets
raw_environmental_1 <- read_excel("raw_battle_rst_data.xlsx", 
                                  sheet = "UBC Environmental 03-20") %>% glimpse()
raw_environmental_2 <- read_excel("raw_battle_rst_data.xlsx", 
                                  sheet = "UBC Environmental 20-21") %>% glimpse()
```

## Data transformations

```{r}
raw_rst_environmental <- bind_rows(raw_environmental_1, raw_environmental_2) %>% 
  glimpse()

cleaner_rst_environmental <- raw_rst_environmental %>%
  janitor::clean_names() %>%
  rename() %>%
  mutate(trap_start_date = as.Date(trap_start_date),
         trap_start_time = hms::as_hms(trap_start_time),
         sample_date = as.Date(sample_date),
         sample_time = hms::as_hms(sample_time)) %>%
  select(-station_code, -user_name, -user_name2, 
         -sample_weight) %>% # remove sample weight because it is defined to be the same as cone
  glimpse

```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_rst_environmental %>% select_if(is.numeric) %>% colnames 
```

### Variable: `counter`

Number on the cone revolution counter at SampleDate and SampleTime							

**Plotting distribution of counter**

```{r}
cleaner_rst_environmental %>% 
  filter(counter < 25000) %>% # filter out 5 values > 25000
  ggplot() +
  geom_histogram(aes(x = counter), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_boxplot(aes(x = counter, y = as.factor(year(sample_date))), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Year")
```



**Numeric Summary of counter over Period of Record**

```{r}
summary(cleaner_rst_environmental$counter)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$counter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `counter` column are NA. 

### Variables: `flow_start_meter`, `flow_end_meter`

The read out number on the mechanical counter of the flow meter at the start and end of the velocity test							


**Plotting distribution of flow start meter and flow end meter**

```{r}

cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = flow_start_meter), fill = "blue", alpha = .5) +
  geom_histogram(aes(x = flow_end_meter), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```

Start times and end times appear to have a similar distribution.

**Numeric Summary**

```{r}
summary(cleaner_rst_environmental$flow_start_meter)
summary(cleaner_rst_environmental$flow_end_meter)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$flow_start_meter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `flow_start_meter` column are NA. 
* `r round(sum(is.na(cleaner_rst_environmental$flow_end_meter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `flow_end_meter` column are NA. 


### Variable: `flow_set_time`

How long the General Oceanics mechanical flow meter (Oceanic ® Model 2030) was in the water taking a reading, used to calculate water velocity in front of the cone

Time is in seconds 

**Plotting distribution of flow set time**

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = flow_set_time)) +
  theme_minimal() +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

Flow set times appear mainly between 100 and 1000 with a few values up to almost 4000. 

**Numeric Summary of flow set time over Period of Record**

```{r}
summary(cleaner_rst_environmental$flow_set_time)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$counter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `counter` column are NA. 


### Variable: `velocity`

Calculated water velocity in front of the cone using a General Oceanics mechanical flow meter (Oceanic ® Model 2030) = ( (Flowmeter end - flow meter begin)/time in seconds)*.0875							

Since velocity is just a function of flowmeter_end, flowmeter_start, and flowmeter_set_time we probably do not need to include all of these. 

**Plotting distribution of velocity**

```{r}
cleaner_rst_environmental %>% 
  filter(velocity < 25) %>% # filter out values greater than 25
  ggplot() +
  geom_histogram(aes(x = velocity), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>%
  mutate(wy = factor(ifelse(month(sample_date) %in% 10:12, year(sample_date) + 1, year(sample_date))),
         fake_year = 2000,
         fake_year = ifelse(month(sample_date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(sample_date), day(sample_date)))) %>%
  ggplot(aes(x = fake_date, y = velocity)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("") +
  facet_wrap(~wy, scales = "free") + 
  theme_minimal()
```



**Numeric Summary of velocity**

```{r}
summary(cleaner_rst_environmental$velocity)
```

A velocity of 155.090 seems out of the range of possibilities. It seems like every velocity greater than 7 is probably a mistake that should be scaled down or filtered out. Most likely these were caused by the outliers in flow_start_time, flow_end_time or flow_set_time since this is a calculatted value. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$velocity))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `velocity` column are NA.

### Variable: `turbidity`

Turbidity result from a grab sample taken at the trap on the SampleDate and SampleTime							

**Plotting distribution of turbidity**
```{r}
cleaner_rst_environmental %>% 
  filter(counter < 200) %>% # filter out values < 200
  ggplot() +
  geom_histogram(aes(x = turbidity), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
  
```{r}

cleaner_rst_environmental %>% 
  group_by(date = as.Date(sample_date)) %>%
  mutate(avg_turbidity_ntu = mean(turbidity)) %>%
  filter(avg_turbidity_ntu < 100) %>%
  ungroup() %>%
  ggplot() + 
  geom_boxplot(aes(x = as.factor(month(date)), y = avg_turbidity_ntu)) + 
  # facet_wrap(~year(date), scales = "free") + 
  # scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Turbidity Measures sumarized by month",
       x = "Month", 
       y = "Average Daily Turbidity NTUs")  
```

```{r}
cleaner_rst_environmental %>%
  mutate(wy = factor(ifelse(month(sample_date) %in% 10:12, year(sample_date) + 1, year(sample_date))),
         fake_year = 2000,
         fake_year = ifelse(month(sample_date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(sample_date), day(sample_date)))) %>%
  ggplot(aes(x = fake_date, y = turbidity)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("") +
  facet_wrap(~wy, scales = "free") + theme_minimal()
```

Most turbidity measures are low but a few outliers going up until 850

**Numeric Summary of turbidity over Period of Record**
  
```{r}
summary(cleaner_rst_environmental$turbidity)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_environmental$turbidity))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `turbidity` column are NA.




### Variables: `cone`

Definition `cone`:
Was the trap fished at cone full-cone (1.0) or half-cone (0.5) setting				
(same as `sample_weight` column removed above)

**Plotting distribution of cone**
```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = cone), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

All sample weights and cone measures are either .5 o 1. 

**Numeric Summary of cone over Period of Record**
  
```{r}
summary(cleaner_rst_environmental$cone)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_environmental$cone))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `cone` column are NA.

### Variable: `river_left_depth`, `river_right_depth`, `river_center_depth`

Unit for depth is feet, definitions of measurements are: 

* River depth from directly in the center of cone off crossbeam #2 (cone crossbeam)							
* River depth from inside of the river left (facing down stream) pontoon off crossbeam #2 (cone crossbeam)							
* River depth from inside of the river right (facing down stream) pontoon off crossbeam #2 (cone crossbeam)					

**Plotting distribution of depth measures**
```{r}
depth_1 <- cleaner_rst_environmental %>% 
  # filter(counter < 200) %>% # filter out values < 200
  ggplot() +
  geom_histogram(aes(x = river_left_depth), fill = "blue", alpha = .5) +
  geom_histogram(aes(x = river_center_depth), fill = "gray", alpha = .75) +
  geom_histogram(aes(x = river_right_depth), fill = "green", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "River Depth: left (blue), right (green), and center (gray)")

depth_2 <- cleaner_rst_environmental %>% 
  filter(river_right_depth > 1) %>% # filter out values < 200
  ggplot() +
  geom_histogram(aes(x = river_left_depth), fill = "blue", alpha = .5) +
  geom_histogram(aes(x = river_center_depth), fill = "gray", alpha = .75) +
  geom_histogram(aes(x = river_right_depth), fill = "green", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x = "River Depth: left (blue), right (green), and center (gray)",
       title = "Filtered Depth Distribution")
gridExtra::grid.arrange(depth_1, depth_2)
```


**Numeric Summary of river depth over Period of Record**
  
```{r}
summary(cleaner_rst_environmental$river_left_depth)
summary(cleaner_rst_environmental$river_center_depth)
summary(cleaner_rst_environmental$river_right_depth)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_environmental$river_left_depth))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `river_left_depth` column are NA.
* `r round(sum(is.na(cleaner_rst_environmental$river_center_depth))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `river_center_depth` column are NA.
* `r round(sum(is.na(cleaner_rst_environmental$river_right_depth))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `river_right_depth` column are NA.


### Variable: `depth_adjust`

The depth of the bottom of the cone	(measured in Inches)
Depth in relation to the cone (not to the surface of the water) - not sure how it is used. 

**Plotting distribution of depth adjustment**

```{r}
cleaner_rst_environmental %>% 
  # filter(counter < 25000) %>% # filter out 5 values > 25000
  ggplot() +
  geom_histogram(aes(x = depth_adjust), fill = "blue", alpha = .5, binwidth = 1) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_boxplot(aes(x = depth_adjust, y = as.factor(year(sample_date))), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Year")
```
Looks like depth adjust varies by year depending on how the trap was positioned. 


**Numeric Summary of depth adjustment over Period of Record**

```{r}
summary(cleaner_rst_environmental$depth_adjust)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$depth_adjust))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `depth_adjust` column are NA.

### Variable: `debris_tubs`

The number of 10-g tubs of debris removed from the trap during the sample period (volumetrically)							
**Plotting distribution of debris_tubs**

```{r}
cleaner_rst_environmental %>% 
  # filter(counter < 25000) %>% # filter out 5 values > 25000
  ggplot() +
  geom_histogram(aes(x = debris_tubs), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_boxplot(aes(x = debris_tubs, y = as.factor(month(sample_date))), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Month")
```

Debris looks relatively evenly distributed throughout the months. 

**Numeric Summary of debris over Period of Record**

```{r}
summary(cleaner_rst_environmental$debris_tubs)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$debris_tubs))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `debris_tubs` column are NA.

### Variable: `avg_time_per_rev`

The average time per cone rotation (average of three rotations)	- units are seconds 					

**Plotting distribution of average time per revolution**

```{r}
cleaner_rst_environmental %>% 
  # filter(counter < 25000) %>% # filter out 5 values > 25000
  ggplot() +
  geom_histogram(aes(x = avg_time_per_rev), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_boxplot(aes(x = avg_time_per_rev, y = as.factor(year(sample_date))), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Year")
```



**Numeric Summary of average time per rev over Period of Record**

```{r}
summary(cleaner_rst_environmental$avg_time_per_rev)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$avg_time_per_rev))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `avg_time_per_rev` column are NA.


### Variable: `baileys_eff`

Trap Efficiency = (Recaptured+1)/(Released+1); used to calculate daily passage							
						

**Plotting distribution of counter**

```{r}
cleaner_rst_environmental %>% 
  # filter(counter < 25000) %>% # filter out 5 values > 25000
  ggplot() +
  geom_histogram(aes(x = baileys_eff), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_boxplot(aes(x = baileys_eff, y = as.factor(year(sample_date))), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Year")
```

**Numeric Summary of baileys efficency over Period of Record**

```{r}
summary(cleaner_rst_environmental$baileys_eff)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$baileys_eff))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `baileys_eff` column are NA.

### Variable: `num_released`

Number fish released					

**Plotting distribution of number released**

```{r}
cleaner_rst_environmental %>% 
  # filter(counter < 25000) %>% # filter out 5 values > 25000
  ggplot() +
  geom_histogram(aes(x = num_released), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_boxplot(aes(x = num_released, y = as.factor(year(sample_date))), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y = "Year")
```



**Numeric Summary of counter over Period of Record**

```{r}
summary(cleaner_rst_environmental$num_released)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$num_released))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `num_released` column are NA.


### Variable: `start_counter`

Beginning cone revolution counter number, usually zero.							
							

All values either 0 or NA. 

**Numeric Summary of counter over Period of Record**

```{r}
summary(cleaner_rst_environmental$start_counter)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$start_counter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `start_counter` column are NA.


### Variable: `report_baileys_eff`

The Bailey's efficiency used in old reports. Bailey's efficiency has been standardized to only four significant digits.							

**Plotting distribution of counter**

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = report_baileys_eff), fill = "blue", alpha = .5, binwidth = .01) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of baileys efficency measures for reports over Period of Record**

```{r}
summary(cleaner_rst_environmental$report_baileys_eff)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$report_baileys_eff))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `report_baileys_eff` column are NA.



## Explore Categorical variables: {.tabset}

```{r}
cleaner_rst_environmental %>% select_if(is.character) %>% colnames
```


### Variable: `sample_id`

The calendar year Julian date and year code for that ~24-h sample period (ddd_yy)	

```{r}
nrow(cleaner_rst_environmental) == length(unique(cleaner_rst_environmental$sample_id))

```
There are `r length(unique(cleaner_rst_environmental$sample_id))` unique sample IDs.

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$sample_id))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `sample_id` column are NA. 


### Variable: `weather_code`

A code for the weather conditions on the SampleDate and SampleTime. See VariableCodesLookUp table			

| code | description |			
| ---- | ------------ |
| CLR |	sunny |
| RAN |	precipitation |
| FOG |	foggy |
| CLD |	overcast |


```{r}
table(cleaner_rst_environmental$weather_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$weather_code <- case_when(
  cleaner_rst_environmental$weather_code %in% c("CLD", "cld", "CLOUDY") ~ "cloudy", 
  cleaner_rst_environmental$weather_code == "CLR" ~ "clear",
  cleaner_rst_environmental$weather_code == "FOG" ~ "fog", 
  cleaner_rst_environmental$weather_code == "PCLD" ~ "partially cloudy", 
  cleaner_rst_environmental$weather_code == "RAN" ~ "precipitation",
  cleaner_rst_environmental$weather_code == "WIN/CLR" ~ "windy and clear",
  cleaner_rst_environmental$weather_code == "WINDY" ~ "windy"
)

table(cleaner_rst_environmental$weather_code) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$weather_code))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `weather_code` column are NA.

### Variable: `lunar_phase`

Stage of the Moon on the SampleDate and SampleTime, from: http://aa.usno.navy.mil/data/docs/MoonFraction.php. See VariableCodesLookUp table	

```{r}
table(cleaner_rst_environmental$lunar_phase) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$lunar_phase <- case_when(
  cleaner_rst_environmental$lunar_phase == "F" ~ "full", 
  cleaner_rst_environmental$lunar_phase == "H" ~ "half", 
  cleaner_rst_environmental$lunar_phase == "N" ~ "new"
)

table(cleaner_rst_environmental$lunar_phase) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$lunar_phase))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `lunar_phase` column are NA.

### Variable: `trap_sample_type`

The type of sample regime, see VariableCodesLookUp table	

| code |  definition  |
| --- | ------------ | 
| N |	non-intensive |
| I | intensive |
| S | sunrise-sunset |
| R | random |


```{r}
table(cleaner_rst_environmental$trap_sample_type) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$trap_sample_type <- case_when(
  cleaner_rst_environmental$trap_sample_type == "N" ~ "non-intensive",
  cleaner_rst_environmental$trap_sample_type == "R" ~ "random"
)

table(cleaner_rst_environmental$trap_sample_type) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$trap_sample_type))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `trap_sample_type` column are NA.

### Variable: `habitat`

The type of flow habitat the trap fished in, see VariableCodesLookUp table	

| code | definition |
| ---- | ---------- |
| P |	plunge pool |
| R |	run |
| G |	glide |
| B |	backwater pool |
| L |	lateral flood |


```{r}
table(cleaner_rst_environmental$habitat) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$habitat <- case_when(
  cleaner_rst_environmental$habitat == "B" ~ "backwater pool",
  cleaner_rst_environmental$habitat == "R" ~ "run"
)

table(cleaner_rst_environmental$habitat) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$habitat))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `habitat` column are NA.

### Variable: `thalweg`

Was trap fishing in the thalweg at SampleDate and SampleTime							

```{r}
table(cleaner_rst_environmental$thalweg) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$thalweg <- case_when(
  cleaner_rst_environmental$thalweg == "Y" ~ TRUE, 
  cleaner_rst_environmental$thalweg == "N" ~ FALSE
)

table(cleaner_rst_environmental$thalweg) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$thalweg))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `thalweg` column are NA.

### Variable: `diel`

The time of day relative to the sun, see VariableCodesLookUp table	


| code | definition |
| ---- | ---------- |
| C1 |	pre-sunrise |
| C2 |	post-sunrise |
| C3 |	pre-sunset |
| C4 |	post-sunset |
| D |	day |
| N |	night |


```{r}
table(cleaner_rst_environmental$diel) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$diel <- case_when(cleaner_rst_environmental$diel == "D" ~ "day", 
                                            cleaner_rst_environmental$diel == "N" ~ "night", 
                                            cleaner_rst_environmental$diel == "C1" ~ "pre-sunrise")
table(cleaner_rst_environmental$diel) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$diel))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `diel` column are NA.

### Variable: `debris_type`

The type of debris found in the live-box, see VariableCodesLookUp table			


| code | definition |
| ----- | -------------- |
| a |	aquatic vegetation |
| l |	large woody debris |
| s |	sticks |
| al |	veg + wood |
| as |	veg + sticks |
| ls |	wood + sticks |
| als |	veg + wood + sticks |


```{r}
table(cleaner_rst_environmental$debris_type) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$debris_type <- tolower(cleaner_rst_environmental$debris_type)
table(cleaner_rst_environmental$debris_type) 
```
**Create lookup rda for gear debris encoding:** 
```{r}
# View description of domain for viewing condition 
battle_rst_debris_type <- c('a','l','s','al','as','ls','als')
names(battle_rst_debris_type) <- c(
  "aquatic vegetation",
  "large woody debris",
  "sticks",
  "veg + wood",
  "veg + sticks",
  "wood + sticks",
  "veg + wood + sticks")
# write_rds(battle_rst_debris_type, "../../../data/battle_rst_debris_type.rds")
tibble(code = battle_rst_debris_type, 
       definitions = names(battle_rst_debris_type))
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$debris_type))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `debris_type` column are NA.

### Variable: `fish_properly`

Was there a problem with the trap at SampleDate and SampleTime							

```{r}
table(cleaner_rst_environmental$fish_properly) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$fish_properly <- case_when(
  cleaner_rst_environmental$fish_properly == "Y" ~ TRUE, 
  cleaner_rst_environmental$fish_properly == "N" ~ FALSE
)

table(cleaner_rst_environmental$fish_properly) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$fish_properly))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `fish_properly` column are NA.

### Variable: `sub_week`

If sample week has more than one efficiency, which part of week is sample from	

Description from Mike: The sub weeks do not describe specific days of the week, they indicate when a stratum (usually one week) is split into substrata based upon the trap efficiency used for the strata or substrata. If a stratum is not split all days in the stratum will be an A. If it is split the days in the first substratum are A’s, the second are B’s, etc.

```{r}
table(cleaner_rst_environmental$sub_week) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$sub_week))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `sub_week` column are NA.

### Variable: `trap_comments`

```{r}
unique(cleaner_rst_environmental$trap_comments)[1:5]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$trap_comments))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `trap_comments` column are NA.

### Variable: `gear_condition_code`

A code for the condition of the trap on the SampleDate and SampleTime; see VariableCodesLookUp table				

| code | definition  |
| ---- | ----------- |
| n	|normal |
| pb|	partial block |
| tb|	total block |
| nr|	not rotating |


```{r}
table(cleaner_rst_environmental$gear_condition_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$gear_condition_code <- case_when(
  cleaner_rst_environmental$gear_condition_code == "BP" ~ "partial block",
  cleaner_rst_environmental$gear_condition_code == "N" ~ "normal", 
  cleaner_rst_environmental$gear_condition_code == "NR" ~ "not rotating",
  cleaner_rst_environmental$gear_condition_code == "TB" ~ "total block"
)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$gear_condition_code))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `gear_condition_code` column are NA.

### Variable: `trap_fishing`

```{r}
table(cleaner_rst_environmental$trap_fishing) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$trap_fishing <- case_when(
  cleaner_rst_environmental$trap_fishing == "Yes" ~ TRUE, 
  cleaner_rst_environmental$trap_fishing == "No" ~ FALSE
)

table(cleaner_rst_environmental$trap_fishing) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$trap_fishing))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `trap_fishing` column are NA.

### Variable: `partial_sample`

```{r}
table(cleaner_rst_environmental$partial_sample) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_environmental$partial_sample <- case_when(
  cleaner_rst_environmental$partial_sample == "Yes" ~ TRUE, 
  cleaner_rst_environmental$partial_sample == "No" ~ FALSE
)

table(cleaner_rst_environmental$partial_sample) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$partial_sample))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `partial_sample` column are NA.


## Summary of identified issues

* Need to figure out what sub_week column values stand for. Answer from Mike: The sub weeks do not describe specific days of the week, they indicate when a stratum (usually one week) is split into substrata based upon the trap efficiency used for the strata or substrata. If a stratum is not split all days in the stratum will be an A. If it is split the days in the first substratum are A's, the second are B's, etc.  
* Outliers in some of the numeric variables: 
  * Velocity (seems like anything greater than 7 needs to be addressed)
  * Turbidity (everything greater than 100 needs to be addressed)
  * Counter (one value way larger than the others)
* There are a few variables that I am unsure of how they would be used (ex `depth_adjust`). Asking Mike and these may not be relevant to us and can be filtered out. 


## Save cleaned data back to google cloud 

```{r}
battle_rst_environmental <- cleaner_rst_environmental %>% glimpse()
```
```{r, eval=FALSE}
# Write to google cloud 
# Name file [watershed]_[data type].csv
f <- function(input, output) write_csv(input, file = output)
gcs_upload(battle_rst_environmental,
           object_function = f,
           type = "csv",
           name = "rst/battle-creek/data/battle_rst_environmental.csv")
```
