---
title: "Battle Creek RST QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
palette <- wesanderson::wes_palette(name = "Moonrise2")
```

# Battle Creek Rotary Screw Trap Data

## Description of Monitoring Data

These data were collected by the U.S. Fish and Wildlife Service's, Red Bluff Fish and Wildlife Office's, Battle Creek Monitoring Program. These data represent environmental conditions for Battle Creek RST. 									

**Timeframe:** 2003 - 2021

**Screw Trap Season:** September - June

**Completeness of Record throughout timeframe:** 

**Sampling Location:** Upper Battle Creek (UBC)

**Data Contact:** [Mike Schraml](mailto:mike_schraml@fws.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "rst/battle-creek/data-raw/UBC Spring.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_battle_rst_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
sheets <- excel_sheets("raw_battle_rst_data.xlsx")
sheets
raw_environmental_1 <- read_excel("raw_battle_rst_data.xlsx", 
                                  sheet = "UBC Environmental 03-20") %>% glimpse()
raw_environmental_2 <- read_excel("raw_battle_rst_data.xlsx", 
                                  sheet = "UBC Environmental 20-21") %>% glimpse()
```

## Data transformations

```{r}
raw_rst_environmental <- bind_rows(raw_environmental_1, raw_environmental_2) %>% 
  glimpse()

cleaner_rst_environmental <- raw_rst_environmental %>%
  janitor::clean_names() %>%
  rename() %>%
  mutate(trap_start_date = as.Date(trap_start_date),
         trap_start_time = hms::as_hms(trap_start_time),
         sample_date = as.Date(sample_date),
         sample_time = hms::as_hms(sample_time)) %>%
  select(-station_code, -user_name, -user_name2) %>%
  glimpse

```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_rst_environmental %>% select_if(is.numeric) %>% colnames 
```

### Variable: `counter`

Number on the cone revolution counter at SampleDate and SampleTime							

**Plotting distribution of counter**

```{r}
cleaner_rst_environmental %>% 
  filter(counter < 25000) %>% # filter out 5 values > 25000
  ggplot() +
  geom_histogram(aes(x = counter), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_boxplot(aes(x = counter, y = as.factor(year(sample_date))), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```



**Numeric Summary of counter over Period of Record**

```{r}
summary(cleaner_rst_environmental$counter)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$counter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `counter` column are NA. 

### Variables: `flow_start_meter`, `flow_end_meter`

The read out number on the mechanical counter of the flow meter at the start and end of the velocity test							


**Plotting distribution of flow start meter and flow end meter**

```{r}

cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = flow_start_meter), fill = "blue", alpha = .5) +
  geom_histogram(aes(x = flow_end_meter), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```

Start times and end times appear to have a similar distribution.

**Numeric Summary**

```{r}
summary(cleaner_rst_environmental$flow_start_meter)
summary(cleaner_rst_environmental$flow_end_meter)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$flow_start_meter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `flow_start_meter` column are NA. 
* `r round(sum(is.na(cleaner_rst_environmental$flow_end_meter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `flow_end_meter` column are NA. 


### Variable: `flow_set_time`

How long the General Oceanics mechanical flow meter (Oceanic ® Model 2030) was in the water taking a reading, used to calculate water velocity in front of the cone							
**Plotting distribution of flow set time**

```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = flow_set_time)) +
  theme_minimal() +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

Flow set times appear mainly between 100 and 1000 with a few values up to almost 4000. 

**Numeric Summary of flow set time over Period of Record**

```{r}
summary(cleaner_rst_environmental$flow_set_time)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$counter))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `counter` column are NA. 


### Variable: `velocity`

Calculated water velocity in front of the cone using a General Oceanics mechanical flow meter (Oceanic ® Model 2030) = ( (Flowmeter end - flow meter begin)/time in seconds)*.0875							

**Plotting distribution of velocity**

```{r}
cleaner_rst_environmental %>% 
  filter(velocity < 25) %>% # filter out values greater than 25
  ggplot() +
  geom_histogram(aes(x = velocity), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
# TODO add another graph
```



**Numeric Summary of velocity**

```{r}
summary(cleaner_rst_environmental$velocity)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$velocity))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `velocity` column are NA.

### Variable: `turbidity`

Turbidity result from a grab sample taken at the trap on the SampleDate and SampleTime							

**Plotting distribution of turbidity**
```{r}
cleaner_rst_environmental %>% 
  filter(counter < 200) %>% # filter out values < 200
  ggplot() +
  geom_histogram(aes(x = turbidity), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
  
```{r}
# TODO fix this 

# Make whatever plot is appropriate 
# maybe 2 plots is appropriate
cleaner_rst_environmental %>% 
  group_by(date = as.Date(sample_date)) %>%
  mutate(avg_turbidity_ntu = mean(turbidity)) %>%
  filter(avg_turbidity_ntu < 100) %>%
  ungroup() %>%
 mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_turbidity_ntu, color = year)) + 
  geom_point(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Turbidity Measures (colored by year)",
       x = "Date", 
       y = "Average Daily Turbidity NTUs")  
```
```{r}
# TODO fix plot
cleaner_rst_environmental %>% 
  filter(year(as.Date(sample_date)) > 1999) %>% # Filter because only a few measure before this date
  mutate(year = as.factor(year(as.Date(sample_date)))) %>%
  ggplot(aes(x = turbidity, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Water Turbidity measures summarized by year",
       x = "Turbidity NTUs") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

Most turbidity measures are low but a few outliers going up until 850

**Numeric Summary of turbidity over Period of Record**
  
```{r}
summary(cleaner_rst_environmental$turbidity)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_environmental$turbidity))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `turbidity` column are NA.




### Variables: `sample_weight`, `cone`

Definition for both `sample_weight` and `cone`:
Was the trap fished at cone full-cone (1.0) or half-cone (0.5) setting							
**Plotting distribution of sample_weight**
```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = sample_weight), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Plotting distribution of cone**
```{r}
cleaner_rst_environmental %>% 
  ggplot() +
  geom_histogram(aes(x = cone), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

All sample weights and cone measures are either .5 o 1. 

**Numeric Summary of turbidity over Period of Record**
  
```{r}
summary(cleaner_rst_environmental$sample_weight)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_environmental$sample_weight))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `sample_weight` column are NA.

### Variable: `turbidity`

Turbidity result from a grab sample taken at the trap on the SampleDate and SampleTime							

**Plotting distribution of turbidity**
```{r}
cleaner_rst_environmental %>% 
  filter(counter < 200) %>% # filter out values < 200
  ggplot() +
  geom_histogram(aes(x = turbidity), fill = "blue", alpha = .5) +
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
  
```{r}
# TODO fix this 

# Make whatever plot is appropriate 
# maybe 2 plots is appropriate
cleaner_rst_environmental %>% 
  group_by(date = as.Date(sample_date)) %>%
  mutate(avg_turbidity_ntu = mean(turbidity)) %>%
  filter(avg_turbidity_ntu < 100) %>%
  ungroup() %>%
 mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_turbidity_ntu, color = year)) + 
  geom_point(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none") + 
  labs(title = "Daily Turbidity Measures (colored by year)",
       x = "Date", 
       y = "Average Daily Turbidity NTUs")  
```
```{r}
# TODO fix plot
cleaner_rst_environmental %>% 
  filter(year(as.Date(sample_date)) > 1999) %>% # Filter because only a few measure before this date
  mutate(year = as.factor(year(as.Date(sample_date)))) %>%
  ggplot(aes(x = turbidity, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Water Turbidity measures summarized by year",
       x = "Turbidity NTUs") + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

Most turbidity measures are low but a few outliers going up until 850

**Numeric Summary of turbidity over Period of Record**
  
```{r}
summary(cleaner_rst_environmental$turbidity)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_environmental$turbidity))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `turbidity` column are NA.


## Explore Categorical variables: {.tabset}

```{r}
cleaner_rst_environmental %>% select_if(is.character) %>% colnames
```


### Variable: `sample_id`

The calendar year Julian date and year code for that ~24-h sample period (ddd_yy)	

```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$sample_id))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `sample_id` column are NA. 


### Variable: `weather_code`

A code for the weather conditions on the SampleDate and SampleTime. See VariableCodesLookUp table							


```{r}
table(cleaner_rst_environmental$weather_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$weather_code))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `weather_code` column are NA.

### Variable: `lunar_phase`

Stage of the Moon on the SampleDate and SampleTime, from: http://aa.usno.navy.mil/data/docs/MoonFraction.php. See VariableCodesLookUp table	

```{r}
table(cleaner_rst_environmental$lunar_phase) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$lunar_phase))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `lunar_phase` column are NA.

### Variable: `trap_sample_type`

The type of sample regime, see VariableCodesLookUp table							

```{r}
table(cleaner_rst_environmental$trap_sample_type) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$trap_sample_type))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `trap_sample_type` column are NA.

### Variable: `habitat`

The type of flow habitat the trap fished in, see VariableCodesLookUp table							

```{r}
table(cleaner_rst_environmental$habitat) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$habitat))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `habitat` column are NA.

### Variable: `thalweg`

Was trap fishing in the thalweg at SampleDate and SampleTime							

```{r}
table(cleaner_rst_environmental$thalweg) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$thalweg))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `thalweg` column are NA.

### Variable: `diel`

The time of day relative to the sun, see VariableCodesLookUp table							

```{r}
table(cleaner_rst_environmental$diel) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$diel))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `diel` column are NA.

### Variable: `debris_type`

The type of debris found in the live-box, see VariableCodesLookUp table							
```{r}
table(cleaner_rst_environmental$debris_type) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$debris_type))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `debris_type` column are NA.

### Variable: `fish_properly`

Was there a problem with the trap at SampleDate and SampleTime							

```{r}
table(cleaner_rst_environmental$fish_properly) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$fish_properly))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `fish_properly` column are NA.

### Variable: `sub_week`

If sample week has more than one efficiency, which part of week is sample from		

```{r}
table(cleaner_rst_environmental$sub_week) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$sub_week))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `sub_week` column are NA.

### Variable: `trap_comments`

```{r}
table(cleaner_rst_environmental$trap_comments) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$trap_comments))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `trap_comments` column are NA.

### Variable: `gear_condition_code`

```{r}
table(cleaner_rst_environmental$gear_condition_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$gear_condition_code))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `gear_condition_code` column are NA.

### Variable: `trap_fishing`

```{r}
table(cleaner_rst_environmental$trap_fishing) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$trap_fishing))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `trap_fishing` column are NA.

### Variable: `partial_sample`

```{r}
table(cleaner_rst_environmental$partial_sample) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_environmental$partial_sample))/nrow(cleaner_rst_environmental), 3) * 100` % of values in the `partial_sample` column are NA.



## Summary of identified issues

* List things that are funcky/bothering us but that we don't feel like should be changed without more investigation

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```

