---
title: "Battle Creek RST QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(knitr)
palette <- wesanderson::wes_palette(name = "Moonrise2")
```

# Battle Creek Rotary Screw Trap Data

## Description of Monitoring Data

These data were collected by the U.S. Fish and Wildlife Service, Red Bluff Fish and Wildlife Office, Battle Creek Monitoring Program. These data represent brood years (BY) 2003-2020. The fish were captured in the upper Battle Creek rotary screw trap site (UBC), Shasta County, California, from September 30, 2003, through June 30, 2021. The catch data were collected using a 5-ft diameter rotary screw trap located at river mile 6.2. The data represent both the raw catch and processed catch values. The trap was only fished through March 25, 2020, during BY2019 sampling.									

**Timeframe:** 2003 - 2021

**Screw Trap Season:** September - June

**Completeness of Record throughout timeframe:** Sample Year tab on excel sheet describes start and end date for trap each year. Sampled every year from 1998 - 2019, some years not fished on weekends or during high flow events. 

**Sampling Location:** Upper Battle Creek (UBC)

**Data Contact:** [Mike Schraml](mailto:mike_schraml@fws.gov)


## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "rst/battle-creek/data-raw/UBC Spring.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_battle_rst_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
sheets <- excel_sheets("raw_battle_rst_data.xlsx")
sheets
raw_rst_count_data <- read_excel("raw_battle_rst_data.xlsx", sheet = "UBC Catch Data") %>% glimpse()
```

## Data transformations

```{r}
cleaner_rst_count <- raw_rst_count_data %>%
  janitor::clean_names() %>%
  rename("date" = sample_date,
         "fork_length" = fork_length,
         "lifestage" = life_stage,
         "count_2" = count,
         "count" = r_catch,
         "interpolated" = interp,
         "run" = fws_race) %>%
  mutate(date = as.Date(date)) %>%
  select(-organism_code, -station_code, -brood_year, 
         -id_week, -race,
         -count_2) %>% # raw catch data, I kept interpolated catch (only 1% of the data is interpolated)
  glimpse()

sum(cleaner_rst_count$interpolated == "YES")/nrow(cleaner_rst_count) * 100 # % of the catch data is interpolated
```
 
Notes: 

* `r_catch` value has an interpolated catch for times the trap did not fish. Not an estimate just for missed days interpolates values based on prior/future days

* I selected only one race, the data has two races, I do not know which one is best to keep, here are details on the two: 

  * fWSRace - USFWS run designation base upon location or emergence timing used in reports and for passage indices, W=winter-run, S=spring-run, F=fall-run, L=late-fall run Chinook Salmon					
  * race - Database generated Sheila Greene run designation of catch, W=winter-run, S=spring-run, F=fall-run, L=late-fall run Chinook Salmon, see RunDesignation and RunChart tables				
  
These columns seem to have very different run designations. 
```{r}
sum(raw_rst_count_data$Race == raw_rst_count_data$FWSRace, na.rm = T)/length(raw_rst_count_data$Race)
```

  
Additional info from Mike describing the race columns: 
We use the FWSRace for our reports, all fish length-designated as fall-run by the Sheila Greene length-at-date (LAD) charts we consider spring-run at the upper Battle and Clear Creek sites.  We consider the barrier weir at Coleman National Fish Hatchery to be fish tight up to 800 cfs, flows rarely exceed that level during the spring-run and early fall-run escapement period. We are fairly confident that no fall-run get above the weir. Likewise, we install a separation (picket) weir in Clear Creek below the upper trap site that excludes fall-run from spawning with spring-run and superimposing redds on top of spring-run redds. At the lower Clear Creek (LCC) site we strictly go by the Sheila Greene LAD chart because there is no way to tell the difference between spring-run and fall-run fish in the field.  There is overlap in the sizes of spring-run and fall-run and we know that the Sheila Greene LAD charts don't capture the overlap. In Clear Creek, some fish that classify as fall-run based on length from our LCC trap are actually spring-run and vise versa.

FWSRace incorporates all the fine-scale decisions about when to use the Sheila Greene LAD chart. We use FWSRace to determine run, especially because this is used to determine passage estimates.

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- cleaner_rst_count %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_rst_count),
                          description = c("Sample date - end date of approximately 24 hour sampling period",
                                          "The calendar year Julian date and year code for that ~24-h sample period (ddd_yy)",
                                          "Run designation using the FWSRace variable",
                                          "Fork length in millimeters",
                                          "Life stage of the catch (yolk-sac fry, fry, parr, silvery parr, smolt)",
                                          "Number of fish caught in trap",
                                          "Indicates mortality (T/F)",
                                          "Indicates interpolated catch for times the trap did not fish (T/F)"),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
```

 
## Explore Numeric Variables: {.tabset}

```{r}
cleaner_rst_count %>% select_if(is.numeric) %>% colnames()
```

### Variable: `fork_length`

Fork length of the fish captured, is zero if the fish was not measured							

**Plotting fork_length**
  
```{r}
cleaner_rst_count %>% filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = fork_length)) + 
  geom_histogram(breaks=seq(0, 200, by=2)) + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  theme_minimal() +
  labs(title = "Fork length distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
cleaner_rst_count %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = fork_length, y = lifestage)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Fork length summarized by lifestage") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of fork_length over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_rst_count$fork_length)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_count$fork_length))/nrow(cleaner_rst_count), 3) * 100` % of values in the `fork_length` column are NA. 

* 0 for `fork_length` seems like an NA or error
```{r}
cleaner_rst_count %>% filter(fork_length == 0 & !is.na(count))

cleaner_rst_count$fork_length <- ifelse(cleaner_rst_count$fork_length == 0, NA, cleaner_rst_count$fork_length)
```


### Variable: `count`

Catch number used to generate passage indices for reports, plus counts are split into races, zero fork lengths have been assigned 

*  Definition given in Spreadsheet metadata, do not know what 0 fork lengths have been assigned means (i know if fish is not measured they give it a fork length of 0 so maybe they are assigning these 0 values lifestages or counts on interpolated days) TODO ask Mike

**Plotting fish counts over period of record**
```{r, include=FALSE}
sac_indices <- waterYearType::water_year_indices %>% 
    filter(location == "Sacramento Valley") %>% 
    transmute(water_year = WY, year_type = as.character(Yr_type))
```

Total daily passage for spring run

```{r}
filter(cleaner_rst_count, run == "S") %>% 
  group_by(date) %>%
  summarise(total_daily_catch = sum(count)) %>%
  filter(year(date) > 2014, year(date) < 2021) %>%
  mutate(water_year = ifelse(month(date) %in% 10:12, year(date) + 1, year(date))) %>% 
  left_join(sac_indices) %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  filter(water_year < 2021) %>%
  ggplot(aes(x = fake_date, y = total_daily_catch, fill = year_type)) + 
  geom_col() + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1900-10-01"), as.Date("1901-06-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Total Spring Run Daily Raw Passage 2015 - 2020",
       y = "Total daily raw catch",
       x = "Date")+ 
  facet_wrap(~water_year, scales = "free") +
  scale_color_manual(palette)
```

Total catch by year for spring run

```{r}
cleaner_rst_count  %>%
  filter(year(date) < 2021, 
         run == "S"
         ) %>% 
  mutate(year = as.factor(year(date))) %>%
  group_by(year, run) %>%
  summarise(total_yearly_catch = sum(count)) %>%
  ggplot(aes(x = year, y = total_yearly_catch)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total fish counted each year for spring run",
       y = "Total fish raw catch") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap(~run, scales = "free_y")
```

**Numeric Summary of counts over Period of Record**
  
  Numerical summary includes all runs
  
```{r}
# Table with summary statistics
summary(cleaner_rst_count$count)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_rst_count$count))/nrow(cleaner_rst_count), 3) * 100` % of values in the `count` column are NA.


## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
cleaner_rst_count %>% select_if(is.character) %>% colnames()
```


### Variable: `sample_id`

The calendar year Julian date and year code for that ~24-h sample period (ddd_yy)							

```{r}
length(unique(cleaner_rst_count$sample_id)) 
```

There are `r length(unique(cleaner_rst_count$sample_id))`  unique sample ID

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_count$sample_id))/nrow(cleaner_rst_count), 3) * 100` % of values in the `sample_id` column are NA.

### Variable: `run`
```{r}
table(cleaner_rst_count$run)
```


Fix inconsistencies with spelling, capitalization, and abbreviations. "N/P" is changed to NA in the case_when statement below.

```{r}
cleaner_rst_count$run <- case_when(cleaner_rst_count$run == "F" ~ "fall", 
                                   cleaner_rst_count$run == "S" ~ "spring",
                                   cleaner_rst_count$run == "L" ~ "late fall",
                                   cleaner_rst_count$run == "W" ~ "winter")
table(cleaner_rst_count$run)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_count$run))/nrow(cleaner_rst_count), 3) * 100` % of values in the `run` column are NA.


### Variable: `lifestage`

Life stage of the catch (CHN = C0 - yolk-sac fry, C1 - fry, C2 - parr, C3 - silvery parr, C4 - smolt, n/p - Not provided; RBT = R1 - yolk-sac fry, R2 - fry, R3 - parr, R4 - silvery parr, R5 - smolt, R6 - adult). "N/P" is changed to NA in the case_when statement below.							


```{r}
table(cleaner_rst_count$lifestage)
```


Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_count$lifestage <- case_when(cleaner_rst_count$lifestage == "C0" ~ "yolk-sac fry", 
                                   cleaner_rst_count$lifestage == "C1" ~ "fry",
                                   cleaner_rst_count$lifestage == "C2" ~ "parr",
                                   cleaner_rst_count$lifestage == "C3" ~ "silvery parr",
                                   cleaner_rst_count$lifestage == "C4" ~ "smolt")
table(cleaner_rst_count$lifestage)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_count$lifestage))/nrow(cleaner_rst_count), 3) * 100` % of values in the `lifestage` column are NA.

### Variable: `dead`

Indicates if the fish was a mortality (YES = mortality)							
							


```{r}
table(cleaner_rst_count$dead)
```


Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_count$dead <- case_when(cleaner_rst_count$dead == "YES" ~ TRUE, 
                                   cleaner_rst_count$dead == "N" |  cleaner_rst_count$dead == "NO" ~ FALSE)
table(cleaner_rst_count$dead)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_count$dead))/nrow(cleaner_rst_count), 3) * 100` % of values in the `dead` column are NA.

### Variable: `interpolated`

Is count value an interpolated catch for times the trap did not fish? Yes = the data is interpolated data and not actual catch data		

```{r}
table(cleaner_rst_count$interpolated)
```


Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_rst_count$interpolated <- case_when(cleaner_rst_count$interpolated == "YES" ~ TRUE, 
                                            cleaner_rst_count$interpolated == "NO" ~ FALSE)
table(cleaner_rst_count$interpolated)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_rst_count$interpolated))/nrow(cleaner_rst_count), 3) * 100` % of values in the `interpolated` column are NA.

## Summary of identified issues

* Count values are interpolated on days where the traps were not fished, to see if interpolated or not refer to `interpolated` column. Recommend providing more information about how interpolation is calculated.
* Investigate what it means for: zero fork lengths have been assigned - 0 `fork_length` values mean fish was not measured. 

## Next steps

### Columns to remove

- `sample_id` may not be needed.

## Save cleaned data back to google cloud 

```{r}
battle_rst_catch <- cleaner_rst_count %>% glimpse()
```

```{r, eval=FALSE}
# Write to google cloud 
# Name file [watershed]_[data type].csv
f <- function(input, output) write_csv(input, file = output)
gcs_upload(battle_rst_catch,
           object_function = f,
           type = "csv",
           name = "rst/battle-creek/data/battle_rst_catch.csv")
```