---
title: "Butte Creek Snorkel Survey 2002 QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
```

# Butte Creek Adult Snorkel Survey: 2002 Holding Data 

## Description of Monitoring Data
Butte Creek snorkel holding data was shared by Claire Bryant. This data was shared in multi tab spreadsheets. 

**Timeframe:** 2002 

**Snorkel Season:** Snorkel Survey is conducted in July or August 

**Completeness of Record throughout timeframe:** 

**Sampling Location:** Butte Creek 

**Data Contact:** 

Claire Bryant 

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
getwd() #to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/butte-creek/data-raw/Butte 2002 Snorkel Modified.xls",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ButteSnorkel2002.xls",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data: 

Butte creek data needs to be transformed before it can be easy reviewed and used. Currently each sheet describes snorkeling for a different site. 
```{r}
butte_snorkel <- readxl::read_excel("ButteSnorkel2002.xls") %>% View()
```
## Create function that transforms each sheet 

```{r}
file_names = c("ButteSnorkel2002.xls", "ButteSnorkel2002.xls", "ButteSnorkel2002.xls")
sheet_names = c(1, 2, 3)
tidy_up_snorkel_data <- function(file_name, sheet_name){
  metadata <- readxl::read_excel(file_name, sheet = sheet_name, range = "A3:J7")
  date_surveyed <- janitor::excel_numeric_to_date(as.numeric(metadata[2,2]))
  raw_data <- readxl::read_excel(file_name, sheet = sheet_name, range = "A13:M70", 
                                 col_names = c("Reach", "Clint Number", "Clint Condition",
                                               "Curtis Number", "Curtis Condition",
                                               "Mike Number", "Mike Condition",
                                               "Adam Number", "Adam Condition", 
                                               "Avg", "Low", "High", "Comments"))
  
  combined_data <- tibble()
  names <- c("Clint", "Curtis", "Mike", "Adam")
  for (i in 1:length(names)) {
    print(i)
    print(names[i])
    cols <- colnames(raw_data)[stringr::str_detect(colnames(raw_data), names[i])]
    i_dat <- raw_data %>% 
      select("reach" = Reach, "fish_count" = cols[1], "why_fish_count_na" = cols[2]) %>%
      mutate(personnel = names[i])
    combined_data <- bind_rows(combined_data, i_dat)
  }
  
  transformed_data <- combined_data %>%
    filter(reach != "TOTAL", reach != "TOTALS", reach != "Total", reach != "RANGE", reach != "Range") %>%
    mutate(why_fish_count_na = case_when(why_fish_count_na == "NGC" ~ "not a good count",
                                why_fish_count_na == "DNS" ~ "did not see", 
                                why_fish_count_na == "DNSw"| why_fish_count_na == "DNSW" ~ "did not swim",
                                why_fish_count_na == "DNC" ~ "did not count", 
                                why_fish_count_na == "FRAB" ~ "from above"),
           date = date_surveyed) %>%
    select(date, reach, personnel, fish_count, why_fish_count_na)
  return(transformed_data)
}
cleaner_data <- bind_rows(purrr::map2(file_names, sheet_names, tidy_up_snorkel_data)) %>% glimpse

```

## Explore Date 

Survey completed in 3 days in August. 
```{r}
unique(cleaner_data$date)
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_data %>% 
  select_if(is.numeric) %>% colnames()
```

### Variable: `fish_count`

**Plotting fish_count by reach**

```{r}
cleaner_data %>% 
  ggplot(aes(x = reach, y = fish_count, color = personnel)) + 
  geom_point(size = 3) + 
  theme_minimal() + 
  labs(x = "Reach", y = "Fish Counts", title = "Fish Counts Per Reach") +
  theme(axis.text.x=element_text(angle=90, hjust=1))
```

**Numeric Summary of fish_count over Period of Record**

```{r}
summary(cleaner_data$fish_count)
```

**NA and Unknown Values**

*  `r round(sum(is.na(cleaner_data$fish_count))/nrow(cleaner_data), 3)*100` % of values in the `fish_count` column are NA.


## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
cleaner_data %>% 
  select_if(is.character) %>% colnames()
```


### Variable: `reach`
```{r}
table(cleaner_data$reach)
```

They do not appear to do the same number of snorkels in each reach. 

**NA and Unknown Values**

*  `r round(sum(is.na(cleaner_data$reach))/nrow(cleaner_data), 3)*100` % of values in the `reach` column are NA.


### Variable: `personnel`
```{r}
table(cleaner_data$personnel)
```

**NA and Unknown Values**

*  `r round(sum(is.na(cleaner_data$personnel))/nrow(cleaner_data), 3)*100` % of values in the `personnel` column are NA.

### Variable: `why_fish_count_na`
```{r}
table(cleaner_data$why_fish_count_na)
```


**NA and Unknown Values**

*  `r round(sum(is.na(cleaner_data$why_fish_count_na))/nrow(cleaner_data), 3)*100` % of values in the `why_fish_count_na` column are NA.


## Summary of identified issues

* Each year is different formatted. Need to create a new markdown for wrangling. 
* Lots of data points where there was not a good count of fish 

## Save cleaned data back to google cloud 

```{r}
butte_holding_2002 <- cleaner_data %>% glimpse
```

```{r, eval= FALSE}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(butte_holding_2002,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/butte-creek/data/butte_holding_2002.csv")


```
