---
title: "Butte Creek Snorkel Survey QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
```

# Butte Creek Adult Snorkel Survey: Holding Data 

## Description of Monitoring Data
Butte Creek snorkel holding data was shared by Claire Bryant. This data was shared in multi tab spreadsheets. 

**Timeframe:** 2001 - 2019 

**Snorkel Season:** Snorkel Survey is conducted in July or August 

**Completeness of Record throughout timeframe:** 

**Sampling Location:** Butte Creek 

**Data Contact:** 

Claire Bryant 

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
getwd() #to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/butte-creek/data-raw/Butte 2001 Snorkel Modified.xls",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ButteSnorkel2021.xls",
               overwrite = TRUE)

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
butte_snorkel <- readxl::read_excel("ButteSnorkel2021.xls") %>% glimpse()
```

## Data transformations

```{r}
# For different excel sheets for each year read in and combine years here
```

```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
```

### Variable: `[name]`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
# Filter clean data to show only categorical variables
```


### Variable: `[name]`
```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Summary of identified issues

* List things that are funcky/bothering us but that we don't feel like should be changed without more investigation

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```