---
title: "Feather River CWT - Broodstock QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River CWT - Broodstock Data

## Description of Monitoring Data 
This dataset includes the CWT data collected so far at the Feather River Hatchery for the 2021 season. 

**Timeframe:** 2021 (partial)

**Data Contact:** [Lea Koerber](mailto:Lea.Koerber@wildlife.ca.gov) 

**Metadata Shared by Lea**

| Sex  |	RMIS Run  |	Condition  |
| ------ | ---------- | ----------- |
| 1 = Male |	1 = Spring |	1 = Killed NOT Spawned |
| 2 = Female |	3 = Fall |	2 = Dead in Pond |
| 9 = Unknown |	7 = Late-fall |	3 = Spawned |


*Additional Note*: Following up with OSP people regarding data for full period of record. 

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
read_from_cloud <- function(year){
  gcs_get_object(object_name = paste0("adult-upstream-passage-monitoring/feather-river/CWT/data-raw/2021-09-", year, "_FRFH_tblOutput.xlsx"),
               bucket = gcs_get_global_bucket(),
               saveToDisk = paste0("raw_feather_cwt_", year, ".xlsx"),
               overwrite = TRUE)
  data <- read_excel(paste0("raw_feather_cwt_", year, ".xlsx"))
}
years <- c(14, 15, 17, 20, 21, 22)
raw_data <- purrr::map(years, read_from_cloud) %>%
  reduce(bind_rows)

write_csv(raw_data, "raw_cwt_data.csv")

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
raw_cwt_data <- read_csv("raw_cwt_data.csv") %>% glimpse
```

## Data transformations

```{r}

```

```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables (this way we know we do not miss any)
```

### Variable: `[name]`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2 plots is appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Explore Categorical variables: {.tabset}

General notes: If there is an opertunity to turn yes no into boolean do so, but not if you loose value 


```{r}
# Filter clean data to show only categorical variables (this way we know we do not miss any)
```


### Variable: `[name]`
```{r}
# table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistancies with catagorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

### Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```
