---
title: "Feather River Hallprint Data QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River hallprint adult broodstock selection and enumeration data (Tagging Data)

## Description of Monitoring Data

Hallprint tagging data only describes fish marked, giving hallprint tag numbers and dates tagged for each fish. It does not include any other sampling of fish like size or genetics.This dataset does not contain any additional information about fish returns. We currently have one file describing fish returns. We can likely acquire more but there are issues with the return data. 

**Timeframe:** 2004 - 2020 
  
**Season:** Tagging typically begins in April and goes through June

**Provisional or QC data** Data is QC  
  
**Sampling Location:** Some tags have a acoustic location attribute. All fish are tagged on the Feather River. 
  
**Data Contact:** [Byron Mache](mailto:Byron.Mache@water.ca.gov) 
  
Note fron Byron: Each fish is given two tags. The `second_tag_number` is NA if the first tag and the second tag have the same number. 
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
gcs_get_object(object_name = 
                 "adult-upstream-passage-monitoring/feather-river/hallprint/data-raw/all_hallprints_all_years.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_feather_hallprint_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
raw_hallprint_data <- read_excel("raw_feather_hallprint_data.xlsx", 
                                 col_types = c("text", "date", "text",
                                               "text", "text", "text",
                                               "text")) %>% glimpse()
```

## Data transformations

```{r}
cleaner_hallprint_data <- raw_hallprint_data %>% 
  rename("date" = Date, "tag_number" = `Tag#`, "second_tag_number" = `2ndTag#`, 
         "color" = Color, "acoustic" = Acoustic, 
         "acoustic_location" = `Acoustic Location`) %>%
  mutate(date = as.Date(date)) %>%
  glimpse() 
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data. 

```{r data_dictionary}
percent_na <- cleaner_hallprint_data %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_hallprint_data),
                          description = c("IDs",
                                          "Tag date",
                                          "Tag number",
                                          "Second tag number",
                                          "color of the tag",
                                          "Acoustic",
                                          "Fish were marked with Hallprint tags and acoustic tags"),
                          percent_na = round(percent_na$value*100)
                          
)
knitr::kable(data_dictionary)
```
## Explore Numeric Variables: {.tabset}

No numeric variables. 
```{r}
# Filter clean data to show only numeric variables 
cleaner_hallprint_data %>% select_if(is.numeric) %>% colnames()
```

## Explore Categorical variables: {.tabset}

```{r}
# Filter clean data to show only categorical variables
cleaner_hallprint_data %>% select_if(is.character) %>% colnames()
```

### Variable: `ID`

All IDs are unique as anticipated:
```{r}
length(unique(cleaner_hallprint_data$ID)) == nrow(cleaner_hallprint_data)
```

**NA and Unknown Values**
```{r}
cleaner_hallprint_data %>% filter(is.na(ID))
```
There are no NA values in `ID`. 

### Variable: `tag_number`

Are all the tags unique? No. In more recent years tag numbera start to be unique and we have less NA values but there are 8 years where tags are not unique. 
```{r}
cleaner_hallprint_data %>% 
  group_by(year = year(date)) %>%
  summarize(count_unique_tags = length(unique(tag_number)), 
            number_tags = n(), 
            na_tags = sum(is.na(tag_number)),
            all_tags_na_or_unique = ifelse(count_unique_tags == number_tags + na_tags, TRUE, FALSE))
```

```{r}
cleaner_hallprint_data %>% 
  group_by(year = year(date)) %>%
  summarise(total_tags_per_year = sum(!is.na(tag_number))) %>% 
  ggplot(aes(x = year, y = total_tags_per_year)) +
  geom_col() + 
  theme_minimal() + 
  labs(title = "Total fish tagged by year",
       y = "Number of Fish Tagged")  
```

More fish were tagged in 2013 than any other year. 
Numbers of fish tagged are not even over the period of record. 

```{r}
cleaner_hallprint_data %>% 
  group_by(year = year(date), month = month(date)) %>%
  summarise(total_tags_per_month = sum(!is.na(tag_number))) %>% 
  ggplot(aes(x = month, y = total_tags_per_month)) +
  geom_col() + 
  facet_wrap(~year) +  
  theme_minimal() + 
  labs(title = "Tags by month",
       y = "Number of Fish Tagged")  
```

Generally tagging occurs in May and June. 

**NA and Unknown Values**
```{r}
nrow(subset(cleaner_hallprint_data, is.na(cleaner_hallprint_data$tag_number)))
```
 * `r round(sum(is.na(cleaner_hallprint_data$tag_number))/nrow(cleaner_hallprint_data), 3) * 100` % of values in the `tag_number` column are NA.

### Variable: `second_tag_number`

Similarly not all unique. Only one year where count of unique tags equals total number of unique tags. 
```{r}
cleaner_hallprint_data %>% 
  group_by(year = year(date)) %>%
  summarize(count_unique_second_tags = length(unique(second_tag_number)), 
            number_second_tags = n(), 
            na_tags = sum(is.na(second_tag_number)),
            all_tags_na_or_unique = ifelse(count_unique_second_tags == number_second_tags + na_tags, TRUE, FALSE))
```

 
**NA and Unknown Values**

 * `r round(sum(is.na(cleaner_hallprint_data$second_tag_number))/nrow(cleaner_hallprint_data), 3)* 100` % of values in the `second_tag_number` column are NA. Byron shared that if the `tag_number` equals the `second_tag_number` they leave `second_tag_number` as NA. 
 
```{r}
# Replace 2nd tag number with first tag number if second tag number is NA
cleaner_hallprint_data$second_tag_number <- if_else(is.na(cleaner_hallprint_data$second_tag_number), 
                                                    cleaner_hallprint_data$tag_number, 
                                                    cleaner_hallprint_data$second_tag_number)
```


### Variable: `color`

Color refers to the color of the tag. Most are actual colors, Hole Punches are used in the place of tags when they run out. (Hole Punch fin as form of tagging)
```{r}
table(cleaner_hallprint_data$color) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
unique(cleaner_hallprint_data$color)
```


```{r}
# Fix any inconsistencies with categorical variables (changes made based on correspondence with Byron)
cleaner_hallprint_data$color <- case_when(
  cleaner_hallprint_data$color == "G" ~ "green",
  cleaner_hallprint_data$color %in% c("R?", "R") ~ "red",
  cleaner_hallprint_data$color %in% c("Y", "YD", "YH") ~ "yellow",
  cleaner_hallprint_data$color == "Hole Punch" ~ "hole punch",
  cleaner_hallprint_data$color == "P" ~ "pink",
  cleaner_hallprint_data$color == "?" ~ "unknown",
  TRUE ~ 'uhoh'
)
table(cleaner_hallprint_data$color)
```

**NA and Unknown Values**

 * `r round(sum(cleaner_hallprint_data$color == "unknown")/nrow(cleaner_hallprint_data), 5)* 100` % of values in the `color` column are Unknown. 
  
### Variable: `acoustic`
```{r}
table(cleaner_hallprint_data$acoustic) 

# Make into a booleen 
cleaner_hallprint_data$acoustic <- case_when(
  is.na(cleaner_hallprint_data$acoustic) ~ NA,
  cleaner_hallprint_data$acoustic == "Y" ~ TRUE,
  cleaner_hallprint_data$acoustic == "N" ~ FALSE,
)

table(cleaner_hallprint_data$acoustic) 
```

**NA and Unknown Values**

  * `r round(sum(is.na(cleaner_hallprint_data$acoustic))/nrow(cleaner_hallprint_data), 3) * 100` % of values in the `acoustic` column are NA. 
  
### Variable: `acoustic_location`

This fish were marked with Hallprint tags and acoustic tags in the lower river.  These fish would be spawned as spring-run broodstock in the hatchery if they were to return in late September.
```{r}
table(cleaner_hallprint_data$acoustic_location) 
```

**Create lookup rda for acoustic_location encoding:** 
```{r}
feather_hallprint_acoustic_location <- c("SRA", "FRFH")
names(feather_hallprint_acoustic_location) <- c("Spring Run Angling Project", "Feather River Fish Hatchery")

write_rds(feather_hallprint_acoustic_location, "../../../data/feather_hallprint_acoustic_location.rds")
```

**NA and Unknown Values**

  * `r round(sum(is.na(cleaner_hallprint_data$acoustic_location))/nrow(cleaner_hallprint_data), 3) * 100` % of values in the `acoustic_location` column are NA. 
  
**Summary of identified issues:**

* One identified issue is that not all `tags_numbers` or `second_tag_numbers` are unique. 
* Each tag is a unique fish, this data may be more helpful as some estimate of fish count, add in a count column?

### Next Steps

* Figure out how to use this dataset for adult upstream passage estimate. The hallprint data is quite different from other upstream passage data - need to figure out which excel sheet and which column is the most useful.

### Save cleaned data back to google cloud 

```{r}
feather_hallprint <- cleaner_hallprint_data %>% glimpse()
```

```{r, eval = FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_hallprint,
           object_function = f,
           type = "csv",
           name = "adult-upstream-passage-monitoring/feather-river/data/feather_hallprint.csv")
```
