---
title: "Feather River Hallprint Data QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River hallprint adult broodstock selection and enumeration data (Return/Carcass Data)

## Description of Monitoring Data

We currently only have one file describing fish returns. We can likely acquire more but the return data is currently very messy and not stored in a consistent format across years. 

The excel workbook that we have has 15 sheets. Most sheets are summaries of hallprint tagging data. This markdown is focused on the `Carcass Recoveries` tab of the workbook. 

It looks like there is additional information on Hallprint returns/tag numbers in the feather river carcass suvey.

**Timeframe:** 2018
  
**Season:** Carcass Recoveries appear to be found from September - December
  
**Completeness of Record throughout timeframe:** Most records are from weeks 4 - 10 however we have records from week 2 - 10 and weeks 12 - 14. 

**Sampling Location:** Locations described by Section, will need more information to map each section to a specific location
   
**Data Contact:** [Byron Mache](mailto:Byron.Mache@water.ca.gov) 
  
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "adult-upstream-passage-monitoring/feather-river/hallprint/data-raw/returns/FRFH HP RUN DATA 2018 as of 9-10-20.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_feather_hallprint_returns_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
sheets <- excel_sheets("raw_feather_hallprint_returns_data.xlsx")
sheets # many sheets, most do not seem useful 
raw_carcass_data <- read_excel("raw_feather_hallprint_returns_data.xlsx", 
                               sheet = "Carcass Recoveries",
                               range = "A1:V209") %>% # additonal columns are summaries
  glimpse()
```

## Data transformations

```{r}
# For different excel sheets for each year read in and combine years here
cleaner_carcass_data <- raw_carcass_data %>% 
  set_names(tolower(colnames(raw_carcass_data))) %>% 
  rename("individual_id" = individualid, 
         "survey_id" = surveyid, 
         "distinct_tag_applied" = disctagapplied, 
         "spawn_status" = spawnstatus, 
         "fork_length" = flmm, 
         "adipose_fin_clip" = adfinclip, 
         "head_number" = headnumber, 
         "scale_number" = scalenumber, 
         "otolith_number" = otolithnumber) %>% 
  mutate(date = as.Date(date),
         survey_id = as.character(survey_id),
         individual_id = as.character(individual_id),
         otolith_number = as.numeric(otolith_number)) %>%
  filter(species == "CHN") %>%
  select(-flcm, -species, -tissuenumber, -colortagapplied, -othermarks) %>% # we already have flcm, tissuenumber and colortagapplied are all NA, species are all chinook 
  glimpse()
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data. 

```{r data_dictionary}
percent_na <- cleaner_carcass_data %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_carcass_data),
                          description = c("Week",
                                          "Date",
                                          "Each section number describes a different portion of the river",
                                          "Each survey ID is a unique ID for each survey day",
                                          "Each Individual ID describes a tagged fish",
                                          "All values are spring run",
                                          "Is the fish chopped or tagged",
                                          "A number describing the distinct tag applied to fish",
                                          "Sex of the fish",
                                          "Spawn status of the fish",
                                          "Fork length of the fish",
                                          "TODO: figure out what condition means",
                                          "Adipose fin clip",
                                          "Number corresponding to head tags",
                                          "Number corresponding to scale tags",
                                          "Number corresponding to otolith",
                                          "Comment describing if a fish is spring green or not"),
                          percent_na = round(percent_na$value*100)
                          
)
knitr::kable(data_dictionary)
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
cleaner_carcass_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `week`

```{r}
cleaner_carcass_data %>%
  ggplot(aes(x = week)) + 
  geom_histogram(breaks=seq(0, 15, by=1)) + 
  scale_x_continuous(breaks=seq(0, 15, by=1)) +
  theme_minimal() +
  labs(title = "Week Distribution") + 
  theme(text = element_text(size = 18)) 
```

Numeric Summary: 

```{r}
summary(cleaner_carcass_data$week)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$week))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `week` column are NA. 

### Variable: `fork_length`

Length of fish in mm							

**Plotting fork_length**
  
```{r}
cleaner_carcass_data %>%
  ggplot(aes(x = fork_length)) + 
  geom_histogram() + 
  # scale_x_continuous(breaks=seq(50, 100, by = 5)) +
  theme_minimal() +
  labs(title = "fork length distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of fork_length over Period of Record**
  
```{r}
# Table with summary statistics
summary(cleaner_carcass_data$fork_length)
```

**NA and Unknown Values**
  
* `r round(sum(is.na(cleaner_carcass_data$fork_length))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `fork_length` column are NA. 



### Variable: `otolith_number`

```{r}
cleaner_carcass_data %>%
  ggplot(aes(x = otolith_number)) + 
  geom_histogram(breaks=seq(0, 225, by=5)) + 
  scale_x_continuous(breaks=seq(0, 225, by=25)) +
  theme_minimal() +
  labs(title = "Otolith Number Distribution") + 
  theme(text = element_text(size = 18)) 
```

Numeric Summary: 

```{r}
summary(cleaner_carcass_data$otolith_number)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$otolith_number))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `otolith_number` column are NA.


## Explore Categorical variables: {.tabset}

```{r}
# Filter clean data to show only categorical variables
cleaner_carcass_data %>% select_if(is.character) %>% colnames()
```

### Variable: `section`

Each section number describes a different portion of the river

```{r}
table(cleaner_carcass_data$section)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$section))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `section` column are NA.

### Variable: `survey_id`

Each survey id is a unique id for each survey day

There are `r length(unique(cleaner_carcass_data$survey_id))` unique individual ids. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$survey_id))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `survey_id` column are NA.

### Variable: `individual_id`

Each Individual ID is an ID for the tagged fish

There are `r length(unique(cleaner_carcass_data$individual_id))` unique individual ids. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$individual_id))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `individual_id` column are NA.

### Variable: `run`

```{r}
table(cleaner_carcass_data$run)
```
All values are spring run. 

```{r}
cleaner_carcass_data$run <- tolower(cleaner_carcass_data$run)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$run))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `run` column are NA.

### Variable: `disposition`

```{r}
table(cleaner_carcass_data$disposition)
```

```{r}
cleaner_carcass_data$disposition <- tolower(cleaner_carcass_data$disposition)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$disposition))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `disposition` column are NA.

### Variable: `distinct_tag_applied`

A number describing the distinct tag applied to fish

There are `r length(unique(cleaner_carcass_data$distinct_tag_applied))` distinct tag numbers. 


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$distinct_tag_applied))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `distinct_tag_applied` column are NA.


### Variable: `sex`

```{r}
table(cleaner_carcass_data$sex)
```

All unknown and not recorded (N/R) become NA.

```{r}
cleaner_carcass_data$sex <- case_when(cleaner_carcass_data$sex == "F" ~ "female",
                                      cleaner_carcass_data$sex == "M" ~ "male")

table(cleaner_carcass_data$sex)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$sex))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `sex` column are NA.


### Variable: `spawn_status`

```{r}
table(cleaner_carcass_data$spawn_status)
```

Change spawn status to TRUE if spawned (Y) and NA for everything else. 

```{r}
cleaner_carcass_data$spawn_status <- ifelse(cleaner_carcass_data$spawn_status == "Y", TRUE, NA)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$spawn_status))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `spawn_status` column are NA.

### Variable: `condition`

TODO figure out what these condition values mean, most are NA 

```{r}
table(cleaner_carcass_data$condition)
```

```{r}
cleaner_carcass_data$condition <- tolower(cleaner_carcass_data$condition)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$condition))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `condition` column are NA.

### Variable: `adipose_fin_clip`

```{r}
table(cleaner_carcass_data$adipose_fin_clip)
```

```{r}
cleaner_carcass_data$adipose_fin_clip <- tolower(ifelse(cleaner_carcass_data$adipose_fin_clip %in% c("Not Recorded", "Unknown"), NA,
                                          cleaner_carcass_data$adipose_fin_clip))

table(cleaner_carcass_data$adipose_fin_clip)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$adipose_fin_clip))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `adipose_fin_clip` column are NA.

### Variable: `head_number`

There are `r length(unique(cleaner_carcass_data$head_number))` unique head numbers corresponding to head tags. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$head_number))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `head_number` column are NA.

### Variable: `scale_number`

There are `r length(unique(cleaner_carcass_data$scale_number))` unique scale numbers corresponding to scale tags. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$scale_number))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `scale_number` column are NA.\

### Variable: `comments`

```{r}
unique(cleaner_carcass_data$comments)
```
All comments just describe if a fish is spring green or not. Change to spring green TRUE, FALSE variable
```{r}
cleaner_carcass_data$spring_green <- ifelse(cleaner_carcass_data$comments == "SPRING GREEN", TRUE, FALSE)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$comments))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `comments` column are NA.

**Summary of identified issues:**

* Need to figure out what the condition codes mean (TODO contact byron)
* Only one year of data, the rest we did not receive because they are all in varying formats. This one is also a many sheets non tidy excel table. I choose to look at this sheet because it seemed the most useful but I am not sure it is the best information. 

### Next Steps

* Hallprint returns data has similar columns as some of the carcass data so it might be more fitting to put it under there. Or hallprint returns data and hallprint data could be its own entity due to how different the data are. Also need to figure out which one of the many excel data sheet under hallprint data is the most useful.

### Save cleaned data back to google cloud 

```{r}
feather_hallprint_returns <- cleaner_carcass_data %>% select(-comments) %>% glimpse()
```

```{r, eval = FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_hallprint_returns,
           object_function = f,
           type = "csv",
           name = "adult-upstream-passage-monitoring/feather-river/data/feather_hallprint_returns.csv")
```
