---
title: "Battle Creek Redd Survey QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Battle Creek Redd Survey

## Description of Monitoring Data

These data were aquired via snorkel and kayak surveys on Battle Creek from 2001 to 2019. Red location, size, substrate and flow were measured. Annual monitoring questions and conditions drove the frequency and detail of individual redd measurements. 						


**Timeframe:** 2001 - 2019

**Survey Season:** 

**Completeness of Record throughout timeframe:** Sampled each year

**Sampling Location:** Battle Creek 

**Data Contact:** [Natasha Wingerter](mailto:natasha_wingerter@fws.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
sheets <- excel_sheets("raw_adult_spawn_hold_carcass.xlsx")
sheets 
raw_redd_data <-read_excel("raw_adult_spawn_hold_carcass.xlsx", sheet = "Redd Survey") %>% glimpse()
```

## Data transformations

```{r}
cleaner_redd_data <- raw_redd_data %>% 
  janitor::clean_names() %>% 
  rename("date" = sample_date,
         "fish_garding" = `for`, 
         "redd_measured" = measure, 
         "why_not_measured" = why_not_me,
         "date_measured" = date_measu, 
         "pre_redd_substrate_size" = pre_sub, 
         "redd_substrate_size" = sides_sub, 
         "tail_substrate_size" = tail_sub,
         "pre_redd_depth" = pre_depth, 
         "redd_pit_depth" = pit_depth, 
         "redd_tail_depth" = tail_depth,
         "redd_length_in" = length_in, 
         "redd_width_in" = width_in
         ) %>%
  mutate(date = as.Date(date)) %>%
  select(-project, -year, -date_measured, 
         -species_run) %>% #All are spring run 
  glimpse()
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_redd_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `longitude`, `latitude`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values


### Variable: `river_mile`

**Plotting [Variable] over Period of Record**

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = river_mile, y = year(date))) +
  geom_point(alpha = .75) + 
  theme_minimal()
```

It looks like river miles 0 - 4 and 11 - 12 most commonly have redds. In most recent years almost all the redds are before mile 5. 

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = river_mile)) +
  geom_histogram(alpha = .75) + 
  theme_minimal()
```

**Numeric Summary of river mile over Period of Record**

```{r}
summary(cleaner_redd_data$river_mile)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$river_mile))/nrow(cleaner_redd_data), 3) * 100` % of values in the `river_mile` column are NA. 


### Variable: `pre_redd_depth`

**Plotting distribution of pre redd depth**

```{r}
cleaner_redd_data %>%
  ggplot(aes(x = pre_redd_depth)) +
  geom_histogram() +
  theme_minimal()
```

**Numeric Summary of pre redd depth over Period of Record**

```{r}
summary(cleaner_redd_data$pre_redd_depth)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$pre_redd_depth))/nrow(cleaner_redd_data), 3) * 100` % of values in the `pre_redd_depth` column are NA. 
* There are a lot of 0 values. Could these also be NA?

### Variable: `redd_pit_depth``

**Plotting distribution of redd pit depth**

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = redd_pit_depth)) +
  geom_histogram() +
  theme_minimal()
```

**Numeric Summary of Redd pit depth over Period of Record**

```{r}
summary(cleaner_redd_data$redd_pit_depth)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$redd_pit_depth))/nrow(cleaner_redd_data), 3) * 100` % of values in the `redd_pit_depth` column are NA. 
* There are a lot of 0 values. Could these be NA?


### Variable: `redd_tail_depth`


**Plotting distribution of redd tail depth**

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = redd_tail_depth)) +
  geom_histogram() +
  theme_minimal()
```

**Numeric Summary of Redd tail depth over Period of Record**

```{r}
summary(cleaner_redd_data$redd_tail_depth)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$redd_tail_depth))/nrow(cleaner_redd_data), 3) * 100` % of values in the `redd_tail_depth` column are NA. 
* There are a lot of 0 values. Could these be NA?


### Variable: `redd_length_in`


**Plotting distribution of redd length inches**

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = redd_length_in)) +
  geom_histogram() +
  theme_minimal()
```

**Numeric Summary of Redd length inches over Period of Record**

```{r}
summary(cleaner_redd_data$redd_length_in)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$redd_length_in))/nrow(cleaner_redd_data), 3) * 100` % of values in the `redd_length_in` column are NA. 
* There are a lot of 0 values. Could these be NA?



### Variable: `redd_width_in``


**Plotting distribution of redd width inches**

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = redd_width_in)) +
  geom_histogram() +
  theme_minimal()
```

**Numeric Summary of Redd width inches over Period of Record**

```{r}
summary(cleaner_redd_data$redd_width_in)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$redd_width_in))/nrow(cleaner_redd_data), 3) * 100` % of values in the `redd_width_in` column are NA. 
* There are a lot of 0 values. Could these be NA?



### Variable: `flow_fps`


**Plotting distribution of flow feet per second**

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = flow_fps)) +
  geom_histogram() +
  theme_minimal()
```

```{r}
cleaner_redd_data %>% 
  ggplot(aes(x = flow_fps, y = reach)) +
  geom_boxplot() +
  theme_minimal()
```
**Numeric Summary of flow over Period of Record**

```{r}
summary(cleaner_redd_data$flow_fps)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$flow_fps))/nrow(cleaner_redd_data), 3) * 100` % of values in the `flow_fps` column are NA. 
* There are a lot of 0 values. Could these be NA?


TODO Add in start time and end, see if I can fix format of them first

## Explore Categorical variables: {.tabset}


```{r}
cleaner_redd_data %>% select_if(is.character) %>% colnames()
```


### Variable: `reach``
```{r}
table(cleaner_redd_data$reach) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$reach))/nrow(cleaner_redd_data), 3) * 100` % of values in the `reach`` column are NA. 


### Variable: `pre_redd_substrate_size`
```{r}
table(cleaner_redd_data$pre_redd_substrate_size) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_redd_data$pre_redd_substrate_size <- if_else(
  cleaner_redd_data$pre_redd_substrate_size == ".1 to 1" | 
  cleaner_redd_data$pre_redd_substrate_size == "0.1-1", "0.1 to 1", cleaner_redd_data$pre_redd_substrate_size
)

cleaner_redd_data$pre_redd_substrate_size <- if_else(
  cleaner_redd_data$pre_redd_substrate_size == "<.1", "<0.1", cleaner_redd_data$pre_redd_substrate_size
)
table(cleaner_redd_data$pre_redd_substrate_size) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$pre_redd_substrate_size))/nrow(cleaner_redd_data), 3) * 100` % of values in the `pre_redd_substrate_size` column are NA. 


### Variable: `redd_substrate_size`
```{r}
table(cleaner_redd_data$redd_substrate_size) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_redd_data$redd_substrate_size <- if_else(
  cleaner_redd_data$redd_substrate_size == ".1 to 1" | 
  cleaner_redd_data$redd_substrate_size == "0.1-1", "0.1 to 1", cleaner_redd_data$redd_substrate_size
)

cleaner_redd_data$redd_substrate_size <- if_else(
  cleaner_redd_data$redd_substrate_size == "<.1", "<0.1", cleaner_redd_data$redd_substrate_size
)

cleaner_redd_data$redd_substrate_size <- ifelse(
  cleaner_redd_data$redd_substrate_size == "NA", NA, cleaner_redd_data$redd_substrate_size
)
table(cleaner_redd_data$redd_substrate_size) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$redd_substrate_size))/nrow(cleaner_redd_data), 3) * 100` % of values in the `redd_substrate_size` column are NA. 





### Variable: `tail_substrate_size`
```{r}
table(cleaner_redd_data$tail_substrate_size) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_redd_data$tail_substrate_size <- if_else(
  cleaner_redd_data$tail_substrate_size == ".1 to 1" | 
  cleaner_redd_data$tail_substrate_size == "0.1-1", "0.1 to 1", cleaner_redd_data$tail_substrate_size
)


cleaner_redd_data$tail_substrate_size <- ifelse(
  cleaner_redd_data$tail_substrate_size == "NA", NA, cleaner_redd_data$tail_substrate_size
)
table(cleaner_redd_data$tail_substrate_size) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$tail_substrate_size))/nrow(cleaner_redd_data), 3) * 100` % of values in the `tail_substrate_size` column are NA. 





### Variable: `fish_garding`
```{r}
table(cleaner_redd_data$fish_garding) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_redd_data$fish_garding <- case_when(
  cleaner_redd_data$fish_garding == "No" | cleaner_redd_data$fish_garding == "NO" ~FALSE, 
  cleaner_redd_data$fish_garding == "Yes" | cleaner_redd_data$fish_garding == "YES" ~TRUE
)

table(cleaner_redd_data$fish_garding) 

```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$fish_garding))/nrow(cleaner_redd_data), 3) * 100` % of values in the `fish_garding` column are NA. 


### Variable: `redd_measured`
```{r}
table(cleaner_redd_data$redd_measured) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_redd_data$redd_measured <- case_when(
  cleaner_redd_data$redd_measured == "NO"  ~ FALSE, 
  cleaner_redd_data$redd_measured == "YES" ~ TRUE
)

table(cleaner_redd_data$redd_measured) 

```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$redd_measured))/nrow(cleaner_redd_data), 3) * 100` % of values in the `redd_measured` column are NA. 



### Variable: `why_not_measured`
```{r}
table(cleaner_redd_data$why_not_measured) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_redd_data$why_not_measured <- case_when(
  cleaner_redd_data$why_not_measured == "Fish on redd" | 
    cleaner_redd_data$why_not_measured == "FISH ON REDD"  ~ "fish on redd", 
  cleaner_redd_data$why_not_measured == "Sub-Sample" | 
    cleaner_redd_data$why_not_measured == "SUB-SAMPLE"  ~ "sub sample", 
  cleaner_redd_data$why_not_measured == "Too Deep" ~ "too deep", 
)

table(cleaner_redd_data$why_not_measured) 

```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$why_not_measured))/nrow(cleaner_redd_data), 3) * 100` % of values in the `why_not_measured` column are NA. 


### Variable: `flow_meter`
```{r}
table(cleaner_redd_data$flow_meter) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_redd_data$flow_meter <- case_when(
  cleaner_redd_data$flow_meter %in% c("flow bomb", "Flow Bomb", "Flow bomb")  ~ "flow bomb", 
  cleaner_redd_data$flow_meter == "Digital" ~ "digital",
  cleaner_redd_data$flow_meter == "Flow Watch"  ~ "flow watch", 
  cleaner_redd_data$flow_meter == "Marsh" ~ "marsh", 
)

table(cleaner_redd_data$flow_meter) 

```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$flow_meter))/nrow(cleaner_redd_data), 3) * 100` % of values in the `flow_meter` column are NA. 


### Variable: `comments`
```{r}
unique(cleaner_redd_data$comments)[1:10]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_redd_data$comments))/nrow(cleaner_redd_data), 3) * 100` % of values in the `comments` column are NA. 

## Summary of identified issues

* there are a lot of zero values for the physical characteristics of redds, I need to figure out if these are not measured values or are actually zero

## Save cleaned data back to google cloud 



```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```