---
title: "Battle Creek Carcass Survey QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Battle Creek Carcass Survey QC

## Description of Monitoring Data

These data were aquired via snorkel and kayak surveys on Battle Creek from 1996 to 2019 and describe spring-run and unknown run Chinook Salmon carcasses found within and along Battle Creek.						


**Timeframe:** 1996 - 2019

**Survey Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:** Battle Creek 

**Data Contact:** [Natasha Wingerter](mailto:natasha_wingerter@fws.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
sheets <- excel_sheets("raw_adult_spawn_hold_carcass.xlsx")
sheets 
raw_carcass_data <- read_excel("raw_adult_spawn_hold_carcass.xlsx", sheet = "Carcass") %>% glimpse()
```

## Data transformations

```{r}
cleaner_carcass_data <- raw_carcass_data %>%
  janitor::clean_names() %>% glimpse()
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
```

### Variable: `[name]`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
# Filter clean data to show only categorical variables
```


### Variable: `[name]`
```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Summary of identified issues

* List things that are funcky/bothering us but that we don't feel like should be changed without more investigation

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```
