---
title: "clear_holding_survey_qc"
author: "Inigo Peng"
date: "11/4/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
```


# Clear Creek Adult Holding Survey 

## Description of Monitoring Data

These data were collected by the U.S. Fish and Wildlife Service's, Red Bluff Fish and Wildlife Office's, Clear Creek Monitoring Program.These data encompass spring-run Chinook Salmon redd index data from 2000 to 2019. Data were collected on lower Clear Creek from Whiskeytown Dam located at river mile 18.1, (40.597786N latitude, -122.538791W longitude [decimal degrees]) to the Clear Creek Video Station located at river mile 0.0 (40.504836N latitude, -122.369693W longitude [decimal degrees]) near the confluence with the Sacramento River.						


**Timeframe:** 2008 - 2019

**Completeness of Record throughout timeframe:** 

**Sampling Location:** Clear Creek 

**Data Contact:** [Ryan Schaefer](mailto:ryan_a_schaefer@fws.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/clear-creek/data-raw/FlowWest SCS JPE Data Request_Clear Creek.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_redd_holding_carcass_data.xlsx")
               # Overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data sheet: 

```{r include=FALSE}
 
raw_holding_data <-readxl::read_excel("raw_redd_holding_carcass_data.xlsx", sheet = "Live") %>% glimpse()
```
## Data Transformation

```{r}
cleaner_data <- raw_holding_data %>% 
  janitor::clean_names() %>% 
  select(-c('survey','method','qc_type','qc_date','inspector','year')) %>% #all method is snorkel, year could be extracted from date, 
  rename('longitude' = 'point_x',
         'latitude' = 'point_y',
         'count' = 'total_fish',
         'jack_count' = 'num_of_jacks') %>% 
  mutate(date = as.Date(date)) %>% 
  glimpse()
```

## Explore Date
```{r}
summary(cleaner_data$date)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$date))/nrow(cleaner_data), 3)*100` % of values in the `date` column are NA.

## Explore Categorical Variables

```{r}
cleaner_data %>% 
  select_if(is.character) %>% colnames()
```

### Variable: `reach`

**Description:** reach surveyed on each survey day

```{r}
table(cleaner_data$reach)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$reach))/nrow(cleaner_data), 3)*100` % of values in the `reach` column are NA.

### Variable: `comments`

```{r}
unique(cleaner_data$comments)[1:5]
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$comments))/nrow(cleaner_data), 3)*100` % of values in the `comments` column are NA.

### Variable: `species`

Change to lower-case capitalization; we are only interested in chinook data.

```{r}

cleaner_data <- cleaner_data %>% 
  mutate(species = tolower(species)) %>% 
  filter(species == 'chinook') 

table(cleaner_data$species)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$species))/nrow(cleaner_data), 3)*100` % of values in the `species` column are NA.

### Variable: `survey_intent`

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(survey_intent = tolower(survey_intent))

table(cleaner_data$survey_intent)

```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$survey_intent))/nrow(cleaner_data), 3)*100` % of values in the `survey_intent` column are NA.

### Variable: `pw_relate`

**Description:** Fish above or below Picket Weir

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(pw_relate = tolower(pw_relate))
table(cleaner_data$pw_relate)
```

## Explore Numerical Variables

```{r}
cleaner_data %>% 
  select_if(is.numeric) %>% colnames()
```


### Variable: `river_mile`

Plotting river mile over Period of Record

```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = river_mile, y = year)) +
  geom_point(alpha = .5, size = 1.5, color = "blue") +
  labs(x = "River Mile", y = "Year", title = "River Mile Over the Years")
```

```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x= river_mile, y = count, color = year))+
  geom_point()+
  theme_minimal()+
  labs(title = "Total Fish Count Per River Mile")
```

**Numeric Summary of river_mile Over the Years**

```{r}
summary(cleaner_data$river_mile)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$river_mile))/nrow(cleaner_data), 3)*100` % of values in the `river_mile` column are NA.


## Variable: `longitude` and `latitude`

**Numeric Summary of longitude over Period of Record**
```{r}
summary(cleaner_data$longitude)
```

**Numeric Summary of latitude over Period of Record**

```{r}
summary(cleaner_data$latitude)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$longitude))/nrow(cleaner_data), 3)*100` % of values in the `longitude` column are NA.

*  `r round(sum(is.na(cleaner_data$latitude))/nrow(cleaner_data), 3)*100` % of values in the `latitude` column are NA.


## Variable: `count`

**Description:** total number of Adult Chinook Salmon encountered including the number of 2-year olds (Jacks/Jills) 

```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date)),
         fake_date = as.Date(paste0("1990", "-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = count)) + 
  geom_col() + 
  facet_wrap(~year(date), scales = "free_y") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 13),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Holding Count", 
       x = "Date")  
```
```{r}
cleaner_data %>% 
  group_by(date) %>%
  summarise(daily_count = sum(count)) %>%
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = year, y = daily_count)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Daily Count Summarized by Year") + 
  theme(text = element_text(size = 13),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))   + 
  scale_y_continuous(limits = c(0, 101))
```

```{r}
cleaner_data  %>%
  mutate(year = as.factor(year(date))) %>%
  group_by(year(date)) %>%
  mutate(total_catch = sum(count)) %>%
  ungroup() %>%
  ggplot(aes(x = year, y = total_catch)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Yearly Fish Count",
       y = "Total fish count") + 
  theme(text = element_text(size = 13),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of count over Period of Record**
```{r}
summary(cleaner_data$count)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$count))/nrow(cleaner_data), 3)*100` % of values in the `count` column are NA.


## Variable: `jack_count`

**Description:** total number of 2-year old Chinook Salmon (Jacks/Jills) encountered

```{r}
cleaner_data %>% 
  ggplot(aes(x = jack_count)) +
  geom_histogram(bins = 4) +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Numeric Summary of jack count over Period of Record**
```{r}
summary(cleaner_data$jack_count)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$jack_count))/nrow(cleaner_data), 3)*100` % of values in the `jack_count` column are NA.


## Variable: `pw_location_rm`

**Description:** location of the Picket Weir

```{r}
cleaner_data %>% 
  ggplot(aes(x = pw_location_rm)) +
  geom_histogram(bins = 4) +
  theme_minimal() + 
  theme(text = element_text(size = 15))+
  labs(title = "Distribution of pw_location_rm")
```



**Numeric Summary of pw_location_rm over Period of Record**
```{r}

summary(cleaner_data$pw_location_rm)

```

Seems like there are two locations

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$pw_location_rm))/nrow(cleaner_data), 3)*100` % of values in the `pw_location_rm` column are NA.


## Save cleaned data back to google cloud

```{r}
clear_holding <- cleaner_data  
```

```{r, eval=FALSE}
gcs_list_objects()
f <- function(input, output) write_csv(input, file = output)
gcs_upload(clear_holding,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_holding.csv")
```
