---
title: "Clear Carcass Survey QC"
author: "Inigo Peng"
date: "11/4/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
```


# Clear Creek Adult Carcass Survey 

## Description of Monitoring Data

This carcass data were collected by the U.S. Fish and Wildlife Service's, Red Bluff Fish and Wildlife Office's, Clear Creek Monitoring Program.This data encompass spring-run Chinook Salmon carcasses retrieved on redd and escapement index surveys from 2008 to 2019 on Clear Creek. Data were collected on lower Clear Creek from Whiskeytown Dam located at river mile 18.1, (40.597786N latitude, -122.538791W longitude) to the Clear Creek Video Station located at river mile 0.0 (40.504836N latitude, -122.369693W longitude) near the confluence with the Sacramento River.						


**Timeframe:** 2008 - 2019

**Completeness of Record throughout timeframe:** 

Data available for all years.

**Sampling Location:** Clear Creek 

**Data Contact:** [Ryan Schaefer](mailto:ryan_a_schaefer@fws.gov)

This [report](https://www.fws.gov/redbluff/CC%20BC/Battle%20Creek%20Monitoring%20Final%20Reports/2019%20Battle%20Creek%20Adult%20Monitoring%20Report.pdf) gives additional information on Adult Chinook monitoring and carcass collection. 

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/clear-creek/data-raw/FlowWest SCS JPE Data Request_Clear Creek.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_redd_holding_carcass_data.xlsx")
               # Overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data sheet: 

```{r}
 
raw_carcass_data <-readxl::read_excel("raw_redd_holding_carcass_data.xlsx", sheet = "Carcass") %>% glimpse()
```
```{r}
cleaner_data <- raw_carcass_data %>% 
  janitor::clean_names() %>% 
  rename('longitude' = 'point_x',
         'latitude' = 'point_y',
         'condition' = 'condit',
         'sex' = 'gender',
         'why_sex_unknown' = 'why_gender',
         'head_retrieved' = 'head_tak',
         'spawn_status' = 'spawn_st',
         'brood_year' = 'by',
         'fork_length' = 'fork_len') %>% 
  mutate(date = as.Date(date),
         fork_length = as.numeric(fork_length),
         age = as.numeric(age),
         brood_year = as.numeric(brood_year),
         sample_id = as.character(sample_id)) %>% 
    filter(species == "Chinook") %>%
    select(-c('survey','qc_type','qc_date','inspector','year', 'year_id', 'species')) %>% 
  glimpse()

```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data. 

```{r data_dictionary}
percent_na <- cleaner_data %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_data),
                          description = c("Survey type", 
                                          "Date of sampling",
                                          "GPS X point",
                                          "GPS Y point",
                                          "Reach number (1-7); other location",
                                          "River mile number",
                                          "True if the fish was observed and not sampled (T/F)",
                                          "Unique sample ID number",
                                          "Adipose status (absent, present, unknown)",
                                          "Fork length of fish in mm",
                                          "Condition of fish (fresh or not fresh)",
                                          "? TODO", 
                                          "? TODO", 
                                          "True if a scale sample was taken", 
                                          "True if an otolith sample was taken",
                                          "Male, Female, Unknown",
                                          "Why fish sex is unknown (decomposed, predation, NA)",
                                          "Spawning status (partial, spawned, unspawned)",
                                          "Why not spawned (decomposed, male, male always unknown, most eggs present, observed only, predation, prespawn, too decomposed",
                                          "True if fish head was retrieved", 
                                          "Type of tag, if any present (Floy, external mark)",
                                          "Code describing if a photo of the carcass was taken, TODO get code definitions", 
                                          "General comments from survey crew",
                                          "Coded wire tag number",
                                          "Run of Chinook (spring, fall, late-fall, winter, hybrid, unknown)",
                                          "Brood year of carcass found", 
                                          "Location where hatchery smolts were released, TODO figure out how this is relevent to carcass data", 
                                          "Hatchery that produced the fish (Coleman hatchery or feather river hatchery)", 
                                          "Age of fish", 
                                          "Rate at which Hatchery Fish were marked, TODO figure out what this means",
                                          "CWT code and relevant information, TODO clean this up",
                                          "Run call based on field data, TODO differnciate from other run",
                                          "Type of genetics taken, TODO elaborate on options"),
                          
                          percent_na = round(percent_na$value*100)
                          
)
knitr::kable(data_dictionary)
```

## Explore date

```{r}
summary(cleaner_data$date)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$date))/nrow(cleaner_data), 3) * 100` % of values in the `date` column are NA.

## Explore Categorical Data

```{r}
cleaner_data %>% select_if(is.character) %>% colnames()
```
### Variable: `type`

**Description:** Survey Type (Kayak, Rotary Screw Trap, CCVS, Etc.)

TODO: metadata question: what is psam and ccvs? 

```{r}
cleaner_data$type <- tolower(cleaner_data$type)
table(cleaner_data$type)
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$survey))/nrow(cleaner_data), 3) * 100` % of values in the `survey` column are NA.


### Variable: `reach`

**Description:** Reach surveyed on each survey day

```{r}
table(cleaner_data$reach)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$reach))/nrow(cleaner_data), 3) * 100` % of values in the `reach` column are NA.

### Variable: `obs_only`
```{r}

cleaner_data <- cleaner_data %>%
  mutate(obs_only = tolower(obs_only),
         obs_only = case_when(
           obs_only == "no" ~ FALSE,
           obs_only == "yes" ~ TRUE
         ))
table(cleaner_data$obs_only)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$obs_only))/nrow(cleaner_data), 3) * 100` % of values in the `obs_only` column are NA.

### Variable: `adipose`

**Description:** Indicates the presence or absence of an adipose fin
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(adipose = tolower(adipose))
table(cleaner_data$adipose)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$adipose))/nrow(cleaner_data), 3) * 100` % of values in the `adipose` column are NA.

### Variable: `condition`

**Description:** fresh or not fresh

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(condition = tolower(condition),
         condition = ifelse(condition == "unknown", NA_character_, condition))
table(cleaner_data$condition)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$condition))/nrow(cleaner_data), 3) * 100` % of values in the `condition` column are NA.


### Variable: `tis_eth`

**Description:** Tissue for ethanol storage? 
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(tis_eth = tolower(tis_eth))
table(cleaner_data$tis_eth)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$tis_eth))/nrow(cleaner_data), 3) * 100` % of values in the `tis_eth` column are NA.

### Variable: `tis_dry`

**Description:** Dry tissue sample? Yes or No
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(tis_dry = tolower(tis_dry))
table(cleaner_data$tis_dry)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$tis_dry))/nrow(cleaner_data), 3) * 100` % of values in the `tis_dry` column are NA.

### Variable: `scale`

**Description:** Scales collected? Yes or No

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(scale = tolower(scale),
         scale = case_when(
           scale == "no" ~ FALSE,
           scale == "yes" ~ TRUE
         ))
table(cleaner_data$scale)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$scale))/nrow(cleaner_data), 3) * 100` % of values in the `scale` column are NA.

### Variable: `otolith_st`

**Description:** Otoliths collected? Yes or No

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(otolith_st = tolower(otolith_st),
         otolith_st = case_when(
           otolith_st == "no" ~ FALSE,
           otolith_st == "yes" ~ TRUE
         ))
table(cleaner_data$otolith_st)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$otolith_st))/nrow(cleaner_data), 3) * 100` % of values in the `otolith_st` column are NA.

### Variable: `sex`

**Description:** sex of carcass
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(sex = tolower(sex),
         sex = ifelse(sex == "unknown", NA_character_, sex))
table(cleaner_data$sex)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$sex))/nrow(cleaner_data), 3) * 100` % of values in the `sex` column are NA.

### Variable: `why_sex_unknown`

**Description:** Indicates why sex is unknown
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(why_sex_unknown = tolower(why_sex_unknown))
table(cleaner_data$why_sex_unknown)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$why_sex_unknown))/nrow(cleaner_data), 3) * 100` % of values in the `why_sex_unknown` column are NA.

### Variable: `spawn_status`

**Description:** Spawning status (spawned, unspawned, unknown)
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(spawn_status = tolower(spawn_status),
         spawn_status = ifelse(spawn_status == "unknown", NA_character_, spawn_status))
table(cleaner_data$spawn_status)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$spawn_status))/nrow(cleaner_data), 3) * 100` % of values in the `spawn_status` column are NA.

### Variable: `why_not_sp`

**Description:** If unspawned, why unspawned (Predation, etc)

TODO what is male always unknown?

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(why_not_sp = tolower(why_not_sp),
         why_not_sp = case_when(
           why_not_sp == "decomposition" ~ "decomposed",
           why_not_sp == "unknown"| why_not_sp == "n/a"| why_not_sp == "na" ~ NA_character_,
           why_not_sp == "male-always unknown" | why_not_sp == "males-always unk" ~ "male always unknown",
           TRUE ~ as.character(why_not_sp)
           ))
table(cleaner_data$why_not_sp)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$spawn_status))/nrow(cleaner_data), 3) * 100` % of values in the `spawn_status` column are NA.

### Variable: `head_retrieved`

**Description:** Indicates if head was retreived for coded wire tag extraction
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(head_retrieved = tolower(head_retrieved),
         head_retrieved = case_when(
           head_retrieved == "no" ~ FALSE,
           head_retrieved == "yes" ~ TRUE
         ))
table(cleaner_data$head_retrieved)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$head_retrieved))/nrow(cleaner_data), 3) * 100` % of values in the `head_retrieved` column are NA.

### Variable: `tag_type`

**Description:** Type of tag, if any present (Floy, external mark)

```{r}
cleaner_data<- cleaner_data %>% 
  mutate(tag_type = tolower(tag_type),
         tag_type = case_when(
           tag_type == "floy, grey" ~ "floy grey",
           tag_type == "floy, yellow, 309,310" ~ "floy yellow 309 310",
           TRUE ~ as.character(tag_type)
         )
         )
table(cleaner_data$tag_type)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$tag_type))/nrow(cleaner_data), 3) * 100` % of values in the `tag_type` column are NA.

### Variable: `photo`



**Description:** Photo taken of the carcass? 

```{r}
cleaner_data<- cleaner_data %>% 
  mutate(photo = case_when(
    photo == "No"| photo == "NO"| photo == "No picture" | photo == "Cant find" | photo == "missing picture" ~ "no",
    photo == "YES" | photo == "Yes" ~ "yes",
    TRUE ~ as.character(photo)
    
  ))
table(cleaner_data$photo)
```

TODO: metadata issue - what do the numbers stand for? Number of pictures?

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$photo))/nrow(cleaner_data), 3) * 100` % of values in the `photo` column are NA.

### Variable: `comments`
```{r}
unique(cleaner_data$comments[1:5])
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$comments))/nrow(cleaner_data), 3) * 100` % of values in the `comments` column are NA.

### Variable: `cwt_code`

TODO: metadata - what is "LT", "NTD"

```{r}
table(cleaner_data$cwt_code)
```

There are `r length(unique(cleaner_data$cwt_code))` unique coded wire tags.

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$cwt_code))/nrow(cleaner_data), 3) * 100` % of values in the `cwt_code` column are NA.

### Variable: `run`

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(run = tolower(run),
         run = case_when(
           run == 'hybrid (5)' ~ 'hybrid',
           TRUE ~ as.character(run)
         ))
table(cleaner_data$run)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$run))/nrow(cleaner_data), 3) * 100` % of values in the `run` column are NA.


### Variable: `release_location`

**Description:** Where hatchery smolts were released

Note: why is smolts location part of the data for carcass?

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(release_location = str_to_title(release_location),
         release_location = case_when(
           release_location == 'Sac R Lake Redding' ~ 'Sac R Lake Redding Park',
           TRUE ~ as.character(release_location)
         ))
table(cleaner_data$release_location)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$release_location))/nrow(cleaner_data), 3) * 100` % of values in the `release_location` column are NA.

### Variable: `hatchery`

**Description:** Hatchery that produced the fish

```{r}
# table(cleaner_data$hatchery)
cleaner_data <- cleaner_data %>% 
  mutate(hatchery = tolower(hatchery),
         hatchery = case_when(
           hatchery == "feather r hatchery" | hatchery == "feather river" | hatchery == "frh" ~ "feather river hatchery",
           hatchery == "cnfh" ~ "coleman nfh", 
           TRUE ~ as.character(hatchery)
           ))
table(cleaner_data$hatchery)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$hatchery))/nrow(cleaner_data), 3) * 100` % of values in the `hatchery` column are NA.

### Variable: `verification_and_cwt_comments`

**Description:** CWT code and relevant information
```{r}
unique(cleaner_data$verification_and_cwt_comments)[1:5]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$verification_and_cwt_comments))/nrow(cleaner_data), 3) * 100` % of values in the `verification_and_cwt_comments` column are NA.

### Variable: `sample_id`

**Description:** Tissue sample identification number

```{r}
unique(cleaner_data$sample_id)[1:5]
```

There are `r length(unique(cleaner_data$sample_id))` unique sample IDs.

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$sample_id))/nrow(cleaner_data), 3) * 100` % of values in the `sample_id` column are NA.


### Variable: `run_call`

**Description:** Run call based on field data

TODO: how is this different from run?

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(run_call = tolower(run_call))
  
table(cleaner_data$run_call)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$run_call))/nrow(cleaner_data), 3) * 100` % of values in the `run_call` column are NA.

### Variable: `genetic`

**Description:** Type of genetics taken
```{r}
cleaner_data$genetic <- tolower(cleaner_data$genetic)
table(cleaner_data$genetic)
```

TODO: metadata need descriptions on these variables

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$genetic))/nrow(cleaner_data), 3) * 100` % of values in the `genetic` column are NA.

## Explore Numerical Data

```{r}
cleaner_data %>% select_if(is.numeric) %>% colnames()
```
### Variable: `longitude`, `latitude`

```{r}
summary(cleaner_data$longitude)
```

```{r}
summary(cleaner_data$latitude)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$longitude))/nrow(cleaner_data), 3) * 100` % of values in the `longitude` column are NA.

* `r round(sum(is.na(cleaner_data$latitude))/nrow(cleaner_data), 3) * 100` % of values in the `latitude` column are NA.

## Variable: `river_mile`

```{r}
cleaner_data %>%
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = river_mile, y = year))+
  geom_point(alpha = 1.2, color = 'blue')+
  labs(title = 'River Mile over the Years')+
  theme_minimal()
```


**Numeric Summary of river_mile over Period of Time**

```{r}
summary(cleaner_data$river_mile)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$river_mile))/nrow(cleaner_data), 3) * 100` % of values in the `river_mile` column are NA.

### Variable: `brood_year`

```{r}
cleaner_data %>% 
  ggplot(aes(x = brood_year))+
  geom_histogram()+
  theme_minimal()+
  scale_x_continuous(breaks = pretty_breaks())+
  labs(title = "Distribution of Brood Year")
```

**Numeric Summary of brood_year over Period of Time**
```{r}
summary(cleaner_data$brood_year)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$brood_year))/nrow(cleaner_data), 3) * 100` % of values in the `brood_year` column are NA.

### Variable: `fork_length`

**Description:** Fork length of recovered carcasses

```{r}
cleaner_data %>% 
  ggplot(aes(x = fork_length))+
  geom_histogram()+
  theme_minimal()+
  labs(title = "Distribution of Fork Length")
```

**Numeric Summary of fork_length over Period of Time**

```{r}
summary(cleaner_data$fork_length)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$fork_length))/nrow(cleaner_data), 3) * 100` % of values in the `fork_length` column are NA.

### Variable: `age`

```{r}
cleaner_data %>% 
  ggplot(aes(x= age))+
  geom_histogram()+
  theme_minimal()+
  labs(title = "Distribution of Age")
```

**Numeric Summary of Age Over Period of Time**
```{r}
summary(cleaner_data$age)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$age))/nrow(cleaner_data), 3) * 100` % of values in the `age` column are NA.

### Variable: `mark_rate`

**Description:** Rate at which Hatchery Fish were marked
```{r}
cleaner_data %>% 
  ggplot(aes(x=mark_rate))+
  geom_histogram(binwidth = 0.1)+
  theme_minimal()+
  labs(title = "Distribution of Mark Rate")
```

```{r}
summary(cleaner_data$mark_rate)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_data$mark_rate))/nrow(cleaner_data), 3) * 100` % of values in the `mark_rate` column are NA.

### Summary of Identified Issues

* Some columns need better metadata description (tis_eth, tis_dry, photo, release_location, mark_rate, verification_and_cwt_comments, run_call)
* Some columns such as 'run' and 'run_call' seem to contain the same type of information but have very different values. TODO need to contact Ryan for more details on these columns. 

## Next steps

* Work on data modeling to identify important variables needed for carcass datasets.

### Columns to remove

* Suggest removing some of the location variables we currently have: `longitude`, `latitude`, `reach` and `river_mile`. 
* Suggest removing one of the run columns, either `run` or `run_call` should be removed 
* `genetics`, `verification_and_cwt_comments`, `mark_rate`, `hatchery`, `release_location`, `brood_year`, `comments` and `photo` all contain a lot of NA values. These may not be needed.

### Save Cleaned data back to google cloud
```{r}
clear_carcass <- cleaner_data %>% glimpse
```

```{r, eval=FALSE}
# gcs_list_objects()
f <- function(input, output) write_csv(input, file = output)
gcs_upload(clear_carcass,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_carcass.csv")
```