---
title: "Lower Sacramento (Knight's Landing) RST data QC - trap data"
author: "Ashley Vizek"
date: "10/22/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
```

## Description of Monitoring Data

**Timeframe:** 

10/04/2002 through 05/31/2021

**Completeness of Record throughout timeframe:** 

- The 2013 sampling period is not complete (Oct 2012 through Dec 2012)
- We are missing data for 2017 and 2018. 

**Sampling Location:**

Knight's Landing

**Data Contact:** 

[Jeanine Phillips](mailto::Jeanine.Phillips@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()

get_data <- function(name) {
gcs_get_object(object_name = paste0("rst/lower-sac-river/data/knights-landing/knl_combine_", name,".csv"),
               bucket = gcs_get_global_bucket(),
               saveToDisk = paste0("knl_combine_", name, ".csv"),
               overwrite = TRUE)
}

get_data("rst")
```

Read in data from google cloud, glimpse raw data and domain description sheet.
Need to change date, start_date, stop_date into appropriate format. Make changes
in changes section at end.

```{r}
# read in data to clean 
knl_rst <- read.csv("knl_combine_rst.csv")
glimpse(knl_rst)
```

## Data transformations

This work was done in other rmd (1_knights_landing_initial_clean.Rmd, 2_knights_landing_combined_data.Rmd)

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables (this way we know we do not miss any)
knl_rst %>%
  select_if(is.numeric) %>%
  colnames()
```

### Variable: `fork_length_max_mm`

**Plotting `fork_length_max_mm` over Period of Record**

```{r}
ggplot(knl_rst, aes(y = fork_length_max_mm, x = factor(year(date)))) +
  geom_point() 
```

There are some datapoints that extend to 200. These may be spring run salmon but unable to
tell because min/max fork length is collected for all unmarked salmon/runs and not separated by run.

**Numeric Summary of `fork_length_max_mm` over Period of Record**

There are a lot of NAs for this variable because it is not collected for marked salmon and for marked/unmarked steelhead.

```{r}
knl_rst %>%
  group_by(year(date)) %>%
  summarize(mean = mean(fork_length_max_mm, na.rm = T),
            median = median(fork_length_max_mm, na.rm = T),
            min = min(fork_length_max_mm, na.rm = T),
            max = max(fork_length_max_mm, na.rm = T),
            na = length(which(is.na(fork_length_max_mm))))
```

I do not think that zeros should exist and should be NAs. Only one day when fork length was 0 and count was not 0 and I think it just wasn't recorded on this day rather than being a real 0.

```{r}
filter(knl_rst, fork_length_max_mm == 0 & count > 0)
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_rst$fork_length_max_mm))/nrow(knl_rst), 3) * 100` % of values in the `fork_length_max_mm` column are NA.

### Variable: `fork_length_min_mm`

**Plotting `fork_length_min_mm` over Period of Record**

```{r}
ggplot(knl_rst, aes(y = fork_length_min_mm, x = factor(year(date)))) +
  geom_point() 
```

**Numeric Summary of `fork_length_min_mm` over Period of Record**

```{r}
knl_rst %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(fork_length_min_mm, na.rm = T),2),
            median = round(median(fork_length_min_mm, na.rm = T),2),
            min = round(min(fork_length_min_mm, na.rm = T),2),
            max = round(max(fork_length_min_mm, na.rm = T),2),
            na = length(which(is.na(fork_length_min_mm))))
```

I do not think that zeros should exist and should be NAs. Only one day when fork length was 0 and count was not 0 and I think it just wasn't recorded on this day rather than being a real 0.

```{r}
filter(knl_rst, fork_length_min_mm == 0 & count > 0)
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_rst$fork_length_min_mm))/nrow(knl_rst), 3) * 100` % of values in the `fork_length_min_mm` column are NA.

### Variable: `count`

**Plotting `count` over Period of Record**

*Natural Chinook*

Data is dominated by fall run. There are some spring run peaks 2019-2021.

```{r}
knl_rst %>%
  filter(species == "Chinook", marked == F) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
ggplot(aes(y = count, x = fake_date, color = at_capture_run)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line() +
  facet_wrap(~wy, scales = "free_y") +
  xlab("") +
  theme(legend.position="bottom")
```

*Marked Chinook*

Marked Chinook are not recorded by run until 2019-2021 datasets. Much fewer counts than natural run.

```{r}
knl_rst %>%
  filter(species == "Chinook", marked == T) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
ggplot(aes(y = count, x = fake_date, color = at_capture_run)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line() +
  facet_wrap(~wy, scales = "free_y") +
  xlab("")
```

*Natural Steelhead*

Few Steelhead are recorded.

```{r}
knl_rst %>%
  filter(species == "Steelhead", marked == F) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
ggplot(aes(y = count, x = fake_date)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line() +
  facet_wrap(~wy, scales = "free_y") +
  xlab("")
```

*Marked Steelhead*

More marked than natural Steelhead recorded.

```{r}
knl_rst %>%
  filter(species == "Steelhead", marked == T) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
ggplot(aes(y = count, x = fake_date)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line() +
  facet_wrap(~wy, scales = "free_y") +
  xlab("")
```

**Numeric Summary of `count` over Period of Record**

Only providing summary for natural Chinook. 

Summary by water year

```{r}
knitr::kable(knl_rst %>%
  filter(species == "Chinook", marked == F) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy) %>%
  summarize(mean = round(mean(count, na.rm = T),2),
            median = round(median(count, na.rm = T),2),
            min = round(min(count, na.rm = T),2),
            max = round(max(count, na.rm = T),2),
            na = length(which(is.na(count)))))
```

Total counts by year and run

```{r}
knitr::kable(knl_rst %>%
  filter(species == "Chinook", marked == F) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date)))) %>%
  group_by(wy, at_capture_run) %>%
  summarize(cumulative = sum(count, na.rm = T)) %>%
  pivot_wider(names_from = at_capture_run, values_from = cumulative))
```

Some of the max counts are in the tens of thousands. Either there was an error in some data transformations or perhaps some of these are passage estimates and not raw counts. Check raw data to see if these large counts exist. If not, may need to reach back out to data contact.
These values exist in the data. Contacted Jeanine and she said they seem right to her.

```{r}
filter(knl_rst, count > 5000)

# 2004-02-06
filter(knl_rst, count == 15759)
# 2014-03-03
filter(knl_rst, count == 28948)
# 2014-03-04
filter(knl_rst, count == 12560)
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_rst$count))/nrow(knl_rst), 3) * 100` % of values in the `count` column are NA.

### Variable: `cpue`

**Plotting `cpue` over Period of Record**

*Natural Chinook*

Catch per unit effort seems very high for 2003 and 2004. Could these be passage estimates? Or maybe cpue is being miscalculated? These high values correspond to the high count values identified. 

cpue only calculated for natural species.

cpue for fall/spring and spue late fall/winter was calculated prior to 2015 when they were calculate separately.

```{r}
knl_rst %>%
  filter(species == "Chinook", marked == F) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
ggplot(aes(y = cpue, x = fake_date, color = at_capture_run)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(alpha = 0.5) +
  facet_wrap(~wy, scales = "free_y") +
  xlab("")
```

*Natural Steelhead*

```{r}
knl_rst %>%
  filter(species == "Steelhead", marked == F) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
ggplot(aes(y = cpue, x = fake_date)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(alpha = 0.5) +
  facet_wrap(~wy, scales = "free_y") +
  xlab("")
```

**Numeric Summary of `cpue` over Period of Record**

Only provide summary for natural Chinook

```{r}
knl_rst %>%
  filter(species == "Chinook", marked == F) %>%
  group_by(year(date), at_capture_run, species, marked, lifestage) %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(cpue, na.rm = T),2),
            median = round(median(cpue, na.rm = T),2),
            min = round(min(cpue, na.rm = T),2),
            max = round(max(cpue, na.rm = T),2),
            na = length(which(is.na(cpue))))
```

**NA and Unknown Values**

Only calculated for natural species so there is a high amount of NAs.

Percent of NA:

`r round(sum(is.na(knl_rst$cpue))/nrow(knl_rst), 3) * 100` % of values in the `cpue` column are NA.
  
## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

Need to change dates to appropriate format!

```{r}
knl_rst %>%
  select_if(is.character) %>%
  colnames()
```


### Variable: `location`

```{r}
table(knl_rst$location) 
```

Some years do not have location variable filled in. Fill in for these years.

```{r}
filter(knl_rst, is.na(location)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_rst$location))/nrow(knl_rst), 3) * 100` % of values in the `location` column are NA.

### Variable: `species`

```{r}
table(knl_rst$species) 
```

```{r}
filter(knl_rst, is.na(species)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_rst$species))/nrow(knl_rst), 3) * 100` % of values in the `species` column are NA.

### Variable: `at_capture_run`


```{r}
table(knl_rst$at_capture_run) 
```

Sometimes marked salmon were not recorded by run. Older juvenile salmon were not given a run. Steelhead do not have a run associated.

```{r}
filter(knl_rst, is.na(at_capture_run)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_rst$at_capture_run))/nrow(knl_rst), 3) * 100` % of values in the `at_capture_run` column are NA.

### Variable: `lifestage`

```{r}
table(knl_rst$lifestage) 
```

Only one lifestage called older juvenile. If not in this lifestage then none recorded.

```{r}
filter(knl_rst, is.na(lifestage)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

Lifestage is NA unless older juvenile.

`r round(sum(is.na(knl_rst$lifestage))/nrow(knl_rst), 3) * 100` % of values in the `lifestage` column are NA.

## Make any changes identified

```{r make_changes}
knl_rst_changes <- knl_rst %>%
  # fix typo in water_t_f
  mutate(location = "KL",
         # fix date formatting
         date = as.Date(date),
         start_date = ymd_hms(start_date),
         stop_date = ymd_hms(stop_date),
         # change 0s for fork length to NAs
         fork_length_max_mm = case_when(fork_length_max_mm == 0 ~ NA_real_,
                                        T ~ as.numeric(fork_length_max_mm)),
         fork_length_min_mm = case_when(fork_length_min_mm == 0 ~ NA_real_,
                                        T ~ as.numeric(fork_length_min_mm))) %>%
  rename(catch_per_unit_effort = cpue)
```

## Questions for follow up

Very large counts on 2014-03-03, 2014-03-04, and 2004-02-06. These exist in the raw data. Follow up with contact to double check that these are raw counts or if these might be typos.

### Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
# f <- function(input, output) write_csv(input, file = output)
# 
# upload_data <- function(data, name) {
# gcs_upload(data,
#            object_function = f,
#            type = "csv",
#            name = paste0("rst/lower-sac-river/data/knights-landing/knl_combine_", name, ".csv"))
# }
# 
# upload_data(knl_rst_changes, "rst_clean")
```
