---
title: "[Watershed] [Data type] [QC]"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
```

# Clear Creek Adult Upstream Video Data 

## Description of Monitoring Data
TODO add current knows here

TODO investigate/talk about outages, is a 0 a real zero or explained by other factors
TODO get location of video monitoring station

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
```

## Data transformations

```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables (this way we know we do not miss any)
```

### Variable: `[name]`

**Plotting Passage Counts Moving Up over Period of Record**

```{r}

```

**Numeric Summary of Passage Counts Moving Up over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

`r round(sum(is.na(cleaner_video_data$up))/nrow(cleaner_video_data), 3)` % of values in the `up` column are NA. 

However, there are clearly gaps in data. More investigation needs to be done to see if 0 is a real 0 or if it can be explained by other factors (outages).


## Explore Categorical variables: {.tabset}

```{r}
# Filter clean data to show only categorical variables (this way we know we do not miss any)
```


### Variable: `[name]`
```{r}
table() 
```

**Create lookup rda for [variable] encoding:** 
```{r}

```

**NA and Unknown Values**
  

### Save cleaned data back to google cloud 

```{r}

```