---
title: "Combine Temperature Data"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10, 
                      root.dir = rprojroot::find_rstudio_root_file())
Sys.setenv("GCS_AUTH_FILE" = "config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")
library(tidyverse)
library(googleCloudStorageR)
library(CDECRetrieve)
library(lubridate)
library(hms)
library(weathermetrics) 
library(forecast)
```

FlowWest aggregated environmental data for temperature data from XXXX
monitoring programs:

-   Feather River (source: RST monitoring data)
-   Battle Creek (source: FWS)
-   Butte Creek (source: CDEC)
-   Clear Creek (source: FWS)
-   Deer Creek (source: CDEC)
-   Mill Creek (source: CDEC)
-   Sacramento River (source for Tisdale: USGS; source for Knights
    Landing: RST monitoring data)
-   Yuba River (TODO)

### Rational for Data Sources:

The following decision heirarchyu was used to determine which datasets
were used in the final temperature data aggregation:

1.  Is there publicly available data available through USGS or CDEC?
    This data can easily be built into a data pipeline because it's
    available through a query and publicly available. The data is also
    QC'd and available at precise temporal scales like hourly.
2.  If no, is there data available through Fish and Wildlife?
3.  If no, is there RST temperature monitoring data available?
4.  If no, perform linear regression on air temperature to get modeled
    stream temperature for site

## Standard format for Temperature Data

Data dictionary for standard format:

(B - Battle Creek, Bu - Butte Creek, C - Clear Creek, F - Feather River,
D - Deer Creek, F - Feather River, M - Mill Creek, S- Sacramento River,
Y - Yuba River)

| column name       | stream collects   | definition                                          |
|:-----------------|:-------------------|:---------------------------------|
| date              | B, Bu, C, D, M, F | date that measurement occurs                        |
| stream            | B, Bu, C, D, M, F | unique stream name associated with temperature data |
| site              | B, Bu, C, D, M, F | unique site identifier where measurement occurs     |
| source            | B, Bu, C, D, M, F | data source                                         |
| location          | F                 | differentiates between `high flow` and `low flow`   |
| mean_daily_temp_c | B, Bu, C, D, M, F | average daily temperature (degrees Celsius)         |
| max_daily_temp_c  | B, C              | maximum daily temperature (degrees Celsius)         |

# Pull in data and format {.tabset}

### All RST Environmental Data 

```{r}
# gcs_get_object(object_name = "standard-format-data/standard_RST_environmental.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/standard_RST_environmental.csv",
#                overwrite = TRUE)

rst_enviro <- read_csv('../../data/rst/standard_RST_environmental.csv') %>% glimpse()
```

## Battle Creek

-   One site from the Fish and Wildlife Service provides data from
    `2003-01-07` to `2021-02-2021`

```{r}
ubc_temp_raw <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 4)

battle_format <- ubc_temp_raw %>% 
  rename(date = DT,
         temp_c = TEMP_C) %>%
  mutate(stream = "battle creek", 
         site = "ubc",
         date = as.Date(date),
         source = "FWS") %>%
  group_by(date, stream, site, source) %>%
  summarise(mean_temp_degC = mean(temp_c),
            max_temp_degC = max(temp_c)) %>% 
  filter(mean_temp_degC > 0) %>%
  glimpse

summary(battle_format)

battle_format %>% ggplot() + 
  geom_line(aes(x = date, y = mean_temp_degC, color = site)) + 
  # geom_point(aes(x = date, y = mean_temp_degC), size = .001) +
  theme_minimal()

```

## Butte Creek

-   The CDEC gage `BCK` was used to aggregate temperature for sites
    `adams dam` and `okie dam` from `1998-09-16` to `2022-06-30`.

-   TODO: add max temp value

```{r}
# source CDEC
butte <- read_rds(here::here("data-raw", "standard-format-data-prep","temp_data", "butte_temp.rds"))
butte_format <- butte %>%
  mutate(site = "adams dam") %>%
  bind_rows(butte %>%
              mutate(site = "okie dam")) %>%
  mutate(stream = "butte creek",
         source = "CDEC BCK") %>%
  filter(mean_temp_degC > 0 & mean_temp_degC < 35) %>% # Filter out outliners TODO confirm with ashley that this is an appropriate range
glimpse()

summary(butte_format$date)

butte_format %>% ggplot() + 
  geom_line(aes(x = date, y = mean_temp_degC, color = site)) + 
  # geom_point(aes(x = date, y = mean_temp_degC), size = .001) +
  theme_minimal() 
```

## Clear Creek

-   Clear Creek data was aggregated from Fish Wildlife Service's site
    `lcc` from `2001-02-16` to `2021-12-31`

```{r}
ucc_temp_raw <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 2)

ucc_temp <- ucc_temp_raw %>% 
  rename(date = DT,
         temp_c = TEMP_C) %>%
  mutate(stream = "clear creek", 
         site = "ucc",
         date = as.Date(date)) %>%
  glimpse

lcc_temp_raw <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 3)
lcc_temp <- lcc_temp_raw %>% 
  rename(date = DT,
         temp_c = TEMP_C) %>% 
  mutate(stream = "clear creek", 
         site = "lcc",
         date = as.Date(date)) %>%
  glimpse

clear_format <- bind_rows(lcc_temp, ucc_temp) %>% 
  group_by(date, stream, site) %>%
  summarise(mean_temp_degC = mean(temp_c),
            max_temp_degC = max(temp_c)) %>% 
  mutate(source = "FWS") %>% glimpse

summary(clear_format$date)

clear_format %>% ggplot() + 
  geom_line(aes(x = date, y = mean_temp_degC, color = site)) + 
  # geom_point(aes(x = date, y = mean_temp_degC), size = .001) +
  theme_minimal()
```

## Deer Creek

-   Deer Creek data was aggregated from CDEC's site `DCV` from
    `1998-10-01` to `2022-07-13`

-   TODO: add max temperature

```{r}
# source USGS
deer <- read_rds(here::here("data-raw", "standard-format-data-prep","temp_data", "deer_temp.rds"))
deer_format <- deer %>%
  mutate(stream = "deer creek",
         site = stream,
         source = "CDEC DCV") %>%
  filter(mean_temp_degC > 0 & mean_temp_degC < 35) %>% 
  glimpse()

summary(deer_format$date)

deer_format %>% ggplot() + 
  geom_line(aes(x = date, y = mean_temp_degC), color = "gray") + 
  theme_minimal()
```

## Feather River

-   Chose to use RST environmental data instead of CDEC or USGS because
    a significant data gap exists in publicly available data between
    2006-09-05 and 2018-09-30. The two publicly available sites are also
    spatially far apart.

-   Feather River data can be aggregate by high and low flows using the
    `location` column

```{r}
# Two gages according to RST shiny app: 
# - CDEC GRL
# - CDEC FRA
#cdec_datasets("GRL")
#GRL_CDEC <- cdec_query(station = "GRL", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")

#save(GRL_CDEC, file = "temp_data/GRL_cdec.Rdata")
load('temp_data/GRL_cdec.Rdata')

feather_river_temp_GRL <- GRL_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC) %>%
  group_by(date) %>%
  summarise(mean_temp_degC = mean(temp_degC, na.rm = TRUE)) %>% 
  mutate(stream = "feather river", 
         site = "GRL", 
         source = "CDEC") %>%
  filter(mean_temp_degC < 23 & mean_temp_degC > 2) # remove outlier

# ggplot() + 
#   geom_line(data = feather_river_temp_GRL, aes(x = date, y = mean_temp_degC)) +
#   theme_minimal()
# 
# summary(feather_river_temp_GRL$date)

# FRA
#cdec_datasets("FRA")

# FRA_CDEC <- cdec_query(station = "FRA", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")

#save(FRA_CDEC, file = "temp_data/FRA_cdec.Rdata")
load('temp_data/FRA_cdec.Rdata')

feather_river_temp_FRA <- FRA_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC) %>%
  group_by(date) %>%
  summarise(mean_temp_degC = mean(temp_degC, na.rm = TRUE)) %>% 
  mutate(stream = "feather river", 
         site = "FRA", 
         source = "CDEC") 

# ggplot() + 
#   geom_line(data = feather_river_temp_FRA, aes(x = date, y = mean_temp_degC)) +
#   theme_minimal()
# 
# summary(feather_river_temp_FRA$date)
```

### Combine Gage data for Feather River

-   significant data gap exists

```{r}

feather_cdec_gages  <- bind_rows(feather_river_temp_FRA,
                             feather_river_temp_GRL)

ggplot() +
  geom_line(data = feather_cdec_gages, aes(x = date, y = mean_temp_degC, color = site))

```

#### RST Environmental Data for Feather River 

-   Chose to use this dataset since it is more complete

-   Data can be aggregated up to high and low flows using `location`
    column

```{r}
# low flow:
low_flow <- c('steep riffle', 'eye riffle', 'gateway riffle')

# high flow
high_flow <- c('herringer riffle', 'live oak', 'sunset pumps', 'shawns beach')

feather_rst <- rst_enviro %>% 
  filter(stream == "feather creek") %>%
  select(stream, site, date, temperature) %>%
  rename(mean_temp_degC = temperature) %>%
  mutate(location = case_when(site %in% low_flow ~ "low flow", 
                              site %in% high_flow ~ 'high_flow'),
         source = "RST environmental") %>%
  filter(mean_temp_degC > 0) %>%
  glimpse

ggplot() +
  geom_line(data = feather_rst, aes(x = date, y = mean_temp_degC, color = site)) + 
  facet_wrap(~location) +
  theme_minimal()
```

## Mill Creek

-   Mill Creek temperature data is sourced from CDEC's site `MLM` from
    `1998-10-07` to `2022-07-13`

```{r}
# source USGS
mill <- read_rds(here::here("data-raw", "standard-format-data-prep","temp_data", "mill_temp.rds"))
mill_format <- mill %>%
  mutate(stream = "mill creek",
         site = stream, 
         source = "CDEC MLM") %>%
   filter(mean_temp_degC > 0 & mean_temp_degC < 35) %>% 
glimpse()

summary(mill_format$date)

mill_format %>% ggplot() + 
  geom_line(aes(x = date, y = mean_temp_degC), color = "gray") + 
  # geom_point(aes(x = date, y = mean_temp_degC), size = .001) +
  theme_minimal()
```

## Sacramento

-   combined dataset spans `1998-10-01` to `2022-07-25` and includes RST
    environmental data from Knights Landing and USGS data from Tisdale

#### Gage Data for Tisdale

-   Data is from `USGS 11390500` and exists from `1998-10-01` to
    `2022-07-25` with a significant data gap between `1998-09-30` and
    `2016-10-01` .

```{r}

### USGS 11390500
# Read created CSV
sac_tisdale_usgs <- read_csv("WLK_USGS.csv")

# Format to make tidier
sac_tisdale_usgs_format <- sac_tisdale_usgs %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  as_tibble() %>% 
  rename(date = Date,
         mean_temp_degC = temp_degC) %>% 
  mutate(stream = "sacramento river", 
         site = "USGS 11390500", 
         source = "USGS") %>%
  #filter(year(date) > 2000) %>% # filter to greater than 2000
  glimpse

# Plot
ggplot() +
  geom_line(data = sac_tisdale_usgs_format, aes(x = date, y = mean_temp_degC), color = "black") +
  labs(x = "date",
       y = "Temperature (deg C)", 
       caption = "USGS 11390500")
```

#### Sacramento RST Environmental Data 

-   One site: `knights landing`

-   Date range from `2002-10-04` to `2021-05-31` with a two year gap
    between `2016-06-20` and `2018-08-28`

```{r}

sac_rst_format <- rst_enviro %>% 
  filter(stream == "sacramento river") %>% 
  select(stream, site, date, temperature) %>%
  rename(mean_temp_degC = temperature) %>%
  mutate(source = "RST environmental") %>%
  glimpse

summary(sac_rst_format$date)

ggplot() +
  geom_line(data = sac_rst_format, aes(x = date, y = mean_temp_degC))

```

#### Combine Sacramento data

-   combined dataset spans `1998-10-01` to `2022-07-25` and includes RST
    environmental data from Knights Landing and USGS data from Tisdale

```{r}

sac_format <- bind_rows(sac_rst_format, # rst
                        sac_tisdale_usgs_format) %>% glimpse # usgs

summary(sac_format$date)

ggplot() +
  geom_line(data = sac_format, aes(x = date, y = mean_temp_degC, color = site)) 
  #facet_wrap(~site)

```

## Yuba

In progress... 

```{r eval=FALSE, include=FALSE}
#TODO need to do air temp modeling, check temperature_data_prep for potential gages (if not there use noaa data)
# TODO: pull temperature from RST data 

yuba_rst_format <- rst_enviro %>% 
  filter(stream == "yuba river") %>% 
  select(stream, site, date, temperature) %>%
  rename(mean_temp_degC = temperature) %>%
  mutate(source = "RST environmental") %>%
  glimpse

summary(yuba_rst_format$date)

ggplot() +
  geom_line(data = yuba_rst_format, aes(x = date, y = mean_temp_degC))
```

<!-- #### Other data available? -->

<!-- * CDEC YRS - spatially far away from RST -->

```{r, eval=FALSE, include=FALSE}
# limited USGS gage data 
# USGS 11417500 S YUBA R A JONES BAR NR GRASS VALLEY CA
# site_no=11417500

#cdec_datasets('YRS')
cdec_yrs <- cdec_query(station = "YRS", dur_code = "H", sensor_num = "25", start_date = "1996-01-01") 

## TODO: convert to C from F
cdec_yrs_format <- cdec_yrs %>% 
  rename(source = agency_cd,
         site = location_id, 
         temperature_c = parameter_value,
         date = datetime) %>%
  mutate(date = as_date(date),
         temperature_c = (temperature_c-32 * 5/9)) %>%
  group_by(date) %>%
  mutate(mean_temp_degC = mean(temperature_c, na.rm = TRUE),
         max_temp_degC = max(temperature_c, na.rm = TRUE)) %>%
  filter(mean_temp_degC > 0) %>%
  glimpse
  
summary(cdec_yrs_format)
#save(cdec_yrs_format, file = "temp_data/yrs_cdec.Rdata")
load('temp_data/yrs_cdec.Rdata')


ggplot() +
  geom_line(data = cdec_yrs_format, aes(x = date, y = mean_temp_degC))
```

<!-- #### Compare to Air Data Linear Regression -->

```{r, eval=FALSE, include=FALSE}

yuba_monthly_agg <- yuba_rst_format %>% 
  mutate(year = year(date), month = month(date)) %>%
  group_by(year, month) %>%
  summarise(mean_temp_degC = mean(mean_temp_degC, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(date = ymd(paste(year, month, '01', sep = '-'))) %>%
  filter(!is.na(date)) %>%
  select(date, mean_temp_degC) %>% glimpse

yuba_air <- read_rds(here::here("data-raw", "standard-format-data-prep","temp_data", "yuba_temp_air.rds")) %>% 
  mutate(year = year(date), month = month(date)) %>%
  group_by(year, month) %>%
  summarise(air_tempC = mean(air_tempC, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(date = ymd(paste(year, month, '01', sep = '-'))) %>%
  filter(!is.na(date)) %>%
  select(date, air_tempC) %>%
  glimpse

ggplot(yuba_air, aes(x=date, y = air_tempC)) +
  geom_line(color = "darkred") +
  labs(y = "Air temperature (deg C)") +
  theme(axis.text = element_text(size = 12))

# combine water and air

yuba_air_water <- yuba_monthly_agg %>%
  left_join(yuba_air) %>%
  filter(!is.na(air_tempC)) %>% glimpse()

yuba_air_water %>%
  ggplot(aes(x = air_tempC, y = mean_temp_degC)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)

yuba_temp_model <- lm(mean_temp_degC ~ air_tempC, data = yuba_air_water)
summary(yuba_temp_model)

yuba_temp_model$coefficients

# air temp thresholds
# MR: I don't really understand this section 
y <- c(13, 14)
air_temp_thresholds <- (y - yuba_temp_model$coefficients[[1]]) / yuba_temp_model$coefficients[[2]]

yuba_air_water %>%
  ggplot(aes(x = air_tempC, y = mean_temp_degC)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  geom_vline(xintercept = air_temp_thresholds[1], alpha = .2) +
  geom_vline(xintercept = air_temp_thresholds[2], alpha = .2) +
  geom_hline(yintercept = 13) +
  geom_hline(yintercept = 14)

pred <- broom::augment(yuba_temp_model) %>% pull(.fitted)
truth <- yuba_monthly_agg$mean_temp_degC
#xtab <- table(pred > 11, truth > 11)
#xtab <- table(pred > 12, truth > 12)

# new stuff
yuba_air_more_time <- yuba_air %>%
  bind_rows(
    tibble(date = seq.Date(ymd('2000-01-01'), ymd('2020-12-01'), by = 'month'),
           air_tempC = 0)
  ) %>%
  group_by(date) %>%
  summarise(air_tempC = max(air_tempC, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(air_tempC = ifelse(air_tempC == 0, NA, air_tempC)) %>% glimpse


ts_yuba <- ts(yuba_air_more_time$air_tempC, start = c(2000, 1), end = c(2020, 12), frequency = 12)

na.interp(ts_yuba) %>% autoplot(series = 'Interpolated') +
  forecast::autolayer(ts_yuba, series = 'Original')

yuba_air_temp_c <- tibble(
  date = seq.Date(ymd('2000-01-01'), ymd('2020-12-01'), by = 'month'),
  air_tempC = as.numeric(na.interp(ts_yuba)))


yuba_air_temp_c %>%
  ggplot(aes(x = date, y = air_tempC)) +
  geom_col(fill = 'darkgoldenrod2') +
  geom_col(data = yuba_air_temp_c, aes(x = date, y = air_tempC)) +
  theme_minimal() +
  labs(y = 'monthly mean air temperature (°C)')

yuba_pred_water_temp <- predict(yuba_temp_model, yuba_air_temp_c)

yuba_water_temp_c <- tibble(
  date = seq.Date(ymd('2000-01-01'), ymd('2020-12-01'), by = 'month'),
  stream = 'yuba river',
  mean_temp_degC = yuba_pred_water_temp) %>% 
  mutate(source = "predicted water temperature") %>%
  glimpse

yuba_water_temp_c %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = mean_temp_degC), color = 'red') +
  geom_line(data = yuba_air_temp_c, aes(y = air_tempC))

summary(yuba_water_temp_c)

yuba_format <-  yuba_water_temp_c

```

# Combine temperature data {.tabset}

```{r}
combined_temp <- bind_rows(battle_format,
                           butte_format,
                           clear_format,
                           deer_format,
                           feather_rst,
                           mill_format,
                           sac_format,
                           #yuba_format,
                           #yuba_rst_format
                           ) %>%
  rename(mean_daily_temp_c  = mean_temp_degC) %>%
  glimpse()
```

## QA/QC

```{r}
unique(combined_temp$stream)
unique(combined_temp$site)
ggplot(combined_temp, aes(x = date, y = mean_daily_temp_c, color = site)) +
  theme_minimal() +
  geom_line() +
  facet_wrap(~stream)
filter(combined_temp, is.na(date))
combined_flow_clean <- filter(combined_temp, !is.na(date))
```

#TODO save to cloud in standard format data

```{r, eval = FALSE}
# f <- function(input, output) write_csv(input, file = output)
# gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
# gcs_upload(combined_flow_clean,
#            object_function = f,
#            type = "csv",
#            name = "standard-format-data/standard_temp.csv",
#            predefinedAcl = "bucketLevel")
# write_csv(combined_flow_clean, "../../data/standard-format-data/standard_temp.csv")
```
