---
title: "Covariate processing for stock recruit model"
author: "ashley vizek"
output: 
  html_document:
    theme: flatly
date: "2023-12-21"
---

```{r setup, include=FALSE}
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = F, messages = F, warning = F)

library(tidyverse)
library(zoo)
library(googleCloudStorageR)
library(gsheet)
library(kableExtra)
```

# Overview

The goal of this document is to process covariates for use in SR JPE stock recruitment modeling. Initial literature review and planning has been completed to outline covariates that are expected to be important based on past research. This work can be found [here](https://docs.google.com/spreadsheets/d/1Q4VUBE72KdPq0x65y_vUoDMj5dDHp8XwOyDNIyIJKpE/edit#gid=0).

*Table 1. Summary of covariates to include in stock recruit modeling based on preliminary literature review*

```{r}
covariate_references <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Q4VUBE72KdPq0x65y_vUoDMj5dDHp8XwOyDNIyIJKpE/edit#gid=0")

kable(covariate_references) |> 
  kable_classic(full_width = T)
```

# Temperature

Temperature has been found to influence spawning and rearing. Therefore it is expected to affect the translation from stock to recruit. Temperature can be included in multiple different formats including:

- Number of degree days
- Day of year when 7DADM is above 13 C
- Maximum weekly stream temperature
- *In development: Emergence date (calculated from temperature and spawning date)*

```{r, include=F}
# google cloud set up
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_get_object(object_name = "standard-format-data/standard_temperature.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard-format-data/standard_temperature.csv",
               overwrite = TRUE)
standard_temperature <- read_csv("data/standard-format-data/standard_temperature.csv")
```

## Number of degree days

Sum of the daily mean temperatures between August and December (spawning time period) by year and stream. 

*Note that for streams with multiple locations the max daily mean was selected.*

```{r, include = F}
# 1. Filter data to the spawning and incubation time period
# 2. Calculate the DD (use 0 as threshold for now)

# ck_complete_years <- standard_temperature |> 
#   group_by(year(date), stream) |> 
#   summarize(min = min(date),
#             month_min = month(min)) |> 
#   filter(month_min > 8)
# Need to remove following years because incomplete: 1997 feather river, 1998 deer mill, 1999 butte and yuba
std_temp_dd_spawn <- standard_temperature |> 
  mutate(year = year(date),
         rm = case_when(stream %in% c("feather river") & year == 1997 ~ "incomplete",
                        stream %in% c("deer creek", "mill creek") & year == 1998 ~ "incomplete",
                        stream %in% c("butte creek", "yuba river") & year == 1999 ~ "incomplete",
                        T ~ "complete")) |> 
  filter(month(date) %in% c(8:12), stream != "sacramento river", rm != "incomplete") |> # filter to the spawning and incubation time period, not including sacramento because spawning is happening in tributaries
  select(date, year, stream, site, subsite, statistic, value) |> 
  pivot_wider(id_cols = c(date, year, stream, site, subsite), values_from = "value", names_from = "statistic") |> 
  group_by(date, stream, year) |> 
  summarize(max_mean = max(mean, na.rm = T)) |> 
  group_by(year, stream) |> 
  summarize(gdd_spawn = sum(max_mean)) |> glimpse()
```


```{r}

std_temp_dd_spawn |> 
  ggplot(aes(x = year, y = gdd_spawn, color = stream)) + 
  geom_line() +
  labs(x = "",
       y = "Sum of daily mean temperatures during spawning/incubation time period (Aug-Dec)") +
  theme_minimal()

# years for each stream, data are complete
# std_temp_dd_spawn |> 
#   ggplot(aes(x = year, y = stream)) + 
#   geom_point()
```

*Figure 1. Annual sum of daily mean temperatures during spawning and incubation time period (Aug - Dec)*

## Day of year when 7DADM drops above/below 13 C

Day of the year and week when the 7DADM is above 13C

- Data are not filtered to the spawning time period (Aug-Dec) as this threshold is typically met before August
- Below 13C is not included as this does not make sense in terms of impact on salmon. Temperatures are rarely below 13C during the spawning time period and if they are, unlikely to be significantly cold enough to have an impact
- TODO - Currently do not have max daily temperatures for the Sacramento so mainstem is not included
 
```{r, include = F}
# 1. hypothesized for spawning so select spawning time period 
# 2. calculate the 7DADM
# 3. identify the min date when drops below/above 13 C

std_temp_7dadm <- standard_temperature |> 
  filter(statistic == "max") |>  # can only calculate using the max temperature
  #filter(month(date) %in% c(8:12)) |> 
  mutate(year = year(date)) |> # need to add year as grouping so that the rollingmean will apply to each year separately
  group_by(year, stream, site, subsite) |> 
  arrange(stream, site, subsite, date) |> 
  mutate(roll_7 = rollapply(value, 7, mean, align = "center", fill = NA))

# Use this to summarize rolling mean by week to fill in Nas
std_temp_7dadm_week <- std_temp_7dadm |> 
  mutate(week = week(date)) |> 
  group_by(week, stream, site, subsite, year) |> 
  summarize(mean_7dadm = mean(roll_7, na.rm = T))

# Fill in Nas with weekly mean
std_temp_7dadm_interpolated <- std_temp_7dadm |> 
  mutate(week = week(date)) |> 
  left_join(std_temp_7dadm_week) |> 
  mutate(roll_7 = ifelse(is.na(roll_7), mean_7dadm, roll_7))

#ck <- filter(std_temp_7dadm_interpolated, is.na(roll_7))
std_temp_7dadm_above_13 <- std_temp_7dadm_interpolated |> 
  filter(roll_7 > 13) |> 
  group_by(year, stream) |> 
  slice_min(date) |> 
  select(date, year, week, stream) |> 
  rename(above_13 = week) |> glimpse()

std_temp_7dadm_below_13 <- std_temp_7dadm_interpolated |> 
  filter(roll_7 < 13) |> 
  group_by(year, stream) |> 
  slice_min(date) |> 
  select(date, year, week, stream) |> 
  rename(below_13 = week)
```

**Plot of week where threshold is met (above or below 13) for all years and streams**

```{r}
# Explore data
std_temp_7dadm_threshold <- full_join(std_temp_7dadm_above_13, 
                                      std_temp_7dadm_below_13) |> 
  pivot_longer(cols = c(above_13, below_13), values_to = "week", names_to = "threshold")
ggplot(std_temp_7dadm_threshold, aes(x = year, y = week, color = threshold)) +
  geom_point() +
  facet_wrap(~stream) +
  theme_minimal()

```

*Figure 2. Plot of week where threshold is met (above or below 13) for all years and streams*

```{r, include=F}
std_temp_7dadm_final <- std_temp_7dadm_above_13 |> 
  rename(above_13_temp_week = above_13) |> 
  mutate(above_13_temp_day = yday(date)) |> 
  select(stream, above_13_temp_day, above_13_temp_week, year) |> glimpse()
```

## Max temperature

- Summarizes the weekly maximum temperature for each stream (meaning it finds the max across all sites/subsites) across years. Note that the Sacramento River temperature data does not currently include a daily maximum so the weekly max is the max of the daily mean.
- Summarize the weekly max for each stream and year by finding the mean, median, and max.

```{r, include = F}
std_temp_weekly <- standard_temperature |> 
  filter(month(date) %in% c(8:12)) |> 
  mutate(week = week(date),
         year = year(date)) |> 
  group_by(week, year, stream) |> 
  summarize(w_max = max(value, na.rm = T)) |>  # find the max weekly temperature
  ungroup() |>
  group_by(stream, year) |> 
  summarize(weekly_max_temp_max = max(w_max, na.rm = T),
            weekly_max_temp_mean = mean(w_max, na.rm = T),
            weekly_max_temp_median = median(w_max, na.rm = T)) |> glimpse()
  
```

```{r}
std_temp_weekly |> 
  ggplot(aes(x = year, y = weekly_max_temp_max)) +
  geom_line() +
  facet_wrap(~stream) +
  labs(x = "",
       y = "Weekly max temperatures (Celcius)") +
  theme_minimal()
```

*Figure 3. Plot of mean, median, max of the maximum weekly temperature*

## Emergence date

*In development*

Emergence data encompasses the effect of temperature and is expected to affect growth and survival of juveniles. Modeled following [Kaylor et al. 2022](https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecs2.4160). Daily Ei values are summed from the spawn date, and emergence is assumed to occur on the first day where Ei exceeds 1.

Ei = 1 / exp [loge a - loge(Ti - b)]

logea = 6.872
b = -0.332
Ti = mean daily temperature
Ei = daily contribution to development (ranging from 0 to 1)

Constants derived from [Beacham and Murray (1990)](https://afspubs.onlinelibrary.wiley.com/doi/epdf/10.1577/1548-8659%281990%29119%3C0927%3ATESADO%3E2.3.CO%3B2)

# Flow

Precipitation and streamflow have been found to influence spawning and rearing. Therefore it is expected to affect the translation from stock to recruit. Streamflow can be included in multiple different formats including:
- Mean, median, max precipitation (or streamflow)

```{r, include=F}
gcs_get_object(object_name = "standard-format-data/standard_flow.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard-format-data/standard_flow.csv",
               overwrite = TRUE)
standard_flow <- read_csv("data/standard-format-data/standard_flow.csv")
```

## Mean, median, max flow

```{r, include = F}
std_flow_spawn_summary <- standard_flow |> 
  mutate(year = year(date)) |> 
  filter(month(date) %in% c(8:12), stream != "sacramento river") |> 
  group_by(date, year, stream) |> 
  summarize(flow_cfs = max(flow_cfs)) |> 
  group_by(year, stream) |> 
  summarize(mean = mean(flow_cfs, na.rm = T),
            median = median(flow_cfs, na.rm = T),
            max = max(flow_cfs, na.rm = T),
            min = min(flow_cfs, na.rm = T)) |> 
  pivot_longer(cols = mean:min, values_to = "value", names_to = "statistic") |> 
  mutate(value = ifelse(is.nan(value) | is.infinite(value), NA_real_, value))
```

```{r}
std_flow_spawn_summary |> 
  ggplot(aes(x = year, y = value, color = statistic)) +
  geom_line() +
  facet_wrap(~stream, scales = "free_y")
```

*Figure 4. Plot of annual summarized (mean, median, max, min) max temperatures during the spawning and incubation time period (Aug - Dec)*

```{r, include = F}
std_flow_rear_summary <- standard_flow |> 
  mutate(year = ifelse(month(date) %in% c(10:12), year(date) + 1, year(date))) |> 
  filter(month(date) %in% c(11:12, 1:7), stream != "sacramento river") |> 
  group_by(date, year, stream) |> 
  summarize(flow_cfs = max(flow_cfs)) |> 
  group_by(year, stream) |> 
  summarize(mean = mean(flow_cfs, na.rm = T),
            median = median(flow_cfs, na.rm = T),
            max = max(flow_cfs, na.rm = T),
            min = min(flow_cfs, na.rm = T)) |> 
  pivot_longer(cols = mean:min, values_to = "value", names_to = "statistic") |> 
  mutate(value = ifelse(is.nan(value) | is.infinite(value), NA_real_, value))
```

```{r}
std_flow_rear_summary |> 
  ggplot(aes(x = year, y = value, color = statistic)) +
  geom_line() +
  facet_wrap(~stream, scales = "free_y")
```

*Figure 5. Plot of annual summarized (mean, median, max, min) max temperatures during the rearing time period (Nov - July)*

```{r, include = F}
std_flow_final <- std_flow_spawn_summary |> 
  pivot_wider(id_cols = c(year, stream), values_from = "value", names_from = "statistic") |> 
  rename(mean_flow = mean,
         median_flow = median,
         max_flow = max,
         min_flow = min) |> 
  mutate(lifestage = "spawning and incubation") |> 
  bind_rows(std_flow_rear_summary |> 
  pivot_wider(id_cols = c(year, stream), values_from = "value", names_from = "statistic") |> 
  rename(mean_flow = mean,
         median_flow = median,
         max_flow = max,
         min_flow = min) |> 
  mutate(lifestage = "rearing")) |> glimpse()
  
```

```{r, include = F}

# Combine covariate data
stock_recruit_covariates <- std_temp_dd_spawn |> 
  mutate(lifestage = "spawning and incubation",
         covariate_type = "temperature") |> 
  full_join(std_temp_7dadm_final |> 
              mutate(lifestage = NA_character_,
                     covariate_type = "temperature")) |> 
  full_join(std_temp_weekly |> 
              mutate(lifestage = "spawning and incubation",
                     covariate_type = "temperature")) |> 
  full_join(std_flow_final |> 
              mutate(covariate_type = "flow")) |>  
  pivot_longer(cols = c(gdd_spawn, above_13_temp_day, above_13_temp_week, weekly_max_temp_max, weekly_max_temp_mean, weekly_max_temp_median, mean_flow, max_flow, min_flow, median_flow), values_to = "value", names_to = "covariate_structure") |> 
  filter(!is.na(value)) |> 
    glimpse()
```

```{r, include = F}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(stock_recruit_covariates,
           object_function = f,
           type = "csv",
           name = "jpe-model-data/stock_recruit_covariates.csv",
           predefinedAcl = "bucketLevel")
write_csv(stock_recruit_covariates, "data/model-data/stock_recruit_covariates.csv")
```

