---
title: "pull flow data for trap locations"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(CDECRetrieve)
library(dataRetrieval)
library(hms)
```

# Trap location and flow gage lookup

```{r}
# We need flow data for each of the sites included in the trap_locations table
# for all of the years listed.
# We do not need to go to subsite scale of resolution.
# For now this table is for reference to find the years that data is needed.
# This table is also helpful for understanding years where we need flow data for each location:
# https://github.com/FlowWest/JPE-datasets/tree/main/data-raw/qc-markdowns/rst

# trap_locations_raw <- read_csv("data-raw/standard-format-data-prep/rst_sites.csv")
trap_locations_raw <- read_csv("rst_sites.csv")

trap_locations <- trap_locations_raw %>%
  select(tributary, site_name, river_location, latitude, longitude, year) %>%
  distinct()
  # identify years

# gage_locations_raw <- read_csv("data-raw/standard-format-data-prep/gage_sites.csv")
gage_locations_raw <- read_csv("gage_sites.csv")

# Pull flow data from each of the gage sites here 
gage_locations <- gage_locations_raw %>%
  filter(flow_gage == T) %>%
  select(tributary, site_name, agency, latitude, longitude, identifier) %>%
  distinct() %>%
  rename(gage_latitude = latitude,
         gage_longitude = longitude)
  # get codes
```

# Pull in flow data {.tabset}

## Battle Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("BAT") 

BAT_CDEC <- cdec_query(station = "BAT", dur_code = "H", sensor_num = "20", start_date = "2003-01-01")

# Format data
BAT_hourly_flows <- BAT_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
BAT_USGS <- readNWISdv(11376550, "00060")

# Format to make tidier
BAT_daily_flows <- BAT_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 2003) %>%
  as_tibble() %>% 
  rename(date = Date)
```
#### QA/QC

```{r}
# plots to check the quality of data. no data below zero.
ggplot(BAT_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
```

```{r}
# Clean negative values
BAT_hourly_flows_clean <- BAT_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

BAT_daily_flows_clean <- BAT_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

ggplot(BAT_daily_flows_clean, aes(x = date, y = flow_cfs)) +
  geom_line()

```
#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
BAT_combined = BAT_hourly_flows_clean %>% 
  full_join(BAT_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(BAT_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
BAT_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

## Butte Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("BCK") 

BCK_CDEC <- cdec_query(station = "BCK", dur_code = "H", sensor_num = "20", start_date = "1995-01-01")

# Format data
BCK_hourly_flows <- BCK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
BCK_USGS <- readNWISdv(11372000, "00060")

# Format to make tidier
BCK_daily_flows <- BCK_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
# Clean negative values
BCK_hourly_flows_clean <- BCK_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

BCK_daily_flows_clean <- BCK_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
BCK_combined = BCK_hourly_flows_clean %>% 
  full_join(BCK_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(BCK_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
BCK_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

## Clear Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)
 
# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("IGO") 

IGO_CDEC <- cdec_query(station = "IGO", dur_code = "H", sensor_num = "20", start_date = "2003-01-01")

# Format data
IGO_hourly_flows <- IGO_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
IGO_USGS <- readNWISdv(11372000, "00060")

# Format to make tidier
IGO_daily_flows <- IGO_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 2003) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
# Clean negative values
IGO_hourly_flows_clean <- IGO_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

IGO_daily_flows_clean <- IGO_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
IGO_combined = IGO_hourly_flows_clean %>% 
  full_join(IGO_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(IGO_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
IGO_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

## Deer Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both

cdec_datasets("DCV") 

DCV_CDEC <- cdec_query(station = "DCV", dur_code = "H", sensor_num = "20", start_date = "1995-01-01")

# Format data
DCV_hourly_flows <- DCV_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
  
```

### USGS data

```{r}
DCV_USGS <- readNWISdv(11383500, "00060")

# Format to make tidier
DCV_daily_flows <- DCV_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
# Clean negative values
DCV_hourly_flows_clean <- DCV_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

DCV_daily_flows_clean <- DCV_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
DCV_combined = DCV_hourly_flows_clean %>% 
  full_join(IGO_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(DCV_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

#### USGS vs CDEC Mean data

```{r}
DCV_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

## Feather River

### High flow channel sites

```{r}
# We already identified the code for the CDEC gage of interest here: GRL
# Use this code to see what data are available
cdec_datasets("GRL") 
# For reference - this mapping tool shows CDEC sites: https://cdec.water.ca.gov/webgis/?appid=cdecstation

# Use this code to pull the data from CDEC. 
gridley_CDEC <- cdec_query(station = "GRL", dur_code = "H", sensor_num = "20", start_date = "1996-01-01")

# Format data
gridley_hourly_flows <- gridley_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
  
```

#### QA/QC 

It looks like there are some negative values. We should make these NA.

It would be important to note any gaps in data collection.

```{r}
ggplot(gridley_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
```
#### Clean data

```{r}
gridley_hourly_flows_clean <- gridley_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))
```

### Low flow channel sites

```{r}
# For reference - this mapping tool shows USGS sites: https://maps.waterdata.usgs.gov/mapper/index.html
# You can use the mapper to search for sites and find the right parameter code. I use 0060 below for daily discharge. 
# Hourly or instaneous would be better but it is not available.
# Documentation for dataRetrieval: https://github.com/USGS-R/dataRetrieval

oroville_USGS <- readNWISdv(11407000, "00060")

# Format to make tidier
oroville_hourly_flows <- oroville_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1996) %>%
  as_tibble() %>% 
  rename(date = Date)
```


#### QA/QC

```{r}
ggplot(oroville_daily_flows, aes(x = date, y = flow_cfs)) +
  geom_line()
```

```{r}
# Clean negative values
gridley_hourly_flows_clean <- gridley_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

oroville_daily_flows_clean <- oroville_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
gridley_combined = gridley_hourly_flows_clean %>% 
  full_join(oroville_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(gridley_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

#### USGS vs CDEC Mean data

```{r}
gridley_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

#### USGS vs CDEC Mean data

```{r}
gridley_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

## Mill Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both

cdec_datasets("MLM") 

MLM_CDEC <- cdec_query(station = "MLM", dur_code = "H", sensor_num = "20", start_date = "1996-01-01")

# Format data
MLM_hourly_flows <- MLM_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
  
```

### USGS data

```{r}
MLM_USGS <- readNWISdv(11381500, "00060")

# Format to make tidier
MLM_daily_flows <- MLM_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1996) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
# Clean negative values
MLM_hourly_flows_clean <- MLM_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

MLM_daily_flows_clean <- MLM_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
MLM_combined = MLM_hourly_flows_clean %>% 
  full_join(MLM_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(MLM_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
MLM_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
## Yuba River

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("MRY") 

MRY_CDEC <- cdec_query(station = "MRY", dur_code = "H", sensor_num = "20", start_date = "1999-01-01")

# Format data
MRY_hourly_flows <- MRY_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
MRY_USGS <- readNWISdv(11421000, "00060")

# Format to make tidier
MRY_daily_flows <- MRY_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1999) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
# Clean negative values
MRY_hourly_flows_clean <- MRY_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

MRY_daily_flows_clean <- MRY_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
MRY_combined = MRY_hourly_flows_clean %>% 
  full_join(MRY_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(MRY_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
MRY_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

## Sacramento River - Knights Landing

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("WLK") 

WLK_CDEC <- cdec_query(station = "WLK", dur_code = "H", sensor_num = "20", start_date = "1995-01-01")

# Format data
WLK_hourly_flows <- WLK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
WLK_USGS <- readNWISdv(11390500, "00060")

# Format to make tidier
WLK_daily_flows <- WLK_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
# Clean negative values
WLK_hourly_flows_clean <- WLK_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

WLK_daily_flows_clean <- WLK_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
WLK_combined = WLK_hourly_flows_clean %>% 
  full_join(WLK_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(WLK_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

#### USGS vs CDEC Mean data

```{r}
WLK_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

## Sacramento River - Tisdale

### CDEC data

```{r}
# Only CDEC data is available for Tisdale
cdec_datasets("TIS") 

TIS_CDEC <- cdec_query(station = "TIS", dur_code = "H", sensor_num = "20", start_date = "2010-01-01")

# Format data
TIS_hourly_flows <- TIS_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)

```

#### QA/QC

```{r}
# plots to check the quality of data. no data below zero.
ggplot(TIS_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
```

```{r}
TIS_hourly_flows_clean <- TIS_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

ggplot(TIS_hourly_flows_clean, aes(x = datetime, y = flow_cfs)) +
  geom_line()

```


## Recommendations

Battle Creek : no significant differences
Buttle Creek : Recommended to use CDEC to not lose significant fluctuations
Clear Creek :  no significant differences
Feather River : Recommend to use CDEC to capture significant fluctuations
Mill Creek : Recommend to use USGS because CDEC has no available data from 2008-2020.
Yuba : Recommend to use USGS because CDEC has no available data from 2003-2020.
Sacramento River- Knights Landing :  no significant difference
Sacramento Tisdale : only CDEC data is available 
# ##



