---
title: "Create standard format for Holding data"
output: html_document
---
```{r, include = F}
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(knitr)
library(hms)
```

Data were checked and cleaned in scripts available [here](https://github.com/FlowWest/JPE-datasets/tree/main/data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys).
Cleaned data were saved on the jpe-dev-bucket on google cloud.

```{r, data_pull, include = F, echo = F}
# # Data pull ---------------------------------------------------------------

# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# define files to pull in (in some cases there is just one catch file while others
# have sample and environmental data)
files_battle <- tibble(path = rep("adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_holding"),
                       name = c(""),
                       save = rep("data/redd-carcass-holding/battle_holding"))
files_butte <- tibble(path = rep("adult-holding-redd-and-carcass-surveys/butte-creek/data/butte_",17),
                       name = c("holding_2001",
                                "holding_2002", "holding_2003", "holding_2004", "holding_2005", "holding_2006",
                                "holding_2007", "holding_2008", "holding_2009", "holding_2010", "holding_2011",
                                "holding_2012", "holding_2013", "holding_2014", "holding_2015", "holding_2016",
                                "holding_2017"),
                      save = "data/redd-carcass-holding/butte_")
files_clear <- tibble(path = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_holding",
                      name = c(""),
                      save = rep("data/redd-carcass-holding/clear_holding"))
files_deer <- tibble(path = rep("adult-holding-redd-and-carcass-surveys/deer-creek/data/deer_",2),
                     name = c("adult_holding_1986_to_1996",
                              "adult_holding_1997_to_2020"),
                     save = "data/redd-carcass-holding/deer_")
# function to save file to disk
get_data <- function(path, name, save) {
  gcs_get_object(object_name = paste0(path, name, ".csv"),
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = paste0(save, name, ".csv"),
                 overwrite = TRUE)
}
# apply function to each set of files
pmap(files_battle, get_data)
pmap(files_butte, get_data)
pmap(files_clear, get_data)
pmap(files_deer, get_data)


# load in data
# catch data
battle_holding <- read_csv("data/redd-carcass-holding/battle_holding.csv")
butte_holding_01 <- read_csv("data/redd-carcass-holding/butte_holding_2001.csv")
butte_holding_02 <- read_csv("data/redd-carcass-holding/butte_holding_2002.csv")
butte_holding_03 <- read_csv("data/redd-carcass-holding/butte_holding_2003.csv")
butte_holding_04 <- read_csv("data/redd-carcass-holding/butte_holding_2004.csv")
butte_holding_05 <- read_csv("data/redd-carcass-holding/butte_holding_2005.csv")
butte_holding_06 <- read_csv("data/redd-carcass-holding/butte_holding_2006.csv")
butte_holding_07 <- read_csv("data/redd-carcass-holding/butte_holding_2007.csv")
butte_holding_08 <- read_csv("data/redd-carcass-holding/butte_holding_2008.csv")
butte_holding_09 <- read_csv("data/redd-carcass-holding/butte_holding_2009.csv")
butte_holding_10 <- read_csv("data/redd-carcass-holding/butte_holding_2010.csv")
butte_holding_11 <- read_csv("data/redd-carcass-holding/butte_holding_2011.csv")
butte_holding_12 <- read_csv("data/redd-carcass-holding/butte_holding_2012.csv")
butte_holding_13 <- read_csv("data/redd-carcass-holding/butte_holding_2013.csv")
butte_holding_14 <- read_csv("data/redd-carcass-holding/butte_holding_2014.csv")
butte_holding_15 <- read_csv("data/redd-carcass-holding/butte_holding_2015.csv")
butte_holding_16 <- read_csv("data/redd-carcass-holding/butte_holding_2016.csv")
butte_holding_17 <- read_csv("data/redd-carcass-holding/butte_holding_2017.csv")
clear_holding <- read_csv("data/redd-carcass-holding/clear_holding.csv")
deer_holding_1997_2020 <- read_csv("data/redd-carcass-holding/deer_adult_holding_1997_to_2020.csv")
deer_holding_1986_1996 <- read_csv("data/redd-carcass-holding/deer_adult_holding_1986_to_1996.csv")

```

# Data formatting {.tabset}

## Battle Creek

```{r, battle_holding_data_dictionary}
battle_data_dictionary <- read_rds("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/data-dictionaries/battle_holding_data_dictionary.rds")
kable(battle_data_dictionary)
```
```{r, battle_clean}
# There are multiple observations per reach. Checked raw data and there is not a 
# field for crew name etc to indicate that there are duplicates
battle_holding_clean <- battle_holding %>%
  select(-notes) %>%
  mutate(stream = "battle creek",
         year = lubridate::year(date),
         adipose_clipped = "not recorded",
         run = "spring") # either spring or unknown, but not recorded in raw data
  

```

### Varibles removed

- notes: internal use

## Clear Creek

```{r, clear_holding_data_dictionary}
clear_data_dictionary <- read_rds("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/data-dictionaries/clear_holding_data_dictionary.rds")
kable(clear_data_dictionary)
```
```{r, clear_clean}
# There are multiple observations per reach. Checked raw data and there is not a 
# field for crew name etc to indicate that there are duplicates
clear_holding_clean <- clear_holding %>%
  select(-comments) %>%
  rename(jacks = jack_count) %>%
  mutate(stream = "clear creek",
         year = lubridate::year(date),
         run = "spring",
         adipose_clipped = "not recorded")
```

### Variables removed

- comments: internal use


## Deer Creek

```{r, deer_holding_data_dictionary}
deer_data_dictionary <- read_rds("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/data-dictionaries/deer_holding_data_dictionary.rds")
kable(deer_data_dictionary)
```

```{r, deer_clean}
deer_holding_clean <- bind_rows(deer_holding_1986_1996,
                                deer_holding_1997_2020) %>%
  rename(reach = location) %>%
  mutate(stream = "deer creek",
         date = as_date(paste0(year, "-08-01")),
         run = "spring",
         adipose_clipped = as.character(F)) # no hatchery, any adipose clipped fish would be strays
```

### Variables removed

- none

## Butte Creek

```{r, butte_holding_data_dictionary}
butte_data_dictionary <- read_rds("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/data-dictionaries/butte_holding_data_dictionary.rds")
kable(butte_data_dictionary)
```


```{r, butte_clean}
butte_holding_clean <- bind_rows(butte_holding_01 %>% select(-comments),
                                 butte_holding_02,
                                 butte_holding_03,
                                 butte_holding_04,
                                 butte_holding_05,
                                 butte_holding_06,
                                 butte_holding_07, 
                                 butte_holding_08,
                                 butte_holding_09,
                                 butte_holding_10,
                                 butte_holding_11,
                                 butte_holding_12,
                                 butte_holding_13,
                                 butte_holding_14,
                                 butte_holding_15,
                                 butte_holding_16,
                                 butte_holding_17) %>%
  group_by(date, reach) %>%
  summarize(avg_count = round(mean(fish_count, na.rm = T), digits = 0)) %>%
  rename(count = avg_count) %>% 
  mutate(stream = "butte creek",
         year = lubridate::year(date),
         run = "spring",
         adipose_clipped = "not recorded") %>% 
  glimpse()
```

### Variables removed
- comments, personnel removed
- Butte creek data has multiple observations from different personnels per day - to prevent double counting, the average of each day at each reach is used as the daily fish count

# Standard format {.tabset}
```{r, combined}
combined_holding <- bind_rows(battle_holding_clean,
                              butte_holding_clean,
                              clear_holding_clean,
                              deer_holding_clean)
```

## Save file without standardized reaches
We need all raw reaches to pull into `analysis/adult-reach-analysis.R`

# Save data
```{r, save_data}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(combined_holding,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_holding_raw_reaches.csv",
           predefinedAcl = "bucketLevel")
```
## Survey Reach Standardization

Survey reach naming conventions changed over time for several monitoring programs. A standardized reach lookup table was created in `analysis/adult-reach-analysis.R` and is merged in here to standardize all reaches.

```{r echo=TRUE, message=FALSE, warning=FALSE}
gcs_get_object(
  object_name = "jpe-model-data/standard_reach_lookup.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "standard-format-data", "standard-reach-lookup.csv"),
  overwrite = TRUE
)

standard_reach_lookup <- read_csv(here::here("data", "standard-format-data", "standard-reach-lookup.csv")) |> 
  # join sub-reach for butte
  mutate(standardized_reach = ifelse(stream == "butte creek", standardized_sub_reach, standardized_reach),
         reach = ifelse(stream == "butte creek", sub_reach, reach)) |>
  select(stream, reach, standardized_reach)

combined_holding <- combined_holding |> 
  left_join(standard_reach_lookup, by = c("stream", "reach")) |>
  select(-reach) |> 
  rename(reach = standardized_reach) |> 
  relocate(reach, .before = river_mile) |> 
  glimpse()
```


## Data dictionary

```{r, data_dictionary_prep, include = F}
colnames(combined_holding)
holding_catch_data_dictionary <-
  tibble(
    variable_name = colnames(combined_holding),
    description = c(
      "Date survey was conducted",
      "Reach number within stream",
      "River mile number",
      "Number of fish observed: Butte Creek data has multiple observations from different personnels per day - to prevent double counting, the average of each day at each reach is used as the daily fish count",
      "Number of jacks observed",
      "Latitude measurement of adult holding",
      "Longitude measurement of adult holding",
      "Stream name",
      "Year survey was conducted",
      "Whether or not adipose fin was clipped.",
      "Run designation of the surveyed fish. Nearly all are spring run - Battle Creek contains spring and unknown runs",
      "Reason survey was conducted",
      "Location of the Picket Weir - this is unique to Clear Creek and is either 7.4 or 8.2",
      "Fish above or below Picket Weir"
    ),
    encoding = c(
      "YYYY-MM-DD",
      NA,
      NA,
      NA,
      NA,
      NA,
      NA,
      NA,
      "YYYY",
      NA,
      NA,
      "august index, spawning",
      NA,
      "above, below"
    ),
    collected = c(
      "All but Deer",
      "All",
      "Battle and Clear",
      "All",
      "Battle and Clear",
      "Battle and Clear",
      "Battle and Clear",
      "All",
      "All",
      "None",
      "All",
      "Clear",
      "Clear",
      "Clear"
    )
  )
```
```{r, data_dictionary}
kable(holding_catch_data_dictionary)
```

## Exploratory plots

Total count per year for each waterbody

```{r}
combined_holding %>% 
  group_by(year, stream) %>%
  ggplot(aes(x = year, y = count, fill = stream))+
  geom_col()
```

Average count per month. This chart shows Battle, Butte and Clear; Deer does not have monthly count.

```{r}
combined_holding %>% 
  mutate(month = lubridate::month(date)) %>% 
  group_by(month, stream) %>% 
  summarize(average_count = mean(count, na.rm = T)) %>%
  ggplot(aes(x = month, y = average_count, fill = stream)) +
  geom_col()
```

Average jacks per month. Jacks is only collected at Battle and Clear.
```{r}
combined_holding %>% 
  mutate(month = lubridate::month(date)) %>% 
  group_by(month, stream) %>% 
  summarize(average_jacks = mean(jacks, na.rm = T)) %>% 
  ggplot(aes(x = month, y = average_jacks, fill = stream)) + 
  geom_col()
```

River mile distribution for Battle and Clear

```{r}
combined_holding %>% 
  filter(!is.na(river_mile)) %>% 
  ggplot(aes(x = stream, y = river_mile)) + 
  geom_boxplot(
  )
```

# QA/QC {.tabset}

# Categorical Variables

```{r, categorical_variables}
unique(combined_holding$reach)
unique(combined_holding$stream)
unique(combined_holding$survey_intent)
unique(combined_holding$picket_weir_relate)
```

# Count

Monthly count is not recorded for Deer.

Figure below shows the percentage count by month for each waterbody except Deer.
```{r, count_qaqc}

total_count <- combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(stream) %>% 
  summarize(total = sum(count, na.rm = T))

combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(stream, month) %>% 
  summarize(count = sum(count, na.rm = T)) %>% 
  left_join(total_count) %>% 
  mutate(percent = (count/total)*100) %>% 
  ggplot(aes(x= month, y = percent, fill = stream)) +
  geom_col() +
  facet_wrap(~stream, scales = "free_y")
```

*  `r round(sum(is.na(combined_holding$count))/nrow(combined_holding), 3)*100` % of values in the `count` column are NA.

## Jacks
Jacks are not recorded for Deer and Butte

Figure below show the percentage count by month for Battle and Clear.

```{r, jack_qa_qc}
total_jacks <- combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(stream) %>% 
  summarize(total = sum(jacks, na.rm = T))

combined_holding %>% 
  mutate(month = month(date)) %>% 
  group_by(stream, month) %>% 
  summarize(count = sum(jacks, na.rm = T)) %>% 
  left_join(total_jacks) %>% 
  mutate(percent = (count/total)*100) %>% 
  ggplot(aes(x= month, y = percent, fill = stream)) +
  geom_col() +
  facet_wrap(~stream, scales = "free_y")
```

*  `r round(sum(is.na(combined_holding$jacks))/nrow(combined_holding), 3)*100` % of values in the `jacks` column are NA.

# Save data
```{r, save_data}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(combined_holding,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_holding.csv",
           predefinedAcl = "bucketLevel")

```

