---
title: "[Watershed] [Data type] [QC]"
author: "Badhia Yunes Katz"
date: "01/25/2024"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(RODBC)
library(chron)
library(weathermetrics)
library(CDECRetrieve)
library(hms)
library(dplyr)
```

# Mill Creek [Data type long version]

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** no data gaps

**Sampling Location:** Mill Creek

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
gcs_get_object(object_name = "environmental/data-raw/MASTER_TEMPERATURE.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "MASTER_TEMPERATURE.mdb",
               overwrite = TRUE)


master_temp_deer_mill <- odbcConnectAccess2007(here::here("data-raw", "standard-format-data-prep", "MASTER_TEMPERATURE.mdb"))

```

## Data transformations

```{r}
# For different excel sheets for each year read in and combine years here

tables <- sqlTables(master_temp_deer_mill)


mill_1 <- sqlFetch(master_temp_deer_mill,"Mill 8-0 Mouth") |> 
  mutate(date = as_date(Date)) |>                                            #Transforming columns to appropriate types
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") # Remove redundant columns
# Finding Logger mean "Mill 8-0 Mouth"
mill_1_monthly_mean <- mill_1 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))


mill_2 <- sqlFetch(master_temp_deer_mill, "Mill 8-1 VIDEO Station")|> 
    mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-1 VIDEO Station"
mill_2_monthly_mean <- mill_2 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))


mill_3 <- sqlFetch(master_temp_deer_mill, "Mill 8-2 Little Mill Confluence")|> 
  mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-1 VIDEO Station"
mill_3_monthly_mean <- mill_3 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))



mill_4 <- sqlFetch(master_temp_deer_mill, "Mill 8-3 Rancheria")|>
  mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-3 Rancheria"
mill_4_monthly_mean <- mill_4 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))


mill_5 <- sqlFetch(master_temp_deer_mill, "Mill 8-4 Black Rock")|> 
  mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-4 Black Rock"
mill_5_monthly_mean <- mill_5 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))



mill_6 <- sqlFetch(master_temp_deer_mill, "Mill 8-5 Sooner")|>
  mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-5 Sooner"
mill_6_monthly_mean <- mill_6 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))




mill_7 <- sqlFetch(master_temp_deer_mill, "Mill 8-6 Canyon Camp")|> 
  mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-6 Canyon Camp"
mill_7_monthly_mean <- mill_7 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))


mill_8 <- sqlFetch(master_temp_deer_mill, "Mill 8-7 Hole in Ground")|> 
  mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-7 Hole in Ground"
mill_8_monthly_mean <- mill_8 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))


mill_9 <- sqlFetch(master_temp_deer_mill, "Mill 8-8 Little Hole in Ground")|> 
  mutate(date = as_date(Date)) |>                                            
  mutate(temperature = fahrenheit.to.celsius(Temperature, round = 1)) |> 
  mutate(year = lubridate::year(Date)) |> 
  select(-Date, -Time, -"Primary ID", -"Sort ID", -"Logger ID", -"Place ID") 
# Finding Logger mean "Mill 8-8 Little Hole in Ground"
mill_9_monthly_mean <- mill_9 |> 
  group_by(month = month(date),
           year = year(date)) |> 
  mutate(month = factor(month)) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))


#There are these tables too that we won't use to compare data because they are from CDEC stations
#Mill 8-05 MCH lower CDEC, Mill 8-15 MLM upper CDEC

combined_table_mill <- bind_rows(mill_1, mill_2, mill_3 ,mill_4, mill_5, mill_6, mill_7, mill_8, mill_9, mill_10, .id = "mill_site")

#adding CDEC data (with removed outliers from CDEC using 0-40C range)
MLM_CDEC <- cdec_query(station = "MLM", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")


MLM_hourly_temps <- MLM_CDEC |> 
  mutate(date = as_date(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) |>
  filter(temp_degC < 37, temp_degC > 0) |> #Filtered out outliers using temp range of 0-37 degrees Celsius 
  rename(temperature = temp_degC)

```

```{r}
# Snake case, 
# Columns are appropriate types



# Remove redundant columns

```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
#checking temporal coverage across deer sites from logger
mill_plot_logger <- ggplot(combined_table_mill, aes(x = date, y = temperature, color = "Logger data")) +
  geom_line(color = "darkred") +
  labs(title = "Mill temp plot from Logger data", y = "Temperature (deg C)", color = "Data Source")



```

### Variable: `Temperature`

**Plotting [Variable] over Period of Record**


```{r}
# Create the combined plot of all logger Mill sites and CDEC
combined_CDEC_logger_plot <- deer_plot_logger +
   geom_line(data = MLM_hourly_temps, aes(x = date, y = temperature, color = "CDEC data")) +
   scale_color_manual(values = c("blue", "darkred")) +
   labs(title = "Combined Plot to See Temporal coverage - Mill Logger and CDEC Data", y = "Temperature (deg C)", color = "Data Source")
print(combined_CDEC_logger_plot)

#CDEC mean temp
MLM_mean_temp <- MLM_hourly_temps |> 
  group_by(month = month(date),
           year = format(date, "%Y")) |> 
  summarise(mean_temp = mean(temperature, na.rm = TRUE))
```

```{r}
#plotting means from Mill 8-0 Mouth + CDEC 
MLM_filtered1 <- MLM_mean_temp |> 
  mutate(month = factor(month)) |> 
  filter(year >= 2013, year <= 2018)
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-0 Mouth (2013-2018)",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_1_monthly_mean, aes(x = month, y = mean_temp), fill = "red")

#plotting means from Mill 8-1 VIDEO Station + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-1 VIDEO Station",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_2_monthly_mean, aes(x = month, y = mean_temp), fill = "red")
#TODO check data from this logger 


#plotting means from Mill 8-2 Little Mill Confluence + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-2 Little Mill Confluence",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_3_monthly_mean, aes(x = month, y = mean_temp), fill = "red")


#plotting means from Mill 8-3 Rancheria + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-3 Rancheria",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_4_monthly_mean, aes(x = month, y = mean_temp), fill = "red")


#plotting means from Mill 8-4 Black Rock + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-4 Black Rock",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_5_monthly_mean, aes(x = month, y = mean_temp), fill = "red")


#plotting means from Mill 8-5 Sooner + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-5 Sooner",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_6_monthly_mean, aes(x = month, y = mean_temp), fill = "red")


#plotting means from Mill 8-6 Canyon Camp  + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-6 Canyon Camp",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_7_monthly_mean, aes(x = month, y = mean_temp), fill = "red")


#plotting means from Mill 8-7 Hole in Ground + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-7 Hole in Ground",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_8_monthly_mean, aes(x = month, y = mean_temp), fill = "red")


#plotting means from Mill 8-8 Little Hole in Ground + CDEC 
ggplot() +
  geom_boxplot(data = MLM_filtered1, aes(x = month, y = mean_temp)) +
  labs(title = "Monthly Mean Temperatures Comparison CDEC and Mill 8-8 Little Hole in Ground",
       y = "Mean Temperature (deg C)",
       x = "Month") +
  geom_boxplot(data = mill_9_monthly_mean, aes(x = month, y = mean_temp), fill = "red")


# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
# Filter clean data to show only categorical variables
```


### Variable: `Temperature`
```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Summary of identified issues

* List things that are funcky/bothering us but that we don't feel like should be changed without more investigation

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```