---
title: "Create standard format for RST hours fished"
output: html_document
---

Hours fished is not included as a variable in the data and needs to be calculated.
This file calculates hours fished by date, stream and site.

```{r, include = F, warning = FALSE, message = FALSE}
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(hms)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

Data were checked and cleaned in scripts available [here](https://github.com/FlowWest/JPE-datasets/tree/main/scripts/rst).
Cleaned data were saved on the jpe-dev-bucket on google cloud.

```{r, include = F}
gcs_get_object(object_name = "standard-format-data/standard_rst_trap.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard_rst_trap.csv",
               overwrite = TRUE)
trap <- read_csv("data/standard_rst_trap.csv")

gcs_get_object(object_name = "standard-format-data/standard_rst_catch.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard_rst_catch.csv",
               overwrite = TRUE)
catch <- read_csv("data/standard_rst_catch.csv")

# pull in hours fished for knights landing pre 2006
gcs_get_object(object_name = "rst/lower-sac-river/data/knights-landing/knl_combine_sampling_effort_clean.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/rst/knights_combine_sampling_effort_clean.csv",
               overwrite = TRUE)

knights_effort_pre2006 <- read_csv(here::here("data", "rst", "knights_combine_sampling_effort_clean.csv"))

# pull in effort for knights landing pre 2002
gcs_get_object(object_name = "rst/lower-sac-river/data/knights-landing/pre-2002-trap.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/rst/knights_pre_2002_trap.csv",
               overwrite = TRUE)

knights_effort_pre2002 <- read_csv(here::here("data", "rst", "knights_pre_2002_trap.csv"))
```

# Method

We explored three primary methods to calculate hours fished. Methods were chosen based on data collected at the RST site. When no times, trap revolutions, or trap RPMs were collected we assume 24 hours for each day that trap is fishing. The trap revolution method is commonly used by RST programs but we were not able to get reasonable results for most programs we tried it on so we ended up relying on trap times instead of revolutions. This may be due to different methods for determining cone rotation rate. For instance, Battle measures the average time per rotation whereas Feather River and Knights Landing measure the revolutions per minute. Using total revolutions and the cone rotation rate to calculate sampling period may be useful in the future when methods are more similar.  

1) Use start and stop time. 
2) Use rpms start, rpms end, and total revolutions
3) Use start time only 

Below is the code for each of the methods described above: 

1) If the start and stop date/time of the sampling period are available, simply subtract the difference in time.

```{r, helpers}
# calculating hours fished when have start and stop datetime
hours_fished <- function(dat){
  dat %>%
    filter(!is.na(trap_stop_time), !is.na(trap_start_time)) %>%
    mutate(start_datetime = ymd_hms(paste(trap_start_date, trap_start_time)),
           stop_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time)),
           hours_fished = round(difftime(stop_datetime, start_datetime, units = "hours"), 2))
}
```

2) If the RPMs start, RPMs end, and sample period revolutions are available, take the average cone RPMs and divide `sample_period_revolutions`/average RMP/60 to get total hours fished.

```{r, helpers2}
# calculating hours fished when have start and stop datetime
revolution_calculated_hours_fished <- function(dat){
  dat %>%
    filter(!is.na(rpms_start) | !is.na(rpms_end), !is.na(sample_period_revolutions)) %>% 
    mutate(prior_day_rpm = ifelse(is.na(rpms_end), NA, lag(rpms_end)),
           sum_rpms = ifelse(is.na(prior_day_rpm), rpms_start, rpms_start + prior_day_rpm), 
           cone_rpms = sum_rpms / 2,
           hours_fished = case_when(is.na(rpms_start) ~ round(sample_period_revolutions/rpms_end/60, 2),
                                    is.na(rpms_end) ~ round(sample_period_revolutions/rpms_start/60, 2), 
                                    TRUE ~ round(sample_period_revolutions/cone_rpms/60, 2)))
}
```

3) If only one date is available, assume that the end date is the date sampled on the following day.

```{r, helper3}
# calculating hours fished when have only date and time
hours_fished_one_date <- function(dat) {
  dat %>%
  arrange(site, subsite, start_datetime) %>%
  mutate(end_datetime = lead(start_datetime),
         end_datetime = case_when(difftime(end_datetime, start_datetime, units = "hours") > 120 ~ start_datetime + hours(24), 
                                  T ~ end_datetime),
         hours_fished = round(difftime(end_datetime, start_datetime, units = "hours"), 2))
}
```

## Other Assumptions

Based on protocols and information from monitoring programs, sampling periods are typically 24 hours,
unless flows are very high in which traps are checked more frequently. Calculating hours fished based
on time and date provided in data provides more specificity, however, there are typos and errors resulting
in nonsensical hours fished. In these cases, we assume 24 hour sampling periods. 

- If only a date is available and no time, assume 24 hour sampling periods for each date.
- If information is missing or there are typos, assume 24 hour sampling periods.

# Hours fished {.tabset}

## battle creek

Battle Creek has data available to calculate the hours fished. There are a few
cases in 2021 when start date is NA. For these we fill in the start date assuming
it is the sample date of the previous record.

```{r, battle}
## Battle ####
### Hours fished ####
# filter(trap, stream == "battle creek", is.na(trap_start_time))
# # these are typos that need to be fixed
# filter(trap, stream == "battle creek", trap_stop_date == "2004-02-13")
# filter(trap, stream == "battle creek", trap_stop_date == "2005-01-06")

battle_hours_fished <- filter(trap, stream == "battle creek") %>% 
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, 
         stream, site, subsite, trap_visit_id) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  mutate(hours_fished_methodology = "using start time and stop time") %>% 
  hours_fished()


knitr::kable(head(battle_hours_fished, 5))
```

### QC

```{r}
battle_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

battle_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# filter(battle_hours_fished, hours_fished > 5000) %>%  tally()
# filter(trap, stream == "battle creek", trap_stop_date == "2004-02-13")
# filter(trap, stream == "battle creek", trap_stop_date == "2005-01-06")
# filter(trap, stream == "battle creek", trap_stop_date == "2020-07-25")
# 
# # these seem large. it is possible so i will leave as is.
filter(battle_hours_fished, hours_fished > 50)
# filter(battle_hours_fished, hours_fished < 0)
# ck <- filter(trap, stream == "battle creek")

# TODO we can go into trap and fix double date at one trap 
# - would be educated guess as to which one to keep and leave
battle_hours_fished %>% filter(trap_start_date == as_date("2003-10-22"))

battle_hours_fished %>% filter(trap_start_date == as_date("2018-09-20")) # leave as is given everything looks good with start and end date
trap %>% filter(site == "ubc", trap_start_date == as_date("2018-09-20")) # everything looks fine with this data
trap %>% filter(site == "ubc", trap_start_date == as_date("2018-10-01"))# also fine here 
```
### Compare methodologies 

Tested out trap revolution methodology - more variation in hours fished. There is inconsistent data on RPMs start / RPMs end. It makes most sense to use the start date - end date method. 

```{r}
#test out using trap rotations methodology 
rev_method <- filter(trap, stream == "battle creek") %>% 
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, rpms_start, rpms_end, sample_period_revolutions, stream, site, subsite, trap_visit_id) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  mutate(hours_fished_methodology = "using rpm and revolutions") %>% 
  revolution_calculated_hours_fished() %>% glimpse()

# hours fished appear to be more spread out using this revolution method - do not link well with trap start date/trap start time 
rev_method %>% 
  ggplot(aes(x = trap_stop_date, y = hours_fished)) +
  geom_point()

rev_method %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()
```


## butte creek

Hours fished are calculated by taking the difference between the date/time at each 
observation. 

```{r, butte}
## Butte ####
### Hours fished ####
# There are no missing dates but we only have one data and time, no start/stop
# filter(trap, stream == "butte creek", is.na(trap_stop_date)) %>% 
#   tally()
# 
# filter(trap, stream == "butte creek", trap_stop_date == "2008-03-24")

butte_hours_fished <- filter(trap, stream == "butte creek") %>%
  distinct(trap_stop_date, trap_stop_time, site, subsite, stream) %>%
  mutate(start_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time))) %>%
  mutate(hours_fished_methodology = "using only start time") %>% 
  hours_fished_one_date() %>% 
  mutate(end_datetime = case_when(trap_stop_date == "2007-05-03" ~ ymd_hms("2007-05-04 08:00:00"),
                                  trap_stop_date == "2008-04-23" ~ ymd_hms("2008-04-24 07:15:00"),
                                  trap_stop_date == "2015-06-03" ~ ymd_hms("2015-06-04 08:00:00"),
                                  T ~ end_datetime),
         hours_fished = round(difftime(end_datetime, start_datetime, units = "hours"), 2),
         hours_fished = ifelse(hours_fished < 0, 24, hours_fished))

# tried treating post 2015 differnt since it is cAMP data and has start and end dates but caused a lot of huge outliers

knitr::kable(head(butte_hours_fished, 5))
```

### QC

```{r}
butte_hours_fished %>% 
  ggplot(aes(x = trap_stop_time, y = hours_fished)) +
  geom_point()


# trying to figure out where to set cut of for effort
# decided to set as 5 days - 120 hours
butte_hours_fished %>% 
  filter(hours_fished < 200) %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# butte_hours_fished %>% 
#   filter(hours_fished > 5000) %>% tally()
butte_hours_fished %>%
  filter(hours_fished > 50)
# big hour days do not seem to correspond to weekends - trap fished effort unclear here 
# TODO see if there are more dates that have 2 records for the same site
trap %>%  filter(stream == "butte creek", trap_stop_date == as_date("2014-11-28")) 
# ck <- filter(trap, stream == "butte creek")
```

## clear creek

Clear Creek has data available to calculate the hours fished. There are a few
cases when start date is NA. For these we fill in the start date assuming
it is the sample date of the previous record.

```{r, clear}
## Clear ####
### Hours fished ####

# missing some start dates
# filter(trap, stream == "clear creek", is.na(trap_start_date)) %>% 
#   tally()
# 
# filter(trap, stream == "clear creek", is.na(trap_stop_date)) %>% 
#   tally()
# 
# # swapped start and stop time for both the UCC and LCC
# filter(trap, stream == "clear creek", trap_stop_date == "2020-02-24")
# # duplicates for UCC, remove one with missing trap_start_time and missing counter
# filter(trap, stream == "clear creek", trap_stop_date == "2005-01-16")
# # typo
# filter(trap, stream == "clear creek", trap_stop_date == "2006-01-07")

clear_hours_fished <- filter(trap, stream == "clear creek") %>%
  select(stream, site, subsite, trap_start_date, trap_start_time, trap_stop_date, trap_stop_time) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  # fill in missing start dates with the stop date from previous day
  mutate(trap_start_date = case_when(is.na(trap_start_date) ~ lag(trap_stop_date),
                                T ~ trap_start_date),
         trap_start_time = case_when(is.na(trap_start_time) ~ lag(trap_stop_time), 
                                T ~ trap_start_time)) %>% 
  mutate(hours_fished_methodology = "using start time and stop time") %>% 
  hours_fished()

knitr::kable(head(clear_hours_fished, 5))
```

### QC

```{r}
clear_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

clear_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# filter(clear_hours_fished, hours_fished > 5000) %>% 
#   tally()
# filter(clear_hours_fished, hours_fished < 0) %>% 
#   tally()
# 
# filter(trap, stream == "clear creek", trap_stop_date == "2006-01-07")
```

### Compare methodologies 

Tested out trap revolution methodology - more variation in hours fished. There is inconsistent data on RPMs start / RPMs end. It makes most sense to use the start date - end date method. 

```{r}
#test out using trap rotations methodology 
rev_method <- filter(trap, stream == "clear creek") %>% 
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, rpms_start, rpms_end, sample_period_revolutions, stream, site, subsite, trap_visit_id) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  mutate(hours_fished_methodology = "using rpm and revolutions") %>% 
  revolution_calculated_hours_fished() %>% glimpse()

# hours fished appear to be more spread out using this revolution method - do not link well with trap start date/trap start time 
rev_method %>% 
  ggplot(aes(x = trap_stop_date, y = hours_fished)) +
  geom_point()

rev_method %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()
```


## deer creek

Deer Creek does not have data available (no time variable) to calculate hours
Deer creek also does not have sample period revolutions so we cannot use trap revolutions 
fished so assume 24 hours for every day with data. 

```{r, deer}
## Deer #####
deer_hours_fished <- filter(trap, stream == "deer creek") %>% 
  distinct(trap_stop_date, stream, site, subsite) %>% 
  mutate(hours_fished_methodology = "24 hour assumption") %>% 
  mutate(hours_fished = 24)

knitr::kable(head(deer_hours_fished, 5))
```

## feather river

Feather River has data available to calculate the hours fished. We only calculate
the hours fished when fish are processed. Does not look like they have any sample period revolutions so cannot try out total revolution technique. 

#TODO it does look like feather has a counter at end which should correspond to total revs (might need some manipulation)
```{r, feather}
## Feather ####

# # it looks like all cases where start date is missing and fish were processed are typos or issues with data
# filter(trap, stream == "feather river", is.na(trap_start_date) & fish_processed == "processed fish")
# # are there multiple subsites per site?
# filter(trap, stream == "feather river") %>% 
#   group_by(site) %>% 
#   distinct(subsite)
# # multiple subsites fish on same day? yes
# filter(trap, stream == "feather river") %>% 
#   group_by(trap_stop_date, site) %>% 
#   tally() %>% 
#   filter(n > 1)
# filter(trap, stream == "feather river") %>% 
# filter(trap_stop_date == "2019-01-02")
# filter(trap, stream == "feather river") %>% 
# filter(trap_stop_date == "2000-01-20")
# filter(trap, stream == "feather river") %>% 
# filter(trap_start_date == "2015-02-14")
# calculate hours fished only for period of times when trap was fishing otherwise there will be duplicates that don't make sense
# typically a trap is not processed fish when service but this happens 2x. stop date/time filled in with the visit date
feather_hours_fished <- filter(trap, stream == "feather river", 
                               # Tested out move lower so that we still know the next day/day before even if fish were not processed on that day, doesn't look like it makes a difference 
                               fish_processed %in% c("processed fish", "not recorded", "no fish caught"), include == T, 
                               visit_type != "start trapping") %>%
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, 
         trap_visit_time, trap_visit_date, stream, site, subsite) %>% 
  arrange(stream, site, subsite, trap_stop_date, trap_stop_time) %>% 
  # fix typos
  mutate(
  # fill in missing start dates with the stop date from previous day
         trap_start_date = case_when(is.na(trap_start_date) ~ lag(trap_stop_date),
                                T ~ trap_start_date),
         trap_start_time = case_when(is.na(trap_start_time) ~ lag(trap_stop_time),
                                T ~ trap_start_time)) %>%
  mutate(hours_fished_methodology = "using start time and stop time") %>% 
  hours_fished() %>%  
  select(-trap_visit_date, -trap_visit_time)

knitr::kable(head(feather_hours_fished, 5))

```

### QC

```{r}
feather_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point() 
  # geom_point(data = og_feather, aes(x = stop_datetime, y = hours_fished), color = "red", alpha = .5)

feather_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# filter(feather_hours_fished, hours_fished > 1000)%>% 
#   tally()
# filter(feather_hours_fished, hours_fished < 0) %>% 
#   tally()
# filter(feather_hours_fished, is.na(hours_fished))%>% 
#   tally()

```


## knights landing

Knights Landing has date and time data available to calculate the hours fished. We do not have sample_period_revolutions so we cannot try out the trap revolution method. 

```{r, knights}
## Knights Landing ####
# there seem to be quite a few missing effort calculations pre 2002. for the ones that are missing,
# we will apply the one date metholody
knights_hours_fished_pre2006_calculated <- filter(catch, site == "knights landing", year(date) < 2006, year(date) >= 2002) %>%
  distinct(date, site, subsite, stream) %>%
  rename(trap_stop_date = date) |> 
  mutate(start_datetime = ymd_hms(paste(trap_stop_date, "00:00:00"))) %>%
  mutate(hours_fished_methodology = "using only start time") %>% 
  hours_fished_one_date() |> 
  mutate(hours_fished = as.numeric(hours_fished)) |> 
  rename(hours_fished_calculated = hours_fished)
# prior to 2006 data stored in excel spreadsheets where hours fished was calculated
knights_hours_fished_pre2006 <- knights_effort_pre2006 %>% 
  select(date, hrs_fished, cone_id) %>% 
  rename(trap_stop_date = date,
         hours_fished = hrs_fished,
         subsite = cone_id) %>% 
  mutate(stream = "sacramento river",
         site = "knights landing") %>% 
  group_by(stream, site, trap_stop_date) %>% 
  summarize(hours_fished = sum(hours_fished, na.rm = T)) %>% 
  mutate(subsite = "knights landing",
         hours_fished_methodology = "calculated by knights landing") |> # catch data pre 2006 is aggregated across traps
  full_join(knights_hours_fished_pre2006_calculated |> 
              select(trap_stop_date, stream, site, subsite, hours_fished_calculated)) |> 
  mutate(hours_fished = ifelse(is.na(hours_fished), hours_fished_calculated, hours_fished),
         hours_fished_methodology = ifelse(is.na(hours_fished_methodology),"using only start time", hours_fished_methodology)) |> select(-hours_fished_calculated)

# there seem to be quite a few missing effort calculations pre 2002. for the ones that are missing,
# we will apply the one date metholody
knights_hours_fished_pre2002_calculated <- filter(catch, site == "knights landing", year(date) < 2002) %>%
  distinct(date, site, subsite, stream) %>%
  rename(trap_stop_date = date) |> 
  mutate(start_datetime = ymd_hms(paste(trap_stop_date, "00:00:00"))) %>%
  mutate(hours_fished_methodology = "using only start time") %>% 
  hours_fished_one_date() |> 
  mutate(hours_fished = as.numeric(hours_fished)) |> 
  rename(hours_fished_calculated = hours_fished)
# prior to 2002 data stored in database files where hours fished was calculated

knights_hours_fished_pre2002 <- knights_effort_pre2002 %>% 
  # if effort is.na() then assume 24 hours
  mutate(effort = ifelse(is.na(effort), 24, effort)) |> 
  select(date, trapid, effort) %>% 
  rename(trap_stop_date = date,
         hours_fished = effort,
         subsite = trapid) %>% 
  mutate(stream = "sacramento river",
         site = "knights landing") %>% 
  group_by(stream, site, subsite, trap_stop_date) %>% 
  summarize(hours_fished = sum(hours_fished, na.rm = T)) %>% 
  mutate(subsite = ifelse(is.na(subsite), "knights landing", as.character(subsite)), # there are a few subsites that are NA so fill in with Knights Landing
         hours_fished_methodology = "calculated by knights landing") |> 
  full_join(knights_hours_fished_pre2002_calculated |> 
              select(trap_stop_date, stream, site, subsite, hours_fished_calculated)) |> 
  mutate(hours_fished = ifelse(is.na(hours_fished), hours_fished_calculated, hours_fished),
         hours_fished_methodology = ifelse(is.na(hours_fished_methodology),"using only start time", hours_fished_methodology)) |> select(-hours_fished_calculated)

### Hours fished ####
# calculate hours fished only for period of times when trap was fishing otherwise there will be duplicates that don't make sense
knights_hours_fished <- trap %>% 
  filter(stream == "sacramento river", site == "knights landing", 
         fish_processed == "processed fish", include == T, visit_type != "start trapping"
         ) %>%
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time,
         stream, site, subsite, fish_processed, 
         include, visit_type) %>% 
  arrange(subsite, trap_stop_date, trap_stop_time) %>% 
  mutate(hours_fished_methodology = "using start time and stop time") %>% 
  hours_fished() %>% 
  mutate(hours_fished = as.numeric(hours_fished)) %>% 
  bind_rows(knights_hours_fished_pre2006,
            knights_hours_fished_pre2002)

knitr::kable(head(knights_hours_fished, 5))
```

### QC

```{r}
knights_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

knights_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

knights_hours_fished %>% filter(hours_fished > 50)
```

## tisdale

```{r, tisdale}
## Tisdale ####
# filter(trap, stream == "sacramento river", site == "tisdale", trap_start_date == "2019-07-24")

tisdale_hours_fished <- filter(trap, stream == "sacramento river", site == "tisdale", 
                               fish_processed %in% c("processed fish", "not recorded", "no fish caught"), include == T, 
                               visit_type != "start trapping") %>%
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, 
         trap_visit_time, trap_visit_date, subsite, site, stream) %>% 
  arrange(subsite, trap_stop_date, trap_stop_time) %>% 
  mutate(hours_fished_methodology = "using start time and stop time") %>% 
  hours_fished() 

knitr::kable(head(tisdale_hours_fished, 5))
```



### QC

```{r}
tisdale_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

tisdale_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# filter(tisdale_hours_fished, hours_fished > 1000) %>%  tally()
```


## mill creek

Mill Creek does not have data available (no time variable) to calculate hours
fished. Assume that effort is 24 hours for every day with data.

```{r, mill}
## Mill ####
mill_hours_fished <- filter(trap, stream == "mill creek") %>% 
  distinct(trap_stop_date, stream, site, subsite) %>% 
  mutate(hours_fished_methodology = "24 hour assumption") %>% 
  mutate(hours_fished = 24)


knitr::kable(head(mill_hours_fished, 5))
```

## yuba river

Hours fished are calculated by taking the difference between the date/time for 
each observation. 

```{r, yuba}
## Yuba ####
# filter(trap, stream == "yuba river", is.na(trap_stop_date)) %>% 
#   tally()
### Hours fished ####
yuba_hours_fished <- filter(trap, stream == "yuba river") %>%
  distinct(trap_stop_date, trap_stop_time, subsite, site, stream) %>%
  mutate(start_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time))) %>%
  mutate(hours_fished_methodology = "using only start time") %>% 
  hours_fished_one_date() %>% 
  # add hours across traps at a site
  # group_by(stream, site, date) %>% 
  # summarize(hours_fished = sum(hours_fished)) %>% 
  mutate(hours_fished = case_when(is.na(hours_fished) | hours_fished < 0 ~ 24,
                                  T ~ as.numeric(hours_fished)))

knitr::kable(head(yuba_hours_fished, 5))
```

### QC

```{r}
# there are some errors: hours > 5000
yuba_hours_fished %>% 
  ggplot(aes(x = end_datetime, y = hours_fished)) +
  geom_point()

yuba_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# filter(yuba_hours_fished, hours_fished < 0)
```

### Compare methodologies 

There are a few really large values when using the rev methodology, some comments in the trap data reflect that the counter may have been tampered with in some of these instances. To clean up the data on these really large days we change all instances where hours fished > 200 and assume a 24 hour fish. 

```{r}
#test out using trap rotations methodology 
yuba_rev_method <- filter(trap, stream == "yuba river") %>% 
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, rpms_start, rpms_end, sample_period_revolutions, stream, site, subsite, trap_visit_id) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  mutate(hours_fished_methodology = "using rpm and revolutions") %>% 
  revolution_calculated_hours_fished() %>% 
  mutate(hours_fished = case_when(hours_fished > 200 ~ 24,
                                  T ~ as.numeric(hours_fished)))

# hours fished appear to be more spread out using this revolution method - do not link well with trap start date/trap start time 
yuba_rev_method %>% 
  ggplot(aes(x = trap_stop_date, y = hours_fished)) +
  geom_point()

yuba_rev_method %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()
```

## redd bluff

```{r}
rbdd_hours_fished <- filter(trap, site == "red bluff diversion dam") %>%
  distinct(trap_start_date, trap_start_time, site, subsite, stream) %>%
  rename(trap_stop_date = trap_start_date,
         trap_stop_time = trap_start_time) |>
  mutate(start_datetime = ymd_hms(paste0(trap_stop_date, " ", trap_stop_time))) %>%
  mutate(hours_fished_methodology = "using only start time") %>% 
  hours_fished_one_date() |> 
  mutate(hours_fished = as.numeric(hours_fished),
         hours_fished = ifelse(hours_fished < 0, 24, hours_fished),
         hours_fished = ifelse(hours_fished > 200, 24, hours_fished)) 

```

### QC

```{r}
rbdd_hours_fished %>% 
  filter(hours_fished < 72 & hours_fished > 0) |> 
  ggplot(aes(x = end_datetime, y = hours_fished)) +
  geom_point()

rbdd_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

filter(rbdd_hours_fished, hours_fished > 80)
```


# Combine data

```{r}
# all wee need is stop_date, hours_fished, stream, site
# do not sum by subsites. this can be done in further analysis
battle_hours_fished %>% glimpse()
battle_final <- battle_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

butte_hours_fished %>%  glimpse()
butte_final <- butte_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

clear_hours_fished %>% glimpse()
clear_final <- clear_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

deer_hours_fished %>% glimpse()
deer_final <- deer_hours_fished 

feather_hours_fished %>% glimpse()
feather_final <- feather_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

knights_hours_fished %>% glimpse() 
knights_final <- knights_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time,  hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

tisdale_hours_fished %>% glimpse()
tisdale_final <- tisdale_hours_fished %>% 
   select(stream, site, subsite, trap_stop_date, trap_stop_time,  hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

mill_hours_fished %>% glimpse()
mill_final <- mill_hours_fished

yuba_hours_fished %>% glimpse()
yuba_final <- yuba_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

rbdd_hours_fished |> glimpse()
rbdd_final <- rbdd_hours_fished |> 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) |> 
  mutate(hours_fished = as.numeric(hours_fished))

hours_fished_combined <- bind_rows(battle_final,
                                   butte_final,
                                   clear_final,
                                   deer_final,
                                   feather_final,
                                   knights_final,
                                   tisdale_final,
                                   mill_final,
                                   yuba_final,
                                   rbdd_final) %>% 
  # if hours fished NA assumed 24 hours
  mutate(hours_fished = ifelse(is.na(hours_fished), 24, hours_fished)) %>% 
  rename(date = trap_stop_date) %>% 
  group_by(stream, site, subsite, date) %>% 
  summarize(hours_fished = sum(hours_fished)) %>% 
  glimpse()
```

## Summarized by week

```{r}
# filter(hours_fished_combined, is.na(trap_stop_date)) %>%  tally()
# filter(hours_fished_combined, is.na(hours_fished)) %>%  tally()
# filter(hours_fished_combined, trap_stop_date == "2008-03-24")
# filter(hours_fished_combined, hours_fished < 0)
#filter(butte_hours_fished, date == "2015-06-03")
#filter(clear_hours_fished, stop_date == "2021-05-10")
#filter(yuba_hours_fished, date == "2009-08-18")

hours_fished_weekly <- hours_fished_combined %>% 
  mutate(week = week(date),
         year = year(date)) %>% 
  group_by(stream, site, subsite, week, year) %>% 
  summarize(hours_fished = sum(hours_fished)) %>% glimpse()

hours_fished_combined %>% filter(site == "lower feather river", date == as_date("2022-02-22")) 
```

### QC

```{r}
hours_fished_weekly %>% 
  ggplot(aes(x = week, y = hours_fished, color = year)) +
  geom_point() + 
  facet_wrap(~stream)

hours_fished_weekly %>% 
  ggplot(aes(x = hours_fished, fill = stream)) +
  geom_histogram(alpha = .75, position = 'identity')
```
```{r, include = F}
check_dates <- function(dat) {
  dat |> 
    group_by(stream, site, subsite) |> 
    summarize(min_date = min(date),
              max_date = max(date))
}

hours_fished_combined |> 
  check_dates() |> 
  view()

```

# Save data

```{r, save_data}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(hours_fished_combined,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_rst_effort.csv",
           predefinedAcl = "bucketLevel")
```

