---
title: "Create standard format for RST catch data"
output: html_document
---

```{r, include = F}
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(knitr)
library(hms)
```

Data were checked and cleaned in scripts available [here](https://github.com/FlowWest/JPE-datasets/tree/main/scripts/rst)
Cleaned data were saved on the jpe-dev-bucket on google cloud with the exception
of Knights Landing data from CAMP which has not yet been saved to google cloud.

```{r, data_pull}
# Data pulled in from google cloud in rst_catch_standard_format
# load in data
# catch data
battle_rst <- read_csv("data/rst/battle_rst_catch.csv")
butte_rst <- read_csv("data/rst/butte_rst.csv") # contains trap data too
clear_rst <- read_csv("data/rst/clear_rst_catch.csv")
deer_rst <- read_csv("data/rst/deer_rst.csv") # contains trap data too
feather_rst <- read_csv("data/rst/feather_rst.csv")
# knights_rst <- read_csv("data/rst/knights_rst_clean.csv") # contains cpue info
tisdale_rst <- read_csv("data/rst/tisdale_rst.csv")
mill_rst <- read_csv("data/rst/mill_rst.csv") # contains trap data too
yuba_rst <- read_csv("data/rst/yuba_rst.csv")

# sampling effort and environmental
battle_environmental <- read_csv("data/rst/battle_rst_environmental.csv")
clear_environmental <- read_csv("data/rst/clear_rst_environmental.csv")
feather_effort <- read_csv("data/rst/feather_rst_effort.csv")
knights_effort <- read_csv("data/rst/knights_sampling_effort_clean.csv")

# Knights landing data pulled from camp
# TODO add to cloud (this dataset was shared later and has not yet been added to cloud)
knights_landing_rst <- read_rds("data/rst/knights_landing_raw_catch.rds")
```

# Data formatting {.tabset}

## Battle Creek

```{r, battle_clean}
# counter means number on cone revolution counter at sample_date, sample_time
battle_environmental %>% str
unique(battle_environmental$trap_sample_type)
unique(battle_environmental$fish_properly)
unique(battle_environmental$partial_sample)
unique(battle_environmental$debris_tubs)
unique(battle_environmental$gear_condition_code)

# check to see why some dates are NA. I think this is an error
# because these contain no data. Decided to remove. 
# filter(battle_trap_clean, is.na(start_date))

battle_trap_clean <- battle_environmental %>%
  select(-c(sample_id, baileys_eff, num_released, report_baileys_eff, sub_week, sample_weight)) %>%
  # select(trap_start_date, trap_start_time, sample_date, sample_time,
  #        counter, flow_start_meter, flow_end_meter, flow_set_time, velocity,
  #        turbidity, cone, trap_sample_type, thalweg, depth_adjust, avg_time_per_rev,
  #        fish_properly, partial_sample) %>%
  # sample_date and sample_time are described as the end of 24 hour sampling period
  # rename to match others
  rename(start_date = trap_start_date,
         start_time = trap_start_time,
         end_date = sample_date,
         end_time = sample_time,
         sample_period_revolutions = counter,
         comments = trap_comments,
         cone_setting = cone) %>%
  mutate(site = "Battle Creek",
         watershed = "Battle Creek") %>%
  filter(!is.na(start_date)) %>%
  data.frame() %>%
  distinct()
 
# check to make sure one row per day/time 
# battle_trap_clean %>%
#   group_by(start_date, start_time) %>%
#   tally() %>%
#   filter(n < 1)
```

### Variables removed

- sample_id: doesn't map to data outside of CAMP
- baileys_eff and report_baileys_eff are calculated values
- num_released included in mark and recapture data
- sub_week: is related to efficiency and included in efficiency table
- sample_weight: same as cone variable

## Butte Creek

```{r, butte_clean}
unique(butte_rst$trap_status)
unique(butte_rst$gear_id)
unique(butte_rst$station)
butte_rst %>%
  group_by(station, gear_id) %>%
  tally()

# there are multiple obs per day which i think is a data entry issue.
# group by date, station, time, gear_id, trap_status then average
butte_trap_clean <- butte_rst %>%
  select(-c(dead, count, fork_length, weight, mark_code, lifestage)) %>%
  # select(date, station, trap_status, time, gear_id, temperature, turbidity, velocity,
  #        trap_revolutions, rpms_start, rpms_end) %>%
  group_by(date, station, trap_status, time, gear_id) %>%
  summarize(temperature = mean(temperature, na.rm = T),
            turbidity = mean(turbidity, na.rm = T),
            velocity = mean(velocity, na.rm = T),
            trap_revolutions = mean(trap_revolutions, na.rm = T),
            rpms_start = mean(rpms_start, na.rm = T),
            rpms_end = mean(rpms_end, na.rm = T)) %>%
  rename(site = station,
         sample_period_revolutions = trap_revolutions) %>%
  mutate(watershed = "Butte Creek") %>%
  distinct()

# check to make sure one row per day/time 
# butte_trap_clean %>%
#   group_by(date, time, site, gear_id, trap_status) %>%
#   tally() %>%
#   filter(n > 1)

```

### Variables removed

All variables removed are included in the catch table.

## Clear Creek

```{r, clear_clean}
unique(clear_environmental$gear_condition_code)
unique(clear_environmental$station_code)

# there are NA start_dates which don't have any data. need to remove.

clear_trap_clean <- clear_environmental %>%
  select(-c(sample_id, num_released, baileys_eff, sub_week)) %>%
  rename(site = station_code,
         start_date = trap_start_date,
         start_time = trap_start_time,
         end_date = sample_date,
         end_time = sample_time,
         sample_period_revolutions = counter,
         comments = trap_comments,
         cone_setting = cone) %>%
  mutate(watershed = "Clear Creek") %>%
  filter(!is.na(start_date)) %>%
  distinct()

# check to make sure one row per day/time 
# clear_trap_clean %>%
#   group_by(start_date, start_time, site) %>%
#   tally() %>%
#   filter(n > 1)
```

### Variables removed

- sample_id: doesn't map to data outside of CAMP
- baileys_eff is a calculated value
- num_released included in mark and recapture data
- sub_week: is related to efficiency and included in efficiency table

## Deer Creek

```{r, deer_clean}
unique(deer_rst$location)
unique(deer_rst$trap_condition_code)

deer_trap_clean <- deer_rst %>%
  select(-c(count, fork_length, weight)) %>%
  # select(date, location, flow, time_for_10_revolutions, trap_condition_code, turbidity,
  #        water_temperature_celsius) %>%
  group_by(date, location, trap_condition_code) %>%
  summarize(flow = mean(flow, na.rm = T),
            time_for_10_revolutions = mean(time_for_10_revolutions, na.rm = T),
            turbidity = mean(turbidity, na.rm = T),
            water_temperature_celsius = mean(water_temperature_celsius, na.rm =T)) %>%
  rename(site = location,
         temperature = water_temperature_celsius,
         gear_condition_code = trap_condition_code) %>%
  mutate(watershed = "Deer Creek") %>%
  distinct()


# check to make sure one row per day/time 
# deer_trap_clean %>%
#   group_by(date,site, trap_condition_code) %>%
#   tally() %>%
#   filter(n > 1)
# 
# filter(deer_trap_clean, date == "2001-11-10")
```

### Variables removed

All variables removed are included in the catch table.

## Feather River

```{r, feather_clean}
unique(feather_effort$trap_functioning)
unique(feather_effort$sub_site_name)
unique(feather_effort$fish_processed)
# date, site_name, visit_time, visit_type, trap_functioning, water_temp_c,
# turbidity_ntu, latitude, longitude
feather_trap_clean <- feather_effort %>%
  select(-c(river_location, river_mile, latitude, longitude)) %>%
  # select(date, site_name, sub_site_name, visit_time, visit_type, trap_functioning,
  #        water_temp_c, turbidity_ntu, latitude, longitude) %>%
  rename(site = site_name,
         subsite = sub_site_name,
         time = visit_time,
         temperature = water_temp_c,
         turbidity = turbidity_ntu,
         trap_status = visit_type,
         gear_condition_code = trap_functioning) %>%
  mutate(time = as_hms(time),
         watershed = "Feather River") %>%
  distinct()

# check to make sure one row per day/time 
# feather_trap_clean %>%
#   group_by(date,time, site, subsite) %>%
#   tally() %>%
#   filter(n > 1)
```

### Variables removed

All variables removed are included in the location table.

## Mill Creek

```{r, mill_clean}
unique(mill_rst$location)
unique(mill_rst$trap_condition_code)

mill_trap_clean <- mill_rst %>%
  select(-c(count, fork_length, weight, location)) %>%
  # select(date, location, flow, time_for_10_revolutions, trap_condition_code,
  #        water_temperature, turbidity) %>%
  group_by(date, trap_condition_code) %>%
  summarize(flow = mean(flow, na.rm = T),
            time_for_10_revolutions = mean(time_for_10_revolutions, na.rm = T),
            turbidity = mean(turbidity, na.rm = T),
            water_temperature = mean(water_temperature, na.rm =T)) %>%
  rename(temperature = water_temperature,
         gear_condition_code = trap_condition_code) %>%
  mutate(watershed = "Mill Creek",
         site = "Mill Creek") %>%
  distinct()

# check to make sure one row per day/time 
# mill_trap_clean %>%
#   group_by(date,site, trap_condition_code) %>%
#   tally() %>%
#   filter(n > 1)
# filter(mill_trap_clean, date == "2002-02-07")
```

### Variables removed

- location: all are Mill Creek RSTR

All other removed variables included in the catch table.


## Yuba River

```{r, yuba_clean}
unique(yuba_rst$method)
unique(yuba_rst$trap_status)
unique(yuba_rst$location)

yuba_trap_clean <- yuba_rst %>%
  select(-c(organism_code, fork_length, weight, lifestage, count, run)) %>%
  # select(date, time, method, temperature, turbidity, velocity, trap_status,
  #        rpms_before, rpms_after, location) %>%
  rename(site = location,
         sample_period_revolutions = trap_revolutions,
         start_counter = trap_revolutions2,
         rpms_start = rpms_before,
         rpms_end = rpms_after) %>%
  mutate(watershed = "Yuba River") %>%
  distinct()

# check to make sure one row per day/time 
# yuba_trap_clean %>%
#   group_by(date,time, site) %>%
#   tally() %>%
#   filter(n > 1)
```

### Variables removed

All variables removed are included in the catch table.

## Lower Sacramento - Knights Landing

```{r, knights_clean}
unique(knights_effort$gear)
unique(knights_effort$cone_sampling_effort)
# date, start_date, stop_date, number_traps, hrs_fished, flow_cfs, turbidity,
# water_t_f, cone_sampling_effort, cone_id, cone_rpm, total_cone_rev
knights_trap_clean <- knights_effort %>%
  select(-c(date, location, gear, sampling_period_hrs, turbidity_units, cone_sampling_effort)) %>%
  # convert temp from F to C
  mutate(temperature = (water_t_f - 32) * (5/9),
         watershed = "Lower Sacramento - Knights Landing",
         site = "Lower Sacramento - Knights Landing",
         cone_id = as.character(cone_id)) %>%
  select(-water_t_f) %>%
  rename(end_date = stop_date,
         end_time = stop_time,
         flow = flow_cfs,
         sample_period_revolutions = total_cone_rev,
         secchi = secchi_ft,
         gear_id = cone_id) %>%
  distinct()
# 
# # check to make sure one row per day/time 
# knights_trap_clean %>%
#   group_by(start_date,start_time, cone_id, date) %>%
#   tally() %>%
#   filter(n > 1)
```

### Variables removed

- date: use start_date and end_date
- location: all are KL for Knights Landing
- gear: all are two 8in cones
- sampling_period_hrs: unclear how this variable is different than hrs_fished
- turbidity_units: all are NTU except in 2014 when they are FTU
- cone_sampling_effort: Variable is unclear and is redundant of number of traps and hrs fished

## Lower Sacramento - Tisdale

No effort data is currently available for Tisdale.

# Standard format {.tabset}

```{r, combined}
# TODO fill in run method
# TODO decide if we need the "random" variable. It only applies to Tisdale
combined_trap_raw <- bind_rows(battle_trap_clean,
                          butte_trap_clean,
                          clear_trap_clean,
                          deer_trap_clean,
                          feather_trap_clean,
                          mill_trap_clean,
                          yuba_trap_clean,
                          knights_trap_clean)
colnames(combined_trap_raw)
combined_rst <- combined_rst_raw %>%
  run_encoding() %>%
  lifestage_encoding() %>%
  # all should be chinook salmon
  mutate(species = ifelse(is.na(species), "chinook salmon", species),
         run_method = case_when(is.na(run_method) ~ "not recorded",
                                run == "not recorded" | count == 0 ~ NA_character_,
                                T ~ tolower(run_method)))

unique(combined_trap_raw$fish_processed)
unique(combined_trap_raw$trap_status)
unique(combined_trap_raw$gear_condition_code)

gear_condition_encoding <- function(dat) {
  dat %>%
    mutate(gear_condition_code = case_when(gear_condition_code %in% c("not rotating", "total block", 
                                                                      "total blockage", "cone stopped", 
                                                                      "Trap not in service", "Trap stopped functioning") ~ "out of service",
                                           gear_condition_code %in% c("partial block", "partial blockage", "Trap functioning, but not normally") ~ "abnormal",
                                           gear_condition_code %in% c("normal", "Trap functioning normally") ~ "normal"))
}

trap_visit_type_encoding <- function(dat) {
  dat %>%
    mutate(trap_visit_type = case_when(trap_status %in% ))
}
```
