---
title: "Standard release fish"
author: "Ashley Vizek"
date: "10/6/2022"
output: 
  html_document:
    theme: flatly
---
This script creates a dataset with individual fork length measurements for
all fish released in efficiency trial.

Knights Landing is the only location where data is available.

```{r, include = F}
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(hms)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

```{r, include = F}
gcs_get_object(object_name = "rst/feather-river/data-raw/feather_release_fish_camp.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/rst/feather_release_fish.csv",
               overwrite = TRUE)
feather_release_fish <- read_csv("data/rst/feather_release_fish.csv")

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/knights_release_fish_camp.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/rst/knights_release_fish.csv",
               overwrite = TRUE)
knights_release_fish <- read_csv("data/rst/knights_release_fish.csv")

gcs_get_object(object_name = "rst/lower-sac-river/data-raw/tisdale/tisdale_release_fish_camp.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/rst/tisdale_release_fish.csv",
               overwrite = TRUE)
tisdale_release_fish <- read_csv("data/rst/tisdale_release_fish.csv")
```

# Tisdale

Tisdale and Lower Feather do not have fork lengths for individual fish released

```{r}
filter(tisdale_release_fish, !is.na(forkLength))
```

# Feather

Feather does not have fork lengths for individual fish released

```{r}
filter(feather_release_fish, !is.na(forkLength))
```

# Knights

Knights Landing does have fork lengths for individual fish released

```{r}
filter(knights_release_fish, !is.na(forkLength))

knights_release_fish_clean <- knights_release_fish %>% 
  select(-projectDescriptionID) %>% 
  mutate(stream = "sacramento river",
         site = "knights landing",
         subsite = "knights landing") %>% 
  rename(release_id = releaseID,
         count = nMarked,
         fork_length = forkLength)
```

# Combined data

```{r}
combined_release_fish <- knights_release_fish_clean
```

# QC

## fork_length

Knights Landing only measures 50-100 released fish although they may release more

```{r}
combined_release_fish %>% 
  ggplot(aes(fork_length)) +
  geom_histogram()

ggplot(data = combined_release_fish, aes(
  x = as.factor(release_id),
  y = fork_length,
  color = stream
)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 8, angle = 45))
```


## weight

Weight is measured in one release trial

```{r}
filter(combined_release_fish, !is.na(weight))

ggplot(data = combined_release_fish, aes(
  x = as.factor(release_id),
  y = weight,
  color = stream
)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(size = 8, angle = 45))
```


# Save data

```{r}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(combined_release_fish,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_rst_release_fish.csv",
           predefinedAcl = "bucketLevel")
```
