---
title: "Create standard format for RST hours fished"
output: html_document
---

Hours fished is not included as a variable in the data and needs to be calculated.
This file calculates hours fished by date, stream and site.

```{r, include = F}
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(hms)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

Data were checked and cleaned in scripts available [here](https://github.com/FlowWest/JPE-datasets/tree/main/scripts/rst).
Cleaned data were saved on the jpe-dev-bucket on google cloud.

```{r, include = F}
gcs_get_object(object_name = "standard-format-data/standard_rst_trap.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/standard_rst_trap.csv",
               overwrite = TRUE)
trap <- read_csv("data/standard_rst_trap.csv")
```

# Method

If the start and stop date/time of the sampling period are available, simply subtract the difference in time.

```{r, helpers}
# calculating hours fished when have start and stop datetime
hours_fished <- function(dat){
  dat %>%
    filter(!is.na(trap_stop_time), !is.na(trap_start_time)) %>%
    mutate(start_datetime = ymd_hms(paste(trap_start_date, trap_start_time)),
         stop_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time)),
         hours_fished = round(difftime(stop_datetime, start_datetime, units = "hours"), 2))
}
```

If the start and stop date/time of the sampling period are available, simply subtract the difference in time.

```{r, helpers}
# calculating hours fished when have start and stop datetime
revolution_calculated_hours_fished <- function(dat){
  dat %>%
    filter(!is.na(rpms_start), !is.na(rpms_end), !is.na(sample_period_revolutions)) %>% 
    mutate(prior_day_rpm = lag(rpms_end),
           sum_rpms = sum(prior_day_rpm, rpms_start, na.rm = T), 
           cone_rpms = prior_day_rpm + rpms_start / sum_rpms,
           hours_fished = round(sample_period_revolutions/cone_rpms/60, 2))
}
```

If only one date is available, assume that the end date is the date sampled following day.

```{r, helper2}
# calculating hours fished when have only date and time
hours_fished_one_date <- function(dat) {
  dat %>%
  arrange(site, subsite, start_datetime) %>%
  mutate(end_datetime = lead(start_datetime),
         end_datetime = case_when(difftime(end_datetime, start_datetime, units = "hours") > 120 ~ start_datetime + 1, 
                                  T ~ end_datetime),
         hours_fished = round(difftime(end_datetime, start_datetime, units = "hours"), 2))
}
```

## Other Assumptions

Based on protocols and information from monitoring programs, sampling periods are typically 24 hours,
unless flows are very high in which traps are checked more frequently. Calculating hours fished based
on time and date provided in data provides more specificity, however, there are typos and errors resulting
in nonsensical hours fished. In these cases, we assume 24 hour sampling periods. 

- If only a date is available and no time, assume 24 hour sampling periods for each date.
- If information is missing or there are typos, assume 24 hour sampling periods.

# Hours fished {.tabset}

## battle creek

Battle Creek has data available to calculate the hours fished. There are a few
cases in 2021 when start date is NA. For these we fill in the start date assuming
it is the sample date of the previous record.

```{r, battle}
## Battle ####
### Hours fished ####
# filter(trap, stream == "battle creek", is.na(trap_start_time))
# # these are typos that need to be fixed
# filter(trap, stream == "battle creek", trap_stop_date == "2004-02-13")
# filter(trap, stream == "battle creek", trap_stop_date == "2005-01-06")

battle_hours_fished <- filter(trap, stream == "battle creek") %>%
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, stream, site, subsite, trap_visit_id) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  hours_fished()
```

### QC

```{r}
battle_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

battle_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# filter(battle_hours_fished, hours_fished > 5000) %>%  tally()
# filter(trap, stream == "battle creek", trap_stop_date == "2004-02-13")
# filter(trap, stream == "battle creek", trap_stop_date == "2005-01-06")
# filter(trap, stream == "battle creek", trap_stop_date == "2020-07-25")
# 
# # these seem large. it is possible so i will leave as is.
# filter(battle_hours_fished, hours_fished > 50)
# filter(battle_hours_fished, hours_fished < 0)
# ck <- filter(trap, stream == "battle creek")
```

## butte creek

Hours fished are calculated by taking the difference between the date/time at each 
observation. 

```{r, butte}
## Butte ####
### Hours fished ####
# There are no missing dates but we only have one data and time, no start/stop
# filter(trap, stream == "butte creek", is.na(trap_stop_date)) %>% 
#   tally()
# 
# filter(trap, stream == "butte creek", trap_stop_date == "2008-03-24")

butte_hours_fished <- filter(trap, stream == "butte creek") %>%
  distinct(trap_stop_date, trap_stop_time, site, subsite, stream) %>%
  mutate(start_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time))) %>%
  hours_fished_one_date() %>% 
  mutate(end_datetime = case_when(trap_stop_date == "2007-05-03" ~ ymd_hms("2007-05-04 08:00:00"),
                                  trap_stop_date == "2008-04-23" ~ ymd_hms("2008-04-24 07:15:00"),
                                  trap_stop_date == "2015-06-03" ~ ymd_hms("2015-06-04 08:00:00"),
                                  T ~ end_datetime),
         hours_fished = round(difftime(end_datetime, start_datetime, units = "hours"), 2),
         hours_fished = ifelse(hours_fished < 0, 24, hours_fished))
```

### QC

```{r}
butte_hours_fished %>% 
  ggplot(aes(x = end_datetime, y = hours_fished)) +
  geom_point()

# trying to figure out where to set cut of for effort
# decided to set as 5 days - 120 hours
butte_hours_fished %>% 
  filter(hours_fished < 200) %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# butte_hours_fished %>% 
#   filter(hours_fished > 5000) %>% tally()
# butte_hours_fished %>% 
#   filter(hours_fished < 0)
# ck <- filter(trap, stream == "butte creek")
```

## clear creek

Clear Creek has data available to calculate the hours fished. There are a few
cases when start date is NA. For these we fill in the start date assuming
it is the sample date of the previous record.

```{r, clear}
## Clear ####
### Hours fished ####

# missing some start dates
# filter(trap, stream == "clear creek", is.na(trap_start_date)) %>% 
#   tally()
# 
# filter(trap, stream == "clear creek", is.na(trap_stop_date)) %>% 
#   tally()
# 
# # swapped start and stop time for both the UCC and LCC
# filter(trap, stream == "clear creek", trap_stop_date == "2020-02-24")
# # duplicates for UCC, remove one with missing trap_start_time and missing counter
# filter(trap, stream == "clear creek", trap_stop_date == "2005-01-16")
# # typo
# filter(trap, stream == "clear creek", trap_stop_date == "2006-01-07")

clear_hours_fished <- filter(trap, stream == "clear creek") %>%
  select(stream, site, subsite, trap_start_date, trap_start_time, trap_stop_date, trap_stop_time) %>% 
  arrange(site, subsite, trap_stop_date, trap_stop_time) %>% 
  # fill in missing start dates with the stop date from previous day
  mutate(trap_start_date = case_when(is.na(trap_start_date) ~ lag(trap_stop_date),
                                T ~ trap_start_date),
         trap_start_time = case_when(is.na(trap_start_time) ~ lag(trap_stop_time), 
                                T ~ trap_start_time)) %>% 
  hours_fished()
```

### QC

```{r}
clear_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

clear_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# filter(clear_hours_fished, hours_fished > 5000) %>% 
#   tally()
# filter(clear_hours_fished, hours_fished < 0) %>% 
#   tally()
# 
# filter(trap, stream == "clear creek", trap_stop_date == "2006-01-07")
```

## deer creek

Deer Creek does not have data available (no time variable) to calculate hours
fished so assume 24 hours for every day with data. 

```{r, deer}
## Deer #####
deer_hours_fished <- filter(trap, stream == "deer creek") %>% 
  distinct(trap_stop_date, stream, site, subsite) %>% 
  mutate(hours_fished = 24)
```

## feather river

Feather River has data available to calculate the hours fished. We only calculate
the hours fished when fish are processed. 

```{r, feather}
## Feather ####

# # it looks like all cases where start date is missing and fish were processed are typos or issues with data
# filter(trap, stream == "feather river", is.na(trap_start_date) & fish_processed == "processed fish")
# # are there multiple subsites per site?
# filter(trap, stream == "feather river") %>% 
#   group_by(site) %>% 
#   distinct(subsite)
# # multiple subsites fish on same day? yes
# filter(trap, stream == "feather river") %>% 
#   group_by(trap_stop_date, site) %>% 
#   tally() %>% 
#   filter(n > 1)
# filter(trap, stream == "feather river") %>% 
# filter(trap_stop_date == "2019-01-02")
# filter(trap, stream == "feather river") %>% 
# filter(trap_stop_date == "2000-01-20")
# filter(trap, stream == "feather river") %>% 
# filter(trap_start_date == "2015-02-14")
# calculate hours fished only for period of times when trap was fishing otherwise there will be duplicates that don't make sense
# typically a trap is not processed fish when service but this happens 2x. stop date/time filled in with the visit date
feather_hours_fished <- filter(trap, stream == "feather river", fish_processed == "processed fish", include == T, visit_type != "start trapping") %>%
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, trap_visit_time, trap_visit_date, stream, site, subsite) %>% 
  arrange(stream, site, subsite, trap_stop_date, trap_stop_time) %>% 
  # fix typos
  mutate(
  # fill in missing start dates with the stop date from previous day
         trap_start_date = case_when(is.na(trap_start_date) ~ lag(trap_stop_date),
                                T ~ trap_start_date),
         trap_start_time = case_when(is.na(trap_start_time) ~ lag(trap_stop_time),
                                T ~ trap_start_time)) %>%
  hours_fished() %>% 
  select(-trap_visit_date, -trap_visit_time)

```

### QC

```{r}
feather_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

feather_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# these are typos and will be fixed by assuming 24 hours fished
# filter(feather_hours_fished, hours_fished > 1000)%>% 
#   tally()
# filter(feather_hours_fished, hours_fished < 0) %>% 
#   tally()
# filter(feather_hours_fished, is.na(hours_fished))%>% 
#   tally()

```

## knights landing

Knights Landing has data available to calculate the hours fished. 

```{r, knights}
## Knights Landing ####

### Hours fished ####
# calculate hours fished only for period of times when trap was fishing otherwise there will be duplicates that don't make sense
knights_hours_fished <- filter(trap, stream == "sacramento river", site == "knights landing", fish_processed == "processed fish", include == T, visit_type != "start trapping") %>%
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, stream, site, subsite) %>% 
  arrange(subsite, trap_stop_date, trap_stop_time) %>% 
  hours_fished() 
```

### QC

```{r}
knights_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

knights_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()
```

## tisdale & lower feather

```{r, tisdale}
## Tisdale ####
# filter(trap, stream == "sacramento river", site == "tisdale", trap_start_date == "2019-07-24")

tisdale_hours_fished <- filter(trap, stream == "sacramento river", site == "tisdale", fish_processed == "processed fish", include == T, visit_type != "start trapping") %>%
  select(trap_start_date, trap_start_time, trap_stop_date, trap_stop_time, trap_visit_time, trap_visit_date, subsite, site, stream) %>% 
  arrange(subsite, trap_stop_date, trap_stop_time) %>% 
  hours_fished() 

```

### QC

```{r}
tisdale_hours_fished %>% 
  ggplot(aes(x = stop_datetime, y = hours_fished)) +
  geom_point()

tisdale_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# filter(tisdale_hours_fished, hours_fished > 1000) %>%  tally()
```

## mill creek

Mill Creek does not have data available (no time variable) to calculate hours
fished. Assume that effort is 24 hours for every day with data.

```{r, mill}
## Mill ####
mill_hours_fished <- filter(trap, stream == "mill creek") %>% 
  distinct(trap_stop_date, stream, site, subsite) %>% 
  mutate(hours_fished = 24)
```

## yuba river

Hours fished are calculated by taking the difference between the date/time for 
each observation. 

```{r, yuba}
## Yuba ####
# filter(trap, stream == "yuba river", is.na(trap_stop_date)) %>% 
#   tally()
### Hours fished ####
yuba_hours_fished <- filter(trap, stream == "yuba river") %>%
  distinct(trap_stop_date, trap_stop_time, subsite, site, stream) %>%
  mutate(start_datetime = ymd_hms(paste(trap_stop_date, trap_stop_time))) %>%
  hours_fished_one_date() %>% 
  # add hours across traps at a site
  # group_by(stream, site, date) %>% 
  # summarize(hours_fished = sum(hours_fished)) %>% 
  mutate(hours_fished = case_when(is.na(hours_fished) | hours_fished < 0 ~ 24,
                                  T ~ as.numeric(hours_fished)))
yuba_hours_fished |> glimpse()
yuba_2 <- dat |> revolution_calculated_hours_fished() |> glimpse()

```

### QC

```{r}
# there are some errors: hours > 5000
yuba_hours_fished %>% 
  ggplot(aes(x = end_datetime, y = hours_fished)) +
  geom_point()

yuba_hours_fished %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()

# filter(yuba_hours_fished, hours_fished < 0)
```

# Combine data

```{r}
# all wee need is stop_date, hours_fished, stream, site
# do not sum by subsites. this can be done in further analysis
battle_hours_fished %>% glimpse()
battle_final <- battle_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

butte_hours_fished %>%  glimpse()
butte_final <- butte_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

clear_hours_fished %>% glimpse()
clear_final <- clear_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

deer_hours_fished %>% glimpse()
deer_final <- deer_hours_fished 

feather_hours_fished %>% glimpse()
feather_final <- feather_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

knights_hours_fished %>% glimpse() 
knights_final <- knights_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time,  hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

tisdale_hours_fished %>% glimpse()
tisdale_final <- tisdale_hours_fished %>% 
   select(stream, site, subsite, trap_stop_date, trap_stop_time,  hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

mill_hours_fished %>% glimpse()
mill_final <- mill_hours_fished

yuba_hours_fished %>% glimpse()
yuba_final <- yuba_hours_fished %>% 
  select(stream, site, subsite, trap_stop_date, trap_stop_time, hours_fished) %>% 
  mutate(hours_fished = as.numeric(hours_fished))

hours_fished_combined <- bind_rows(battle_final,
                                   butte_final,
                                   clear_final,
                                   deer_final,
                                   feather_final,
                                   knights_final,
                                   tisdale_final,
                                   mill_final,
                                   yuba_final) %>% 
  # if hours fished NA assumed 24 hours
  mutate(hours_fished = ifelse(is.na(hours_fished), 24, hours_fished)) %>% glimpse()
```

## Summarized by week

```{r}
# filter(hours_fished_combined, is.na(trap_stop_date)) %>%  tally()
# filter(hours_fished_combined, is.na(hours_fished)) %>%  tally()
# filter(hours_fished_combined, trap_stop_date == "2008-03-24")
# filter(hours_fished_combined, hours_fished < 0)
#filter(butte_hours_fished, date == "2015-06-03")
#filter(clear_hours_fished, stop_date == "2021-05-10")
#filter(yuba_hours_fished, date == "2009-08-18")

hours_fished_weekly <- hours_fished_combined %>% 
  mutate(week = week(trap_stop_date),
         year = year(trap_stop_date)) %>% 
  group_by(stream, site, subsite, week, year) %>% 
  summarize(hours_fished = sum(hours_fished)) %>% glimpse()
```

### QC

```{r}
hours_fished_weekly %>% 
  ggplot(aes(x = week, y = hours_fished, color = stream)) +
  geom_point()

hours_fished_weekly %>% 
  ggplot(aes(x = hours_fished)) +
  geom_histogram()
```

# Save data

```{r, save_data}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(hours_fished_combined,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_rst_effort.csv",
           predefinedAcl = "bucketLevel")
```

