---
title: "Standardize RST Environmental Data"
author: "Erin Cain"
date: "4/20/2022"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)

color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")
```

```{r, include = FALSE}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

## RST Environmental Data Standardization

FlowWest received environmental data for rst from 9 monitoring programs:

-   Battle Creek
-   Butte Creek
-   Clear Creek
-   Deer Creek
-   Feather River
-   Mill Creek
-   Knights Landing
-   Yuba River
-   Tisdale

## Standard format for RST Environmental Data

Data dictionary for standard format:

(B - Battle Creek, Bu - Butte Creek, C - Clear Creek, F - Feather River,
D - Deer Creek, M - Mill Creek, KL - Knights Landing,
T - Tisdale, Y - Yuba River)

| column name   | stream collects         | definition                                               |
|:-----------------|:-------------------|:---------------------------------|
| stream      | B, Bu, C, D, F, M, KL, T, Y | Which stream the RST is located on                       |
| site          | B, Bu, C, D, F, M, KL, T, Y | site name                                                |
| subsite       | B, Bu, C, F, Y, KL, T          | sub site information name                                |
| date          | B, Bu, C, D, F, M, KL, T, Y | date that samples are taken                              |
| parameter     | B, Bu, C, D, F, M, KL, T, Y           | type of parameter being measured: velocity (ft/s), turbidity (NTU), weather (text), habitat (text), river depth left (m), river depth center (m), river depth right (m), temperature (degrees C), light penetration (secchi disk), discharge (cfs), water depth (m; measured at trap location)                     |
| text          | B, Bu, C, D, F, M, KL, T, Y               | text value for categorical parameters                                             |
| value  | B, Bu, C, D, F, M, KL, T, Y      | numeric value for numeric parameters                   |

## Read in data {.tabset}

### Battle Creek

#### Columns Removed

The dataset included operations and environmental data. Below only
columns for environmental data are selected.

- velocity
- turbidity
- weather
- habitat

We did not include lunar phase since that stream can be found elsewhere. 


```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/battle-creek/data/battle_rst_environmental.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "battle_rst_environmental.csv"),
               overwrite = TRUE)

battle_env <- read_csv(here::here("data", "rst", "battle_rst_environmental.csv")) %>% 
  select(trap_visit_id = sample_id, date = sample_date, 
         velocity, turbidity, weather = weather_code, habitat, `river depth left` = river_left_depth, `river depth center` = river_center_depth, `river depth right` = river_right_depth) %>%
  mutate(stream = "battle creek",
         site = "ubc",
         subsite = "ubc") %>%
  distinct() %>%
  mutate_if(is.numeric, as.character) %>% 
  pivot_longer(cols = c("velocity", "turbidity", "weather", "habitat", "river depth left", "river depth center", "river depth right"), values_to = "text", names_to = "parameter") %>% 
  mutate(value = as.numeric(text),
         text = ifelse(!is.na(value), NA_character_, text)) %>% 
  glimpse
```

### Butte Creek

#### Columns Removed

The dataset included catch, operations and environmental data. Below
only columns for environmental data are selected.

- weather
- temperature
- turbidity
- secchi
- velocity

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/butte-creek/data/butte_rst.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "butte_rst_environmental.csv"),
               overwrite = TRUE)

butte_env <- read_csv(here::here("data", "rst", "butte_rst_environmental.csv"), 
                      col_types = list(
                      "D", "c", "c", "l", "n", "n", "n", "n", "c", "t", 
                      "c", "c", "n", "n", "n", "n", "l", "l", "n", "c", 
                      "n", "n", "c"
                      )) %>%
  select(date, site = station, 
         weather, temperature, turbidity, `light penetration` = secchi, velocity) %>%
  mutate(stream = "butte creek",
         subsite = tolower(site), 
         site = ifelse(site == "Adams Dam", "adams dam", "okie dam")) %>%
  distinct() %>%
  mutate_if(is.numeric, as.character) %>% 
  pivot_longer(cols = c("velocity", "turbidity", "weather", "temperature", "light penetration"), values_to = "text", names_to = "parameter") %>% 
  mutate(value = as.numeric(text),
         text = ifelse(!is.na(value), NA_character_, text)) %>% 
  glimpse
```

### Clear Creek

#### Columns Removed

The dataset included operations and environmental data. Below only
columns for environmental data are selected.

- velocity
- turbidity
- weather
- habitat

We did not include lunar phase since that stream can be found elsewhere. 

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/clear-creek/data/clear_rst_environmental.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "clear_rst_environmental.csv"),
               overwrite = TRUE)

clear_env <- read_csv(here::here("data", "rst", "clear_rst_environmental.csv")) %>% 
  select(trap_visit_id = sample_id, date = sample_date, 
         velocity, turbidity, weather = weather_code, habitat, station_code, `river depth left` = river_left_depth, `river depth center` = river_center_depth, `river depth right` = river_right_depth) %>%
  mutate(stream = "clear creek", 
         site = ifelse(station_code == "ucc", "ucc", "lcc"),
         subsite = site) %>%
  select(-station_code) %>%
  distinct() %>%
  mutate_if(is.numeric, as.character) %>% 
  pivot_longer(cols = c("velocity", "turbidity", "weather", "habitat", "river depth left", "river depth center", "river depth right"), values_to = "text", names_to = "parameter") %>% 
  mutate(value = as.numeric(text),
         text = ifelse(!is.na(value), NA_character_, text)) %>% 
  glimpse
```

### Deer Creek

#### Columns Removed

The dataset included operations and environmental data. Below only
columns for environmental data are selected.

- turbidity
- weather
- temperature
- flow 

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/deer-creek/data/deer_rst.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "deer_rst_environmental.csv"),
               overwrite = TRUE)

deer_env <- read_csv(here::here("data", "rst", "deer_rst_environmental.csv"), 
                     col_types = list("D", "c", "n", "n", "n", "n", "n", "n", "c", 
                                      "n", "c", "c", "n")) %>% 
  select(date, site = location, discharge = flow, turbidity, weather, temperature = water_temperature_celsius) %>%
  mutate(stream = "deer creek",
         site = "deer creek", # All the streams are really the same site
         subsite = "deer creek") %>%
  distinct() %>%
  mutate_if(is.numeric, as.character) %>% 
  pivot_longer(cols = c("discharge", "turbidity", "weather", "temperature"), values_to = "text", names_to = "parameter") %>% 
  mutate(value = as.numeric(text),
         text = ifelse(!is.na(value), NA_character_, text)) %>% 
  glimpse

```

### Feather River

#### Columns Removed

The dataset included operations and environmental data. Below only
columns for environmental data are selected.

- discharge
- velocity
- temperature
- turbidity

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/feather-river/data-raw/feather_environmental_camp.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "feather_environmental_camp.csv"),
               overwrite = TRUE)

feather_env_raw <- read_csv(here::here("data", "rst", "feather_environmental_camp.csv")) 

filter(feather_env_raw, parameter == "velocity") %>%  ggplot(aes(value)) + geom_histogram()
# check those with different units
# looks like for those with different units is typo
# assume all flow in cfs
feather_env <- feather_env_raw %>% 
  # convert meters to feet and get rid of nonsensical values which are errors
  mutate(value = case_when(unit == "meters/second (m/sec)" & parameter == "velocity" ~ value*3.28,
                           # convert F to C
                           unit == "°F" & parameter == "temperature" ~ (value-32)*(5/9),
                           T ~ value)) %>% 
  group_by(date, site, subsite, parameter) %>% 
  summarise(value = mean(value, na.rm = T),
            value = ifelse(is.nan(value), NA_real_, value)) %>% 
  mutate(stream = "feather river")
```

### Lower Sac - Knights Landing

#### Columns Removed

The dataset included operations and environmental data. Below only
columns for environmental data are selected.

*knights landing has 2 subsites 8.3 and 8.4. The trap environmental data that we received did not include subsite. We listed subsite as NA for these data. 

- discharge
- light penetration
- temperature
- turbidity


```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/lower-sac-river/data/knights-landing/knl_combine_sampling_effort_clean.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "knights_rst_environmental_pre2006.csv"),
               overwrite = TRUE)

knights_landing_env_pre2006 <- read_csv(here::here("data", "rst", "knights_rst_environmental_pre2006.csv")) %>% 
  filter(date < as.Date("2006-10-02")) %>% 
  select(date, location, discharge = flow_cfs, `light penetration` = secchi_ft, temperature = water_t_f, turbidity, cone_id) %>%
  mutate(temperature = round((temperature - 32) * (5/9), 1), 
         site = "knights landing", 
         subsite = as.character(cone_id), 
         stream = "sacramento river") %>%
  select(-location, -cone_id) %>%
  distinct() %>%
  pivot_longer(cols = c("discharge", "turbidity", "light penetration", "temperature"), values_to = "value", names_to = "parameter") %>% 
  glimpse()
```

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/knights-landing/knights_environmental_camp.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst",  "knights_environmental_camp.csv"),
               overwrite = TRUE)

knights_env_raw <- read_csv(here::here("data", "rst", "knights_environmental_camp.csv")) 

knights_env <- knights_env_raw %>% 
  # convert meters to feet and get rid of nonsensical values which are errors
  mutate(value = case_when(unit == "meters/second (m/sec)" & parameter == "velocity" ~ value*3.28,
                           parameter == "velocity" & value > 50 ~ NA_real_,
                           # convert F to C
                           unit == "°F" & parameter == "temperature" ~ (value-32)*(5/9),
                           T ~ value)) %>% 
  group_by(date, site, subsite, parameter) %>% 
  summarise(value = mean(value, na.rm = T),
            value = ifelse(is.nan(value), NA_real_, value)) %>% 
  mutate(stream = "sacramento river",
         site = "knights landing",
         subsite = as.character(subsite)) %>% 
  bind_rows(knights_landing_env_pre2006)
```

### Lower Sac - Tisdale

#### Columns Removed

The dataset was pulled from the Tisdale CAMP Access database.

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/lower-sac-river/data-raw/tisdale/tisdale_environmental_camp.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "tisdale_rst_environmental.csv"),
               overwrite = TRUE)

tisdale_env_raw <- read_csv(here::here("data", "rst", "tisdale_rst_environmental.csv")) 

tisdale_env <- tisdale_env_raw %>% 
  # convert meters to feet and get rid of nonsensical values which are errors
  mutate(value = case_when(unit == "meters/second (m/sec)" & parameter == "velocity" ~ value*3.28,
                           parameter == "velocity" & value > 50 ~ NA_real_,
                           # convert F to C
                           unit == "°F" & parameter == "temperature" ~ (value-32)*(5/9),
                           # convert ft to m
                           parameter == "water depth" ~ value*0.3048,
                           T ~ value)) %>% 
  group_by(date, site, subsite, parameter) %>% 
  summarise(value = mean(value, na.rm = T),
            value = ifelse(is.nan(value), NA_real_, value)) %>% 
  mutate(subsite = tolower(subsite)) %>% 
  # removed because large majority did not recored units
  filter(!parameter %in% c("dissolved oxygen", "barometer", "conductivity"))
```

### Mill Creek

#### Columns Removed

The dataset included operations and environmental data. Below only
columns for environmental data are selected.

- flow
- temperature
- turbidity
- weather

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/mill-creek/data/mill_rst.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "mill_rst_environmental.csv"),
               overwrite = TRUE)

mill_env <- read_csv(here::here("data", "rst", "mill_rst_environmental.csv")) %>% 
  select(date, site = location, discharge = flow, temperature = water_temperature, turbidity, weather) %>%
  mutate(temperature = round((temperature - 32) * (5/9), 1),
         stream = "mill creek",
         site = "mill creek", #only one site 
         subsite = "mill creek") %>%
  distinct() %>%
  mutate_if(is.numeric, as.character) %>% 
  pivot_longer(cols = c("discharge", "turbidity", "weather", "temperature"), values_to = "text", names_to = "parameter") %>% 
  mutate(value = as.numeric(text),
         text = ifelse(!is.na(value), NA_character_, text)) %>% 
  glimpse()

```

### Yuba River

#### Columns Removed

The dataset included operations and environmental data. Below only
columns for environmental data are selected.

*Yuba River has 3 subsites. They also reported data that did not link to a specific subsite. For these data we listed subsite as NA. 

- temperature
- turbidity
- velocity

```{r, message=FALSE, warning=FALSE}
gcs_get_object(object_name = "rst/yuba-river/data/yuba_rst.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data", "rst", "yuba_rst_environmental.csv"),
               overwrite = TRUE)

yuba_env <- read_csv(here::here("data", "rst", "yuba_rst_environmental.csv"), 
                     col_types = list("T", "t", "c", "n", "n", "n", "c", "n", 
                                      "n", "c", "n", "n", "c", "c", "n", "n", 
                                      "c", "n", "c", "c")) %>% 
  mutate(date = lubridate::as_date(date)) %>%
  select(date, temperature, turbidity, velocity, site = location) %>%
  mutate(stream = "yuba river",
         subsite = case_when(site == "Yuba River" ~ "yub",
                             site == "RST 1 at Hallwood on Yuba River" ~ "hal",
                             site == "RST 2 at Hallwood" ~ "hal2",
                             site == "5 foot RST at Hallwood" ~ "hal3"),
         site = ifelse(subsite == "yub", "yuba river", "hallwood"),
         temperature = ifelse(temperature > 40, round((temperature - 32) * (5/9), 1), temperature)) %>% # appears to be F and C temps at yuba 
  distinct() %>%
  pivot_longer(cols = c("turbidity", "velocity", "temperature"), values_to = "value", names_to = "parameter") %>% 
  glimpse()

unique(yuba_env$subsite)
unique(yuba_env$site)
```

## Combine data:

```{r}
combined_env <- bind_rows(battle_env, butte_env, clear_env, deer_env,
                          feather_env, knights_env, tisdale_env, 
                          mill_env, yuba_env) %>% 
  mutate(text = case_when(parameter == "weather" & text == "partially cloudy" ~ "partly cloudy",
                          parameter == "weather" & text == "foggy" ~ "fog",
                          parameter == "weather" & text %in% c("precipitation", "rain", "rainy", "windy and precipitation") ~ "precipitation",
                          parameter == "weather" & text %in% c("sunny", "clear") ~ "clear",
                          parameter == "weather" & text %in% c("windy", "windy and clear") ~ "windy",
                          T ~ text),
         # remove temperature above 40 because errors
         value = case_when(parameter == "temperature" & value > 40 ~ NA_real_,
                           T ~ value)) %>% 
  # we do not need to keep rows for parameters where no data was collected because it does not provide any additional information
  filter(!is.na(value) | !is.na(text)) %>% 
  # decided to remove this because not needed for standard datasets
  select(-trap_visit_id)
```

## Explore Varibles {.tabset}

### date

```{r}
combined_env %>% 
  ggplot() + 
  geom_point(aes(x = date, y = site, color = stream), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) + 
  theme(legend.position = "none", 
        text = element_text(size = 18)) 
```

### velocity

Battle, Butte, Clear, and Yuba provide velocity data.

```{r}
combined_env %>% 
  filter(parameter == "velocity") %>% 
  ggplot() + 
  geom_boxplot(aes(x = value, y = stream, color = stream), alpha = .5, binwidth = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

Zoom in on 0 - 10 velocity range:

```{r}
combined_env %>% 
  filter(parameter == "velocity") %>% 
  filter(value < 10) %>%
  ggplot() + 
  geom_histogram(aes(x = value, fill = stream), alpha = .5, binwidth = .3, position="identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### turbidity

All watersheds provide turbidity data

```{r}
combined_env %>% 
  filter(parameter == "turbidity") %>% 
  ggplot() + 
  geom_boxplot(aes(x = value, y = stream, color = stream), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

Zoom in on 0 - 25 turbidity range:

```{r}
combined_env %>% 
  filter(parameter == "turbidity") %>% 
  filter(value < 50) %>% # 1,647 values above 50 (274,049 rows total)
  ggplot() + 
  geom_histogram(aes(x = value, fill = stream), alpha = .5, binwidth = .3, position="identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### weather

Battle, Butte, Clear Deer, and Mill collect weather data:

```{r}
unique(combined_env$weather)
combined_env %>% 
  filter(parameter == "weather") %>% 
  ggplot() + 
  geom_bar(aes(x = text, fill = stream), position = "dodge") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

### habitat

Only battle and clear give habitat:

```{r}
combined_env %>% 
  filter(parameter == "habitat") %>% 
  ggplot() + 
  geom_bar(aes(x = text, fill = stream), position = "dodge") +
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))


```

### stream

```{r}
table(combined_env$stream, useNA = "ifany")
```

### temperature

Butte, Deer, Feather, Knights Landing, Mill, and Yuba provide
temperature data

```{r}
combined_env %>% 
  filter(parameter == "temperature") %>% 
  ggplot() + 
  geom_boxplot(aes(x = value, y = stream, color = stream), alpha = .5, binwidth = .5) + 
  theme_minimal() +
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

Zoom in on 0 - 25 temperature range:

```{r}
combined_env %>% 
  filter(parameter == "temperature", value < 30) %>% # Filter out outliers
  ggplot() + 
  geom_histogram(aes(x = value, fill = stream), alpha = .5, binwidth = 2.5, position = "identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### secchi

```{r}
combined_env %>% 
  filter(parameter == "light penetration") %>% 
  ggplot() + 
  geom_boxplot(aes(x = value, y = stream, color = stream), alpha = .5, binwidth = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

```{r}
combined_env %>% 
  filter(parameter == "light penetration") %>% 
  ggplot() + 
  geom_histogram(aes(x = value, fill = stream), binwidth = 2.5, position = "identity", alpha = .5) + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### site

```{r}
table(combined_env$site, useNA = "ifany")
```

### flow

```{r}
combined_env %>% 
  filter(parameter == "discharge") %>% 
  ggplot() + 
  geom_boxplot(aes(x = value, y = stream, color = stream), alpha = .5, binwidth = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

```{r}
combined_env %>% 
  filter(parameter == "discharge") %>% 
  ggplot() + 
  geom_histogram(aes(x = value, fill = stream), binwidth = 10) + 
  theme_minimal() + 
  facet_wrap(~stream, ncol = 1, drop = TRUE, scales = "free") + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

### subsite

There are no missing subsites

```{r}
filter(combined_env, is.na(subsite))
table(combined_env$subsite, useNA = "ifany")
```

## Save data to google cloud 

```{r, eval = FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(reordered_env,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_RST_environmental.csv")
write_csv(reordered_env, "../../data/standard-format-data/standard_RST_environmental.csv")
```
