---
title: "Standardize Adult Redd Datasets"
author: "Maddee Rubenson (FlowWest)"
date: '2022-07-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
library(knitr)
color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

## Adult Redd Data Standardization

FlowWest received Adult Redd data from four monitoring programs:

-   Feather River
-   Battle Creek
-   Clear Creek
-   Mill Creek
-   Yuba River

Redd data were pulled together at a daily level. Mill Creek conducts one annual survey per year and produces an annual count, and so those data were added to the daily data with a pseudo date as October 1. This is to allow Mill Creek data to be added to the database without creating a separate table. 

Previously a standard annual table was created in this markdown but this is now created and documented in `SRJPEdata`.

## Standard format for Adult All Redd Data

Data dictionary for standard format for `redd_data_final`:

(B - Battle Creek, C - Clear Creek, M - Mill Creek,
F - Feather River, Y - Yuba River)

|column_name              |tributary_collects |definitions                                                        |
|:------------------------|:------------------|:------------------------------------------------------------------|
|date                     |B, C, F, M, Y      |Date redd data was collected                                       |
|latitude                 |B, C, F, Y         |Latitude of redd location                                          |
|longitude                |B, C, F, Y         |Longitude of redd location                                         |
|reach                    |B, C, F, M         |Reach sampled                                                      |
|river_mile               |B, C               |River mile associated with redd measured                           |
|redd_id                  |B, C, Y            |Unique redd ID                                                     |
|fish_guarding            |B, C               |Whether or not fish were observed guarding the redd (T/F)          |
|redd_measured            |B, C               |Whether or not redd was measured (T/F)                             |
|redd_width               |B, C, F            |Width of redd measured (m)                                         |
|redd_length              |B, C, F            |Length of redd measured (m)                                        |
|velocity                 |B, C, F, Y         |Measured stream velocity (ft/s)                                    |
|age                      |B, C               |Age of redd                                                        |
|age_index                |B, C               |Number of times redd was aged. If 0, redd was sampled but not aged |
|redd_count               |B, C, F, M, Y      |Number of redds sampled                                            |
|stream                   |B, C, F, M, Y      |Stream associated with redd data collection                        |
|year                     |B, C, F, M, Y      |Year data were collected                                           |
|survey_method            |C                  |Survey method                                                      |
|run                      |B, C, F, M, Y      |Run of fish associated with redd                                   |
|species                  |B, C, F, M, Y      |Species associated with redd                                       |
|depth_m                  |F, Y               |Depth of water (m)                                                 |
|starting_elevation_ft    |M                  |Starting elevation (ft)                                            |
|num_of_fish_on_redd      |Y                  |Number of fish on redd                                             |
|redd_substrate_class     |B, C               |Categorical, size class determined by millimeter size of substrate |
|tail_substrate_class     |B, C               |Categorical, size class determined by millimeter size of substrate |
|pre_redd_substrate_class |B, C               |Categorical, size class determined by millimeter size of substrate |                                         |


## Read in data {.tabset}

Below we read in the adult redd data for each monitoring program and rename or select columns so that we can join all the monitoring datasets together in the section below.

### Battle Creek

This contains only spring run chinook.

#### Columns Removed

-   `why_not_measured`: enumerated; `NA`, `fish_on_redd`, `sub sample`, `too deep` Deemed unnecessary.
-   `flow_meter`: type of flow meter used; deemed unnecessary for final dataset
-   `start_number_flow_meter`: deemed unnecessary for final dataset
-   `end_number_flow_meter`: deemed unnecessary for final dataset
-   `flow_meter_time`: deemed unnecessary for final dataset
-   `start_number_flow_meter_80`: deemed unnecessary for final dataset
-   `end_number_flow_meter_80`: deemed unnecessary for final dataset
-   `flow_meter_time_80`: deemed unnecessary for final dataset
-   `redd_pit_depth`: deemed unnecessary for final dataset
-   `redd_tail_depth`: deemed unnecessary for final dataset

columns created:

-   `stream`: used Battle Creek
-   `flow_fps` changed to `velocity`

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "battle_redd.csv"),
  overwrite = TRUE
)
# Redd data for Battle Creek
battle_redd <- read_csv(here::here("data", "redd-carcass-holding", "battle_redd.csv"), 
                        col_names = TRUE)

summary(battle_redd)

table(battle_redd$pre_redd_substrate_size, useNA = "ifany")
table(battle_redd$redd_substrate_size, useNA = "ifany")
table(battle_redd$tail_substrate_size, useNA = "ifany")
unique(battle_redd$why_not_measured)
unique(battle_redd$flow_meter)

clean_battle_redd <- battle_redd  |>
  select(date, latitude, longitude, 
         reach, river_mile, fish_guarding, 
         redd_measured, redd_width, redd_length, JPE_redd_id,
         pre_redd_substrate_size, redd_substrate_size, 
         tail_substrate_size, 
         flow_fps, age, age_index)  |>
  mutate(# adding redd count for only those redds surveyed and not aged, or aged once
         redd_count = ifelse(age_index %in% c(0, 1), 1, 0),
         stream = 'battle creek', 
         year = lubridate::year(date))  |>
  rename(velocity = flow_fps,
         redd_id = JPE_redd_id)  |>
  relocate(redd_id, .before = fish_guarding) |> 
  glimpse()

```

### Clear Creek

FlowWest acquired raw Clear Creek redd data in November 2023.

#### Columns Removed

```{r}
# data directly from Natasha
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "clear_redd.csv"),
  overwrite = TRUE
) 

# Redd data for Clear Creek
clear_redd <- read_csv(here::here("data", "redd-carcass-holding", "clear_redd.csv"), 
                        col_names = TRUE, 
                        col_types = c("c", "D", "c", "n", "n", "n", "n", "n", "c",
                                      "n", "c", "c", "c", "n", "n", "c", "c", "c",
                                      "c", "c", "l", "l", "c", "D", "n", "n", "n",
                                      "n", "n", "n", "n", "n", "n", "n", "n", "n",
                                      "c", "n", "n")) |> 
                          glimpse()

summary(clear_redd)

table(clear_redd$pre_redd_substrate_size, useNA = "ifany")
table(clear_redd$redd_substrate_size, useNA = "ifany")
table(clear_redd$tail_substrate_size, useNA = "ifany")
unique(clear_redd$why_not_measured)
unique(clear_redd$flow_meter)

clean_clear_redd <- clear_redd  |>
  select(date, latitude, longitude, 
         reach = surveyed_reach, river_mile, JPE_redd_id, 
         fish_guarding = fish_on_redd, redd_measured = measured, redd_width,
         redd_length, age, age_index, pre_redd_substrate_size, 
         redd_substrate_size, tail_substrate_size, velocity,
         survey_method, run, species)  |>
  mutate(stream = 'clear creek', 
         # adding redd count for only those redds surveyed and not aged, or aged once
         redd_count = ifelse(age_index %in% c(0, 1), 1, 0),
         year = lubridate::year(date))  |>
  rename(redd_id = JPE_redd_id) |> 
  glimpse() 

```

### Mill Creek

-   changed `location` to `reach` to align with other datasets

#### Columns Removed

- `surveyed` column

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/mill-creek/data/mill_redd_survey.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "mill_redd.csv"),
  overwrite = TRUE
)
mill_redd <- read_csv(here::here("data", "redd-carcass-holding", "mill_redd.csv"))  |>  glimpse()

summary(mill_redd)

clean_mill_redd <- mill_redd  |>
  rename(reach = location)  |>
  filter(surveyed %in% c(TRUE, NA)) |> # remove years where reach was not surveyed
  mutate(stream = 'mill creek',
         # pseudo date to add into daily data
         date = as_date(paste0(year, "-10-01")))  |>
  select(-surveyed) |> 
  glimpse()
```

### Feather River

-   changed `location` to `reach` to align with other datasets
-   `velocity_m_per_s` converted to feet per second and named velocity

#### Columns Removed

-   `type`: enumerated as `Area`, `Point`, `Questionable Redds`, or `NA`; deemed unnecessary
-   `salmon_count`: integer ranging from 0 to 200; deemed unnecessary
-   `pot_depth_m`: numeric ranging from 0 to 50; deeemd unnecessary
-   `percent_fine_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_small_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_medium_substrate`: numeric ranging from 0 to 95; mostly `NA` and deemed unnecessary
-   `percent_large_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_boulder`: numeric ranging from 0 to 70; mostly `NA` and deemed unnecessary

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "feather_redd.csv"),
  overwrite = TRUE
)
feather_redd <- read_csv(here::here("data", "redd-carcass-holding", "feather_redd.csv"), 
                         col_names = TRUE)

summary(feather_redd)
table(feather_redd$type, useNA = "ifany")

clean_feather_redd <- feather_redd  |>
  select(date, location, redd_count, 
         latitude, longitude, 
         redd_width_m, redd_length_m, velocity, depth_m)  |> 
  rename(reach = location,
         redd_width = redd_width_m, 
         redd_length = redd_length_m)  |>
  mutate(stream = "feather river",
         year = lubridate::year(date),
         date = as.Date(date),
         velocity = velocity / 0.3048)  |> # convert to feet per second 
  glimpse()

```

### Yuba River

#### Columns Created:

-   `stream` set to Yuba River

#### Columns Removed:

None

```{r warning=FALSE}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/yuba-river/data/yuba_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "yuba_redd.csv"),
  overwrite = TRUE
)
yuba_river_raw <- read_csv(here::here("data", "redd-carcass-holding", "yuba_redd.csv")) |> 
  glimpse()

clean_yuba_redd <- yuba_river_raw |> 
  mutate(stream = "yuba river",
         redd_count = 1) # each row is a redd observation 
```

## Combine data:

This is the daily tabular redd data. 

```{r}
all_redd_data <- bind_rows(clean_battle_redd, 
                           clean_clear_redd,
                           clean_feather_redd,
                           clean_mill_redd,
                           clean_yuba_redd)  |> 
  mutate(species = case_when(species %in% c("chinook","CHN") ~ "chinook",
                             species %in% c("UNK", "Unknown", "Salmonid") ~ "unknown",
                             species %in% c("SH", "O. mykiss") ~ "steelhead",
                             is.na(species) ~ "not recorded"),
         run = ifelse(is.na(run), "not recorded", run))
```

## Save file without standardized reaches
We need all raw reaches to pull into `analysis/adult-reach-analysis.R`

# Save data
```{r, save_data}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_redd_data,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_redd_raw_reaches.csv",
           predefinedAcl = "bucketLevel")
```

## Survey Reach Standardization

Survey reach naming conventions changed over time for several monitoring programs. A standardized reach lookup table was created in `analysis/adult-reach-analysis.R` and is merged in here to standardize all reaches.

```{r echo=TRUE, message=FALSE, warning=FALSE}
gcs_get_object(
  object_name = "jpe-model-data/standard_reach_lookup.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "standard-format-data", "standard-reach-lookup.csv"),
  overwrite = TRUE
)

standard_reach_lookup <- read_csv(here::here("data", "standard-format-data", "standard-reach-lookup.csv")) |> 
  select(stream, reach, standardized_reach)

redd_data_final <- all_redd_data |> 
  left_join(standard_reach_lookup, by = c("stream", "reach")) |> 
  select(-reach) |> 
  rename(reach = standardized_reach) |> 
  relocate(reach, .before = river_mile) |> 
  glimpse()
```

## Substrate Size Standardization

Substrate sizes were given names and classes following the Wentworth Scale, created by W.C. Krumbein. This scale is what is commonly used in the United States. The substrate size provided for redd data was standardized using this scale. When the size range fell into two categories, they were rounded down.

| Class              | Size Range (mm) |
|--------------------|-----------------|
| Boulder            | \>256           |
| Cobble             | 64-256          |
| Very coarse gravel | 32-64           |
| Coarse gravel      | 16-32           |
| Medium gravel      | 8-16            |
| Fine gravel        | 4-8             |
| Very fine gravel   | 2-4             |
| Very coarse sand   | 1-2             |
| Coarse sand        | 0.5-1           |
| Medium sand        | 0.25-0.5        |
| Fine sand - clay   | \<0.25          |

```{r echo=TRUE, message=FALSE, warning=FALSE}
# standarized size ranges lookup
substrate_class = data.frame("standardized_size_range" = c("<0.25", 
                                                           "0.25-0.5",
                                                           "0.5-1",
                                                           "1-2",
                                                           "2-4", 
                                                           "4-8", 
                                                           '8-16', 
                                                           ">16"),
                              "redd_substrate_class" = c("fine sand",
                                                    "medium sand",
                                                    "coarse sand", "very coarse sand",
                                                    "very fine gravel", "fine gravel",
                                                    "medium gravel", 
                                                    "coarse gravel to boulder"))

unique(redd_data_final$redd_substrate_size)

redd_substrate_size_lookup <- 
  data.frame("redd_substrate_size" = unique(redd_data_final$redd_substrate_size),
             "standardized_size_range" = c(NA, "1-2", "2-4", "2-4", "2-4", 
                                           "0.5-1", "2-4","2-4", '8-16', "4-8", 
                                           "1-2", "1-2", "4-8", "0.5-1", "1-2", 
                                           "2-4", "0.5-1", "1-2", "2-4", "2-4", 
                                           "<0.25", "4-8", "4-8", "0.5-1", "4-8",
                                           "8-16")) |> 
  left_join(substrate_class)
                    
tail_substrate_size_lookup <- redd_data_final  |> 
  ungroup()  |>
  select(tail_substrate_size)  |> 
  full_join(redd_substrate_size_lookup  |> 
              rename(tail_substrate_size = redd_substrate_size))  |> 
  distinct()  |>
  rename(tail_substrate_class = redd_substrate_class)

pre_redd_substrate_size_lookup <- redd_data_final  |> 
  ungroup()  |>
  select(pre_redd_substrate_size)  |> 
  full_join(redd_substrate_size_lookup  |> 
              rename(pre_redd_substrate_size = redd_substrate_size))  |> 
  distinct()  |>
  rename(pre_redd_substrate_class = redd_substrate_class) 

redd_data_final <- redd_data_final  |> 
  left_join(redd_substrate_size_lookup)  |>
  select(-standardized_size_range)  |>
  left_join(tail_substrate_size_lookup)  |>
  select(-standardized_size_range)  |>
  left_join(pre_redd_substrate_size_lookup)  |>
  select(-standardized_size_range, -pre_redd_substrate_size, 
         -redd_substrate_size, -tail_substrate_size)  

```

## Explore Variables {.tabset}

### stream

```{r}
table(redd_data_final$stream)
```

### date

```{r}
summary(redd_data_final$date)

redd_data_final  |>
  group_by(stream)  |>
  summarise(year_range = paste0(min(year, na.rm = TRUE), " : ", max(year, na.rm = TRUE)), 
            months = paste0(unique(sort(lubridate::month(date))), collapse = ", "))  |>
  ungroup()  |>
  distinct()  |> knitr::kable()

```

### velocity

```{r}
summary(redd_data_final$velocity)

ggplot() + 
  geom_point(data = redd_data_final, aes(x = date, y = velocity, color = reach)) + 
  facet_wrap(~ stream) +
  theme(legend.position = "none")
```

### fish_guarding

```{r}
table(redd_data_final$fish_guarding, useNA = "ifany")
```

### redd_measured

```{r}
table(redd_data_final$redd_measured, useNA = "ifany")

```

### redd_width and redd_length

```{r}

summary(redd_data_final$redd_width)
summary(redd_data_final$redd_length)

ggplot() + 
  geom_histogram(data = redd_data_final, aes(redd_width)) 
  
ggplot() +   
  geom_histogram(data = redd_data_final, aes(redd_length))

```

### pre_redd_substrate_class

```{r}
table(redd_data_final$pre_redd_substrate_class, useNA = "ifany")

```

### redd_substrate_class

```{r}
table(redd_data_final$redd_substrate_class, useNA = "ifany")
```

### tail_substrate_class

```{r}
table(redd_data_final$tail_substrate_class, useNA = "ifany")

```


## Exploratory Plots

```{r}
# a plot to help understand sampling frequency across streams 
redd_data_final  |>
  group_by(date, stream)  |>
  summarise(n_reaches = length(unique(reach)))  |>
  ggplot() + 
  geom_point(aes(x = date, y = n_reaches)) + 
  ylab('number of samples per reach and date') + 
  facet_wrap(~ stream)
```

## Save Cleaned Data to Google Cloud

```{r}
knitr::kable(redd_data_final  |> head())
```

```{r, eval=FALSE}
write_csv(redd_data_final, here::here("data","redd-carcass-holding","standard_all_redd_data.csv"))
```

```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(redd_data_final,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_daily_redd.csv",
           predefinedAcl = "bucketLevel")
```
