---
title: "Standardize Adult Upstream Passage Datasets"
author: "Erin Cain"
date: '2022-05-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")
```

## Adult Upstream Passage Data Standardization

FlowWest received Adult Upstream Passage data from 6 monitoring programs:

-   Battle Creek
-   Butte Creek
-   Clear Creek
-   Deer Creek
-   Mill Creek
-   Yuba River

## Standard format for Adult Upstream Passage Data

Data dictionary for standard format:

(B - Battle Creek, Bu - Butte Creek, C - Clear Creek,  
D - Deer Creek, M - Mill Creek, Y - Yuba River)

| column name        | tributary collects | definition                                                                                             |
|:------------------|:-----------------|:-----------------------------------|
| stream             |                    | Which Spring Run JPE stream is the data from                                                           |
| location           |                    | what is the location name                                                                              |
| date               |                    | date of video footage                                                                                  |
| time               |                    | time of video footage                                                                                  |
| hours_viewed       |                    | number of hours viewed by day                                                                          |
| count              |                    | number of fish observed                                                                                |
| adipose            |                    | if adipose fin is present (TRUE/FALSE)                                                                 |
| run                |                    | run designation                                                                                        |
| run_method         |                    | method used to identify run                                                                            |
| passage_direction  |                    | direction of fish passage                                                                              |
| sex                |                    | sex of fish observed                                                                                   |
| length             |                    | length of fish observed in feet TODO check units                                                       |
| viewing_condition  |                    | direction of fish observed (normal, readable, not readable, weir is flooded)                           |
| spawning_condition |                    | description of spawning status based on coloration (none, energetic, spawning colors, fungus, unknown) |

## Read in data {.tabset}

Below we read in the adult upstream passage data for each monitoring program and rename or select columns so that we can join all the monitoring datasets together in the section below.

### Battle Creek

#### Columns Removed 

No columns are removed from battle creek upstream passage video data. 
```{r}
# Set your authentication using gcs_auth
# gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# # Set global bucket
# gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
# download file as csv
# gcs_get_object(object_name = "adult-upstream-passage-monitoring/battle-creek/data/battle_passage_video.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/adult-upstream-passage-monitoring/battle_adult_upstream.csv",
#                overwrite = TRUE)

battle_upstream_passage <- read_csv("../../data/adult-upstream-passage-monitoring/battle_adult_upstream.csv") %>% 
  mutate(stream = "Battle Creek") %>% 
  glimpse
```

### Clear Creek

#### Columns Removed 

* **time_bloc**: Time block was removed because it refers to the time block being viewed and this is not relevent if we only care speicifically about when the fish passes. We retained the time_passed column instead to keep this information 

```{r}
# download file as csv
# gcs_get_object(object_name = "adult-upstream-passage-monitoring/clear-creek/data/clear_passage.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/adult-upstream-passage-monitoring/clear_adult_upstream.csv",
#                overwrite = TRUE)

clear_upstream_passage <- read_csv("../../data/adult-upstream-passage-monitoring/clear_adult_upstream.csv") %>% 
  select(-time_block) %>%
  mutate(stream = "Clear Creek") %>% 
  glimpse
```

### Deer Creek

#### Columns Removed 

```{r}
# download file as csv
# gcs_get_object(object_name = "adult-upstream-passage-monitoring/deer-creek/data/deer_upstream_passage_estimate.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/adult-upstream-passage-monitoring/deer_adult_upstream.csv",
#                overwrite = TRUE)

deer_upstream_passage <- read_csv("../../data/adult-upstream-passage-monitoring/deer_adult_upstream.csv") %>% 
    rename(count = passage_estimate) %>% # confirm with matt johnson that this is a count not a modeled passage
    mutate(stream = "Deer Creek") %>% 
    glimpse
```

### Mill Creek 

#### Columns Removed 

```{r}

# download file as csv
# gcs_get_object(object_name = "adult-upstream-passage-monitoring/mill-creek/data/mill_upstream_passage_estimate.csv",
#                 bucket = gcs_get_global_bucket(),
#                 saveToDisk = "../../data/adult-upstream-passage-monitoring/mill_adult_upstream.csv",
#                 overwrite = TRUE)

mill_upstream_passage <- read_csv("../../data/adult-upstream-passage-monitoring/mill_adult_upstream.csv") %>% 
  rename(count = passage_estimate) %>% # confirm with matt johnson that this is a count not a modeled passage
  mutate(stream = "Mill Creek") %>% 
  glimpse
```


### Yuba River

#### Columns Removed 

```{r}

# download file as csv
# gcs_get_object(object_name = "adult-upstream-passage-monitoring/yuba-river/data/yuba_upstream_passage.csv",
#                 bucket = gcs_get_global_bucket(),
#                 saveToDisk = "../../data/adult-upstream-passage-monitoring/yuba_adult_upstream.csv",
#                 overwrite = TRUE)

yuba_upstream_passage <- read_csv("../../data/adult-upstream-passage-monitoring/yuba_adult_upstream.csv") %>% 
  select(-category, -length_cm, -speed_m_per_s, -depth_m, -position_in_frame) %>% # No clear methodology shared by yuba on these data and how they should be used  
  mutate(stream = "Yuba River") %>% 
  glimpse
```


## Combine data: 
```{r}
combined_upstream_passage <- bind_rows(battle_upstream_passage, clear_upstream_passage, 
                                       deer_upstream_passage, mill_upstream_passage, 
                                       yuba_upstream_passage) %>% 
  glimpse
```

## Explore Varibles{.tabset} 
