---
title: "Standardize Carcass Data"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)

color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")
```

## Carcass Data Standardization

FlowWest received carcass data from 5 monitoring programs:

-   Battle Creek
-   Butte Creek
-   Clear Creek
-   Feather River
-   Yuba River

## Standard format Carcass Data

Data dictionary for standard format:

(B - Battle Creek, Bu - Butte Creek, C - Clear Creek, F - Feather River, Y - Yuba River)

| column name   | stream collects         | definition                                               |
|:-----------------|:-------------------|:---------------------------------|
| longitude      | B, C | Longitude of survey site                    |
| latitude         | B, C | Latitude of survey site     |
| river_mile | B, C, Y | River mile where the survey was performed |
| date | B, Bu, F, C, Y | Date the survey was performed |
| survey_method | B, C | Method by which the survey was performed (i.e. snorkel survey, barrier weir trap, kayak survey, fish rescue, etc.)
| location | B | Butte Creek's variable for location of survey |
|sex | B, Bu, F, C, Y | Sex of the carcass |
|fork_length | B, Bu, F, C, Y | Fork length of the carcass (mm) |
| adipose | B, Bu, F, C, Y | Whether or not the adipose fin was clipped |
| carcass_condition | B, Bu, F, C, Y | Condition of the carcass when sampled (i.e. fresh, decayed, non-fresh, bright, etc.) |
| spawn_condition | B, Bu, F, C, Y | Spawning condition of the carcass sampled (i.e. spawned, unspawned, etc.) |
| run | B, F, C | Run of the carcass sampled, if assigned |
| stream | B, Bu, F, C, Y | Stream from which the data was collected |
| chop_count | B, Bu, F, C, Y | Count of chopped carcasses |
| survey_id | Bu, F | ID of the survey performed |
| section_cd | Bu | Butte creek-specific location information, section of river surveyed |
| way_pt | Bu | Butte creek-specific location information, way point of survey |
| tag_count | Bu, F | Count of tagged carcasses |
| tag_col | F, C | Color of tag (generally corresponds to week of survey) |
| sec | F | Feather river-specific location information, section of river surveyed |
| unit | F | Feather river-specific location information, unit of river surveyed |
| chan | F | Feather river-specific location information, channel of river surveyed |
| min | F | Feather river-specific location information |
| chan_id | F | Feather river-specific location information, ID of channel surveyed |
| reach | C | Clear creek-specific location information, reach of river surveyed |



## Read in data {.tabset}

### Battle Creek

#### Columns Removed

`comments`, `year`, `observed_only`, `cwt_code`

```{r, message=FALSE, warning=FALSE, include = FALSE}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_carcass.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/redd-carcass-holding/battle_carcass.csv",
               overwrite = TRUE)
```
```{r, message = FALSE, warning = FALSE}

battle_carcass <- read_csv("data/redd-carcass-holding/battle_carcass.csv") |> 
  mutate(stream = "battle creek") |> 
  rename(carcass_condition = carcass_live_status,
         survey_method = method) |> 
  mutate(count = 1) |> 
  select(-c(comments, year, observed_only, cwt_code)) |> 
  glimpse()
# each row is a carcass in table
```

### Butte Creek

#### Columns Removed

`disc_tag_applied`, `scale_nu`, `tissue_nu`, `otolith_nu`, `comments`, `year`

```{r, message=FALSE, warning=FALSE, include = FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/butte-creek/data/butte_carcass_chops.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/redd-carcass-holding/butte_carcass.csv",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/butte-creek/data/butte_carcass_2014-2016.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/redd-carcass-holding/butte_carcass_2014-2016.csv",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/butte-creek/data/butte_carcass_2017-2020.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/redd-carcass-holding/butte_carcass-2017-2020.csv",
               overwrite = TRUE)
```
```{r, message = FALSE, warning = FALSE}

# this may contain carcasses that were chopped but did not have any biological information recorded
butte_carcass_bulk_chops <- read_csv("data/redd-carcass-holding/butte_carcass.csv") |> 
  rename(sect = section_cd,
         mark_recapture = disposition,
         count = chop_count,
         adipose_fin_clipped = ad_fin_clip,
         carcass_condition = condition) |> 
  mutate(source_table = "chops") |> 
  glimpse()
# each row is an individual for these
butte_carcass_2014_2016 <- read_csv("data/redd-carcass-holding/butte_carcass_2014-2016.csv") 
butte_carcass_2017_2020 <- read_csv("data/redd-carcass-holding/butte_carcass-2017-2020.csv")

butte_carcass_individuals <- bind_rows(butte_carcass_2014_2016, butte_carcass_2017_2020) |> 
  select(-c(survey, year, comments)) |> 
  rename(mark_recapture = disposition,
         tag_id = disc_tag_applied,
         spawn_condition = spawning_status,
         adipose_fin_clipped = ad_fin_clip_cd,
         otolith_sample = otolith_nu,
         tissue_sample = tissue_nu,
         fork_length = fork_length_mm,
         carcass_condition = condition,
         sect = section_cd) |> 
  mutate(spawn_condition = case_when(spawn_condition == "yes" ~ "spawned",
                                     spawn_condition == "no" ~ "unspawned",
                                     spawn_condition == "p" ~ "partial",
                                     TRUE ~ spawn_condition),
         carcass_condition = case_when(carcass_condition == "f" ~ "fresh",
                                       carcass_condition == "d" ~ "decayed",
                                       carcass_condition == "not recorded" ~ NA_character_,
                                       TRUE ~ carcass_condition),
         source_table = "carcass_individual",
         count = 1,
         scale_sample = as.character(scale_nu)) |> 
  select(-c(scale_nu)) |> 
  glimpse()

# combine bulk chops and individual chops
butte_carcass_combined <- bind_rows(butte_carcass_individuals, butte_carcass_bulk_chops) |> 
  mutate(stream = "butte creek") |> glimpse()

```

### Clear Creek


```{r, message=FALSE, warning=FALSE, include = FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_carcass.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/redd-carcass-holding/clear_carcass.csv",
               overwrite = TRUE)
```
```{r, message = FALSE, warning = FALSE}

clear_carcass <- read_csv("data/redd-carcass-holding/clear_carcass.csv")  |> 
  select(-c(obs_only, tis_eth, why_sex_unknown, why_not_sp, photo, comments,
            cwt_code, brood_year, release_location, hatchery, age, mark_rate,
            verification_and_cwt_comments, run_call, genetic)) |> 
  rename(survey_method = type,
         tag_id = sample_id, 
         carcass_condition = condition,
         tissue_sample = tis_dry,
         scale_sample = scale,
         otolith_sample = otolith_st,
         spawn_condition = spawn_status,
         head_sample = head_retrieved,
         tag_col = tag_type) |> 
  mutate(adipose_fin_clipped = case_when(adipose == "present" ~ TRUE,
                                         adipose == "absent" ~ FALSE,
                                         is.na(adipose) ~ NA),
         stream = "clear creek",
         count = 1,
         otolith_sample = as.character(otolith_sample),
         scale_sample = as.character(scale_sample)) |> 
  select(-c(adipose)) |> 
  glimpse()
```

### Feather River

We removed a lot of variables from the feather carcass data in order to combine it with the other streams. However, some of these may still be helpful in some analyses. 

```{r, message=FALSE, warning=FALSE, include = FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/redd-carcass-holding/feather_carcass.csv",
               overwrite = TRUE)
```
```{r, message = FALSE, warning = FALSE}

feather_carcass <- read_csv("data/redd-carcass-holding/feather_carcass.csv") |>
  filter(species %in% c("chinook salmon", NA)) |>  
  mutate(stream = "feather river",
         sect = as.character(sect),
         otolith_sample = as.character(otolith_sample),
         scale_sample = as.character(scale_sample)) |>
  rename(spawn_condition = spawning_condition) |> 
  select(-c(rel_loc, hall_print2, tag_color, hallprint_color, year, lifestage, flow, vis, weather, min, recov_color, hallprint1)) |> glimpse()
```

### Yuba River


```{r, message=FALSE, warning=FALSE, include = FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/yuba-river/data/yuba_carcass.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data/redd-carcass-holding/yuba_carcass.csv",
               overwrite = TRUE)
```

```{r, message = FALSE, warning = FALSE}
yuba_carcass <- read_csv("data/redd-carcass-holding/yuba_carcass.csv") |> 
  rename(adipose_fin_clipped = adipose_clip,
         spawn_condition = spawned,
         count = chop_count) |> 
  mutate(stream = "yuba river") |> glimpse()
```
    

### Combined
#### remove location variables so they are not in the standard dataset but can be easily included again.

```{r}
carcass_combined <- bind_rows(battle_carcass, 
                              butte_carcass_combined,
                              feather_carcass,
                              clear_carcass,
                              yuba_carcass) |> 
  select(-c(species, tissue_sample, otolith_sample, source_table, scale_sample,
            samples_collected, head_sample, way_pt, sect, longitude, latitude,
            river_mile, location, unit, chan, river_section, reach)) |> 
  glimpse()
```
### Upload standardized data to google cloud
```{r, include = FALSE}
f <- function(input, output) write_csv(input, file = output)

# combined chops and tags
gcs_upload(carcass_combined,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_carcass.csv",
           predefinedAcl = "bucketLevel")
```

