---
title: "Standardize Adult Redd Datasets"
author: "Maddee Rubenson (FlowWest)"
date: '2022-07-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
library(knitr)
color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

## Adult Redd Data Standardization

FlowWest received Adult Redd data from four monitoring programs:

-   Feather River
-   Battle Creek
-   Clear Creek
-   Mill Creek
-   Yuba River

Due to the format of the data received, two redd datasets were developed. The first dataset includes Feather, Battle, Yuba, and Clear Creek daily data and the second dataset includes Feather, Battle, Clear, Yuba, and Mill Creek and has a yearly maximum redd counts value.

## Standard format for Adult All Redd Data

Data dictionary for standard format for `all_redd_data`:

(B - Battle Creek, C - Clear Creek,\
F - Feather River, Y - Yuba River)

| column name              | tributary collects | definition                                                         |
|:-----------------|:-----------------|:------------------------------------|
| date                     | B, C, F, Y         | Date in which redd data was collected                              |
| latitude                 | B, C, F, Y         | Latitude of redd location                                          |
| longitude                | B, C, F, Y         | Longitude of redd location                                         |
| reach                    | B, C, F            | Unique name identifier of reach sampled                            |
| river_mile               | B, C               | River mile associated with redd measured                           |
| redd_measured            | B, C               | Whether or not a redd was measured (T/F)                           |
| age                      | C                  | integer ranging from 1 - 5; redd age                               |
| fish_guarding            | B, C               | Whether or not a fish is on redd when measured ??                  |
| redd_length              | B, C, F            | Length of measured redd (m)                                        |
| redd_width               | B, C, F            | Width of measured redd (m)                                         |
| run                      | C                  | `spring`, `fall`, `late-fall`                                      |
| stream                   | B, C, F, Y         | Unique stream associated with redd data collected                  |
| redd_substrate_class     | B, C               | categorical; size class determined by millimeter size of substrate |
| pre_redd_substrate_class | B, C               | categorical; size class determined by millimeter size of substrate |
| tail_substrate_class     | B, C               | categorical; size class determined by millimeter size of substrate |
| year                     | B, C, F, Y         | Calendar year sampling occurred                                    |
| velocity                 | B, C, F, Y         | measured stream velocity (ft/s)                                    |
| depth_m                  | F, Y               | Depth of water (meters)                                            |

## Standard format for Adult Redd Data Aggregated by Year

This dataset includes `max_yearly_redd_count` which is the maximum number of redds counted on a single day per year per reach.

Data dictionary for standard format for `max_yearly_redd_count`:

| column name           | tributary collects | definition                                                          |
|:-----------------|:-----------------|:-----------------------------------|
| date                  | B, C, F, Y         | Date in which redd data was collected                               |
| reach                 | B, C, F, M         | Unique name identifier of reach sampled                             |
| stream                | B, C, F, M         | Unique stream associated with redd data collected                   |
| year                  | B, C, F, M, Y      | Calendar year sampling occurred                                     |
| max_yearly_redd_count | B, C, F, M, Y      | maximum number of redds counted on a day in a unique reach per year |
| run                   | C                  | `spring`, `fall`, `late-fall`                                       |

## Read in data {.tabset}

Below we read in the adult redd data for each monitoring program and rename or select columns so that we can join all the monitoring datasets together in the section below.

### Battle Creek

#### Columns Removed

-   `why_not_measured`: enumerated; `NA`, `fish_on_redd`, `sub sample`, `too deep` Deemed unnecessary.
-   `flow_meter`: type of flow meter used; deemed unnecessary for final dataset
-   `start_number_flow_meter`: deemed unnecessary for final dataset
-   `end_number_flow_meter`: deemed unnecessary for final dataset
-   `flow_meter_time`: deemed unnecessary for final dataset
-   `start_number_flow_meter_80`: deemed unnecessary for final dataset
-   `end_number_flow_meter_80`: deemed unnecessary for final dataset
-   `flow_meter_time_80`: deemed unnecessary for final dataset
-   `redd_pit_depth`: deemed unnecessary for final dataset
-   `redd_tail_depth`: deemed unnecessary for final dataset
-   `comments`: deemed unnecessary for final dataset

columns created:

-   `stream`: used Battle Creek
-   `flow_fps` changed to `velocity`

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "battle_redd.csv"),
  overwrite = TRUE
)
# Redd data for Battle Creek
battle_redd <- read_csv(here::here("data", "redd-carcass-holding", "battle_redd.csv"), 
                        col_names = TRUE, 
                        col_types = list("n", "n", "D", "c", "n", "c", 
                                         "c", "c", "l", "l", "l", "n", 
                                         "n", "n", "n", "n", "c", "n", 
                                         "n", "n", "n", "n", "n", "n", 
                                         "c"))  |> glimpse()

summary(battle_redd)

table(battle_redd$pre_redd_substrate_size, useNA = "ifany")
table(battle_redd$redd_substrate_size, useNA = "ifany")
table(battle_redd$tail_substrate_size, useNA = "ifany")
unique(battle_redd$why_not_measured)
unique(battle_redd$flow_meter)

clean_battle_redd <- battle_redd  |>
  select(date, latitude, longitude, 
         reach, river_mile, fish_guarding, 
         redd_measured, redd_width, redd_length, 
         pre_redd_substrate_size, redd_substrate_size, 
         tail_substrate_size, 
         flow_fps)  |>
  group_by(date, reach)  |>
  mutate(redd_count = n(),
         stream = 'Battle Creek', 
         year = lubridate::year(date))  |>
  rename(velocity = flow_fps)  |>
  glimpse()

```

### Clear Creek

FlowWest acquired Clear Creek redd data for late fall and spring runs through CVPIA. This data was integrated into this dataset by taking the maximum date of data provided to JPE spring run and filtering the CVPIA-provided spring run to only include newer data. Late fall run data was bound to this dataset.

O. mykiss data is included in this dataset.

#### Columns Removed

-   `survey`: integer ranging from 1 to 12; deemed unnecessary for final dataset
-   `X1000ftbreak`: deemed unnecessary for final dataset
-   `ucc_relate`: enumerated value of `above` or `below`; deemed unnecessary for final dataset
-   `picket_weir_location`: numeric value ranging from 7.4 to 8.2; deemed unnecessary for final dataset
-   `picket_weir_relation`: this could be used to determine the run type, if needed; enumerated as `above` or `below`
-   `redd_id`: unique identifier to each redd; deemed unnecessary for final redd dataset
-   `gravel`: enumerated value of `combination`, `injection`, or `native`; deemed unnecessary for final redd dataset
-   `fish_on_redd`: TRUE/FALSE
-   `why_not_measured`: description of why redd was not measured; deemed unnecessary for final redd dataset
-   `date_measured`: chose to use `date` instead of this column
-   `pre_redd_depth`: numeric range from 0.13 to 2.24; deemed unnecessary
-   `redd_pit_depth`: numeric range from 0.15 to 2.41; deemed unnecessary
-   `redd_tail_depth`: numeric range from 0.076 to 2.13; deemed unnecessary
-   `velocity`: numeric range from 0.25 to 11.1; deemed unnecessary
-   `start_60`: numeric; deemed unnecessary
-   `end_60`: numeric; deemed unnecessary
-   `sec_60`: numeric; deemed unnecessary
-   `start_80`: numeric; deemed unnecessary
-   `end_80`: numeric; deemed unnecessary
-   `sec_80`: numeric; deemed unnecessary
-   `bomb_vel60`: numeric; deemed unnecessary
-   `bomb_vel80`: numeric; deemed unnecessary
-   `comments`: unique to dataset and deemed unnecessary
-   `run`: enumerated as `spring`, `fall`, `late-fall` though mostly `NA`; deemed unnecessary.
-   `observation_reach` -- Incomplete and matches `surveyed_reach`
-   `observation_date` -- Incomplete and matches `date`
-   `observation_age` -- Incomplete and matches `age`
-   `survey_observed`: TRUE/FALSE though mostly `NA`
-   `method`: mostly snorkel surveys, method deemed unnecessary for final dataset

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "clear_redd.csv"),
  overwrite = TRUE
)
# spring run Chinook Salmon from USFW - Ryan Shaffer 
clear_redd <- read_csv(here::here("data", "redd-carcass-holding","clear_redd.csv"), 
                       col_names = TRUE,
                       col_types = list("c", "n", "n", "n", "n", "n", 
                                        "c", "n", "c", "D", "c", "c", "n", "c", 
                                        "c", "c", "c", "c", "c", "l", "l", "l", 
                                        "D", "n", "n", "n", "n", "n", "n", "n", 
                                        "n", "n", "n", "n", "n", "c", "c", "c",
                                        "D", "n", "l", "n", "n"))  |> glimpse()

summary(clear_redd)
table(clear_redd$method, useNA = "ifany")
table(clear_redd$ucc_relate, useNA = "ifany")
table(clear_redd$picket_weir_relation, useNA = "ifany")
table(clear_redd$run, useNA = "ifany")


clean_clear_redd <- clear_redd  |>
  select(surveyed_reach, date, river_mile, 
         latitude, longitude, 
         measured, fish_on_redd,
         redd_length, redd_width, run,
         pre_redd_substrate_size, redd_substrate_size, 
         tail_substrate_size, 
         velocity)  |> # feet per second
  group_by(date, surveyed_reach)  |>
  mutate(redd_count = length(measured),
         stream = 'Clear Creek', 
         year = lubridate::year(date))  |>  
  rename(reach = surveyed_reach,
         redd_measured = measured,
         fish_guarding = fish_on_redd)  |>
  glimpse()

# load in aggregated clear creek data from other runs 
# this data was pulled from CVPIA-EDA repo (CVPIA-EDA/jim_data_requests/redd_monitoring_data/all_redd_data_for_clear_creek.csv)
# and saved to jpe google cloud for easy access
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/clear-creek/data/all_redd_data_for_clear_creek.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "all_redd_data_for_clear_creek.csv"),
  overwrite = TRUE
)
clear_all_runs <- read_csv(here::here("data", "redd-carcass-holding", "all_redd_data_for_clear_creek.csv"))

max_date <- max(clear_redd$date)
print(max_date)

spring_filter <- clear_all_runs  |> 
  filter(run == "spring")  |>
  filter(date > max_date)
clear_fall_and_spring <- clear_all_runs  |> 
  filter(run == "late fall")  |>
  bind_rows(spring_filter)  |>
  mutate(stream = "Clear Creek")  |>
  bind_rows(clean_clear_redd)  |>
  mutate(run = case_when(run == "late-fall" ~ "late fall",
                         TRUE ~ as.character(run)))

# replace name 
clean_clear_redd <- clear_fall_and_spring  |> glimpse()
```

### Mill Creek

-   changed `location` to `reach` to align with other datasets

#### Columns Removed

-   `starting_elevation_ft`: numeric and ranging from 1600 to 5200; deemed unnecessary for final dataset

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/mill-creek/data/mill_redd_survey.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "mill_redd.csv"),
  overwrite = TRUE
)
mill_redd <- read_csv(here::here("data", "redd-carcass-holding", "mill_redd.csv"))  |>  glimpse()

summary(mill_redd)

clean_mill_redd <- mill_redd  |>
  rename(reach = location)  |>
  select(-starting_elevation_ft)  |>
  mutate(stream = 'Mill Creek')  |>
  glimpse()
```

### Feather River

-   changed `location` to `reach` to align with other datasets
-   `velocity_m_per_s` converted to feet per second and named velocity

#### Columns Removed

-   `type`: enumerated as `Area`, `Point`, `Questionable Redds`, or `NA`; deemed unnecessary
-   `salmon_count`: integer ranging from 0 to 200; deemed unnecessary
-   `pot_depth_m`: numeric ranging from 0 to 50; deeemd unnecessary
-   `percent_fine_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_small_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_medium_substrate`: numeric ranging from 0 to 95; mostly `NA` and deemed unnecessary
-   `percent_large_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_boulder`: numeric ranging from 0 to 70; mostly `NA` and deemed unnecessary

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "feather_redd.csv"),
  overwrite = TRUE
)
feather_redd <- read_csv(here::here("data", "redd-carcass-holding", "feather_redd.csv"), 
                         col_names = TRUE, 
                         col_types = list("D", "c", "c", "n",
                                       "n", "n", "n", "n", "n",
                                       "n", "n", "n", "n", "n",
                                       "n", "n", "n"))  |> glimpse()

summary(feather_redd)
table(feather_redd$type, useNA = "ifany")

clean_feather_redd <- feather_redd  |>
  select(date, location, redd_count, 
         latitude, longitude, 
         redd_width_m, redd_length_m, velocity_m_per_s, depth_m)  |> 
  rename(reach = location,
         redd_width = redd_width_m, 
         redd_length = redd_length_m)  |>
  mutate(stream = "Feather River",
         year = lubridate::year(date),
         date = as.Date(date),
         velocity = velocity_m_per_s / 0.3048)  |> # convert to feet per second 
  select(-velocity_m_per_s)  |>
  glimpse()

```

### Yuba River

#### Columns Created:

-   `stream` set to Yuba River

#### Columns Removed:

None

```{r warning=FALSE}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/yuba-river/data/yuba_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "yuba_redd.csv"),
  overwrite = TRUE
)
yuba_river_raw <- read_csv(here::here("data", "redd-carcass-holding", "yuba_redd.csv")) |> 
  glimpse()

clean_yuba_redd <- yuba_river_raw |> 
  mutate(stream = "Yuba River") 
```

## Combine data:

This is the daily tabular redd data. Since Mill Creek is already aggregated to yearly counts, that stream is removed from this dataset.

```{r}
all_redd_data <- bind_rows(clean_battle_redd, 
                           clean_clear_redd,
                           clean_feather_redd,
                           clean_mill_redd,
                           clean_yuba_redd)  |> 
  select(-redd_count)  |>
  # remove mill creek
  filter(stream != "Mill Creek")  
```

## Substrate Size Standardization

Substrate sizes were given names and classes following the Wentworth Scale, created by W.C. Krumbein. This scale is what is commonly used in the United States. The substrate size provided for redd data was standardized using this scale. When the size range fell into two categories, they were rounded down.

| Class              | Size Range (mm) |
|--------------------|-----------------|
| Boulder            | \>256           |
| Cobble             | 64-256          |
| Very coarse gravel | 32-64           |
| Coarse gravel      | 16-32           |
| Medium gravel      | 8-16            |
| Fine gravel        | 4-8             |
| Very fine gravel   | 2-4             |
| Very coarse sand   | 1-2             |
| Coarse sand        | 0.5-1           |
| Medium sand        | 0.25-0.5        |
| Fine sand - clay   | \<0.25          |

```{r echo=TRUE, message=FALSE, warning=FALSE}
# standarized size ranges lookup
substrate_class = data.frame("standardized_size_range" = c("<0.25", 
                                                           "0.25-0.5",
                                                           "0.5-1",
                                                           "1-2",
                                                           "2-4", "4-8", '8-16', ">16"),
                              "redd_substrate_class" = c("fine sand",
                                                    "medium sand",
                                                    "coarse sand", "very coarse sand",
                                                    "very fine gravel", "fine gravel",
                                                    "medium gravel", 
                                                    "coarse gravel to boulder"))

unique(all_redd_data$redd_substrate_size)

redd_substrate_size_lookup <- 
  data.frame("redd_substrate_size" = unique(all_redd_data$redd_substrate_size),
                     "standardized_size_range" = c(NA, "1-2", "2-4", "1-2", "2-4",
                                                   "2-4", "1-2", "2-4", "0.25-0.5", 
                                                   "<0.25", "4-8", "4-8", ">16", "2-4", 
                                                   "2-4", "1-2", "2-4", "1-2", "2-4",
                                                   "0.25-0.5", "2-4", "2-4", "4-8", 
                                                   "4-8", "8-16"))  |> 
  left_join(substrate_class)
                    
tail_substrate_size_lookup <- all_redd_data  |> 
  ungroup()  |>
  select(tail_substrate_size)  |> 
  full_join(redd_substrate_size_lookup  |> 
              rename(tail_substrate_size = redd_substrate_size))  |> 
  distinct()  |>
  rename(tail_substrate_class = redd_substrate_class)

pre_redd_substrate_size_lookup <- all_redd_data  |> 
  ungroup()  |>
  select(pre_redd_substrate_size)  |> 
  full_join(redd_substrate_size_lookup  |> 
              rename(pre_redd_substrate_size = redd_substrate_size))  |> 
  distinct()  |>
  rename(pre_redd_substrate_class = redd_substrate_class) 

all_redd_data <- all_redd_data  |> 
  left_join(redd_substrate_size_lookup)  |>
  select(-standardized_size_range)  |>
  left_join(tail_substrate_size_lookup)  |>
  select(-standardized_size_range)  |>
  left_join(pre_redd_substrate_size_lookup)  |>
  select(-standardized_size_range, -pre_redd_substrate_size, 
         -redd_substrate_size, -tail_substrate_size)  

```

## Explore Variables {.tabset}

### stream

```{r}
table(all_redd_data$stream)
```

### date

```{r}
summary(all_redd_data$date)

all_redd_data  |>
  group_by(stream)  |>
  summarise(year_range = paste0(min(year, na.rm = TRUE), " : ", max(year, na.rm = TRUE)), 
            months = paste0(unique(sort(lubridate::month(date))), collapse = ", "))  |>
  ungroup()  |>
  distinct()  |> knitr::kable()

```

### velocity

```{r}
summary(all_redd_data$velocity)

ggplot() + 
  geom_point(data = all_redd_data, aes(x = date, y = velocity, color = reach)) + 
  facet_wrap(~ stream) +
  theme(legend.position = "none")
```

### fish_guarding

```{r}
table(all_redd_data$fish_guarding, useNA = "ifany")
```

### redd_measured

```{r}
table(all_redd_data$redd_measured, useNA = "ifany")

```

### redd_width and redd_length

```{r}

summary(all_redd_data$redd_width)

summary(all_redd_data$redd_length)

ggplot() + 
  geom_histogram(data = all_redd_data, aes(redd_width)) 
  
ggplot() +   
  geom_histogram(data = all_redd_data, aes(redd_length))

```

### pre_redd_substrate_class

```{r}
table(all_redd_data$pre_redd_substrate_class, useNA = "ifany")

```

### redd_substrate_class

```{r}
table(all_redd_data$redd_substrate_class, useNA = "ifany")

```

### tail_substrate_class

```{r}
table(all_redd_data$tail_substrate_class, useNA = "ifany")

```

## By Year Aggregation

All redd data was aggregated by sampled reach and date to get a count of redds. In order to not potentially double count, the maximum of a daily count is reported as the yearly redd count. Since Mill Creek is aggregated by year already, this dataset includes Mill Creek.

```{r}
redd_count_by_year <- bind_rows(clean_battle_redd, 
                           clean_clear_redd,
                           clean_feather_redd, 
                           clean_yuba_redd)  |> 
  group_by(reach, date)  |>
  mutate(sum_redd = n())  |>
  # how many redds were counted on each day? 
  group_by(year, stream)  |>
  mutate(max_yearly_redd_count = max(sum_redd))  |>  # maximum per year 
  bind_rows(clean_mill_redd  |> 
              group_by(year, reach)  |>
              mutate(max_yearly_redd_count = max(redd_count, na.rm = TRUE))) |>
  select(year, reach, stream, species, run, max_yearly_redd_count)  |>
  distinct()  


```

## Exploratory Plots

```{r}
# a plot to help understand sampling frequency across streams 
all_redd_data  |>
  group_by(date, stream)  |>
  summarise(n_reaches = length(unique(reach)))  |>
  ggplot() + 
  geom_point(aes(x = date, y = n_reaches)) + 
  ylab('number of samples per reach and date') + 
  facet_wrap(~ stream)

# Note: based on plot, count value could be double counting for Mill Creek
redd_count_by_year  |> 
  select(year, max_yearly_redd_count, stream)  |>
  distinct()  |>
ggplot() + 
  geom_col(aes(x = year, y = max_yearly_redd_count, fill = stream)) + 
  facet_wrap(~stream) +
  scale_fill_manual(values = color_pal) + 
  theme_minimal()
```

## Save Cleaned Data to Google Cloud

```{r}
knitr::kable(all_redd_data  |> head())

knitr::kable(redd_count_by_year  |> head())

```

```{r, eval=FALSE}

write_csv(all_redd_data, here::here("data","redd-carcass-holding","standard_all_redd_data.csv"))
write_csv(redd_count_by_year, here::here("data","redd-carcass-holding","standard_redd_data_by_year.csv"))
f <- function(input, output) write_csv(input, file = output)
gcs_upload(all_redd_data,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_daily_redd.csv")
gcs_upload(redd_count_by_year,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_annual_redd.csv")
```
