---
title: "Standardize Adult Redd Datasets"
author: "Maddee Rubenson (FlowWest)"
date: '2022-07-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
library(knitr)
color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

## Adult Redd Data Standardization

FlowWest received Adult Redd data from four monitoring programs:

-   Feather River
-   Battle Creek
-   Clear Creek
-   Mill Creek
-   Yuba River

Due to the format of the data received, two redd datasets were developed. The first dataset includes Feather, Battle, Yuba, and Clear Creek daily data and the second dataset includes Feather, Battle, Clear, Yuba, and Mill Creek and has a yearly maximum redd counts value.

## Standard format for Adult All Redd Data

Data dictionary for standard format for `all_redd_data`:

(B - Battle Creek, C - Clear Creek,\
F - Feather River, Y - Yuba River)

| column name              | tributary collects | definition                                                         |
|:-----------------|:-----------------|:-----------------------------------|
| date                     | B, C, F, Y         | Date in which redd data was collected                              |
| latitude                 | B, C, F, Y         | Latitude of redd location                                          |
| longitude                | B, C, F, Y         | Longitude of redd location                                         |
| reach                    | B, C, F            | Unique name identifier of reach sampled                            |
| river_mile               | B, C               | River mile associated with redd measured                           |
| redd_measured            | B, C               | Whether or not a redd was measured (T/F)                           |
| age                      | B, C               | integer ranging from 1 - 5; redd age                               |
| fish_guarding            | B, C               | Whether or not a fish is on redd when measured ??                  |
| redd_length              | B, C, F            | Length of measured redd (m)                                        |
| redd_width               | B, C, F            | Width of measured redd (m)                                         |
| run                      | B, C               | `spring`, `fall`, `late-fall`                                      |
| stream                   | B, C, F, Y         | Unique stream associated with redd data collected                  |
| redd_substrate_class     | B, C               | categorical; size class determined by millimeter size of substrate |
| pre_redd_substrate_class | B, C               | categorical; size class determined by millimeter size of substrate |
| tail_substrate_class     | B, C               | categorical; size class determined by millimeter size of substrate |
| year                     | B, C, F, Y         | Calendar year sampling occurred                                    |
| velocity                 | B, C, F, Y         | measured stream velocity (ft/s)                                    |
| depth_m                  | F, Y               | Depth of water (meters)                                            |

## Standard format for Adult Redd Data Aggregated by Year

This dataset includes `max_yearly_redd_count` which is the maximum number of redds counted on a single day per year per reach.

Data dictionary for standard format for `max_yearly_redd_count`:

| column name           | tributary collects | definition                                                          |
|:-----------------|:-----------------|:----------------------------------|
| date                  | B, C, F, Y         | Date in which redd data was collected                               |
| reach                 | B, C, F, M         | Unique name identifier of reach sampled                             |
| stream                | B, C, F, M         | Unique stream associated with redd data collected                   |
| year                  | B, C, F, M, Y      | Calendar year sampling occurred                                     |
| max_yearly_redd_count | B, C, F, M, Y      | maximum number of redds counted on a day in a unique reach per year |
| run                   | C                  | `spring`, `fall`, `late-fall`                                       |

## Read in data {.tabset}

Below we read in the adult redd data for each monitoring program and rename or select columns so that we can join all the monitoring datasets together in the section below.

### Battle Creek

#### Columns Removed

-   `why_not_measured`: enumerated; `NA`, `fish_on_redd`, `sub sample`, `too deep` Deemed unnecessary.
-   `flow_meter`: type of flow meter used; deemed unnecessary for final dataset
-   `start_number_flow_meter`: deemed unnecessary for final dataset
-   `end_number_flow_meter`: deemed unnecessary for final dataset
-   `flow_meter_time`: deemed unnecessary for final dataset
-   `start_number_flow_meter_80`: deemed unnecessary for final dataset
-   `end_number_flow_meter_80`: deemed unnecessary for final dataset
-   `flow_meter_time_80`: deemed unnecessary for final dataset
-   `redd_pit_depth`: deemed unnecessary for final dataset
-   `redd_tail_depth`: deemed unnecessary for final dataset
-   `comments`: deemed unnecessary for final dataset

columns created:

-   `stream`: used Battle Creek
-   `flow_fps` changed to `velocity`

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "battle_redd.csv"),
  overwrite = TRUE
)
# Redd data for Battle Creek
battle_redd <- read_csv(here::here("data", "redd-carcass-holding", "battle_redd.csv"), 
                        col_names = TRUE, 
                        col_types = list("D", "c", "n", "n", "c", "c", 
                                         "n", "c", "c", "c", "c", "c", 
                                         "c", "l", "l", "c", "n", "n", 
                                         "n", "n", "n", "c", "n", "n", 
                                         "n", "n", "n", "n", "n", "n", 
                                         "n", "n")) |> 
                          glimpse()

summary(battle_redd)

table(battle_redd$pre_redd_substrate_size, useNA = "ifany")
table(battle_redd$redd_substrate_size, useNA = "ifany")
table(battle_redd$tail_substrate_size, useNA = "ifany")
unique(battle_redd$why_not_measured)
unique(battle_redd$flow_meter)

# TODO do distinct redd id
clean_battle_redd <- battle_redd  |>
  select(date, latitude, longitude, 
         reach, river_mile, fish_guarding, redd_id,
         redd_measured, redd_width, redd_length, age,
         age_index, pre_redd_substrate_size, 
         redd_substrate_size, tail_substrate_size, flow_fps,
         survey_method, run)  |>
  mutate(stream = 'battle creek', 
         # adding redd count for only those redds surveyed and not aged, or aged once
         redd_count = ifelse(age_index %in% c(0, 1), 1, 0),
         year = lubridate::year(date))  |>
  rename(velocity = flow_fps)  |>
  glimpse()

```

### Clear Creek

FlowWest acquired Clear Creek redd data for late fall and spring runs through CVPIA.

O. mykiss data is included in this dataset.

#### Columns Removed

```{r}
# TODO confirm that we want to use the original clear data and not the CVPIA? CVPIA does not seem to have redd age

# load in aggregated clear creek data from other runs 
# this data was pulled from CVPIA-EDA repo (CVPIA-EDA/jim_data_requests/redd_monitoring_data/all_redd_data_for_clear_creek.csv)
# and saved to jpe google cloud for easy access

# data directly from Natasha
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/clear-creek/data/clear_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "clear_redd.csv"),
  overwrite = TRUE
) 

# Redd data for Clear Creek
clear_redd <- read_csv(here::here("data", "redd-carcass-holding", "clear_redd.csv"), 
                        col_names = TRUE, 
                        col_types = list("D", "c", "n", "n", "n", "n", 
                                         "n", "c", "n", "c", "c", "c",
                                         'c', "n", "n", "c", "c", "c",
                                         "c", "c", "l", "l", "c", "D",
                                         "n", "n", "n", "n", "n", "n",
                                         "n", "n", "n", "n", "n", "n",
                                         "c", "c", "n", "n")) |> 
                          glimpse()

summary(clear_redd)

table(clear_redd$pre_redd_substrate_size, useNA = "ifany")
table(clear_redd$redd_substrate_size, useNA = "ifany")
table(clear_redd$tail_substrate_size, useNA = "ifany")
unique(clear_redd$why_not_measured)
unique(clear_redd$flow_meter)

clean_clear_redd <- clear_redd  |>
  select(date, latitude, longitude, 
         reach = surveyed_reach, river_mile, fish_guarding = fish_on_redd, 
         redd_id, redd_measured = measured, redd_width, redd_length, age,
         age_index, pre_redd_substrate_size, 
         redd_substrate_size, tail_substrate_size, velocity,
         survey_method, run)  |>
  mutate(stream = 'clear creek', 
         # adding redd count for only those redds surveyed and not aged, or aged once
         redd_count = ifelse(age_index %in% c(0, 1), 1, 0),
         year = lubridate::year(date))  |>
  glimpse()

```

### Mill Creek

-   changed `location` to `reach` to align with other datasets

#### Columns Removed

None

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/mill-creek/data/mill_redd_survey.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "mill_redd.csv"),
  overwrite = TRUE
)
mill_redd <- read_csv(here::here("data", "redd-carcass-holding", "mill_redd.csv"))  |>  glimpse()

summary(mill_redd)

clean_mill_redd <- mill_redd  |>
  rename(reach = location)  |>
  mutate(stream = 'mill creek')  |>
  glimpse()
```

### Feather River

-   changed `location` to `reach` to align with other datasets
-   `velocity_m_per_s` converted to feet per second and named velocity

#### Columns Removed

-   `type`: enumerated as `Area`, `Point`, `Questionable Redds`, or `NA`; deemed unnecessary
-   `salmon_count`: integer ranging from 0 to 200; deemed unnecessary
-   `pot_depth_m`: numeric ranging from 0 to 50; deeemd unnecessary
-   `percent_fine_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_small_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_medium_substrate`: numeric ranging from 0 to 95; mostly `NA` and deemed unnecessary
-   `percent_large_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_boulder`: numeric ranging from 0 to 70; mostly `NA` and deemed unnecessary

```{r}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "feather_redd.csv"),
  overwrite = TRUE
)
feather_redd <- read_csv(here::here("data", "redd-carcass-holding", "feather_redd.csv"), 
                         col_names = TRUE, 
                         col_types = list("D", "c", "c", "n",
                                       "n", "n", "n", "n", "n",
                                       "n", "n", "n", "n", "n",
                                       "n", "n", "n"))  |> glimpse()

summary(feather_redd)
table(feather_redd$type, useNA = "ifany")

clean_feather_redd <- feather_redd  |>
  select(date, location, redd_count, 
         latitude, longitude, 
         redd_width_m, redd_length_m, velocity_m_per_s, depth_m)  |> 
  rename(reach = location,
         redd_width = redd_width_m, 
         redd_length = redd_length_m)  |>
  mutate(stream = "feather river",
         year = lubridate::year(date),
         date = as.Date(date),
         velocity = velocity_m_per_s / 0.3048)  |> # convert to feet per second 
  select(-velocity_m_per_s)  |>
  glimpse()

```

### Yuba River

#### Columns Created:

-   `stream` set to Yuba River

#### Columns Removed:

None

```{r warning=FALSE}
gcs_get_object(
  object_name = "adult-holding-redd-and-carcass-surveys/yuba-river/data/yuba_redd.csv",
  bucket = gcs_get_global_bucket(),
  saveToDisk = here::here("data", "redd-carcass-holding", "yuba_redd.csv"),
  overwrite = TRUE
)
yuba_river_raw <- read_csv(here::here("data", "redd-carcass-holding", "yuba_redd.csv")) |> 
  glimpse()

clean_yuba_redd <- yuba_river_raw |> 
  mutate(stream = "yuba river",
         redd_count = 1) # each row is a redd observation 
```

## Combine data:

This is the daily tabular redd data. Since Mill Creek is already aggregated to yearly counts, that stream is removed from this dataset.

```{r}
all_redd_data <- bind_rows(clean_battle_redd, 
                           clean_clear_redd,
                           clean_feather_redd,
                           clean_mill_redd,
                           clean_yuba_redd)  |> 
  select(-redd_count)  |>
  # remove mill creek
  filter(stream != "mill creek")  |> 
  mutate(species = case_when(species %in% c("chinook","CHN") ~ "chinook",
                             species %in% c("UNK", "Unknown", "Salmonid") ~ "unknown",
                             species == "SH" ~ "steelhead",
                             is.na(species) ~ "not recorded"),
         run = ifelse(is.na(run), "not recorded", run))
```

## Substrate Size Standardization

Substrate sizes were given names and classes following the Wentworth Scale, created by W.C. Krumbein. This scale is what is commonly used in the United States. The substrate size provided for redd data was standardized using this scale. When the size range fell into two categories, they were rounded down.

| Class              | Size Range (mm) |
|--------------------|-----------------|
| Boulder            | \>256           |
| Cobble             | 64-256          |
| Very coarse gravel | 32-64           |
| Coarse gravel      | 16-32           |
| Medium gravel      | 8-16            |
| Fine gravel        | 4-8             |
| Very fine gravel   | 2-4             |
| Very coarse sand   | 1-2             |
| Coarse sand        | 0.5-1           |
| Medium sand        | 0.25-0.5        |
| Fine sand - clay   | \<0.25          |

```{r echo=TRUE, message=FALSE, warning=FALSE}
# standarized size ranges lookup
substrate_class = data.frame("standardized_size_range" = c("<0.25", 
                                                           "0.25-0.5",
                                                           "0.5-1",
                                                           "1-2",
                                                           "2-4", 
                                                           "4-8", 
                                                           '8-16', 
                                                           ">16"),
                              "redd_substrate_class" = c("fine sand",
                                                    "medium sand",
                                                    "coarse sand", "very coarse sand",
                                                    "very fine gravel", "fine gravel",
                                                    "medium gravel", 
                                                    "coarse gravel to boulder"))

unique(all_redd_data$redd_substrate_size)

redd_substrate_size_lookup <- 
  data.frame("redd_substrate_size" = unique(all_redd_data$redd_substrate_size),
             "standardized_size_range" = c(NA, "1-2", "1-2", "2-4", "2-4", 
                                           "0.5-1", "2-4", "2-4", ">16", "4-6",
                                           "2-4", "1-2", "4-8", "0.5-1", "0.5-1",
                                           "4-8", "8-16")) |> 
  left_join(substrate_class)
                    
tail_substrate_size_lookup <- all_redd_data  |> 
  ungroup()  |>
  select(tail_substrate_size)  |> 
  full_join(redd_substrate_size_lookup  |> 
              rename(tail_substrate_size = redd_substrate_size))  |> 
  distinct()  |>
  rename(tail_substrate_class = redd_substrate_class)

pre_redd_substrate_size_lookup <- all_redd_data  |> 
  ungroup()  |>
  select(pre_redd_substrate_size)  |> 
  full_join(redd_substrate_size_lookup  |> 
              rename(pre_redd_substrate_size = redd_substrate_size))  |> 
  distinct()  |>
  rename(pre_redd_substrate_class = redd_substrate_class) 

all_redd_data <- all_redd_data  |> 
  left_join(redd_substrate_size_lookup)  |>
  select(-standardized_size_range)  |>
  left_join(tail_substrate_size_lookup)  |>
  select(-standardized_size_range)  |>
  left_join(pre_redd_substrate_size_lookup)  |>
  select(-standardized_size_range, -pre_redd_substrate_size, 
         -redd_substrate_size, -tail_substrate_size)  

```

## Explore Variables {.tabset}

### stream

```{r}
table(all_redd_data$stream)
```

### date

```{r}
summary(all_redd_data$date)

all_redd_data  |>
  group_by(stream)  |>
  summarise(year_range = paste0(min(year, na.rm = TRUE), " : ", max(year, na.rm = TRUE)), 
            months = paste0(unique(sort(lubridate::month(date))), collapse = ", "))  |>
  ungroup()  |>
  distinct()  |> knitr::kable()

```

### velocity

```{r}
summary(all_redd_data$velocity)

ggplot() + 
  geom_point(data = all_redd_data, aes(x = date, y = velocity, color = reach)) + 
  facet_wrap(~ stream) +
  theme(legend.position = "none")
```

### fish_guarding

```{r}
table(all_redd_data$fish_guarding, useNA = "ifany")
```

### redd_measured

```{r}
table(all_redd_data$redd_measured, useNA = "ifany")

```

### redd_width and redd_length

```{r}

summary(all_redd_data$redd_width)

summary(all_redd_data$redd_length)

ggplot() + 
  geom_histogram(data = all_redd_data, aes(redd_width)) 
  
ggplot() +   
  geom_histogram(data = all_redd_data, aes(redd_length))

```

### pre_redd_substrate_class

```{r}
table(all_redd_data$pre_redd_substrate_class, useNA = "ifany")

```

### redd_substrate_class

```{r}
table(all_redd_data$redd_substrate_class, useNA = "ifany")

```

### tail_substrate_class

```{r}
table(all_redd_data$tail_substrate_class, useNA = "ifany")

```

## By Year Aggregation

Redd data for Feather River and Yuba River was aggregated by sampled reach and date to get a count of redds. In order to not potentially double count, the maximum of a daily count is reported as the yearly redd count. 

For Battle and Clear Creeks, `redd_count` is 1 for every row where the redd was aged either `0` times or `1` time (recorded in `age_index` column). These were then summed by reach and year.

Since Mill Creek is aggregated by year already, this dataset includes Mill Creek.

```{r}
redd_count_by_year <- bind_rows(clean_feather_redd, 
                                clean_yuba_redd)  |> 
  # find the redd count by date/stream/reach/species/run
  group_by(reach, date, stream, year, species)  |>
  summarize(redd_count = sum(redd_count))  |>
  ungroup() |>
  # since some redds may be double counted, summarize yearly count by date
  # as we learn more about methods we can summarize using a better method
  group_by(year, stream, reach, species)  |>
  summarize(max_yearly_redd_count = max(redd_count, na.rm = T))  |>  # maximum per year 
  bind_rows(clean_mill_redd  |> 
              rename(max_yearly_redd_count = redd_count) |> 
              select(-starting_elevation_ft)) |> 
  bind_rows(clean_battle_redd |> 
              group_by(reach, date, stream, year, run) |> 
              summarise(annual_redd_count = sum(redd_count)) |> 
              ungroup(), 
            clean_clear_redd |> 
              group_by(reach, date, stream, year, run) |> 
              summarise(annual_redd_count = sum(redd_count)) |> 
              ungroup()) |> 
  mutate(annual_redd_count = ifelse(stream %in% c("mill creek", "yuba river", "feather river"),
                                    max_yearly_redd_count, annual_redd_count),
         annual_summarization_method = ifelse(stream %in% c("mill creek", "yuba river", "feather river"), "maximum yearly redd count", "summed by redd id")) |> 
  mutate(species = case_when(species %in% c("chinook","CHN") ~ "chinook",
                             species %in% c("UNK", "Unknown", "Salmonid") ~ "unknown",
                             species == "SH" ~ "steelhead",
                             is.na(species) ~ "not recorded"),
         run = ifelse(is.na(run), "not recorded", run))

```

## Exploratory Plots

```{r}
# a plot to help understand sampling frequency across streams 
all_redd_data  |>
  group_by(date, stream)  |>
  summarise(n_reaches = length(unique(reach)))  |>
  ggplot() + 
  geom_point(aes(x = date, y = n_reaches)) + 
  ylab('number of samples per reach and date') + 
  facet_wrap(~ stream)

# Note: based on plot, count value could be double counting for Mill Creek
# I spent some time trying to figure out why feather and mill are so much higher
# feather does survey a lot of reaches. this may be an artificat of the methodology
redd_count_by_year  |> 
  group_by(year, stream)  |> # add across reaches
  summarize(max_yearly_redd_count = sum(max_yearly_redd_count, na.rm = T)) |> 
ggplot() + 
  geom_col(aes(x = year, y = max_yearly_redd_count, fill = stream)) + 
  facet_wrap(~stream) +
  scale_fill_manual(values = color_pal) + 
  theme_minimal()
```

## Save Cleaned Data to Google Cloud

```{r}
knitr::kable(all_redd_data  |> head())

knitr::kable(redd_count_by_year  |> head())

```

```{r, eval=FALSE}

write_csv(all_redd_data, here::here("data","redd-carcass-holding","standard_all_redd_data.csv"))
write_csv(redd_count_by_year, here::here("data","redd-carcass-holding","standard_redd_data_by_year.csv"))
f <- function(input, output) write_csv(input, file = output)
gcs_upload(all_redd_data,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_daily_redd.csv",
           predefinedAcl = "bucketLevel")
gcs_upload(redd_count_by_year,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_annual_redd.csv",
           predefinedAcl = "bucketLevel")
```
