---
title: "Standardize Mark Recapture Data"
author: "Erin Cain"
date: "4/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
```

## Mark Recapture Data Standardization 

FlowWest received mark recapture data from 4 monitoring programs: 

* Battle Creek 
* Clear Creek 
* Feather River
* Knights Landing 

Tisdale has conducted efficiency tests but these are in paper format. 
Mill and Deer did not historically conduct efficiency tests but will do so moving forward. 

TODO Butte & Yuba ? do we have any idea here 

## Standard format for Mark Recapture Data

Data dictionary for standard format: 

| Column Name | Definition | 
| :------------------------------------ | :---------------------------------------------------------------- | 
| tributary | Which Spring Run JPE is the data from |
| date_released | date that marked fish are released |
| time_released | time that marked fish are released |
| location_released | subsite on tributary where fish are released |
| number_released | Count of fish released |
| date_recaptured | Date that fish were recaptured |
| number_recaptured  | Count of fish recaptured in RST on a specific recapture date |
| efficiency | Efficiency of mark recapture trial, calculated by (number_recaps + 1)/(number_released + 1) |
| median_fork_length_released | median fork length of group of fish released|
| median_fork_length_recaptured | median fork length of group of fish recaptured |



All released and recaptured fish are chinook salmon. 

## Read in data {.tabset} 

### Battle Creek  
```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/battle-creek/data/battle_mark_reacpture.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/battle_mark_recapture.csv",
#                overwrite = TRUE)

battle_mark_recap <- read_csv("../../data/rst/battle_mark_recapture.csv") %>% 
  mutate(tributary = "Battle Creek") %>%
  rename(number_released = no_released, number_recaptured = recaps,
         median_fork_length_recaptured = mark_med_fork_length_mm, 
         median_fork_length_released = recap_med_fork_length_mm) %>%
  select(-no_marked, -mortality, -mean_flow_day_of_rel, -mean_temp_day_of_rel, -number_recaptured) %>% 
  pivot_longer(cols = caught_day_1:caught_day_5, names_to = "recaptured_date", values_to = "number_recaptured") %>%
  mutate(recaptured_date = case_when(recaptured_date == "caught_day_1" ~ release_date + 1, 
                                    recaptured_date == "caught_day_2" ~ release_date + 2,
                                    recaptured_date == "caught_day_3" ~ release_date + 3,
                                    recaptured_date == "caught_day_4" ~ release_date + 4,
                                    recaptured_date == "caught_day_5" ~ release_date + 5,)) %>%
  glimpse
```

### Clear Creek 
```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/clear-creek/data/clear_mark_reacpture.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/clear_mark_recapture.csv",
#                overwrite = TRUE)

clear_mark_recap <- read_csv("../../data/rst/clear_mark_recapture.csv") %>% 
  mutate(tributary = "Clear Creek") %>% 
  rename(number_released = no_released, number_recaptured = recaps,
         median_fork_length_recaptured = mark_med_fork_length_mm, 
         median_fork_length_released = recap_med_fork_length_mm) %>%
  select(-no_marked, -mean_flow_day_of_rel, -mean_temp_day_of_rel, -number_recaptured) %>% 
  pivot_longer(cols = caught_day_1:caught_day_5, names_to = "recaptured_date", values_to = "number_recaptured") %>%
  mutate(recaptured_date = case_when(recaptured_date == "caught_day_1" ~ release_date + 1, 
                                    recaptured_date == "caught_day_2" ~ release_date + 2,
                                    recaptured_date == "caught_day_3" ~ release_date + 3,
                                    recaptured_date == "caught_day_4" ~ release_date + 4,
                                    recaptured_date == "caught_day_5" ~ release_date + 5,)) %>%
  glimpse
```

### Feather River
```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/feather-river/data/feather_river_mark_recapture_data.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/feather_mark_recapture.csv",
#                overwrite = TRUE)

feather_mark_recap <- read_csv("../../data/rst/feather_mark_recapture.csv") %>% 
  mutate(tributary = "Feather River") %>% 
  rename(number_released = nReleased) %>% 
  select(-efficiency) %>%
  glimpse

```

### Knights Landing
```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/lower-sac-river/data/knights-landing/knights_landing_mark_recapture_data.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/knights_landing_mark_recapture.csv",
#                overwrite = TRUE)

knights_landing_mark_recap <- read_csv("../../data/rst/knights_landing_mark_recapture.csv") %>% 
  mutate(tributary = "Lower Sac - Knights Landing") %>% 
  rename(number_released = nReleased) %>% 
  select(-efficiency) %>%
  glimpse
```

## Combine data: 
```{r}
combined_mark_recapture <- bind_rows(battle_mark_recap, clear_mark_recap, feather_mark_recap, knights_landing_mark_recap) %>% 
  relocate(tributary, release_site_id = releaseSiteID, release_id = releaseID, release_date, release_time, number_released, 
           recaptured_date, number_recaptured, median_fork_length_released, median_fork_length_recaptured, day_or_night_release, 
           days_held_post_mark, flow_at_release = flow_release, temperature_at_release = release_temp, turbidity_at_release = release_turbidity, 
           origin, cone_status) %>% glimpse
```

## Explore Varibles{.tabset} 

### tributary 

```{r}
table(combined_mark_recapture$tributary)
```

### release_site_id 

Site ids can be joined to additional locational data using RST_trap_locations.rds
```{r}
table(combined_mark_recapture$release_site_id)
```

Battle and Clear Creek do not provide a site id 
```{r}
filter(combined_mark_recapture, tributary %in% c("Battle Creek", "Clear Creek")) %>% pull(release_site_id) %>% unique()
```

### release_id 

Release id are unique identifiers for an efficiency trial on a specific tributary. There are 843 unique release IDs  
```{r}
length(unique(combined_mark_recapture$release_id))
```

Battle and Clear Creek do not provide a site id 
```{r}
filter(combined_mark_recapture, tributary %in% c("Battle Creek", "Clear Creek")) %>% pull(release_id) %>% unique()
```

### release_date     


```{r}
combined_mark_recapture %>% ggplot() + 
  geom_point(aes(x = release_date, y = tributary, color = tributary), alpha = .1) + 
  theme_minimal() + 
  theme(legend.position = "none")
```


### release_time      

#TODO see if we can get feather and knights landing release time
```{r}
combined_mark_recapture %>% ggplot() + 
  geom_point(aes(x = release_time, y = tributary, color = tributary), alpha = .1) + 
  theme_minimal() + 
  scale_x_time() +
  theme(legend.position = "none")
```


### number_released                

Most release trials release less than 2000 fish. 
```{r}
combined_mark_recapture %>% ggplot() + 
  geom_histogram(aes(x = number_released, fill = tributary), alpha = .5) + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```
### recaptured_date  

```{r}
combined_mark_recapture %>% ggplot() + 
  geom_point(aes(x = recaptured_date, y = tributary, color = tributary), alpha = .1) + 
  theme_minimal() + 
  theme(legend.position = "none")
```

### number_recaptured     

Most trials only recapture less than 200 fish. 

```{r}
combined_mark_recapture %>% 
  filter(number_recaptured < 250) %>% 
  ggplot() + 
  geom_histogram(aes(x = number_recaptured, fill = tributary), alpha = .5, binwidth = 10) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

There are 10 values where number_recaptured is greater than 250
```{r}
combined_mark_recapture %>% 
  filter(number_recaptured > 250)
```
### median_fork_length_released

All fork length values are summarized to just show medium fork length. Feather data does not
provide data on the fork length of released fish. Knights Landing data are summarized to show 
median fork length per recapture day. Battle and 
Clear Creek only show median fork length per efficiency trial.

### median_fork_length_recaptured

All fork length values are summarized to just show medium fork length. Feather and 
Knights Landing are summarized to show median fork length per recapture day. Battle and 
Clear Creek only show median fork length per efficiency trial. 


### day_or_night_release           
Only battle and clear provide timing of release (day or night)
```{r}
table(combined_mark_recapture$day_or_night_release)
```

```{r}
filter(combined_mark_recapture, tributary %in% c("Lower Sac - Knights Landing", "Feather River")) %>% pull(day_or_night_release) %>% unique()
```

### days_held_post_mark    
Only battle and clear provide the days held post mark (pre release), most fish are held one day at most before release
```{r}
combined_mark_recapture %>% 
  ggplot() + 
  geom_histogram(aes(x = days_held_post_mark, fill = tributary), alpha = .5, binwidth = 1) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

### flow_at_release                      
Only battle and clear provide flow at release
```{r}
combined_mark_recapture %>% 
  ggplot() + 
  geom_histogram(aes(x = flow_at_release, fill = tributary), alpha = .5, binwidth = 25) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

### temperature_at_release  
Only battle and clear provide temperature at release, temperature values mostly fall between 40 and 60 degrees C
```{r}
combined_mark_recapture %>% 
  ggplot() + 
  geom_histogram(aes(x = temperature_at_release, fill = tributary), alpha = .5, binwidth = 2.5) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```
### turbidity_at_release
Only battle and clear provide turbidity measures, most turbidity measures fall below 25 NTU 
```{r}
combined_mark_recapture %>% 
  ggplot() + 
  geom_histogram(aes(x = turbidity_at_release, fill = tributary), alpha = .5, binwidth = 2) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```
### origin   
Only battle and clear provide
```{r}
table(combined_mark_recapture$origin)
```


```{r}
filter(combined_mark_recapture, tributary %in% c("Lower Sac - Knights Landing", "Feather River")) %>% pull(origin) %>% unique()
```

### cone_status
Only battle and clear provide a cone_status
```{r}
table(combined_mark_recapture$cone_status)
```

```{r}
filter(combined_mark_recapture, tributary %in% c("Lower Sac - Knights Landing", "Feather River")) %>% pull(cone_status) %>% unique()
```

## Summarize data by mark recapture data {.tabset}

To calculate the efficiency of each mark recapture trial the data is now summarized by trial and total fish recaptured per trial is calculated below. 
```{r}
summarized_mark_recapture <- combined_mark_recapture %>% 
  group_by(tributary, release_id, release_date, number_released) %>%
  summarize(total_fish_recaptured = sum(number_recaptured, na.rm = T), 
            median_fork_length_released = median(median_fork_length_released, na.rm = T),
            median_fork_length_recaptured = median(median_fork_length_recaptured, na.rm = T),
            flow_at_release = mean(flow_at_release, na.rm = T),
            temperature_at_release = mean(temperature_at_release, na.rm = T),
            turbidity_at_release = mean(turbidity_at_release, na.rm = T),
            efficiency = (number_recaptured + 1)/(number_released + 1)) %>% glimpse

```

### Efficiency Measures 
```{r}
summarized_mark_recapture %>% 
  filter(efficiency < 1) %>% 
  ggplot() + 
  geom_histogram(aes(x = efficiency, fill = tributary), alpha = .5) + 
  facet_wrap(~tributary, scales ="free") +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

### Flow vs Efficiency 
No points for knights landing and feather since we do not have flow

```{r}
summarized_mark_recapture %>% 
  filter(efficiency < 1, tributary %in% c("Battle Creek", "Clear Creek")) %>%  # one efficiency outlier (TODO check and remove)
  ggplot() + 
  geom_point(aes(x = flow_at_release, y = efficiency, color = tributary), alpha = .5) + 
  facet_wrap(~tributary) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

### Temperature vs Efficiency 

No points for knights landing and feather since we do not have temp

```{r}
summarized_mark_recapture %>% 
  filter(efficiency < 1, tributary %in% c("Battle Creek", "Clear Creek")) %>% # one efficiency outlier (TODO check and remove)
  ggplot() + 
  geom_point(aes(x = temperature_at_release, y = efficiency, color = tributary), alpha = .5) + 
  facet_wrap(~tributary) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```


### Turbidity vs Efficiency 
No points for knights landing and feather since we do not have turbidity

```{r}
summarized_mark_recapture %>% 
  filter(efficiency < 1, tributary %in% c("Battle Creek", "Clear Creek")) %>%
  ggplot() + 
  geom_point(aes(x = turbidity_at_release, y = efficiency, color = tributary), alpha = .5) + 
  facet_wrap(~tributary) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```



## Save unsumarized data to google cloud
```{r}

```