---
title: "Standardize Adult Redd Datasets"
author: "Maddee Rubenson (FlowWest)"
date: '2022-07-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
library(knitr)
color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")
```

## Adult Redd Data Standardization

FlowWest received Adult Redd data from four monitoring programs:

## Standard format for Adult Redd Data

Data dictionary for standard format:

(B - Battle Creek, C - Clear Creek,\
F - Feather River, M - Mill Creek)

| column name             | tributary collects | definition                                                 |
|:-----------------------|:-----------------------|:-----------------------|
| date                    | B, C, F            | Date in which redd data was collected                      |
| latitude                | B, C, F, M         | Latitude of redd location                                  |
| longitude               | B, C, F, M         | Longitude of redd location                                 |
| reach                   | B, C, F, M         | Unique name identifer of reach sampled                     |
| river_mile              | B, C               | River mile associated with redd measured                   |
| fish_guarding           | B                  | Whether or not a fish is guarding redd when measured (T/F) |
| redd_measured           | B, C               | Whether or not a redd was measured (T/F)                   |
| age                     | C                  | ??                                                         |
| fish_on_redd            | C                  | Whether or not a fish is on redd when measured ??          |
| redd_count              | B, C, F, M         | Amount of redds counted in given reach on given date       |
| redd_length             | B, C, F            | Length of measured redd (m)                                |
| redd_width              | B, C, F            | Width of measured redd (m)                                 |
| run ??                  | C                  |                                                            |
| stream                  | B, C, F, M         | Unique stream associated with redd data collected          |
| redd_substrate_size     | B, C               |                                                            |
| pre_redd_substrate_size | B, C               |                                                            |
| tail_substrate_size     | B, C               |                                                            |

## Read in data {.tabset}

Below we read in the adult redd data for each monitoring program and rename or select columns so that we can join all the monitoring datasets together in the section below.

### Battle Creek

#### Columns Removed

-   `why_not_measured`: all NA
-   `flow_meter`: type of flow meter used; deemed unnecessary for final dataset
-   `flow_fps`: measured flow (feet per second); deemed unnecessary for final dataset
-   `start_number_flow_meter`: deemed unnecessary for final dataset
-   `end_number_flow_meter`: deemed unnecessary for final dataset
-   `flow_meter_time`: deemed unnecessary for final dataset
-   `start_number_flow_meter_80`: deemed unnecessary for final dataset
-   `end_number_flow_meter_80`: deemed unnecessary for final dataset
-   `flow_meter_time_80`: deemed unnecessary for final dataset
-   `redd_pit_depth`: deemed unnecessary for final dataset
-   `redd_tail_depth`: deemed unnecessary for final dataset
-   `comments`: deemed unecessary for final dataset

columns created:

-   `redd_count`: based on the count of `redd_measured` for alike days and reach identifier
-   `stream`: used Battle Creek

```{r}
# Adult upstream passage data for Battle Creek
battle_redd <- read_csv("../../data/redd-carcass-holding/battle_redd.csv", 
                        col_names = TRUE, 
                        col_types = list("n", "n", "D", "c", "n", "c", 
                                         "c", "c", "l", "l", "l", "n", 
                                         "n", "n", "n", "n", "c", "n", 
                                         "n", "n", "n", "n", "n", "n", 
                                         "c")) %>% glimpse

summary(battle_redd)

table(battle_redd$pre_redd_substrate_size, useNA = "ifany")
table(battle_redd$redd_substrate_size, useNA = "ifany")
table(battle_redd$tail_substrate_size, useNA = "ifany")
unique(battle_redd$why_not_measured)
unique(battle_redd$flow_meter)

clean_battle_redd <- battle_redd %>%
  select(date, latitude, longitude, 
         reach, river_mile, fish_guarding, 
         redd_measured, redd_width, redd_length, 
         pre_redd_substrate_size, redd_substrate_size, 
         tail_substrate_size) %>%
  group_by(date, reach) %>%
  mutate(redd_count = length(redd_measured),
         stream = 'Battle Creek', 
         year = lubridate::year(date)) %>% 
  glimpse


```

### Clear Creek

#### Columns Removed

-   `survey`: integer ranging from 1 to 12; deemed unnecessary for final dataset
-   `X1000ftbreak`: deemed unnecessary for final dataset
-   `ucc_relate`: enumerated value of `above` or `below`; deemed unnessary for final dataset
-   `picket_weir_location`: numeric value ranging from 7.4 to 8.2; deemed unnessary for final dataset
-   `picket_weir_relation`: this could be used to determine the run type, if needed; enumerated as `above` or `below`
-   `redd_id`: unique identifier to each redd; deemed unessary for final redd dataset
-   `age`: integer ranging form 1 to 5; deemed unessary for final redd dataset
-   `gravel`: enumerated value of `combination`, `injection`, or `native`; deemed unessary for final redd dataset
-   `fish_on_redd`: ??? TRUE/FALSE
-   `why_not_measured`: all NAs
-   `date_measured`: chose to use `date` instead of this column 
-   `pre_redd_depth`: numeric range from 0.13 to 2.24; deemed unnecessary 
-   `redd_pit_depth`: numeric range from 0.15 to 2.41; deemed unnecessary 
-   `redd_tail_depth`: numeric range from 0.076 to 2.13; deemed unnecessary 
-   `velocity`: numeric range from 0.25 to 11.1; deemed unnecessary 
-   `start_60`: numeric; deemed unnecessary 
-   `end_60`: numeric; deemed unnecessary 
-   `sec_60`: numeric; deemed unnecessary 
-   `start_80`: numeric; deemed unnecessary 
-   `end_80`: numeric; deemed unnecessary 
-   `sec_80`: numeric; deemed unnecessary 
-   `bomb_vel60`: numeric; deemed unnecessary 
-   `bomb_vel80`: numeric; deemed unnecessary 
-   `comments`: unique to dataset and deemed unnecessary
-   `run`: enumerated as `spring`, `fall`, `late-fall` though mostly `NA`; deemed unnecessary. 
-   `observation_reach` -- Incomplete and matches `surveyed_reach`
-   `observation_date` -- Incomplete and matches `date`
-   `observation_age` -- Incomplete and matches `age`
-   `survey_observed`: TRUE/FALSE though mostly `NA`
-   `method`: mostly snorkel surveys, method deemed unnecessary for final dataset 

```{r}
# spring run Chinook Salmon from USFW - Ryan Shaffer 
clear_redd <- read_csv("../../data/redd-carcass-holding/clear_redd.csv", 
                       col_names = TRUE,
                       col_types = list("c", "n", "n", "n", "n", "n", 
                                        "c", "n", "c", "D", "c", "c", "n", "c", 
                                        "c", "c", "c", "c", "c", "l", "l", "l", 
                                        "D", "n", "n", "n", "n", "n", "n", "n", 
                                        "n", "n", "n", "n", "n", "c", "c", "c",
                                        "D", "n", "l", "n", "n")) %>% glimpse()

summary(clear_redd)
table(clear_redd$method, useNA = "ifany")
table(clear_redd$ucc_relate, useNA = "ifany")
table(clear_redd$picket_weir_relation, useNA = "ifany")
table(clear_redd$run, useNA = "ifany")


clean_clear_redd <- clear_redd %>%
  select(surveyed_reach, date, river_mile, 
         latitude, longitude, 
         measured, fish_on_redd,
         redd_length, redd_width, 
         pre_redd_substrate_size, redd_substrate_size, 
         tail_substrate_size) %>%
  group_by(date, surveyed_reach) %>%
  mutate(redd_count = length(measured),
         stream = 'Clear Creek', 
         year = lubridate::year(date)) %>%  
  rename(reach = surveyed_reach,
         redd_measured = measured) %>%
  glimpse()


```

### Mill Creek

-   changed `location` to `reach` to align with other datasets

#### Columns Removed

-   `starting_elevation_ft`: numeric and ranging from 1600 to 5200; deemed unnecessary for final dataset

```{r}

mill_redd <- read_csv("../../data/redd-carcass-holding/mill_redd.csv") %>% glimpse()

summary(mill_redd)

clean_mill_redd <- mill_redd %>%
  rename(reach = location) %>%
  select(-starting_elevation_ft) %>%
  mutate(stream = 'Mill Creek') %>%
  glimpse
```

### Feather River

-   changed `location` to `reach` to align with other datasets

#### Columns Removed

-   `type`: enumerated as `Area`, `Point`, `Questionable Redds`, or `NA`; deemed unnecessary
-   `salmon_count`: integer ranging from 0 to 200; deemed unnecessary
-   `pot_depth_m`: numeric ranging from 0 to 50; deeemd unnecessary
-   `velocity_m_per_s`: numeric ranging from -0.04 to 3.5; mostly `NA` and deemed unnecessary
-   `percent_fine_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_small_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_medium_substrate`: numeric ranging from 0 to 95; mostly `NA` and deemed unnecessary
-   `percent_large_substrate`: numeric ranging from 0 to 90; mostly `NA` and deemed unnecessary
-   `percent_boulder`: numeric ranging from 0 to 70; mostly `NA` and deemed unnecessary
-   `depth_m`: numeric ranging from 0.02 to 1.5; mostly NA and deemed unnecessary

```{r}
feather_redd <- read_csv("../../data/redd-carcass-holding/feather_redd.csv", 
                         col_names = TRUE, 
                         col_types = list("D", "c", "c", "n",
                                       "n", "n", "n", "n", "n",
                                       "n", "n", "n", "n", "n",
                                       "n", "n", "n")) %>% glimpse()

summary(feather_redd)
table(feather_redd$type, useNA = "ifany")

clean_feather_redd <- feather_redd %>%
  select(date, location, redd_count, 
         latitude, longitude, 
         redd_width_m, redd_length_m) %>% 
  rename(reach = location,
         redd_width = redd_width_m, 
         redd_length = redd_length_m) %>%
  mutate(stream = "Feather River",
         year = lubridate::year(date),
         date = as.Date(date)) %>%
  glimpse

```

## Combine data:

TODO: is fish_guarding and fish_on_redd the same metric? 

```{r}
all_redd_data <- bind_rows(clean_battle_redd, 
                           clean_clear_redd,
                           clean_feather_redd,
                           clean_mill_redd) %>% glimpse
```

## Explore Variables {.tabset}

### stream

```{r}
table(all_redd_data$stream)
```

### date

```{r}
summary(all_redd_data$date)

all_redd_data %>%
  group_by(stream) %>%
  summarise(year_range = paste0(min(year, na.rm = TRUE), " : ", max(year, na.rm = TRUE)), 
            months = paste0(unique(sort(lubridate::month(date))), collapse = ", ")) %>%
  ungroup() %>%
  distinct %>% kable()

```

### fish_guarding

```{r}
table(all_redd_data$fish_guarding, useNA = "ifany")
```

### redd_measured

```{r}
table(all_redd_data$redd_measured, useNA = "ifany")

```

### redd_width and redd_length

```{r}

summary(all_redd_data$redd_width)

summary(all_redd_data$redd_length)

ggplot() + 
  geom_histogram(data = all_redd_data, aes(redd_width)) 
  
ggplot() +   
  geom_histogram(data = all_redd_data, aes(redd_length))

```

### redd_count

```{r}
summary(all_redd_data$redd_count)

# redd data without Mill Creek
all_redd_data %>% 
  ungroup() %>%
 # filter(stream != "Mill Creek") %>% 
  select(redd_count) %>% 
  summary()

all_redd_data %>% 
ggplot() + 
  geom_histogram(aes(redd_count)) 

# ggplot() + 
#   geom_point(data = all_redd_data %>% filter(stream != "Mill Creek"), aes(x = date, y = redd_count)) +
#   facet_wrap(~ stream)

# I'm not positive that this is comparing apples-to-apples because Mill Creek is aggregated 
# by year
ggplot() + 
  geom_point(data = all_redd_data , aes(x = (year), y = redd_count)) +
  facet_wrap(~ stream) +
  coord_flip()
```

### fish_on_redd

-   Clear Creek is the only dataset containing this variable

```{r}
summary(all_redd_data$fish_on_redd)

which_streams_included <- all_redd_data %>% 
  filter(!is.na(fish_on_redd)) 
unique(which_streams_included$stream)
```

### pre_redd_substrate_size

* TODO: standardize!!!!! 

```{r}
table(all_redd_data$pre_redd_substrate_size)

# tmp <- all_redd_data %>% 
#   separate(pre_redd_substrate_size, c('A', 'B'), sep = "to", remove = FALSE) %>%
#   mutate(tmp = case_when(!is.na(pre_redd_substrate_size) ~ paste0(A,"-", B),
#                                              TRUE ~ as.character(pre_redd_substrate_size)))


```

### redd_substrate_size

* TODO: standardize!!!!! 

```{r}
table(all_redd_data$redd_substrate_size)

```

### tail_substrate_size

```{r}
table(all_redd_data$tail_substrate_size)

```

## Save Cleaned Data to Google Cloud

```{r}
# clean_passage <- combined_upstream_passage
# 
# knitr::kable(clean_passage %>% head)
```

```{r, eval=FALSE}
# Write to google cloud 
# Name file [watershed]_[data type].csv

# write_csv(clean_passage, "../../data/standard-format-data/standard_adult_upstream_passage.csv")
# f <- function(input, output) write_csv(input, file = output)
# gcs_upload(clean_passage,
#            object_function = f,
#            type = "csv",
#            name = "standard-format-data/standard_adult_upstream_passage.csv")
```
