---
title: "Standardize Mark Recapture Data"
author: "Erin Cain"
date: "4/20/2022"
output: 
  html_document:
  theme: flatly

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)

color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")
```

## Mark Recapture Data Standardization 

FlowWest received environmental data for rst from []monitoring programs: 
  
* Battle Creek 
* Butte Creek 
* Clear Creek 
* Deer Creek
* Feather River
* Mill Creek
* Knights Landing 
* Yuba River

No data is provided for tisdale

## Standard format for Mark Recapture Data

Data dictionary for standard format: 

(B - Battle Creek, Bu - Butte Creek, C - Clear Creek, F - Feather River, 
  D - Deer Creek, F - Feather River, M - Mill Creek, K - Knights Landing,
  Y - Yuba River)

| column name | tributary collects | definition | 
  | :------------------------------------ | :------------ |  :---------------------------------------------------------------- | 
  | tributary | B, Bu, C, D, F, M, KL, Y | Which Spring Run JPE is the data from |
  | date | B, Bu, C, D, F, M, KL, Y | date that samples are taken |
  | sample_id | B, C | sample id that links with catch or trap condition tables |
  | velocity | B, Bu, C, Y | velocity measure in feet per seccond |
  | temperature |  Bu, D, F, KL, M, Y | temperature measure in degrees celcius|
  | turbidity | B, Bu, C, D, F, M, KL, Y | turbidity measure in NTU |
  | secchi | Bu, KL | secchi measure in ft |
  | weather  | B, Bu, C, D, M | Weather condition at time of trapping  |
  | lunar_phase | B, C | Phase of moon (full, half, new) |
  | habitat | B, C | Habitat trap is located in (backwater pool, glide, run) |
  | thalweg | B, C | If the trap is fishing in the thawleg |
  | flow_start_meter | B, C | Start time of flow meter | 
  | flow_end_meter | B, C | End time of flow meter |
  | flow_set_time | B, C | Flow set time ? |
  | flow | D, KL, M | flow in cfs |
  | site | Y, Bu, F | site name  |
  | sub_site | F | sub site name |
  
  All released and recaptured fish are chinook salmon. 

## Read in data {.tabset} 

### Battle Creek  

#### Columns Removed 
The dataset included operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
# gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# gcs_get_object(object_name = "rst/battle-creek/data/battle_rst_environmental.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/battle_rst_environmental.csv",
#                overwrite = TRUE)

battle_env <- read_csv("../../data/rst/battle_rst_environmental.csv") %>%
  select(sample_id, date = sample_date, flow_start_meter, flow_end_meter, flow_set_time, 
         velocity, turbidity, weather = weather_code, lunar_phase, habitat, thalweg) %>%
  mutate(tributary = "Battle Creek") %>%
  glimpse
```
### Butte Creek  

#### Columns Removed 
The dataset included catch, operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/butte-creek/data/butte_rst.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/butte_rst_environmental.csv",
#                overwrite = TRUE)

butte_env <- read_csv("../../data/rst/butte_rst_environmental.csv", 
                      col_types = list(
                      "D", "c", "c", "l", "n", "n", "n", "n", "c", "t", 
                      "c", "c", "n", "n", "n", "n", "l", "l", "n", "c", 
                      "n", "n", "c"
                      )) %>%
  select(date, site = station, weather, temperature, turbidity, secchi, velocity) %>%
  mutate(tributary = "Butte Creek") %>%
  glimpse
```


### Clear Creek  

#### Columns Removed 
The dataset included operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/clear-creek/data/clear_rst_environmental.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/clear_rst_environmental.csv",
#                overwrite = TRUE)

clear_env <- read_csv("../../data/rst/clear_rst_environmental.csv") %>%
  select(sample_id, date = sample_date, flow_start_meter, flow_end_meter, flow_set_time, 
         velocity, turbidity, weather = weather_code, lunar_phase, habitat, thalweg) %>%
  mutate(tributary = "Clear Creek") %>%
  glimpse
```

### Deer Creek

#### Columns Removed 
The dataset included operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/deer-creek/data/deer_rst.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/deer_rst_environmental.csv",
#                overwrite = TRUE)

deer_env <- read_csv("../../data/rst/deer_rst_environmental.csv", 
                     col_types = list("D", "c", "n", "n", "n", "n", "n", "n", "c", 
                                      "n", "c", "c", "n")) %>% 
  select(date, site = location, flow, turbidity, weather, temperature = water_temperature_celsius) %>%
  mutate(temperature = round((temperature - 32) * (5/9), 1),
         tributary = "Deer Creek") %>%
  glimpse

```

### Feather River

#### Columns Removed 
The dataset included operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/feather-river/data/feather_rst_effort.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/feather_rst_environmental.csv",
#                overwrite = TRUE)

feather_env <- read_csv("../../data/rst/feather_rst_environmental.csv") %>% 
  select(date, site = site_name, sub_site = sub_site_name, temperature = water_temp_c, 
         turbidity = turbidity_ntu) %>%
  mutate(tributary = "Feather Creek") %>%
  glimpse

```

### Lower Sac - Knights Landing

#### Columns Removed 
The dataset included operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/lower-sac-river/data/knights-landing/knl_combine_sampling_effort_clean.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/knights_landing_rst_environmental.csv",
#                overwrite = TRUE)

knights_landing_env <- read_csv("../../data/rst/knights_landing_rst_environmental.csv") %>% 
  select(date, site = location, flow = flow_cfs, secchi = secchi_ft, temperature = water_t_f, turbidity) %>%
  mutate(temperature = round((temperature - 32) * (5/9), 1), 
         tributary = "Lower Sac - Knights Landing") %>%
  glimpse

```


### Mill Creek

#### Columns Removed 
The dataset included operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/mill-creek/data/mill_rst.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/mill_rst_environmental.csv",
#                overwrite = TRUE)

mill_env <- read_csv("../../data/rst/mill_rst_environmental.csv") %>% 
  select(date, site = location, flow, temperature = water_temperature, turbidity, weather) %>%
  mutate(temperature = round((temperature - 32) * (5/9), 1),
         tributary = "Mill Creek") %>%
  glimpse
mill_env$weather %>% unique()
```
### Yuba River

#### Columns Removed 
The dataset included operations and environmental data. Below only columns 
for environmental data are selected. 

```{r, message=FALSE, warning=FALSE}
# gcs_get_object(object_name = "rst/yuba-river/data/yuba_rst.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/rst/yuba_rst_environmental.csv",
#                overwrite = TRUE)

yuba_env <- read_csv("../../data/rst/yuba_rst_environmental.csv", 
                     col_types = list("T", "t", "c", "n", "n", "n", "c", "n", 
                                      "n", "c", "n", "n", "c", "c", "n", "n", 
                                      "c", "n", "c", "c")) %>% 
  mutate(date = lubridate::as_date(date)) %>%
  select(date, temperature, turbidity, velocity, site = location) %>%
  mutate(tributary = "Yuba River") %>%
  glimpse

```

## Combine data: 
```{r}
combined_env <- bind_rows(battle_env, butte_env, clear_env, deer_env,
                          feather_env, knights_landing_env, mill_env, 
                          yuba_env) %>% glimpse


```

## Explore Varibles{.tabset} 

### sample_id 
Battle and Clear Creeks are the only tribs that give a sample ID. The rest of the tribs are entirely NA for this column. 

```{r}
combined_env %>% group_by(tributary) %>%
  summarize(unique_sample_id = length(unique(sample_id)))
```


### date 

```{r}
combined_env %>% 
  ggplot() + 
  geom_point(aes(x = date, y = tributary, color = tributary), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) + 
  theme(legend.position = "none", 
        text = element_text(size = 18)) 
```

### flow_start_meter 

Battle and Clear Creeks are the only tribs that give a flow_start_meter. The rest of the tribs are entirely NA for this column. 

```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = flow_start_meter, y = tributary, color = tributary), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

### flow_start_meter 

Battle and Clear Creeks are the only tribs that give a flow_end_meter. The rest of the tribs are entirely NA for this column. 

```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = flow_end_meter, y = tributary, color = tributary), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

### flow_set_time   

Battle and Clear Creeks are the only tribs that give a flow_set_time. The rest of the tribs are entirely NA for this column.

```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = flow_set_time, y = tributary, color = tributary), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

### velocity   

Battle, Butte, Clear, and Yuba provide velocity data. 
```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = velocity, y = tributary, color = tributary), alpha = .5, binwidth = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```
Zoom in on 0 - 10 velocity range: 
```{r}
combined_env %>% 
  filter(velocity < 10) %>%
  ggplot() + 
  geom_histogram(aes(x = velocity, fill = tributary), alpha = .5, binwidth = .3, position="identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "buttom", 
        text = element_text(size = 18))
```


### turbidity   

All watersheds provide turbidity data

```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = turbidity, y = tributary, color = tributary), alpha = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```
Zoom in on 0 - 25 velocity range: 
```{r}
combined_env %>% 
  filter(turbidity < 50) %>% # 1,647 values above 50 (274,049 rows total)
  ggplot() + 
  geom_histogram(aes(x = turbidity, fill = tributary), alpha = .5, binwidth = .3, position="identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```
### weather   

Battle, Butte, Clear Deer, and Mill collect weather data: 
```{r}
unique(combined_env$weather)
combined_env %>% 
  ggplot() + 
  geom_bar(aes(x = weather, fill = tributary), position = "dodge") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Fix encodings of weather data 
```{r}
combined_env <- combined_env %>% 
  mutate(weather = case_when(weather %in% c("precipitation", "rain", "rainy", "windy and precipitation") ~ "precipitation",
                             weather %in% c("sunny", "clear") ~ "clear",
                             weather %in% c("partially cloudy", "partly cloudy") ~ "partially cloudy",
                             weather %in% c("fog", "foggy") ~ "clear",
                             weather %in% c("windy", "windy and clear") ~ "windy",
                             weather == "cloudy" ~ "cloudy")) %>% glimpse

table(combined_env$weather, useNA = "ifany")

combined_env %>% 
  ggplot() + 
  geom_bar(aes(x = weather, fill = tributary), position = "dodge") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### lunar_phase 

```{r}
table(combined_env$lunar_phase, useNA = "ifany")
```
Only battle and clear give lunar_phase: 

```{r}
combined_env %>% 
  ggplot() + 
  geom_bar(aes(x = lunar_phase, fill = tributary), position = "dodge") +
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))

```
### habitat
```{r}
table(combined_env$habitat, useNA = "ifany")
```
Only battle and clear give habitat: 

```{r}
combined_env %>% 
  ggplot() + 
  geom_bar(aes(x = habitat, fill = tributary), position = "dodge") +
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))


```
### thawleg
```{r}
table(combined_env$thalweg, useNA = "ifany")
```
Only battle and clear give thawleg: 

```{r}
combined_env %>% 
  ggplot() + 
  geom_bar(aes(x = thalweg, fill = tributary), position = "dodge") +
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))

```

### tributary 
```{r}
table(combined_env$tributary, useNA = "ifany")
```
### temperature

TODO fix funky temp stuff 

Butte, Deer, Feather, Knights Landing, Mill, and Yuba provide temperature data

```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = temperature, y = tributary, color = tributary), alpha = .5, binwidth = .5) + 
  theme_minimal() +
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```
Zoom in on 0 - 25 temperature range: 
```{r}
combined_env %>% 
  filter(temperature < 100) %>% # Filter out outliers
  ggplot() + 
  geom_histogram(aes(x = temperature, fill = tributary), alpha = .5, binwidth = 2.5, position = "identity") + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### secchi

Knights Landing and Butte provide secchi data

```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = secchi, y = tributary, color = tributary), alpha = .5, binwidth = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

```{r}
combined_env %>% 
  ggplot() + 
  geom_histogram(aes(x = secchi, fill = tributary), binwidth = 2.5, position = "identity", alpha = .5) + 
  theme_minimal() + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "bottom", 
        text = element_text(size = 18))
```

### site
```{r}
table(combined_env$site, useNA = "ifany")
```
### flow 

Mill, Deer, and Knights Landing provide flow data
```{r}
combined_env %>% 
  ggplot() + 
  geom_boxplot(aes(x = flow, y = tributary, color = tributary), alpha = .5, binwidth = .5) + 
  theme_minimal() + 
  scale_color_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

```{r}
combined_env %>% 
  filter(tributary %in% c("Mill Creek", "Lower Sac - Knights Landing", "Deer Creek")) %>%
  ggplot() + 
  geom_histogram(aes(x = flow, fill = tributary), binwidth = 10) + 
  theme_minimal() + 
  facet_wrap(~tributary, ncol = 1, drop = TRUE, scales = "free") + 
  scale_fill_manual(values = color_pal) +
  theme(legend.position = "none", 
        text = element_text(size = 18))
```

### sub_site 
```{r}
table(combined_env$sub_site, useNA = "ifany")
```



## Save unsumarized data to google cloud


```{r}
knitr::kable(align = 'c', head(combined_env, 5)) 
```


```{r, eval = FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(combined_env,
           object_function = f,
           type = "csv",
           name = "standard-format-data/standard_RST_environmental.csv")

```