---
title: "Standardize Carcass Data"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)

color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")
```

## Carcass Data Standardization

FlowWest received carcass data from 4 monitoring programs:

-   Battle Creek
-   Butte Creek
-   Clear Creek
-   Feather River

## Standard format Carcass Data

Data dictionary for standard format:

(B - Battle Creek, Bu - Butte Creek, C - Clear Creek, F - Feather River,
D - Deer Creek, F - Feather River, M - Mill Creek, S- Sacramento River,
Y - Yuba River)

| column name   | stream collects         | definition                                               |
|:-----------------|:-------------------|:---------------------------------|
| stream      | | Which stream the RST is located on                       |
| site          | | site name                                                |


## Read in data {.tabset}

### Battle Creek

#### Columns Removed

```{r, message=FALSE, warning=FALSE}
# gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# # Set global bucket 
# gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
# 
# gcs_list_objects()
# 
# gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_carcass.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "../../data/redd-carcas-holding/battle_carcass.csv",
#                overwrite = TRUE)

battle_carcass <- read_csv("../../data/redd-carcas-holding/battle_carcass.csv") %>% glimpse
```

### Butte Creek

#### Columns Removed


```{r, message=FALSE, warning=FALSE}

```

### Clear Creek

#### Columns Removed


```{r, message=FALSE, warning=FALSE}

```

### Feather River

#### Columns Removed


```{r, message=FALSE, warning=FALSE}

```
