---
title: "Feather River Hallprint Data QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River hallprint adult broodstock selection and enumeration data

## Description of Monitoring Data

Hallprint tagging data only describes fish marked, giving hallprint tag numbers and dates tagged for each fish. It does not include any other sampling of fish like size or genetics. The data that we have does not contain any additional information about fish returns. We currently have one file describing fish returns. We can likely acquire more but there are issues with the return data. 

**Timeframe:** 2004 - 2020 
  
**Season:** Tagging typically begins in April and goes through June

**Provisional or QC data** Data is QC  
  
**Sampling Location:** Some tags have a acoustic location attribute. 
  
**Data Contact:** [Byron Mache](mailto:Byron.Mache@water.ca.gov) 
  
Each fish is given two tags. The `second_tag_number` is NA if the first tag and the second tag have the same number. 
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
gcs_get_object(object_name = 
                 "adult-upstream-passage-monitoring/feather-river/hallprint/data-raw/all_hallprints_all_years.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_feather_hallprint_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
raw_hallprint_data <- read_excel("raw_feather_hallprint_data.xlsx", 
                                 col_types = c("text", "date", "text",
                                               "text", "text", "text",
                                               "text")) %>% glimpse()
```

## Data transformations

```{r}
cleaner_hallprint_data <- raw_hallprint_data %>% 
  rename("date" = Date, "tag_number" = `Tag#`, "second_tag_number" = `2ndTag#`, 
         "color" = Color, "acoustic" = Acoustic, 
         "acoustic_location" = `Acoustic Location`) %>%
  mutate(date = as.Date(date)) %>%
  glimpse() 
```


## Explore Numeric Variables: {.tabset}

No numeric variables. 
```{r}
# Filter clean data to show only numeric variables 
cleaner_hallprint_data %>% select_if(is.numeric) %>% colnames()
```

## Explore Categorical variables: {.tabset}

```{r}
# Filter clean data to show only categorical variables
cleaner_hallprint_data %>% select_if(is.character) %>% colnames()
```

### Variable: `ID`

All IDs are unique as anticipated:
```{r}
length(unique(cleaner_hallprint_data$ID)) == nrow(cleaner_hallprint_data)
```

**NA and Unknown Values**
```{r}
cleaner_hallprint_data %>% filter(is.na(ID))
```
There are no NA values in `ID`. 

### Variable: `tag_number`

Are all the tags unique? No. In more recent years tag numbera start to be unique and we have less NA values but there are 8 years where tags are not unique. 
```{r}
cleaner_hallprint_data %>% 
  group_by(year = year(date)) %>%
  summarize(count_unique_tags = length(unique(tag_number)), 
            number_tags = n(), 
            na_tags = sum(is.na(tag_number)),
            all_tags_na_or_unique = ifelse(count_unique_tags == number_tags + na_tags, TRUE, FALSE))
```

```{r}
cleaner_hallprint_data %>% 
  group_by(year = year(date)) %>%
  mutate(total_tags_per_year = sum(!is.na(tag_number))) %>% 
  ungroup() %>% 
  ggplot(aes(x = year, y = total_tags_per_year)) +
  geom_col() + 
  theme_minimal() + 
  labs(title = "Total fish tagged by year",
       y = "Number of Fish Tagged")  
```

More fish were tagged in 2013 than any other year. 
Numbers of fish tagged are not even over the period of record. 

```{r}
# TODO Could also do a visual of when in the year this is happening (tags by month) - facet on year (some portion of years or average over years)
cleaner_hallprint_data %>% 
  group_by(month = month(date)) %>%
  mutate(total_tags_per_month = sum(!is.na(tag_number))) %>% 
  ungroup() %>% 
  ggplot(aes(x = month, y = total_tags_per_month)) +
  geom_col() + 
  facet_wrap(~year(date)) +  
  theme_minimal() + 
  labs(title = "Tags by month",
       y = "Number of Fish Tagged")  
```

Generally tagging occurs in May and June. 

**NA and Unknown Values**
```{r}
nrow(subset(cleaner_hallprint_data, is.na(cleaner_hallprint_data$tag_number)))
```
 * `r round(sum(is.na(cleaner_hallprint_data$tag_number))/nrow(cleaner_hallprint_data), 3)` % of values in the `tag_number` column are NA.

### Variable: `second_tag_number`

Similarly not all unique. Only one year where count of unique tags equals total number of unique tags. 
```{r}
cleaner_hallprint_data %>% 
  group_by(year = year(date)) %>%
  summarize(count_unique_second_tags = length(unique(second_tag_number)), 
            number_second_tags = n(), 
            na_tags = sum(is.na(second_tag_number)),
            all_tags_na_or_unique = ifelse(count_unique_second_tags == number_second_tags + na_tags, TRUE, FALSE))
```

 
**NA and Unknown Values**
 * `r round(sum(is.na(cleaner_hallprint_data$second_tag_number))/nrow(cleaner_hallprint_data), 3)` % of values in the `second_tag_number` column are NA. Byron shared that if the `tag_number` equals the `second_tag_number` they leave `second_tag_number` as NA. 
 
```{r}
# Replace 2nd tag number with first tag number if second tag number is NA
cleaner_hallprint_data$second_tag_number <- if_else(is.na(cleaner_hallprint_data$second_tag_number), 
                                                    cleaner_hallprint_data$tag_number, 
                                                    cleaner_hallprint_data$second_tag_number)
```


### Variable: `color`
```{r}
# TODO email Byron to ask about color 
table(cleaner_hallprint_data$color) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**
  
Provide a stat on NA or unknown values
  
### Variable: `acoustic`
```{r}
table(cleaner_hallprint_data$acoustic) 

# Make into a booleen 
cleaner_hallprint_data$acoustic <- case_when(
  is.na(cleaner_hallprint_data$acoustic) ~ NA,
  cleaner_hallprint_data$acoustic == "Y" ~ TRUE,
  cleaner_hallprint_data$acoustic == "N" ~ FALSE,
)

table(cleaner_hallprint_data$acoustic) 
```

**NA and Unknown Values**

  * `r round(sum(is.na(cleaner_hallprint_data$acoustic))/nrow(cleaner_hallprint_data), 3)` % of values in the `acoustic` column are NA. 
### Variable: `acoustic_location`
```{r}
table(cleaner_hallprint_data$acoustic_location) 
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

  * `r round(sum(is.na(cleaner_hallprint_data$acoustic_location))/nrow(cleaner_hallprint_data), 3)` % of values in the `acoustic_location` column are NA. 
  
**Summary of identified issues:**
* One identified issue is that not all `tags_numbers` or `second_tag_numbers` are unique. 

### Save cleaned data back to google cloud 

```{r}
cleaner_hallprint_data %>% glimpse()
```

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv

```
