---
title: "Deer Creek Adult Upstream Count QC"
author: "Inigo Peng"
date: "10/19/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
```

# Deer Creek Adult Upstream Passage Monitoring Data 2014 to 2023 

**Description of Monitoring Data**

Adult spring run salmon passage daily estimates based on data collected at SVRIC Dam on Deer River via video monitoring.

**QC/Raw or Count** QC Counts

**Timeframe:** 

2014 to 2023

**Completeness of Record throughout timeframe:**

Some missing data for variable flow

**Sampling Location:** 

* SVRIC Dam   

**Data Contact:**  [Ryan Revnak](mailto:Ryan.Revnak@wildlife.ca.gov) and [Doug Killam](mailto:Doug.Killam@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# get filenames for deer creek
deer_creek_filenames <- gcs_list_objects() |> 
  filter(str_detect(name, "adult-upstream-passage-monitoring/deer-creek/data-raw/")) |> 
  filter(str_detect(name, "DCVS|DSVS|DNVS")) |> 
  pull(name)

# define function to download all files
get_files_from_cloud <- function(filename) {
  print(substr(filename, 55, nchar(filename)))
  gcs_get_object(object_name = filename,
                 bucket = gcs_get_global_bucket(),
                 saveToDisk = paste0("data-raw/qc-markdowns/adult-upstream-passage-monitoring/deer-creek/", substr(filename, 55, nchar(filename))),
                 overwrite = TRUE)
  message(paste0("downloaded ", filename))
}

lapply(deer_creek_filenames, get_files_from_cloud)

# 2020-2021 passage data
# gcs_get_object(object_name = "adult-upstream-passage-monitoring/deer-creek/data-raw/Copy of DCVS both 2021-22 Spring FINAL 8-17-22.xlsx",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "data-raw/qc-markdowns/adult-upstream-passage-monitoring/deer-creek/deer_creek_passage_counts_2021.xlsx",
#                overwrite = TRUE)
# # 2022-2023 passage data
# gcs_get_object(object_name = "adult-upstream-passage-monitoring/deer-creek/data-raw/DCVS both 2022-23 Spring DRAFT 8-26-23.xlsx",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "data-raw/qc-markdowns/adult-upstream-passage-monitoring/deer-creek/deer_creek_passage_counts_2022.xlsx",
               # overwrite = TRUE)
```

## Read in and clean raw count data for each year

```{r, message = FALSE, warning = FALSE}
save_to_prefix <- "data-raw/qc-markdowns/adult-upstream-passage-monitoring/deer-creek/"
```

This requires a different approach for each year because the file formats change. We also need to distinguish between files for one ladder (north or south) vs. files for both ladders. 

### 2013-2014

```{r, message=FALSE, warning=FALSE}
deer_2013_raw <- readxl::read_xls(paste0(save_to_prefix, "DCVS 2013-2014 as of 4-14-17.xls"), 
                                  sheet = "Sheet1",
                                  skip = 24) |> 
  glimpse()

deer_2013 <- deer_2013_raw |> 
  janitor::clean_names() |> 
  filter(!date %in% c("Date", "Totals:")) |> 
  # this converts "no data" to NA
  mutate(across(south_ladder_daily_spring_run_chinook_total:post_diversion_average_daily_flow_cfs, as.numeric)) |> 
  select(date, south_ladder_daily_spring_run_chinook_total, north_ladder_daily_spring_run_chinook_total, avg_flow = post_diversion_average_daily_flow_cfs) |> 
  pivot_longer(south_ladder_daily_spring_run_chinook_total:north_ladder_daily_spring_run_chinook_total,
               names_to = "ladder", 
               values_to = "count") |> 
  mutate(ladder = ifelse(ladder == "south_ladder_daily_spring_run_chinook_total", "south", "north"),
         date = mdy(date)) |> 
  glimpse()
```

### 2014-2015
- in separate files for `north` and `south` ladders

```{r, message=FALSE, warning=FALSE}
deer_2014_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DNVS 2014-2015 as of  6-30-2015.xlsx"), 
                                  sheet = "DATA ENTRY",
                                  skip = 4) |> 
  mutate(ladder = "north") |> 
  glimpse()

deer_2014_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DSVS 2014-2015 as of 6-30-2015.xlsx"), 
                                  sheet = "DATA ENTRY",
                                  skip = 4) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2014 <- bind_rows(deer_2014_raw_north, 
                       deer_2014_raw_south) |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```

### 2015-2016
- in separate files for `north` and `south` ladders

```{r, message=FALSE, warning=FALSE}
deer_2015_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DNVS 2015-2016 as of 8-12-2016.xlsx"), 
                                  sheet = "DATA ENTRY",
                                  skip = 5) |> 
  mutate(ladder = "north") |> 
  glimpse()

deer_2015_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DSVS 2015-2016 as of 8-12-2016.xlsx"), 
                                  sheet = "DATA ENTRY",
                                  skip = 5) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2015 <- bind_rows(deer_2015_raw_north, 
                             deer_2015_raw_south) |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```

### 2016-2017
- `north` and `south` ladders in different tabs

```{r, message=FALSE, warning=FALSE}
deer_2016_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2016-2017 as of  9-19-18.xlsx"), 
                                  sheet = "DNVS 16-17 Spring",
                                  skip = 5) |> 
  mutate(ladder = "north",
         across(`UP Salmon`:`DN Trout`, as.numeric)) |> 
  glimpse()

deer_2016_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2016-2017 as of  9-19-18.xlsx"), 
                                         sheet = "DSVS 16-17 Spring",
                                  
                                  skip = 5) |> 
  mutate(ladder = "south",
         across(`UP Salmon`:`DN Trout`, as.numeric),
         `Viewing Adjust!!` = as.character(`Viewing Adjust!!`)) |> 
  glimpse()

deer_2016 <- bind_rows(deer_2016_raw_north,
                       deer_2016_raw_south) |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```


### 2017-2018
- `north` and `south` ladders in different tabs

```{r, message=FALSE, warning=FALSE}
deer_2017_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2017-2018 as of 9-14-18.xlsx"), 
                                  sheet = "DEER-North-17-18",
                                  skip = 5) |> 
  mutate(ladder = "north") |>
  glimpse()

deer_2017_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2017-2018 as of 9-14-18.xlsx"), 
                                         sheet = "DEER-South-17-18",
                                  
                                  skip = 5) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2017 <- bind_rows(deer_2017_raw_north,
                       deer_2017_raw_south) |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```

### 2018-2019
- `north` and `south` ladders in different tabs

```{r, message=FALSE, warning=FALSE}
deer_2018_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2018-2019 as of 12-14-20.xlsx"), 
                                  sheet = "DEER-North-18-19",
                                  skip = 5) |> 
  mutate(ladder = "north") |>
  glimpse()

deer_2018_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2018-2019 as of 12-14-20.xlsx"), 
                                         sheet = "DEER-South-18-19",
                                  
                                  skip = 5) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2018 <- bind_rows(deer_2018_raw_south,
                       deer_2018_raw_north) |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```
### 2019-2020
- `north` and `south` ladders in different tabs

```{r, message=FALSE, warning=FALSE}
deer_2019_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2019-2020 as of 9-9-21.xlsx"), 
                                  sheet = "DEER-North-19-20",
                                  skip = 5) |> 
  mutate(ladder = "north") |>
  glimpse()

deer_2019_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2019-2020 as of 9-9-21.xlsx"), 
                                         sheet = "DEER-South-19-20",
                                  
                                  skip = 5) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2019 <- bind_rows(deer_2019_raw_south,
                       deer_2019_raw_north) |> 
  janitor::clean_names() |> 
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```

### 2020-2021
- `north` and `south` ladders in different tabs

```{r, message=FALSE, warning=FALSE}
deer_2020_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS Both 2020-2021 as of 10-13-22.xlsx"), 
                                  sheet = "DEER-North 20-21-DK ",
                                  skip = 2) |> 
  mutate(ladder = "north") |>
  glimpse()

deer_2020_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS Both 2020-2021 as of 10-13-22.xlsx"), 
                                         sheet = "DEER-South-20-21",
                                  
                                  skip = 4) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2020 <- bind_rows(deer_2020_raw_north,
                       deer_2020_raw_south) |> 
  janitor::clean_names() |>
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```

### 2021-2022
- `north` and `south` ladders in different tabs

```{r, message=FALSE, warning=FALSE}
deer_2021_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2021-22 Spring FINAL 3-3-23.xlsx"), 
                                  sheet = "DEER-North-21-22",
                                  skip = 5) |> 
  mutate(ladder = "north") |>
  glimpse()

deer_2021_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2021-22 Spring FINAL 3-3-23.xlsx"), 
                                         sheet = "DEER-South-21-22",
                                  
                                  skip = 5) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2021 <- bind_rows(deer_2021_raw_north,
                       deer_2021_raw_south) |> 
  janitor::clean_names() |>
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```
### 2022-2023
- `north` and `south` ladders in different tabs

```{r, message=FALSE, warning=FALSE}
deer_2022_raw_north <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2022-2023 as of 10-16-23.xlsx"), 
                                  sheet = "DEER-North-22-23",
                                  skip = 5) |> 
  mutate(ladder = "north") |>
  glimpse()

deer_2022_raw_south <- readxl::read_xlsx(paste0(save_to_prefix, "DCVS both 2022-2023 as of 10-16-23.xlsx"), 
                                         sheet = "DEER-South-22-23",
                                  
                                  skip = 5) |> 
  mutate(ladder = "south") |> 
  glimpse()

deer_2022 <- bind_rows(deer_2022_raw_north,
                       deer_2022_raw_south) |> 
  janitor::clean_names() |>
  mutate(across(up_salmon:dn_jacks, as.numeric),
         start_time = format(start_time, "%H:%M:%S"),
         viewing_adjust = as.character(viewing_adjust)) |> 
  select(date, start_time, up_salmon, dn_salmon, 
         up_jacks, dn_jacks, viewing_adjust, ladder) |> 
  pivot_longer(up_salmon:dn_jacks,
               names_to = "count_type", 
               values_to = "count") |> 
  mutate(passage_direction = ifelse(count_type == "up_salmon", "up", "down"),
         jack = ifelse(str_detect(count_type, "jacks"), TRUE, FALSE)) |> 
  glimpse()
```


## Data transformations

```{r, message=FALSE, warning=FALSE}
all_years_raw <- bind_rows(deer_2013, deer_2014, deer_2015, deer_2016,
                           deer_2017, deer_2018, deer_2019, deer_2020,
                           deer_2021, deer_2022) |> 
  glimpse()

all_deer_data <- all_years_raw |> 
  select(-count_type) |> 
  # pulled from excel tabs
  mutate(viewing_adjust = case_when(is.na(viewing_adjust) | viewing_adjust == 0 ~ "VAKI",
                                    viewing_adjust == 1 ~ "turbid",
                                    viewing_adjust == 3 ~ "equipment fail",
                                    viewing_adjust == 4 ~ "DIDSON",
                                    viewing_adjust == 6 ~ "GV",
                                    viewing_adjust == 7 ~ "Hwell",
                                    TRUE ~ NA)) |> 
  relocate(date, start_time, count, passage_direction, jack, viewing_adjust, flow = avg_flow) |> 
  filter(!is.na(date)) |> 
  mutate(date = as.Date(date),
         count = ifelse(is.na(count), 0, count)) |> 
  glimpse()
```


## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data. 

```{r data_dictionary}
percent_na <- all_deer_data |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(all_deer_data),
                          description = c("Date of sampling",
                                          "Start time of sampling",
                                          "Count of fish",
                                          "Direction of fish passage",
                                          "Whether or not the fish was jack sized (fewer than 24 inches)",
                                          "Reason camera needed to be adjusted",
                                          "Flow in CFS",
                                          "Ladder (north or south) data came from"),
                          
                          percent_na = round(percent_na$value*100)
                          
)
knitr::kable(data_dictionary)
```

## Explore `date`

Check for outlier and NA values

```{r}
summary(all_deer_data$date)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_deer_data$date))/nrow(all_deer_data), 3)*100` % of values in the `date` column are NA.

## Explore Numerical Values

```{r}
all_deer_data |> select_if(is.numeric) |> colnames()
```


### Variable:`count` 
```{r}
all_deer_data |> 
  mutate(year = as.factor(year(date))) |> 
  ggplot(aes(x=date, y = count))+
  geom_line()+
  facet_wrap(~year, scales = "free")+
  theme_minimal()+
  labs(title = "Daily Count From 2014 - 2023")
```


```{r}
all_deer_data |> 
  mutate(year = as.factor(year(date))) |> 
  group_by(year) |> 
  summarise(total = sum(count, na.rm = T)) |>
  ggplot(aes(x = year, y = total, group = 1))+
  geom_line()+
  geom_point(aes(x=year, y = total))+
  theme_minimal()+
  labs(title = "Count from 2014 - 2023",
       y = "Total count")
```

**Numeric Summary of count From 2014 to 2023**
```{r}
summary(all_deer_data$count)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_deer_data$count))/nrow(all_deer_data), 3)*100` % of values in the `count` column are NA.

### Variable:`flow`

Flow in cfs, only collected in `2014`

```{r fig.height=8, fig.width=8}
all_deer_data |> 
  group_by(date) |>
  mutate(flow = mean(flow, na.rm = T)) |>
  ungroup() |> 
  mutate(year = as.factor(year(date)),
         fake_year = 1900,
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) |> 
  ggplot(aes(x = fake_date, y = flow, color = year)) + 
  scale_color_brewer(palette = "Dark2")+
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Water Flow (colored by year)",
       y = "Average Daily Flow", 
       x = "Date")  
```
```{r fig.height=8, fig.width=8}
all_deer_data |> 
  mutate(year = as.factor(year(date))) |> 
  ggplot(aes(x=flow, fill = year))+
  scale_fill_brewer(palette = "Dark2")+
  geom_histogram()+
  theme_minimal()+
  labs(title = "Distribution of Flow")
```

**Numeric Summary of flow From 2014 to 2023**
```{r}
summary(all_deer_data$flow)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_deer_data$flow))/nrow(all_deer_data), 3)*100` % of values in the `flow` column are NA.

### Variable:`passage_direction` 

```{r}
all_deer_data |> 
  mutate(year = as.factor(year(date))) |> 
  ggplot(aes(x=date, y = count, fill = passage_direction))+
  geom_col()+
  scale_fill_brewer(palette = "Dark2")+
  facet_wrap(~year, scales = "free")+
  theme_minimal() +
  labs(title = "Daily Count From 2014 - 2023")
```


```{r}
all_deer_data |> 
  mutate(year = as.factor(year(date))) |> 
  group_by(date, passage_direction) |> 
  summarise(total_passage = sum(count)) |> 
  ggplot(aes(x=date, y = total_passage, color = passage_direction))+
  geom_line(linewidth = 0.8, alpha = 0.5)+
  scale_color_brewer(palette = "Dark2")+
  facet_wrap(~year(date), scales = "free")+
  theme_minimal() +
  labs(title = "Daily Count From 2014 - 2023")
```

**Numeric Summary of passage_direction from 2014 to 2023**
```{r}
table(all_deer_data$passage_direction)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_deer_data$passage_direction))/nrow(all_deer_data), 3)*100` % of values in the `passage_direction` column are NA.

### Variable:`jack` 

```{r}
all_deer_data |> 
  group_by(date, jack) |> 
  summarise(total_count = sum(count)) |> 
  ggplot(aes(x=date, y = total_count, fill = jack)) +
  geom_col(position = "dodge")+
  facet_wrap(~year(date), scales = "free")+
  theme_minimal()+
  labs(title = "Daily Count From 2014 - 2023")
```

**Numeric Summary of jack From 2014 to 2023**
```{r}
table(all_deer_data$jack)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_deer_data$jack))/nrow(all_deer_data), 3)*100` % of values in the `jack` column are NA.

### Variable:`viewing_adjust` 

```{r}
all_deer_data |> 
  group_by(date, viewing_adjust) |> 
  summarise(total_count = sum(count)) |> 
  ggplot(aes(x=date, y = total_count, fill = viewing_adjust))+
  geom_col(position = "dodge")+
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~year(date), scales = "free")+
  theme_minimal()+
  labs(title = "Daily Count From 2014 - 2023")
```

**Numeric Summary of viewing_adjust From 2014 to 2023**
```{r}
all_deer_data |> 
  group_by(year(date), viewing_adjust) |> 
  tally()
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_deer_data$viewing_adjust))/nrow(all_deer_data), 3)*100` % of values in the `viewing_adjust` column are NA.

### Variable:`ladder` 

```{r}
all_deer_data |> 
  ggplot(aes(x=date, y = count, color = ladder))+
  geom_line()+
  facet_wrap(~year(date), scales = "free")+
  theme_minimal()+
  labs(title = "Daily Count From 2014 - 2023")
```


**Numeric Summary of ladder From 2014 to 2023**
```{r}
table(all_deer_data$ladder)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(all_deer_data$ladder))/nrow(all_deer_data), 3)*100` % of values in the `ladder` column are NA.


### Notes and Issues

* Numbers are really low for 2023

## Next steps

* Do they have flow/temp for those missing years?

### Add cleaned data back onto google cloud
```{r}
deer_upstream_counts <- all_deer_data |> glimpse()
```

```{r, eval = FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(deer_upstream_counts,
           object_function = f,
           type = "csv",
           name = "adult-upstream-passage-monitoring/deer-creek/data/deer_upstream_counts.csv",
           predefinedAcl = "bucketLevel")
```