---
title: "deer-holding-1986-to-1996-qc-checklist"
author: "Inigo Peng"
date: "10/19/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
library(knitr)
```

# Deer Creek Adult Holding Survey Data 1986 to 1996 

**Description of Monitoring Data**

The adult holding survey data is in two separate files: first one is from 1986 to 1996, second one is 1997 to 2020. The data is collected via snorkel counts.The data has been QCed and summarized.

The data from 1986 to 1996 has been reformatted for better access. Both original file and reformatted file could be found on google cloud.

**Timeframe:** 

1986 to 1996

**Completeness of Record throughout timeframe:**

* From 1986 to 1988 sampling reach was Hwy 32 to A-Line.
* From 1989 to 1991 sampling reach was Lower Falls to A-Line.
* From 1992 to 1996 sampling reach increased to 11 locations.


**Sampling Location:** 

* Hwy 32 to A-Line
* Lower Falls to A-Line
* Upper Falls to Potato Patch
* Potato Patch to Hwy 32
* Hwy 32 to Lower Falls
* Lower Falls to A-Line
* A-Line to Wilson
* Wilson to Polk Springs
* Polk Springs to Murphy
* Murphy to Beaver
* Beaver to Ponderosa
* Ponderosa to Homestead
* Homestead to 2E17

**Other Notes:**

In 1989 the section from Hwy 32 to Lower Falls was omitted from the indicator reach.  From 1992 to 1996, an average of 2% of the total salmon observed were in the omitted section: Hwy 32 to lower falls.  This leaves 29%  of the salmon  holding  in the reduced indicator reach.

**Data Contact:** [Matt Johnson](mailto:Matt.Johnson@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/deer-creek/data-raw/reformatted_deer_creek_snorkel_counts_by_reach_1986_1996.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "deer_creek_holding_raw.xlsx")
               # Overwrite = TRUE)
```

```{r}
raw_data = readxl::read_excel('deer_creek_holding_raw.xlsx')
glimpse(raw_data)
```

## Data Transformations
```{r}
cleaner_data <- raw_data %>% 
  set_names(tolower(colnames(raw_data))) %>% 
  glimpse()
```
## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- cleaner_data %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_data),
                          description = c("Sample year",
                                          "Description of location sampled",
                                          "Number of fish observed"),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
```


## Explore Categorical Variables

### Variable `location`
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(location = str_to_title(location),
         location = gsub("-", " ", location))
table(cleaner_data$location)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$location))/nrow(cleaner_data), 3)*100` % of values in the `location` column are NA.

## Explore Numeric Variables

### Variable `count`
```{r fig.height=7, fig.width=10}
#Find the most distinctive colours for visual
colourCount = length(unique(cleaner_data$location))
getPalette = colorRampPalette(brewer.pal(12, "Paired"))

cleaner_data %>%
  mutate(date =lubridate::ymd(year, truncated = 2L), .keep = "unused") %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = year, y = count, fill = location))+
  scale_fill_manual(values = getPalette(colourCount))+
  geom_col()+
  theme_minimal()+
  labs(title = "Adult Holding Count By Year")+
  ylim(0, 1500)
```
```{r}
cleaner_data %>% 
  ggplot(aes(x = count, y = location))+
  geom_boxplot() +
  theme_minimal()+
  labs(title = "Adult Holding Count By Reach")
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$count))/nrow(cleaner_data), 3)*100` % of values in the `count` column are NA.

**Numeric Annual Summary of count From 1986 to 1996**
```{r}
cleaner_data %>%
  group_by(year) %>%
  summarise(count = sum(count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```
## Next steps

- Recommend adding numerical location variable such as latitude, longitude; or maybe four variables (starting latitude, starting longitude, ending latitude, ending longitude)

## Add cleaned data back onto google cloud
```{r}
deer_adult_holding_survey_1986_to_1996 <- cleaner_data %>% glimpse()
```
```{r}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(deer_adult_holding_survey_1986_to_1996,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/deer-creek/data/deer_adult_holding_1986_to_1996.csv")
```
