---
title: "deer-holding-1997-to-2020-qc-checklist"
author: "Inigo Peng"
date: "10/19/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
library(knitr)
```

# Deer Creek Adult Holding Survey Data 1997 to 2020 

**Description of Monitoring Data**

The adult holding survey data is in three files: first one is from 1986 to 1996, second one is 1997 to 2020, and a third is 1997-2023. Originally the data file from 1997-2020 was used, but this is updated to use the 1997-2023 data as of `2024-05-1`. The data is collected via snorkel counts. The data has been qced and summed up.

**Timeframe:** 

1997 to 2023

**Completeness of Record throughout timeframe:**

* Some data lacks for Trail 2E17 to Dillon Cove

**Sampling Location:** 

* Upper Falls to Potato Patch
* Potato Patch to Lower Falls
* Lower Falls to A-Line
* A-Line to Wilson Cove
* Wilson Cove to Polk Springs
* Polk Springs to Murphy Trail
* Murphy Trail to Ponderosa Way
* Ponderosa Way to Trail 2E17
* Trail 2E17 to Dillon Cove     


**Data Contact:** [Matt Johnson](mailto:Matt.Johnson@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# 1997-2020 data (for comparison)
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/deer-creek/data-raw/Deer Creek SRCS snorkel survey counts by section 1997-2020.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "deer_creek_holding_1997_raw.xlsx",
               Overwrite = TRUE)

# 1997-2023 data
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/deer-creek/data-raw/Deer Creek Spring-Run Snorkel Counts by Section 1997-2023.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/deer-creek/deer_creek_snorkel_1997-2023_raw.xlsx")
```
### 1997-2020 data
```{r}
raw_data = readxl::read_excel('deer_creek_holding_1997_raw.xlsx', 
                              col_types = c("text", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric",
                                            "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric",
                                            "numeric", "numeric", "numeric", "numeric", "numeric"))
glimpse(raw_data)
```

### 1997-2023 data
```{r}
raw_data_2023 <- readxl::read_xlsx(path = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/deer-creek/deer_creek_snorkel_1997-2023_raw.xlsx",
                                   sheet = "Deer SRCS Snorkel 1997-2022",
                                   skip = 5) |> 
  glimpse()
```
## Data Transformations

### Older dataset
```{r}
cleaner_data <- raw_data |> 
  slice(1:(n()-1)) |> 
  pivot_longer(!'Survey Reach', names_to = "year", values_to = "count") |> 
  rename('location' = 'Survey Reach') |> 
  glimpse()
```

### Updated dataset (with 2022 and 2023)
```{r}
cleaner_data_2023 <- raw_data_2023 |> 
  rename(location = `Survey Reach`) |> 
  filter(!location %in% c(NA, "Total Count:")) |> 
  filter(!str_detect(location, "not surveyed")) |> 
  mutate(across(`1997`:`2023`, as.character)) |> 
  pivot_longer(!"location", names_to = "year", values_to = "count") |> 
  mutate(surveyed = ifelse(count %in% c("n/s", "*"), FALSE, TRUE),
         count = ifelse(count %in% c("n/s", "*"), NA_integer_, as.numeric(count)),
         year = as.numeric(year)) |> 
  glimpse()
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- cleaner_data_2023 |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_data_2023),
                          description = c("Survey reach",
                                          "Sample year",
                                          "Number of fish observed",
                                          "Whether or not that reach was surveyed in a given year"),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
```


## Explore Categorical Variables

### Variable `location`
```{r}
table(cleaner_data_2023$location)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data_2023$location))/nrow(cleaner_data_2023), 3)*100` % of values in the `location` column are NA.

## Explore Numeric Variables

### Variable `count`
```{r fig.height=7, fig.width=10}
#Find the most distinctive colours for visual
colourCount = length(unique(cleaner_data_2023$location))
getPalette = colorRampPalette(brewer.pal(12, "Paired"))

cleaner_data_2023 |>
  mutate(year = as.factor(year)) |> 
  ggplot(aes(x = year, y = count, fill = location))+
  scale_fill_manual(values = getPalette(colourCount))+
  geom_col()+
  theme_minimal()+
  labs(title = "Adult Holding Count By Year")+
  theme(axis.text.x = element_text(angle = 90))+
  ylim(0, 3000)
```
```{r}
cleaner_data_2023 |> 
  ggplot(aes(x = count, y = location))+
  geom_boxplot() +
  theme_minimal()+
  labs(title = "Adult Holding Count By Reach")
```

**Numeric Annual Summary of count From 1997 to 2023**
```{r}
cleaner_data_2023 |>
  group_by(year) |>
  summarise(count = sum(count, na.rm = T)) |>
  pull(count) |>
  summary()
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data_2023$count))/nrow(cleaner_data_2023), 5)*100` % of values in the `count` column are NA. 

## Next steps


## Add cleaned data back onto google cloud
```{r}
deer_adult_holding_survey_1997_to_2023 <- cleaner_data_2023 |> 
  glimpse()
```
```{r}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(deer_adult_holding_survey_1997_to_2023,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/deer-creek/data/deer_adult_holding_1997_to_2023.csv")
```