---
title: "mill 2021-2022 data"
output: html_document
date: "2023-05-09"
---


```{r, include = F}
library(dtplyr)
library(data.table)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(knitr)
library(hms)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
f <- function(input, output) write_csv(input, file = output)
```

```{r, include = F}
gcs_get_object(object_name = 
              "rst/mill-creek/data-raw/2021_2022/mill_deer_2021_data_Liz.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_2021_rst_deer_and_mill_liz.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = 
              "rst/mill-creek/data-raw/2021_2022/mill_deer_2021_data_av.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_2021_rst_deer_and_mill_ashley.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = 
              "rst/mill-creek/data-raw/2021_2022/mill_deer_2021_data_erin.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_2021_rst_deer_and_mill_erin.xlsx",
               overwrite = TRUE)
```

## Catch

TODO 
- standardize species and lifestage 
- add morts to new rows

```{r, include = F}
catch_liz <- readxl::read_excel("raw_2021_rst_deer_and_mill_liz.xlsx", sheet = "Catch") |> 
  mutate(is_plus_count = ifelse(is_plus_count == "T", T, F))
catch_erin <- readxl::read_excel("raw_2021_rst_deer_and_mill_erin.xlsx", sheet = "Catch") |> 
  mutate(fork_length = as.numeric(fork_length),
         weight = as.numeric(weight),
         count = as.numeric(count))
catch_ashley <- readxl::read_excel("raw_2021_rst_deer_and_mill_ashley.xlsx", sheet = "Catch") |> 
  mutate(is_plus_count = ifelse(is_plus_count == "T", T, F))
```

```{r}
catch_combined <- bind_rows(catch_liz, catch_erin, catch_ashley) |> 
  # standardize chinook - others can worry about later
  mutate(species = case_when(species == "CHISAL" ~ "chinook salmon",
                             T ~ species),
         count = ifelse(!is.na(fork_length) & is.na(count), 1, count)) |> 
  select(-pdf_page_number, -ad_clip) |> 
  # there was one NA count decided to filter out
  filter(!is.na(count)) |> 
  glimpse()
```

### simple checks: stream, date

```{r}
unique(catch_combined$stream)
range(catch_combined$date)
```

### species

```{r}
catch_combined |> 
  group_by(species) |> 
  summarize(count = sum(count))
```

### fork length

```{r}
catch_combined |> 
  filter(species == "chinook salmon") |> 
  group_by(fork_length) |> 
  summarize(count = sum(count)) |> 
  ggplot(aes(x = fork_length, y = count)) +
  geom_point()
```

```{r}
catch_combined |> 
  filter(species == "chinook salmon") |> 
  group_by(date) |> 
  summarize(fork_length = mean(fork_length, na.rm = T)) |> 
  ggplot(aes(x = date, y = fork_length)) +
  geom_point()
```
### lifestage

```{r}
catch_combined |> 
  group_by(lifestage) |> 
  summarize(count = sum(count))
```

##  Recapture

```{r}
recap_liz <- readxl::read_excel("raw_2021_rst_deer_and_mill_liz.xlsx", sheet = "Recapture") |> glimpse()
recap_erin <- readxl::read_excel("raw_2021_rst_deer_and_mill_erin.xlsx", sheet = "recapture_fish") |> glimpse()
recap_ashley <- readxl::read_excel("raw_2021_rst_deer_and_mill_ashley.xlsx", sheet = "recapture_fish") |> glimpse()

recap_combined <- bind_rows(recap_liz, recap_ashley) |> 
  mutate(lifestage = ifelse(lifestage == "p", "parr", lifestage))


```


## Trap

TODO
- Make sure all debris_gallons are in gallons and not tubs; remove debris_code
- Make sure water temp is all in the same unit or just remove and use the gage temp
- Find weather code and encode

```{r}
trap_liz <- readxl::read_excel("raw_2021_rst_deer_and_mill_liz.xlsx", sheet = "Trap Operations & Env") |> 
  rename(debris_gal = `debris_code...6`,
         debris_code = `debris_code...13`) |> 
  relocate(debris_code, .before = debris_gal)
trap_erin <- readxl::read_excel("raw_2021_rst_deer_and_mill_erin.xlsx", sheet = "Trap Operations & Env") |> 
  mutate(hours_fished = as.numeric(hours_fished),
         debris_gal = as.numeric(debris_gal),
         weather_code = as.numeric(weather_code),
         staff_gage_ft = as.numeric(staff_gage_ft),
         water_velocity_m_per_s = as.numeric(water_velocity_m_per_s),
         turbidity_ntu = as.numeric(turbidity_ntu),
         total_revs = as.numeric(total_revs),
         before_rpms = as.numeric(before_rpms),
         after_rpms = as.numeric(after_rpms)) |> 
  relocate(debris_code, .before = debris_gal) |> 
  rename(flow_cfs = flow_csf) |> 
  select(-page_number)
trap_ashley <- readxl::read_excel("raw_2021_rst_deer_and_mill_ashley.xlsx", sheet = "Trap Operations & Env") |> 
  relocate(debris_code, .before = debris_gal) |> 
  rename(flow_cfs = flow_csf)

trap_combined <- bind_rows(trap_liz, trap_erin, trap_ashley) |> 
  mutate(trap_location = "upper mill creek",
         flow_cfs = ifelse(is.na(flow_cfs), staff_gage_ft, flow_cfs)) |> 
  # all are NA
  select(-c(water_velocity_m_per_s, staff_gage_ft))


```


## Release

```{r}
release_liz <- readxl::read_excel("raw_2021_rst_deer_and_mill_liz.xlsx", sheet = "Release Summary") |> glimpse()
release_erin <- readxl::read_excel("raw_2021_rst_deer_and_mill_erin.xlsx", sheet = "Efficiency summary") |> glimpse()
release_ashley <- readxl::read_excel("raw_2021_rst_deer_and_mill_ashley.xlsx", sheet = "Release Summary") |> glimpse()

release_combined <- release_erin


```

## Release site location

```{r}
release_site <- readxl::read_excel("raw_2021_rst_deer_and_mill_erin.xlsx", sheet = "release site location") |> glimpse()


```

## Upload to google cloud

```{r}
gcs_upload(catch_combined,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2021_catch.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(recap_combined,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2021_recaptures.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(trap_combined,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2021_trap.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(release_combined,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2021_release.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(release_site,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2021_release_site.csv",
           predefinedAcl = "bucketLevel")
```
