---
title: "mill_2022_2023_qc"
output: html_document
date: "2023-10-10"
---

```{r, include = F}
library(dtplyr)
library(data.table)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(knitr)
library(hms)

root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
f <- function(input, output) write_csv(input, file = output)
```

```{r, include = F}
gcs_get_object(object_name = 
              "rst/mill-creek/data-raw/2022_2023/raw_2022_rst_mill_liz.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_2022_rst_mill_liz.xlsx",
               overwrite = TRUE)
```

## Catch

TODO 
- clarify non-standard species codes
- clarify the date recorded as 12-23-2023
- standardize lifestage codes vs. text
- add morts to new rows

```{r, include = F}
catch_liz <- readxl::read_excel("raw_2022_rst_mill_liz.xlsx", sheet = "Catch") |> 
  mutate(is_plus_count = ifelse(is_plus_count == "T", T, F))
```

```{r}
catch_clean <- catch_liz |> 
  # standardize chinook - others can worry about later
  mutate(species = ifelse(species == "chisal", "chinook salmon", species),
         count = ifelse(!is.na(fork_length) & is.na(count), 1, count),
         lifestage = case_when(lifestage == "s" ~ "smolt",
                               lifestage == "sp" ~ "silvery parr",
                               lifestage == "sf" ~ "sac fry",
                               lifestage == "f" ~ "fry",
                               lifestage == "p" ~ "parr",
                               TRUE ~ lifestage)) |> 
  filter(date != as.Date("2023-12-13")) |> # TODO find out what this date actually is
  select(-c(ad_clip)) |> 
  glimpse()
```

### simple checks: stream, date

```{r}
unique(catch_clean$stream)
range(catch_clean$date)
```

### species

```{r}
catch_clean |> 
  group_by(species) |> 
  summarize(count = sum(count))
```

### fork length

```{r}
catch_clean |> 
  filter(species == "chinook salmon") |> 
  group_by(fork_length) |> 
  summarize(count = sum(count)) |> 
  ggplot(aes(x = fork_length, y = count)) +
  geom_point()
```

```{r}
catch_clean |> 
  filter(species == "chinook salmon") |> 
  group_by(date) |> 
  summarize(fork_length = mean(fork_length, na.rm = T)) |> 
  ggplot(aes(x = date, y = fork_length)) +
  geom_point()
```
### lifestage

```{r}
catch_clean |> 
  group_by(lifestage) |> 
  summarize(count = sum(count))
```

##  Recapture

```{r}
recap_liz <- readxl::read_excel("raw_2022_rst_mill_liz.xlsx", sheet = "Recapture") |> 
  glimpse()

recap_clean <- recap_liz |> 
  mutate(lifestage = case_when(lifestage == "p" ~ "parr",
                               lifestage == "s" ~ "smolt",
                               TRUE ~ lifestage))

```


## Trap

TODO
- Make sure all debris_gallons are in gallons and not tubs; remove debris_code
- Make sure water temp is all in the same unit or just remove and use the gage temp
- Find weather code and encode
- Fill in flow_cfs with staff_gage (in PDFs, often written in the staff gage area but with cfs added in pencil)

```{r}
trap_liz <- readxl::read_excel("raw_2022_rst_mill_liz.xlsx", sheet = "Trap Operations & Env") |> 
  mutate(hours_fished = as.numeric(hours_fished),
         debris_gal = ifelse(is.na(debris_gal), debris_code, debris_gal),
         flow_cfs = ifelse(is.na(flow_cfs), staff_gage_ft, flow_cfs)) |> 
  select(-c(debris_code, water_velocity_m_per_s, turbidity_ntu))

```


## Release

No release sheet.

## Release site location

No release site sheet.

## Upload to google cloud

```{r}
gcs_upload(catch_clean,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2022_catch.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(recap_clean,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2022_recaptures.csv",
           predefinedAcl = "bucketLevel")

gcs_upload(trap_liz,
           object_function = f,
           type = "csv",
           name = "rst/mill-creek/data/mill_2022_trap.csv",
           predefinedAcl = "bucketLevel")
```