---
title: "Feather River Seine Data 2015 - 2021 QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River Seine and Snorkel Data 

## Description of Monitoring Data

Feather River Seine data 2015 - 2021. All new data will be stored in this format in the current access database. 

**Timeframe:** 2015 - 2021

**Seine Season:** ? TODO 

**Completeness of Record throughout timeframe:** 

Sampling occurred every year between 2015 - 2021

**Sampling Location:** Feather River

**Data Contact:** [Casey Campos](mailto:Casey.Campos@water.ca.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/all_fields_seine_2008-2014.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_seine_2008-2014.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/all_fields_seine_2014-2021.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_seine_2015-2021.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data: 
```{r, warning=FALSE}
# read in data to clean 
raw_seine_2015 <- read_xlsx("raw_seine_2015-2021.xlsx") %>% glimpse()
```

## Data transformations

```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns
cleaner_seine_data <- raw_seine_2015 %>% 
  janitor::clean_names() %>% 
  rename("bs_depth_half" = bs_depth_1_2, 
         "lifestage" = salmonid_life_stage,
         "latitude" = ) %>%
  filter(species %in% c("CHN", "CHNF", "CHNFT", "CHNI",
                        "CHNL", "CHNS", "CHNST", "CHNT","CHNW")) %>%
  mutate(date = as.Date(date),
         time = hms::as_hms(time),
         total_length = as.numeric(total_length),
         weight_g = as.numeric(weight_g),
         sample_id = as.character(sample_id),
         id = as.character(id),
         gear_code = as.character(gear_code),
         condition_code = as.character(condition_code),
         lifestage = as.character(lifestage),
         substrate_1 = as.character(substrate_1), # Should I change these to the actual cover type ex: fine_substrate T/F values?
         substrate_2 = as.character(substrate_2),
         substrate_3 = as.character(substrate_3),
         objectid = as.character(objectid),
         run = case_when(species %in% c("CHNF", "CHNFT") ~ "fall",
                         species %in% c("CHNS", "CHNST") ~ "spring",
                         species == "CHNW" ~ "winter",
                         species == "CHNI" ~ "inland", 
                         species %in% c("CHN", "CHNT") ~ "unknown"),
         tagged = if_else(grepl('T$', species), TRUE, FALSE)) %>% 
  select(-catch_tbl_comments, -dissolved_oxygen, -ec, -gear_height, -substrate_4, -substrate_5, -cover_6,
         -bs_velocity_full, -bs_velocity_1_2, -bs_depth_full, -species) %>% # Remove because all values are NA, species redundant
  glimpse()

unique(cleaner_seine_data$flow)
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables
cleaner_seine_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `count`

**Plotting Count over Period of Record**

```{r}
cleaner_seine_data %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = count)) + 
  geom_col() + 
  facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Seine Count All Runs", 
       x = "Date")  
```

```{r}
cleaner_seine_data %>% group_by(date) %>%
  mutate(daily_count_upstream = sum(count)) %>%
  mutate(year = as.factor(year(date))) %>% 
  ungroup() %>%
  ggplot(aes(x = year, y = daily_count_upstream)) + 
  geom_boxplot() + 
  theme_minimal() +
  theme(text = element_text(size = 23)) + 
  labs(title = "Daily Count of Seine Catch Sumarized by Year All Runs") 
```
```{r}
cleaner_seine_data  %>%
  mutate(year = as.factor(year(date))) %>%
  filter(run %in% c("fall", "spring", "winter")) %>% # Filter to only show runs that have more than one data point and are not NA/Unknown
  ggplot(aes(x = year, y = count)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Yearly Fish Counts by Run",
       y = "Total fish count") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_grid(~run)
```
Not a lot of spring run fish caught in comparison to fall run fish caught. 
**Numeric Summary of Count over Period of Record**

```{r}
# daily numeric summary 
cleaner_seine_data %>% group_by(date) %>%
  summarise(daily_count = sum(count, na.rm = T)) %>%
  pull(daily_count) %>%
  summary()

# daily numeric summary SR
cleaner_seine_data %>% 
  filter(run == "spring") %>%
  group_by(date) %>%
  summarise(daily_count = sum(count, na.rm = T)) %>%
  pull(daily_count) %>%
  summary()
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$count))/nrow(cleaner_seine_data), 3) * 100` % of values in the `count` column are NA. However, there are clearly gaps in data. 



### Variable: `fork_length`

**Plotting fork_length over Period of Record**

```{r}

cleaner_seine_data %>%  
  ggplot(aes(x = fork_length)) + 
  geom_histogram(breaks=seq(0, 200, by=2)) + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  theme_minimal() +
  labs(title = "Fork length distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
lifestage_code <- c(1, 2, 3, 4, 5, 6, 7, 8)
names(lifestage_code)  <- c("yolk-sac fry", "fry", "parr", "silvery parr", "smolt", "adult", "not provided", NA)

cleaner_seine_data %>% 
  mutate(year = as.factor(year(date)),
         lifestage_name = names(lifestage_code[as.numeric(lifestage)])) %>%
  ggplot(aes(x = fork_length, y = lifestage_name)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Fork length summarized by lifestage") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of fork_length over Period of Record**

```{r}
summary(cleaner_seine_data$fork_length)
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$fork_length))/nrow(cleaner_seine_data), 3) * 100` % of values in the `fork_length` column are NA. 

### Variable: `total_length`

**Plotting total_length over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of total_length over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

### Variable: `weight_g`

**Plotting weight_g over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of weight_g over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

### Variable: `temperature`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values


### Variable: `bs_start_length`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

### Variable: `bs_distance_out`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

### Variable: `bs_depth_half`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values


### Variables: `utm_northing`, `utm_easting` 

These are the latitude and longitude columns. 
**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values



## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
# Filter clean data to show only categorical variables
```


### Variable: `[name]`
```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Summary of identified issues

* List things that are funcky/bothering us but that we don't feel like should be changed without more investigation

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```