---
title: "Feather River Seine Data 2008 - 2014 QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River Seine and Snorkel Data 

## Description of Monitoring Data

Feather River Seine data 2008 - 2014. 

**Timeframe:** 2008 - 2014

**Seine Season:** Appears to be Feb - August (but not consistant across years sampled)

**Completeness of Record throughout timeframe:** 

No sampling occurred in the following years: 

* 2010 
* 2011
* 2012

**Sampling Location:** Feather River

**Data Contact:** [Casey Campos](mailto:Casey.Campos@water.ca.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/all_fields_seine_2008-2014.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_seine_2008-2014.xlsx",
               overwrite = TRUE)

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r, warning=FALSE}
# read in data to clean 
raw_seine_2008 <- read_xlsx("raw_seine_2008-2014.xlsx") %>% glimpse

```

## Data transformations

```{r}
cleaner_seine_data <- raw_seine_2008 %>% 
  janitor::clean_names() %>% 
  rename("bs_depth_half" = bs_depth_1_2, 
         "lifestage" = salmonid_life_stage,
         "stream_features" = rpg_ru,
         "condition" = condition_code,
         "weather" = weather_code,
         "gear" = gear_code) %>%
  filter(species %in% c("CHN", "CHNF", "CHNFT", "CHNI",
                        "CHNL", "CHNS", "CHNST", "CHNT","CHNW")) %>%
  mutate(date = as.Date(date),
         time = hms::as_hms(time),
         weight_g = as.numeric(weight_g),
         sample_id = as.character(sample_id),
         id = as.character(id),
         condition = as.character(condition),
         gear = as.character(gear),
         lifestage = as.character(lifestage),
         substrate_1 = as.character(substrate_1), # Should I change these to the actual cover type ex: fine_substrate T/F values?
         substrate_2 = as.character(substrate_2),
         substrate_3 = as.character(substrate_3),
         run = case_when(species %in% c("CHNF", "CHNFT") ~ "fall",
                         species %in% c("CHNS", "CHNST") ~ "spring",
                         species == "CHNW" ~ "winter",
                         species == "CHNI" ~ "inland", 
                         species %in% c("CHN", "CHNT") ~ "unknown"),
         tagged = if_else(grepl('T$', species), TRUE, FALSE)) %>% 
  select(-dissolved_oxygen, -gear_height, -weight_g, # Remove because all values are NA
         -bs_velocity_full, -bs_velocity_1_2, - efbs_velocity_top, -efbs_velocity_bottom, 
         -species, # Remove species because we filtered for chinook so it is redundant
         -total_length, # Remove total length because only one value, and fork_length already measures size
         -site_id, -l_id, -objectid) %>% # Remove site_id/l_id/objectid because already joined with attributes relevant to these id (have site_name, lat/long)
  glimpse()
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_seine_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `count`

**Plotting Count over Period of Record**

```{r}
cleaner_seine_data %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = count)) + 
  geom_col() + 
  facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 20),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Seine Count All Runs", 
       x = "Date")  
```

```{r}
cleaner_seine_data %>% group_by(date) %>%
  mutate(daily_count_upstream = sum(count)) %>%
  mutate(year = as.factor(year(date))) %>% 
  ungroup() %>%
  ggplot(aes(x = year, y = daily_count_upstream)) + 
  geom_boxplot() + 
  theme_minimal() +
  theme(text = element_text(size = 23)) + 
  labs(title = "Daily Count of Seine Catch Sumarized by Year All Runs") 
```
```{r}
cleaner_seine_data  %>%
  mutate(year = as.factor(year(date))) %>%
  filter(run %in% c("fall", "spring", "winter")) %>% # Filter to only show runs that have more than one data point and are not NA/Unknown
  ggplot(aes(x = year, y = count)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Yearly Fish Counts by Run",
       y = "Total fish count") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_grid(~run)
```

Not a lot of spring run fish caught in comparison to fall run fish caught. 

**Numeric Summary of Count over Period of Record**

```{r}
# daily numeric summary 
cleaner_seine_data %>% group_by(date) %>%
  summarise(daily_count = sum(count, na.rm = T)) %>%
  pull(daily_count) %>%
  summary()

# daily numeric summary SR
cleaner_seine_data %>% 
  filter(run == "spring") %>%
  group_by(date) %>%
  summarise(daily_count = sum(count, na.rm = T)) %>%
  pull(daily_count) %>%
  summary()
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$count))/nrow(cleaner_seine_data), 3) * 100` % of values in the `count` column are NA. However, there are clearly gaps in data. 


### Variable: `fork_length`

**Plotting fork_length over Period of Record**

```{r}

cleaner_seine_data %>%  
  ggplot(aes(x = fork_length)) + 
  geom_histogram(breaks=seq(0, 200, by=2)) + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  theme_minimal() +
  labs(title = "Fork length distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
lifestage_code <- c(1, 2, 3, 4, 5, 6, 7, 8)
names(lifestage_code)  <- c("yolk-sac fry", "fry", "parr", "silvery parr", "smolt", "adult", "not provided", NA)

cleaner_seine_data %>% 
  mutate(year = as.factor(year(date)),
         lifestage_name = names(lifestage_code[as.numeric(lifestage)])) %>%
  ggplot(aes(x = fork_length, y = lifestage_name)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Fork length summarized by lifestage") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of fork_length over Period of Record**

```{r}
summary(cleaner_seine_data$fork_length)
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$fork_length))/nrow(cleaner_seine_data), 3) * 100` % of values in the `fork_length` column are NA. 

### Variable: `temperature`


**Plotting temperature over Period of Record**

Daily average water temperature measures appear to be lower in Feb - April and then increase May - August. They appear to typically range from 9 - 23 degrees C.
```{r}
cleaner_seine_data %>% 
  group_by(date = as.Date(date)) %>%
  mutate(avg_temp = mean(temperature)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  geom_point(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Daily Water Temperature (colored by year)",
       y = "Average daily temp", 
       x = "Date")  
```

**Plotting Distribution of temperature measures**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = temperature)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Temperature distribution (celcius)") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of flow over Period of Record**

```{r}
# Table with summary statistics
summary(cleaner_seine_data$temperature)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$temperature))/nrow(cleaner_seine_data), 3) * 100` % of values in the `temperature` column are NA. 


### Variable: `flow`

**Plotting flow over Period of Record**

```{r}
cleaner_seine_data %>% 
  group_by(date = as.Date(date)) %>%
  mutate(avg_flow = mean(flow)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_flow, color = year)) + 
  geom_point(alpha = .75) + 
  geom_line(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Daily Average Flow (colored by year)",
       y = "Flow (CFS)", 
       x = "Date")  
```

Flow measurements are not taken frequently. It looks like about ~ 1 - 8 days out of the year. 

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = flow)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Flow distribution (cfs)") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of temperature over Period of Record**

```{r}
summary(cleaner_seine_data$flow)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$flow))/nrow(cleaner_seine_data), 3) * 100` % of values in the `flow` column are NA. 
 
### Variable: `temperature`


**Plotting temperature over Period of Record**

Daily average water temperature measures appear to be lower in Feb - April and then increase May - August. They appear to typically range from 9 - 23 degrees C.
```{r}
cleaner_seine_data %>% 
  group_by(date = as.Date(date)) %>%
  mutate(avg_temp = mean(temperature)) %>%
  ungroup() %>%
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  geom_point(alpha = .25) + 
  # facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Daily Water Temperature (colored by year)",
       y = "Average daily temp", 
       x = "Date")  
```

**Plotting Distribution of temperature measures**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = temperature)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Temperature distribution (celcius)") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of temperature over Period of Record**

```{r}
# Table with summary statistics
summary(cleaner_seine_data$temperature)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$temperature))/nrow(cleaner_seine_data), 3) * 100` % of values in the `temperature` column are NA. 


### Variable: `efbs_length`

**Plotting efbs_length distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = efbs_length)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "EFBS length") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of efbs_length over Period of Record**

```{r}
summary(cleaner_seine_data$efbs_length)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$efbs_length))/nrow(cleaner_seine_data), 3) * 100` % of values in the `efbs_length` column are NA. 

### Variable: `efbs_width`

**Plotting efbs_width distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = efbs_width)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "EFBS width") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of efbs_width over Period of Record**

```{r}
summary(cleaner_seine_data$efbs_width)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$efbs_width))/nrow(cleaner_seine_data), 3) * 100` % of values in the `efbs_width` column are NA. 

### Variable: `efbs_depth_top`

**Plotting efbs_depth_top distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = efbs_depth_top)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "EFBS Depth Top") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of efbs_depth_top over Period of Record**

```{r}
summary(cleaner_seine_data$flow)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$efbs_depth_top))/nrow(cleaner_seine_data), 3) * 100` % of values in the `efbs_depth_top` column are NA. 

### Variable: `efbs_depth_bottom`

**Plotting efbs_depth_bottom distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = efbs_depth_bottom)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "EFBS depth bottom") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of efbs_depth_bottom over Period of Record**

```{r}
summary(cleaner_seine_data$efbs_depth_bottom)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$efbs_depth_bottom))/nrow(cleaner_seine_data), 3) * 100` % of values in the `efbs_depth_bottom` column are NA. 

### Variable: `bs_start_length`

**Plotting bs_start_length distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = bs_start_length)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Beach Seine Start length") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of bs_start_length over Period of Record**

```{r}
summary(cleaner_seine_data$bs_start_length)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$bs_start_length))/nrow(cleaner_seine_data), 3) * 100` % of values in the `bs_start_length` column are NA. 

### Variable: `bs_close_width`

**Plotting bs_close_width distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = bs_close_width)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Beach Seine Close Width") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of bs_close_width over Period of Record**

```{r}
summary(cleaner_seine_data$bs_close_width)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$bs_close_width))/nrow(cleaner_seine_data), 3) * 100` % of values in the `bs_close_width` column are NA. 

### Variable: `bs_distance_out`

**Plotting bs_distance_out distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = bs_distance_out)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Beach Seine distance out") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of bs_distance_out over Period of Record**

```{r}
summary(cleaner_seine_data$bs_distance_out)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$bs_distance_out))/nrow(cleaner_seine_data), 3) * 100` % of values in the `bs_distance_out` column are NA. 

### Variable: `bs_depth_half`

**Plotting bs_depth_half distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = bs_depth_half)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Beach Seine half depth") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of bs_depth_half over Period of Record**

```{r}
summary(cleaner_seine_data$bs_depth_half)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$bs_depth_half))/nrow(cleaner_seine_data), 3) * 100` % of values in the `bs_depth_half` column are NA. 

### Variable: `bs_depth_full`

**Plotting bs_depth_full distribution**

```{r}
cleaner_seine_data %>%  
  ggplot(aes(x = bs_depth_full)) + 
  geom_histogram() + 
  scale_x_continuous() +
  theme_minimal() +
  labs(title = "Beach Seine full depth") +  
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of bs_depth_full over Period of Record**

```{r}
summary(cleaner_seine_data$bs_depth_full)
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$bs_depth_full))/nrow(cleaner_seine_data), 3) * 100` % of values in the `bs_depth_full` column are NA. 

### Variable: `x_coord`, `y_coord`

These are the location columns. TODO

```{r}
# Check that they are in consistent format
# Map
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$x_coord))/nrow(cleaner_seine_data), 3) * 100` % of values in the `x_coord` column are NA. 
* `r round(sum(is.na(cleaner_seine_data$y_coord))/nrow(cleaner_seine_data), 3) * 100` % of values in the `y_coord` column are NA. 


## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
cleaner_seine_data %>% select_if(is.character) %>% colnames()
```

### Variable: `sample_id`, `id`
```{r}
length(unique(cleaner_seine_data$id)) == nrow(cleaner_seine_data)
```

Each id is unique as anticipated. 

Each `sample_id` is not unique because multiple catches can occur in the same beach seine sample. 

```{r}
length(unique(cleaner_seine_data$sample_id)) == nrow(cleaner_seine_data)
```

**NA and Unknown Values**

No na values
* `r round(sum(is.na(cleaner_seine_data$sample_id))/nrow(cleaner_seine_data), 3) * 100` % of values in the `sample_id` column are NA.
* `r round(sum(is.na(cleaner_seine_data$id))/nrow(cleaner_seine_data), 3) * 100` % of values in the `id` column are NA.

### Variable: `lifestage`
```{r}
table(cleaner_seine_data$lifestage)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
lifestage_code <- c(1, 2, 3, 4, 5, 6, 7)
names(lifestage_code)  <- c("yolk-sac fry", "fry", "parr", "silvery parr", "smolt", "adult", "not provided")

cleaner_seine_data$lifestage <- names(lifestage_code[as.numeric(cleaner_seine_data$lifestage)])
cleaner_seine_data$lifestage <- ifelse(cleaner_seine_data$lifestage == "not provided", NA, cleaner_seine_data$lifestage) # Change not provided to NA

table(cleaner_seine_data$lifestage)
```

**Create lookup rda for [variable] encoding:** 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$lifestage))/nrow(cleaner_seine_data), 3) * 100` % of values in the `lifestage` column are NA.

### Variable: `gear`
```{r}
table(cleaner_seine_data$gear)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. Update gear codes with gear names:

* SEIN25 - Gear description 25 foot beach seine with bag
* SEIN50 - Gear description 25 foot beach seine with bag

```{r}
cleaner_seine_data$gear <- case_when(cleaner_seine_data$gear == "1" ~ "SEIN25",
                                          cleaner_seine_data$gear == "3" ~ "SEIN50")

table(cleaner_seine_data$gear)
```

**NA and Unknown Values**


* `r round(sum(is.na(cleaner_seine_data$gear))/nrow(cleaner_seine_data), 3) * 100` % of values in the `gear` column are NA.

### Variable: `condition`
```{r}
table(cleaner_seine_data$condition)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

Descriptions: 

* Good - The haul went smoothly without getting hung up on anything or net rolling.
* Fair - The haul went relatively smoothly, but you had to stop once or twice to get it unsnagged or take big rocks out of the net.
* Poor - The haul didn’t go well; you had to stop numerous times while the poles were still a good distance from the shore to free the seine from snags or big rocks.

All other values (0, 4) corresond to NA

```{r}
cleaner_seine_data$condition <- case_when(cleaner_seine_data$condition == "1" ~ "good",
                                               cleaner_seine_data$condition == "2" ~ "fair",
                                               cleaner_seine_data$condition == "3" ~ "poor")


table(cleaner_seine_data$condition)
```

**NA and Unknown Values**


* `r round(sum(is.na(cleaner_seine_data$condition))/nrow(cleaner_seine_data), 3) * 100` % of values in the `condition` column are NA.

### Variable: `weather`
```{r}
table(cleaner_seine_data$weather)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_seine_data$weather <- case_when(cleaner_seine_data$weather == "CLD" ~ "overcast",
                                             cleaner_seine_data$weather == "CLR" ~ "sunny",
                                             cleaner_seine_data$weather == "FOG" ~ "foggy",
                                             cleaner_seine_data$weather == "RAN" ~ "precipitation")


table(cleaner_seine_data$weather)
```

**NA and Unknown Values**


* `r round(sum(is.na(cleaner_seine_data$weather))/nrow(cleaner_seine_data), 3) * 100` % of values in the `weather_code` column are NA.

### Variable: `substrate_1`, `substrate_2`, `substrate_3`

TODO fix substrate stuff 

| SubstrateCode |	Substrate |
| -------------- | ---------------------- |
| 1 |	Fine - small gravel (0-50mm) (0-2in.) |
| 2 |	Small - medium gravel (50-150mm) (2-6in.) |
| 3 |	Medium - large cobble (150-300mm) (6-12in.) |
| 4 |	Boulder (>300mm) (>12in.) |
| 5 |	Pavement (Boat Ramp) |

```{r}
table(cleaner_seine_data$substrate_1)
table(cleaner_seine_data$substrate_2)
table(cleaner_seine_data$substrate_3)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$substrate_1))/nrow(cleaner_seine_data), 3) * 100` % of values in the `substrate_1` column are NA.
* `r round(sum(is.na(cleaner_seine_data$substrate_2))/nrow(cleaner_seine_data), 3) * 100` % of values in the `substrate_2` column are NA.
* `r round(sum(is.na(cleaner_seine_data$substrate_3))/nrow(cleaner_seine_data), 3) * 100` % of values in the `substrate_3` column are NA.


### Variable: `cover_1`, `cover_2`, `cover_3`

TODO fix cover stuff 

```{r}
table(cleaner_seine_data$cover_1)
table(cleaner_seine_data$cover_2)
table(cleaner_seine_data$cover_3)

```

**Create lookup rda for cover encoding:** 

```{r}
feather_seine_cover <- c("A", "B", "C", "D", "E", "BD", "DE" )
names(feather_seine_cover) <- c("no apparent cover",
"small instream objects",
"large instream objects",
"overhead objects",
"submerged aquatic vegetation",
"small instream objects and overhead objects",
"overhead objects and submerged aquatic vegetation")

tibble(code = feather_seine_cover, 
       definitions = names(feather_seine_cover))

# write_rds(feather_seine_cover, "../../../data/feather_seine_cover.rds")
```
**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$cover_1))/nrow(cleaner_seine_data), 3) * 100` % of values in the `cover_1` column are NA.
* `r round(sum(is.na(cleaner_seine_data$cover_2))/nrow(cleaner_seine_data), 3) * 100` % of values in the `cover_2` column are NA.
* `r round(sum(is.na(cleaner_seine_data$cover_3))/nrow(cleaner_seine_data), 3) * 100` % of values in the `cover_3` column are NA.

### Variable: `stream_features`
```{r}
table(cleaner_seine_data$stream_features)
```

**Create lookup rda for stream_features encoding:** 
```{r}
cleaner_seine_data$stream_features <- ifelse(cleaner_seine_data$stream_features == "run", "RU", cleaner_seine_data$stream_features)
feather_seine_stream_features <- c("G", "GP", "GR", "GRRU", "GRU", "P", "PRU", "R", "RRU", "RU")
names(feather_seine_stream_features) <- c( "Glide", "Glide & Pool", "Glide & Riffle", "Glide & Riffle & Run", "Glide & Run", "Pool",
                                           "Pool & Run", "Riffle", "Riffle & Run", "Run")

tibble(code = feather_seine_stream_features, 
       definitions = names(feather_seine_stream_features))

# write_rds(feather_seine_stream_features, "../../../data/feather_seine_stream_features.rds")
```
**NA and Unknown Values**


* `r round(sum(is.na(cleaner_seine_data$stream_features))/nrow(cleaner_seine_data), 3) * 100` % of values in the `stream_features` column are NA.


### Variable: `site_name`
```{r}
table(cleaner_seine_data$site_name)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
format_site_name <- function(string) {
  clean <- str_replace_all(string, "1/2", "half") %>%
    str_replace_all("1/4", "quarter") %>%
    str_replace_all("S.C.", "SC") %>%
    str_replace_all("'", "") %>%
    str_replace_all("G-95", "G95") %>% 
    str_replace_all("[^[:alnum:]]", " ") %>% 
    trimws() %>% 
    stringr::str_squish() %>%
    stringr::str_to_title()
}

cleaner_seine_data$site_name <- format_site_name(cleaner_seine_data$site_name)
table(cleaner_seine_data$site_name)
```


**NA and Unknown Values**


* `r round(sum(is.na(cleaner_seine_data$site_name))/nrow(cleaner_seine_data), 3) * 100` % of values in the `site_name` column are NA.

### Variable: `gear_type`

There are two distinct gear types: 

* SEIN: standard seine 
* EF_SE: backpack shock into any seine 

```{r}
table(cleaner_seine_data$gear_type)
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_seine_data$seine))/nrow(cleaner_seine_data), 3) * 100` % of values in the `seine` column are NA.

### Variable: `run`
```{r}
table(cleaner_seine_data$run)
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_seine_data$run <- ifelse(cleaner_seine_data$run  == "unknown", NA, cleaner_seine_data$run)
table(cleaner_seine_data$run)
```

**NA and Unknown Values**


* `r round(sum(is.na(cleaner_seine_data$run))/nrow(cleaner_seine_data), 3) * 100` % of values in the `run` column are NA.

## Summary of identified issues

* Not a lot of days sampled per year
* Only 10 rows that include spring data. 

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```
