---
title: "Feather Carcass QC 2002"
author: "Inigo Peng"
date: '2022-07-25'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data
```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```
```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2002.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2002.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2002.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopChannel_1 <- sqlFetch(con, "ChopChannelTBL")
  ChopChannel_2<- sqlFetch(con, "ChopChannelTBL2")
  ChopRecov_1 <- sqlFetch(con, "ChopRecovTBL")
  ChopRecov_2 <- sqlFetch(con, "ChopRecovTBL2")
  ChopHeader_1 <- sqlFetch(con, "ChopHeaderTBL")
  ChopHeader_2 <- sqlFetch(con, "ChopHeaderTBL2")
  cwt_1 <- sqlFetch(con, "CWTTagTBL")  |> 
    select(-WeekNum)
  cwt_2 <- sqlFetch(con, "CWTTagTBL1") |> 
    mutate(SampNum = as.character(SampNum))
  cwt_header_1 <- sqlFetch(con, "CWTHeaderTBL")
  cwt_header_2 <- sqlFetch(con, "CWTHeaderTBL1")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  ChopChannel_1 <- mdb.get(filepath, "ChopChannelTBL")
  ChopChannel_2 <- mdb.get(filepath, "ChopChannelTBL2")
  ChopRecov_1 <- mdb.get(filepath, "ChopRecovTBL")
  ChopRecov_2 <- mdb.get(filepath, "ChopRecovTBL2")
  ChopHeader_1 <- mdb.get(filepath, "ChopHeaderTBL")
  ChopHeader_2 <- mdb.get(filepath, "ChopHeaderTBL2")
  cwt_1 <- mdb.get(filepath, "CWTTagTBL") |> 
    select(-WeekNum)
  cwt_2 <- mdb.get(filepath, "CWTTagTBL1") |> 
    mutate(SampNum = as.character(SampNum))
  cwt_header_1 <- mdb.get(filepath, "CWTHeaderTBL")
  cwt_header_2 <- mdb.get(filepath, "CWTHeaderTBL1")
  # Steelhead Data table contains biological sampling data (n = 7) for steelhead
  # TagColLU contains metadata about tag colors

  detach(package:Hmisc)
}

# combine some tables before writing
ChopChannel_raw <- bind_rows(ChopChannel_1, ChopChannel_2)
ChopRecov_raw <- bind_rows(ChopRecov_1, ChopRecov_2)
ChopHeader_raw <- bind_rows(ChopHeader_1, ChopHeader_2)
cwt_raw <- bind_rows(cwt_1, cwt_2)
cwt_header_raw <-  bind_rows(cwt_header_1, cwt_header_2)

write_csv(ChopChannel_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopChannel_2002.csv"))
write_csv(ChopRecov_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2002.csv"))
write_csv(ChopHeader_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2002.csv"))
write_csv(cwt_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "cwt_2002.csv"))
write_csv(cwt_header_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2002.csv"))
```


## Raw Data Glimpse: {.tabset}

### ChopEnv_Raw
```{r}
ChopChannel_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                       "adult-holding-redd-and-carcass-surveys",
                                       "feather-river","ChopChannel_2002.csv")) |>
  glimpse()
```

### ChopRecov_raw
```{r}
ChopRecov_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                     "adult-holding-redd-and-carcass-surveys",
                                     "feather-river","ChopRecov_2002.csv")) |> glimpse()
```

### ChopHeader_Raw
```{r}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                      "adult-holding-redd-and-carcass-surveys",
                                      "feather-river","ChopHeader_2002.csv")) |>
  rename(HeaderID = headerID) |> 
  glimpse()
```

### cwt_raw
```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys",
                               "feather-river","cwt_2002.csv")) |> glimpse()
```

### CWT_Header_raw
```{r}
cwt_header_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                      "adult-holding-redd-and-carcass-surveys",
                                      "feather-river","CWTHeader_2002.csv")) |> glimpse()
```

## Data transformations: {.tabset}

### Counts

The `chop_channel` table contains carcass counts by sex, grilise and chop/tagged
```{r}
#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw |> 
                                 select(HeaderID, TagCol, Date, Time),
                               ChopChannel_raw,
                               by = "HeaderID") |> 
  clean_names() |> 
  rename("male_chop" = male,
         "female_chop" = fem,
         "grilse" = gril) |> glimpse()
```

### Recovery

The `chop_recovery` table contains recovered carcass counts by color

```{r}
# HeaderID connects ChopChannel to ChopHeader; ChanID joins ChopChannel to ChopRecov
chop_recovery_join <- full_join(chop_channel_join |> 
                                  select(date, time, sect, chan, min, chan_id, header_id,
                                         tag_col), ChopRecov_raw |> 
                                  rename(chan_id = ChanID), 
                                by = "chan_id") |> 
  clean_names() |> 
  glimpse()
# according to Access relationships, ChanID connects the header to the Recovery table
# chop_recovery_join <- full_join(ChopHeader_raw |> 
#                                   select(HeaderID, Date, Time),
#                                 ChopRecov_raw |> 
#                                   rename(HeaderID = ChanID))  |> # join on HeaderID <-> ChanID
#   clean_names()
```

### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopHeader_raw |> 
  clean_names()
```

### CWT

The `cwt` table contains coded wire tag information. 
```{r}
cwt <- full_join(cwt_raw |> clean_names(), cwt_header_raw |> clean_names()) |> 
  clean_names() |> 
  glimpse()
```

## Explore Numeric Variables: {.tabset}

#### Chop Channel Variable:`id`, `chan_id`, `sect`, `min`

```{r}
# Filter clean data to show only numeric variables 
chop_channel_join |> 
  select_if(is.numeric) |>
  colnames()
```

```{r}
summary(chop_channel_join$header_id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

* `r round(sum(is.na(chop_channel_join$header_id))/nrow(chop_channel_join), 3) * 100` % of values in the `id` column are NA.
* `r round(sum(is.na(chop_channel_join$chan_id))/nrow(chop_channel_join), 3) * 100` % of values in the `chan_id` column are NA.
* `r round(sum(is.na(chop_channel_join$sect))/nrow(chop_channel_join), 3) * 100` % of values in the `sect` column are NA.
* `r round(sum(is.na(chop_channel_join$min))/nrow(chop_channel_join), 3) * 100` % of values in the `min` column are NA.

### Chop Channel Variable: `female_chop`, `male_chop`,`grilise`, `tag_count`

**Numeric Summary of `female_chop`, `male_chop`,`unknown_chops`, `tag_count` over Period of Record**

```{r}
summary(chop_channel_join$female_chop)
```

```{r}
summary(chop_channel_join$male_chop)
```

```{r}
summary(chop_channel_join$grilse)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$male_chop))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$female_chop))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$grilse))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$male_chop))/nrow(chop_channel_join), 3) * 100` % of values in the `male_chop` column are NA.
* `r round(sum(is.na(chop_channel_join$female_chop))/nrow(chop_channel_join), 3) * 100` % of values in the `female_chop` column are NA.
* `r round(sum(is.na(chop_channel_join$grilse))/nrow(chop_channel_join), 3) * 100` % of values in the `unknown_chops` column are NA.

**Plotting total chops (male, female, grilse) over Period of Record**

```{r}
# Make whatever plot is appropriate 
chop_channel_join |> 
  group_by(date) |> 
  summarise(total = sum(male_chop, female_chop, grilse)) |> 
  ggplot(aes(x=date, y=total)) +
  geom_col()
```



**Plotting Total chops over Period of Record by sex**

```{r}
chop_channel_join |> glimpse()

total_chops <- chop_channel_join|>
  mutate(male_chop = ifelse(is.na(male_chop), 0, male_chop), # fill na
         female_chop = ifelse(is.na(female_chop), 0, female_chop),
         grilse_chop = ifelse(is.na(grilse), 0, grilse),
         total_chops = grilse_chop + male_chop + female_chop) |> 
  select(date, male_chop, female_chop, grilse_chop, total_chops)

# Case check: checking for duplicates
total_chops_test <- total_chops |> # get total chops
  filter(date == as_date("2002-09-20"))

# Check if NAs introduced in creating total chop columns
nrow(total_chops_test) == (sum(total_chops_test$total_chops == (total_chops_test$male_chop + total_chops_test$female_chop + total_chops_test$grilse_chops)))

# total number of chops for the day
total_chops_summary <- total_chops |>
  group_by(date) |>
  summarise(total_chops = sum(total_chops, na.rm = T),
            male_chops = sum(male_chop, na.rm = T),
            female_chops = sum(female_chop, na.rm = T),
            grilse_chops = sum(grilse_chop, na.rm = T))
```

```{r}
total_chops_summary |> 
  pivot_longer(cols = c(male_chops, female_chops, grilse_chops), names_to = "sex", values_to = "count") |> 
  mutate(proportions = (count / total_chops)) |> 
  ggplot(aes(x = date, y = proportions, fill = sex)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Sex", 
                    labels = c("Female", "Male", "Grilse")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```

This plot shows the proportion of chops based on sex and grilse for each day over the the period of 2002-09-03 to 2002-12-19. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. 

```{r}
summary(chop_channel_join$tag_count)
```

```{r}
round(sum(is.na(chop_channel_join$tag_count))/nrow(chop_channel_join), 3) * 100
```

* `r round(sum(is.na(chop_channel_join$tag_count))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_count` column are NA.

**Plotting tag_count over Period of Record**

```{r}
total_tags_summary <- chop_channel_join |>
  group_by(date) |>
  summarise(total_tags = sum(tag_count, na.rm = T))

total_tags_summary |> 
  ggplot(aes(x = date, y = total_tags)) +
  geom_col() +
  theme_minimal() +
  labs(y = "Daily Total Tags", x = "Date")

```

**Comparing tags and chops**

```{r}
total_carcass <- full_join(total_chops_summary, total_tags_summary,
                           by = "date") |> 
  select(date, total_chops, total_tags) |> 
  pivot_longer(c(total_chops, total_tags), names_to = "mark")

proportion_tags <- total_carcass |> 
  pivot_wider(names_from = mark, 
              values_from = value) |> 
  group_by(date) |> 
  mutate(total_collected = sum(total_chops, total_tags)) |> 
  summarise(proportion_tagged = total_tags/total_collected) |> 
  summary()

# Note: the tags recovered on a specific date are likely from a previous date (not necessarily
# linked with chops on that day)

ggplot(total_carcass, aes(x = date, y = value, fill = mark)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Date", y = "Count")

```
Tags are significantly less applied compared to chops.

### Chop Recovery Variable: `header_id`, `recov_id`, `recov_count`

```{r}
chop_recovery_join |> 
  select_if(is.numeric) |> 
  colnames()
```

```{r}
summary(chop_recovery_join$header_id)
```

```{r}
summary(chop_recovery_join$recov_id)
```

```{r}
summary(chop_recovery_join$recov_count)
```

**Plotting recov_count over Period of Record**

```{r}
chop_recovery_join |> 
  ggplot(aes(x = date, y = recov_count)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Date", y = "Recovered Chop Count")
```
This chart shows the recovered chops. The chart is not showing count values that do not have dates or times associated with them.

### Chop Header Variable: `header_id`
```{r}
chop_header |> 
  select_if(is.numeric) |> 
  colnames()
```
```{r}
summary(chop_header$header_id)
```

### CWT Variable: `tag_id`, `sect`, `fl`, `head_tag`, `header_id`, `week_num`
```{r}
cwt |> 
  select_if(is.numeric) |> 
  colnames()
```
```{r}
summary(cwt$tag_id)
```
```{r}
summary(cwt$sect)
```
```{r}
summary(cwt$fl)
```
* `r round(sum(is.na(cwt$tag_id))/nrow(cwt), 3) * 100` % of values in the `tag_id` column are NA.
* `r round(sum(is.na(cwt$sect))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.
* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
#Is this graph helpful?
unique(cwt$sex)
cwt_count <- cwt |> 
  mutate(count = 1) |>
  mutate(sex = case_when(sex == "nd"|is.na(sex) ~ "U",
                         TRUE ~ sex)) |> 
  pivot_wider(names_from = sex, values_from = count, values_fill = list(0)) |> 
  unnest() |> 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) |> glimpse()

total_cwt_summary <- cwt_count |> 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) |> 
  group_by(month(date)) |> 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary |> 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") |> 
  mutate(proportions = (count / total_cwt)) |> 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```
**Plotting fork length of each sex**

```{r}
cwt |> 
  mutate(sex = case_when(sex == "nd"|is.na(sex) ~ "Unknown",
                         TRUE ~ sex))|> 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```
## Explore Categorical variables: {.tabset}

### Chop Channel Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_channel_join |> 
  select_if(is.character) |>
  colnames()
```
```{r}
unique(chop_channel_join$tag_col)
```

```{r}
#Remove 5?
unique(chop_channel_join$chan)
```

```{r}
chop_channel_cleaner <- chop_channel_join |>
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time))

chop_channel_cleaner
```

### Chop Recovery Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_recovery_join |> 
  select_if(is.character) |> 
  colnames()
```

```{r}
unique(chop_recovery_join$recov_color)
```
```{r}
chop_recovery_cleaner <- chop_recovery_join |> 
  mutate(date = as_date(date)) |>
  mutate(recov_color = str_replace_all(recov_color, " ", "")) |> 
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time))

chop_recovery_cleaner
```

### Chop Header Clean Data

```{r}
chop_header |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$recorder)
```
```{r}
unique(chop_header$weather)
```


```{r}
chop_header_cleaner <- chop_header |>
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  mutate(crew = str_replace_all(crew, " ", ""),
         tag_col = str_replace_all(tag_col, " ", ""),
         recorder = str_replace_all(recorder, " ", ""),
         weather = case_when(weather == "ran" ~ "rain",
                             weather == "cld" ~ "cloudy",
                             TRUE ~ weather)) |> 
  select(-c(time))

chop_header_cleaner
```

### CWT Clean Data

```{r}
cwt |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(cwt$rel_loc)
```
```{r}
unique(cwt$spawn)
```
```{r}
unique(cwt$clip)
```
```{r}
unique(cwt$samp_type)

```
```{r}
unique(cwt$carc_stat)
```

```{r}
unique(cwt$sex)
```

```{r}
cwt_cleaner <- cwt |> 
  mutate_if(is.character, str_to_lower) |>
  mutate(crew = str_replace_all(crew, " ", ""),
         tag_col = str_replace_all(tag_col, " ", ""),
         sex = case_when(sex == "nd"|is.na(sex) ~ "uk",
                         TRUE ~ sex),
         rel_loc = ifelse(rel_loc == "no data", NA, rel_loc),  
         spawn = ifelse(spawn == "no data", NA, spawn),
         clip = case_when(clip == "no data" ~ NA_character_,
                          clip == "y" ~ "yes",
                          clip == "n" ~ "no",
                          TRUE ~ clip),
         samp_type = case_when(samp_type == "no data"| samp_type == "none" ~ NA_character_, 
                               TRUE ~ samp_type),
         carc_stat = ifelse(carc_stat == "no data", NA, carc_stat)) |> 
  glimpse()
```

## Comments

* CWT data has a many new columns - not sure what is needed
* Theres no Tag data tble 

## Data Dictionaries

### Channel

```{r}
percent_na <- chop_channel_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("ID",
                                          "Color of tag applied to carcass",
                                          "Date of survey",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped and male",
                                          "Carcass that were chopped and female",
                                          "Carcass that were grilse",
                                          "Carcass that were tagged"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
percent_na <- chop_recovery_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


recovery_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("Date of survey",
                                          "Sect",
                                          "Channel",
                                          "Min",
                                          "Channel ID",
                                          "Header ID",
                                          "Tag color",
                                          "Recovery ID",
                                          "Recovery Color",
                                          "Count of recovered tags"),
                          percent_na = round(percent_na$value*100))

kable(recovery_data_dictionary)
```

### Survey

```{r}
# 
percent_na <- chop_header_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())
# 
survey_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Date of survey",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number"),
                          percent_na = round(percent_na$value*100))

kable(survey_data_dictionary)
```

### CWT
```{r}

percent_na <- cwt_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())

cwt_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Sect",
                                          "Release Location",
                                          "Fork length",
                                          "Sex",
                                          "Spawn",
                                          "Clip",
                                          "Sample type",
                                          "Carcass stat",
                                          "Sample number",
                                          "Head tag",
                                          "Comment",
                                          "Header ID",
                                          "Date",
                                          "Tag colour",
                                          "Crew",
                                          "Week number"),
                          percent_na = round(percent_na$value*100))

kable(cwt_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)


```{r}
feather_carcass_chops_and_tags_2002 <- chop_channel_cleaner |> glimpse()
feather_carcass_chop_recovery_2002 <- chop_recovery_cleaner |> glimpse()
feather_carcass_cwt_2002 <-cwt_cleaner |> glimpse()
feather_carcass_chop_header_2002 <- chop_header_cleaner |> glimpse()
```
```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_carcass_chops_and_tags_2002,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chops_and_tags_2002.csv")

gcs_upload(feather_carcass_chop_recovery_2002,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_recovery_2002.csv")

gcs_upload(feather_carcass_cwt_2002,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_cwt_2002.csv")
gcs_upload(feather_carcass_chop_header_2002,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_header_2002.csv")

```
