---
title: "Feather Carcass QC 2012"
author: "Inigo Peng"
date: '2022-07-21'
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```
```{r}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/Chops_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "Chops_2012.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/ChopHeader_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2012.xlsx",
               overwrite = TRUE)

# 
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/CWTHeader_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTHeader_2012.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/CWT_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWT_2012.xlsx",
               overwrite = TRUE)
# 
Chops_raw <- read_excel("Chops_2012.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  select(-`Chop ID`) %>% 
  glimpse()


ChopHeader_raw <- read_excel("ChopHeader_2012.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  glimpse()


#Coded Wiretag Info

cwt_raw <- read_excel("CWT_2012.xlsx") %>% 
  rename("ID" = `CWT Header ID`) %>% 
  glimpse

cwt_header_raw <- read_excel("CWTHeader_2012.xlsx") %>% 
  rename("ID" = `CWT Header ID` ) %>% 
  glimpse
```

## Data transformations

```{r}

#1. chop table (with dates and tag color)
#2. Header has no time
chop_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, Date),
                               Chops_raw) %>% 
  clean_names() %>% 
  rename(sec = "section",
         min = "minutes",
         count = "total_count") %>% glimpse

chop_header <- ChopHeader_raw %>% 
  clean_names()%>%  glimpse

cwt_join <- full_join(cwt_header_raw, cwt_raw) %>% 
  glimpse()
```

### Counts

The `chop_join` table contains carcass counts 

### Survey

The `chop_header` table contains survey metadata and covariates

### CWT

The `cwt_join` table contains coded wire tag data

## Explore Numeric Variables: {.tabset}

```{r}
chop_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

### Variable: `id`, `min`

```{r}
summary(chop_join$id)
```

```{r}
summary(chop_join$min)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100
round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100
```

* `r round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100` % of values in the `id` column are NA.
* `r round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100` % of values in the `min` column are NA.

### Variable: `count`
```{r}
summary(chop_join$count)
```

* `r round(sum(is.na(chop_join$count))/nrow(chop_join), 3) * 100` % of values in the `count` column are NA.

**Plotting total_count over Period of Record**

```{r}
chop_join %>% 
  ggplot(aes(x = date, y = count)) + 
  geom_point() +
  theme_minimal()
```
```{r}
chop_join %>%
  group_by(date) %>%
  summarise(total_count = sum(count, na.rm = T)) %>%
  ggplot(aes(x = date, y = total_count)) +
  geom_col() +
  theme_minimal()
```
```{r}
cwt_join %>% 
  select_if(is.numeric) %>% 
  colnames()
```

## Explore Categorical variables: {.tabset}

```{r}
chop_join %>% 
  select_if(is.character) %>%
  colnames()
```
```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```
```{r}
cwt_join %>% 
  select_if(is.character) %>% 
  colnames()
```

## Clean data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_cleaner <- chop_join %>%
  # mutate(datetime = paste(as_date(date), time)) %>%
  mutate_if(is.character, str_to_lower) 
  # select(-c(date, time))

chop_cleaner
```
```{r}
chop_header_cleaner <- chop_header %>%
  # mutate(datetime = paste(as_date(date), time)) %>%
  mutate_if(is.character, str_to_lower) %>% 
  mutate(crew = str_replace_all(crew, "/", ","))
         # crew = str_replace_all(crew, " ",","))
  # select(-c(date, time))

chop_header_cleaner
```

```{r}
cwt_cleaner
```

## Data Dictionaries

# Count
```{r}
percent_na <- chop_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_cleaner),
                          description = c("ID",
                                          "Date of survey",
                                          "Sect", 
                                          "Min", 
                                          "Count",
                                          "Comments"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```
### Survey

```{r}

percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
counts_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("Date of survey",
                                          "Week Number",
                                          "Weather",
                                          "Crew memeber initials that collected",
                                          "Comments",
                                          "ID"),
                          percent_na = round(percent_na$value*100))
# 
kable(counts_data_dictionary)
```

### CWT
```{r}
cwt_clean
```

