---
title: "Feather Carcass QC 2009"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2009.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2009.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2009.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  Chops_raw <- sqlFetch(con, "ChopChannelTBL1")
  ChopHeader_raw <- sqlFetch(con, "ChopHeaderTBL")
  ChopRecov_raw <- sqlFetch(con, "ChopRecovTBL")
  cwt_raw <- sqlFetch(con, "CWTTagTBL")
  cwt_header_raw <- sqlFetch(con, "CWTHeaderTBL")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  # ChopHallprintTBL: TagID, sex, clip, headtag, fl, carc_stat, HeaderID
  # ScaleSPLs: empty
  # TagColLU_06: metadata about tag colors
  # Carcass Hallprint 2009: hallprint IDs, date, sex, FL, sec
  # FL by sex: empty
  # headtag: empty
  Chops_raw <- mdb.get(filepath, "ChopChannelTBL1")
  ChopHeader_raw <- mdb.get(filepath, "ChopHeaderTBL")
  ChopRecov_raw <- mdb.get(filepath, "ChopRecovTBL")
  cwt_raw <- mdb.get(filepath, "CWTTagTBL")
  cwt_header_raw <- mdb.get(filepath, "CWTHeaderTBL")
  detach(package:Hmisc)
}

write_csv(Chops_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "Chops_2009.csv"))
write_csv(ChopHeader_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2009.csv"))
write_csv(ChopRecov_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2009.csv"))
write_csv(cwt_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWT_2009.csv"))
write_csv(cwt_header_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2009.csv"))
```

## Raw Data Glimpse: {.tabset}

### ChopChannel_raw

```{r}
ChopChannel_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "Chops_2009.csv")) |> 
  glimpse()
```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopHeader_2009.csv")) |> 
  glimpse()
```

### ChopRecov_raw
```{r}
ChopRecov_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopRecov_2009.csv")) |> 
  glimpse()
```
### cwt_raw
```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "CWT_2009.csv")) |> 
  glimpse()
```

### cwt_header_raw

```{r}
cwt_header_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "CWTHeader_2009.csv")) |> 
  glimpse()
```





## Data transformations: {.tabset}

### CWT
The `cwt` table contains sampling information on individual carcasses. It is not joined with the other tables. Added a `count` column assuming each row represents one individual. This table does not contain `disposition` information (usually in the `carc_stat` variable). Added `source_table` column for tracking. Each `tag_id` is represented only once in this table.

```{r}
cwt <- cwt_raw |>  
  left_join(cwt_header_raw, by = c("HeaderID" = "headerID")) |> 
  clean_names() |>
  mutate(source_table = "CWT",
         count = 1) |>  
  rename(adipose_fin_clip = clip) |> 
  glimpse()
```

### Bulk chop counts

The `chops_and_tags` table has a chop count and tag count column. In earlier years (2000-2006) this table also contained either information about sex, lifestage, or adipose fin clip but this year no additional information is included. `disposition` column added to distinguish between chopped and tagged counts.

```{r}
chops_and_tags <- left_join(ChopChannel_raw, 
                            ChopHeader_raw |> select(HeaderID = headerID, TagCol, Date, Time),
                            by = "HeaderID") |>
  clean_names() |> 
  pivot_longer(c(chops, tags), 
               values_to = "count",
               names_to = "disposition") |> 
  mutate(disposition = if_else(disposition == "chops", "chopped", "tagged"),
         source_table = "ChopChannel") |> 
  glimpse()
```

### Recovered tags and chops

The `chop_recovery` table contains recovered carcass counts by color. According to Access relationships, `chan_id` connects the header table to the recovery table. Here we get header information from `chops_and_tags` above and join those to the `chop_recovery` table on `chan_id`. Renames `recov_count` to `count`. Adds `disposition` and `source_table` columns.

```{r}
chan_id_lookup <- chops_and_tags |> distinct(chan_id, date, time, sect, chan, min, chan_id, header_id, tag_col)
# HeaderID connects ChopChannel to ChopHeader; ChanID joins ChopChannel to ChopRecov
chop_recovery_join <- left_join(ChopRecov_raw |> rename(chan_id = ChanID),
                                chan_id_lookup, 
                                by = "chan_id") |> 
  clean_names() |> 
  rename(count = recov_count) |> 
  mutate(disposition = "chopped_recovered",
         source_table = "ChopRecov") |> 
  glimpse()
```

### Combine tables

```{r}
all_carcasses <- bind_rows(chops_and_tags, chop_recovery_join, cwt) |> glimpse()
```

### Survey

The `chop_header` table contains survey metadata and covariates 
```{r}
chop_header <- ChopHeader_raw |> 
  clean_names() |> glimpse()
```


## Explore Numeric Variables: {.tabset}

### All Carcasses Variables: `chan_id`, `sect`, `min`, `header_id`, `tag_col`, `count`, `recov_id`,  `recov_color`, `tag_id`, `fl`, `head_tag`, `scale_sample`, `week_num`

```{r}
# Filter clean data to show only numeric variables 
all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
```

```{r}
summary(all_carcasses$chan_id)
summary(all_carcasses$sect)
summary(all_carcasses$min)
summary(all_carcasses$header_id)
summary(all_carcasses$tag_col)
summary(all_carcasses$recov_id)
summary(all_carcasses$recov_color)
summary(all_carcasses$tag_id)
summary(all_carcasses$fl)
summary(all_carcasses$head_tag)
summary(all_carcasses$scale_sample)
summary(all_carcasses$week_num)
```

### All Carcasses Variable: `count`

```{r}
summary(all_carcasses$count)
```

** Plot carcass counts by sex **

```{r}
all_carcasses |>  
  mutate(sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         sex %in% c("ND", "UK") ~ "unknown",
                         TRUE ~ sex)) |> 
  group_by(date, sex) |>  
  summarize(count = sum(count, na.rm = T)) |> 
  ggplot(aes(x = date, y = count, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  theme(text = element_text(size = 15)) +
  ylim(c(0, 100))
```
This plot shows carcass over period of record by sex and is helpful for understanding the seasonality of the survey, i.e. chop counts peak in Oct-Nov. The population crashed around this time (source: Casey Campos, meeting). Because the bulk chops did not have sex information this year, most of the data have sex recorded as `NA`.

**Plotting Total Adult Chops over Period of Record**

```{r}
total_adult_chops <- all_carcasses |> 
  filter(disposition != "tagged") |> 
  select(date, count, disposition, sex) |> 
  glimpse()

total_adult_chops_summary <- total_adult_chops |> 
  group_by(date, sex) |> 
  summarise(total_adult_chops = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_adult_chops_summary |>
  ggplot(aes(x = date, y = total_adult_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

This plot shows the sex proportion of adult chops (no `tagged` carcasses) for each day over the period of `r min(all_carcasses$date, na.rm = T)` to `r max(all_carcasses$date, na.rm = T)`. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. This year there was no sex information.


**Plotting grilse over Period of Record**
This year did not have lifestage information.

**Plotting all carcasses over Period of Record**
Carcasses in the `all_carcasses` dataset are marked as either `tagged`, `chopped_recovered`, or  `chopped`. Carcasses from the `CWT` table are marked as `chopped`, `tagged`, or `NA` and have individual-level sampling information (i.e sex, fork length). `chopped` carcasses from `ChopChannel` do not have individual-level sampling data. `chopped_recovered` carcasses come from the `ChopRecov` data table and have no individual-level sampling information. `tagged` carcasses come from the `ChopChannel` data table and do not have individual-level sampling information.

```{r}
total_carcasses_summary <- all_carcasses |> 
  group_by(date, disposition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_carcasses_summary |> 
  ggplot(aes(x = date, y = total_carcasses, fill = disposition)) + 
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Individual-level sampling information

**Plotting fork length of each sex**

```{r}
summary(all_carcasses$fl)
```

```{r}
all_carcasses |> 
  mutate(sex = case_when(sex %in% c("F", "fem") ~ "female",
                         sex == "M" ~ "male",
                         sex %in% c("UK", "ND") ~ "unknown")) |>
  ggplot(aes(x = sex, y = fl)) +
  geom_boxplot() +
  facet_wrap(~disposition) +
  theme_minimal() + 
  labs(y = "fork length (mm)", x = "sex")
```

**Plotting sex proportion of counts within each disposition**

```{r}
all_carcasses |> 
  mutate(sex = case_when(sex %in% c("F", "fem") ~ "female",
                         sex == "M" ~ "male",
                         sex %in% c("UK", "ND") ~ "unknown")) |> 
  ggplot(aes(y = count, x = disposition, fill = sex)) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Disposition") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Chop Header Variable: `header_id`, `week_num`, `tag_col`

```{r}
chop_header |>  
  select_if(is.numeric) |>  
  colnames()
```
```{r}
summary(chop_header$header_id)
summary(chop_header$week_num)
summary(chop_header$tag_col)
```


## Explore Categorical variables: {.tabset}

### All Carcasses

Fix inconsistencies with spelling, capitalization, and dates

```{r}
all_carcasses |>  
  select_if(is.character) |>  
  colnames()
```

```{r}
unique(all_carcasses$chan)
unique(all_carcasses$disposition)
unique(all_carcasses$source_table)
unique(all_carcasses$sex)
unique(all_carcasses$spawn)
unique(all_carcasses$adipose_fin_clip)
unique(all_carcasses$samp_type)
unique(all_carcasses$rel_loc)
unique(all_carcasses$run)
```

```{r}
all_carcasses_cleaner <- all_carcasses |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(sex = case_when(sex %in% c("f", "female", "fem") ~ "female",
                         sex == "m" ~ "male",
                         sex %in% c("nd", "uk") ~ "unknown", 
                         TRUE ~ sex),
         spawn = case_when(spawn == "u" ~ "unspawned",
                           spawn == "uk" ~ "unknown",
                           spawn %in% c("no data", "nd") ~ NA_character_,
                           spawn == "s" ~ "spawned",
                           TRUE ~ spawn),
         adipose_fin_clip = case_when(adipose_fin_clip %in% c("no data", "na", "uk") ~ NA_character_, 
                                      adipose_fin_clip == "y" ~ "TRUE",
                                      adipose_fin_clip == "n" ~ "FALSE"),
         adipose_fin_clip = as.logical(adipose_fin_clip)) |>
  select(-c(time)) |>  # time is removed because no stop time provided so information is not useful. chan is all NAs
  glimpse()
```

### Chop Header Clean Data

```{r}
chop_header |>  
  select_if(is.character) |>  
  colnames()
```
```{r}
unique(chop_header$crew)
unique(chop_header$recorder)
unique(chop_header$weather)
```


```{r}
chop_header_cleaner <- chop_header |> 
  mutate_if(is.character, str_to_lower) |>  
  mutate(crew = str_replace_all(crew, " ", ""),
         crew = str_replace_all(crew, "/", ","),
         crew = str_replace_all(crew, "&", ","),
         weather = case_when(weather == "ran" ~ "rain",
                             weather == "cld" ~ "cloudy",
                             TRUE ~ weather)) |>  
  select(-time) # stop is all NA and start is not useful without stop

```

## Comments

* Table layout/structure same as 2002-2008.
* `Hallprint` variable in `cwt` (began collecting in 2004).
* `run` variable available in `cwt` (began collecting in 2008)
* `ChopChannel` table this year does not contain sex, lifestage, or fin clip information.
* what is `samp_num`, `rel_loc`, `head_tag`?
* can we join recovered tags on a specific `id` value?

## Data Dictionaries

### All Carcasses
```{r}
percent_na <- all_carcasses_cleaner |> 
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |> 
  pivot_longer(cols = everything())


all_carcasses_data_dictionary <- tibble(variables = colnames(all_carcasses_cleaner),
                          description = c("Channel ID",
                                          "Section",
                                          "Channel",
                                          "Minute",
                                          "Header ID",
                                          "Tag color",
                                          "Date of survey",
                                          "Disposition",
                                          "Count",
                                          "Table from which the data came",
                                          "Recovery ID",
                                          "Recovered tag color",
                                          "Tag ID",
                                          "Fork length (mm)",
                                          "Sex",
                                          "Spawning condition",
                                          "Adipose fin clip status",
                                          "Sample type",
                                          "Sample Number",
                                          "Head Tag",
                                          "Comment",
                                          "Hall Print 1",
                                          "Hall Print 2",
                                          "Scale sample",
                                          "Run",
                                          "Crew",
                                          "Week Number"),
                          percent_na = round(percent_na$value*100))

kable(all_carcasses_data_dictionary)
```

## Save cleaned data back to google cloud (TBA)

```{r}
all_carcasses_2009 <- all_carcasses_cleaner |> glimpse()
```

```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_carcasses_2009,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2009.csv")
```








## Data transformations: {.tabset}

### Counts

The `chop_channel` table contains carcass counts by chop/tagged based on clips


```{r}
#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw |> 
                                 select(HeaderID = headerID, TagCol, Date, Time),
                               ChopChannel_raw, 
                               by = "HeaderID") |> 
  clean_names()
```

### Recovery

The `chop_recovery` table contains recovered carcass counts by color

```{r}
# 2. choprecovery table (with dates)
# should join headerID to chopchannel headerID, and then chanID from chop channel to ChopRecovery
# HeaderID connects ChopChannel to ChopHeader; ChanID joins ChopChannel to ChopRecov
chop_recovery_join <- full_join(chop_channel_join |> 
                                  select(date, time, sect, chan, min, chan_id, header_id,
                                         tag_col), ChopRecov_raw |> 
                                  rename(chan_id = ChanID), 
                                by = "chan_id") |> 
  clean_names() |> 
  glimpse()

# chop_recovery_join <- full_join(ChopHeader_raw |> 
#                                   select(ID, Date, Time),
#                                 ChopRecov_raw |> 
#                                   rename(ID = RecovID), by = "ID")  |> 
#   clean_names()
```

### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopHeader_raw |> 
  clean_names()
```

### CWT
The `cwt` table contains biological sampling data for carcasses. It does not contain information on whether these carcasses were chopped or tagged. Each `tag_id` is only represented once in the table.

```{r}
cwt <- full_join(cwt_raw |> clean_names(), cwt_header_raw |> clean_names(),
                 by = "header_id") |>  
  glimpse()
```

## Explore Numeric Variables: {.tabset}

### Variable: `header_id`, `chan_id`, `sect`, `min`

```{r}
chop_channel_join |> 
  select_if(is.numeric) |>
  colnames()
```

```{r}
summary(chop_channel_join$header_id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

### Chop Channel Variable: `chops`, `tags`

```{r}
summary(chop_channel_join$chops)
```

```{r}
summary(chop_channel_join$tags)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$chops))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$tags))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$chops))/nrow(chop_channel_join), 3) * 100` % of values in the `chops` column are NA.
* `r round(sum(is.na(chop_channel_join$tags))/nrow(chop_channel_join), 3) * 100` % of values in the `tags` column are NA.

**Plotting chops over Period of Record**

```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(chops = sum(chops)) |> 
  ggplot(aes(x = date, y = chops)) +
  geom_col(fill = "blue") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```


**Plotting tags over Period of Record**

```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(tags = sum(tags)) |> 
  ggplot(aes(x = date, y = tags)) +
  geom_col(fill = "red") + 
  labs(x = "Date", 
       y = "Tags") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Comparing tags and chops**
```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(chops_total = sum(chops, na.rm = T),
            tags_total = sum(tags, na.rm = T)) |>
  pivot_longer(cols = c(chops_total, tags_total), names_to = "tags_or_chops", values_to = "count") |> 
  ggplot(aes(x = date, y = count, fill = tags_or_chops)) +
  geom_col() + 
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

Tags appear to be less commonly applied compared to chops.

### Chop Recovery Variable: `header_id`, `chan_id`, `recov_count`

```{r}
# Filter clean data to show only numeric variables 
chop_recovery_join |> 
  select_if(is.numeric) |>
  colnames()
```

```{r}
summary(chop_recovery_join$header_id)
```
```{r}
summary(chop_recovery_join$recov_count)
```

* `r round(sum(is.na(chop_recovery_join$recov_count))/nrow(chop_recovery_join$recov_count), 3) * 100` % of values in the `recov_count` column are NA.

**Plotting chop_recovery over Period of Time**

```{r}
chop_recovery_join |> 
  ggplot(aes(x = date, y = recov_count)) + 
  geom_col() +
  theme_minimal()
```

### Chop Header Variable: `header_id`

```{r}
chop_header |> 
  select_if(is.numeric) |> 
  colnames()
```
```{r}
summary(chop_header$header_id)
```

### CWT Variable: `tag_id`, `sect`, `fl`, `header_id`, `week_num`
```{r}
cwt |> 
  select_if(is.numeric) |> 
  colnames()
```
```{r}
summary(cwt$tag_id)
```
```{r}
summary(cwt$sect)
```
```{r}
summary(cwt$fl)
```
* `r round(sum(is.na(cwt$tag_id))/nrow(cwt), 3) * 100` % of values in the `tag_id` column are NA.
* `r round(sum(is.na(cwt$sect))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.
* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
cwt_count <- cwt |> 
  mutate(count = 1) |>
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "U",
                         TRUE ~ sex)) |> 
  pivot_wider(names_from = sex, values_from = count, values_fill = 0) |> 
  # unnest() |> 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) |> glimpse()

total_cwt_summary <- cwt_count |> 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) |> 
  group_by(month(date)) |> 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary |> 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") |> 
  mutate(proportions = (count / total_cwt)) |> 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Month") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

**Plotting fork length of each sex**

```{r}
cwt |> 
  filter(!is.na(sex)) |> 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```


## Explore Categorical variables: {.tabset}

### Chop Channel Clean Data

Fix inconsistencies with spelling, capitalization, and dates
```{r}
# Filter clean data to show only categorical variables
chop_channel_join |> 
  select_if(is.character) |>
  colnames()
```

```{r}
unique(chop_channel_join$chan)
```
```{r}
chop_channel_cleaner <- chop_channel_join |>
  mutate(date = as_date(date),
         tag_col = str_replace_all(tag_col, " ", "")) |>
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time)) |> 
  glimpse()
```
### Chop Recovery Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_recovery_join |> 
  select_if(is.character) |> 
  colnames()
```

```{r}
chop_recovery_cleaner <- chop_recovery_join |> 
  mutate(date = as_date(date),
         recov_color = str_replace_all(recov_color, " ", "")) |>
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time)) |> 
  glimpse()
```

### Chop Header Clean Data

```{r}
chop_header |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$recorder)
```
```{r}
unique(chop_header$weather)
```



```{r}
chop_header_cleaner <- chop_header |>
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         weather = case_when(weather == "cld" ~ "cloudy",
                             weather == "ran" ~ "rain",
                             TRUE ~ weather)) |> 
  select(-c(time)) |> 
  glimpse()
```

### CWT Clean Data

```{r}
cwt |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(cwt$spawn)
unique(cwt$clip)
unique(cwt$samp_type)
```


```{r}
cwt_cleaner <- cwt |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         sex = case_when(sex == "m" ~ "male",
                         sex == "f" ~ "female",
                         TRUE ~ sex), 
         spawn = case_when(spawn == "no data" ~ NA_character_,
                           spawn == "uk" ~ "unknown",
                           spawn == "u" ~ "unspawned",
                           spawn == "s" ~ "spawned",
                           TRUE ~ spawn),
         clip = case_when(clip %in% c("no data", "uk") ~ NA_character_,
                          clip == "y" ~ "yes",
                          clip == "n" ~ "no", 
                          TRUE ~ clip
                          ),
         samp_type = ifelse(samp_type == "none", NA, samp_type)) |> 
  glimpse()

```

## Comment

* Similar to previous years dataset

## Data Dictionaries

# Channel
```{r}
percent_na <- chop_channel_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("Header ID",
                                          "Color of tag applied to carcass",
                                          "Date of survey",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped",
                                          "Carcass that were tagged"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
percent_na <- chop_recovery_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


recovery_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("Date of Survey",
                                          "Sect",
                                          "Chan",
                                          "Min",
                                          "Chan ID",
                                          "Header ID",
                                          "Color of tag recovered from carcass",
                                          "Count of recovery",
                                          "Color of recovery tag", 
                                          "Count of recvered tags"),
                          percent_na = round(percent_na$value*100))

kable(recovery_data_dictionary)
```

### Survey

```{r}

percent_na <- chop_header_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())
# 
header_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Date of survey",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number"),
                          percent_na = round(percent_na$value*100))
# 
kable(header_data_dictionary)
```

### CWT
```{r}
percent_na <- cwt_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())

cwt_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Sect",
                                          "Fork length",
                                          "Sex",
                                          "Spawn",
                                          "Clip",
                                          "Sample type",
                                          "Sample number",
                                          "Head tag",
                                          "Comment",
                                          "Header ID",
                                          "Hall print 1",
                                          "Hall print 2",
                                          "Scale sample",
                                          "Run",
                                          "Date",
                                          "Tag colour",
                                          "Crew",
                                          "Week number"),
                          percent_na = round(percent_na$value*100))

kable(cwt_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)

## Save cleaned data back to google cloud (TBA)

```{r}
#Ignore chop recovery as its 100% NA
feather_carcass_chops_and_tags_2009 <- chop_channel_cleaner |> glimpse()
feather_carcass_cwt_2009 <- cwt_cleaner |> glimpse()
feather_carcass_chop_header_2009 <- chop_header_cleaner |> glimpse()
```
```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_carcass_chops_and_tags_2009,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chops_and_tags_2009.csv")
gcs_upload(feather_carcass_cwt_2009,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_cwt_2009.csv")
gcs_upload(feather_carcass_chop_header_2009,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_header_2009.csv")
```

|> |> |> |> |> |> |> 