---
title: "Feather Carcass QC 2006"
author: "Inigo Peng"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data



```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx


```

```{r}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2006/ChopChannelTBL1_2006.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopChannel_2006.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2006/ChopHeaderTBL_2006.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2006.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2006/ChopRecovTBL_2006.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopRecov_2006.xlsx",
               overwrite = TRUE)

```
```{r}
ChopChannel_raw <- read_excel("ChopChannel_2006.xlsx") %>% 
  rename("ID" = HeaderID) %>%
  glimpse()

ChopHeader_raw <- read_excel("ChopHeader_2006.xlsx") %>% 
  rename("ID" = headerID) %>%
  glimpse()

ChopRecov_raw <- read_excel("ChopRecov_2006.xlsx") %>% 
  glimpse()
```


## Data transformations

```{r}
# 1. we need to link chopheader and chopchannel to get the dates and tag color applied
# 2. we need to link chopheader and choprecov to get the dates

#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, TagCol, Date, Time),
                               ChopChannel_raw) %>% 
  clean_names()

# 2. choprecovery table (with dates)

chop_recovery_join <- full_join(ChopHeader_raw %>% 
                                  select(ID, Date, Time),
                                ChopRecov_raw %>% 
                                  rename(ID = RecovID))  %>% 
  clean_names()

chop_header <- ChopHeader_raw %>% 
  clean_names()
```

### Counts

The `chop_channel` table contains carcass counts by sex and chop/tagged

### Recovery

The `chop_recovery` table contains recovered carcass counts by color

### Survey

The `chop_header` table contains survey metadata and covariates


## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
chop_channel_join %>% 
  select_if(is.numeric) %>%
  colnames()
```
### Variable: `id`, `chan_id`, `sect`, `min`

```{r}
summary(chop_channel_join$id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

### Variable: `chop_clip`, `chop_n_clip`,`chop_uncheck`

**Numeric Summary of `chop_clip`, `chop_n_clip`,`chop_uncheck` over Period of Record**

```{r}
summary(chop_channel_join$chop_clip)
```

```{r}
summary(chop_channel_join$chop_n_clip)
```

```{r}
summary(chop_channel_join$chop_uncheck)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$chop_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$chop_n_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$chop_uncheck))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$chop_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `chop_clip` column are NA.
* `r round(sum(is.na(chop_channel_join$chop_n_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `chop_n_clip` column are NA.
* `r round(sum(is.na(chop_channel_join$chop_uncheck))/nrow(chop_channel_join), 3) * 100` % of values in the `chop_uncheck` column are NA.

**Plotting chop_clip over Period of Record**

```{r}
 
chop_channel_join %>% 
  ggplot(aes(x = date, y = chop_clip)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Chop With Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```


**Plotting chop_n_clip over Period of Record**

```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = chop_n_clip)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Date", 
       y = "Chop Without Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting chop_uncheck over Period of Record**

```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = chop_uncheck)) +
  geom_point(size = 1.4, alpha = .5, color = "black") + 
  labs(x = "Date", 
       y = "Chop Not Checked") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting Total chops over Period of Record**

```{r}
chop_channel_join %>% glimpse()

total_chops <- chop_channel_join%>%
  mutate(chop_clip = ifelse(is.na(chop_clip), 0, chop_clip), # fill na
         chop_n_clip = ifelse(is.na(chop_n_clip), 0, chop_n_clip),
         chop_uncheck = ifelse(is.na(chop_uncheck), 0, chop_uncheck),
         total_chops = chop_clip + chop_n_clip + chop_uncheck) %>% 
  select(date, chop_clip, chop_n_clip, chop_uncheck, total_chops)

# Case check: checking for duplicates
total_chops_test <- total_chops %>% # get total chops
  filter(date == as_date("2006-09-15"))

# Check if NAs introduced in creating total chop columns
nrow(total_chops_test) == (sum(total_chops_test$total_chops == (total_chops_test$chop_clip + total_chops_test$chop_n_clip + total_chops_test$chop_uncheck)))

# total number of chops for the day
total_chops_summary <- total_chops %>%
  group_by(date) %>%
  summarise(total_chops = sum(total_chops, na.rm = T),
            chop_clip = sum(chop_clip, na.rm = T),
            chop_n_clip = sum(chop_n_clip, na.rm = T),
            chop_uncheck = sum(chop_uncheck, na.rm = T))



# check if NAs are introduced in merge
A = left_join(ChopChannel_raw, ChopHeader_raw, by = "ID")$ChanID
B = ChopRecov_raw$ChanID
sum(B %in% A) # they are
# yes I noticed that. there are some IDs in the chop_channel table that are not in the chop_header table. there are A LOT of IDs in the recovery table that are not in header table. at this point i would assume this is a data entry issue and we can leave as is but clearly document. this is something we may learn more about and then be able to fix
```
```{r}
total_chops_summary %>% 
  pivot_longer(cols = c(chop_clip, chop_n_clip, chop_uncheck), names_to = "clip", values_to = "count") %>% 
  mutate(proportions = (count / total_chops)) %>% 
  ggplot(aes(x = date, y = proportions, fill = clip)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Sex", 
                    labels = c("Clip", "No Clip", "Not Checked")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```
This plot shows the  proportion of chops with, without clips and unchecked for each day over the the period of 2006-09-05 to 2006-12-22. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. We see that the majority of the chops were unchecked for clips.

"tag_clip"     "tag_n_clip"   "tag_unk"

Comments: the dataset from 2006 did not record for sex and instead, recorded for clip(?) or not.

###Variable: `tag_clip`, `tag_n_clip`, `tag_unk`

**Numeric Summary of `tag_clip`, `tag_n_clip`, `tag_unk` over Period of Record**

```{r}
summary(chop_channel_join$tag_clip)
```

```{r}
summary(chop_channel_join$tag_n_clip)
```

```{r}
summary(chop_channel_join$tag_unk)
```


**NA and Unknown Values**
Provide a stat on NA or unknown values

```{r}
round(sum(is.na(chop_channel_join$tag_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$tag_n_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$tag_unk))/nrow(chop_channel_join), 3) * 100
```

* `r round(sum(is.na(chop_channel_join$tag_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_clip` column are NA.
* `r round(sum(is.na(chop_channel_join$tag_n_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_n_clip` column are NA.
  * `r round(sum(is.na(chop_channel_join$tag_unk))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_unk` column are NA.

**Plotting tag_clip over Period of Record**
```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = tag_clip)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Tag Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting tag_n_clip over Period of Record**

```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = tag_n_clip)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Date", 
       y = "Tag No Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting tag_unk over Period of Record**
```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = tag_unk)) +
  geom_point(size = 1.4, alpha = .5, color = "black") + 
  labs(x = "Date", 
       y = "Unknown Tags") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting Total Tags over Period of Record**

```{r}
# Percentage by Gender of Daily Chops
# Address NA values for chops
total_tags <- chop_channel_join %>%
  mutate(tag_clip = ifelse(is.na(tag_clip), 0, tag_clip), # fill na
         tag_n_clip = ifelse(is.na(tag_n_clip), 0, tag_n_clip),
         tag_unk = ifelse(is.na(tag_unk), 0, tag_unk),
         total_tags = tag_unk + tag_clip + tag_n_clip) %>% 
  select(date, tag_clip, tag_n_clip, tag_unk, total_tags)

# total number of chops for the day
total_tags_summary <- total_tags %>%
  group_by(date) %>%
  summarise(total_tags = sum(total_tags, na.rm = T),
            tag_clip = sum(tag_clip, na.rm = T),
            tag_n_clip = sum(tag_n_clip, na.rm = T),
            tag_unk = sum(tag_unk, na.rm = T))

# Plot to identify proportions by sex for each day
total_tags_summary %>% 
  pivot_longer(cols = c(tag_n_clip, tag_clip, tag_unk), names_to = "tags", values_to = "count") %>% 
  mutate(proportions = (count / total_tags)) %>% 
  ggplot(aes(x = date, y = proportions, fill = tags)) +
  geom_bar(stat = "identity", position = "stack")

# Make Graph Prettier
total_tags_summary %>% 
  pivot_longer(cols = c(tag_n_clip, tag_clip, tag_unk), names_to = "tags", values_to = "count") %>% 
  mutate(proportions = (count / total_tags)) %>% 
  ggplot(aes(x = date, y = proportions, fill = tags)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Sex", 
                    labels = c("Tags Without Clips", "Tags With Clips", "Tags Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

**Comparing tags and chops**

```{r}
# from ashley: please improve on this plot I got started, make it look better, describe it, double check that the numbers make sense
total_carcass <- full_join(total_chops_summary, total_tags_summary) %>% 
  select(date, total_chops, total_tags) %>% 
  pivot_longer(c(total_chops, total_tags), names_to = "mark")

full_join(total_chops_summary, total_tags_summary) %>% 
  select(date, total_chops, total_tags) %>% 
  pivot_longer(c(total_chops, total_tags), names_to = "mark")

ggplot(total_carcass, aes(x = date, y = value, fill = mark)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```

Tags appear to be less commonly applied compared to chops.

```{r}
# Filter clean data to show only numeric variables 
chop_recovery_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

```{r}
summary(chop_recovery_join$id)
```

```{r}
summary(chop_recovery_join$chan_id)
```

```{r}
summary(chop_recovery_join$recov_count)
```

Note: Cannot plot recovery count over period of record because none of the recovered count has date associated to it.

```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(chop_header$id)
```

## Explore Categorical variables: {.tabset}

```{r}
# Filter clean data to show only categorical variables
chop_channel_join %>% 
  select_if(is.character) %>%
  colnames()
```
```{r}
chop_recovery_join %>% 
  select_if(is.character) %>% 
  colnames()
```

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```

## Clean data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_channel_cleaner <- chop_channel_join %>%
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_channel_cleaner
```

```{r}
chop_recovery_cleaner <- chop_recovery_join %>% 
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_recovery_cleaner
```

```{r}
chop_header_cleaner <- chop_header %>%
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_header_cleaner
```

## Data Dictionaries

### Channel

```{r}
percent_na <- chop_channel_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("ID",
                                          "Color of tag applied to carcass",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped and has clip",
                                          "Carcass that were chopped and has no clip",
                                          "Carcass that were chopped and not checked for clip",
                                          "Carcass that were tagged and has clip",
                                          "Carcass that were tagged and has no clip",
                                          "Carcass that were tagged and unknown",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
# from ashley: please insert data dictionary for cleaned recovery table
percent_na <- chop_recovery_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("ID",
                                          "Color of tag recovered from carcass",
                                          "Count of recovery",
                                          "Channel ID",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Survey

```{r}
# 
percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
counts_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))
# 
kable(counts_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)

```{r}
# Name file [watershed]_[data type].csv
```

## Save cleaned data back to google cloud (TBA)

```{r}
# Name file [watershed]_[data type].csv
```
