---
title: "Feather Carcass QC 2002"
author: "Inigo Peng"
date: '2022-07-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

```{r}
#Connect to microsoft access through RODBC
#Use 
DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
MDBPATH <- "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/Carcass 2002.mdb"
PATH <- paste0(DRIVERINFO, "DBQ=", MDBPATH)

con <- odbcDriverConnect(PATH)

# Show the different table names 
sqlTables(con)$TABLE_NAME

#Download table and write csv
ChopChannel_1 <- sqlFetch(con, "ChopChannelTBL")
ChopChannel_2 <- sqlFetch(con, "ChopChannelTBL2")
ChopChannel_raw <- bind_rows(ChopChannel_1, ChopChannel_2)
write_csv(ChopChannel_raw, "ChopChannel_2002.csv")

ChopRecov_1 <- sqlFetch(con,"ChopRecovTBL")
ChopRecov_2 <- sqlFetch(con, "ChopRecovTBL2")
ChopRecov_raw <- bind_rows(ChopRecov_1, ChopRecov_2)
write_csv(ChopRecov_raw, "ChopRecov_2002.csv")

ChopHeader_1 <- sqlFetch(con,"ChopHeaderTBL")
ChopHeader_2 <- sqlFetch(con, "ChopHeaderTBL2")
ChopHeader_raw <- bind_rows(ChopHeader_1, ChopHeader_2)
write_csv(ChopHeader_raw, "ChopHeader_2002.csv")


cwt_1 <- sqlFetch(con, "CWTTagTBL") %>% 
  select(-WeekNum)
cwt_2 <- sqlFetch(con, "CWTTagTBL1") %>% 
  mutate(SampNum = as.character(SampNum))
cwt_raw <- bind_rows(cwt_1, cwt_2)
write_csv(cwt_raw, "cwt_2002.csv")

cwt_header_1 <- sqlFetch(con, "CWTHeaderTBL")
cwt_header_2 <- sqlFetch(con, "CWTHeaderTBL1")
cwt_header_raw <-  bind_rows(cwt_header_1, cwt_header_2)
write_csv(cwt_header_raw, "CWTHeader_2002.csv")

#Other tables include Steelhead Data and TagColLu
```

```{r}
ChopChannel_raw <- read_csv("ChopChannel_2002.csv") %>% 
  rename("ID" = HeaderID) %>% 
  glimpse()

ChopRecov_raw <- read_csv("ChopRecov_2002.csv") %>% glimpse

ChopHeader_raw <- read_csv("ChopHeader_2002.csv") %>% 
  rename("ID" = headerID) %>%
  glimpse()

cwt_raw <- read_csv("cwt_2002.csv") %>% glimpse()

cwt_header_raw <- read_csv("CWTHeader_2002.csv")
```
## Data transformations

```{r}
#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, TagCol, Date, Time),
                               ChopChannel_raw) %>% 
  clean_names() %>% 
  rename("male_chop" = male,
         "female_chop" = fem,
         "grilse" = gril) %>% glimpse()

chop_recovery_join <- full_join(ChopHeader_raw %>% 
                                  select(ID, Date, Time),
                                ChopRecov_raw %>% 
                                  rename(ID = RecovID))  %>% 
  clean_names()

chop_header <- ChopHeader_raw %>% 
  clean_names()
```

### Counts

The `chop_channel` table contains carcass counts by sex, grilise and chop/tagged

### Recovery

The `chop_recovery` table contains recovered carcass counts by color

### Survey

The `chop_header` table contains survey metadata and covariates


## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
chop_channel_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

### Variable: `id`, `chan_id`, `sect`, `min`

```{r}
summary(chop_channel_join$id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

### Variable: `female_chop`, `male_chop`,`grilise`, `tag_count`

**Numeric Summary of `female_chop`, `male_chop`,`unknown_chops`, `tag_count` over Period of Record**

```{r}
summary(chop_channel_join$female_chop)
```

```{r}
summary(chop_channel_join$male_chop)
```

```{r}
summary(chop_channel_join$unknown_chops)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$male_chop))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$female_chop))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$grilse))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$male_chop))/nrow(chop_channel_join), 3) * 100` % of values in the `male_chop` column are NA.
* `r round(sum(is.na(chop_channel_join$female_chop))/nrow(chop_channel_join), 3) * 100` % of values in the `female_chop` column are NA.
* `r round(sum(is.na(chop_channel_join$grilse))/nrow(chop_channel_join), 3) * 100` % of values in the `unknown_chops` column are NA.

**Plotting female_chop over Period of Record**

```{r}
# Make whatever plot is appropriate 
filter(chop_channel_join) %>% 
  ggplot(aes(x = date, y = female_chop)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Female Chop") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```


**Plotting male_chop over Period of Record**
```{r}
filter(chop_channel_join) %>% 
  ggplot(aes(x = date, y = male_chop)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Date", 
       y = "Male Chop") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting grilse over Period of Record**
```{r}
filter(chop_channel_join) %>% 
  ggplot(aes(x = date, y = grilse)) +
  geom_point(size = 1.4, alpha = .5, color = "black") + 
  labs(x = "Date", 
       y = "Grilse") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting Total chops over Period of Record**

```{r}
chop_channel_join %>% glimpse()

# Percentage by Gender of Daily Chops
# Address NA values for chops
total_chops <- chop_channel_join%>%
  mutate(male_chop = ifelse(is.na(male_chop), 0, male_chop), # fill na
         female_chop = ifelse(is.na(female_chop), 0, female_chop),
         grilse_chop = ifelse(is.na(grilse), 0, grilse),
         total_chops = grilse_chop + male_chop + female_chop) %>% 
  select(date, male_chop, female_chop, grilse_chop, total_chops)

# Case check: checking for duplicates
total_chops_test <- total_chops %>% # get total chops
  filter(date == as_date("2002-09-20"))

# Check if NAs introduced in creating total chop columns
nrow(total_chops_test) == (sum(total_chops_test$total_chops == (total_chops_test$male_chop + total_chops_test$female_chop + total_chops_test$grilse_chops)))

# total number of chops for the day
total_chops_summary <- total_chops %>%
  group_by(date) %>%
  summarise(total_chops = sum(total_chops, na.rm = T),
            male_chops = sum(male_chop, na.rm = T),
            female_chops = sum(female_chop, na.rm = T),
            grilse_chops = sum(grilse_chop, na.rm = T))
```

```{r}
total_chops_summary %>% 
  pivot_longer(cols = c(male_chops, female_chops, grilse_chops), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count / total_chops)) %>% 
  ggplot(aes(x = date, y = proportions, fill = sex)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Sex", 
                    labels = c("Female", "Male", "Grilse")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```

This plot shows the proportion of chops based on sex and grilse for each day over the the period of 2002-09-03 to 2002-12-19. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. 

```{r}
summary(chop_channel_join$tag_count)
```

```{r}
round(sum(is.na(chop_channel_join$tag_count))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$tag_count))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_count` column are NA.

**Plotting tag_count over Period of Record**

```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = tag_count)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Date", 
       y = "Tag Count") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

```{r}
total_tags_summary <- chop_channel_join %>%
  group_by(date) %>%
  summarise(total_tags = sum(tag_count, na.rm = T))

total_tags_summary %>% 
  ggplot(aes(x = date, y = total_tags)) +
  geom_col() +
  theme_minimal() +
  labs(y = "Daily Total Tags", x = "Date")

```
**Comparing tags and chops**

```{r}
# from ashley: please improve on this plot I got started, make it look better, describe it, double check that the numbers make sense
total_carcass <- full_join(total_chops_summary, total_tags_summary) %>% 
  select(date, total_chops, total_tags) %>% 
  pivot_longer(c(total_chops, total_tags), names_to = "mark")

full_join(total_chops_summary, total_tags_summary) %>% 
  select(date, total_chops, total_tags) %>% 
  pivot_longer(c(total_chops, total_tags), names_to = "mark")

ggplot(total_carcass, aes(x = date, y = value, fill = mark)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```
Tags are significantly less applied compared to chops.

```{r}
chop_recovery_join %>% 
  select_if(is.numeric) %>% 
  colnames()
```

```{r}
summary(chop_recovery_join$id)
```

```{r}
summary(chop_recovery_join$chan_id)
```

```{r}
summary(chop_recovery_join$recov_count)
```

**Plotting recov_count over Period of Record**

```{r}
chop_recovery_join %>% 
  ggplot(aes(x = date, y = recov_count)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Date", y = "Recovered Chop Count")
```
This chart shows the recovered chops. The chart is not showing count values that do not have dates or times associated with them.

```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(chop_header$id)
```

## Explore Categorical variables: {.tabset}

```{r}
chop_channel_join %>% 
  select_if(is.character) %>%
  colnames()
```
```{r}
chop_recovery_join %>% 
  select_if(is.character) %>% 
  colnames()
```

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```

## Clean data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_channel_cleaner <- chop_channel_join %>%
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_channel_cleaner
```

```{r}
chop_recovery_cleaner <- chop_recovery_join %>% 
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_recovery_cleaner
```

```{r}
chop_header_cleaner <- chop_header %>%
  # mutate(time = case_when(is.na(chop_header$time) ~ " ",
                          # TRUE ~ as_hms(chop_header$time))) %>% glimpse
  # mutate(time = as_hms(chop_header$time)) %>% 
  # mutate(time = ifelse(is.na(chop_header$time), " ", as_hms(chop_header$time))) %>% glimpse
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_header_cleaner
```

## Data Dictionaries

### Channel

```{r}
percent_na <- chop_channel_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("ID",
                                          "Color of tag applied to carcass",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped and male",
                                          "Carcass that were chopped and female",
                                          "Carcass that were grilse",
                                          "Carcass that were tagged",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
# from ashley: please insert data dictionary for cleaned recovery table
percent_na <- chop_recovery_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("ID",
                                          "Color of tag recovered from carcass",
                                          "Count of recovery",
                                          "Channel ID",
                                          "Datetime"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Survey

```{r}
# 
percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
counts_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number",
                                          "Datetime"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

## Save cleaned data back to google cloud (TBA)

```{r}
# Name file [watershed]_[data type].csv
```