---
title: "Feather Carcass QC 2009"
author: "Inigo Peng"
date: '2022-07-21'
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```
```{r}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2009/ChopChannelTBL1_2009.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopChannel_2009.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2009/ChopHeaderTBL_2009.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2009.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2009/ChopRecovTBL_2009.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopRecov_2009.xlsx",
               overwrite = TRUE)
# 
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2009/CWTHeaderTBL_2009.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTHeader_2009.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2009/CWTTagTBL_2009.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTTag_2009.xlsx",
               overwrite = TRUE)
# 
ChopChannel_raw <- read_excel("ChopChannel_2009.xlsx") %>% 
  rename("ID" = HeaderID) %>%
  glimpse()


ChopHeader_raw <- read_excel("ChopHeader_2009.xlsx") %>% 
  rename("ID" = headerID) %>%
  glimpse()

ChopRecov_raw <- read_excel("ChopRecov_2009.xlsx") %>% 
  glimpse()

#Coded Wiretag Info

cwt_raw <- read_excel("CWTTag_2009.xlsx") %>% 
  glimpse
```

## Data transformations

```{r}

#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, TagCol, Date, Time),
                               ChopChannel_raw) %>% 
  clean_names() %>% glimpse
# 2. choprecovery table (with dates)

chop_recovery_join <- full_join(ChopHeader_raw %>% 
                                  select(ID, Date, Time),
                                ChopRecov_raw %>% 
                                  rename(ID = RecovID))  %>% 
  clean_names()%>% glimpse

chop_header <- ChopHeader_raw %>% 
  clean_names()%>%  glimpse
```
### Variable: `id`, `chan_id`, `sect`, `min`

```{r}
summary(chop_channel_join$id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

### Variable: `chop`, `tag`


### Counts

The `chop_channel` table contains carcass counts by sex and chop/tagged

### Recovery

The `chop_recovery` table contains recovered carcass counts by color

### Survey

The `chop_header` table contains survey metadata and covariates


## Explore Numeric Variables: {.tabset}

**Numeric Summary of `chops`, `tags` over Period of Record**

```{r}
summary(chop_channel_join$chops)
```

```{r}
summary(chop_channel_join$tags)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$chops))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$tags))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$chops))/nrow(chop_channel_join), 3) * 100` % of values in the `chops` column are NA.
* `r round(sum(is.na(chop_channel_join$tags))/nrow(chop_channel_join), 3) * 100` % of values in the `tags` column are NA.

**Plotting chops over Period of Record**

```{r}
 
chop_channel_join %>% 
  ggplot(aes(x = date, y = chops)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```


**Plotting tags over Period of Record**

```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = tags)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Date", 
       y = "Tags") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Comparing tags and chops**
```{r}
chop_channel_join %>% 
  group_by(date) %>% 
  summarise(chops_total = sum(chops, na.rm = T),
            tags_total = sum(tags, na.rm = T)) %>%
  pivot_longer(cols = c(chops_total, tags_total), names_to = "tags_or_chops", values_to = "count") %>% 
  ggplot(aes(x = date, y = count, fill = tags_or_chops)) +
  geom_col() + 
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

Tags appear to be less commonly applied compared to chops.

```{r}
# Filter clean data to show only numeric variables 
chop_recovery_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

```{r}
summary(chop_recovery_join$id)
```
```{r}
summary(chop_recovery_join$recov_count)
```

* `r round(sum(is.na(chop_recovery_join$recov_count))/nrow(recov_count), 3) * 100` % of values in the `recov_id` column are NA.


```{r}
chop_recovery_join %>% 
  ggplot(aes(x = date, y = recov_count)) + 
  geom_col() +
  theme_minimal()
```


```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(chop_header$id)
```

## Explore Categorical variables: {.tabset}

```{r}
chop_channel_join %>% 
  select_if(is.character) %>%
  colnames()
```
```{r}
chop_recovery_join %>% 
  select_if(is.character) %>% 
  colnames()
```

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```

## Clean data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_channel_cleaner <- chop_channel_join %>%
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_channel_cleaner
```

```{r}
chop_recovery_cleaner <- chop_recovery_join %>% 
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_recovery_cleaner
```

```{r}
chop_header_cleaner <- chop_header %>%
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>%
  mutate(crew = str_replace_all(crew, ",", ""),
         crew = str_replace_all(crew, " ",",")) %>%
  select(-c(date, time))

chop_header_cleaner
```

## Data Dictionaries

# Channel
```{r}
percent_na <- chop_channel_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("ID",
                                          "Color of tag applied to carcass",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped",
                                          "Carcass that were tagged",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
percent_na <- chop_recovery_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("ID",
                                          "Color of tag recovered from carcass",
                                          "Count of recovery",
                                          "Channel ID",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Survey

```{r}

percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
counts_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))
# 
kable(counts_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)

```{r}
# Name file [watershed]_[data type].csv
```

