---
title: "Feather Carcass QC 2000"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```

# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

Read in data from google cloud, glimpse raw data and domain description sheet:
```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2000.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2000.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2000.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopEnv_raw <- sqlFetch(con, "ChopEnvTBL")
  RecovTag_raw <- sqlFetch(con, "RecovTagTBL")
  ChopRecov_raw <- sqlFetch(con, "ChopRecovTBL")
  TagCol <- sqlFetch(con, "TagColLU")
  TagData <- sqlFetch(con, "TagDataTBL")
  TagEnv <- sqlFetch(con, "TagEnvTBL")
  cwt <- sqlFetch(con, "cwtTBL")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  # TagColLU contains metadata about tag colors
  # Jason's Week #'s contains information about week # and date
  ChopEnv_raw <- mdb.get(filepath, tables = "ChopEnvTBL")
  RecovTag_raw <- mdb.get(filepath, tables = "RecovTagTBL")
  ChopRecov_raw <- mdb.get(filepath, tables = "ChopRecovTBL")
  TagCol <- mdb.get(filepath, tables = "TagColLU")
  TagData <- mdb.get(filepath, tables = "TagDataTBL")
  TagEnv <- mdb.get(filepath, tables = "TagEnvTBL")
  cwt <- mdb.get(filepath, tables = "cwtTBL")
  detach(package:Hmisc)
}

write_csv(ChopEnv_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopEnv_2000.csv"))
write_csv(RecovTag_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "RecovTag_2000.csv"))
write_csv(ChopRecov_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2000.csv"))
write_csv(TagCol, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "TagCol_2000.csv"))
write_csv(TagData, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "TagData_2000.csv"))
write_csv(TagEnv, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "TagEnvTBL_2000.csv"))
write_csv(cwt, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "cwt_2000.csv"))
```

## Raw Data Glimpse: {.tabset}

### ChopEnv_Raw
Heading information for bulk chops data.
```{r}
# Read Data
ChopEnv_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopEnv_2000.csv"),
                                 col_types = list("d", "c", "T", "c", "T", "c", "c")) |>  
  glimpse()
```

### ChopRecov_raw
Recovered chop counts with no sampling information.
```{r}
#Data collection for Chopping and Spawning to determine how many carcasses are being captured (link by recovID)
ChopRecov_raw <-read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopRecov_2000.csv")) |>  
  rename("Number Spawned" = Spawn) |>  
  glimpse()
```

### RecovTag_Raw
```{r}
#Data for Recovered Tags
#What is RelStat and MultRecov? 
RecovTag_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "RecovTag_2000.csv")) |>  
  glimpse()
```

### TagData_raw
```{r}
#Data for Tagging DataSheet - when export access table to excel, Tag Number is automatically changed to TagID - converting it back here
TagData_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "TagData_2000.csv")) |>  
  rename("TagNum" = "TagID") |> 
  glimpse()
```

### TagEnv_raw
```{r}
#Heading Info for Tagging Data Sheet
TagEnv_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "TagEnvTBL_2000.csv")) |>  
  glimpse()
```

### cwt_raw
This table contains all carcasses with sampling information.

```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "cwt_2000.csv")) |>  
  glimpse()
```

## Data transformations: {.tabset}

### CWT

The `cwt` table contains sampling information on individual carcasses. It is not joined with the other tables.

```{r}
cwt <- cwt_raw |>  
  clean_names() |>
  mutate(disposition = "chopped",
         source_table = "CWT",
         count = 1) |> 
  glimpse()
```

### Recovered and chopped carcasses

The `chop_recovery_join` table contains recovered carcass counts. Add `disposition` column to mark as chops.

```{r}
#1. Link ChopRecov with ChopEnv to get the date
chop_recovery_join <- full_join(ChopEnv_raw |>  
                                  select(ChopEnvID, Date, Start, TagCol), ChopRecov_raw, 
                                by = "ChopEnvID") |>  
  clean_names() |>  
  rename(male_chop = "adult_m",
         female_chop = "adult_f",
         grilse_male = "gril_m",
         grilse_female = "gril_f") |>  
  pivot_longer(c(male_chop, female_chop, grilse_male, grilse_female), 
               values_to = "count",
               names_to = "sex") |> 
  mutate(lifestage = if_else(str_detect(sex, "grilse"), "grilse", "adult"),
         sex = case_when(sex %in% c("male_chop", "grilse_male") ~ "male", 
                         sex %in% c("female_chop", "grilse_female") ~ "female"),
         disposition = "chopped_recovered",
         source_table = "ChopRecov") |> 
  glimpse()
```

### Tags

The `tag_join` table contains tagged carcasses with survey metadata added.

```{r}
tag_join <- left_join(TagData_raw, TagEnv_raw,
                         by = "TagEnvID") |> 
  mutate(disposition = "tagged",
         source_table = "TagData",
         count = 1) |> 
  clean_names() |> 
  glimpse()
# tag_join <- left_join(TagData_raw, RecovTag_raw) |>  
#   clean_names() |>  
#   mutate(disposition = "tagged") |> 
#   select(c(tag_num, sect, unit, fl, sex, egg_ret, recov_id, disposition)) |>  glimpse()
```

### Recovered tagged carcasses

The `tag_recovery` table contains recovered tagged carcasses. Add `disposition` column to take `rel_stat` column and convert to `recovered`, `chopped`, or `NA`. Rename `mult_recov` to `count`.
```{r}
tag_recovery <- RecovTag_raw |> 
  clean_names() |> 
  mutate(disposition = case_when(rel_stat == "R" ~ "recovered",
                                 rel_stat == "C" ~ "chopped",
                                 rel_stat == "X" ~ NA_character_),
         source_table = "RecovTag") |> 
  rename(count = mult_recov) |> 
  select(-rel_stat) |> 
  glimpse() 
```
### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopEnv_raw |>  
  clean_names() |>  glimpse()
```



### Combine tables
```{r}
all_carcasses <- bind_rows(tag_join, chop_recovery_join, cwt, tag_recovery) |> glimpse()
```
## Explore Numeric Variables: {.tabset}

### All Carcasses Variables: 

### Chop Recovery Variable: `tag_num`, `sect`, `unit`, `fl`, `float`, `tag_env_id`, `rel_loc2`, `week`, `chop_env_id`, `count`, `number_spawned`, `recov_id`, `cwt_id`, `ht_num`, `recov_tag_id`

```{r}
# Filter clean data to show only numeric variables 
all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
```

```{r}
summary(all_carcasses$tag_num)
summary(all_carcasses$sect)
summary(all_carcasses$unit)
summary(all_carcasses$fl)
summary(all_carcasses$float)
summary(all_carcasses$tag_env_id)
summary(all_carcasses$rel_loc2)
summary(all_carcasses$week)
summary(all_carcasses$chop_env_id)
summary(all_carcasses$number_spawned)
summary(all_carcasses$recov_id)
summary(all_carcasses$cwt_id)
summary(all_carcasses$ht_num)
summary(all_carcasses$recov_tag_id)
```


### All Carcasses Variable: `count`

```{r}
summary(all_carcasses$count)
```

** Plot carcass counts by sex **

```{r}
all_carcasses |>  
  mutate(sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         sex == "X" ~ "unknown")) |> 
  group_by(date, sex) |>  
  summarize(count = sum(count, na.rm = T)) |> 
  ggplot(aes(x = date, y = count, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```
This plot shows carcass over period of record by sex and is helpful for understanding the seasonality of the survey, i.e. chop counts peak in Oct-Nov. Because some chops are recorded without sampling information, much of the data has sex recorded as `NA`.

**Plotting Total Adult Chops over Period of Record**

```{r}
total_adult_chops <- all_carcasses |> 
  filter(disposition %in% c("chopped_recovered", "chopped", "recovered")) |> 
  filter(lifestage != "grilse") |> 
  select(date, count, disposition, sex) |> 
  glimpse()

total_adult_chops_summary <- total_adult_chops |> 
  group_by(date, sex) |> 
  summarise(total_adult_chops = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_adult_chops_summary |>
  ggplot(aes(x = date, y = total_adult_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

This plot shows the sex proportion of adult chops for each day over the the period of 2000-09-05 to 2000-12-14. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. 


**Plotting grilse chops over Period of Record**

```{r}
total_grilse_chops <- all_carcasses |> 
  filter(disposition %in% c("chopped_recovered", "chopped", "recovered")) |> 
  filter(lifestage == "grilse") |> 
  select(date, count, disposition, sex) |> 
  glimpse()

total_grilse_chops_summary <- total_grilse_chops |> 
  group_by(date, sex) |> 
  summarise(total_grilse_chops = sum(count, na.rm = T)) |> 
  glimpse()

total_grilse_chops_summary |>
  ggplot(aes(x = date, y = total_grilse_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2"))
```
This plot shows the sex proportion of grilse chops for each day over the the period of 2000-09-05 to 2000-12-14. There are more data gaps for the grilse data compared to the normal chops data. All of the grilse chop data have been identified for their sexes - there is a significant higher proportion of males compared to the females.


**Plotting all carcasses over Period of Record**
Carcasses in the `all_carcasses` dataset are marked as either `tagged`, `chopped_recovered`, `recovered` and `chopped`. `chopped` carcasses have come from the `CWT` data table and contain individual-level sampling information (i.e. sex, fork length). `chopped_recovered` carcasses come from the `ChopRecov` data table and have sex and lifestage information, but no other individual-level sampling information. `tagged` carcasses come from the `TagData` data table and have individual-level sampling information. `recovered` carcasses come from the `RecovTag` data table and do not have individual-level sampling information, though they could potentially be joined by `recovID.`

```{r}
total_carcasses_summary <- all_carcasses |> 
  mutate(sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male")) |> 
  group_by(date, disposition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_carcasses_summary |> 
  ggplot(aes(x = date, y = total_carcasses, fill = disposition)) + 
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```
This plot shows the proportion of the carcasses that were either chopped, tagged, or recovered and chopped on a given day over the period of 2000-09-05 to 2000-12-14. 
 
```{r}
summary(chop_recovery_join$number_spawned)
```

* `r round(sum(is.na(chop_recovery_join$number_spawned))/nrow(chop_recovery_join), 3) * 100` % of values in the `number_spawned` column are NA.

**Plotting number_spawned over a Period of Record**

```{r}
all_carcasses |>  
  group_by(date) |>  
  summarize(number_spawned = sum(number_spawned, na.rm = T)) |>  
  ggplot(aes(x = date, y = number_spawned)) +
  geom_col(fill = "blue") + 
  labs(x = "Date", 
       y = "spawn") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```
This plot shows the number spawned for each day over the the period of 2000-09-05 to 2000-12-14. Data on number spawned was not present for all dates and was only collected in the `ChopRecov` table, but was collected on 09-05-2000 and 09-21-2000 (seen in the two columns). 


### Individual-level sampling information

**Plotting fork length of each sex**

```{r}
summary(all_carcasses$fl)
```

```{r}
all_carcasses |> 
  mutate(sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         sex == "X" ~ "unknown")) |> 
  ggplot(aes(x = sex, y = fl)) +
  geom_boxplot() +
  facet_wrap(~disposition) +
  theme_minimal() + 
  labs(y = "fork length (mm)", x = "sex")
```

**Plotting sex proportion of counts within each disposition**

```{r}
all_carcasses |> 
  mutate(sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         sex == "X" ~ "unknown")) |> 
  ggplot(aes(y = count, x = disposition, fill = sex)) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Disposition") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Chop Header Variable: `chop_env_id`, `week`, 

```{r}
chop_header |>  
  select_if(is.numeric) |>  
  colnames()
```
```{r}
summary(chop_header$chop_env_id)
```

```{r}
summary(chop_header$week)
```


## Explore Categorical variables: {.tabset}

### All Carcasses

Fix inconsistencies with spelling, capitalization, and dates

```{r}
all_carcasses |>  
  select_if(is.character) |>  
  colnames()
```

```{r}
unique(all_carcasses$sex)
unique(all_carcasses$egg_ret)
unique(all_carcasses$rel_loc)
unique(all_carcasses$recorder)
unique(all_carcasses$tag_col)
unique(all_carcasses$disposition)
unique(all_carcasses$source_table)
unique(all_carcasses$lifestage)
```

```{r}
all_carcasses_cleaner <- all_carcasses |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(sex = case_when(sex == "f" ~ "female",
                         sex == "m" ~ "male",
                         sex == "x" ~ "unknown"),
         egg_ret = case_when(egg_ret == "u" ~ "unspawned",
                             egg_ret == "s" ~ "spawned",
                             egg_ret == "x" ~ "unknown",
                             egg_ret == "p" ~ "prespawned",
                             TRUE ~ egg_ret)) |> 
  select(-c(start, cwt_code)) |>  # start has time data and is removed because no stop time provided so information is not useful. cwt_code is all NAs
  glimpse()
```

### Chop Header Clean Data
```{r}
chop_header |>  
  select_if(is.character) |>  
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$recorder)
```
```{r}
unique(chop_header$weather)
```
```{r}
#Not sure what turbot is
unique(chop_header$turbot)
```

```{r}
chop_header_cleaner <- chop_header |>  
  mutate(date = as_date(date))|>  
  mutate_if(is.character, str_to_lower) |>  
  mutate(crew = str_replace_all(crew, " ", ""),
         crew = str_replace_all(crew, "/", ","),
         crew = str_replace_all(crew, "&", ","),
         weather = case_when(weather == "ran" ~ "rain",
                             weather == "unk" ~ NA_character_,
                             TRUE ~ weather)) |>  
  select(-c(start, stop, week, turbot)) # stop is all NA and start is not useful without stop; turbot does not have any data

```


## Comments

* We combined all tables - `ChopRecov`, `TagData`, `RecovTag` and `CWT` into `all_carcasses`. Data from `RecovTag` may be associated with `TagData`, so there could be some double counting here (i.e. one carcass represented more than once in the dataset). This could be addressed by filtering to a unique identifier (TagNum?)
* We assigned `disposition` based on the table from which the data came from.
* We assigned spawning condition based on values in egg_ret. May need to confirm.
* We kept `recov_id` because it may be able to connect some tagged individuals with chopped/recovered_chopped individuals.
* What is `float`, `rel_loc`, `rel_loc2`, 'ht_num', and `cwt_id`?

## Data Dictionaries

### All Carcasses
```{r}
percent_na <- all_carcasses_cleaner |> 
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |> 
  pivot_longer(cols = everything())


all_carcasses_data_dictionary <- tibble(variables = colnames(all_carcasses_cleaner),
                          description = c("Tag number",
                                          "Section of survey",
                                          "Unit of survey",
                                          "Fork length (mm)", 
                                          "Sex",
                                          "Egg retention (spawning condition)",
                                          "Float",
                                          "Rel_loc",
                                          "Comment",
                                          "Tag environment ID",
                                          "Rel_loc 2",
                                          "Date of survey",
                                          "Crew member recording data",
                                          "Tag color",
                                          "Week of survey",
                                          "Disposition of carcass",
                                          "Table from which the data came",
                                          "Count",
                                          "Chop environment ID",
                                          "Number spawned",
                                          "Recovery ID",
                                          "Lifestage",
                                          "CWT ID",
                                          "HT Number",
                                          "Recovery Tag ID"),
                          percent_na = round(percent_na$value*100))

kable(all_carcasses_data_dictionary)
```

## Save cleaned data back to google cloud (TBA)

```{r}
all_carcasses_2000 <- all_carcasses_cleaner |> glimpse()
```
```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_carcasses_2000,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2000.csv")
```

