---
title: "Feather Carcass QC 2008"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2008.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2008.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2008.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopChannel <- sqlFetch(con, "ChopChannelTBL1")
  ChopRecov <- sqlFetch(con, "ChopRecovTBL")
  ChopHeader <- sqlFetch(con, "ChopHeaderTBL")
  cwt <- sqlFetch(con, "CWTTagTBL")
  cwt_header <- sqlFetch(con, "CWTHeaderTBL")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  ChopChannel <- mdb.get(filepath, "ChopChannelTBL1")
  ChopRecov <- mdb.get(filepath, "ChopRecovTBL")
  ChopHeader <- mdb.get(filepath, "ChopHeaderTBL")
  cwt <- mdb.get(filepath, "CWTTagTBL")
  cwt_header <- mdb.get(filepath, "CWTHeaderTBL")
  
  # FL by sex is empty
  # TagColLU_06 contains metadata about tag color
  # headtag is empty
  # ChopHallprintTBL contains TagID, sect, fork length, sex, clip, carc_stat, headerID, hallprint1, hallprint2, headtag, otolith, scale
  # ScaleSPLs is empty
  detach(package:Hmisc)
}
```

```{r}
write_csv(ChopChannel, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopChannel_2008.csv"))
write_csv(ChopRecov, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2008.csv"))
write_csv(ChopHeader, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2008.csv"))
write_csv(cwt, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTTag_2008.csv"))
write_csv(cwt_header, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2008.csv"))
```



## Raw Data Glimpse: {.tabset}

### ChopChannel_raw
```{r}
ChopChannel_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopChannel_2008.csv")) |> 
  glimpse()
```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopHeader_2008.csv")) |> 
  glimpse()
```

### ChopRecov_raw
```{r}
ChopRecov_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopRecov_2008.csv")) |> 
  glimpse()
```
### cwt_raw
```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "CWTTag_2008.csv")) |> 
  glimpse()
```

### cwt_header_raw

```{r}
cwt_header_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "CWTHeader_2008.csv")) |> 
  glimpse()
```


## Data transformations: {.tabset}

### CWT
The `cwt` table contains sampling information on individual carcasses. It is not joined with the other tables. Added a `count` column assuming each row represents one individual. This table does not contain `disposition` information (usually in the `carc_stat` variable). Added `source_table` column for tracking. Each `tag_id` is represented only once in this table.

```{r}
cwt <- cwt_raw |>  
  clean_names() |>
  mutate(source_table = "CWT",
         count = 1,
         clip = case_when(clip == "y" ~ "Y",
                          clip == "NO DATA" ~ NA_character_,
                          TRUE ~ clip)) |>  
  rename(adipose_fin_clip = clip) |> 
  select(-c(samp_num, header_id)) |> 
  glimpse()
```

### Bulk chop counts

The `chops_and_tags` table has a chop count and tag count column. In earlier years (2000-2006) this table also contained either information about sex, lifestage, or adipose fin clip but this year no additional information is included. `disposition` column added to distinguish between chopped and tagged counts.

```{r}
chops_and_tags <- left_join(ChopChannel_raw, 
                            ChopHeader_raw |> select(HeaderID = headerID, TagCol, Date, Time),
                            by = "HeaderID") |>
  clean_names() |> 
  pivot_longer(c(chops, tags), 
               values_to = "count",
               names_to = "disposition") |> 
  mutate(disposition = if_else(disposition == "chops", "chopped", "tagged"),
         source_table = "ChopChannel") |> 
  glimpse()
```

### Recovered tags and chops

The `chop_recovery` table contains recovered carcass counts by color. According to Access relationships, `chan_id` connects the header table to the recovery table. Here we get header information from `chops_and_tags` above and join those to the `chop_recovery` table on `chan_id`. Renames `recov_count` to `count`. Adds `disposition` and `source_table` columns.

```{r}
chan_id_lookup <- chops_and_tags |> distinct(chan_id, date, time, sect, chan, min, chan_id, header_id, tag_col)
# HeaderID connects ChopChannel to ChopHeader; ChanID joins ChopChannel to ChopRecov
chop_recovery_join <- left_join(ChopRecov_raw |> rename(chan_id = ChanID),
                                chan_id_lookup, 
                                by = "chan_id") |> 
  clean_names() |> 
  rename(count = recov_count,
         tag_id = recov_id) |> 
  mutate(disposition = "chopped_recovered",
         source_table = "ChopRecov") |> 
  select(-c(chan_id, header_id)) |> 
  glimpse()
```

### Combine tables

```{r}
all_carcasses <- bind_rows(chops_and_tags |> 
                             select(-c(header_id, chan_id)), chop_recovery_join, cwt) |> glimpse()
```

### Survey

The `chop_header` table contains survey metadata and covariates 
```{r}
chop_header <- ChopHeader_raw |> 
  clean_names() |> glimpse()
```


## Explore Numeric Variables: {.tabset}

### All Carcasses Variables: `chan_id`, `sect`, `min`, `header_id`, `tag_col`, `count`, `recov_id`,  `recov_color`, `tag_id`, `fl`, `head_tag`, `scale_sample`

```{r}
# Filter clean data to show only numeric variables 
all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
```

```{r}
summary(all_carcasses$chan_id)
summary(all_carcasses$sect)
summary(all_carcasses$min)
summary(all_carcasses$header_id)
summary(all_carcasses$tag_col)
summary(all_carcasses$recov_id)
summary(all_carcasses$recov_color)
summary(all_carcasses$tag_id)
summary(all_carcasses$fl)
summary(all_carcasses$head_tag)
summary(all_carcasses$scale_sample)
```

### All Carcasses Variable: `count`

```{r}
summary(all_carcasses$count)
```

** Plot carcass counts by sex **

```{r}
all_carcasses |>  
  mutate(sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         sex %in% c("ND", "UK") ~ "unknown",
                         TRUE ~ sex)) |> 
  group_by(date, sex) |>  
  summarize(count = sum(count, na.rm = T)) |> 
  ggplot(aes(x = date, y = count, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  theme(text = element_text(size = 15)) +
  ylim(c(0, 500))
```
This plot shows carcass over period of record by sex and is helpful for understanding the seasonality of the survey, i.e. chop counts peak in Oct-Nov. The population crashed this year (source: Casey Campos, meeting). Because the bulk chops did not have sex information this year, most of the data have sex recorded as `NA`.

**Plotting Total Adult Chops over Period of Record**

```{r}
total_adult_chops <- all_carcasses |> 
  filter(disposition != "tagged") |> 
  select(date, count, disposition, sex) |> 
  glimpse()

total_adult_chops_summary <- total_adult_chops |> 
  group_by(date, sex) |> 
  summarise(total_adult_chops = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_adult_chops_summary |>
  ggplot(aes(x = date, y = total_adult_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

This plot shows the sex proportion of adult chops (no `tagged` carcasses) for each day over the period of `r min(all_carcasses$date, na.rm = T)` to `r max(all_carcasses$date, na.rm = T)`. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. This year there was no sex information.


**Plotting grilse over Period of Record**
This year did not have lifestage information.

**Plotting all carcasses over Period of Record**
Carcasses in the `all_carcasses` dataset are marked as either `tagged`, `chopped_recovered`, or  `chopped`. Carcasses from the `CWT` table are marked as `chopped`, `tagged`, or `NA` and have individual-level sampling information (i.e sex, fork length). `chopped` carcasses from `ChopChannel` do not have individual-level sampling data. `chopped_recovered` carcasses come from the `ChopRecov` data table and have no individual-level sampling information. `tagged` carcasses come from the `ChopChannel` data table and do not have individual-level sampling information.

```{r}
total_carcasses_summary <- all_carcasses |> 
  group_by(date, disposition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_carcasses_summary |> 
  ggplot(aes(x = date, y = total_carcasses, fill = disposition)) + 
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Individual-level sampling information

**Plotting fork length of each sex**

```{r}
summary(all_carcasses$fl)
```

```{r}
all_carcasses |> 
  mutate(sex = case_when(sex %in% c("F", "fem") ~ "female",
                         sex == "M" ~ "male",
                         sex %in% c("UK", "ND") ~ "unknown")) |>
  ggplot(aes(x = sex, y = fl)) +
  geom_boxplot() +
  facet_wrap(~disposition) +
  theme_minimal() + 
  labs(y = "fork length (mm)", x = "sex")
```

**Plotting sex proportion of counts within each disposition**

```{r}
all_carcasses |> 
  mutate(sex = case_when(sex %in% c("F", "fem") ~ "female",
                         sex == "M" ~ "male",
                         sex %in% c("UK", "ND") ~ "unknown")) |> 
  ggplot(aes(y = count, x = disposition, fill = sex)) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Disposition") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Chop Header Variable: `header_id`, `week_num`, `tag_col`

```{r}
chop_header |>  
  select_if(is.numeric) |>  
  colnames()
```
```{r}
summary(chop_header$header_id)
summary(chop_header$week_num)
summary(chop_header$tag_col)
```


## Explore Categorical variables: {.tabset}

### All Carcasses

Fix inconsistencies with spelling, capitalization, and dates

```{r}
all_carcasses |>  
  select_if(is.character) |>  
  colnames()
```

```{r}
unique(all_carcasses$chan)
unique(all_carcasses$disposition)
unique(all_carcasses$source_table)
unique(all_carcasses$sex)
unique(all_carcasses$spawn)
unique(all_carcasses$adipose_fin_clip)
unique(all_carcasses$samp_type)
unique(all_carcasses$rel_loc)
unique(all_carcasses$run)
```

```{r}
all_carcasses_cleaner <- all_carcasses |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(sex = case_when(sex %in% c("f", "female", "fem") ~ "female",
                         sex == "m" ~ "male",
                         sex %in% c("nd", "uk") ~ "unknown"),
         spawn = case_when(spawn == "u" ~ "unspawned",
                           spawn == "uk" ~ "unknown",
                           spawn %in% c("no data", "nd") ~ NA_character_,
                           spawn == "s" ~ "spawned",
                           TRUE ~ spawn),
         adipose_fin_clip = case_when(adipose_fin_clip %in% c("no data", "na", "uk") ~ NA_character_, 
                                      adipose_fin_clip == "y" ~ "TRUE",
                                      adipose_fin_clip == "n" ~ "FALSE"),
         adipose_fin_clip = as.logical(adipose_fin_clip)) |>
  select(-c(time)) |>  # time is removed because no stop time provided so information is not useful. chan is all NAs
  glimpse()
```

### Chop Header Clean Data

```{r}
chop_header |>  
  select_if(is.character) |>  
  colnames()
```
```{r}
unique(chop_header$crew)
unique(chop_header$recorder)
unique(chop_header$weather)
```


```{r}
chop_header_cleaner <- chop_header |> 
  mutate_if(is.character, str_to_lower) |>  
  mutate(crew = str_replace_all(crew, " ", ""),
         crew = str_replace_all(crew, "/", ","),
         crew = str_replace_all(crew, "&", ","),
         weather = case_when(weather == "ran" ~ "rain",
                             weather == "cld" ~ "cloudy",
                             TRUE ~ weather)) |>  
  select(-time) # stop is all NA and start is not useful without stop

```

## Comments

* Table layout/structure same as 2002-2007.
* `Hallprint` variable in `cwt` (began collecting in 2004).
* `run` variable available in `cwt` (first year it has shown up)
* `ChopChannel` table this year does not contain sex, lifestage, or fin clip information.
* what is `samp_num`, `rel_loc`, `head_tag`?
* can we join recovered tags on a specific `id` value?

## Data Dictionaries

### All Carcasses
```{r}
percent_na <- all_carcasses_cleaner |> 
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |> 
  pivot_longer(cols = everything())


all_carcasses_data_dictionary <- tibble(variables = colnames(all_carcasses_cleaner),
                          description = c("Section",
                                          "Channel",
                                          "Minute",
                                          "Tag color",
                                          "Date of survey",
                                          "Disposition",
                                          "Count",
                                          "Table from which the data came",
                                          "Tag ID",
                                          "Recovered tag color",
                                          "Fork length (mm)",
                                          "Sex",
                                          "Spawning condition",
                                          "Adipose fin clip status",
                                          "Sample type",
                                          "Head Tag",
                                          "Comment",
                                          "Hall Print 1",
                                          "Hall Print 2", 
                                          "Scale sample",
                                          "Run"),
                          percent_na = round(percent_na$value*100))

kable(all_carcasses_data_dictionary)
```

## Save cleaned data back to google cloud (TBA)

```{r}
all_carcasses_2008 <- all_carcasses_cleaner |> glimpse()
```

```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_carcasses_2008,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2008.csv")
```
