---
title: "Feather Carcass QC 2010"
author: "Jonathan Fernandez"
date: "06/17/2022"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)

```

# [Watershed] [Data type long version]

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Google Cloud Set up
# Sys.setenv("GCS_AUTH_FILE" = "config.json")
# Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")
# gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
# gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/ChopChannelTBL1.xlsx",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "ChopChannelTBL1.xlsx",
#                overwrite = TRUE)
# 
# gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/ChopHeaderTBL.xlsx",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "ChopHeaderTBL.xlsx",
#                overwrite = TRUE)
# 
# gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/ChopRecovTBL.xlsx",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "Cç",
#                overwrite = TRUE)

# Read Data
ChopChannel_raw <- read_excel("ChopChannelTBL1.xlsx") %>% 
  rename("ID" = HeaderID)
ChopHeader_raw <- read_excel("ChopHeaderTBL.xlsx") %>% 
  rename("ID" = headerID)
ChopRecov_raw <-read_excel("ChopRecovTBL.xlsx")
```
Read in data from Microsoft Access, glimpse raw data and domain description sheet: 
```{r}
# TODO: Attempt, but not needed
library(RODBC)
feather_camp <- odbcConnectAccess2007("CAMP_Escapement_20210412.mdb")
```
## Data transformations

```{r}
# TODO: Clean Names
# TODO: Consider combining tables
# Snake case, 
# Columns are appropriate types
# Remove redundant columns

# Checking Join
# chopTBL_test1 <- left_join(ChopChannel_raw, ChopHeader_raw, by = "ID")
# chopTBL_test2 <- full_join(chopTBL_test1, ChopRecov_raw, by = "ChanID")

chopTBL <- left_join(ChopChannel_raw, ChopHeader_raw, by = "ID") %>%
  full_join(ChopRecov_raw, by = "ChanID")

cleaner_chopTBL <- chopTBL %>% 
  clean_names() %>% 
  glimpse()

# TODO: Any columns to remove? Check in with Ashley on this.
 # cleaner_chopTBL %>% 
 #  select(-sect,chan,crew,week_num) %>%
 #  glimpse()
```

## Data Dictionary
```{r}
percent_na <- cleaner_chopTBL %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_chopTBL),
                          description = c("Channel ID", # TODO: Identify linkage, Lat/long?
                                          "Section", # TODO: Identify encoding
                                          "Channel", # TODO: Identify encoding
                                          "Minimum", # What exactly?
                                          "Head ID", # TODO: Rename to head_ID
                                          "Observed unknown sex chops",
                                          "Observed unknown sex tags",
                                          "Observed male chops",
                                          "Observed female chops",
                                          "Observed female tags",
                                          "Observed male tags",
                                          "Color of tag applied to carcass", # color?
                                          "Date",
                                          "Initials of crew conducting survey", 
                                          "Time of Day",
                                          "Initials of person recording data",
                                          "Weather (sunny, cloudy, rainy)",
                                          "Comments",
                                          "Week Number"),
                          percent_na = round(percent_na$value*100))
kable(data_dictionary)
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
cleaner_chopTBL %>% 
  select_if(is.numeric) %>%
  colnames()
```

### Variable: `female_chop`, `male_chop`,`unknown_chop`

**Numeric Summary of `female_chop`, `male_chop`,`unknown_chops` over Period of Record**

```{r}
summary(cleaner_chopTBL$female_chop)
```

```{r}
summary(cleaner_chopTBL$male_chop)
```

```{r}
summary(cleaner_chopTBL$unknown_chops)
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

NA and Unknown Values
  - <1 % of values in the female_chop column are NA.
  - <1 % of values in the male_chop column are NA.
  - <1 % of values in the unknown_chops column are NA.

**Plotting female_chop over Period of Record**

### TODO: compare Male, female, unknowntogether -> histogram?

```{r}
# Make whatever plot is appropriate 
filter(cleaner_chopTBL) %>% 
  ggplot(aes(x = female_chop, y = date)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Female Chop", 
       y = "Date") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```


**Plotting male_chop over Period of Record**
```{r}
# Make whatever plot is appropriate 
filter(cleaner_chopTBL) %>% 
  ggplot(aes(x = male_chop, y = date)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Male Chop", 
       y = "Date") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting unknown_chops over Period of Record**
```{r}
# Make whatever plot is appropriate 
filter(cleaner_chopTBL) %>% 
  ggplot(aes(x = unknown_chops, y = date)) +
  geom_point(size = 1.4, alpha = .5, color = "black") + 
  labs(x = "Unknown Chops", 
       y = "Date") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting female_tag over Period of Record**

```{r}
# Make whatever plot is appropriate 
filter(cleaner_chopTBL) %>% 
  ggplot(aes(x = female_tag, y = date)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Female Tag", 
       y = "Date") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting male_tag over Period of Record**

```{r}
# Make whatever plot is appropriate 
filter(cleaner_chopTBL) %>% 
  ggplot(aes(x = male_tag, y = date)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Male Tag", 
       y = "Date") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting unknown_tags over Period of Record**
```{r}
# Make whatever plot is appropriate 
filter(cleaner_chopTBL) %>% 
  ggplot(aes(x = unknown_tags, y = date)) +
  geom_point(size = 1.4, alpha = .5, color = "black") + 
  labs(x = "Unknown Tags", 
       y = "Date") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
# Filter clean data to show only categorical variables
```


### Variable: `[name]`
```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values



## Summary of identified issues

* List things that are funcky/bothering us but that we don't feel like should be changed without more investigation

## Save cleaned data back to google cloud 

```{r}
# TODO: Request Permission Access
# Write to google cloud 
# Name file [watershed]_[data type].csv
```
