---
title: "Feather Carcass QC 2017"
author: "Elizabeth Stebbins"
date: '2022-12-21'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```

# Feather River Carcass Data

## Description of Monitoring Data

See Feather Carcass Survey QC Document

## Access Cloud Data

```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# get data and save as xlsx
```

Read in data from google cloud, glimpse raw data and domain description sheet. 
```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2017_2021/CAMP_Escapement_20210412.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2017_2021.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2017_2021.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopHeader_raw <- sqlFetch(con, "CarcassSurvey")
  Chops_raw <- sqlFetch(con, "CarcassChops")
  Carcass_individual_raw <- sqlFetch(con, "CarcassIndividual")
  ChopRecov_raw <- sqlFetch(con, "CarcassRecover")
  disposition_LU = sqlFetch(con, "luDisposition")
  color_LU = sqlFetch(con, "luColor")
  condition_LU = sqlFetch(con, "luCondition")
  cwtstatus_LU = sqlFetch(con, "luCWTStatus")
  spawnstatus_LU = sqlFetch(con, "luSpawnStatus")
  marktype_LU = sqlFetch(con, "luMarkType")
  species_LU = sqlFetch(con, "luSpecies")
  adfin_LU = sqlFetch(con, "luAdFinStatus")
  choptype_LU = sqlFetch(con, "luChopType")
  sex_LU = sqlFetch(con, "luSex")
  weather_LU = sqlFetch(con, "luWeather")
  run_LU = sqlFetch(con, "luRun")
  choptype_LU = sqlFetch(con, "luChopType")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  Chops_raw <- mdb.get(filepath, "CarcassChops")
  CWT_raw <- mdb.get(filepath, "CarcassIndividual")
  ChopRecov_raw <- mdb.get(filepath, "CarcassRecover")
  ChopHeader_raw <- mdb.get(filepath, "CarcassSurvey")
  # lookup tables
  disposition_LU = mdb.get(filepath, "luDisposition")
  color_LU = mdb.get(filepath, "luColor")
  condition_LU = mdb.get(filepath, "luCondition")
  cwtstatus_LU = mdb.get(filepath, "luCWTStatus")
  spawnstatus_LU = mdb.get(filepath, "luSpawnStatus")
  marktype_LU = mdb.get(filepath, "luMarkType")
  species_LU = mdb.get(filepath, "luSpecies")
  adfin_LU = mdb.get(filepath, "luAdFinStatus")
  choptype_LU = mdb.get(filepath, "luChopType")
  sex_LU = mdb.get(filepath, "luSex")
  weather_LU = mdb.get(filepath, "luWeather")
  run_LU = mdb.get(filepath, "luRun")
  location_LU = mdb.get(filepath, "Location")
  subsample_LU = mdb.get(filepath, "luSubsample")
  # Other tables include luActive, luBoat,
  # luBodyPart, luCentralValleyStreams, 
  # luDates, luEstimator, luLocationType, luMarkType,
  # luOtherMarkDisposition, luOtherObservations, luPeople, luProject,
  # luSizeClass, luSubsample,
  # OptionLocationSubSections, OptionSelections, CarcassOtherMarks,
  # SurveyDocuments, TrapCWTS, TrapOtherSpecies, TrapSurvey, Version, 
  # luNoYes, luTransportType, TrapSalmon, CarcassOtherObsv,
  # SurveyMetadata
  detach(package:Hmisc)
}

write_csv(Chops_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "Chops_2017_2021.csv"))
write_csv(ChopHeader_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2017_2021.csv"))
write_csv(ChopRecov_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2017_2021.csv"))
write_csv(CWT_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWT_2017_2021.csv"))

```

## Raw Data Glimpse: {.tabset}

### Chops_raw
```{r, message = FALSE}
Chop_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river","Chops_2017_2021.csv")) |> 
  glimpse()
```

### ChopHeader_raw

```{r, message = FALSE}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river","ChopHeader_2017_2021.csv")) |>
  mutate(Date = as.Date(SurveyDate)) |> 
  select(-SurveyDate) |> 
  filter(year(Date) >= 2017) |> 
  glimpse()
```

### ChopRecov_raw

```{r, message = FALSE}
ChopRecov_raw <- read_csv(here::here("data-raw", "qc-markdowns", 
                                  "adult-holding-redd-and-carcass-surveys",
                                  "feather-river", "ChopRecov_2017_2021.csv")) |> 
  glimpse()
```

### CWT raw

```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river","CWT_2017_2021.csv")) |>
  glimpse()
```


## Data transformations: {.tabset}

Notes about 2017 data and onwards:
* Starting in 2017, Feather River began storing data in the CAMP database
* Prior to 2021, Feather River carcass surveys would recapture a carcass multiple times before chopping (after 2021, they moved to chopping upon recapture)
* `ChopHeader` now contains environmental variables, but they are empty
* The `CWT` table contains individual carcasses, chopped and tagged
* The `chop` table contains bulk carcass counts by week
* The `chop_recovery` table contains recovered carcass counts by color


### Survey

The `chop_header` table contains survey metadata and covariates. The `chop_header` table does not contain an entry for `survey_meta_id == 70002`, which is included in many of the raw tables.

* Joined lookup tables and removed empty columns

```{r}
chop_header <- ChopHeader_raw |> 
  clean_names() |> 
  left_join(location_LU |> 
              select(LocationID, Location),
            by = c("location_id" = "LocationID")) |> 
  left_join(weather_LU |> select(WeatherID, Weather),
            by = c("weather_id" = "WeatherID")) |> 
  left_join(subsample_LU |> select(SubsampleID, Subsample),
            by = c("subsample_id" = "SubsampleID")) |> 
  mutate(time_in = as.POSIXct(gsub("[()]", "", time_in), 
                                format = "%m/%d/%y %H:%M:%S"),
         time_out = as.POSIXct(gsub("[()]", "", time_out), 
                                    format = "%m/%d/%y %H:%M:%S"),
         edit_time = as.POSIXct(gsub("[()]", "", edit_time), 
                                format = "%m/%d/%y %H:%M:%S"),
         creation_time = as.POSIXct(gsub("[()]", "", creation_time), 
                                    format = "%m/%d/%y %H:%M:%S")) |> 
  select(-c(location_id, transport_type_id, subsample_id, 
            weather_id, water_temp_f, water_temp_c, 
            secchi_cm, secchi_feet, flow_cfs, gps_unit,
            turbidity_ntu)) |> 
  clean_names() |> 
  glimpse()
```

### CWT

The `cwt` table contains sampling information on individual carcasses. This table also contains sex, spawning condition, and adipose fin clip status. It is not joined with the other tables. Added a `count` column assuming each row represents one individual. This table contains `disposition` information in the `recapture_or_chop` variable (in previous years this was the `carc_stat` variable, and in 2012 this was in three separate recap ID columns). Added `source_table` column for tracking.

* removed columns that were all `NA`: `way_pt`, `r_mile`, `color_tag_applied_id`, `dn_anu`, `transport_type_id`, `gps_unit`, `turbidity_ntu`, `water_temp_f`, `water_temp_c`, `secchi_cm`, `secchi_feet`, `flow_cfs`, `tissue_nu`

```{r}
cwt <- left_join(cwt_raw |> clean_names(), 
                 chop_header |> select(-c(data_recorder, comments, 
                             creation_time, editor, edit_time)),
                 by = c("survey_meta_id", "survey_id")) |> 
  rename(fl_mm = "f_lmm",
         fl_cm = "f_lcm",
         cwt_cd = "cw_tcd") |> 
  select(-c(way_pt, r_mile, color_tag_applied_id,
            dn_anu, tissue_nu)) |> 
  # add lookup tables
  left_join(species_LU |> 
              mutate(species_id = as.numeric(TaxonID),
                     species = CommonName) |> 
              select(species_id, species), by = "species_id") |>
  left_join(run_LU |> select(runID, run), 
            by = c("run_id" = "runID")) |>
  left_join(disposition_LU |> select(DispositionID, 
                                     disposition = Disposition), 
            by = c("disposition_id" = "DispositionID")) |> 
  left_join(condition_LU |> 
              select(ConditionID, condition = Condition), 
            by = c("condition_id" = "ConditionID")) |> 
  left_join(spawnstatus_LU |> select(SpawnedID, spawned = Spawned), 
            by = c("spawned_id" = "SpawnedID")) |> 
  left_join(adfin_LU |> select(AdFinClipID, ad_fin_clip_status = AdFinClip),
            by = c("ad_fin_clip_id" = "AdFinClipID")) |> 
  left_join(cwtstatus_LU |> select(CWTStatusID, cwt_status = CWTStatus),
            by = c("cwt_status_id" = "CWTStatusID")) |> 
  left_join(sex_LU |> select(SexID, sex = Sex), 
            by = c("sex_id" = "SexID")) |> 
  select(-c(condition_id, ad_fin_clip_id, sex_id, cwt_status_id, 
            spawned_id, disposition_id, run_id, species_id)) |> 
  clean_names() |> 
  mutate(source_table = "CWT",
         count = 1,
         edit_time = as.POSIXct(gsub("[()]", "", edit_time), 
                                format = "%m/%d/%y %H:%M:%S"),
         creation_time = as.POSIXct(gsub("[()]", "", creation_time), 
                                    format = "%m/%d/%y %H:%M:%S")) |> 
  glimpse()
```

### Bulk chop counts

The `chops` table has bulk chop counts by week. There is no additional information. `disposition` column added assuming all were `chopped`.

```{r}
chop_join <- left_join(Chop_raw |> clean_names(), 
                 chop_header |> select(-c(data_recorder, comments, 
                             creation_time, editor, edit_time)),
                                by = c("survey_meta_id", "survey_id")) |> 
  left_join(species_LU |> 
            mutate(species_id = as.numeric(TaxonID),
                   species = CommonName) |> 
            select(species_id, species), by = "species_id") |>
  left_join(run_LU |> select(runID, run), 
            by = c("run_id" = "runID")) |>
  select(-c(species_id, run_id, r_mile, way_pt)) |> 
  rename(count = chop_count,
         disposition = chop_type_cd) |> 
  mutate(edit_time = as.POSIXct(gsub("[()]", "", edit_time), 
                                format = "%m/%d/%y %H:%M:%S"),
         creation_time = as.POSIXct(gsub("[()]", "", creation_time), 
                                    format = "%m/%d/%y %H:%M:%S"),
         source_table = "ChopJoin") |> 
  glimpse()
```


### Recovered tags and chops

Added a `count` column here assuming each row represents one individual. 

```{r}
chop_recovery_join <- left_join(ChopRecov_raw |> clean_names(), 
                                chop_header |> select(-c(data_recorder, 
                                                         comments,
                                                         creation_time,
                                                         editor, edit_time)),
                                by = c("survey_meta_id", "survey_id")) |> 
  left_join(disposition_LU |> select(DispositionID, disposition = Disposition),
            by = c("disposition_id" = "DispositionID")) |>
  left_join(color_LU |> 
              select(colorID, color), 
            by = c("color_tag_recovered_id" = "colorID")) |>  
  left_join(sex_LU |> select(SexID, sex = Sex),
            by = c("sex_id" = "SexID")) |> 
  mutate(edit_time = as.POSIXct(gsub("[()]", "", edit_time), 
                                format = "%m/%d/%y %H:%M:%S"),
         creation_time = as.POSIXct(gsub("[()]", "", creation_time), 
                                    format = "%m/%d/%y %H:%M:%S"),
         fl_mm = as.numeric(f_lmm),
         fl_cm = as.numeric(f_lcm),
         source_table = "ChopRecov",
         disposition = "recovered",
         count = 1) |> 
  select(-c(f_lmm, f_lcm, way_pt, r_mile, disposition_id,
            color_tag_recovered_id,
            sex_id)) |> 
  rename(tag_col = color) |> 
  glimpse()
```



### Combine tables
* standardize disposition
* all rows with `survey_meta_id == 70002` have no `date` because it is not in the `SurveyMetaID` lookup table.

```{r}
all_carcasses <- bind_rows(chop_join, cwt, chop_recovery_join) |> 
  mutate(disposition = if_else(disposition == "Chop", "Chopped", disposition)) |> 
           glimpse()
```


## Explore Numeric Variables: {.tabset}

### All Carcasses Variables: `survey_meta_id`, `chops_id`, `survey_id`, `count`, `section_id`, `survey_week`, `individual_id`, `disc_tag_applied`, `fl_mm`, `fl_cm`, `head_nu`, `scale_nu`, `otolith_nu`, `cwt_cd`, `recover_id`, `tag_recovered`

```{r}
# Filter clean data to show only numeric variables 
all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
```

```{r}
numeric_cols <- all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
sapply(all_carcasses[,numeric_cols], summary)
```



### All Carcasses Variable: `count`

```{r}
summary(all_carcasses$count)
```

** Plot carcass counts by sex **

```{r}
all_carcasses |>  
  filter(!is.na(date)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  group_by(fake_date, sex, year) |> 
  summarize(count = sum(count, na.rm = T)) |> 
  ggplot(aes(x = fake_date, y = count, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  facet_wrap(~year) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  theme(text = element_text(size = 15)) +
  ylim(c(0, 3000))
```
This plot shows carcass over period of record by sex and is helpful for understanding the seasonality of the survey, i.e. chop counts peak in Oct-Nov.

**Plotting Total Adult Chops over Period of Record**

```{r}
total_adult_chops <- all_carcasses |> 
  filter(disposition != "tagged") |>
  filter(!is.na(date)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  select(fake_date, year, count, disposition, sex) |> 
  glimpse()

total_adult_chops_summary <- total_adult_chops |> 
  group_by(fake_date, year, sex) |> 
  summarise(total_adult_chops = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_adult_chops_summary |> 
  ggplot(aes(x = fake_date, y = total_adult_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  facet_wrap(~year) +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

This plot shows the sex proportion of adult chops (no `tagged` or `pre_carcass_survey` carcasses) for each day over the period of `r min(all_carcasses$date, na.rm = T)` to `r max(all_carcasses$date, na.rm = T)`. The plot is faceted by `year.` The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period.


**Plotting grilse over Period of Record**
2017-2020 do not have lifestage information.

**Plotting all carcasses over Period of Record**
Carcasses in the `all_carcasses` dataset are marked as either `Chopped`, `Tagged`, `recovered`, or `NA`. Carcasses from the `CWT` table are marked as `Tagged` or `Chopped` and have individual-level sampling information (i.e sex, fork length, adipose fin clip status). `chopped` carcasses from `ChopJoin` do not have individual-level sampling data. All carcasses from `ChopRecov` are marked `recovered` and do not have individual-level sampling data.

```{r}
total_carcasses_summary <- all_carcasses |> 
  filter(!is.na(date)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  group_by(fake_date, year, disposition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_carcasses_summary |> 
  ggplot(aes(x = fake_date, y = total_carcasses, fill = disposition)) + 
  facet_wrap(~year) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Individual-level sampling information

**Plotting fork length of each sex**
* major outlier

```{r}
summary(all_carcasses$fl_mm)
```

```{r}
all_carcasses |> 
  filter(fl_mm < 1000) |> 
  ggplot(aes(x = sex, y = fl_mm)) +
  geom_boxplot() +
  facet_wrap(~disposition) +
  theme_minimal() + 
  labs(y = "fork length (mm)", x = "sex")
```

**Plotting sex proportion of counts within each disposition**

```{r}
all_carcasses |> 
  ggplot(aes(y = count, x = disposition, fill = sex)) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Disposition") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Chop Header Variable: `survey_meta_id`, `survey_id`, `section_id`, `survey_week`


```{r}
chop_header |>  
  select_if(is.numeric) |>  
  colnames()
```
```{r}
summary(chop_header$survey_meta_id)
summary(chop_header$survey_id)
summary(chop_header$section_id)
summary(chop_header$survey_week)
```


## Explore Categorical variables: {.tabset}

### All Carcasses

Fix inconsistencies with spelling, capitalization, and dates

```{r}
all_carcasses |>  
  select_if(is.character) |>  
  colnames()
```

```{r}
character_cols <- all_carcasses |>  
  select_if(is.character) |>  
  colnames()

sapply(all_carcasses[, character_cols[! character_cols %in% c("comments", "field_recorder", "crew")]], unique)
```

```{r}
all_carcasses_cleaner <- all_carcasses |> 
  mutate_if(is.character, str_to_lower) |> 
  select(-tag_col) |> # is empty
  mutate(boat = if_else(boat == "n/a", NA_character_, boat)) |> 
  glimpse()
```

### Chop Header Clean Data

```{r}
chop_header |>  
  select_if(is.character) |>  
  colnames()
```
```{r}
unique(chop_header$boat)
unique(chop_header$field_recorder)
unique(chop_header$data_recorder)
unique(chop_header$editor)
unique(chop_header$location)
unique(chop_header$weather)
unique(chop_header$subsample)
```


```{r}
chop_header_cleaner <- chop_header |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(boat = if_else(boat == "n/a", NA_character_, boat)) |> 
  glimpse()
```

## Comments

* These data have the same format for 2017-2020.
* This is a new table/layout structure comopared to 2011-2016. There is a bulk chops table, an individual carcass sampling table (CWT), and a recovered chops table. 
* Same table/layout structure as 2011-2014. There is only a bulk chops table, `ChopJoin`, and a CWT table with individual carcass sampling data.
* No more `hallprint` variable collection (collected for 2004 - 2014 and 2016; not collected in 2015).
* collects `run`
* Using `recovered` instead of `recaptured`.
* How do we know which `id` to use to figure out the Cormack Jolly Seber model structure?

## Data Dictionaries

### All Carcasses
```{r}
percent_na <- all_carcasses_cleaner |> 
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |> 
  pivot_longer(cols = everything())


all_carcasses_data_dictionary <- tibble(variables = colnames(all_carcasses_cleaner),
                          description = c("Header ID",
                                          "Chop ID",
                                          "Survey ID",
                                          "Disposition",
                                          "Count",
                                          "Crew member recording the data in the database",
                                          "Time of data table creation",
                                          "Editor of data table",
                                          "Time of editing",
                                          "Section ID",
                                          "Boat",
                                          "Week of survey",
                                          "Time data collection began",
                                          "Tiem data collection ended",
                                          "Crew member recording data in field",
                                          "Crew present during survey",
                                          "Date of survey",
                                          "Location of survey",
                                          "Weather",
                                          "Sub-sample type",
                                          "Species",
                                          "Run",
                                          "Table from which the data came",
                                          "Individual ID",
                                          "Disc tag applied",
                                          "Fork length (mm)",
                                          "Fork length (cm)",
                                          "Head number",
                                          "Scale number",
                                          "Otolith number",
                                          "Comments",
                                          "CWT code",
                                          "Condition",
                                          "Spawning condition",
                                          "Adipose fin clip status",
                                          "CWT status",
                                          "Sex",
                                          "Recovery ID",
                                          "Tag number recovered"),
                          percent_na = round(percent_na$value*100))

kable(all_carcasses_data_dictionary)
```


## Save cleaned data back to google cloud (TBA)

```{r}
all_carcasses_2017 <- all_carcasses_cleaner |> 
  filter(year(date) == 2017) |> glimpse()

all_carcasses_2018 <- all_carcasses_cleaner |> 
  filter(year(date) == 2018) |> glimpse()

all_carcasses_2019 <- all_carcasses_cleaner |> 
  filter(year(date) == 2019) |> glimpse()

all_carcasses_2020 <- all_carcasses_cleaner |> 
  filter(year(date) == 2020) |> glimpse()
```

```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_carcasses_2017,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2017.csv")

gcs_upload(all_carcasses_2018,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2018.csv")

gcs_upload(all_carcasses_2019,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2019.csv")

gcs_upload(all_carcasses_2020,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2020.csv")
```
