---
title: "Feather Carcass QC 2012"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```
```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2012.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2012.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2012.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopHeader_raw <- sqlFetch(con, "Chop Header")
  Chops_raw <- sqlFetch(con, "Chops")
  cwt_raw <- sqlFetch(con, "CWT")
  cwt_header_raw <- sqlFetch(con, "CWT Header")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  # Switchboard Items, and Sample Weeks are empty. Paste Errors table not relevant.
  Chops_raw <- mdb.get(filepath, "Chops")
  ChopHeader_raw <- mdb.get(filepath, "Chop Header")
  cwt_raw <- mdb.get(filepath, "CWT")
  cwt_header_raw <- mdb.get(filepath, "CWT Header")
  detach(package:Hmisc)
}

write_csv(Chops_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "Chops_2012.csv"))
write_csv(ChopHeader_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2012.csv"))
write_csv(cwt_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWT_2012.csv"))
write_csv(cwt_header_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2012.csv"))
```

## Raw Data Glimpse: {.tabset}

### Chop_raw
```{r}
Chops_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "Chops_2012.csv")) |> 
  rename(Header.ID = `Chop.Header.ID`) |> 
  select(-c(Comments, `Chop.ID`)) |> # all comments are NA
  glimpse()
```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                      "adult-holding-redd-and-carcass-surveys",
                                      "feather-river", "ChopHeader_2012.csv")) |>
  rename(Header.ID = `Chop.Header.ID`) |>
  glimpse()
```

### cwt_raw
```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                               "adult-holding-redd-and-carcass-surveys", "feather-river",
                               "CWT_2012.csv")) |> 
  rename(Header.ID = `CWT.Header.ID`) |> 
  glimpse()
```

### cwt_header_raw

```{r}
cwt_header_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                      "adult-holding-redd-and-carcass-surveys",
                                      "feather-river", "CWTHeader_2012.csv")) |> 
  rename(Header.ID = `CWT.Header.ID` ) |> 
  glimpse()
```




## Data transformations: {.tabset}

Note: for 2011-2016 data, the Access relationships are as follows:
* `CWT$HeaderID` joins to `CWTHeader$HeaderID`
* `CWTHeader$HeaderID` joins to `SampleWeeks$WeekNumber`
* `SampleWeeks$WeekNumber` joins to `ChopHeader$WeekNumber`
* `ChopHeader$HeaderID` joins to `Chops$HeaderID`


### CWT
The `cwt` table contains sampling information on individual carcasses. This table also contains sex, spawning condition, and adipose fin clip status. It is not joined with the other tables. Added a `count` column assuming each row represents one individual. 

This table does not contain `disposition` information in a variable like 2011. Instead, there are three columns for recapture tag IDs. If a tagged carcass was tagged, recaptured, re-released, and recaptured/chopped again, it will be represented multiple times in this table. However, no individuals are marked as `chopped` here because we can't be sure if the last time it was recorded in the table it was re-released or chopped. This is a pretty big assumption in this table.

Added `source_table` column for tracking.

```{r}
cwt <- left_join(cwt_raw |> clean_names(), 
                 cwt_header_raw |> clean_names(),
                 by = "header_id") |>  
  select(-recap_3_id) |> # this column is empty
  pivot_longer(c(tag_id, recap_1_id, recap_2_id),
               names_to = "disposition",
               values_to = "tag_id") |> 
  mutate(disposition = case_when(disposition == "tag_id" ~ "tagged",
                                 disposition %in% c("recap_1_id", 
                                                    "recap_2_id") ~ "recaptured",
                                 TRUE ~ disposition),
         sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         TRUE ~ sex),
         source_table = "CWT",
         count = 1) |> 
  filter(!is.na(tag_id)) |> 
  rename(hallprint1 = hallprint) |> 
  select(-c(morale, lfc, hfc, radio_tag_id, cwt_id, header_id, chop_id, comments, crew))  |> # lfc, hfc, radio tag empty; morale not helpful
  glimpse()

```


### Bulk chop counts

The `chops` table has bulk chop counts by week. There is no additional information. `disposition` column added assuming all were `chopped`.

```{r}
chop_join <- full_join(ChopHeader_raw |> 
                                 select(Header.ID, Date, Week..),
                               Chops_raw,
                       by = "Header.ID") |> 
  clean_names() |> 
  rename(sect = "section",
         min = "minutes",
         count = "total_count") |> 
  mutate(sect = as.numeric(sect),
         disposition = "chopped",
         source_table = "ChopJoin") |> 
  select(-c(header_id)) |> glimpse()
```


### Recovered tags and chops

There is no `chop_recovery` table for this year.

### Combine tables

```{r}
all_carcasses <- bind_rows(chop_join, cwt) |> glimpse()
```

### Survey

The `chop_header` table contains survey metadata and covariates 
```{r}
chop_header <- ChopHeader_raw |> 
  clean_names() |> glimpse()
```


## Explore Numeric Variables: {.tabset}

### All Carcasses Variables: `header_id`, `week`, `sect`, `min`, `count`, `cwt_id`, `river_section`, `chop_id`, `fork_length`, `head_tag`, `scales`, `otoliths`, `hallprint`, `tag_id`

```{r}
# Filter clean data to show only numeric variables 
all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
```

```{r}
numeric_cols <- all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
sapply(all_carcasses[,numeric_cols], summary)
```

### All Carcasses Variable: `count`

```{r}
summary(all_carcasses$count)
```

** Plot carcass counts by sex **

```{r}
all_carcasses |>  
  group_by(date, sex) |> 
  summarize(count = sum(count, na.rm = T)) |> 
  ggplot(aes(x = date, y = count, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  theme(text = element_text(size = 15))
```
This plot shows carcass over period of record by sex and is helpful for understanding the seasonality of the survey, i.e. chop counts peak in Oct-Nov.

**Plotting Total Adult Chops over Period of Record**

```{r}
total_adult_chops <- all_carcasses |> 
  filter(disposition != "tagged") |> 
  select(date, count, disposition, sex) |> 
  glimpse()

total_adult_chops_summary <- total_adult_chops |> 
  group_by(date, sex) |> 
  summarise(total_adult_chops = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_adult_chops_summary |>
  ggplot(aes(x = date, y = total_adult_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

This plot shows the sex proportion of adult chops (no `tagged` carcasses, which in this year were all male) for each day over the period of `r min(all_carcasses$date, na.rm = T)` to `r max(all_carcasses$date, na.rm = T)`. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period.


**Plotting grilse over Period of Record**
This year did not have lifestage information.

**Plotting all carcasses over Period of Record**
Carcasses in the `all_carcasses` dataset are marked as either `tagged`, `recaptured`, or  `chopped`. Carcasses from the `CWT` table are marked as `tagged`, or `recaptured` and have individual-level sampling information (i.e sex, fork length, adipose fin clip status). `chopped` carcasses from `ChopJoin` do not have individual-level sampling data. 

```{r}
total_carcasses_summary <- all_carcasses |> 
  group_by(date, disposition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_carcasses_summary |> 
  ggplot(aes(x = date, y = total_carcasses, fill = disposition)) + 
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Individual-level sampling information

**Plotting fork length of each sex**

```{r}
summary(all_carcasses$fork_length)
```

```{r}
all_carcasses |> 
  ggplot(aes(x = sex, y = fork_length)) +
  geom_boxplot() +
  facet_wrap(~disposition) +
  theme_minimal() + 
  labs(y = "fork length (mm)", x = "sex")
```

**Plotting sex proportion of counts within each disposition**

```{r}
all_carcasses |> 
  ggplot(aes(y = count, x = disposition, fill = sex)) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Disposition") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Chop Header Variable: `header_id`, `week`, `tag_col`

```{r}
chop_header |>  
  select_if(is.numeric) |>  
  colnames()
```
```{r}
summary(chop_header$header_id)
summary(chop_header$week)
```


## Explore Categorical variables: {.tabset}

### All Carcasses

Fix inconsistencies with spelling, capitalization, and dates

```{r}
all_carcasses |>  
  select_if(is.character) |>  
  colnames()
```

```{r}
character_cols <- all_carcasses |>  
  select_if(is.character) |>  
  colnames()

sapply(all_carcasses[, character_cols[! character_cols %in% c("comments", "crew")]], unique)
```

```{r}
all_carcasses_cleaner <- all_carcasses |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(spawning_condition = case_when(spawning_condition == "u" ~ "unspawned",
                                        spawning_condition == "uk" ~ "unknown",
                                        spawning_condition == "s" ~ "spawned",
                                        TRUE ~ spawning_condition),
         adipose_fin_clipped = case_when(adipose_fin_clipped == "y" ~ "TRUE",
                                         adipose_fin_clipped == "n" ~ "FALSE",
                                         adipose_fin_clipped == "uk" ~ NA_character_,
                                         TRUE ~ adipose_fin_clipped),
         adipose_fin_clipped = as.logical(adipose_fin_clipped)) |>
  glimpse()
```

### Chop Header Clean Data

```{r}
chop_header |>  
  select_if(is.character) |>  
  colnames()
```
```{r}
unique(chop_header$crew)
unique(chop_header$weather)
```


```{r}
chop_header_cleaner <- chop_header |> 
  mutate_if(is.character, str_to_lower) |>  
  mutate(crew = str_replace_all(crew, " ", ""),
         crew = str_replace_all(crew, "/", ","),
         crew = str_replace_all(crew, "&", ","),
         weather = case_when(weather %in% c("cld, ran", "cld /ran", "cld. Rain", "cld.rain", "cld,rain") ~ "cloudy, rainy",
                             weather == "blue" ~ "sunny",
                             weather == "cld" ~ "cloudy",
                             weather == "sun/cld/ran" ~ "cloudy, rainy",
                             weather == "ran" ~ "rainy")) |> 
  glimpse()

```

## Comments

* We were unable to identify which individual carcasses in the `cwt` table were `chopped`, so some marked as `recaptured` may actually be `chopped`.
* Same table/layout structure as 2011. There is only a bulk chops table, `ChopJoin`, and a `CWT` table with individual carcass sampling data.
* `Hallprint` variable in `cwt` (began collecting in 2004).
* no more `run` variable (was collected in 2008 and 2009)
* Using `recaptured` instead of `recovered` (which is the value used in previous years) and can change if necessary. Thought it was important to use the variable in original data (original column is "recapture or chop")
* can we join recovered tags on a specific `id` value?

## Data Dictionaries

### All Carcasses
```{r}
percent_na <- all_carcasses_cleaner |> 
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |> 
  pivot_longer(cols = everything())


all_carcasses_data_dictionary <- tibble(variables = colnames(all_carcasses_cleaner),
                          description = c("Date of survey",
                                          "Week of survey",
                                          "Section",
                                          "Minute",
                                          "Count",
                                          "Disposition",
                                          "Table from which the data came",
                                          "River Section",
                                          "Sex",
                                          "Spawning condition",
                                          "Adipose fin clip status",
                                          "Fork length (mm)",
                                          "Head tag number",
                                          "Scales",
                                          "Otoliths",
                                          "Hallprint tag color",
                                          "Hallprint tag number",
                                          "Color of tag",
                                          "Tag ID"),
                          percent_na = round(percent_na$value*100))

kable(all_carcasses_data_dictionary)
```

## Save cleaned data back to google cloud (TBA)

```{r}
all_carcasses_2012 <- all_carcasses_cleaner |> glimpse()
```

```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_carcasses_2012,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2012.csv")
```

