---
title: "Feather Carcass QC 2008"
author: "Inigo Peng"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx


```

```{r}
#Connect to microsoft access through RODBC
# #Use 
# DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
# MDBPATH <- "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/Carcass 2008.mdb"
# PATH <- paste0(DRIVERINFO, "DBQ=", MDBPATH)
# 
# con <- odbcDriverConnect(PATH)
# #
# # # Show the different table names
# sqlTables(con)$TABLE_NAME
# #
# # #Download table and write csv
# ChopChannel_raw <- sqlFetch(con, "ChopChannelTBL1")
# # write_csv(ChopChannel_raw, "ChopChannel_2008.csv")
# #
# ChopRecov_raw <- sqlFetch(con,"ChopRecovTBL")
# write_csv(ChopRecov_raw, "ChopRecov_2004.csv")
#
# ChopHeader_raw <- sqlFetch(con,"ChopHeaderTBL")
# write_csv(ChopHeader_raw, "ChopHeader_2004.csv")
#
#
# cwt_raw <- sqlFetch(con, "CWTTagTBL")
# write_csv(cwt_raw, "cwt_2004.csv")
#
# cwt_header_raw <- sqlFetch(con, "CWTHeaderTBL")
# write_csv(cwt_header_raw, "CWTHeader_2004.csv")
#Other tables include Fish Data_sr_Database, Steelhead Data, TagColLu
```

```{r,, eval=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2008/ChopChannelTBL1_2008.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopChannel_2008.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2008/ChopHeaderTBL_2008.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2008.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2008/ChopRecovTBL_2008.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopRecov_2008.xlsx",
               overwrite = TRUE)
# 
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2008/CWTHeaderTBL_2008.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTHeader_2008.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2008/CWTTagTBL_2008.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTTag_2008.xlsx",
               overwrite = TRUE)
```

## Raw Data Glimpse: {.tabset}

### ChopChannel_raw
```{r}
ChopChannel_raw <- read_excel("ChopChannel_2008.xlsx") %>% 
  rename("ID" = HeaderID) %>%
  glimpse()
```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_excel("ChopHeader_2008.xlsx") %>% 
  rename("ID" = headerID) %>%
  glimpse()
```

### ChopRecov_raw
```{r}
ChopRecov_raw <- read_excel("ChopRecov_2008.xlsx") %>% 
  glimpse()
```
### cwt_raw
```{r}
cwt_raw <- read_excel("CWTTag_2008.xlsx") %>% 
  glimpse
```

### cwt_header_raw

```{r}
cwt_header_raw <- read_excel("CWTHeader_2008.xlsx") %>% 
  glimpse
```



## Data transformations: {.tabset}

### Counts

The `chop_channel` table contains carcass counts by chop/tagged based on clips


```{r}

#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, TagCol, Date, Time),
                               ChopChannel_raw) %>% 
  clean_names()
```

### Recovery

The `chop_recovery` table contains recovered carcass counts by color

```{r}
# 2. choprecovery table (with dates)

chop_recovery_join <- full_join(ChopHeader_raw %>% 
                                  select(ID, Date, Time),
                                ChopRecov_raw %>% 
                                  rename(ID = RecovID))  %>% 
  clean_names()
```

### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopHeader_raw %>% 
  clean_names()
```

### CWT
The `cwt` table contains coded wire tag information. 

```{r}
cwt <- full_join(cwt_raw %>% clean_names(), cwt_header_raw %>% clean_names()) %>%  
  glimpse
```


## Explore Numeric Variables: {.tabset}

### Chop Channel Variable: `id`, `chan_id`, `sect`, `min`
```{r}
# Filter clean data to show only numeric variables 
# We only get chops and tags for this dataset. Not much additional data
chop_channel_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

```{r}
summary(chop_channel_join$id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

### Chop Channel Variable: `chops`, `tags`

```{r}
summary(chop_channel_join$chops)
```

```{r}
summary(chop_channel_join$tags)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$chops))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$tags))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$chops))/nrow(chop_channel_join), 3) * 100` % of values in the `chops` column are NA.
* `r round(sum(is.na(chop_channel_join$tags))/nrow(chop_channel_join), 3) * 100` % of values in the `tags` column are NA.

**Plotting chops over Period of Record**

```{r}
 
chop_channel_join %>% 
  ggplot(aes(x = date, y = chops)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```


**Plotting tags over Period of Record**

```{r}
chop_channel_join %>% 
  ggplot(aes(x = date, y = tags)) +
  geom_point(size = 1.4, alpha = .5, color = "red") + 
  labs(x = "Date", 
       y = "Tags") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Comparing tags and chops**
```{r}
chop_channel_join %>% 
  pivot_longer(cols = c(chops, tags), names_to = "tags_or_chops", values_to = "count") %>% 
  ggplot(aes(x = date, y = count, fill = tags_or_chops)) +
  geom_col() + 
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2"))

  
```

Tags appear to be less commonly applied compared to chops.


### Chop Recovery Variable: `id`, `chan_id`, `recov_count`
```{r}
# Filter clean data to show only numeric variables 
chop_recovery_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

```{r}
summary(chop_recovery_join$id)
```
```{r}
summary(chop_recovery_join$recov_count)
```

* `r round(sum(is.na(chop_recovery_join$recov_count))/nrow(chop_recovery_join$recov_count), 3) * 100` % of values in the `recov_count` column are NA.

Note: no recov_count have dates associated with it

### Chop Header Variable: `id`

```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(chop_header$id)
```

### CWT Variable: `tag_id`, `sect`, `fl`, `header_id`, `week_num`
```{r}
cwt %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(cwt$tag_id)
```
```{r}
summary(cwt$sect)
```
```{r}
summary(cwt$fl)
```
* `r round(sum(is.na(cwt$tag_id))/nrow(cwt), 3) * 100` % of values in the `tag_id` column are NA.
* `r round(sum(is.na(cwt$sect))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.
* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
#Is this graph helpful?
unique(cwt$sex)
cwt_count <- cwt %>% 
  mutate(count = 1) %>%
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "U",
                         TRUE ~ sex)) %>% 
  pivot_wider(names_from = sex, values_from = count, values_fill = 0) %>% 
  # unnest() %>% 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) %>% glimpse

total_cwt_summary <- cwt_count %>% 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) %>% 
  group_by(month(date)) %>% 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary %>% 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count / total_cwt)) %>% 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Month") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```


**Plotting fork length of each sex**

```{r}
cwt %>% 
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "Unknown",
                         TRUE ~ sex))%>% 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```


## Explore Categorical variables: {.tabset}

### Chop Channel Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_channel_join %>% 
  select_if(is.character) %>%
  colnames()
```
```{r}
unique(chop_channel_join$tag_col)
```

```{r}
unique(chop_channel_join$chan)
```
```{r}
chop_channel_cleaner <- chop_channel_join %>%
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date, time))

chop_channel_cleaner
```
### Chop Recovery Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_recovery_join %>% 
  select_if(is.character) %>% 
  colnames()
```

```{r}
#Not sure what the numbers are
unique(chop_recovery_join$recov_color)
```
```{r}
chop_recovery_cleaner <- chop_recovery_join %>% 
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date,time))

chop_recovery_cleaner
```

### Chop Header Clean Data

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$recorder)
```
```{r}
unique(chop_header$weather)
```



```{r}
chop_header_cleaner <- chop_header %>%
  mutate(datetime = paste(as_date(date), as_hms(as.POSIXct(time, format = "%Y-%m-%dT%H:%M:%OSZ")))) %>%
  mutate_if(is.character, str_to_lower) %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         weather = ifelse(weather == "ran", "rain", weather)) %>% 
  select(-c(date, time))

chop_header_cleaner
```

### CWT Clean Data

```{r}
cwt %>% 
  select_if(is.character) %>% 
  colnames
```

```{r}
unique(cwt$spawn)
```
```{r}
unique(cwt$clip)
```
```{r}
unique(cwt$samp_type)
```
```{r}
unique(cwt$tag_col)
```

```{r}
unique(cwt$run)
```

```{r}
cwt_cleaner <- cwt %>% 
  mutate_if(is.character, str_to_lower) %>% 
  rename(datetime = "date") %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         sex = case_when(sex == "nd"|is.na(sex) ~ "unknown",
                         TRUE ~ sex),
         spawn = ifelse(spawn == "no data", NA, spawn),
         clip = case_when(clip == "no data" ~ NA_character_,
                          clip == "y" ~ "yes",
                          clip == "n" ~ "no",
                          TRUE ~ clip),
         samp_type = ifelse(samp_type == "no data", NA, samp_type)) %>% 
  glimpse
```


## Comments
* The carcass count table (chop_channel) only records chops and tags.
* Some changes to the columns of CWT table 

## Data Dictionaries

# Channel

```{r}
percent_na <- chop_channel_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("ID",
                                          "Color of tag applied to carcass",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped",
                                          "Carcass that were tagged",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
percent_na <- chop_recovery_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


recovery_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("ID",
                                          "Color of tag recovered from carcass",
                                          "Count of recovery",
                                          "Channel ID",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))

kable(recovery_data_dictionary)
```

### Survey

```{r}

percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
header_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))
# 
kable(header_data_dictionary)
```

### CWT
```{r}
percent_na <- cwt_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())

cwt_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Sect",
                                          "Fork length",
                                          "Sex",
                                          "Spawn",
                                          "Clip",
                                          "Sample type",
                                          "Head tag",
                                          "Header ID",
                                          "Hall print 1",
                                          "Hall print 2",
                                          "Scale sample",
                                          "Run",
                                          "Datetime",
                                          "Tag colour",
                                          "Crew",
                                          "Week number"),
                          percent_na = round(percent_na$value*100))

kable(cwt_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)

```{r}
# Name file [watershed]_[data type].csv
```