---
title: "Feather Carcass QC 2001"
author: "Inigo Peng"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
```

# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

```{r}
#Connect to microsoft access through RODBC
#Use 
DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
MDBPATH <- "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/Carcass 2001.mdb"
PATH <- paste0(DRIVERINFO, "DBQ=", MDBPATH)

con <- odbcDriverConnect(PATH)

# Show the different table names 
sqlTables(con)$TABLE_NAME

#Download table and write csv
ChopEnv_raw <- sqlFetch(con, "ChopEnvTBL")
write_csv(ChopEnv_raw, "ChopEnv_2001.csv")

RecovTag_raw <- sqlFetch(con, "RecovTagTBL")
write_csv(RecovTag_raw, "RecovTag_2001.csv")

ChopRecov_raw <- sqlFetch(con,"ChopRecovTBL")
write_csv(ChopRecov_raw, "ChopRecov_2001.csv")

cwt_raw <- sqlFetch(con, "cwtTBL")
write_csv(cwt_raw, "cwt_2001.csv")

TagData_raw <- sqlFetch(con, "TagDataTBL")
write_csv(TagData_raw, "TagData_2001.csv")

TagEnv_raw <- sqlFetch(con, "TagEnvTBL")
write_csv(TagEnv_raw, "TagEnv_2001.csv")
```

```{r}
#Heading Info for Chopped Data
#Columns that were not in ChopEnv 2000: Flow and vis
ChopEnv_raw <- read_csv("ChopEnv_2001.csv") %>% 
  # mutate("Start" = as_hms("Start")) %>%
  glimpse()

#Data collection for Chopping and Spawning to determine how many carcasses are being captured (link to the ChopEnv table using ChopEnvID)

ChopRecov_raw <-read_csv("ChopRecov_2001.csv") %>% 
  glimpse()

#There is Tag1 and Tag2- what is the difference?
# In 2000 Carcass data, RecovTag table was joined to TagData table via TagNum; in this table, there are Tag1 and Tag2 - Join using these two columns?
#What is TagComp?
#Could join to ChopRecov_Raw using RecovID
RecovTag_raw <- read_csv("RecovTag_2001.csv") %>% 
  # rename("ID" = HeaderID) %>% 
  glimpse()

#tag colour lookup table
# TagCol_raw <- read_csv("TagCol_2001.csv") %>% 
  # glimpse()

#Data for Tagging DataSheet - for some reason when export access table to excel, Tag Number is automatically changed to TagID
#Link to RecovTag table
TagData_raw <- read_csv("TagData_2001.csv") %>% 
  rename("TagNum" = "TagID") %>%
  glimpse()

#Heading Info for Tagging Data Sheet
# Join via TagEnvID
TagEnv_raw <- read_csv("TagEnv_2001.csv") %>% 
  glimpse()

#Coded Wiretag Info

cwt_raw <- read_csv("cwt_2001.csv") %>% 
  glimpse
```

## Data transformations

```{r}
#1. Link ChopRecov with ChopEnv to get the date
chop_recovery_join <- full_join(ChopEnv_raw %>% select(Date, Time, ChopEnvID), ChopRecov_raw ) %>%
  clean_names() %>%  
  rename(male_chop = "male",
         female_chop = "fem",
         grilse = "gril") %>% glimpse

tag_join <- left_join(TagData_raw, RecovTag_raw) %>% 
  clean_names() %>% 
  select(c(tag_num, sect, fl, sex, egg_ret, recov_id)) %>% 
  glimpse

#Not sure if joining chop_recovery table and tag table add any more insights
chop_tag_join <- left_join(chop_recovery_join, tag_join) %>% 
  glimpse

chop_header <- ChopEnv_raw %>% 
  clean_names()

cwt <- cwt_raw %>% 
  clean_names()
```

# Recovery 

The `chop_recovery_join` table contains recovered carcass counts 

# Tags

The `tag_join` table contains additional tag information 

### Survey

The `chop_header` table contains survey metadata and covariates

#CWT

The `cwt` table contains coded wire tag information. It is not joined with the other tables.

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
chop_recovery_join %>% 
  select_if(is.numeric) %>%
  colnames()
```
### Variable: `chop_env_id`, `sect`, `minute`
```{r}
summary(chop_recovery_join$chop_env_id)
```
```{r}
summary(chop_recovery_join$sect)
```
```{r}
summary(chop_recovery_join$minute)
```
* `r round(sum(is.na(chop_recovery_join$minute))/nrow(chop_recovery_join), 3) * 100` % of values in the `minute` column are NA.

### Variable: `male_chop`, `female_chop`, `grilse`
```{r}
summary(chop_recovery_join$male_chop)
```
```{r}
summary(chop_recovery_join$female_chop)
```

**Plotting male_chop  over Period of Record**

```{r}
chop_recovery_join %>% 
  ggplot(aes(x = date, y = male_chop)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Male Chop") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting female_chop over Period of Record**

```{r}
chop_recovery_join %>% 
  ggplot(aes(x = date, y = female_chop)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Female Chop") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```
```{r}
total_chops <- chop_recovery_join %>% 
  mutate(total_chops = male_chop + female_chop,
         date = as_date(date)) %>% 
  select(date, male_chop, female_chop, total_chops) %>% glimpse()

total_chops_test <- total_chops %>% 
  filter(date == as_date("2001-09-10"))
nrow(total_chops) == (sum(total_chops$total_chops == (total_chops$male_chop + total_chops$female_chop)))

total_chops_summary <- total_chops %>% 
  group_by(date) %>% 
  summarise(total_chops = sum(total_chops, na.rm = T),
            male_chops = sum(male_chop, na.rm = T),
            female_chops = sum(female_chop, na.rm = T))
```
```{r}
total_chops_summary %>% 
  pivot_longer(cols = c(male_chops, female_chops), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count/total_chops)) %>% 
  ggplot(aes(x = date, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red"), 
                    name = "Sex", 
                    labels = c("Female", "Male")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

This plot shows the gender proportion of chops for each day over the the period of 2001-09-10 to 2001-12-13. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. All of the Chop data have been identified for their sexes.

```{r}
#Assume gril is grilse
summary(chop_recovery_join$grilse)
```

**Plotting grilse  over Period of Record**

```{r}
chop_recovery_join %>% 
  ggplot(aes(x = date, y = grilse)) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "Date", 
       y = "Grilse") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting total daily grilse over Period of Record**
```{r}
chop_recovery_join %>% 
  group_by(date) %>% 
  summarise(total_grilse_chops = sum(grilse, na.rm = T)) %>% 
  ggplot(aes(x = date, y = total_grilse_chops)) + 
  geom_point(size = 1.4, alpha = .5, color = "red") +
  labs(x = "Date",
       y = "Total Grilse") +
  theme_minimal() +
  theme(text = element_text(size = 15))
```
```{r}
summary(chop_recovery_join$spawn)
```

```{r}
tag_join %>% 
  select_if(is.numeric) %>% 
  colnames()
```

```{r}
summary(tag_join$tag_num)
```
```{r}
summary(tag_join$sect)
```

```{r}
summary(tag_join$recov_id)
```
* `r round(sum(is.na(tag_join$recov_id))/nrow(tag_join), 3) * 100` % of values in the `recov_id` column are NA.

**Plotting total count of each sex  over Period of Record**
```{r}
unique(tag_join$sex)
#Since each row of tag_join is one count of the tag, a new column count with value of 1 is created. This value is assigned to the count of each sex
#Pivot table to expand sex column to female_tag, male_tag, and unknown_tags file

all_tags <- tag_join %>% 
  mutate(tag_count = 1) %>% 
  pivot_wider(names_from = sex, values_from = tag_count, values_fill = 0) %>% 
  rename("unknown_tags" = X,
         "male_tag" = M,
         "female_tag" = F) 
total_tags_summary <- all_tags%>% 
  mutate(male_tag = ifelse(is.na(male_tag), 0, male_tag), # fill na
         female_tag = ifelse(is.na(female_tag), 0, female_tag),
         unknown_tags = ifelse(is.na(unknown_tags), 0, unknown_tags),
         total_tags = unknown_tags + male_tag + female_tag) %>% 
  select(male_tag, female_tag, unknown_tags, total_tags) %>% 
  summarise(total_tags = sum(total_tags),
            male_tags = sum(male_tag),
            female_tags = sum(female_tag),
            unknown_tags = sum(unknown_tags))
```
```{r}
total_tags_summary %>% 
  pivot_longer(cols = c(male_tags, female_tags, unknown_tags), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count / total_tags)) %>% 
  ggplot(aes(x = sex, y = proportions)) +
  geom_col() +
  theme_minimal() + 
  labs(y = "Proportion", x = "Sex") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```
```{r}
tag_join %>% 
  mutate(sex = ifelse(tag_join$sex == "X", "Unknown Tags", tag_join$sex)) %>% 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```
### Variable: `female_tag`, `male_tag`, `unknown_tag`

**Numeric Summary of `female_tag`, `male_tag`,`unknown_tag`**

```{r}
summary(all_tags$male_tag)
```
* `r round(sum(is.na(total_tags$male_tag))/nrow(total_tags), 3) * 100` % of values in the `male_tag` column are NA.

```{r}
summary(all_tags$female_tag)
```
* `r round(sum(is.na(tag_chop_join$female_tag))/nrow(tag_chop_join), 3) * 100` % of values in the `female_tag` column are NA.

```{r}
summary(all_tags$unknown_tags)
```

* `r round(sum(is.na(tag_chop_join$unknown_tags))/nrow(tag_chop_join), 3) * 100` % of values in the `unknown_tags` column are NA.

```{r}
#Join tags data with chop recovery to visualize tags applied

tags_with_dates <- left_join(all_tags, chop_recovery_join)%>% glimpse
  # filter(!is.na(date)) %>% glimpse

total_tags_with_dates <- tags_with_dates%>% 
  mutate(male_tag = ifelse(is.na(male_tag), 0, male_tag), # fill na
         female_tag = ifelse(is.na(female_tag), 0, female_tag),
         unknown_tags = ifelse(is.na(unknown_tags), 0, unknown_tags),
         total_tags = unknown_tags + male_tag + female_tag) %>% 
  select(date, male_tag, female_tag, unknown_tags, total_tags)

total_tags_with_dates_summary <- total_tags_with_dates %>%
  group_by(date) %>%
  summarise(total_tags = sum(total_tags, na.rm = T),
            male_tags = sum(male_tag, na.rm = T),
            female_tags = sum(female_tag, na.rm = T),
            unknown_tags = sum(unknown_tags, na.rm = T))

total_tags_with_dates_summary %>% 
  pivot_longer(cols = c(male_tags, female_tags, unknown_tags), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count / total_tags)) %>% 
  ggplot(aes(x = date, y = proportions, fill = sex)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Sex", 
                    labels = c("Female", "Male", "Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
#Lots of NA Dates - making plots over period of time will not be as informative
```
