---
title: "Feather Carcass QC 2016"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```

# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

Read in data from google cloud, glimpse raw data and domain description sheet: 

```{r,eval=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2016/Chops_2016.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "Chops_2016.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2016/ChopHeader_2016.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2016.xlsx",
               overwrite = TRUE)

# 
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2016/CWTHeader_2016.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTHeader_2016.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2016/CWT_2016.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWT_2016.xlsx",
               overwrite = TRUE)
```
## Raw Data Glimpse: {.tabset}

### Chop_raw
```{r}
Chop_raw <- read_excel("Chops_2016.xlsx") %>%
  rename("ID" = `Chop Header ID`,
         "Count" = `Total Count`) %>%
  glimpse()

```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_excel("ChopHeader_2016.xlsx") %>% 
  rename(ID = `Chop Header ID`,
         week_number = `Week #`) %>%
  glimpse()
```

### cwt_raw
```{r}
cwt_raw <- read_excel("CWT_2016.xlsx") %>% 
  rename("ID" = `CWT Header ID`) %>% 
  glimpse

```

### cwt_header_raw

```{r}
cwt_header_raw <- read_excel("CWTHeader_2016.xlsx") %>% 
  rename("ID" = `CWT Header ID` ) %>% 
  glimpse
```

## Data transformations: {.tabset}

### Counts

The `chop` table contains carcass counts by chop/tagged based on clips


```{r}

#1. chop table (with dates and tag color)
chop_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, Date),
                               Chop_raw) %>% 
  clean_names() %>% 
  rename(sec = "section",
         min = "minutes") %>% 
  mutate(sec = as.numeric(sec)) %>% glimpse
```

### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopHeader_raw %>% 
  clean_names() %>% glimpse
```

### CWT
The `cwt` table contains coded wire tag information. 

```{r}
cwt <- full_join(cwt_header_raw, cwt_raw) %>% 
  clean_names() %>% 
  rename(fl = "fork_length") %>% 
  mutate(head_tag_number = as.numeric(head_tag_number),
         tag_id_number = as.numeric(tag_id_number),
         scales = as.numeric(scales),
         otoliths = as.numeric(otoliths))%>% glimpse()
```


## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
chop_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

### Chop Join Variable: `id`, `section`, `min`

```{r}
summary(chop_join$id)
```
```{r}
summary(chop_join$sec)
```

```{r}
summary(chop_join$min)
```


### Chop Join Variable: `count`

**Numeric Summary of `count` over Period of Record**


```{r}
summary(chop_join$count)
```

**NA and Unknown Values**

* `r round(sum(is.na(chop_join$count))/nrow(chop_join), 3) * 100` % of values in the `count` column are NA.

**Plotting count over Period of Record**

```{r}

# daily chop count over time
chop_join %>% 
  ggplot(aes(x = date, y = count)) +
  geom_point() + 
  theme_minimal() +
  theme(text = element_text(size = 15))
```

This plot shows the daily chops collected each day from August to December 2016. The data from 2016 does not contain tags information - it only has count.

**Plotting total chops over Period of Record**

```{r}
chop_join %>% 
  group_by(date) %>% 
  summarize(total_chops = sum(count, na.rm = T)) %>% 
  ggplot(aes(x = date, y = total_chops)) + 
  geom_col() + 
  theme_minimal()
```

This plot shows the daily total chops collected between August and December.
### Chop Header Variable: `id`

```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(chop_header$id)
```

### CWT Variable: `ID`, `sect`, `fl`, `header_id`, `week_num`
```{r}
cwt %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(cwt$fl)
```
```{r}
summary(cwt$river_section)
```

* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.
* `r round(sum(is.na(cwt$river_section))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
#Is this graph helpful?
unique(cwt$sex)
cwt_count <- cwt %>% 
  mutate(count = 1) %>%
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "U",
                         TRUE ~ sex)) %>% 
  pivot_wider(names_from = sex, values_from = count, values_fill = 0) %>% 
  # unnest() %>% 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) %>% glimpse

total_cwt_summary <- cwt_count %>% 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) %>% 
  group_by(month(date)) %>% 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary %>% 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count / total_cwt)) %>% 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Month") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```


**Plotting fork length of each sex**

```{r}
cwt %>% 
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "Unknown",
                         TRUE ~ sex))%>% 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```

## Explore Categorical variables: {.tabset}

### Chop Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_join %>% 
  select_if(is.character) %>%
  colnames()
```

```{r}
chop_cleaner <- chop_join %>%
  mutate(date = as_date(date)) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-chop_id)

chop_cleaner

```
### Chop Header Clean Data

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$weather)
```



```{r}
chop_header_cleaner <- chop_header %>%
  mutate_if(is.character, str_to_lower) %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         weather = case_when(weather == "cld" ~ "cloudy",
                             weather == "cld/rain" ~ "cloudy, rain",
                             weather == "sun/cld" ~ "sun, cloudy",
                             weather == "sun/cld/ran" ~ "sun, cloudy, rain",
                             TRUE ~ weather),
         weather = str_replace_all(weather, " ", ""),
         weather = str_replace_all(weather, "/", ",")) %>% 
  select(-c(time))

chop_header_cleaner
```

### CWT Clean Data

```{r}
cwt %>% 
  select_if(is.character) %>% 
  colnames
```

```{r}
unique(cwt$spawning_condition)
```

```{r}
unique(cwt$adipose_fin_clipped)
```

```{r}
unique(cwt$crew)
```
```{r}
unique(cwt$tag_recapture_or_chop)
```

```{r}
unique(cwt$samples_collected)
```
```{r}
unique(cwt$tag_color)
```
```{r}
unique(cwt$hallprint_color)
```


```{r}
#Dropping sectiongroup and morale
cwt_cleaner <- cwt %>% 
  mutate_if(is.character, str_to_lower) %>% 
  rename(datetime = "date") %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         sex = case_when(sex == "nd"|is.na(sex) ~ "unknown",
                         TRUE ~ sex),
         tag_recapture_or_chop = case_when(tag_recapture_or_chop == "T" ~ "tagged",
                                         tag_recapture_or_chop == "R" ~ "recapture",
                                         tag_recapture_or_chop == "C" ~ "chop",
                                         TRUE ~ tag_recapture_or_chop),
         adipose_fin_clipped = case_when(adipose_fin_clipped == "UK" ~ "unknown",
                                         adipose_fin_clipped == "Y" ~ "yes",
                                         adipose_fin_clipped == "N" ~ "no",
                                         TRUE ~ adipose_fin_clipped),
         hallprint_color = case_when(hallprint_color == "grey" ~ "G",
                                     TRUE ~ hallprint_color),
         spawning_condition = case_when(spawning_condition == "UK" ~ "unknown",
                                        TRUE ~ spawning_condition)) %>% 
  select(-c(morale, section_group_1_10, section_group_11_15, section_group_16_21, section_group_22_38)) %>% 
  glimpse
```


## Data Dictionaries

### Chop Count

```{r}
percent_na <- chop_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_cleaner),
                          description = c("ID",
                                          'Date',
                                          "Sect",
                                          "Min", 
                                          "Carcass chopped count"),
                          percent_na = round(percent_na$value*100,
                                             digits = 1))

kable(counts_data_dictionary)
```

### Survey

```{r}
percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())

chop_header_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("Date of survey",
                                          "Week number",
                                          "Weather",
                                          "Crew memeber initials that collected",
                                          "Comments",
                                          "ID"),
                          percent_na = round(percent_na$value*100,
                                             digits = 1))

kable(chop_header_data_dictionary)
```
### CWT
```{r}
percent_na <- cwt_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())

cwt_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Date",
                                          "Crew",
                                          "Week number",
                                          "Tag colour",
                                          "CWT ID",
                                          "River section",
                                          "Tag ID number",
                                          "Tag recapture or chop",
                                          "Sex",
                                          "Spawning condition",
                                          "Adipose fin clipped",
                                          "Samples collected",
                                          "Fork length",
                                          "Head tag number",
                                          "Scales",
                                          "Otoliths",
                                          "Hallprint color",
                                          "Hallprint",
                                          "Comments"),
                          percent_na = round(percent_na$value*100))

kable(cwt_data_dictionary)
```
## Saved cleaned data back to google cloud
```{r}
#Ignore chop recovery as its 100% NA
feather_carcass_chops_2016 <- chop_cleaner %>% glimpse()
feather_carcass_cwt_2016 <- cwt_cleaner %>% glimpse()
feather_carcass_chop_header_2016 <- chop_header_cleaner %>% glimpse()
```
```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_carcass_chops_2016,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chops_and_tags_2016.csv")
gcs_upload(feather_carcass_cwt_2016,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_cwt_2016.csv")
gcs_upload(feather_carcass_chop_header_2016,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_header_2016.csv")
```




