---
title: "Feather Carcass QC 2016"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```

# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2016.accdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2016.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2016.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopHeader_raw<- sqlFetch(con, "Chop Header")
  Chops_raw<- sqlFetch(con, "Chops")
  CWT_raw<- sqlFetch(con, "CWT")
  CWTHeader_raw <- sqlFetch(con, "CWT Header")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  # Otoliths, Switchboard, and Sample Weeks tables contain no data
  ChopHeader_raw<- mdb.get(filepath, "Chop Header")
  Chops_raw<- mdb.get(filepath, "Chops")
  CWT_raw<- mdb.get(filepath, "CWT")
  CWTHeader_raw <- mdb.get(filepath, "CWT Header")
  detach(package:Hmisc)
}

write_csv(ChopHeader_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2016.csv"))
write_csv(Chops_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "Chops_2016.csv"))
write_csv(CWT_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWT_2016.csv"))
write_csv(CWTHeader_raw, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2016.csv"))
```

## Raw Data Glimpse: {.tabset}

### Chops_raw
```{r}
Chop_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river","Chops_2016.csv")) |> 
  rename(Header.ID = `Chop.Header.ID`,
         "Count" = `Total.Count`) |>
  glimpse()

```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river","ChopHeader_2016.csv")) |> 
  rename(Header.ID = `Chop.Header.ID`,
         week_number = `Week..`) |>
  glimpse()
```

### cwt_raw
```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river","CWT_2016.csv")) |> 
  rename(Header.ID = `CWT.Header.ID`) 

# column names appeared to be shifted one to the right when read in from the
# access database (i.e. fork length was orders of magnitude larger)
# so the following code renames columns based on assumed order and type of
# column values expected

colnames(cwt_raw) <- c("CWT.ID", "Header.ID", "River.Section", "Tag.ID", "Recapture.or.Chop", "Sex", "Spawning.Condition", "Adipose.Fin.Clipped.", "Samples.Collected", "Fork.Length", "Head.Tag.Number", "Scales", "Otoliths", "Hallprint.Color", "Hallprint", "Comments", "Tag")
cwt_raw |> glimpse()

```

### cwt_header_raw

```{r}
cwt_header_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river","CWTHeader_2016.csv")) |> 
  rename(Header.ID = `CWT.Header.ID` ) |> 
  glimpse()
```

## Data transformations: {.tabset}

Note: for 2011-2016 data, the Access relationships are as follows:
* `CWT$HeaderID` joins to `CWTHeader$HeaderID`
* `CWTHeader$HeaderID` joins to `SampleWeeks$WeekNumber`
* `SampleWeeks$WeekNumber` joins to `ChopHeader$WeekNumber`
* `ChopHeader$HeaderID` joins to `Chops$HeaderID`

Note: when `week_number == 0`, it is a pre-carcass survey week. In `cwt_raw`, this is recorded in the `morale` column but we add it to the `disposition` column (which is `NA` for week 0 anyway).


### CWT
The `cwt` table contains sampling information on individual carcasses. This table also contains sex, spawning condition, and adipose fin clip status. It is not joined with the other tables. Added a `count` column assuming each row represents one individual. This table contains `disposition` information in the `recapture_or_chop` variable (in previous years this was the `carc_stat` variable, and in 2012 this was in three separate recap ID columns). If a tagged carcass was tagged, recaptured, re-released, and chopped, it will be represented multiple times in this table (i.e. there can be multiple rows for one `tag_id`).  Added `source_table` column for tracking. 

```{r}
cwt <- left_join(cwt_raw |> clean_names(), 
                 cwt_header_raw |> clean_names(),
                 by = "header_id") |> 
  rename(fl = "fork_length",
         week_number = "week") |> 
  mutate(source_table = "CWT",
         count = 1,
         disposition = case_when(recapture_or_chop == "C" ~ "chopped",
                                 recapture_or_chop == "T" ~ "tagged",
                                 recapture_or_chop == "R" ~ "recaptured",
                                 week_number == 0 ~ "pre_carcass_survey",
                                 TRUE ~ recapture_or_chop),
         sex = case_when(sex == "F" ~ "female",
                         sex == "M" ~ "male",
                         TRUE ~ sex)) |>
  select(-c(recapture_or_chop, morale, crew)) |> 
  glimpse()
```


### Bulk chop counts

The `chops` table has bulk chop counts by week. There is no additional information. `disposition` column added assuming all were `chopped`.

Note: changed week numbers marked as `pre_carcass` to 0 and recorded the `disposition` of those rows as `pre_carcass_survey` to align with `CWT` table.

```{r}
chop_join <- full_join(ChopHeader_raw |> 
                                 select(Header.ID, Date, Time, week_number),
                               Chop_raw,
                       by = "Header.ID") |> 
  clean_names() |> 
  rename(sect = "section",
         min = "minutes") |> 
  mutate(sect = as.numeric(sect),
         disposition = "chopped",
         source_table = "ChopJoin",
         week_number = case_when(week_number %in% c("pre-carcass", "pre carcass") ~ "0",
                                 TRUE ~ week_number),
         week_number = as.numeric(week_number),
         disposition = if_else(week_number == 0, "pre_carcass_survey", disposition)) |> 
  glimpse()
```


### Recovered tags and chops

There is no `chop_recovery` table for this year.

### Combine tables

```{r}
all_carcasses <- bind_rows(chop_join, cwt) |> glimpse()
```

### Survey

The `chop_header` table contains survey metadata and covariates 
```{r}
chop_header <- ChopHeader_raw |> 
  clean_names() |> glimpse()
```


## Explore Numeric Variables: {.tabset}

### All Carcasses Variables: `header_id`, `week_number`, `sect`, `min`, `count`, `cwt_id`, `river_section`, `tag_id`, `fl`, `head_tag_number`, `scales`, `otoliths`

```{r}
# Filter clean data to show only numeric variables 
all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
```

```{r}
numeric_cols <- all_carcasses |>  
  select_if(is.numeric) |> 
  colnames()
sapply(all_carcasses[,numeric_cols], summary)
```

### All Carcasses Variable: `count`

```{r}
summary(all_carcasses$count)
```

** Plot carcass counts by sex **

```{r}
all_carcasses |>  
  group_by(date, sex) |> 
  summarize(count = sum(count, na.rm = T)) |> 
  ggplot(aes(x = date, y = count, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(x = "Date", 
       y = "Chops") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  theme(text = element_text(size = 15))
```
This plot shows carcass over period of record by sex and is helpful for understanding the seasonality of the survey, i.e. chop counts peak in Oct-Nov.

**Plotting Total Adult Chops over Period of Record**

```{r}
total_adult_chops <- all_carcasses |> 
  filter(!disposition %in% c("tagged", "pre_carcass_survey")) |> 
  select(date, count, disposition, sex) |> 
  glimpse()

total_adult_chops_summary <- total_adult_chops |> 
  group_by(date, sex) |> 
  summarise(total_adult_chops = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_adult_chops_summary |>
  ggplot(aes(x = date, y = total_adult_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

This plot shows the sex proportion of adult chops (no `tagged` or `pre_carcass_survey` carcasses) for each day over the period of `r min(all_carcasses$date, na.rm = T)` to `r max(all_carcasses$date, na.rm = T)`. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period.


**Plotting grilse over Period of Record**
This year did not have lifestage information.

**Plotting all carcasses over Period of Record**
Carcasses in the `all_carcasses` dataset are marked as either `tagged`, `recaptured`, `chopped`, or `pre_carcass_survey`. Carcasses from the `CWT` table are marked as `chopped`, `tagged`, `recaptured`, `pre_carcass_survey`, or `NA` and have individual-level sampling information (i.e sex, fork length, adipose fin clip status). `chopped` carcasses from `ChopJoin` do not have individual-level sampling data. 

```{r}
total_carcasses_summary <- all_carcasses |> 
  group_by(date, disposition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_carcasses_summary |> 
  ggplot(aes(x = date, y = total_carcasses, fill = disposition)) + 
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Individual-level sampling information

**Plotting fork length of each sex**

```{r}
summary(all_carcasses$fl)
```

```{r}
all_carcasses |> 
  ggplot(aes(x = sex, y = fl)) +
  geom_boxplot() +
  facet_wrap(~disposition) +
  theme_minimal() + 
  labs(y = "fork length (mm)", x = "sex")
```

**Plotting sex proportion of counts within each disposition**

```{r}
all_carcasses |> 
  ggplot(aes(y = count, x = disposition, fill = sex)) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Disposition") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

### Chop Header Variable: `header_id`, `week`


```{r}
chop_header |>  
  select_if(is.numeric) |>  
  colnames()
```
```{r}
summary(chop_header$header_id)
```


## Explore Categorical variables: {.tabset}

### All Carcasses

Fix inconsistencies with spelling, capitalization, and dates

```{r}
all_carcasses |>  
  select_if(is.character) |>  
  colnames()
```

```{r}
character_cols <- all_carcasses |>  
  select_if(is.character) |>  
  colnames()

sapply(all_carcasses[, character_cols[! character_cols %in% c("comments")]], unique)
```

```{r}
all_carcasses_cleaner <- all_carcasses |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(spawning_condition = case_when(spawning_condition == "u" ~ "unspawned",
                                        spawning_condition == "uk" ~ "unknown",
                                        spawning_condition == "s" ~ "spawned",
                                        TRUE ~ spawning_condition),
         adipose_fin_clipped = case_when(adipose_fin_clipped == "y" ~ "TRUE",
                                         adipose_fin_clipped == "n" ~ "FALSE",
                                         adipose_fin_clipped == "uk" ~ NA_character_,
                                         TRUE ~ adipose_fin_clipped),
         adipose_fin_clipped = as.logical(adipose_fin_clipped)) |> 
  select(-c(section_group_1_10, section_group_11_15, section_group_16_21,
            section_group_22_38, tag, hallprint)) |>  # Not sure section_group variables are useful, tag & hallprint are empty
  glimpse()
```

### Chop Header Clean Data

```{r}
chop_header |>  
  select_if(is.character) |>  
  colnames()
```
```{r}
unique(chop_header$crew)
unique(chop_header$week_number)
unique(chop_header$weather)
```


```{r}
chop_header_cleaner <- chop_header |> 
  mutate_if(is.character, str_to_lower) |>  
  mutate(crew = str_replace_all(crew, " ", ""),
         crew = str_replace_all(crew, "/", ","),
         crew = str_replace_all(crew, "&", ","),
         week_number = if_else(week_number == "pre-carcass", "pre carcass", week_number),
         weather = case_when(weather == "cld" ~ "cloudy",
                             weather == "clr" ~ "sun",
                             weather %in% c("cld/rain", "cld,rain") ~ "rain",
                             weather == "sun/cld" ~ "sunny, cloudy",
                             weather == "bf,tj" ~ NA_character_,
                             weather == "sun/cld/ran" ~ "sunny, cloudy",
                             TRUE ~ weather)) |> 
  glimpse()

```

## Comments

* This table includes a pre-carcass survey in week 0. 
* Same table/layout structure as 2011-2014. There is only a bulk chops table, `ChopJoin`, and a CWT table with individual carcass sampling data.
* `Hallprint` variable is collected in `cwt` (collected for 2004 - 2014 and not for 2015).
* no more `run` variable (was collected in 2008 and 2009)
* Using `recaptured` instead of `recovered` (which is the value used in 2000-2010) and can change if necessary. Thought it was important to use the variable in original data (original column is "recapture or chop")
* can we join recovered tags on a specific `id` value?

## Data Dictionaries

### All Carcasses
```{r}
percent_na <- all_carcasses_cleaner |> 
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |> 
  pivot_longer(cols = everything())


all_carcasses_data_dictionary <- tibble(variables = colnames(all_carcasses_cleaner),
                          description = c("Header ID",
                                          "Date of survey",
                                          "Time of survey",
                                          "Week",
                                          "Chop ID",
                                          "Section",
                                          "Minute",
                                          "Count",
                                          "Disposition",
                                          "Table from which the data came",
                                          "CWT ID",
                                          "River Section",
                                          "Tag ID",
                                          "Sex",
                                          "Spawning condition",
                                          "Adipose fin clip status",
                                          "Samples collected",
                                          "Fork length (mm)",
                                          "Head tag number",
                                          "Scales",
                                          "Otoliths",
                                          "Hallprint tag color",
                                          "Comments",
                                          "Color of tag"),
                          percent_na = round(percent_na$value*100))

kable(all_carcasses_data_dictionary)
```

## Save cleaned data back to google cloud (TBA)

```{r}
all_carcasses_2016 <- all_carcasses_cleaner |> glimpse()
```

```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(all_carcasses_2016,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_all_carcasses_2016.csv")
```
