---
title: "Feather Carcass QC 2012"
author: "Inigo Peng"
date: '2022-07-21'
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```
```{r, eval = FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/Chops_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "Chops_2012.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/ChopHeader_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2012.xlsx",
               overwrite = TRUE)

# 
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/CWTHeader_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTHeader_2012.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2012/CWT_2012.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWT_2012.xlsx",
               overwrite = TRUE)

```
## Raw Data Glimpse: {.tabset}

### Chop_raw
```{r}
Chops_raw <- read_excel("Chops_2012.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  select(-`Chop ID`) %>% 
  glimpse()
```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_excel("ChopHeader_2012.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  glimpse()
```

### cwt_raw
```{r}
cwt_raw <- read_excel("CWT_2012.xlsx") %>% 
  rename("ID" = `CWT Header ID`) %>% 
  glimpse

```

### cwt_header_raw

```{r}
cwt_header_raw <- read_excel("CWTHeader_2012.xlsx") %>% 
  rename("ID" = `CWT Header ID` ) %>% 
  glimpse
```



## Data transformations: {.tabset}

### Counts

The `chop` table contains carcass counts by chop/tagged based on clips


```{r}

#1. chop table (with dates and tag color)
chop_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, Date),
                               Chops_raw) %>% 
  clean_names() %>% 
  rename(sec = "section",
         min = "minutes",
         count = "total_count") %>% 
  mutate(sec = as.numeric(sec)) %>% glimpse
```

### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopHeader_raw %>% 
  clean_names() %>% glimpse
```

### CWT
The `cwt` table contains coded wire tag information. 

```{r}
cwt <- full_join(cwt_raw %>% clean_names(), cwt_header_raw %>% clean_names()) %>%  
  rename(fl = "fork_length") %>% 
  glimpse
```


## Explore Numeric Variables: {.tabset}

### Chop Join Variable: `id`, `min`, `sec`
```{r}
chop_join %>% 
  select_if(is.numeric) %>%
  colnames()
```


```{r}
summary(chop_join$id)
```

```{r}
summary(chop_join$min)
```
```{r}
summary(chop_join$sec)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100
round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100
round(sum(is.na(chop_join$sec))/nrow(chop_join), 3) * 100
```

* `r round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100` % of values in the `id` column are NA.
* `r round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100` % of values in the `min` column are NA.
* `r round(sum(is.na(chop_join$sec))/nrow(chop_join), 3) * 100` % of values in the `sec` column are NA.

### Chop Join Variable: `count`
```{r}
summary(chop_join$count)
```

* `r round(sum(is.na(chop_join$count))/nrow(chop_join), 3) * 100` % of values in the `count` column are NA.

**Plotting total_count over Period of Record**

```{r}
chop_join %>% 
  ggplot(aes(x = date, y = count)) + 
  geom_point() +
  theme_minimal()
```
```{r}
chop_join %>%
  group_by(date) %>%
  summarise(total_count = sum(count, na.rm = T)) %>%
  ggplot(aes(x = date, y = total_count)) +
  geom_col() +
  theme_minimal()
```
### Chop Header Variable: `id`

```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(chop_header$id)
```

### CWT Variable: `ID`, `sect`, `fl`, `header_id`, `week_num`
```{r}
cwt %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(cwt$fl)
```
```{r}
summary(cwt$river_section)
```

* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.
* `r round(sum(is.na(cwt$river_section))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
#Is this graph helpful?
unique(cwt$sex)
cwt_count <- cwt %>% 
  mutate(count = 1) %>%
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "U",
                         TRUE ~ sex)) %>% 
  pivot_wider(names_from = sex, values_from = count, values_fill = 0) %>% 
  # unnest() %>% 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) %>% glimpse

total_cwt_summary <- cwt_count %>% 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) %>% 
  group_by(month(date)) %>% 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary %>% 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count / total_cwt)) %>% 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Month") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```


**Plotting fork length of each sex**

```{r}
cwt %>% 
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "Unknown",
                         TRUE ~ sex))%>% 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```

## Explore Categorical variables: {.tabset}

### Chop Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_join %>% 
  select_if(is.character) %>%
  colnames()
```

```{r}
chop_cleaner <- chop_join %>%
  mutate(datetime = as_date(date)) %>%
  mutate_if(is.character, str_to_lower) %>% 
  select(-c(date))

chop_cleaner

```
### Chop Header Clean Data

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$weather)
```



```{r}
chop_header_cleaner <- chop_header %>%
  mutate_if(is.character, str_to_lower) %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         weather = ifelse(weather == "cld", "cloudy", weather),
         weather = ifelse(weather == "ran", "rain", weather),
         weather = str_replace_all(weather, " ", ""),
         weather = str_replace_all(weather, "/", ","),
         weather = str_replace_all(weather, ".",",")) 

chop_header_cleaner
```

### CWT Clean Data

```{r}
cwt %>% 
  select_if(is.character) %>% 
  colnames
```

```{r}
unique(cwt$spawning_condition)
```

```{r}
unique(cwt$adipose_fin_clipped)
```

```{r}
unique(cwt$crew)
```

```{r}
unique(cwt$tag_color)
```

```{r}
#Dropping sectiongroup and morale
cwt_cleaner <- cwt %>% 
  mutate_if(is.character, str_to_lower) %>% 
  rename(datetime = "date") %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         sex = case_when(sex == "nd"|is.na(sex) ~ "Unknown",
                         TRUE ~ sex)) %>% 
  select(-morale) %>% 
  glimpse
```


## Data Dictionaries

# Count
```{r}
percent_na <- chop_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_cleaner),
                          description = c("ID",
                                          "Date of survey",
                                          "Sect", 
                                          "Min", 
                                          "Count",
                                          "Comments"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```
### Survey

```{r}

percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
counts_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("Date of survey",
                                          "Week Number",
                                          "Weather",
                                          "Crew memeber initials that collected",
                                          "Comments",
                                          "ID"),
                          percent_na = round(percent_na$value*100))
# 
kable(counts_data_dictionary)
```

### CWT
```{r}
percent_na <- cwt_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())

chop_header_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("CWT ID",
                                          "ID",
                                          "River section",
                                          "Tag ID number",
                                          "Fork length",
                                          "Sex",
                                          "Recap 1 ID number",
                                          "Recap 2 ID number",
                                          "Recap 3 ID number",
                                          "Chop ID number",
                                          "Spawning condition",
                                          "Adipose fin clipped",
                                          "Head tag number",
                                          "Scales",
                                          "Otoliths",
                                          "Hallprint color",
                                          "Hallprint number",
                                          "Radio tag ID",
                                          "Comments",
                                          "Date",
                                          "Crew",
                                          "Week number",
                                          "Tag colour",
                                          "LFC",
                                          "HFC"),
                          percent_na = round(percent_na$value*100))

kable(chop_header_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)

```{r}
# Name file [watershed]_[data type].csv
```
