---
title: "Feather Carcass QC 2015"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```
```{r, eval=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2015/Chops_2015.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "Chops_2015.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2015/ChopHeader_2015.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2015.xlsx",
               overwrite = TRUE)

# 
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2015/CWTHeader_2015.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTHeader_2015.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2015/CWT_2015.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWT_2015.xlsx",
               overwrite = TRUE)

```
## Raw Data Glimpse: {.tabset}

### Chop_raw
```{r}
Chops_raw <- read_excel("Chops_2015.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  select(-`Chop ID`) %>% 
  glimpse()
```

### ChopHeader_raw

```{r}
  ChopHeader_raw <- read_excel("ChopHeader_2015.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  glimpse()
```

### cwt_raw
```{r}
cwt_raw <- read_excel("CWT_2015.xlsx") %>% 
  rename("ID" = `CWT Header ID`) %>% 
  glimpse

```

### cwt_header_raw

```{r}
cwt_header_raw <- read_excel("CWTHeader_2015.xlsx") %>% 
  rename("ID" = `CWT Header ID` ) %>% 
  glimpse
```

## Data transformations: {.tabset}

### Counts

The `chop` table contains carcass counts by chop/tagged based on clips


```{r}

#1. chop table (with dates and tag color)
chop_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, Date),
                               Chops_raw) %>% 
  clean_names() %>% 
  rename(sec = "section",
         min = "minutes",
         count = "total_count") %>% 
  mutate(sec = as.numeric(sec)) %>% glimpse
```

### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopHeader_raw %>% 
  clean_names() %>% glimpse
```

### CWT
The `cwt` table contains coded wire tag information. 

```{r}
cwt <- full_join(cwt_header_raw, cwt_raw) %>% 
  clean_names() %>% 
  rename(fl = "fork_length") %>% 
  mutate(head_tag_number = as.numeric(head_tag_number),
         tag_id_number = as.numeric(tag_id_number),
         scales = as.numeric(scales),
         otoliths = as.numeric(otoliths))%>% glimpse()
```


## Explore Numeric Variables: {.tabset}

### Chop Variable: `id`, `min`, `sec`

```{r}
chop_join %>% 
  select_if(is.numeric) %>%
  colnames()
```


```{r}
summary(chop_join$id)
```

```{r}
summary(chop_join$min)
```

```{r}
summary(chop_join$sec)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100
round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100
round(sum(is.na(chop_join$sec))/nrow(chop_join), 3) * 100

```

* `r round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100` % of values in the `id` column are NA.
* `r round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100` % of values in the `min` column are NA.
* `r round(sum(is.na(chop_join$sec))/nrow(chop_join), 3) * 100` % of values in the `sec` column are NA.

### Chop Variable: `count`
```{r}
summary(chop_join$count)
```

* `r round(sum(is.na(chop_join$count))/nrow(chop_join), 3) * 100` % of values in the `count` column are NA.

**Plotting count over Period of Record**

```{r}
chop_join %>% 
  ggplot(aes(x = date, y = count)) + 
  geom_point() +
  theme_minimal()
```
```{r}
chop_join %>%
  group_by(date) %>%
  summarise(total_count = sum(count, na.rm = T)) %>%
  ggplot(aes(x = date, y = total_count)) +
  geom_col() +
  theme_minimal()
```
```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()

```
```{r}
summary(chop_header$id)
```


### Chop Header Variable: `id`

```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(chop_header$id)
```

### CWT Variable: `ID`, `sect`, `fl`, `header_id`, `week_num`
```{r}
cwt %>% 
  select_if(is.numeric) %>% 
  colnames()
```
```{r}
summary(cwt$fl)
```
```{r}
summary(cwt$river_section)
```

* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.
* `r round(sum(is.na(cwt$river_section))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
#Is this graph helpful?
unique(cwt$sex)
cwt_count <- cwt %>% 
  mutate(count = 1) %>%
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "U",
                         TRUE ~ sex)) %>% 
  pivot_wider(names_from = sex, values_from = count, values_fill = 0) %>% 
  # unnest() %>% 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) %>% glimpse

total_cwt_summary <- cwt_count %>% 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) %>% 
  group_by(month(date)) %>% 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary %>% 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") %>% 
  mutate(proportions = (count / total_cwt)) %>% 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Month") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```


**Plotting fork length of each sex**

```{r}
cwt %>% 
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "Unknown",
                         TRUE ~ sex))%>% 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```


## Explore Categorical variables: {.tabset}

### Chop Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_join %>% 
  select_if(is.character) %>%
  colnames()
```

```{r}
chop_cleaner <- chop_join %>%
  mutate(date = as_date(date)) %>%
  mutate_if(is.character, str_to_lower)

chop_cleaner

```
### Chop Header Clean Data

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
#clr = clear? or cloud
unique(chop_header$weather)
```



```{r}
chop_header_cleaner <- chop_header %>%
  mutate(date = as_date(date)) %>% 
  mutate_if(is.character, str_to_lower) %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         weather = case_when(weather == "cld" ~ "cloudy",
                             weather == "cld ran" ~ "cloudy, rain",
                             weather == "ran" ~ "rain",
                             weather == "sun/cld" ~ "sun, cloud",
                             TRUE ~ weather),
         weather = str_replace_all(weather, " ", ""),
         weather = str_replace_all(weather, "/", ",")) %>% 
  select(-c(time))

chop_header_cleaner
```

### CWT Clean Data

```{r}
cwt %>% 
  select_if(is.character) %>% 
  colnames
```

```{r}
unique(cwt$spawning_condition)
```

```{r}
unique(cwt$adipose_fin_clipped)
```

```{r}
unique(cwt$crew)
```
```{r}
unique(cwt$tag_recapture_or_chop)
```

```{r}
unique(cwt$samples_collected)
```
```{r}
unique(cwt$tag_color)
```

```{r}
#Dropping sectiongroup and morale
cwt_cleaner <- cwt %>% 
  mutate_if(is.character, str_to_lower) %>% 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         sex = case_when(sex == "nd"|is.na(sex) ~ "unknown",
                         TRUE ~ sex),
         tag_recapture_or_chop = case_when(tag_recapture_or_chop == "T" ~ "tagged",
                                         tag_recapture_or_chop == "R" ~ "recapture",
                                         tag_recapture_or_chop == "C" ~ "chop",
                                         TRUE ~ tag_recapture_or_chop),
         adipose_fin_clipped = case_when(adipose_fin_clipped == "UK" ~ "unknown",
                                         adipose_fin_clipped == "Y" ~ "yes",
                                         adipose_fin_clipped == "N" ~ "no",
                                         TRUE ~ adipose_fin_clipped),
         hallprint_color = case_when(hallprint_color == "grey" ~ "G",
                                     TRUE ~ hallprint_color),
         spawning_condition = case_when(spawning_condition == "UK" ~ "unknown",
                                        TRUE ~ spawning_condition),
         tag_color = ifelse(is.na(tag_color), NA, "Blue")) %>% 
  select(-c(morale, section_group_1_10, section_group_11_15, section_group_16_21, section_group_22_38)) %>% 
  glimpse
```

## Data Dictionaries

# Count
```{r}
percent_na <- chop_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_cleaner),
                          description = c("ID",
                                          "Date of survey",
                                          "Sect", 
                                          "Min", 
                                          "Count"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```
### Survey

```{r}

percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
chop_header_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("Date of survey",
                                          "Week Number",
                                          "Weather",
                                          "Crew memeber initials that collected",
                                          "Comments",
                                          "ID"),
                          percent_na = round(percent_na$value*100))
# 
kable(chop_header_data_dictionary)
```

### CWT
```{r}
percent_na <- cwt_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())

cwt_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Date",
                                          "Crew memeber initials that collected",
                                          "Week number",
                                          "Tag colour",
                                          "CWT ID",
                                          "River section",
                                          "Tag ID Number",
                                          "Carcass tagged, recaptured, or chopped",
                                          "Sex of the carcass",
                                          "Spawning condition",
                                          "Adipose fin clipped or not",
                                          "Sample collected",
                                          "Fork length",
                                          "Head tag number",
                                          "Scales",
                                          "Otoliths",
                                          "Hallprint colour",
                                          "Hallprint",
                                          "Comments"),
                          percent_na = round(percent_na$value*100))
# 
kable(cwt_data_dictionary)
```
# Saved cleaned data back to google cloud
```{r}
feather_carcass_chops_2015 <- chop_cleaner %>% glimpse()
feather_carcass_cwt_2015 <- cwt_cleaner %>% glimpse()
feather_carcass_chop_header_2015 <- chop_header_cleaner %>% glimpse()
```
```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_carcass_chops_2015,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chops_and_tags_2015.csv")
gcs_upload(feather_carcass_cwt_2015,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_cwt_2015.csv")
gcs_upload(feather_carcass_chop_header_2015,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_header_2015.csv")
```

