---
title: "Feather Carcass QC 2007"
author: "Inigo Peng"
date: '2022-07-21'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```
```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2007.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2007.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2007.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopChannel <- sqlFetch(con, "ChopChannelTBL1")
  ChopRecov <- sqlFetch(con, "ChopRecovTBL")
  ChopHeader <- sqlFetch(con, "ChopHeaderTBL")
  cwt <- sqlFetch(con, "CWTTagTBL")
  cwt_header <- sqlFetch(con, "CWTHeaderTBL")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  ChopChannel <- mdb.get(filepath, "ChopChannelTBL1")
  ChopRecov <- mdb.get(filepath, "ChopRecovTBL")
  ChopHeader <- mdb.get(filepath, "ChopHeaderTBL")
  cwt <- mdb.get(filepath, "CWTTagTBL")
  cwt_header <- mdb.get(filepath, "CWTHeaderTBL")
  
  # FL by sex is empty
  # TagColLU_06 contains metadata about tag color
  # headtag contains location, fork length, run, sex, date, recovery, head_tag ID
  # FL by sex (self explanatory), TagColLU_05 (metadata about tag colors)
  # ChopHallprintTBL contains TagID, sect, fork length, sex, spawn, clip, carc_stat, comment, headerID, hallprint1, hallprint2, headtag
  # ScaleSPLs contains date, fork length, sex, spawn, clip, sample type, head tag
  # Hallprint_Sect contains hallprint1, hallprint2, date, sect
  detach(package:Hmisc)
}

write_csv(ChopChannel, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopChannel_2007.csv"))
write_csv(ChopRecov, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2007.csv"))
write_csv(ChopHeader, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2007.csv"))
write_csv(cwt, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTTag_2007.csv"))
write_csv(cwt_header, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2007.csv"))
```


## Raw Data Glimpse: {.tabset}

### ChopChannel_raw
```{r}
ChopChannel_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopChannel_2007.csv")) |> 
  glimpse()
```

### ChopHeader_raw

```{r}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopHeader_2007.csv")) |> 
  glimpse()
```

### ChopRecov_raw
```{r}
ChopRecov_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "ChopRecov_2007.csv")) |> 
  glimpse()
```
### cwt_raw
```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "CWTTag_2007.csv")) |> 
  glimpse()
```

### cwt_header_raw

```{r}
cwt_header_raw <- read_csv(here::here("data-raw", "qc-markdowns",
                                 "adult-holding-redd-and-carcass-surveys",
                                 "feather-river", "CWTHeader_2007.csv")) |> 
  glimpse()
```



## Data transformations: {.tabset}

### Counts

The `chop_channel` table contains carcass counts by chop/tagged based on clips


```{r}

#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw |> 
                                 select(HeaderID = headerID, TagCol, Date, Time),
                               ChopChannel_raw,
                               by = "HeaderID") |> 
  clean_names()
# ChopChannel_raw has multiple rows per "ID". This translates to the
# chop_channel_join table.
```

### Recovery

The `chop_recovery` table contains recovered carcass counts by color

```{r}
# 2. choprecovery table (with dates)
# should join headerID to chopchannel headerID, and then chanID from chop channel to ChopRecovery
# HeaderID connects ChopChannel to ChopHeader; ChanID joins ChopChannel to ChopRecov
chop_recovery_join <- full_join(chop_channel_join |> 
                                  select(date, time, sect, chan, min, chan_id, header_id,
                                         tag_col), ChopRecov_raw |> 
                                  rename(chan_id = ChanID), 
                                by = "chan_id") |> 
  clean_names() |> 
  glimpse()

# chop_recovery_join <- full_join(ChopHeader_raw |> 
#                                   select(ID, Date, Time),
#                                 ChopRecov_raw |> 
#                                   rename(ID = RecovID), by = "ID")  |> 
#   clean_names()
```

### Survey

The `chop_header` table contains survey metadata and covariates

```{r}
chop_header <- ChopHeader_raw |> 
  clean_names()
```

### CWT
The `cwt` table contains coded wire tag information. 

```{r}
cwt <- full_join(cwt_raw |> clean_names(), cwt_header_raw |> clean_names(),
                 by = "header_id") |>  
  glimpse()
```


## Explore Numeric Variables: {.tabset}

### Chop Channel Variable: `header_id`, `chan_id`, `sect`, `min`
```{r}
# Filter clean data to show only numeric variables 
chop_channel_join |> 
  select_if(is.numeric) |>
  colnames()
```

```{r}
summary(chop_channel_join$header_id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

### Chop Channel Variable: `chop_clip`, `chop_n_clip`,`chop_uncheck`

**Numeric Summary of `chop_clip`, `chop_n_clip`,`chop_uncheck` over Period of Record**

```{r}
summary(chop_channel_join$chop_clip)
```

```{r}
summary(chop_channel_join$chop_n_clip)
```

```{r}
summary(chop_channel_join$chop_n_clip)
```

```{r}
summary(chop_channel_join$chop_uncheck)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$chop_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$chop_n_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$chop_uncheck))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$chop_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `chop_clip` column are NA.
* `r round(sum(is.na(chop_channel_join$chop_n_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `chop_n_clip` column are NA.
* `r round(sum(is.na(chop_channel_join$chop_uncheck))/nrow(chop_channel_join), 3) * 100` % of values in the `chop_uncheck` column are NA.


**Plotting chop_clip over Period of Record**

```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(chop_clip = sum(chop_clip)) |> 
  ggplot(aes(x = date, y = chop_clip)) +
  geom_col(fill = "blue") + 
  labs(x = "Date", 
       y = "Chop With Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```


**Plotting chop_n_clip over Period of Record**

```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(chop_n_clip = sum(chop_n_clip)) |> 
  ggplot(aes(x = date, y = chop_n_clip)) +
  geom_col(fill = "red") + 
  labs(x = "Date", 
       y = "Chop Without Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting chop_uncheck over Period of Record**

```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(chop_uncheck = sum(chop_uncheck)) |> 
  ggplot(aes(x = date, y = chop_uncheck)) +
  geom_col(fill = "black") + 
  labs(x = "Date", 
       y = "Chop Not Checked") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting Total chops over Period of Record**

```{r}
total_chops <- chop_channel_join|>
  mutate(chop_clip = ifelse(is.na(chop_clip), 0, chop_clip), # fill na
         chop_n_clip = ifelse(is.na(chop_n_clip), 0, chop_n_clip),
         chop_uncheck = ifelse(is.na(chop_uncheck), 0, chop_uncheck),
         total_chops = chop_clip + chop_n_clip + chop_uncheck) |> 
  select(date, chop_clip, chop_n_clip, chop_uncheck, total_chops)

# Case check: checking for duplicates
total_chops_test <- total_chops |> # get total chops
  filter(date == as_date("2007-09-15"))

# Check if NAs introduced in creating total chop columns
nrow(total_chops_test) == (sum(total_chops_test$total_chops == (total_chops_test$chop_clip + total_chops_test$chop_n_clip + total_chops_test$chop_uncheck)))

# total number of chops for the day
total_chops_summary <- total_chops |>
  group_by(date) |>
  summarise(total_chops = sum(total_chops, na.rm = T),
            chop_clip = sum(chop_clip, na.rm = T),
            chop_n_clip = sum(chop_n_clip, na.rm = T),
            chop_uncheck = sum(chop_uncheck, na.rm = T))



# check if NAs are introduced in merge
A = left_join(ChopChannel_raw, ChopHeader_raw |> rename(HeaderID = headerID), 
              by = "HeaderID")$ChanID
B = ChopRecov_raw$ChanID
sum(B %in% A) # they are
# yes I noticed that. there are some IDs in the chop_channel table that are not in the chop_header table. 
```
```{r}
total_chops_summary |> 
  pivot_longer(cols = c(chop_clip, chop_n_clip, chop_uncheck), names_to = "clip", values_to = "count") |> 
  mutate(proportions = (count / total_chops)) |> 
  ggplot(aes(x = date, y = proportions, fill = clip)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Clip", 
                    labels = c("Clip", "No Clip", "Not Checked")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```


The dataset from 2007 did not record for sex and instead, recorded for clip(?) or not.This plot shows the  proportion of chops with, without clips and unchecked for each day over the the period of 2007-09-04 to 2007-12-05. Most of the chops after October were not checked for clips.
The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. We see that the majority of the chops were unchecked for clips.

### Chop Channel Variable: `tag_count`, `tag_clip`, `tag_n_clip`, `tag_unk`

**Numeric Summary of `tag_clip`, `tag_n_clip`, `tag_unk` over Period of Record**

```{r}
summary(chop_channel_join$tag_clip)
```

```{r}
summary(chop_channel_join$tag_n_clip)
```

```{r}
summary(chop_channel_join$tag_unk)
```


**NA and Unknown Values**
Provide a stat on NA or unknown values

```{r}
round(sum(is.na(chop_channel_join$tag_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$tag_n_clip))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$tag_unk))/nrow(chop_channel_join), 3) * 100
```

* `r round(sum(is.na(chop_channel_join$tag_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_clip` column are NA.
* `r round(sum(is.na(chop_channel_join$tag_n_clip))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_n_clip` column are NA.
  * `r round(sum(is.na(chop_channel_join$tag_unk))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_unk` column are NA.

**Plotting tag_clip over Period of Record**
```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(tag_clip = sum(tag_clip)) |> 
  ggplot(aes(x = date, y = tag_clip)) +
  geom_col(fill = "blue") + 
  labs(x = "Date", 
       y = "Tag Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting tag_n_clip over Period of Record**

```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(tag_n_clip = sum(tag_n_clip)) |> 
  ggplot(aes(x = date, y = tag_n_clip)) +
  geom_col(fill = "red") + 
  labs(x = "Date", 
       y = "Tag No Clip") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting tag_unk over Period of Record**
```{r}
chop_channel_join |> 
  group_by(date) |> 
  summarise(tag_unk = sum(tag_unk)) |> 
  ggplot(aes(x = date, y = tag_unk)) +
  geom_col(fill = "black") + 
  labs(x = "Date", 
       y = "Unknown Tags") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Plotting Total Tags over Period of Record**

```{r}
# Percentage by Gender of Daily Chops
# Address NA values for chops
total_tags <- chop_channel_join |>
  mutate(tag_clip = ifelse(is.na(tag_clip), 0, tag_clip), # fill na
         tag_n_clip = ifelse(is.na(tag_n_clip), 0, tag_n_clip),
         tag_unk = ifelse(is.na(tag_unk), 0, tag_unk),
         total_tags = tag_unk + tag_clip + tag_n_clip) |> 
  select(date, tag_clip, tag_n_clip, tag_unk, total_tags)

# total number of chops for the day
total_tags_summary <- total_tags |>
  group_by(date) |>
  summarise(total_tags = sum(total_tags, na.rm = T),
            tag_clip = sum(tag_clip, na.rm = T),
            tag_n_clip = sum(tag_n_clip, na.rm = T),
            tag_unk = sum(tag_unk, na.rm = T))

# Make Graph to identify clip proportion
total_tags_summary |> 
  pivot_longer(cols = c(tag_n_clip, tag_clip, tag_unk), names_to = "tags", values_to = "count") |> 
  mutate(proportions = (count / total_tags)) |> 
  ggplot(aes(x = date, y = proportions, fill = tags)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Clips", 
                    labels = c("Tags Without Clips", "Tags With Clips", "Tags Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

**Comparing tags and chops**

```{r}
total_carcass <- full_join(total_chops_summary, total_tags_summary) |> 
  select(date, total_chops, total_tags) |> 
  pivot_longer(c(total_chops, total_tags), names_to = "mark")

proportion_tags <- total_carcass |> 
  pivot_wider(names_from = mark,
              values_from = value) |> 
  mutate(prop_tags = total_tags/total_chops) |> 
  select(prop_tags)
summary(proportion_tags)

ggplot(total_carcass, aes(x = date, y = value, fill = mark)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```

Tags appear to be less commonly applied compared to chops.

### Chop Recovery Variable: `header_id`, `chan_id`, `recov_count`

```{r}
# Filter clean data to show only numeric variables 
chop_recovery_join |> 
  select_if(is.numeric) |>
  colnames()
```
```{r}
summary(chop_recovery_join$header_id)
```
```{r}
summary(chop_recovery_join$recov_count)
```

* Note: all recov_count do not have dates associated with them

### Chop Header Variable: `header_id`

```{r}
chop_header |> 
  select_if(is.numeric) |> 
  colnames()
```
```{r}
summary(chop_header$header_id)
```

### CWT Variable: `tag_id`, `sect`, `fl`, `header_id`, `week_num`
```{r}
cwt |> 
  select_if(is.numeric) |> 
  colnames()
```

```{r}
summary(cwt$tag_id)
```
```{r}
summary(cwt$sect)
```
```{r}
summary(cwt$fl)
```
* `r round(sum(is.na(cwt$tag_id))/nrow(cwt), 3) * 100` % of values in the `tag_id` column are NA.
* `r round(sum(is.na(cwt$sect))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.
* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
cwt_count <- cwt |> 
  mutate(count = 1) |>
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "U",
                         TRUE ~ sex)) |> 
  pivot_wider(names_from = sex, values_from = count, values_fill = 0) |> 
  # unnest() |> 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) |> glimpse()

total_cwt_summary <- cwt_count |> 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) |> 
  group_by(month(date)) |> 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary |> 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") |> 
  mutate(proportions = (count / total_cwt)) |> 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Month") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```


**Plotting fork length of each sex**

```{r}
cwt |> 
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "Unknown",
                         TRUE ~ sex))|> 
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```


## Explore Categorical variables: {.tabset}

### Chop Channel Clean Data

Fix inconsistencies with spelling, capitalization, and dates
```{r}
# Filter clean data to show only categorical variables
chop_channel_join |> 
  select_if(is.character) |>
  colnames()
```

```{r}
unique(chop_channel_join$chan)
```
```{r}
chop_channel_cleaner <- chop_channel_join |>
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time, male, fem, gril, tag_count)) # male, female, gril, tag_count columns are NAs

chop_channel_cleaner
```
### Chop Recovery Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_recovery_join |> 
  select_if(is.character) |> 
  colnames()
```

```{r}
chop_recovery_cleaner <- chop_recovery_join |> 
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time)) |> glimpse()
```

### Chop Header Clean Data

```{r}
chop_header |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$recorder)
```
```{r}
unique(chop_header$weather)
```



```{r}
chop_header_cleaner <- chop_header |>
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         weather = case_when(weather == "cld" ~ "cloudy",
                             TRUE ~ weather)) |> 
  select(-c(time)) |> glimpse()
```

### CWT Clean Data

```{r}
cwt |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(cwt$rel_loc)
```
```{r}
unique(cwt$spawn)
unique(cwt$clip)
unique(cwt$samp_type)
unique(cwt$carc_stat)

```

```{r}
cwt_cleaner <- cwt |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         sex = case_when(is.na(sex) ~ "unknown",
                         TRUE ~ sex),
         spawn = case_when(spawn %in% c("uk", "no data") ~ NA_character_,
                           TRUE ~ spawn),
         clip = case_when(clip %in% c("no data", "uk") ~ NA_character_,
                          clip == "y" ~ "yes",
                          clip == "n" ~ "no",
                          TRUE ~ clip),
         samp_type = ifelse(samp_type == "none", NA, samp_type),
         carc_stat = ifelse(carc_stat == "no data", NA, carc_stat)) |> 
  glimpse()
```

## Comments

* Data set similar to 2007 and 2007 data

## Data Dictionaries

### Channel

```{r}
percent_na <- chop_channel_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("ID",
                                          "Color of tag applied to carcass",
                                          "Date of survey",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped and has clip",
                                          "Carcass that were chopped and has no clip",
                                          "Carcass that were chopped and not checked for clip",
                                          "Carcass that were tagged and has clip",
                                          "Carcass that were tagged and has no clip",
                                          "Carcass that were tagged and unknown"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
percent_na <- chop_recovery_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


recovery_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("Date of Survey",
                                          "Sect",
                                          "Chan",
                                          "Min",
                                          "Chan ID",
                                          "Header ID",
                                          "Color of tag recovered from carcass",
                                          "Count of recovery",
                                          "Color of recovery tag", 
                                          "Count of recvered tags"),
                          percent_na = round(percent_na$value*100))

kable(recovery_data_dictionary)
```

### Survey

```{r}

percent_na <- chop_header_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())
# 
header_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Date of survey",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number"),
                          percent_na = round(percent_na$value*100))
# 
kable(header_data_dictionary)
```
### CWT
```{r}

percent_na <- cwt_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())

cwt_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Sect",
                                          "Release Location",
                                          "Fork length",
                                          "Sex",
                                          "Spawn",
                                          "Clip",
                                          "Sample type",
                                          "Carcass stat",
                                          "Sample number",
                                          "Head tag",
                                          "Comment",
                                          "Header ID",
                                          "Hall print 1",
                                          "Hall print 2",
                                          "Date",
                                          "Tag colour",
                                          "Crew",
                                          "Week number"),
                          percent_na = round(percent_na$value*100))

kable(cwt_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)

```{r}
#Ignore chop recovery as its 100% NA
feather_carcass_chops_and_tags_2007 <- chop_channel_cleaner |> glimpse()
feather_carcass_cwt_2007 <- cwt_cleaner |> glimpse()
feather_carcass_chop_header_2007 <- chop_header_cleaner |> glimpse()
```
```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_carcass_chops_and_tags_2007,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chops_and_tags_2007.csv")
gcs_upload(feather_carcass_cwt_2007,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_cwt_2007.csv")
gcs_upload(feather_carcass_chop_header_2007,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_header_2007.csv")
```

