---
title: "Feather Carcass QC 2004"
author: "Inigo Peng"
date: '2022-07-25'
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data
```{r, include=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```

```{r, include=FALSE}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/Carcass_2004.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2004.mdb",
               overwrite = TRUE)

filepath <- "data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_2004.mdb"

operating_system <- ifelse(grepl("Mac", Sys.info()['nodename']) | grepl("MBP", Sys.info()['nodename']), "mac", "pc")

# Mac and PC need to run different code to pull data from Access db
if(operating_system == "pc") {
  DRIVERINFO <- "Driver={Microsoft Access Driver (*.mdb, *.accdb)};"
  PATH <- paste0(DRIVERINFO, "DBQ=", filepath)
  con <- odbcDriverConnect(PATH)
  # sqlTables(con)$TABLE_NAME
  ChopChannel <- sqlFetch(con, "ChopChannelTBL")
  ChopRecov <- sqlFetch(con, "ChopRecovTBL")
  ChopHeader <- sqlFetch(con, "ChopHeaderTBL")
  cwt <- sqlFetch(con, "CWTTagTBL")
  cwt_header <- sqlFetch(con, "CWTHeaderTBL")
} else{
  library(Hmisc)
  mdb.get(filepath, tables = TRUE) # check for name differences
  ChopChannel <- mdb.get(filepath, "ChopChannelTBL")
  ChopRecov <- mdb.get(filepath, "ChopRecovTBL")
  ChopHeader <- mdb.get(filepath, "ChopHeaderTBL")
  cwt <- mdb.get(filepath, "CWTTagTBL")
  cwt_header <- mdb.get(filepath, "CWTHeaderTBL")
  
  # Steelhead is empty
  # Fish Data_SR_Database has information on TagID (doesn't match with any other datasets), Sample ID (same as TagID), adipose clip (0/1), Tissue (0/1), Fork Length
  # TagColLU contains metadata about tag colors
  # Sample table SR Database - mortality and sample ID that is in same range
  # as Fish Data_SR_Database SampleID
  detach(package:Hmisc)
}

write_csv(ChopChannel, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopChannel_2004.csv"))
write_csv(ChopRecov, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2004.csv"))
write_csv(ChopHeader, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2004.csv"))
write_csv(cwt, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTTag_2004.csv"))
write_csv(cwt_header, here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2004.csv"))
```


## Raw Data Glimpse: {.tabset}

### cwt_raw
```{r}
cwt_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTTag_2004.csv")) |> glimpse()
```

### CWT_Header_raw
```{r}
cwt_header_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "CWTHeader_2004.csv")) |> glimpse()
```

### ChopChannel_Raw
```{r}
ChopChannel_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopChannel_2004.csv")) |> 
  glimpse()
```

### ChopRecov_raw
```{r}
ChopRecov_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopRecov_2004.csv")) |> glimpse()
```

### ChopHeader_Raw
```{r}
ChopHeader_raw <- read_csv(here::here("data-raw", "qc-markdowns", "adult-holding-redd-and-carcass-surveys", "feather-river", "ChopHeader_2004.csv")) |> 
  glimpse()
```



## Data transformations: {.tabset}

### Counts

The `chop_channel` table contains carcass counts by sex, grilise and chop/tagged

```{r}
#1. chopchannel table (with dates and tag color)
chop_channel_join <- full_join(ChopHeader_raw |> 
                                 select(HeaderID = headerID, TagCol, Date, Time),
                               ChopChannel_raw, 
                               by = "HeaderID") |> 
  clean_names() |> 
  rename("male_chop" = male,
         "female_chop" = fem,
         "grilse" = gril) |> glimpse()
```
### Recovery

The `chop_recovery` table contains recovered carcass counts by color

```{r}
# HeaderID connects ChopChannel to ChopHeader; ChanID joins ChopChannel to ChopRecov
chop_recovery_join <- full_join(chop_channel_join |> 
                                  select(date, time, sect, chan, min, chan_id, header_id,
                                         tag_col), ChopRecov_raw |> 
                                  rename(chan_id = ChanID), 
                                by = "chan_id") |> 
  clean_names() |> 
  glimpse()
# chop_recovery_join <- full_join(ChopHeader_raw |> 
#                                   select(headerID, Date, Time),
#                                 ChopRecov_raw |> 
#                                   rename(headerID = ChanID),
#                                 by = "headerID")  |> # ChanID <-> HeaderID - not correct
#   clean_names() |> glimpse()
```

### Survey

The `chop_header` table contains survey metadata and covariates 
```{r}
chop_header <- ChopHeader_raw |> 
  clean_names() |> glimpse()
```
  
### CWT
The `cwt` table contains carcass sampling information on carcasses, both chops and tagged carcasses. Each `tag_id` is represented only once in this table.

```{r}
cwt <- full_join(cwt_raw |> clean_names(), cwt_header_raw |> clean_names()) |> 
  clean_names() |> 
  glimpse()
```


## Explore Numeric Variables: {.tabset}

#### Chop Channel Variable:`id`, `chan_id`, `sect`, `min`
```{r}
# Filter clean data to show only numeric variables 
chop_channel_join |> 
  select_if(is.numeric) |>
  colnames()
```


```{r}
summary(chop_channel_join$header_id)
```
```{r}
summary(chop_channel_join$chan_id)
```
```{r}
summary(chop_channel_join$sect)
```

```{r}
summary(chop_channel_join$min)
```

### Chop Channel Variable: `female_chop`, `male_chop`,`grilise`, `tag_count`

**Numeric Summary of `female_chop`, `male_chop`,`unknown_chops`, `tag_count` over Period of Record**

```{r}
summary(chop_channel_join$female_chop)
```

```{r}
summary(chop_channel_join$male_chop)
```

```{r}
summary(chop_channel_join$grilse)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_channel_join$male_chop))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$female_chop))/nrow(chop_channel_join), 3) * 100
round(sum(is.na(chop_channel_join$grilse))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$male_chop))/nrow(chop_channel_join), 3) * 100` % of values in the `male_chop` column are NA.
* `r round(sum(is.na(chop_channel_join$female_chop))/nrow(chop_channel_join), 3) * 100` % of values in the `female_chop` column are NA.
* `r round(sum(is.na(chop_channel_join$grilse))/nrow(chop_channel_join), 3) * 100` % of values in the `grilse` column are NA.

**Plotting total chops over Period of Record**

```{r}
# Make whatever plot is appropriate 
chop_channel_join |> 
  group_by(date) |> 
  summarise(total_chop = sum(female_chop, male_chop, grilse)) |> 
  ggplot(aes(x = date, y = total_chop)) +
  geom_col() +
  theme_minimal() + 
  labs(x = "Date",
       y = "Total Chops") + 
  theme(text = element_text(size = 15)) 

```

**Plotting Total chops over Period of Record**

```{r}
chop_channel_join |> glimpse()

# Percentage by Gender of Daily Chops
# Address NA values for chops
total_chops <- chop_channel_join|>
  mutate(male_chop = ifelse(is.na(male_chop), 0, male_chop), # fill na
         female_chop = ifelse(is.na(female_chop), 0, female_chop),
         grilse_chop = ifelse(is.na(grilse), 0, grilse),
         total_chops = grilse_chop + male_chop + female_chop) |> 
  select(date, male_chop, female_chop, grilse_chop, total_chops)

# Case check: checking for duplicates
total_chops_test <- total_chops |> # get total chops
  filter(date == as_date("2004-09-20"))

# Check if NAs introduced in creating total chop columns
nrow(total_chops_test) == (sum(total_chops_test$total_chops == (total_chops_test$male_chop + total_chops_test$female_chop + total_chops_test$grilse_chops)))

# total number of chops for the day
total_chops_summary <- total_chops |>
  group_by(date) |>
  summarise(total_chops = sum(total_chops, na.rm = T),
            male_chops = sum(male_chop, na.rm = T),
            female_chops = sum(female_chop, na.rm = T),
            grilse_chops = sum(grilse_chop, na.rm = T))
```

```{r}
total_chops_summary |> 
  pivot_longer(cols = c(male_chops, female_chops, grilse_chops), names_to = "sex", values_to = "count") |> 
  mutate(proportions = (count / total_chops)) |> 
  ggplot(aes(x = date, y = proportions, fill = sex)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("Blue","Red", "Black"), 
                    name = "Sex", 
                    labels = c("Female", "Male", "Grilse")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))

```

This plot shows the proportion of chops based on sex and grilse for each day over the the period of 2004-09-03 to 2004-12-19. The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period. 

```{r}
summary(chop_channel_join$tag_count)
```

```{r}
round(sum(is.na(chop_channel_join$tag_count))/nrow(chop_channel_join), 3) * 100
```
* `r round(sum(is.na(chop_channel_join$tag_count))/nrow(chop_channel_join), 3) * 100` % of values in the `tag_count` column are NA.

**Plotting tag_count over Period of Record**

```{r}
total_tags_summary <- chop_channel_join |>
  group_by(date) |>
  summarise(total_tags = sum(tag_count, na.rm = T))
total_tags_summary |> 
  ggplot(aes(x = date, y = total_tags)) +
  geom_col(fill = "blue") +
  theme_minimal() +
  labs(y = "Daily Total Tags", x = "Date") + 
  theme(text = element_text(size = 15)) 
```

**Comparing tags and chops**

```{r}
# from ashley: please improve on this plot I got started, make it look better, describe it, double check that the numbers make sense
total_carcass <- full_join(total_chops_summary, total_tags_summary) |> 
  select(date, total_chops, total_tags) |> 
  pivot_longer(c(total_chops, total_tags), names_to = "mark")

proportion_tags <- total_carcass |> 
  pivot_wider(names_from = mark, 
              values_from = value) |> 
  group_by(date) |> 
  mutate(total_collected = sum(total_chops, total_tags)) |> 
  summarise(proportion_tagged = total_tags/total_collected)
summary(proportion_tags$proportion_tagged)

ggplot(total_carcass, aes(x = date, y = value, fill = mark)) +
  geom_col() +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  labs(x = "Date", y = "Count")

```
Tags are significantly less applied compared to chops.

### Chop Recovery Variable: `id`, `chan_id`, `recov_count`

```{r}
chop_recovery_join |> 
  select_if(is.numeric) |> 
  colnames()
```

```{r}
summary(chop_recovery_join$header_id)
```

```{r}
summary(chop_recovery_join$chan_id)
```

```{r}
summary(chop_recovery_join$recov_count)
```


### Chop Header Variable: `id`

```{r}
chop_header |> 
  select_if(is.numeric) |> 
  colnames()
```
```{r}
summary(chop_header$header_id)
```

### CWT Variable: `tag_id`, `sect`, `fl`, `header_id`, `week_num`
```{r}
cwt |> 
  select_if(is.numeric) |> 
  colnames()
```
```{r}
summary(cwt$tag_id)
```
```{r}
summary(cwt$sect)
```
```{r}
summary(cwt$fl)
```
* `r round(sum(is.na(cwt$tag_id))/nrow(cwt), 3) * 100` % of values in the `tag_id` column are NA.
* `r round(sum(is.na(cwt$sect))/nrow(cwt), 3) * 100` % of values in the `sect` column are NA.
* `r round(sum(is.na(cwt$fl))/nrow(cwt), 3) * 100` % of values in the `fl` column are NA.

```{r}
#Create a cwt_count column
#Pivot table to expand sex column to female_cwt, male_cwt, and unknown_cwt 
#Is this graph helpful?
cwt_count <- cwt |> 
  mutate(count = 1) |>
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "U",
                         TRUE ~ sex)) |> 
  pivot_wider(names_from = sex, values_from = count, values_fill = 0) |> 
  # unnest() |> 
  rename("male_cwt" = M,
         "female_cwt" = F,
         "unknown_cwt" = U) |> glimpse()

total_cwt_summary <- cwt_count |> 
  mutate(male_cwt = ifelse(is.na(male_cwt), 0, male_cwt), # fill na
         female_cwt = ifelse(is.na(female_cwt), 0, female_cwt),
         unknown_cwt = ifelse(is.na(unknown_cwt), 0, unknown_cwt),
         total_cwt = unknown_cwt + male_cwt + female_cwt) |> 
  group_by(month(date)) |> 
  summarise(total_cwt = sum(total_cwt),
            male_cwt = sum(male_cwt),
            female_cwt = sum(female_cwt),
            unknown_cwt = sum(unknown_cwt))
```

```{r}
total_cwt_summary |> 
  pivot_longer(cols = c(male_cwt, female_cwt, unknown_cwt), names_to = "sex", values_to = "count") |> 
  mutate(proportions = (count / total_cwt)) |> 
  ggplot(aes(x = `month(date)`, y = proportions, fill = sex)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(name = "chops", 
                    labels = c("CWT Male", "CWT Female", "CWT Unknown")) +
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```

**Plotting fork length of each sex**

```{r}
cwt |> 
  mutate(sex = case_when(sex == "ND"|is.na(sex)|sex =="UK" ~ "Unknown",
                         TRUE ~ sex))|> 
  filter(fl < 600) |> # one individual has fl = 777
  ggplot(aes(x = sex, y = fl)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(y = "FL", x = "Sex")
```

## Explore Categorical variables: {.tabset}

### Chop Channel Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_channel_join |> 
  select_if(is.character) |>
  colnames()
```
```{r}
unique(chop_channel_join$tag_col)
```

```{r}
unique(chop_channel_join$chan)
```

```{r}
chop_channel_cleaner <- chop_channel_join |>
  mutate(date = as_date(date), ) |>
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time))

chop_channel_cleaner
```

### Chop Recovery Clean Data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_recovery_join |> 
  select_if(is.character) |> 
  colnames()
```

```{r}
unique(chop_recovery_join$recov_color)
```
```{r}
chop_recovery_cleaner <- chop_recovery_join |> 
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  select(-c(time))

chop_recovery_cleaner
```

### Chop Header Clean Data

```{r}
chop_header |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(chop_header$crew)
```
```{r}
unique(chop_header$recorder)
```
```{r}
unique(chop_header$weather)
```



```{r}
chop_header_cleaner <- chop_header |>
  mutate(date = as_date(date)) |>
  mutate_if(is.character, str_to_lower) |> 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         weather = case_when(weather == "ran" ~ "rain",
                             weather == "cld" ~ "cloudy",
                             TRUE ~ weather)) |>  
  select(-c(time))

chop_header_cleaner
```

### CWT Clean Data

```{r}
cwt |> 
  select_if(is.character) |> 
  colnames()
```
```{r}
unique(cwt$rel_loc)
```
```{r}
unique(cwt$spawn)
```
```{r}
unique(cwt$clip)
```
```{r}
unique(cwt$samp_type)
```
```{r}
unique(cwt$carc_stat)
```


```{r}
cwt_cleaner <- cwt |> 
  mutate_if(is.character, str_to_lower) |> 
  mutate(crew = str_replace_all(crew, " ", ","),
         crew = str_replace_all(crew, ",,", ","),
         sex = case_when(sex %in% c("nd", "uk") ~ "unknown",
                         sex == "m" ~ "male",
                         sex == "f" ~ "female",
                         TRUE ~ sex),
         rel_loc = ifelse(rel_loc == "no data", NA, rel_loc),  
         spawn = case_when(spawn %in% c("no data") ~ NA_character_,
                           spawn == "uk" ~ "unknown",
                           spawn == "u" ~ "unspawned",
                           spawn == "s" ~ "spawned",
                           TRUE ~ spawn),
         clip = case_when(clip %in% c("no data") ~ NA_character_,
                          clip == "y" ~ "yes",
                          clip == "n" ~ "no",
                          clip == "uk" ~ "unknown",
                          TRUE ~ clip),
         samp_type = ifelse(samp_type == "no data", NA, samp_type),
         carc_stat = case_when(carc_stat == "no data" ~ NA_character_,
                               carc_stat == "c" ~ "chopped",
                               carc_stat == "t" ~ "tagged",
                               TRUE ~ carc_stat)) |> 
  glimpse()
```

## Data Dictionaries

### Channel

```{r}
percent_na <- chop_channel_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_channel_cleaner),
                          description = c("ID",
                                          "Color of tag applied to carcass",
                                          "Date of survey",
                                          "Channel ID",
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Carcass that were chopped and male",
                                          "Carcass that were chopped and female",
                                          "Carcass that were grilse",
                                          "Carcass that were tagged"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```

### Recovery

```{r}
# from ashley: please insert data dictionary for cleaned recovery table
percent_na <- chop_recovery_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())


chop_recovery_data_dictionary <- tibble(variables = colnames(chop_recovery_cleaner),
                          description = c("Date of survey", 
                                          "Sect", 
                                          "Chan", 
                                          "Min", 
                                          "Channel ID",
                                          "Header ID",
                                          "Color of tag recovered from carcass",
                                          "Recovery ID",
                                          "Color of recovery tag",
                                          "Count of recovery"),
                          percent_na = round(percent_na$value*100))

kable(chop_recovery_data_dictionary)
```

### Survey

```{r}
# 
percent_na <- chop_header_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())
# 
survey_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Color of tag of carcass",
                                          "Date of survey",
                                          "Crew memeber initials that collected",
                                          "Individual of crew member who recorded",
                                          "Weather",
                                          "Comments",
                                          "Week Number"),
                          percent_na = round(percent_na$value*100))

kable(survey_data_dictionary)
```
### CWT
```{r}

percent_na <- cwt_cleaner |>
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |>
  pivot_longer(cols = everything())

cwt_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Sect",
                                          "Release Location",
                                          "Fork length",
                                          "Sex",
                                          "Spawn",
                                          "Clip",
                                          "Sample type",
                                          "Carcass stat",
                                          "Sample number",
                                          "Head tag",
                                          "Comment",
                                          "Header ID",
                                          "Date",
                                          "Tag colour",
                                          "Crew",
                                          "Week number"),
                          percent_na = round(percent_na$value*100))

kable(cwt_data_dictionary)
```
## Save cleaned data back to google cloud (TBA)


```{r}
feather_carcass_chops_and_tags_2004 <- chop_channel_cleaner |> glimpse()
feather_carcass_chop_recovery_2004 <- chop_recovery_cleaner |> glimpse()
feather_carcass_cwt_2004 <- cwt_cleaner |> glimpse()
feather_carcass_chop_header_2004 <- chop_header_cleaner |> glimpse()
```
```{r, include=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_carcass_chops_and_tags_2004,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chops_and_tags_2004.csv")
gcs_upload(feather_carcass_chop_recovery_2004,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_recovery_2004.csv")
gcs_upload(feather_carcass_cwt_2004,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_cwt_2004.csv")
gcs_upload(feather_carcass_chop_header_2004,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_chop_header_2004.csv")

```
