---
title: "Feather Carcass QC 2014"
author: "Inigo Peng"
date: '2022-07-21'
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```


# Feather River Carcass Data

## Description of Monitoring Data

**Timeframe:** 
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
  
**Sampling Location:**
  
**Data Contact:** 
  
Any additional info?
  
## Access Cloud Data
  
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
Sys.setenv("GCS_AUTH_FILE" = "C:/Users/InigoPeng/Projects/jpe/JPE-datasets/config.json")
Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```
```{r}
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2014/Chops_2014.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "Chops_2014.xlsx",
               overwrite = TRUE)
#
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2014/ChopHeader_2014.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "ChopHeader_2014.xlsx",
               overwrite = TRUE)

# 
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2014/CWTHeader_2014.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWTHeader_2014.xlsx",
               overwrite = TRUE)

gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/feather-river/data-raw/carcass/2014/CWT_2014.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "CWT_2014.xlsx",
               overwrite = TRUE)
# 
Chops_raw <- read_excel("Chops_2014.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  select(-`Chop ID`) %>%
  glimpse()


ChopHeader_raw <- read_excel("ChopHeader_2014.xlsx") %>% 
  rename("ID" = `Chop Header ID`) %>%
  glimpse()


#Coded Wiretag Info

cwt_raw <- read_excel("CWT_2014.xlsx") %>% 
  rename("ID" = `CWT Header ID`) %>% 
  glimpse

cwt_header_raw <- read_excel("CWTHeader_2014.xlsx") %>% 
  rename("ID" = `CWT Header ID` ) %>% 
  glimpse
```

## Data transformations

```{r}
chop_join <- full_join(ChopHeader_raw %>% 
                                 select(ID, Date, Time),
                               Chops_raw) %>% 
  clean_names() %>% 
  rename(sec = "section",
         min = "minutes",
         count = "total_count") %>% 
  mutate(sec = as.numeric(sec)) %>% glimpse

chop_header <- ChopHeader_raw %>% 
  clean_names()%>%  glimpse

cwt_join <- full_join(cwt_header_raw, cwt_raw) %>% 
  clean_names() %>% 
  mutate(head_tag_number = as.numeric(head_tag_number),
         tag_id_number = as.numeric(tag_id_number),
         scales = as.numeric(scales),
         otoliths = as.numeric(otoliths))%>% glimpse()
```

### Counts

The `chop_join` table contains carcass counts 

### Survey

The `chop_header` table contains survey metadata and covariates

### CWT

The `cwt_join` table contains coded wire tag data

## Explore Numeric Variables: {.tabset}

```{r}
chop_join %>% 
  select_if(is.numeric) %>%
  colnames()
```

### Variable: `id`, `min`, `sec`

```{r}
summary(chop_join$id)
```

```{r}
summary(chop_join$min)
```

```{r}
summary(chop_join$sec)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100
round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100
round(sum(is.na(chop_join$sec))/nrow(chop_join), 3) * 100

```

* `r round(sum(is.na(chop_join$id))/nrow(chop_join), 3) * 100` % of values in the `id` column are NA.
* `r round(sum(is.na(chop_join$min))/nrow(chop_join), 3) * 100` % of values in the `min` column are NA.
* `r round(sum(is.na(chop_join$sec))/nrow(chop_join), 3) * 100` % of values in the `sec` column are NA.

### Variable: `count`
```{r}
summary(chop_join$count)
```

* `r round(sum(is.na(chop_join$count))/nrow(chop_join), 3) * 100` % of values in the `count` column are NA.

**Plotting count over Period of Record**

```{r}
chop_join %>% 
  ggplot(aes(x = date, y = count)) + 
  geom_point() +
  theme_minimal()
```
```{r}
chop_join %>%
  group_by(date) %>%
  summarise(total_count = sum(count, na.rm = T)) %>%
  ggplot(aes(x = date, y = total_count)) +
  geom_col() +
  theme_minimal()
```
```{r}
chop_header %>% 
  select_if(is.numeric) %>% 
  colnames()

```
```{r}
summary(chop_header$id)
```


```{r}
cwt_join %>% 
  select_if(is.numeric) %>% 
  colnames()
```
### Variable: `id`, `river_section`

```{r}
summary(cwt_join$id)
```

```{r}
summary(cwt_join$river_section)
```

**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(cwt_join$id))/nrow(cwt_join), 3) * 100
round(sum(is.na(cwt_join$river_section))/nrow(cwt_join), 3) * 100
```

* `r round(sum(is.na(cwt_join$id))/nrow(cwt_join), 3) * 100` % of values in the `id` column are NA.
* `r round(sum(is.na(cwt_join$river_section))/nrow(cwt_join), 3) * 100` % of values in the `river_section` column are NA.

### Variable: `fork_length`
```{r}
summary(cwt_join$fork_length)
```
```{r}
round(sum(is.na(cwt_join$fork_length))/nrow(cwt_join), 3) * 100
```
* `r round(sum(is.na(cwt_join$fork_length))/nrow(cwt_join), 3) * 100` % of values in the `fork_length` column are NA.

**Fork Length based on Sex**

```{r}
cwt_join %>% 
  mutate(sex = ifelse(is.na(sex), "Unknown", sex)) %>% 
  ggplot(aes(x = sex, y = fork_length)) + 
  geom_boxplot() +
  theme_minimal()
```


## Explore Categorical variables: {.tabset}

```{r}
chop_join %>% 
  select_if(is.character) %>%
  colnames()
```

```{r}
chop_header %>% 
  select_if(is.character) %>% 
  colnames()
```

```{r}
cwt_join %>% 
  select_if(is.character) %>% 
  colnames()
```

```{r}
unique(cwt_join$crew)
```
```{r}
unique(cwt_join$spawning_condition)
cwt_join <- cwt_join %>% 
  mutate(spawning_condition = case_when(spawning_condition == "UK" ~ "Unknown",
                                        TRUE ~ spawning_condition)) %>% 
  glimpse
```

```{r}
unique(cwt_join$tag_color)
```

```{r}
unique(cwt_join$sex)
cwt_join <- cwt_join %>% 
  mutate(sex = case_when(is.na(sex) ~ "unknown",
                        TRUE ~ sex)) %>% 
  glimpse
```

```{r}
unique(cwt_join$tag_recapture_or_chop)
cwt_join <- cwt_join %>% 
  mutate(tag_recapture_or_chop = case_when(is.na(adipose_fin_clipped) ~ "unknown",
                                         tag_recapture_or_chop == "T" ~ "tagged",
                                         tag_recapture_or_chop == "R" ~ "recapture",
                                         tag_recapture_or_chop == "C" ~ "chop",
                                         TRUE ~ tag_recapture_or_chop)) %>% 
  glimpse
```

```{r}
unique(cwt_join$adipose_fin_clipped)
cwt_join <- cwt_join %>% 
  mutate(adipose_fin_clipped = case_when(adipose_fin_clipped == "UK" ~ "Unknown",
                                         adiopose_fin_clipped == "Y" ~ "yes",
                                         adiopose_fin_clipped == "N" ~ "no",
                                         TRUE ~ adipose_fin_clipped)) %>% 
  glimpse
```
```{r}
unique(cwt_join$samples_collected)
```
```{r}
unique(cwt_join$hallprint_color)
cwt_join <- cwt_join %>% 
  mutate(hallprint_color = case_when(hallprint_color == "grey" ~ "G",
                                     hallprint_color == "4751" ~ NA_character_,
                                     TRUE ~ hallprint_color)) %>% 
  glimpse()
```
```{r}
unique(cwt_join$comments)
```



## Clean data

Fix inconsistencies with spelling, capitalization, and dates

```{r}
chop_cleaner <- chop_join %>%
  mutate(datetime = paste(as_date(date), time)) %>%
  mutate_if(is.character, str_to_lower) %>%  
  select(-c(date, time))

chop_cleaner
```
```{r}
chop_header_cleaner <- chop_header %>%
  mutate(datetime = paste(as_date(date), time)) %>%
  mutate_if(is.character, str_to_lower) %>% 
  mutate(crew = str_replace_all(crew, "/", ","),
         crew = str_replace_all(crew, " ","")) %>% 
  select(-c(date, time))

chop_header_cleaner
```

```{r}
cwt_cleaner <- cwt_join %>% 
  mutate_if(is.character, str_to_lower) %>%
  mutate(crew = str_replace_all(crew, " ", ",")) %>%
  select(-c(morale, section_group_1_10, section_group_11_15, section_group_16_21, section_group_22_38, )) %>% 
  glimpse
```

## Data Dictionaries

# Count
```{r}
percent_na <- chop_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())


counts_data_dictionary <- tibble(variables = colnames(chop_cleaner),
                          description = c("ID",
                                          "Sect", 
                                          "Min", 
                                          "Count",
                                          "Date and Time of Survey"),
                          percent_na = round(percent_na$value*100))

kable(counts_data_dictionary)
```
### Survey

```{r}

percent_na <- chop_header_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
# 
counts_data_dictionary <- tibble(variables = colnames(chop_header_cleaner),
                          description = c("ID",
                                          "Week Number",
                                          "Weather",
                                          "Crew memeber initials that collected",
                                          "Comments",
                                          "Date and time of survey"),
                          percent_na = round(percent_na$value*100))
# 
kable(counts_data_dictionary)
```

### CWT
```{r}
percent_na <- cwt_cleaner %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())

counts_data_dictionary <- tibble(variables = colnames(cwt_cleaner),
                          description = c("ID",
                                          "Date of survey",
                                          "Crew memeber initials that collected",
                                          "Week number",
                                          "Tag colour",
                                          "CWT ID",
                                          "River section",
                                          "Tag ID Number",
                                          "Carcass tagged, recaptured, or chopped",
                                          "Sex of the carcass",
                                          "Spawning condition",
                                          "Adipose fin clipped or not",
                                          "Sample collected",
                                          "Fork length",
                                          "Head tag number",
                                          "Scales",
                                          "Otoliths",
                                          "Hallprint colour",
                                          "Hallprint",
                                          "Comments"),
                          percent_na = round(percent_na$value*100))
# 
kable(counts_data_dictionary)
```

