---
title: "Feather Carcass QC "
author: "Inigo Peng"
date: '2022-08-18'
output: rmarkdown::github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)

library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
library(janitor)
library(hms) #?as_hms()
library(RODBC)
library(knitr)
library(wesanderson)
```

# Feather River Carcass Data

## Description of Monitoring Data
The Feather River Carcass monitoring program has slightly differing methodology and data entry across the period of record (2000-2020). 

According to later sampling protocol (and meetings indicate this hasn't changed too much over the years), one of three processes occurs when a carcass is encountered: if the carcass is fresh and not tagged, it is tagged and returned to the river (this is a `tagged` carcass); if the carcass is not fresh and not tagged, it is chopped (`chopped`); finally, if the carcass is tagged and fresh, the carcass is recovered and re-released (`recovered`/`recaptured`) or recovered and chopped (`recovered, chopped` or `chopped`). 

Generally, the table layout/structure is similar across four periods: 
* 2000-2001: `RecovTag`, `ChopRecov`, `CWT`, `TagData`
* 2002-2009: `ChopChannel`, `ChopRecov`, `CWT`
* 2010-2016: `Chops`, `CWT` (`ChopRecov` only in 2010)
* 2017-2020: `Chops`, `ChopRecov`, `CWT`

Though the naming changes across these periods, each year generally has a table of individual carcass samples with biological sampling data like `adipose_fin_clip`, `sex`, and `fork_length`; a bulk chops table that contains bulk chop counts without individual sampling information; and sometimes a chop recovery table that contains recovered carcass data.

To standardize across these periods, we created a `source_table` variable to track the different tables and a `disposition` column to track whether the carcass was `chopped`, `tagged`, or `recovered`/`recaptured`.

The goal is to use the data in a modified Cormack Jolly Seber model, which requires three tables as input:
* Individual carcasses and their recapture history
* Bulk chops by week
* Covariates (fork length, sex, etc.)

The CAMP database creates these tables automatically, but data contacts created their own version of these tables in Excel prior to 2017.


**Timeframe:** 2000 to 2020
  
**Video Season:** 
  
**Completeness of Record throughout timeframe:** 
Data is complete across years. Methodology and data storage differs.
  
**Sampling Location:** Various sampling locations on Feather River. Some river miles have a different protocol wherein carcasses are not individually sampled but are always chopped.
  
**Data Contact:** 
Casey Campos


## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx
```


```{r}
read_from_cloud <- function(carcass_data_type, year){
  gcs_get_object(object_name = paste0("adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass_",carcass_data_type, "_",year, ".csv"),
               bucket = gcs_get_global_bucket(),
               saveToDisk = paste0("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_",carcass_data_type,"_",year, ".csv"),
               overwrite = TRUE)
  data <- read.csv(paste0("data-raw/qc-markdowns/adult-holding-redd-and-carcass-surveys/feather-river/feather_carcass_",carcass_data_type, "_", year, ".csv"))  
}
```

## All Carcasses

## 2000
* rename `ht_num` to `head_tag`
* rename `egg_ret` to `spawn`
* rename `fl` to `fork_length`
```{r, message = FALSE}
feather_carcass_2000 <- read_from_cloud("all_carcasses", 2000) |> 
  rename(head_tag = ht_num,
         spawning_condition = egg_ret,
         fork_length = fl) |> 
  select(-c(float, rel_loc2, recorder, number_spawned)) |> 
  mutate(date = as.Date(date)) |> 
  glimpse() 
```

## 2001
* rename `ht_num` to `head_tag`
* rename `egg_ret` to `spawn`
```{r, message = FALSE}
feather_carcass_2001 <- read_from_cloud("all_carcasses", 2001) |> 
  mutate(date = as.Date(date)) |> 
  rename(head_tag = ht_num,
         spawning_condition = egg_ret,
         week = week_num,
         fork_length = fl,
         min = minute) |> 
  select(-c(tag_comp)) |> 
  glimpse() 
```

## 2002
```{r, message = FALSE}
feather_carcass_2002 <- read_from_cloud("all_carcasses", 2002) |> 
  mutate(date = as.Date(date)) |> 
  rename(fork_length = fl,
         samples_collected = samp_type) |> 
  glimpse() 
```

## 2003
```{r, message = FALSE}
feather_carcass_2003 <- read_from_cloud("all_carcasses", 2003)  |> 
  mutate(date = as.Date(date)) |> 
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         samples_collected = samp_type) |> 
  select(-c(comment)) |> 
  glimpse() 
```

## 2004
* recast `head_tag` to integer
```{r, message = FALSE}
feather_carcass_2004 <- read_from_cloud("all_carcasses", 2004) |> 
  mutate(head_tag = as.integer(head_tag),
         date = as.Date(date)) |> 
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         samples_collected = samp_type) |> 
  select(-c(comment)) |> 
  glimpse() 
```

## 2005
```{r, message = FALSE}
feather_carcass_2005 <- read_from_cloud("all_carcasses", 2005) |>
  mutate(date = as.Date(date)) |> 
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         hallprint1 = hall_print1,
         samples_collected = samp_type) |> 
  glimpse() 
```

## 2006
* recast `tag_col` to character
```{r, message = FALSE}
feather_carcass_2006 <- read_from_cloud("all_carcasses", 2006) |> 
  mutate(date = as.Date(date),
         tag_col = as.character(tag_col),
         recov_color = as.character(recov_color)) |> 
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         hallprint1 = hall_print1,
         samples_collected = samp_type) |> 
  glimpse() 
```

## 2007
* recast `tag_col` to character
```{r, message = FALSE}
feather_carcass_2007 <- read_from_cloud("all_carcasses", 2007) |> 
  mutate(date = as.Date(date),
         tag_col = as.character(tag_col),
         recov_color = as.character(recov_color)) |>
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         hallprint1 = hall_print1,
         samples_collected = samp_type) |> 
  glimpse() 
```

## 2008
* recast `tag_col` to character
```{r, message = FALSE}
feather_carcass_2008 <- read_from_cloud("all_carcasses", 2008) |> 
  mutate(date = as.Date(date),
         tag_col = as.character(tag_col),
         recov_color = as.character(recov_color)) |>
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         hallprint1 = hall_print1,
         samples_collected = samp_type) |>  
  glimpse() 
```

## 2009
* recast `tag_col` to character
```{r, message = FALSE}
feather_carcass_2009 <- read_from_cloud("all_carcasses", 2009) |> 
  mutate(date = as.Date(date),
         tag_col = as.character(tag_col),
         recov_color = as.character(recov_color)) |>
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         week = week_num,
         spawning_condition = spawn,
         hallprint1 = hall_print1,
         samples_collected = samp_type) |> 
  select(-c(comment)) |> 
  glimpse() 
```

## 2010
* recast `tag_col` to character
```{r, message = FALSE}
feather_carcass_2010 <- read_from_cloud("all_carcasses", 2010) |> 
  mutate(date = as.Date(date),
         tag_col = as.character(tag_col),
         recov_color = as.character(recov_color)) |>
  rename(fork_length = fl,
         adipose_fin_clipped = adipose_fin_clip,
         week = week_num,
         spawning_condition = spawn,
         hallprint1 = hall_print1,
         samples_collected = samp_type) |> 
  glimpse() 
```

## 2011
```{r, message = FALSE}
feather_carcass_2011 <- read_from_cloud("all_carcasses", 2011) |> 
  mutate(date = as.Date(date)) |> 
  rename(fork_length = fl,
         otolith_sample = otoliths,
         scale_sample = scales) |> 
  glimpse() 
```

## 2012
```{r, message = FALSE}
feather_carcass_2012 <- read_from_cloud("all_carcasses", 2012) |> 
  mutate(date = as.Date(date),
         hallprint1 = as.character(hallprint1),
         tag_col = tag_color) |> 
  rename(scale_sample = scales,
         otolith_sample = otoliths) |> 
  glimpse() 
```

## 2013
* rename `head_tag_number` to `head_tag`
```{r, message = FALSE}
feather_carcass_2013 <- read_from_cloud("all_carcasses", 2013) |> 
  mutate(date = as.Date(date),
         hallprint1 = as.character(hallprint1)) |> 
  rename(head_tag = head_tag_number,
         tag_col = tag_color,
         fork_length = fl,
         scale_sample = scales,
         otolith_sample = otoliths) |> 
  glimpse() 
```

## 2014
* rename `head_tag_number` to `head_tag`

```{r, message = FALSE}
feather_carcass_2014 <- read_from_cloud("all_carcasses", 2014) |> 
  mutate(date = as.Date(date),
         hallprint1 = as.character(hallprint1)) |> 
  rename(head_tag = head_tag_number,
         tag_col = tag_color,
         fork_length = fl,
         scale_sample = scales,
         otolith_sample = otoliths) |> 
  glimpse() 
```

## 2015
* rename `head_tag_number` to `head_tag`

```{r, message = FALSE}
feather_carcass_2015 <- read_from_cloud("all_carcasses", 2015) |> 
  mutate(date = as.Date(date)) |> 
  rename(head_tag = head_tag_number,
         tag_col = tag_color,
         fork_length = fl,
         scale_sample = scales,
         otolith_sample = otoliths) |> 
  select(-time) |> 
  glimpse() 
```

## 2016
* rename `head_tag_number` to `head_tag`

```{r, message = FALSE}
feather_carcass_2016 <- read_from_cloud("all_carcasses", 2016) |> 
  mutate(date = as.Date(date)) |> 
  rename(head_tag = head_tag_number,
         tag_col = tag_color,
         week = week_number,
         fork_length = fl,
         scale_sample = scales,
         otolith_sample = otoliths) |> 
  select(-time) |> 
  glimpse() 
```

## 2017
```{r, message = FALSE}
feather_carcass_2017 <- read_from_cloud("all_carcasses", 2017) |>
  mutate(date = as.Date(date),
         adipose_fin_clipped = case_when(ad_fin_clip_status %in% c("unknown", "not recorded") ~ "NA",
                                         ad_fin_clip_status == "yes" ~ "TRUE",
                                         ad_fin_clip_status == "no" ~ "FALSE",
                                         TRUE ~ ad_fin_clip_status),
         adipose_fin_clipped = as.logical(adipose_fin_clipped)) |>
  select(-ad_fin_clip_status) |>  
  rename(week = survey_week,
         sect = section_id,
         head_tag = head_nu,
         scale_sample = scale_nu,
         otolith_sample = otolith_nu,
         fork_length = fl_mm,
         spawning_condition = spawned) |> 
  glimpse() 
```

## 2018
```{r, message = FALSE}
feather_carcass_2018 <- read_from_cloud("all_carcasses", 2018) |>
  mutate(date = as.Date(date),
         adipose_fin_clipped = case_when(ad_fin_clip_status %in% c("unknown", "not recorded") ~ "NA",
                                         ad_fin_clip_status == "yes" ~ "TRUE",
                                         ad_fin_clip_status == "no" ~ "FALSE",
                                         TRUE ~ ad_fin_clip_status),
         adipose_fin_clipped = as.logical(adipose_fin_clipped)) |>
  select(-ad_fin_clip_status) |>  
  rename(week = survey_week,
         sect = section_id,
         head_tag = head_nu,
         scale_sample = scale_nu,
         otolith_sample = otolith_nu,
         fork_length = fl_mm,
         spawning_condition = spawned) |> 
  glimpse() 
```

## 2019
```{r, message = FALSE}
feather_carcass_2019 <- read_from_cloud("all_carcasses", 2019) |>
  mutate(date = as.Date(date),
         adipose_fin_clipped = case_when(ad_fin_clip_status %in% c("unknown", "not recorded") ~ "NA",
                                         ad_fin_clip_status == "yes" ~ "TRUE",
                                         ad_fin_clip_status == "no" ~ "FALSE",
                                         TRUE ~ ad_fin_clip_status),
         adipose_fin_clipped = as.logical(adipose_fin_clipped)) |>
  select(-ad_fin_clip_status) |>  
  rename(week = survey_week,
         sect = section_id,
         head_tag = head_nu,
         scale_sample = scale_nu,
         otolith_sample = otolith_nu,
         fork_length = fl_mm,
         spawning_condition = spawned) |> 
  glimpse() 
```

## 2020
```{r, message = FALSE}
feather_carcass_2020 <- read_from_cloud("all_carcasses", 2020) |>
  mutate(date = as.Date(date),
         adipose_fin_clipped = case_when(ad_fin_clip_status %in% c("unknown", "not recorded") ~ "NA",
                                         ad_fin_clip_status == "yes" ~ "TRUE",
                                         ad_fin_clip_status == "no" ~ "FALSE",
                                         TRUE ~ ad_fin_clip_status),
         adipose_fin_clipped = as.logical(adipose_fin_clipped)) |>
  select(-ad_fin_clip_status) |>  
  rename(week = survey_week,
         sect = section_id,
         head_tag = head_nu,
         scale_sample = scale_nu,
         otolith_sample = otolith_nu,
         fork_length = fl_mm,
         spawning_condition = spawned) |> 
  glimpse() 
```

## Combine all years
```{r}
feather_carcass_combined_raw <- 
  bind_rows(feather_carcass_2000,
            feather_carcass_2001,
            feather_carcass_2002,
            feather_carcass_2003,
            feather_carcass_2004,
            feather_carcass_2005,
            feather_carcass_2006,
            feather_carcass_2007,
            feather_carcass_2008,
            feather_carcass_2009,
            feather_carcass_2010,
            feather_carcass_2011,
            feather_carcass_2012,
            feather_carcass_2013,
            feather_carcass_2014,
            feather_carcass_2015,
            feather_carcass_2016,
            feather_carcass_2017,
            feather_carcass_2018,
            feather_carcass_2019,
            feather_carcass_2020) |> glimpse()
```
### Remove variables that are uninformative
* filter out `NA` dates
* create `year` column
* Carcasses in the `feather_carcass_combined` dataset are marked as `tagged`,
`recovered`, `chopped`, `chopped_recovered`/`recaptured, chopped`, `recaptured`/`recovered`, or `NA`. The differences in naming conventions come from different years and source tables.
* we standardize the `disposition` variable here

```{r}
feather_carcass_combined <- feather_carcass_combined_raw |> 
  select(-c(spawn, adipose_fin_clip, time, comment)) |> # spawn and adipose_fin_clip are from survey_meta_id == 70002 and are empty oncne we filter out NA dates 
  mutate(year = year(date),
         disposition = case_when(disposition == "recaptured, chopped" ~ "chopped",
                                 disposition == "chopped_recovered" ~ "chopped",
                                 disposition == "recaptured" ~ "recovered",
                                 TRUE ~ disposition),
         spawning_condition = case_when(spawning_condition == "yes" ~ "spawned",
                                        spawning_condition == "no" ~ "unspawned",
                                        spawning_condition %in% c("not recorded", "unknown") ~ NA_character_,
                                        TRUE ~ spawning_condition),
         samples_collected = case_when(samples_collected == "s&o" ~ "scales, otoliths",
                                       samples_collected == "h" ~ "head",
                                       samples_collected == "o" ~ "otoliths",
                                       samples_collected == "h&s" ~ "head, scales",
                                       samples_collected == "h&o" ~ "head, otoliths",
                                       samples_collected == "s" ~ "scales",
                                       samples_collected == "none" ~ NA_character_,
                                       samples_collected == "h&s&o" ~ "head, scales, otoliths",
                                       TRUE ~ samples_collected)) |> 
  filter(!is.na(date)) |> 
  rename(carcass_condition = condition,
         mark_recapture = disposition) |> 
  glimpse()
```

### Summarization of variables by year
```{r} 
# number of observations
feather_carcass_combined |> 
  count(year)

# print nas
feather_carcass_combined |> 
  group_by(year) |> 
  summarize_all(funs(sum(is.na(.)) / length(.)))

```
### Data Transformation

### Counts

The `feather_carcass_combined` table has carcass counts stored in the `count` variable.

## Explore Numeric Variables: {.tabset}

### Variable: `tag_id`, `sect`, `unit`, `fork_length`, `week`, `count`, `head_tag`, `flow`, `min`, `scale_scample`, `river_section`, `otolith_sample`, `year`

```{r}
feather_carcass_combined %>% 
  select_if(is.numeric) %>%
  colnames()
```

```{r}
numeric_cols <- feather_carcass_combined |>  
  select_if(is.numeric) |> 
  colnames()
sapply(feather_carcass_combined[,numeric_cols], summary)
```

### Carcass counts

```{r}
summary(feather_carcass_combined$count)
```


**NA and Unknown Values**
Provide a stat on NA or unknown values.

```{r}
round(sum(is.na(feather_carcass_combined$count))/nrow(feather_carcass_combined), 3) * 100

```
* `r round(sum(is.na(feather_carcass_combined$count))/nrow(feather_carcass_combined), 3) * 100` % of values in the `count` column are NA.

**Plotting chops over Period of Record**

```{r}
feather_carcass_combined |> 
  ggplot(aes(x = date, y = count)) +
  geom_col() +
  facet_wrap(~year, scales = "free") +
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Daily Carcass count from 2000 to 2020")
```

### Carcass variables: `sex`, `spawning_condition`, `rel_loc`, `tag_col`, `mark_recapture`, `source_table`, `lifestage`, `vis`, `weather`, `chan`, `recov_color`, `samples_collected`, `hallprint1`, `hall_print2`, `run`, `tag_color`, `hallprint_color`, `location`, `species`, `carcass_condition`


```{r}
feather_carcass_combined %>% 
  select_if(is.character) %>%
  colnames()
```

```{r}
character_cols <- feather_carcass_combined |>  
  select_if(is.character) |> 
  colnames()

sapply(feather_carcass_combined[, character_cols[! character_cols %in% c("hallprint", "hallprint1", "hall_print2")]], unique)
```

**Plotting total adult chops over period of record**

```{r}
total_adult_chops <- feather_carcass_combined |> 
  filter(mark_recapture != "tagged") |>
  filter(!is.na(date)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  select(fake_date, year, count, mark_recapture, sex) |> 
  glimpse()

total_adult_chops_summary <- total_adult_chops |> 
  group_by(fake_date, year, sex) |> 
  summarise(total_adult_chops = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_adult_chops_summary |> 
  ggplot(aes(x = fake_date, y = total_adult_chops, fill = sex)) + 
  geom_bar(stat = "identity", position = position_fill()) +
  labs(y = "Proportion", x = "Date") +
  theme_minimal() + 
  facet_wrap(~year) +
  scale_fill_manual(values = wes_palette("Moonrise2"))
```
This plot shows the sex proportion of adult chops (no `tagged` or `pre_carcass_survey` carcasses) for each day over the period of `r min(feather_carcass_combined$date, na.rm = T)` to `r max(feather_carcass_combined$date, na.rm = T)`. The plot is faceted by `year.` The data gaps that we see are a result of the data collection process in which the data was collected 2-4 days each week (with no specific period day of week) over the 4 month period.

**Plotting all carcasses over Period of Record**

```{r}
total_carcasses_summary <- feather_carcass_combined |> 
  filter(!is.na(date)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  group_by(fake_date, year, mark_recapture) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  glimpse()
```

```{r}
total_carcasses_summary |> 
  ggplot(aes(x = fake_date, y = total_carcasses, fill = mark_recapture)) + 
  facet_wrap(~year) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = c(wes_palette("Moonrise2"), "#eeb7d0"))
```
This plot shows the proportion of carcasses by `disposition` by `date`, faceted by `year`. `chopped_recovered` dominates 2000 and 2001, likely due to naming conventions. Across all years, the proportion of `chopped` carcasses tends increase as the sampling seasons goes on, and the proportion of `chopped` carcasses tends to decrease. A small proportion of carcasses across all years are `recovered`. 

*** Plotting all carcasses by spawning condition ***
```{r}
feather_carcass_combined |> 
  filter(!is.na(date)) |> 
  filter(!is.na(spawning_condition)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  group_by(fake_date, year, spawning_condition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  ggplot(aes(x = fake_date, y = total_carcasses, fill = spawning_condition)) + 
  facet_wrap(~year) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = c(wes_palette("Moonrise2"), "#eeb7d0"))
```
This plot shows all carcasses collected, filtered to those data that recorded spawning condition (many were NAs), by proportion of each spawning condition. 

*** Plotting all carcasses by run ***
```{r}
feather_carcass_combined |> 
  filter(!is.na(date)) |>
  filter(!is.na(run)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  group_by(fake_date, year, run) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  ggplot(aes(x = fake_date, y = total_carcasses, fill = run)) + 
  facet_wrap(~year) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = c(wes_palette("Moonrise2"), "#eeb7d0"))
```
This plot shows all carcasses collected, filtered to those data that recorded run (many were NAs), by proportion of carcasses marked as fall or spring. 

*** Plotting all carcasses by condition ***
```{r}
feather_carcass_combined |> 
  filter(!is.na(date)) |>
  filter(!is.na(carcass_condition)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  group_by(fake_date, year, carcass_condition) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  ggplot(aes(x = fake_date, y = total_carcasses, fill = carcass_condition)) + 
  facet_wrap(~year) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = c(wes_palette("Moonrise2"), "#eeb7d0"))
```
This plot shows all carcasses collected, filtered to those data that recorded condition (many were NAs), by proportion of carcasses marked as fresh, decayed, unknown, or not recorded.

*** Plotting all carcasses by adipose fin clip ***
```{r}
feather_carcass_combined |> 
  filter(!is.na(date)) |>
  filter(!is.na(adipose_fin_clipped)) |> 
  mutate(fake_date = paste0("1970-", month(date), "-", day(date)),
         fake_date = ymd(fake_date),
         year = year(date)) |> 
  group_by(fake_date, year, adipose_fin_clipped) |> 
  summarise(total_carcasses = sum(count, na.rm = T)) |> 
  ggplot(aes(x = fake_date, y = total_carcasses, fill = adipose_fin_clipped)) + 
  facet_wrap(~year) +
  geom_bar(stat = "identity", position = position_fill()) + 
  theme_minimal() + 
  labs(y = "Proportion", x = "Date") +
  scale_fill_manual(values = c(wes_palette("Moonrise2"), "#eeb7d0"))
```
This plot shows all carcasses collected, filtered to those data that recorded adipose fin clip (many were NAs), by proportion of carcasses marked as `TRUE` or `FALSE`.

### Individual-level sampling information

**Plotting fork length of each sex**
```{r}
summary(feather_carcass_combined$fork_length)
```
```{r}
feather_carcass_combined |> 
  filter(!is.na(sex)) |> 
  ggplot(aes(x = sex, y = fork_length)) +
  geom_boxplot() +
  facet_wrap(~year, scales = "free") +
  theme_minimal() + 
  labs(y = "fork length (mm)", x = "sex")
```
This plot shows `fork_length` plotted by `sex` and faceted by `year`, for all years where `fork_length` is not `NA`. 2002-2008 `fork_length` was not collected.

## Comments

* Identify variable used to connect recovered carcasses to tagged carcasses
* 
* `tag_id` connects to `recov_tag_id` for 2017-2020 tables

## Data Dictionaries

### All Carcasses
```{r}
percent_na <- feather_carcass_combined |> 
  summarise_all(list(name = ~sum(is.na(.))/length(.))) |> 
  pivot_longer(cols = everything())


all_carcasses_data_dictionary <- tibble(variables = colnames(feather_carcass_combined),
                          description = c("Tag ID",
                                          "Section",
                                          "Unit",
                                          "Fork length (mm)",
                                          "Sex",
                                          "Spawning condition",
                                          "Rel_loc, recorded in 2000 and 2001 TagData table",
                                          "Date of survey",
                                          "Tag Color",
                                          "Week number",
                                          "Mark Recapture Description (chopped, recaptured, tagged)",
                                          "Source table",
                                          "Count",
                                          "Lifestage",
                                          "Head tag number",
                                          "Flow",
                                          "Visibility",
                                          "Weather",
                                          "Channel",
                                          "Minute",
                                          "Color of recovered tag",
                                          "Samples collected",
                                          "Whether the adipose fin was clipped",
                                          "Hallprint tag number",
                                          "Hallprint 2, collected 2009-2010",
                                          "Scale sample",
                                          "Run",
                                          "River section",
                                          "Otolith sample, collected 2011-2014",
                                          "Tag color",
                                          "Color of hallprint tag number",
                                          "Location",
                                          "Species",
                                          "Carcass condition (fresh, decayed, etc.)",
                                          "Year"),
                          percent_na = round(percent_na$value*100))

kable(all_carcasses_data_dictionary)
```


** Upload to google bucket - TODO
```{r, include = FALSE}
f <- function(input, output) write_csv(input, file = output)

# all carcasses
gcs_upload(feather_carcass_combined,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_carcass.csv",
           predefinedAcl = "bucketLevel")
```