---
title: "pull temperature data for trap locations"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(CDECRetrieve)
library(dataRetrieval)
library(hms)
library(weathermetrics) # conversion of temperature units
library(googleCloudStorageR)
library(purrr)
library(zoo)
library(rnoaa)
library(data.table)
library(viridis)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
```

```{r, include = F}
# token for rnoaa is added to Renviron file
#NOAA_KEY <- Sys.getenv("NOAA_KEY")

# google cloud set up
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

# Pull in temperature data {.tabset}

## Battle Creek

### Recommendation

CDEC and USGS do not provide temperature data for Battle Creek. USGS mapper 
indicates as well there is no data available for the requested period.

We requested water temperature data from Battle Creek to fill this data
gap.

### Temperature data collected locally

```{r}
# get data from google cloud
gcs_get_object(object_name = "environmental/data-raw/battle_clear_RSTTEMPS_07052022.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk =  here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"),
               overwrite = TRUE)

ubc_temp <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 4)

battle_temp <- ubc_temp %>% 
  rename(date = DT,
         temp_degC = TEMP_C)

ggplot(battle_temp, aes(x=date, y = temp_degC)) +
  geom_line(color = "darkred") +
  labs(y = "Temperature (deg C)")
```

```{r}
write_rds(battle_temp, here::here("data-raw", "standard-format-data-prep","temp_data", "battle_temp.rds"))
```

## Butte Creek

### Recommendation

We looked at USGS and only mean data is available starting 2016. CDEC has finer resolution hourly data which is why we decided to use CDEC. We use USGS data establish threshold criteria to help filter out outliers in CDEC data.

### CDEC

One station:
* BCK (961-11-01 to present, with a data gap between ~1979-1999): Upper watershed, near Chico, far from confluence with Sacramento River


```{r}
BCK_CDEC <- cdec_query(station = "BCK", dur_code = "H", sensor_num = "25", start_date = "2000-01-01")

BCK_hourly_temps <- BCK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date, time, temp_degC)

# Date Ranges
tail(BCK_hourly_temps$date, n=1)
head(BCK_hourly_temps$date, n=1)

ggplot(BCK_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Buttle Creek", caption = "CDEC")
```

### Visualize CDEC with Outliers Removed

```{r}
# Plot without outliers
BCK_hourly_temps %>%
  filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Buttle Creek", caption = "CDEC")
```

### USGS

One station:
* Site number 11390000 Butte C NR Chico CA (TODO TIME PERIOD): TODO LOCATION DESCRIPTION

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

BCK_USGS <- readNWISdv(11390000, "00010", statCd = "00001") # parameter code 00010: water temperature
glimpse(BCK_USGS)
BCK_USGS_format <- BCK_USGS |> 
  select(Date, temperature=X_00010_00001)
ggplot(BCK_USGS_format, aes(Date, temperature)) +
  geom_line()
BCK_USGS <- readNWISdv(11390000, "00010", statCd = "00002")
BCK_USGS_format <- BCK_USGS |> 
  select(Date, temperature=X_00010_00002)
ggplot(BCK_USGS_format, aes(Date, temperature)) +
  geom_line()
```

### Using USGS min and max to filter out outliers from CDEC

```{r}
#Adding min max datasets
BCK_USGS_max_raw <- readNWISdv(11390000, "00010", statCd = "00001") # parameter code 00010: water temperature

BCK_USGS_max <- BCK_USGS_max_raw |>
  select(Date, temperature=X_00010_00001)
# plotmax <- ggplot(BCK_USGS_max, aes(Date, temperature)) +
#   geom_line()
# 
# 
BCK_USGS_min_raw <- readNWISdv(11390000, "00010", statCd = "00002")
BCK_USGS_min <- BCK_USGS_min_raw |>
  select(Date, temperature=X_00010_00002)
# plotmin <- ggplot(BCK_USGS_min, aes(Date, temperature)) +
#   geom_line()
# 
# #Combined min and max plot
combined_data <- bind_rows(
  mutate(BCK_USGS_max, HourlyTemperatures = "Daily Max"),
  mutate(BCK_USGS_min, HourlyTemperatures = "Daily Min")
)
# plot_minmax <- plotmin + plotmax
ggplot(combined_data, aes(Date, temperature, color = HourlyTemperatures)) +
  scale_color_viridis_d(begin = 1, end = 0) +
  geom_line() +
  #ggtitle("USGS hourly min and max hourly temperatures from Butte Creek near Chico") +
  labs(x = "",
       y = "Temperature (degrees Celcius)",
       color = "") +
  theme_minimal() +
  theme(legend.position = "bottom") 

# Caption for figure: Daily max and min temperatures (degrees Celcius) from 1960 to 2024 using data from USGS gage at Butte Creek near Chico (insert link to data source)."


```


```{r}
#Testing using USGS min and max as a threshold to remove outliers from CDEC
#Max

BCK_maxtemp_threshold <- max(BCK_USGS_max$temperature) |> 
  glimpse()

#Min 
BCK_mintemp_threshold <- min(BCK_USGS_min$temperature) |> 
  glimpse()

BCK_threshold <- BCK_USGS_max |> 
  rename(max = temperature) |> 
  full_join(BCK_USGS_min |> 
              rename(min = temperature)) |> 
  summarize(max = max(max, na.rm = T),
            min = min(min, na.rm = T))

#Add thresholds from min and max to CDEC data
BCK_temps_capped <- BCK_hourly_temps |> 
  filter(temp_degC >= BCK_threshold$min,
         temp_degC <= BCK_threshold$max)

ggplot(BCK_temps_capped, aes(date, temp_degC)) +
  geom_line()
  ggtitle("Filtered CDEC temperatures from Butte Creek near Chico, using USGS min/max threshold")

```


### Clean data and summarize as mean daily

```{r}
# Calculate the daily average temperature
BCK_daily_temps <- BCK_temps_capped %>% 
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC, na.rm = T),
            max_temp_degC = max(temp_degC, na.rm = T),
            min_temp_degC = min(temp_degC, na.rm = T)) 
```

### Save data

```{r}
# save final version of butte data
# from ashley: erin and i have started using the here package, especially in markdowns because it makes it easier for file path reading
write_rds(BCK_daily_temps, here::here("data-raw", "standard-format-data-prep","temp_data", "butte_temp.rds"))
```


## Clear Creek

### Recommendations

-   CDEC data has very high max - I think these are errors or code for
    something. CDEC data is more similar to Upper Clear Creek
-   Use the data from Clear over the CDEC data.

### CDEC

One station:
* IGO (1996-09-06 to present H; 1996-05-13 to 2000-01-01 E): TODO ADD DESCRIPTION OF LCOATION

```{r}
IGO_CDEC <- cdec_query(station = "IGO", dur_code = "H", sensor_num = "25", start_date = "2003-01-01")

IGO_hourly_temps <- IGO_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(IGO_hourly_temps$date, n=1)
head(IGO_hourly_temps$date, n=1)

ggplot(IGO_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y ="Temperature (deg C)", title = "Water Temperature at Clear Creek", caption = "CDEC")
```

#### Visualize CDEC Data with Outliers Removed

```{r}
# Plot without outliers
IGO_hourly_temps %>%
  filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Clear Creek", caption = "CDEC")
```

### USGS

USGS provides no temperature data for Clear Creek. This was confirmed
with mapper tool accessed from waterdata.usgs.gov.

```{r}
# Calculate the daily average temperature
IGO_daily_temps <- IGO_hourly_temps %>% 
  filter(temp_degC < 37, temp_degC > -18) %>%
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC),
            max_temp_degC = max(temp_degC)) 
```

### Temperature data collected locally

```{r}
clear_cdec <- read_rds(here::here("data-raw", "standard-format-data-prep","temp_data", "clear_temp.rds"))
ucc_temp_raw <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 2)

ucc_temp <- ucc_temp_raw %>% 
  rename(date = DT,
         temp_degC = TEMP_C)
lcc_temp_raw <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 3)
lcc_temp <- lcc_temp_raw %>% 
  rename(date = DT,
         temp_degC = TEMP_C)

# compare cdec to upper clear creek
ggplot(filter(clear_cdec, mean_temp_degC < 100), aes(x = date, y = mean_temp_degC)) +
  geom_line(color = "darkred") +
  geom_line(data = ucc_temp, aes(x = as.Date(date), y = temp_degC), color = "darkblue", alpha = 0.1)

# compare cded to lower clear creek
ggplot(filter(clear_cdec, mean_temp_degC < 100), aes(x = date, y = mean_temp_degC)) +
  geom_line(color = "darkred") +
  geom_line(data = lcc_temp, aes(x = as.Date(date), y = temp_degC), color = "darkblue", alpha = 0.1)
```

```{r}
write_rds(ucc_temp, here::here("data-raw", "standard-format-data-prep","temp_data", "upper_clear_temp.rds"))

write_rds(lcc_temp, here::here("data-raw", "standard-format-data-prep","temp_data", "lower_clear_temp.rds"))
```

## Deer Creek

### CDEC
TODO MOVE TO FORMAT LIKE OTHERS
CDEC Temperature Data for Deer Creek ranges from 1998-10-01 to
2022-06-16. Hourly data does range from 1998-present, however there is another record under "event" duration that ranges from 2014-present. There is another gauge on Deer Creek (DVD) called Deer Creek Below Standford Vina Dam. Temp data ranges from 1997-present. It is closer to confluence of the Sac River, below Vina Dam.

```{r}
DCV_CDEC <- cdec_query(station = "DCV", dur_code = "H", sensor_num = "25", start_date = "1995-01-01")

DCV_hourly_temps <- DCV_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(DCV_hourly_temps$date, n = 1)
head(DCV_hourly_temps$date, n = 1)

# Plot
ggplot(DCV_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Water Temperature at Deer Creek", caption = " CDEC")
```

### Visualize CDEC Data with Outliers Removed

```{r}
# Plot without outliers
DCV_hourly_temps %>%
    filter(temp_degC < 37, temp_degC > 0) %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_degC), color ="darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "CDEC Water Temperature at Buttle Creek", caption = "CDEC")
```

### USGS

USGS mean temperature data for deer creek ranges from 2016-10-01 to 2022-05-25. Min and max temp with a wider range from 1998 to present. This gage is also used in DSMtemperature

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

#Mean temperature USGS
DCV_USGS <- readNWISdv(11383500, "00010") # parameter code 00010: water temperature

# Format to make tidier
DCV_daily_temps <- DCV_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)

# Date Ranges
tail(DCV_daily_temps$date,n=1)
head(DCV_daily_temps$date,n=1)

# Plot
ggplot(DCV_daily_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "Water Temperature at Deer Creek", caption = "USGS")

#Pulling Max and Min temperature USGS
#max
DCV_USGS <- readNWISdv(11383500, "00010", statCd = "00001") # parameter code 00010: water temperature
DCV_USGS_max <- DCV_USGS |>
  select(Date, temperature=X_00010_00001)
# plotmax <- ggplot(DCV_USGSS_max, aes(Date, temperature)) +
#   geom_line()

# min
DCV_USGS <- readNWISdv(11383500, "00010", statCd = "00002")
DCV_USGS_min <- DCV_USGS |>
  select(Date, temperature=X_00010_00002)
# plotmin <- ggplot(DCV_USGS_min, aes(Date, temperature)) +
#   geom_line()
# 
# #Combined min and max plot
combined_data <- bind_rows(
  mutate(DCV_USGS_max, HourlyTemperatures = "Max Temp"),
  mutate(DCV_USGS_min, HourlyTemperatures = "Min Temp")
)
# plot_minmax <- plotmin + plotmax
ggplot(combined_data, aes(Date, temperature, color = HourlyTemperatures)) +
  geom_col() +
  ggtitle("USGS hourly min and max hourly temperatures from Deer Creek near Vina")

```

#### Mean Daily USGS vs Daily CDEC data

The plot below shows the daily average to allow for comparison between
CDEC and USGS datasets. This helps it become more apparent that we
should use CDEC data for deer creek due to the availability of the data.

```{r}
DCV_combined = DCV_hourly_temps %>% 
  full_join(DCV_daily_temps, by ="date") %>% 
  rename("temp_degC.CDEC"="temp_degC.x", "temp_degC.USGS"="temp_degC.y")

DCV_combined %>% 
  group_by(date = date(date)) %>% 
  group_by(date) %>% 
  summarise(
    temp_CDEC = median(temp_degC.CDEC, na.rm = TRUE),
    temp_USGS = max(temp_degC.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_CDEC), color = "darkred") +
  geom_line(aes(y = temp_USGS), color = "steelblue", linetype = "twodash") +
  labs(y = "Temperature (deg C)", title = "Water Temperature at Deer Creek")
```

### Recommendation

Use CDEC data because USGS data has a limited range for the mean. Use USGS min and max data to establish threshold criteria that can be used to filter out any outliers in the CDEC data.

### Using USGS min and max to filter out outliers from CDEC

```{r}
#USGS min and max
#Max
DCV_max_threshold <- max(DCV_USGS_max$temperature) |> 
  glimpse()

# TODO for ashley go through summarization with Badhia
#Trying another way to filter max of max
# DCV_USGS_max |>
#   summarise(highest_temperature = max(temperature)) |>
#   glimpse()

#Min 
DCV_min_threshold <- min(DCV_USGS_min$temperature) |> 
  glimpse()

#Add thresholds from min and max to CDEC data
DCV_temps_capped <- DCV_hourly_temps |> 
  filter(temp_degC >= 0.4,
         temp_degC <= 28.5)

ggplot(DCV_temps_capped, aes(date, temp_degC)) +
  geom_line() +
  ggtitle("Filtered CDEC temperatures from Deer Creek, using USGS min/max threshold")

#note that 25 nows were removed, assuming that is because they were outside threshold
```

### Clean data and summarize as mean daily

```{r}
# 1. Calculate the daily average temperature
DCV_daily_temps <- DCV_temps_capped %>% 
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC, na.rm = T),
            max_temp_degC = max(temp_degC, na.rm = T),
            min_temp_degC = min(temp_degC, na.rm = T)) 
```

```{r}
write_rds(DCV_daily_temps, here::here("data-raw", "standard-format-data-prep","temp_data", "deer_temp.rds"))
```

## Feather River

### Recommendation

Both CDEC sites will be used because of the more recent period of time that data are available. If needed, USGS could be appended for a longer time series. Currently only CDEC data will be added to the dataset. GRL will represent the High Flow Channel (HFC) and FRA will represent the Low Flow Channel (LFC).

### CDEC

Two stations:
* GRL (2003-03-05 to 2007-06-01 H; 2020-01-04 to present E): located after Thermalito Afterbay
* FRA (2002-01-01 to present): located between Lake Oroville and Thermalito Afterbay

```{r}
GRL_CDEC <- cdec_query(station = "GRL", dur_code = "H", sensor_num = "25", start_date = "2003-03-05", end_date = "2007-06-01")

GRL_hourly_temps <- GRL_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(GRL_hourly_temps$date, n = 1)
head(GRL_hourly_temps$date, n = 1)

# Plot GRL
ggplot(GRL_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y="Temperature (deg C)", title = "Temperature at Feather River", caption = "CDEC")

#Adding "FRA" dataset
FRA_CDEC <- cdec_query(station = "FRA", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")

FRA_hourly_temps <- FRA_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Plot FRA
ggplot(FRA_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y="Temperature (deg C)", title = "Temperature at FRA Feather River", caption = "CDEC")
```

### Visualize CDEC Data with Outliers Removed

```{r}
# Plot without outliers
GRL_hourly_temps %>%
   filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Feather River", caption = "CDEC")
```

### USGS
Three stations:
* Thermalito Afterbay Rel to Feather R NR Oroville 11406920 (1968-1992): located right where Thermalito rejoins LFC. Time period is old but could be helpful to understand HFC temperature.
* Feather R A Yuba City CA 11407700 (1964-1976): not using this one because it is too low on the stream)
* Feather R A Oroville CA 11407000 (1958-1992): located right below diversion from Oroville into LFC and Thermalito Afterbay. Time period is old but could be helpful to understand the LFC temperature.

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

#Adding Oroville station
# ORO_USGS <- readNWISdv(11406920, "00010", statCd = "00003") # parameter code 00010: water temperature
# 
# # Format to make tidier
# ORO_daily_meantemps <- ORO_USGS %>%
#   select(Date, temp_degC =  X_00010_00003) %>%
#   as_tibble() %>% 
#   rename(date = Date)
# no data for daily statistics mean temp on default statCd = 00003, so using min and max

#max temp
ORO_USGS_max_raw <- readNWISdv(11406920, "00010", statCd = "00001") # parameter code 00010: water temperature

# Format to make tidier
ORO_daily_maxtemps <- ORO_USGS_max_raw %>%
  select(Date, temp_degC =  X_00010_00001) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(ORO_daily_maxtemps)

ggplot(ORO_daily_maxtemps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Max Temperature at Feather Oroville", caption = "USGS")

#min temp
ORO_USGS_min_raw <- readNWISdv(11406920, "00010", statCd = "00002") # parameter code 00010: water temperature

# Format to make tidier
ORO_daily_mintemps <- ORO_USGS_min_raw %>%
  select(Date, temp_degC =  X_00010_00002) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(ORO_daily_mintemps)

ggplot(ORO_daily_mintemps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Min Temperature at Feather Oroville", caption = "USGS")

```

```{r}
#Plot of min and max 
combined_data <- bind_rows(
  mutate(ORO_daily_maxtemps, HourlyTemperatures = "Max Temp"),
  mutate(ORO_daily_mintemps, HourlyTemperatures = "Min Temp")
)
# plot_minmax <- plotmin + plotmax
ggplot(combined_data, aes(date, temp_degC, color = HourlyTemperatures)) +
  geom_col() +
  ggtitle("USGS hourly min and max hourly temperatures Feather River - Oroville")

```



```{r}
# TODO save all the data from the gages that we want to use
# write_rds(GRL_hourly_temps, "../../data-raw/standard-format-data-prep/temp_data/feather_temp.rds")
```

## Mill Creek

### Recommendation
Use CDEC data due to the longer time period and resolution of data. USGS can be used as a threshold criteria if needed. 
* Note that Mill Creek has multiple temperature loggers to represent summer holding habitat. We will be acquiring this data and using in addition to the CDEC data. 

### CDEC
One station:
* MLM (1998-09-30 to 2022-05-27): Upper watershed, near Los Molinos
* MCH (1998-present H; 2005-present E): Lower watershed, near confluence with Sacramento R, Below HWY 99

```{r}

# cdec_datasets("MLM") 

MLM_CDEC <- cdec_query(station = "MLM", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")


MLM_hourly_temps <- MLM_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(MLM_hourly_temps$date, n = 1)
head(MLM_hourly_temps$date, n = 1)

# Plot DCV
ggplot(MLM_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Mill Creek", caption = "CDEC")
```

#### Visualize CDEC Data with Outliers Removed

```{r}
# Plot without outliers
MLM_hourly_temps %>%
    filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "CDEC Water Temperature at Mill Creek", caption = "CDEC")
```

### USGS
One station:
* ADD STATION NAME (2016-10-01 to 2022-05-25 Daily Mean; 1998-2023 Min/Max): ADD LOCATION DESCRIPTION. This gage station data was used on DSMtemperature.

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

MLM_USGS <- readNWISdv(11381500, "00010") # parameter code 00010: water temperature temperature

# Format to make tidier
MLM_daily_temps <- MLM_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  as_tibble() %>% 
  rename(date = Date)

# Date Ranges
tail(MLM_daily_temps$date, n = 1)
head(MLM_daily_temps$date, n = 1)

# Plot DCV
ggplot(MLM_daily_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Temperature at Mill Creek", caption = "USGS")
```

#### Mean Daily USGS vs Daily CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
MLM_combined = MLM_hourly_temps %>% 
  full_join(MLM_daily_temps, by = "date") %>% 
  rename("temp_degC.CDEC" = "temp_degC.x", "temp_degC.USGS" = "temp_degC.y")

MLM_combined %>% 
  group_by(date = date(date)) %>% 
  group_by(date) %>% 
  summarise(
    temp_CDEC = median(temp_degC.CDEC, na.rm = TRUE),
    temp_USGS = max(temp_degC.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_CDEC), color = "darkred") +
  geom_line(aes(y = temp_USGS), color = "steelblue", linetype = "twodash") +
  labs(y = "Temperature (deg C)", title = "Water Temperature at Mill Creek")

```



```{r}

# 1. Calculate the daily average temperature
MLM_daily_temps <- MLM_hourly_temps %>% 
  filter(!is.na(temp_degC),
         temp_degC < 37, temp_degC > -18) %>% 
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC),
            max_temp_degC = max(temp_degC)) 
```

```{r}
# save final version of butte data
# from ashley: erin and i have started using the here package, especially in markdowns because it makes it easier for file path reading
write_rds(MLM_daily_temps, here::here("data-raw", "standard-format-data-prep","temp_data", "mill_temp.rds"))
```

## Yuba River

### CDEC

Two stations:
* YR7 (2019-09-04 to present): Above highway 70 (TODO include more location description)
* YRS TODO ADD TIME PERIOD AND LOCATION DESCRIPTION

```{r}
YRS_CDEC <- cdec_query(station = "YRS", dur_code = "H", sensor_num = "25", start_date = "1995-01-01")
#this data seems weird

YRS_hourly_temps <- YRS_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = parameter_value) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(YRS_hourly_temps$date, n = 1)
head(YRS_hourly_temps$date, n = 1)

# Plot
ggplot(YRS_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Water Temperature at Yuba River", caption = " CDEC")

#Adding another gage
YR7_CDEC <- cdec_query(station = "YR7", dur_code = "E", sensor_num = "146", start_date = "1995-01-01")
#note that this station only has event duration, not hourly
YR7_hourly_temps <- YR7_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = parameter_value) %>%
  select(date,time,temp_degC)

# Plot
ggplot(YR7_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Water Temperature at YR7 Yuba River", caption = " CDEC")

```

### Visualize CDEC Data with Outliers Removed
```{r}
# Plot without outliers
YRS_hourly_temps %>%
    filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "CDEC Water Temperature at Yuba River", caption = "CDEC")
```

## USGS
* daily max and min Temp data from 1964-2003, with some data gaps, for
    station number 11421000 Yuba R NR Marysville (this station data was pulled for va-habitat eda, too)
* daily max and min Temp data from 1972-2005, Station 11418000 Yuba R
    BL Englebright Dam NR Smartsville has
```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

YBAMR_USGS <- readNWISdv(11421000, "00010") # parameter code 00010: water temperature
glimpse(YBAMR_USGS) #no data on 00003 (median), so using min and max

#max
YBAMR_USGS <- readNWISdv(11421000, "00010", statCd = "00001") # parameter code 00010: water temperature

# Format to make tidier
YBAMR_daily_maxtemps <- YBAMR_USGS %>%
  select(Date, temp_degC =  X_00010_00001) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(YBAMR_daily_maxtemps)

#min
YBAMR_USGS <- readNWISdv(11421000, "00010", statCd = "00002") # parameter code 00010: water temperature

# Format to make tidier
YBAMR_daily_mintemps <- YBAMR_USGS %>%
  select(Date, temp_degC =  X_00010_00002) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(YBAMR_daily_mintemps)


#Plot USGS min and max together
combined_data <- bind_rows(
  mutate(YBAMR_daily_maxtemps, HourlyTemperatures = "Max Temp"),
  mutate(YBAMR_daily_mintemps, HourlyTemperatures = "Min Temp")
)
# plot_minmax <- plotmin + plotmax
ggplot(combined_data, aes(date, temp_degC, color = HourlyTemperatures)) +
  geom_col() +
  ggtitle("USGS hourly min and max hourly temperatures Yuba River - Marysville")

```

### Plot without outliers using USGS max/min

```{r}
#Max
YBAMR_maxtemps_threshold <- max(YBAMR_daily_maxtemps$temp_degC) |> 
  glimpse()

#Trying another way to filter max of max
# YBAMR_daily_maxtemps |>
#   summarise(highest_temperature = max(temp_degC)) |>
#   glimpse()

#Min 
YBAMR_mintemps_threshold <- min(YBAMR_daily_mintemps$temp_degC) |> 
  glimpse()
#Trying another way to filter min of min
# YBAMR_daily_mintemps |>
#   summarise(lowest_temperature = min(temp_degC)) |>
#   glimpse()

#Add thresholds from min and max to CDEC YR7 data
YBAMR_temps_capped <- YR7_hourly_temps |> 
  filter(temp_degC >= 4.5,
         temp_degC <= 29.5)

ggplot(YBAMR_temps_capped, aes(date, temp_degC)) +
  geom_line() +
  ggtitle("Filtered CDEC temperatures from Yuba River (YR7), using USGS min/max threshold")

```

## Sacramento River - Knights Landing

```{r}
# cdec_datasets("WLK")

WLK_CDEC <- cdec_query(station = "WLK", dur_code = "E", sensor_num = "25", start_date = "1995-01-01")

WLK_hourly_temps <- WLK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(WLK_hourly_temps$date, n = 1)
head(WLK_hourly_temps$date, n = 1)

# Plot WLK
ggplot(WLK_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "CDEC")
```

CDEC provides temperature data for Sacramento River - Knights Landing
from 2012-11-15 to 2022-06-16

### CDEC Outliers Removed

```{r}
# Plot without outliers
WLK_hourly_temps %>%
  filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "CDEC")
```

### USGS

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

WLK_USGS <- readNWISdv(11390500, "00010") # parameter code 00010: water temperature temperature

# Format to make tidier
WLK_daily_temps <- WLK_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  #filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date) %>% 
  #filter(year(date) > 2000) %>%
  glimpse

# Date Ranges
tail(WLK_daily_temps$date, n = 1)
head(WLK_daily_temps$date, n = 1)

# Plot
ggplot(WLK_daily_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
```

Temperature data for Sacramento River - Knights landing has data from
1995-01-01 to 2022-06-14, but we should explore further the signifcant
data gap and identify that period.

### Data before Missing Data Gap

```{r}
### Identify date data starts to be missing
WLK_daily_temps %>%
  filter(date < "2004-01-01") %>%
  arrange(desc(date))
  # 1998-09-30

# data before missing data gap
WLK_daily_temps %>%
  filter(date < "1999-10-01") %>% 
  arrange(desc(date)) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
```

### Data after missing data gap

```{r}
### Identify date data starts to be missing
WLK_daily_temps %>%
  filter(date > "2015-01-01") %>%
  arrange(date)
  # 2016-10-01

# data after missing data gap
WLK_daily_temps %>%
  filter(date >= "2016-10-01") %>% 
  arrange(desc(date)) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y ="Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
#
```

Temperature data for Sacramento River - Knights landing is missing
significant data from the period of 1998-09-30 to 2016-09-31.

#### Mean Daily USGS vs Daily CDEC data

```{r}
WLK_combined = WLK_hourly_temps %>% 
  full_join(WLK_daily_temps, by = "date") %>% 
  rename("temp_degC.CDEC" = "temp_degC.x", "temp_degC.USGS" = "temp_degC.y")

WLK_combined %>% 
  group_by(date = date(date)) %>% 
  summarise(
    temp_CDEC = median(temp_degC.CDEC, na.rm = TRUE),
    temp_USGS = max(temp_degC.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_CDEC), color = "darkred") +
  geom_line(aes(y = temp_USGS), color = "blue", linetype ="twodash") +
  labs(y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
```
