---
title: "pull temperature data for trap locations"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(tidyverse)
library(lubridate)
library(CDECRetrieve)
library(dataRetrieval)
library(hms)
library(weathermetrics) # conversion of temperature units
library(googleCloudStorageR)
library(purrr)
library(zoo)
library(rnoaa)
library(data.table)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F}
# token for rnoaa is added to Renviron file
#NOAA_KEY <- Sys.getenv("NOAA_KEY")

# google cloud set up
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

# Pull in temperature data {.tabset}

## Battle Creek

```{r}
# Check if CDEC data provides temp data for respective location
# cdec_datasets("BAT") 
# CDEC has no temperature data available
```

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

BAT_USGS <- readNWISdv(11376550, "00010") # parameter code 00010: water temperature

# from ashley: don't need to save this if there is no data.
# write.csv(BAT_USGS,"BAT_USGS.csv", row.names = FALSE)

# Read created CSV
# BAT_USGS <- read_csv("BAT_USGS.csv")
# BAT_USGS
```

CDEC and USGS do not provide temperature data for Battle Creek.
Temperature data is collected by sensor 25, but CDEC data is not
available for this sensor. USGS mapper indicates as well there is no
data available for the requested period.

We requested water temperature data from Battle Creek to fill this data
gap.

### Temperature data collected locally

```{r}
# get data from google cloud
gcs_get_object(object_name = "environmental/data-raw/battle_clear_RSTTEMPS_07052022.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk =  here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"),
               overwrite = TRUE)

ubc_temp <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 4)

battle_temp <- ubc_temp %>% 
  rename(date = DT,
         temp_degC = TEMP_C)

ggplot(battle_temp, aes(x=date, y = temp_degC)) +
  geom_line(color = "darkred") +
  labs(y = "Temperature (deg C)")
```

### Save data

```{r}
write_rds(battle_temp, here::here("data-raw", "standard-format-data-prep","temp_data", "battle_temp.rds"))
```

## Butte Creek

### CDEC

```{r}
# Check if CDEC data provides temp data for respective location
# cdec_datasets("BCK")

# Since this location has sensor that collects temp data, we can make request
BCK_CDEC <- cdec_query(station = "BCK", dur_code = "H", sensor_num = "25", start_date = "1995-01-01")

BCK_hourly_temps <- BCK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date, time, temp_degC)

# Date Ranges
tail(BCK_hourly_temps$date, n=1)
head(BCK_hourly_temps$date, n=1)

ggplot(BCK_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Buttle Creek", caption = "CDEC")
```

CDEC data provides temperature data for Butte Creek. It ranges from
1998-09-16 to 2022-06-16

### CDEC Outliers Removed

```{r}
# Plot without outliers
# TODO review literature on temps on Butte Creek
BCK_hourly_temps %>%
  filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Buttle Creek", caption = "CDEC")
```

### USGS

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

BCK_USGS <- readNWISdv(11390000, "00010", statCd = "00001") # parameter code 00010: water temperature
glimpse(BCK_USGS)
BCK_USGS_format <- BCK_USGS |> 
  select(Date, temperature=X_00010_00001)
ggplot(BCK_USGS_format, aes(Date, temperature)) +
  geom_line()
BCK_USGS <- readNWISdv(11390000, "00010", statCd = "00002")
BCK_USGS_format <- BCK_USGS |> 
  select(Date, temperature=X_00010_00002)
ggplot(BCK_USGS_format, aes(Date, temperature)) +
  geom_line()
# from ashley: don't need to save if no data
#write.csv(BCK_USGS,"BCK_USGS.csv", row.names = FALSE)

# no data available from USGS
```
from Badhia 
on USGS data:
* Site number 11390000 Butte C NR Chico CA contains water data
* Data ranges from 1961-11-01 to present, with a data gap between ~1979-1999 
on CDEC
* Hourly data does range from 1998-present, however there is another record under "event" duration that ranges from 2017-present. Unsure on the difference between datasets 
* Gauge is operating and "present data" is available

TODO We looked at USGS and only mean data is available starting 2016, CDEC has finer resolution hourly data, that is why we decided to use CDEC. 
We are using USGS data to help filter out outliers.

### Using USGS min and max to filter out outliers from CDEC
```{r}
#USGS
BCK_USGS <- readNWISdv(11390000, "00010") # parameter code 00010: water temperature, median 

# Format to make tidier
BCK_daily_temps <- BCK_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)

ggplot(BCK_daily_temps, aes(date, temp_degC)) +
  geom_line()

temp_rangeUSGS <- BCK_daily_temps |> 
  summarise(min_tempUSGS = min(temp_degC),
            max_tempUSGS = max(temp_degC))
glimpse(temp_rangeUSGS)

#TODO also graph adding min max datasets
BCK_USGS <- readNWISdv(11390000, "00010", statCd = "00001") # parameter code 00010: water temperature

# BCK_USGS_format <- BCK_USGS |> 
#   select(Date, temperature=X_00010_00001)
# plotmax <- ggplot(BCK_USGS_format, aes(Date, temperature)) +
#   geom_line()
# 
# 
# BCK_USGS <- readNWISdv(11390000, "00010", statCd = "00002")
# BCK_USGS_format <- BCK_USGS |> 
#   select(Date, temperature=X_00010_00002)
# plotmin <- ggplot(BCK_USGS_format, aes(Date, temperature)) +
#   geom_line()
# 
# #Combined min and max plot


combined_data <- bind_rows(
  mutate(BCK_USGS_format, group = "Stat Code: 00001"),
  mutate(BCK_USGS_format, group = "Stat Code: 00002")
)
# plot_minmax <- plotmin + plotmax
ggplot(combined_data, aes(Date, temperature, color = group)) +
  geom_line() +
  ggtitle("Combined Plot for Stat Code: 00001 and Stat Code: 00002")

#CDEC
BCK_CDEC <- cdec_query(station = "BCK", dur_code = "H", sensor_num = "25", start_date = "1995-01-01") #added start date otherwise defaults to most recent years only


BCK_hourly_temps <- BCK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date, time, temp_degC)

BCK_temps_filtered <- BCK_hourly_temps |> 
  filter(temp_degC >= 3.7,
         temp_degC <= 24.6)

ggplot(BCK_temps_filtered, aes(date, temp_degC)) +
  geom_line()


```


### Clean data and summarize as mean daily

```{r}

# 1. Calculate the daily average temperature
BCK_daily_temps <- BCK_hourly_temps %>% 
  filter(temp_degC < 37, temp_degC > -18) %>%
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC, na.rm = T),
            max_temp_degC = max(temp_degC, na.rm = T),
            min_temp_degC = min(temp_degC, na.rm = T)) 

# 2. Check to see if applying rolling mean would help with outliers.
butte_format <- transform(BCK_daily_temps, rollmean = rollapplyr(mean_temp_degC, 3, mean, partial = T)) %>% 
  mutate(mean_temp_degC_clean = ifelse(mean_temp_degC > 70, rollmean, mean_temp_degC)) 

butte_format %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y = mean_temp_degC_clean), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Buttle Creek", caption = "CDEC")
# doesn't really seem to help much so for transparency don't apply any transformation
filter(butte_format, mean_temp_degC > 70)
```

### Save data

```{r}
# save final version of butte data
# from ashley: erin and i have started using the here package, especially in markdowns because it makes it easier for file path reading
write_rds(BCK_daily_temps, here::here("data-raw", "standard-format-data-prep","temp_data", "butte_temp.rds"))
```

USGS provides no temperature data for Butte Creek. This was confirmed
with mapper tool accessed from waterdata.usgs.gov.

## Clear Creek

### CDEC

```{r}
# Check if CDEC data provides temp data for respective location
# cdec_datasets("IGO")

# Since this location has sensor that collects temp data, we can make request
IGO_CDEC <- cdec_query(station = "IGO", dur_code = "H", sensor_num = "25", start_date = "2003-01-01")

IGO_hourly_temps <- IGO_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(IGO_hourly_temps$date, n=1)
head(IGO_hourly_temps$date, n=1)

ggplot(IGO_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y ="Temperature (deg C)", title = "Water Temperature at Clear Creek", caption = "CDEC")
```

CDEC temeprature data for Clear Cleek ranges from 2002-12-31 to
2022-05-27.

### CDEC Outliers Removed

```{r}
# Plot without outliers
IGO_hourly_temps %>%
  filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Clear Creek", caption = "CDEC")
```

### USGS

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

#IGO_USGS <- readNWISdv(11372000, "00010") # parameter code 00010: water temperature
#write.csv(IGO_USGS,"IGO_USGS.csv", row.names = FALSE)

# Read created CSV
# IGO_USGS <- read_csv("IGO_USGS.csv")
# IGO_USGS
# no data available from USGS
```

USGS provides no temperature data for Clear Creek. This was confirmed
with mapper tool accessed from waterdata.usgs.gov.

Notes from Badhia:
CDEC:
* Temp seems to range from 1996-09-06 to present (for hourly) and 1996-05-13 to 200-01-01 (for event), not 2002-12-31 to 2022-05-27.
USGS:
* No temp

### Clean data and summarize as mean daily

```{r}

# 1. Calculate the daily average temperature
IGO_daily_temps <- IGO_hourly_temps %>% 
  filter(temp_degC < 37, temp_degC > -18) %>%
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC),
            max_temp_degC = max(temp_degC)) 

```

```{r}
# save final version of butte data
# from ashley: erin and i have started using the here package, especially in markdowns because it makes it easier for file path reading
write_rds(IGO_daily_temps, here::here("data-raw", "standard-format-data-prep","temp_data", "clear_temp.rds"))
```

### Temperature data collected locally

-   We received temperature data from clear and battle.
-   CDEC data has very high max - I think these are errors or code for
    something.
-   Use the data from Battle and Clear over the CDEC data.
-   CDEC data is more similar to Upper Clear Creek

```{r}
clear_cdec <- read_rds(here::here("data-raw", "standard-format-data-prep","temp_data", "clear_temp.rds"))
ucc_temp_raw <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 2)

ucc_temp <- ucc_temp_raw %>% 
  rename(date = DT,
         temp_degC = TEMP_C)
lcc_temp_raw <- readxl::read_excel(here::here("data-raw", "standard-format-data-prep", "temp_data", "battle_clear_temp.xlsx"), sheet = 3)
lcc_temp <- lcc_temp_raw %>% 
  rename(date = DT,
         temp_degC = TEMP_C)

# compare cdec to upper clear creek
ggplot(filter(clear_cdec, mean_temp_degC < 100), aes(x = date, y = mean_temp_degC)) +
  geom_line(color = "darkred") +
  geom_line(data = ucc_temp, aes(x = as.Date(date), y = temp_degC), color = "darkblue", alpha = 0.1)

# compare cded to lower clear creek
ggplot(filter(clear_cdec, mean_temp_degC < 100), aes(x = date, y = mean_temp_degC)) +
  geom_line(color = "darkred") +
  geom_line(data = lcc_temp, aes(x = as.Date(date), y = temp_degC), color = "darkblue", alpha = 0.1)
```

```{r}
write_rds(ucc_temp, here::here("data-raw", "standard-format-data-prep","temp_data", "upper_clear_temp.rds"))

write_rds(lcc_temp, here::here("data-raw", "standard-format-data-prep","temp_data", "lower_clear_temp.rds"))
```

## Deer Creek

### CDEC

```{r}

# cdec_datasets("DCV")

DCV_CDEC <- cdec_query(station = "DCV", dur_code = "H", sensor_num = "25", start_date = "1995-01-01")

DCV_hourly_temps <- DCV_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(DCV_hourly_temps$date, n = 1)
head(DCV_hourly_temps$date, n = 1)

# Plot
ggplot(DCV_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Water Temperature at Deer Creek", caption = " CDEC")
```

CDEC Temperature Data for Deer Creek ranges from 1998-10-01 to
2022-06-16.

### CDEC Outliers Removed

```{r}
# Plot without outliers
DCV_hourly_temps %>%
    filter(temp_degC < 37, temp_degC > 0) %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_degC), color ="darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "CDEC Water Temperature at Buttle Creek", caption = "CDEC")
```

### USGS

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

DCV_USGS <- readNWISdv(11383500, "00010") # parameter code 00010: water temperature

# Format to make tidier
DCV_daily_temps <- DCV_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)

# Date Ranges
tail(DCV_daily_temps$date,n=1)
head(DCV_daily_temps$date,n=1)

# Plot
ggplot(DCV_daily_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "Water Temperature at Deer Creek", caption = "USGS")
```

USGS temperature data for deer creek ranges from 2016-10-01 to
2022-05-25.

#### Mean Daily USGS vs Daily CDEC data

```{r}
DCV_combined = DCV_hourly_temps %>% 
  full_join(DCV_daily_temps, by ="date") %>% 
  rename("temp_degC.CDEC"="temp_degC.x", "temp_degC.USGS"="temp_degC.y")

DCV_combined %>% 
  group_by(date = date(date)) %>% 
  group_by(date) %>% 
  summarise(
    temp_CDEC = median(temp_degC.CDEC, na.rm = TRUE),
    temp_USGS = max(temp_degC.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_CDEC), color = "darkred") +
  geom_line(aes(y = temp_USGS), color = "steelblue", linetype = "twodash") +
  labs(y = "Temperature (deg C)", title = "Water Temperature at Deer Creek")
```

The plot below shows the daily average to allow for comparison between
CDEC and USGS datasets. This helps it become more apparent that we
should use CDEC data for deer creek due to the availability of the data.

Notes from Badhia
USGS
* Temp data range from Shinny app 2007-10-01 to present
* 11383500 station has min and mx temp with a wider range than 2016-2022 (1998-present)
* This same gauge data was used on DSMtemperature 
CDEC
* Hourly data does range from 1998-present, however there is another record under "event" duration that ranges from 2014-present. Unsure on the difference between datasets
* There is another gauge on Deer Creek (DVD) called Deer Creek Below Standford Vina Dam. Temp data ranges from 1997-present. It is closer to confluence of the Sac River, below Vina Dam.

TODO use USGS min and max to filter out outlier from CDEC data

### Using USGS min and max to filter out outliers from CDEC

```{r}
library(dplyr)
#USGS
DCV_USGS <- readNWISdv(11383500, "00010") # parameter code 00010: water temperature

# Format to make tidier
DCV_daily_temps <- DCV_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)

temp_rangeUSGS <- DCV_daily_temps |> 
  summarise(min_tempUSGS = min(temp_degC),
            max_tempUSGS = max(temp_degC))
#CDEC
DCV_CDEC <- cdec_query(station = "DCV", dur_code = "H", sensor_num = "25", start_date = "1995-01-01")

DCV_hourly_temps <- DCV_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

DCV_temp_filtered <- DCV_hourly_temps |> 
  filter(temp_degC >= 3.3,
         temp_degC <= 27)

ggplot(DCV_temp_filtered, aes(date, temp_degC)) +
  geom_line()
#TODO add min and max datasets
```



### Clean data and summarize as mean daily

```{r}

# 1. Calculate the daily average temperature
DCV_daily_temps <- DCV_hourly_temps %>% 
  filter(!is.na(temp_degC),
         temp_degC < 37, temp_degC > -18) %>%
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC, na.rm = T),
            max_temp_degC = max(temp_degC, na.rm = T)) 
```

```{r}
# save final version of butte data
# from ashley: erin and i have started using the here package, especially in markdowns because it makes it easier for file path reading
write_rds(DCV_daily_temps, here::here("data-raw", "standard-format-data-prep","temp_data", "deer_temp.rds"))
```

## Feather River

### CDEC

```{r}

# cdec_datasets("GRL")

GRL_CDEC <- cdec_query(station = "GRL", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")

GRL_hourly_temps <- GRL_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(GRL_hourly_temps$date, n = 1)
head(GRL_hourly_temps$date, n = 1)

# Plot GRL
ggplot(GRL_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y="Temperature (deg C)", title = "Temperature at Feather River", caption = "CDEC")

#Adding "FRA" dataset
FRA_CDEC <- cdec_query(station = "FRA", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")

FRA_hourly_temps <- FRA_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Plot FRA
ggplot(FRA_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y="Temperature (deg C)", title = "Temperature at FRA Feather River", caption = "CDEC")
```

CDEC provides temperature data for Feather River from 2003-01-05 to
2007-06-15.

### CDEC Outliers Removed

```{r}
# Plot without outliers
GRL_hourly_temps %>%
   filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "CDEC Water Temperature at Feather River", caption = "CDEC")
```

### USGS

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

# GRL_USGS <- readNWISdv(11407000, "00010") # parameter code 00010: water temperature
# # write.csv(GRL_USGS,"GRL_USGS.csv", row.names = FALSE)
# 
# # Read created CSV
# GRL_USGS <- read_csv("GRL_USGS.csv")
# GRL_USGS

#Adding Oroville station
# ORO_USGS <- readNWISdv(11406920, "00010", statCd = "00003") # parameter code 00010: water temperature
# 
# # Format to make tidier
# ORO_daily_meantemps <- ORO_USGS %>%
#   select(Date, temp_degC =  X_00010_00003) %>%
#   as_tibble() %>% 
#   rename(date = Date)
# no data for daily statistics mean temp on default statCd = 00003, so using min and max


ORO_USGS <- readNWISdv(11406920, "00010", statCd = "00001") # parameter code 00010: water temperature

# Format to make tidier
ORO_daily_maxtemps <- ORO_USGS %>%
  select(Date, temp_degC =  X_00010_00001) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(ORO_daily_maxtemps)

ggplot(ORO_daily_maxtemps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Max Temperature at Feather Oroville", caption = "USGS")

ORO_USGS <- readNWISdv(11406920, "00010", statCd = "00002") # parameter code 00010: water temperature

# Format to make tidier
ORO_daily_mintemps <- ORO_USGS %>%
  select(Date, temp_degC =  X_00010_00002) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(ORO_daily_mintemps)

ggplot(ORO_daily_mintemps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Min Temperature at Feather Oroville", caption = "USGS")

#TODO compare min and max plots
```

USGS does not provide temperature data for Feather River. This was
confirmed with mapper tool accessed from waterdata.usgs.gov. 


Notes from Badhia:
USGS
* data for Feather River across three stations:
* Thermalito Afterbay Rel to Feather R NR Oroville 11406920,
* Feather R A Yuba City CA 11407700, (not using this one because it is too low on the stream) 
* Feather R A Oroville CA 11407000 (the one used).
* Monitoring periods range from 1968-1992, 1964-1976, and 1958-1992 respectively 

CDEC
* only has one station and data from time period stated above was found
* This same station data was used for va-habitat EDA temps

TODO pull data from 11406920. Plot to see all four (2USGS and 1CDEC, plus FRA from CDEC) data sets

The CDEC data does not cover the extent of the monitoring period for
Feather River


```{r}
# write_rds(GRL_hourly_temps, "../../data-raw/standard-format-data-prep/temp_data/feather_temp.rds")
```

## Mill Creek

### CDEC

```{r}

# cdec_datasets("MLM") 

MLM_CDEC <- cdec_query(station = "MLM", dur_code = "H", sensor_num = "25", start_date = "1996-01-01")


MLM_hourly_temps <- MLM_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(MLM_hourly_temps$date, n = 1)
head(MLM_hourly_temps$date, n = 1)

# Plot DCV
ggplot(MLM_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Mill Creek", caption = "CDEC")
```

CDEC provides temperature data for Mill Creek from 1998-09-30 to
2022-05-27

### CDEC Outliers Removed

```{r}
# Plot without outliers
MLM_hourly_temps %>%
    filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "CDEC Water Temperature at Mill Creek", caption = "CDEC")
```

### USGS

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

MLM_USGS <- readNWISdv(11381500, "00010") # parameter code 00010: water temperature temperature

# Format to make tidier
MLM_daily_temps <- MLM_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  as_tibble() %>% 
  rename(date = Date)

# Date Ranges
tail(MLM_daily_temps$date, n = 1)
head(MLM_daily_temps$date, n = 1)

# Plot DCV
ggplot(MLM_daily_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Temperature at Mill Creek", caption = "USGS")
```

USGS provides daily temperature daily for Mill Creek from 2016-10-01 to
2022-05-25

#### Mean Daily USGS vs Daily CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
MLM_combined = MLM_hourly_temps %>% 
  full_join(MLM_daily_temps, by = "date") %>% 
  rename("temp_degC.CDEC" = "temp_degC.x", "temp_degC.USGS" = "temp_degC.y")

MLM_combined %>% 
  group_by(date = date(date)) %>% 
  group_by(date) %>% 
  summarise(
    temp_CDEC = median(temp_degC.CDEC, na.rm = TRUE),
    temp_USGS = max(temp_degC.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_CDEC), color = "darkred") +
  geom_line(aes(y = temp_USGS), color = "steelblue", linetype = "twodash") +
  labs(y = "Temperature (deg C)", title = "Water Temperature at Mill Creek")

```

I would recommend using the CDEC data for Mill Creek due to the
significant data gaps for USGS.

Notes from Badhia:
CDEC
* There is another gauge along Mill (MCH) Mill Creek Bellow HWY 99. Temp Data ranges from 1998-present (hourly) and 2005-present (event)
USGS
* Daily data range is 1998-2023, and daily statistics ranges from 2016-2023 (used) 
* This gage station data was used on DSMtemperature 

## Clean data and summarize as mean daily

```{r}

# 1. Calculate the daily average temperature
MLM_daily_temps <- MLM_hourly_temps %>% 
  filter(!is.na(temp_degC),
         temp_degC < 37, temp_degC > -18) %>% 
  group_by(date) %>% 
  summarize(mean_temp_degC = mean(temp_degC),
            max_temp_degC = max(temp_degC)) 
```

```{r}
# save final version of butte data
# from ashley: erin and i have started using the here package, especially in markdowns because it makes it easier for file path reading
write_rds(MLM_daily_temps, here::here("data-raw", "standard-format-data-prep","temp_data", "mill_temp.rds"))
```

## Yuba River


Notes from Badhia:
USGS
* daily max and min Temp data from 1964-2003, with some data gaps, for
    station number 11421000 Yuba R NR Marysville (this station data was pulled for va-habitat eda, too)
* daily max and min Temp data from 1972-2005, Station 11418000 Yuba R
    BL Englebright Dam NR Smartsville has
CDEC
* Station YR7 Yuba R ABV HWY 70 has temp data (event) from 2019-09-04 to present 
* For va-habitat project a higher upstream gauge was used. Station code = YRS 

TODO pull in those two data sources listed above

##CDEC
```{r}
YRS_CDEC <- cdec_query(station = "YRS", dur_code = "H", sensor_num = "25", start_date = "1995-01-01")

YRS_hourly_temps <- YRS_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = parameter_value) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(YRS_hourly_temps$date, n = 1)
head(YRS_hourly_temps$date, n = 1)

# Plot
ggplot(YRS_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Water Temperature at Yuba River", caption = " CDEC")

#Adding another gage
YR7_CDEC <- cdec_query(station = "YR7", dur_code = "E", sensor_num = "146", start_date = "1995-01-01")
#note that this station only has event duration, not hourly
YR7_hourly_temps <- YR7_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = parameter_value) %>%
  select(date,time,temp_degC)

# Plot
ggplot(YR7_hourly_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)",title = "Water Temperature at YR7 Yuba River", caption = " CDEC")

```

#CDEC Outliers Removed
```{r}
# Plot without outliers
YRS_hourly_temps %>%
    filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "CDEC Water Temperature at Yuba River", caption = "CDEC")
```
### USGS
```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

YBAMR_USGS <- readNWISdv(11421000, "00010") # parameter code 00010: water temperature
glimpse(YBAMR_USGS) #no data on 00003

YBAMR_USGS <- readNWISdv(11421000, "00010", statCd = "00001") # parameter code 00010: water temperature

# Format to make tidier
YBAMR_daily_maxtemps <- YBAMR_USGS %>%
  select(Date, temp_degC =  X_00010_00001) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(YBAMR_daily_maxtemps)

YBAMR_USGS <- readNWISdv(11421000, "00010", statCd = "00002") # parameter code 00010: water temperature

# Format to make tidier
YBAMR_daily_mintemps <- ORO_USGS %>%
  select(Date, temp_degC =  X_00010_00002) %>%
  as_tibble() %>% 
  rename(date = Date)
glimpse(YBAMR_daily_mintemps)
#TODO plot min and max together
```


### Air Gage - code isn't working so comment out for now

```{r}
# use this code to get the data types for the station
# ncdc_datatypes(datasetid='GHCND', stationid='GHCND:USR0000CFOR', token = NOAA_KEY)
# 
# # used this code to figure out units. units for TAVG is in celsius tenths which means we need to divide by 10 to get celsisus
# with_units <- ncdc(datasetid='GHCND', stationid='GHCND:USR0000CFOR', token = NOAA_KEY, datatypeid='TAVG', startdate = '2010-05-01', enddate = '2010-10-31', limit=500, add_units = TRUE)
# head( with_units$data )
# 
# get_temp_yuba <- function(start_date, end_date, year) {
#   query <- rnoaa::ncdc(datasetid = 'GHCND', stationid = 'GHCND:USR0000CFOR', 
#                        datatypeid = 'TAVG', startdate = start_date, 
#                        enddate = end_date, token = NOAA_KEY, limit = 365)
#   # Get columns of interest
#   temp_data <- query$data %>%
#     select(date, value) %>%
#     mutate(date = as.Date(date))
#   # Return
#   write_csv(temp_data, here::here("data-raw", "standard-format-data-prep", "temp_data", paste0(year, "_yuba_noaa_airtemp.csv")))
# }
# 
# # list date ranges, note we can only take yearly data
# # note that there is no data after 2020
# date_ranges <- tibble(year = c(seq(from = 2001, to = 2020))) %>% 
#   mutate(start_date = paste0(year, "-01-01"),
#          end_date = paste0(year, "-12-31"))
#                       
# pmap(date_ranges, get_temp_yuba)
# ```
# 
# ### QC data
# 
# ```{r}
# # read in all the yuba air temp files
# files <- list.files(path = here::here("data-raw", "standard-format-data-prep", "temp_data"), pattern = "_yuba_noaa_airtemp.csv")
# path <- here::here("data-raw", "standard-format-data-prep", "temp_data")
# file_path <- paste0(path, "/",files)
# # fread uses data.table
# yuba_air <- file_path %>% 
#   map_df(~fread(.)) %>% 
#   rename(air_tempC = value) %>% 
#   mutate(date = as.Date(date),
#          air_tempC = air_tempC/10)
# 
# ggplot(yuba_air, aes(x=date, y = air_tempC)) +
#   geom_line(color = "darkred") +
#   labs(y = "Air temperature (deg C)") +
#   theme(axis.text = element_text(size = 12))
```

<!-- ### Save data -->

<!-- ```{r} -->

<!-- write_rds(yuba_air, here::here("data-raw", "standard-format-data-prep","temp_data", "yuba_temp_air.rds")) -->

<!-- ``` -->

## Sacramento River - Knights Landing

```{r}
# cdec_datasets("WLK")

WLK_CDEC <- cdec_query(station = "WLK", dur_code = "E", sensor_num = "25", start_date = "1995-01-01")

WLK_hourly_temps <- WLK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         temp_degC = fahrenheit.to.celsius(parameter_value, round = 1)) %>%
  select(date,time,temp_degC)

# Date Ranges
tail(WLK_hourly_temps$date, n = 1)
head(WLK_hourly_temps$date, n = 1)

# Plot WLK
ggplot(WLK_hourly_temps, aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "CDEC")
```

CDEC provides temperature data for Sacramento River - Knights Landing
from 2012-11-15 to 2022-06-16

### CDEC Outliers Removed

```{r}
# Plot without outliers
WLK_hourly_temps %>%
  filter(temp_degC < 37, temp_degC > -18) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "CDEC")
```

### USGS

```{r}
# Using https://flowwest.shinyapps.io/jpe-eda-app/ to find temperature gage
# Using mapper to check if temp data is available for the site https://maps.waterdata.usgs.gov/mapper/index.html

WLK_USGS <- readNWISdv(11390500, "00010") # parameter code 00010: water temperature temperature

# Format to make tidier
WLK_daily_temps <- WLK_USGS %>%
  select(Date, temp_degC =  X_00010_00003) %>%
  #filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date) %>% 
  #filter(year(date) > 2000) %>%
  glimpse

# Date Ranges
tail(WLK_daily_temps$date, n = 1)
head(WLK_daily_temps$date, n = 1)

# Plot
ggplot(WLK_daily_temps, aes(x=date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year",y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
```

Temperature data for Sacramento River - Knights landing has data from
1995-01-01 to 2022-06-14, but we should explore further the signifcant
data gap and identify that period.

### Data before Missing Data Gap

```{r}
### Identify date data starts to be missing
WLK_daily_temps %>%
  filter(date < "2004-01-01") %>%
  arrange(desc(date))
  # 1998-09-30

# data before missing data gap
WLK_daily_temps %>%
  filter(date < "1999-10-01") %>% 
  arrange(desc(date)) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
```

### Data after missing data gap

```{r}
### Identify date data starts to be missing
WLK_daily_temps %>%
  filter(date > "2015-01-01") %>%
  arrange(date)
  # 2016-10-01

# data after missing data gap
WLK_daily_temps %>%
  filter(date >= "2016-10-01") %>% 
  arrange(desc(date)) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = temp_degC), color = "darkred") +
  labs(x = "Year", y ="Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
#
```

Temperature data for Sacramento River - Knights landing is missing
significant data from the period of 1998-09-30 to 2016-09-31.

#### Mean Daily USGS vs Daily CDEC data

```{r}
WLK_combined = WLK_hourly_temps %>% 
  full_join(WLK_daily_temps, by = "date") %>% 
  rename("temp_degC.CDEC" = "temp_degC.x", "temp_degC.USGS" = "temp_degC.y")

WLK_combined %>% 
  group_by(date = date(date)) %>% 
  summarise(
    temp_CDEC = median(temp_degC.CDEC, na.rm = TRUE),
    temp_USGS = max(temp_degC.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = temp_CDEC), color = "darkred") +
  geom_line(aes(y = temp_USGS), color = "blue", linetype ="twodash") +
  labs(y = "Temperature (deg C)", title = "Temperature at Sacramento River- Knights Landing", caption = "USGS")
```
