---
title: "butte-creek-individual-survey-qc-checklist"
author: "Inigo Peng"
date: "10/21/2021"
output: rmarkdown::github_document
---
---
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
```
# BUtte Creek Individual Survey Data  

## Description of Monitoring Data

**Timeframe:** 2014-2020


**Completeness of Record throughout timeframe:**  



**Sampling Location:** Various sampling locations on Butte Creek.

TODO: Upper survey?


**Data Contact:** [Jessica Nichols](mailto::Jessica.Nichols@Wildlife.ca.gov)


Additional Info:  
The carcass data came in 12 documents for each year. We identified the 'SurveyChops' and 'SurveyIndividuals' datasets as the documents with the most complete information and joined them for all of the years.

## Access Cloud Data
```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()

# git data and save as xlsx
read_from_cloud <- function(year){
  gcs_get_object(object_name = paste0("adult-holding-redd-and-carcass-surveys/butte-creek/data-raw/", year, "_SurveyIndividuals.xlsx"),
               bucket = gcs_get_global_bucket(),
               saveToDisk = paste0(year,"_raw_surveyindividuals.xlsx"),
               overwrite = TRUE)
  data <- readxl::read_excel(paste0(year,"_raw_surveyindividuals.xlsx")) %>% 
    glimpse()
}

open_files <- function(year){
  data <- readxl::read_excel(paste0(year, "_raw_surveyindividuals.xlsx"),
                   col_types = c("numeric",
                                 "text",
                                 "numeric",
                                 "numeric",
                                 "date",
                                 "text",
                                 "text",
                                 "text",
                                 "text",
                                 "numeric",
                                 "text",
                                 "numeric",
                                 "numeric",
                                 "text",
                                 "text",
                                 "text",
                                 "text",
                                 "numeric",
                                 "text",
                                 "text",
                                 "text",
                                 "text",
                                 "text",
                                 "text",
                                 "text",
                                 "text"))
  return (data)
}
years <- c(2014,2015,2016, 2017, 2018, 2020)
# year <- 2020
# raw_data <- purrr::map(years, read_from_cloud)

raw_data <- purrr::map(years, open_files) %>%  
  reduce(bind_rows) %>% 
write_csv(raw_data, "raw_individual_data.csv")
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
raw_individuals_data <- read_csv("raw_individual_data.csv")%>% 
  glimpse()
```

## Data Transformations


```{r}
cleaner_data<- raw_individuals_data %>%
  janitor::clean_names() %>%
  rename('location' = 'location_cd',
         'fork_length_cm' = 'f_lcm',
         'condition' = 'condition_cd',
         'spawning_status' = 'spawned_cd') %>% 
  select(-c('week','year'))%>% 
  mutate(date = as.Date(date)) %>%
  glimpse()
```

## Explore `date`
```{r}
cleaner_data%>%
  ggplot(aes(x = date)) +
  geom_histogram(binwidth = 7, position = 'stack', color = "black") +
  labs(title = "Value Counts For Survey Season Dates")+
  theme(legend.text = element_text(size = 8))
```

```{r}
summary(cleaner_data$date)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$date))/nrow(cleaner_data), 3)*100` % of values in the `date` column are NA.

## Explore Categorical Variables

```{r}
cleaner_data %>% 
  select_if(is.character) %>% colnames()
```
### Variable:`location`
```{r}
table(cleaner_data$location)
```

### Variable:`section_cd`
```{r}
table(cleaner_data$section_cd)
```

### Variable:`way_pt`
```{r}
table(cleaner_data$way_pt)
```
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(way_pt = tolower(way_pt))
table(cleaner_data$way_pt)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$way_pt))/nrow(cleaner_data), 3)*100` % of values in the `way_pt` column are NA.


### Variable:`species_code`
```{r}
table(cleaner_data$species_code)
```
### Variable:`disposition`
```{r}
table(cleaner_data$disposition)
```
### Variable:`sex`
```{r}
table(cleaner_data$sex)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$sex))/nrow(cleaner_data), 3)*100` % of values in the `sex` column are NA.

### Variable:`condition`
```{r}
table(cleaner_data$condition)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$condition))/nrow(cleaner_data), 3)*100` % of values in the `condition` column are NA.

### Variable:`spawning_status`
```{r}
table(cleaner_data$spawning_status)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$spawning+status))/nrow(cleaner_data), 3)*100` % of values in the `spawning_status` column are NA.

### Variable:`tissue_nu`
```{r}
table(cleaner_data$tissue_nu)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$spawning_status))/nrow(cleaner_data), 3)*100` % of values in the `spawning_status` column are NA.

### Variable:`otolith_nu`
```{r}
table(cleaner_data$otolith_nu)
```
