---
title: "Battle Creek Upstream Passage Monitoring Video Data QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Battle Creek Upstream Passage Monitoring Video Data

## Description of Monitoring Data

These data were collected to capture the number of fish that pass the barrier weir at Coleman National Fish Hatchery from 1998 to 2019. Observations were captured via video footage of passing fish. For video data, prior to 2008, fish were viewed via a funnel video setup on the fish ladder. In 2008 the video system was an instream system with viewing via an open section of stream. Beginning in 2009 to the present, fish are viewed via a crowder video setup as they pass through the fish ladder. 						

**Timeframe:** 1998 - 2012

**Video Season:** Video is run all year. 

**Completeness of Record throughout timeframe:** Data for every year in timeframe. 

**Sampling Location:** Battle Creek fish ladder. 

**Data Contact:** [Natasha Wingerter](mailto:natasha_wingerter@fws.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
getwd() #to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "adult-upstream-passage-monitoring/battle-creek/data-raw/battle_creek_upstream_passage_datas.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_battle_creek_passage_data.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
sheets <- readxl::excel_sheets("raw_battle_creek_passage_data.xlsx")
sheets
raw_video_data <- read_excel("raw_battle_creek_passage_data.xlsx", sheet = "Video") %>% glimpse()
```

## Data transformations

```{r}
cleaner_video_data <- raw_video_data %>% 
  janitor::clean_names() %>% 
  filter(species == "Chinook Salmon (CHN)") %>%
  rename("run" = run_call,
         "adipose" = adipose_fin, 
         "date" = sample_date,
         "time" = time_passed,
         "up" = passed_up,
         "down" = passed_down) %>%
  mutate(date = as.Date(date),
         time = hms::as_hms(time)) %>%
  select(-net_total, # net total just equals difference in count up and count down 
         -project, -species, # all species = chn, all project = video 
         -notes) %>% # all notes = NA
  pivot_longer(!c(date, time, adipose:run), 
               names_to = "passage_direction",
               values_to = "count") %>%
  glimpse()
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_video_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `count`

**Plotting Passage Counts over Period of Record**

```{r}
cleaner_video_data %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  ggplot(aes(x = fake_date, y = count, fill = passage_direction)) + 
  geom_col() + 
  facet_wrap(~year(date), scales = "free") + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1901-04-01"), as.Date("1901-09-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 23),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Count of Passage All Runs", 
       x = "Date")  
```

Spring Run Chinook appear to be seen moving upstream April through September.

```{r}
# Boxplots of daily counts by year
cleaner_video_data %>% group_by(date, passage_direction) %>%
  mutate(daily_count_upstream = sum(count)) %>%
  mutate(year = as.factor(year(date))) %>% 
  ungroup() %>%
  ggplot(aes(x = year, y = daily_count_upstream, color = passage_direction)) + 
  geom_boxplot() + 
  theme_minimal() +
  theme(text = element_text(size = 23),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Passage Count Sumarized by Year All Runs") 
```

```{r}
cleaner_video_data  %>%
  mutate(year = as.factor(year(date))) %>%
  filter(run %in% c("FR", "LF", "SR", "WR")) %>% # Filter to only show runs that have more than one data point and are not NA/Unknown
  group_by(year, passage_direction, run) %>%
  summarise(total_count = sum(count)) %>%
  ggplot(aes(x = year, y = total_count, fill = passage_direction)) + 
  geom_col(position = "dodge") + 
  theme_minimal() +
  labs(title = "Total Yearly Fish Counts by Run",
       y = "Total fish count") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  facet_wrap(~run)
```
We mainly have video data describing counts of Spring Run Fish 

**Numeric Summary of Passage Counts over Period of Record**

```{r}
# Table with summary statistics
summary(cleaner_video_data$count)

# daily numeric summary 
cleaner_video_data %>% group_by(date) %>%
  summarise(count = sum(count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_video_data$count))/nrow(cleaner_video_data), 3) * 100` % of values in the `count` column are NA. However, there are clearly gaps in data. More investigation needs to be done to see if 0 is a real 0 or if it can be explained by other factors (outages).


## Explore Categorical variables: {.tabset}

```{r}
cleaner_video_data %>% select_if(is.character) %>% colnames()
```

### Variable: `adipose`
```{r}
table(cleaner_video_data$adipose) 
```
Fix inconsistencies with spelling, capitalization, and abbreviations.  

```{r}
# Fix yes/no/unknown
cleaner_video_data$adipose = tolower(cleaner_video_data$adipose)
table(cleaner_video_data$adipose) 
```
  
**NA or Unknown Values**

* `r round(sum(is.na(cleaner_video_data$adipose))/nrow(cleaner_video_data), 3) * 100` % of values in the `adipose` column are NA. 
* `r round(sum(cleaner_video_data$adipose == "unknown", na.rm = T)/nrow(cleaner_video_data), 3) * 100` % of values in the `adipose` column are`unknown`.

### Variable: `run`

```{r}
table(cleaner_video_data$run) 
# description <- domain_description[which(domain_description$Domain == "Run"), ]$Description
```
Fix inconsistencies with spelling, capitalization, and abbreviations.  

```{r}
# Fix yes/no/unknown
cleaner_video_data$run = if_else(cleaner_video_data$run == "UNK", "unknown", cleaner_video_data$run)
```

**NA or Unknown Values**
  
* `r round(sum(is.na(cleaner_video_data$run))/nrow(cleaner_video_data), 3) * 100` % of values in the `run` column are NA.
* `r round(sum(cleaner_video_data$run == "unknown", na.rm = T)/nrow(cleaner_video_data), 3) * 100` % of values in the `run` column are`unknown`.

### Variable: `comments`

10 unique comments are shown bellow. Comments are only present on 12% of the data. 

```{r}
unique(cleaner_video_data$comments)[1:10]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_video_data$comments))/nrow(cleaner_video_data), 3) * 100` % of values in the `comments` column are NA.

### Variable: `passage_direction`

```{r}
table(cleaner_video_data$passage_direction)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_video_data$passage_direction))/nrow(cleaner_video_data), 3) * 100` % of values in the `passage_direction` column are NA.

## Summary of identified issues

* No info on viewing condition/outages/gaps in sampling 

## Save cleaned data back to google cloud 

```{r}
battle_passage_video <- cleaner_video_data %>% 
  glimpse()
```

```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(battle_passage_video,
           object_function = f,
           type = "csv",
           name = "adult-upstream-passage-monitoring/battle-creek/data/battle_passage_video.csv")
```

