---
title: "deer-creek-adult-upstream-passage-monitoring-qc-checklist"
author: "Inigo Peng"
date: "10/19/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
```

# Mill Creek Adult Holding Survey Data 2012 to 2020 

**Description of Monitoring Data**

**Timeframe:** 

2012 to 2020

**Completeness of Record throughout timeframe:**

TODO

**Sampling Location:** 

* Ward Dam   

**Data Contact:** [Matt Johnson](mailto:Matt.Johnson@wildlife.ca.gov)

TODO: Add issues, descriptions
## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "adult-upstream-passage-monitoring/mill-creek/data-raw/Mill Creek SRCS Daily Video Passage Estimates 2012-2020.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "mill_creek_passage_estimate_raw.xlsx")
```


```{r}
sheets <- readxl::excel_sheets('mill_creek_passage_estimate_raw.xlsx')
list_all <- lapply(sheets, function(x) readxl::read_excel(path = "mill_creek_passage_estimate_raw.xlsx", sheet = x, col_types = c("text", "numeric", "numeric", "numeric", "text")))
raw_data <- dplyr::bind_rows(list_all) %>% 
  glimpse()
```

## Data Transformations

```{r}
cleaner_data <- raw_data %>% 
  set_names(tolower(colnames(raw_data))) %>% 
  select(-"...5") %>% #comments describe dates
  rename("count" =  'adult spring-run passing ward dam',
         "flow" = "avg daily flow below ward dam",
         "temperature"= "avg daily h2o temp below ward dam") %>% 
  filter(date != "Totals:", date != "Total:") %>%
  mutate(date = case_when(
    str_length(date) == 5 ~ as.Date(as.numeric(date), origin="1899-12-30"),
    str_detect(date, '[/]') ~ as.Date(date, "%m/%d/%Y"))
    # TRUE ~ as.Date(date))
  ) %>% 
  glimpse()
```

## Explore `date`
```{r}
cleaner_data%>%
  ggplot(aes(x = date)) +
  geom_histogram(binwidth = 7, position = 'stack', color = "black") +
  labs(title = "Value Counts For Survey Season Dates")+
  theme(legend.text = element_text(size = 8))
```

```{r}
summary(cleaner_data$date)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$date))/nrow(cleaner_data), 3)*100` % of values in the `date` column are NA.

## Explore Numerical Values

### Variable:`count` 
```{r}
cleaner_data %>%
  filter(date != is.na(date)) %>%
  mutate(year = as.factor(year(date))) %>% 
  # glimpse()
  ggplot(aes(x=date, y = count))+
  geom_line()+
  facet_wrap(~year, scales = "free")+
  theme_minimal()+
  labs(title = "Daily Passage Estimate From 2012 - 2020")
```


```{r}
cleaner_data %>% 
  filter(date != is.na(date)) %>%
  mutate(year = as.factor(year(date))) %>%
  # glimpse()
  group_by(year) %>% 
  summarise(total = sum(count, na.rm  = TRUE)) %>%
  # glimpse()
  ggplot(aes(x = year, y = total, group = 1))+
  geom_line()+
  geom_point(aes(x=year, y = total))+
  theme_minimal()+
  labs(title = "Passage Estimate from 2012 - 2020",
       y = "Total Count")
```

**Numeric Monthly Summary of count From 2012 to 2020**
```{r}
cleaner_data %>%
  group_by(month(date)) %>%
  summarise(count = sum(count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$count))/nrow(cleaner_data), 3)*100` % of values in the `date` column are NA.

### Variable:`flow`

```{r fig.height=8, fig.width=8}
cleaner_data %>% 
  filter(date != is.na(date)) %>%
  group_by(date) %>%
  mutate(avg_flow = mean(flow, na.rm = T)) %>%
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = 1900,
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = avg_flow, color = year)) + 
  scale_color_brewer(palette = "Dark2")+
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Water Flow (colored by year)",
       y = "Average Daily Flow", 
       x = "Date")  
```
```{r fig.height=8, fig.width=8}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x=flow, fill = year))+
  scale_fill_brewer(palette = "Dark2")+
  geom_histogram()+
  theme_minimal()+
  labs(title = "Distribution of Flow")
```

**Numeric Annual Summary of flow From 2012 to 2020**
```{r}
cleaner_data %>%
  group_by(year(date)) %>%
  summarise(flow = mean(flow, na.rm = T)) %>%
  pull(flow) %>%
  summary()
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$count))/nrow(cleaner_data), 3)*100` % of values in the `date` column are NA.

### Variable:`temperature`

```{r fig.height=8, fig.width=8}
cleaner_data %>% 
  filter(date != is.na(date)) %>%
  group_by(date) %>%
  mutate(avg_temp = mean(temperature, na.rm = T)) %>%
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = 1900,
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = avg_temp, color = year)) + 
  scale_color_brewer(palette = "Dark2")+
  geom_point(alpha = .25) + 
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Daily Water Temperature (colored by year)",
       y = "Average Daily Temperature", 
       x = "Date")  
```

```{r fig.height=8, fig.width=8}
cleaner_data %>% 
  filter(date != is.na(date)) %>%
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x=temperature, fill = year))+
  scale_fill_brewer(palette = "Dark2")+
  geom_histogram(bin = 5)+
  theme_minimal()+
  labs(title = "Distribution of Temperature")
```

**Numeric Annual Summary of temperature From 2012 to 2020**
```{r}
cleaner_data %>%
  group_by(year(date)) %>%
  summarise(temperature = mean(temperature, na.rm = T)) %>%
  pull(temperature) %>%
  summary()
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$count))/nrow(cleaner_data), 3)*100` % of values in the `date` column are NA.

### Add cleaned data back onto google cloud
```{r}
mill_upstream_estimate <- cleaner_data %>% glimpse()
```

```{r}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(mill_upstream_estimate,
           object_function = f,
           type = "csv",
           name = "adult-upstream-passage-monitoring/mill-creek/data/mill_upstream_estimate.csv")
