---
title: "Feather River Carcass Survey QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
```

# Feather River Carcass Survey QC

## Description of Monitoring Data

The feather river carcass survey was provided to us in two sets of access databases: 

1) Historical data was provided in a "Escapement DBs FlowWest.zip" that contains: 

* **Carcass 2000.mdb** 
  (Identified tables of interest)
  * ChopEnvTBL (ChopEnvID,	Week's Tag Color,	Date,	Time Start,	Time Stop,	Crew,	Recorder,	Turbidity, Bottle Number,	Weather Code,	Comments, Week)
  * ChopRecovTBL (Section	Unit, # of Adult Males Chopped,	# of Adult Females Chopped,	# of Grisle Males Chopped,# of Grilse Females Chopped,	Number Spawned,	RecovID,	Chop Env ID)
  * cwtTBL (Auto Fish ID, Date, Section, Unit, FL, Sex, Eggs, Head Tag Number, CWT Code)
  * RecovTagTBL (RecovID,	Tag Number,	RecovTagID,	RelStat,	MultRecov)
  * TagEnvTBL (Date, Recorder's Initials, Week's Tag Color, TagEnvID, Week)
* **Carcass 2001.mdb**
  * ChopEnvTBL (ChopEnvID,	Date,	Time,	Crew,	Recorder, TagCol,	Flow,	Vis,	Weather,	Comment,	WeekNum)
  * ChopRecovTBL (Sect,	Chan,	Minute,	Male,	Fem,	Gril,	RecovID,	ChopEnvID)
  * cwtTBL (cwtID,	Date,	Sect,	FL,	Sex,	HTNum,	EggRet,	Comment)
  * TagEnvTBL (Date,	Crew,	TagCol,	Recorder,	Flow,	WeekNum,	TagEnvID)
* **Carcass 2002.mdb**
  * Even more tables in this one...

* **Carcass 2003_converted.accdb**
  * ChopChannelTBL (ChanID,	Sect,	Chan,	Min,	Male,	Fem,	Gril,	TagCount,	HeaderID)
  * ChopHeaderTBL (headerID, TagCol, Date, Crew,	Time,	Recorder,	Weather,	Comment,	WeekNum)
  * ChopRecovTBL (RecovID,	RecovColor,	RecovCount,	ChanID)
  * CWTHeaderTBL (headerID,	Date,	TagCol,	Crew,	WeekNum)
  * CWTTagTBL (TagID,	Sect,	RelLoc,	FL,	Sex,	Spawn,	Clip,	SampType,	CarcSta,t	SampNum,	HeadTag,	Comment,	HeaderID)
  
* **Carcass 2004.mdb**
  * 
* **Carcass 2005.mdb**
* **Carcass 2006.mdb**
* **Carcass 2007.mdb**
* **Carcass 2008.mdb**
* **Carcass 2009.mdb**
* **Carcass 2010.mdb**
* **FINAL Carcass 20011.mdb**
* **FINAL Carcass 20013.mdb**
* **FINAL Carcass 20014.accdb**
* **FINAL Carcass 20015.accdb**
* **FINAL Carcass 20016.accdb**

Each databse contains a lot of tables describing carcass surveys on the feather including: carcassSurvey, carcassChops, carcassIndividual...ect

2) Recent data in the CAMP carcass database was sent in a "CAMP DB for FlowWest.zip" that contains: 

* CAMP_Escapement_20210412mdb
* CAMP_EscapementEntry.mdb
* CAMP_EscapementEntry_Backup.mdb

**Timeframe:** 

**Video Season:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

# git data and save as xlsx

```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
```

## Data transformations

```{r}
# For different excel sheets for each year read in and combine years here
```

```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
```

### Variable: `[name]`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
# Filter clean data to show only categorical variables
```


### Variable: `[name]`
```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Summary of identified issues

* List things that are funcky/bothering us but that we don't feel like should be changed without more investigation

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```

