---
title: "feather-river-adult-holding-redd-survey-qc-checklist"
author: "Inigo Peng"
date: "9/30/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(leaflet)
library(knitr)
```
# Feather River Redd Survey Data  

## Description of Monitoring Data

**Timeframe:** 2009 to 2020


**Completeness of Record throughout timeframe:**  

* Longitude and latitude data are not available for 2009-2012, 2019-2020. NA values will be filled in for these data sets in final cleaned data set.


**Sampling Location:** Various sampling locations on Feather River.


**Data Contact:** [Chris Cook](mailto::Chris.Cook@water.ca.gov)


Additional Info:  
1. Latitude and longitude are in NAD 1983 UTM Zone 10N
2. The substrate is observed visually and an estimate of the percentage of 5 size classes is recorded:   

* fine <1cm  
* small 1-5cm  
* medium 6-15cm  
* large 16-30cm  
* boulder >30cm

```{r, eval = FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
```
```{r, include = F}
# git data and save as xlsx

read_from_cloud <- function(year){
  gcs_get_object(object_name = paste0("adult-holding-redd-and-carcass-surveys/feather-river/data/feather_redd_", year, ".csv"),
               bucket = gcs_get_global_bucket(),
               saveToDisk = paste0("feather_redd_", year, ".csv"),
               overwrite = TRUE)
  data <- read.csv(paste0("feather_redd_", year, ".csv"))  
    # glimpse()
}

#Read in data from google cloud, glimpse raw data: 
  
years <- c(2009, 2010, 2011, 2012, 2013, 2014,2015,2016, 2017, 2018, 2019, 2020)
combined_data <- purrr::map(years, read_from_cloud) %>%  
  reduce(bind_rows) 
write_csv(combined_data, "combined_feather_redd.csv")

#combined_data <- read_csv("combined_feather_redd.csv")
```


## Data transformations
```{r}
clean_data <- combined_data %>% 
  mutate('date' = as.Date(date),
         'latitude' = as.numeric(latitude),
         'longitude'= as.numeric(longitude)) %>%
  rename(velocity = velocity_m_per_s) %>%
  glimpse()
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- clean_data %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(clean_data),
                          description = c("Sample date",
                                          "Nominal description of location",
                                          "Type of observation. A = polygon/area, P = point, Q = questionnable redd",
                                          "Number of redds observed",
                                          "Number of salmon observed",
                                          "Depth of water at observation location in meters",
                                          "Depth of the pot in meters",
                                          "Velocity of water at observation location in m/s but standardized to ft/s",
                                          "Percent of fine substrate (fine <1cm)",
                                          "Percent of small substrate (small 1-5cm)",
                                          "Percent of medium substrate (medium 6-15cm)",
                                          "Percent of large substrate (large 16-30cm)",
                                          "Percent boulder (boulder >30cm)",
                                          "Width of redd in meters",
                                          "Length of redd in meters",
                                          "GPS X point",
                                          "GPX Y point"
                                          ),
                          data_type = c("date",
                                          "character",
                                         "factor",
                                          "integer",
                                          "integer",
                                          "numeric",
                                          "numeric",
                                          "numeric",
                                          "integer",
                                          "integer",
                                          "integer",
                                          "integer",
                                          "integer",
                                          "numeric",
                                          "numeric",
                                          "numeric",
                                          "numeric"
                                          ),
                          encoding = c(NA,
                                          NA,
                                         "Area, Point, Questionnable redds",
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA,
                                          NA
                                          ),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
# saveRDS(data_dictionary, file = "data/feather_redd_data_dictionary.rds")
```


## Explore `date`

```{r}
summary(clean_data$date)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(clean_data$date))/nrow(clean_data), 3)*100` % of values in the `date` column are NA.

## Explore Categorical Variables

```{r}
clean_data %>% 
  select_if(is.character) %>% colnames()
```

### Variable:`location`

```{r}
table(clean_data$location)
```
**NA and Unknown Values** 

* `r round(sum(is.na(clean_data$location))/nrow(clean_data), 3)*100` % of values in the `location` column are NA.

## Variable:`type`

# Description:  

* Area - polygon mapped with Trimble GPS unit  

* Point - points mapped with Trimble GPS unit  

* Questionable redds - polygon mapped with Trimble GPS unit where the substrate was disturbed but did not have the proper characteristics to be called a redd - it was no longer recorded after 2011 

```{r}
table(clean_data$type)
```

## Expore Numeric Variables

```{r}
clean_data %>% 
  select_if(is.numeric) %>% colnames()
```


### Variable:`salmon_count`  

```{r}
clean_data %>% 
  filter(is.na(date)==FALSE) %>% 
  ggplot(aes(x = date, y = salmon_count)) + 
  geom_col() +
  facet_wrap(~year(date), scales = "free") +
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Daily Salmon Count From 2012 to 2020")
```
```{r}
clean_data %>% 
  filter(is.na(date)==FALSE) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>%
  summarise(total_salmon_count = sum(salmon_count, na.rm = T)) %>%
  # glimpse()
  ggplot(aes(x = year, y = total_salmon_count)) + 
  geom_col() +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Annual Total Salmon Count From 2012 to 2020",
       x = 'Year',
       y = 'Total Salmon Count')
```


**Numeric Daily Summary of salmon_count from 2009 to 2020**
```{r}
clean_data %>%
  group_by(date) %>%
  summarise(count = sum(salmon_count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$salmon_count))/nrow(clean_data), 5)*100` % of values in the `salmon_count` column are NA. 

### Variable:`redd_count` 

```{r}
clean_data %>% 
  filter(is.na(date)==FALSE) %>% 
  ggplot(aes(x = date, y = redd_count)) + 
  geom_col() +
  facet_wrap(~year(date), scales = "free") +
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Daily Redd Count From 2012 to 2020")
```
```{r}
clean_data %>% 
  filter(is.na(date)==FALSE) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>%
  summarise(total_redd_count = sum(redd_count, na.rm = T)) %>%
  # glimpse()
  ggplot(aes(x = year, y = total_redd_count)) + 
  geom_col() +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10,angle = 90, vjust = 0.5, hjust=0.1)) +
  theme(axis.text.y = element_text(size = 8))+
  labs(title = "Annual Total Redd Count From 2009 to 2020",
       x = 'Year',
       y = 'Total Redd Count')
```

**Numeric Daily Summary of total redd_count From 2009 to 2020**
```{r}
clean_data %>%
  group_by(date) %>%
  summarise(count = sum(redd_count, na.rm = T)) %>%
  pull(count) %>%
  summary()
```

**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$redd_count))/nrow(clean_data), 5)*100` % of values in the `redd_count` column are NA.

### Variable:`redd_width_m` 

```{r}
clean_data %>%
  ggplot(aes(x = redd_width_m)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "grey") +
  scale_x_continuous(breaks = round(seq(min(clean_data$redd_width_m, na.rm = TRUE), max(clean_data$redd_width_m, na.rm = TRUE), by = 0.5),0))+
  labs(title = "Redd Width Distribution")+
  theme_minimal()
```


**Numeric Summary of redd_width_m from 2009 to 2020**
```{r}
summary(clean_data$redd_width_m)
```
**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$'redd_width_m'))/nrow(clean_data), 3)*100` % of values in the `redd_width_m` column are NA.


### Variable: `redd_length_m`  

```{r}
clean_data %>%
  ggplot(aes(x = redd_length_m)) +
  geom_histogram(binwidth = 0.5, color = "black", fill = "grey") +
  scale_x_continuous(breaks = round(seq(min(clean_data$redd_length_m, na.rm = TRUE), max(clean_data$redd_length_m, na.rm = TRUE), by = 1),1))+
  labs(title = "Redd Length Distribution") +
  theme_minimal()
```


**Numeric Summary of redd_length_m 2009 to 2020**

```{r}
summary(clean_data$redd_length_m)
```

**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$'redd_length_m'))/nrow(clean_data), 3)*100` % of values in the `redd_length_m` column are NA.

### Physical Attributes

### Variable:`percent_fine_substrate`
```{r}
# colourCount = length(unique(cleaner_data_2009$location))
# getPalette = colorRampPalette(brewer.pal(12, "Paired"))
clean_data %>%
  ggplot(aes(x = percent_fine_substrate)) +
  # scale_fill_manual(values = getPalette(colourCount))+
  geom_histogram(binwidth = 5, position = 'stack', color = "black") +
  labs(title = "Percent Fine Substrate Distribution")
  # theme(legend.text = element_text(size = 5)) +
  # guides(fill = guide_legend(nrow = 40))
```

**Numeric Summary of percent_fine_substrate from 2009 to 2020**
```{r}
summary(clean_data$percent_fine_substrate)
```
**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$'percent_fine_substrate'))/nrow(clean_data), 3)*100` % of values in the `percent_fine_substrate` column are NA.

### Variable:`percent_small_substrate`
```{r}
clean_data %>%
  ggplot(aes(x = percent_small_substrate)) +
  geom_histogram(binwidth = 10, position = 'stack', color = "black") +
  labs(title = "Percent Small Substrate Distribution")
```

**Numeric Summary of percent_small_substrate from 2009 to 2020**
```{r}
summary(clean_data$percent_small_substrate)
```

**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$'percent_small_substrate'))/nrow(clean_data), 3)*100` % of values in the `percent_small_substrate` column are NA.

### Variable:`percent_medium_substrate`
```{r}
clean_data %>%
  ggplot(aes(x = percent_medium_substrate)) +
  geom_histogram(binwidth = 10, position = 'stack', color = "black") +
  labs(title = "Percent Medium Substrate Distribution")
```

**Numeric Summary of percent_medium_substrate from 2009 to 2020**
```{r}
summary(clean_data$percent_medium_substrate)
```
**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$'percent_medium_substrate'))/nrow(clean_data), 3)*100` % of values in the `percent_medium_substrate` column are NA.

### Variable:`percent_large_substrate`
```{r}
clean_data %>%
  ggplot(aes(x = percent_large_substrate)) +
  geom_histogram(binwidth = 10, position = 'stack', color = "black") +
  labs(title = "Percent Large Substrate Distribution")
```

**Numeric Summary of percent_large_substrate from 2009 to 2020**
```{r}
summary(clean_data$percent_large_substrate)
```
**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$'percent_large_substrate'))/nrow(clean_data), 3)*100` % of values in the `percent_large_substrate` column are NA.

### Variable:`percent_boulder`
```{r}
clean_data %>%
  ggplot(aes(x = percent_boulder)) +
  geom_histogram(binwidth = 7, position = 'stack', color = "black") +
  labs(title = "Percent Boulder Distribution")
```

**Numeric Summary of percent_boulder from 2009 to 2020**
```{r}
summary(clean_data$percent_boulder)
```

**NA and Unknown Values**  

* `r round(sum(is.na(clean_data$'percent_boulder'))/nrow(clean_data), 3)*100` % of values in the `percent_large_substrate` column are NA.

### Summary of Mean Percent Substrate In Each Location

```{r fig.height = 9, fig.width = 8}
clean_data %>% 
  group_by(location) %>% 
  summarise(mean_percent_fine_substrate = mean(percent_fine_substrate, na.rm = TRUE),
            mean_percent_small_substrate = mean(percent_small_substrate, na.rm = TRUE),
            mean_percent_medium_substrate = mean(percent_medium_substrate, na.rm = TRUE),
            mean_percent_large_substrate = mean(percent_large_substrate, na.rm = TRUE),
            mean_percent_boulder = mean(percent_boulder, na.rm = TRUE),
            ) %>% 
  pivot_longer(
    cols = starts_with("mean"),
    names_to = "substrate_type",
    values_to = "percent",
    values_drop_na = TRUE
  ) %>%
  ggplot(aes(fill = substrate_type,
             y = location,
             x = percent))+
  geom_bar(position = 'stack', stat = 'identity', color = 'black')+
  labs(title = "Mean Percent Substrate by Location")+
  theme(axis.text=element_text(size = 7))
```

### Variable: `depth_m`
```{r}
clean_data %>%
  ggplot(aes(x = depth_m)) +
  geom_histogram(binwidth = 0.15, position = 'stack', color = "black") +
  labs(title = "Depth Distribution")
```

**Numeric Summary of depth_m from 2009 to 2020**
```{r}
summary(clean_data$depth_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(clean_data$'depth_m'))/nrow(clean_data), 3)*100` % of values in the `depth_m` column are NA.


### Variable: `pot_depth_m`
```{r}
clean_data %>%
  filter(pot_depth_m < 40) %>% #removed two values for better view of distribution
  ggplot(aes(x = pot_depth_m)) +
  geom_histogram(binwidth = 0.15, position = 'stack', color = "black") +
  labs(title = "Pot Depth Distribution")
```

**Numeric Summary of pot_depth_m from 2009 to 2020**
```{r}
summary(clean_data$pot_depth_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(clean_data$'pot_depth_m'))/nrow(clean_data), 3)*100` % of values in the `pot_depth_m` column are NA.

### Variable: `velocity`

Standardize by converting to ft/s

```{r}
clean_data$velocity <- clean_data$velocity*3.28084
```

```{r}
clean_data %>%
  ggplot(aes(x = velocity)) +
  geom_histogram(binwidth = 0.2, position = 'stack', color = "black") +
  labs(title = "Velocity Distribution")
```

**Numeric Summary of velocity_m_per_s from 2009 to 2020**
```{r}
summary(clean_data$velocity)
```

**NA and Unknown Values**

* `r round(sum(is.na(clean_data$'velocity'))/nrow(clean_data), 3)*100` % of values in the `velocity` column are NA.

### Variable: `latitude and longitude`

**Numeric Summary of latitude from 2009 to 2020**

```{r}
summary(clean_data$latitude)
```

**Numeric Summary of longitude from 2009 to 2020**
```{r}
summary(clean_data$longitude)
```


### Notes and Issues:

* latitude and longitude were not converted to WGS84
* 2012,2013,2015 latitude and longitude shows inconsistencies and large variations (refer to specific year Markdown for more info)

## Next steps

- Most important variables for the JPE are `date`, `latitude`, `longitude`, `type`, `redd_count`, `salmon_count`

### Add cleaned data back onto google cloud
```{r}
feather_redd_survey <- clean_data %>% glimpse()
```
```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(feather_redd_survey,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/feather-river/data/feather_redd.csv")
```