---
title: "RST QA Ranges"
author: "Erin Cain"
date: "4/29/2022"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(googleCloudStorageR)
```


## RST QA/QC

## Load in standard RST Datasets

```{r, include=F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_get_object(object_name = "standard-format-data/standard_RST_environmental.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "standard_RST_environmental.csv",
               overwrite = TRUE)
gcs_get_object(object_name = "standard-format-data/standard_rst_trap.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "standard_rst_trap.csv",
               overwrite = TRUE)
gcs_get_object(object_name = "standard-format-data/standard_rst_catch.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "standard_rst_catch.csv",
               overwrite = TRUE)
gcs_get_object(object_name = "standard-format-data/standard_recapture.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "standard_recapture.csv",
               overwrite = TRUE)
gcs_get_object(object_name = "standard-format-data/standard_flow.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "standard_flow.csv",
               overwrite = TRUE)
```
```{r,message=F, warning=F}
environmental <- read_csv(here::here("data", "standard-format-data", "standard_RST_environmental.csv"))
trap_operation <- read_csv(here::here("data", "standard-format-data", "standard_trap.csv"))
catch <- read_csv(here::here("data", "standard-format-data", "standard_catch.csv"))
mark_recapture <- read_csv(here::here("data", "standard-format-data", "standard_recapture.csv"))
flow <- read_csv(here::here("data", "standard-format-data", "standard_flow.csv"))
```

## Explore Environmental Variables {.tabset}
* `flow`-  flow in cfs  
* `temperature` - temperature in degrees C
* `turbidity` - turbidity measure in NTU 

### Flow

#TODO pull in standard flow

Numeric Summary: 
  - shows lots of NAs
- Max much higher than 3rd quartile (likely some outliers)

```{r}
summary(environmental$flow)
```
Box Plot
```{r}
flow |> 
  ggplot(aes(flow_cfs, color = stream)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```

#### Flow ranges vary by watershed {.tabset}: 
  
##### Mill & Deer 
  
```{r}
flow |> 
  filter(stream %in% c("mill creek", "deer creek")) |> 
  ggplot(aes(flow_cfs, color = stream)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 

flow |> 
  filter(stream %in% c("mill creek", "deer creek")) |> 
  ggplot(aes(flow_cfs, fill = stream)) +
  geom_histogram(position = "identity") +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```

##### Sacramento  

```{r}
flow |> 
  filter(stream == "sacramento river") |> 
  ggplot(aes(flow_cfs, color = stream)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 

flow |> 
  filter(stream == "sacramento river") |> 
  ggplot(aes(flow_cfs, fill = stream)) +
  geom_histogram(position = "identity") +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```

##### Battle 
```{r}
flow |> 
  filter(stream == "battle creek") |> 
  ggplot(aes(flow_cfs, color = stream)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 

flow |> 
  filter(stream == "battle creek") |> 
  ggplot(aes(flow_cfs, fill = stream)) +
  geom_histogram(position = "identity") +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```
##### Butte

```{r}
flow |> 
  filter(stream == "butte creek") |> 
  ggplot(aes(flow_cfs, color = stream)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 

flow |> 
  filter(stream == "butte creek") |> 
  ggplot(aes(flow_cfs, fill = stream)) +
  geom_histogram(position = "identity") +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```

##### Clear 

```{r}
flow |> 
  filter(stream == "clear creek") |> 
  ggplot(aes(flow_cfs, color = stream)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 

flow |> 
  filter(stream == "clear creek") |> 
  ggplot(aes(flow_cfs, fill = stream)) +
  geom_histogram(position = "identity") +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```

##### Feather

```{r}
flow |> 
  filter(stream == "feather river") |> 
  ggplot(aes(flow_cfs, color = site)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 

flow |> 
  filter(stream == "feather river") |> 
  ggplot(aes(flow_cfs, fill = site)) +
  geom_histogram(position = "identity", alpha = .5) +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```

##### Yuba

```{r}
flow |> 
  filter(stream == "yuba river") |> 
  ggplot(aes(flow_cfs, color = stream)) +
  geom_boxplot() +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 

flow |> 
  filter(stream == "yuba river") |> 
  ggplot(aes(flow_cfs, fill = stream)) +
  geom_histogram(position = "identity") +
  labs(title = "Flow (cfs)") + 
  theme_minimal() 
```


#### Max Ranges for QA Recomendations

Sacramento River: 200 - 35000 

Mill & Deer Creeks: 50 - 2000 (min value informed by CDEC ranges)

Battle Creek: 50 - 2500

Butte Creek: 50 - 5000

Clear Creek: 50 - 2000

Feather River: 
HFC:300 - 5000
LFC:500 - 1000 (tightly regulated so should be a fairly small range)

Yuba River: 50 - 3500 (values informed by plots and USGS statistics)


### Temperature

Numeric Summary: 
  - shows lots of NAs

```{r}
summary(environmental$temperature)
```
Box Plot
```{r}
environmental |>
  ggplot(aes(temperature, color = stream)) +
  geom_boxplot() +
  labs(title = "Temperature (C)") + 
  theme_minimal() 
```

All streams fall within a much more reasonable range of each other 

#### Max Range for QA Recomendations

Temperature Range All Streams: > 0 - 30

### Turbidity

Numeric Summary: 
- shows lots of NAs
- Max much higher than 3rd quartile (likely some outliers)

```{r}
summary(environmental$turbidity)
```

Box Plot

```{r}
environmental |>
  ggplot(aes(turbidity, color = stream)) +
  geom_boxplot() +
  labs(title = "Turbidity (NTU)") + 
  theme_minimal() 
```

```{r}
environmental |>
  filter(turbidity < 1000) |> 
  ggplot(aes(turbidity, color = stream)) +
  geom_boxplot() +
  labs(title = "Turbidity (NTU)") + 
  theme_minimal() 
```

#### Max Range for QA Recomendations

Turbidity Range All Streams: 0 - 1000 

(only expect values greater than 25 in runoff events, put 1000 as the upper end of the turbidy measure because most turbidity sensors only accurately go to 1000)

We may want to end up linking turbidity check with flow (You don't usually get this high of a turbidity with x flow...)

## Explore Trap Operation Variables {.tabset}

* `sample_period_revolutions` - total number of revolutions during a sample period
* `debris_volume` - volume of debris
* `rpms_start`- revolutions per minute at start
* `rpms_end` - revolutions per minute at end


### Total Revolutions 

Numeric Summary: 
- shows lots of NAs
- Max much higher than 3rd quartile (likely some outliers)

```{r}
summary(trap_operation$sample_period_revolutions)
```

Box Plot

```{r}
trap_operation |>
  ggplot(aes(sample_period_revolutions, color = stream)) +
  geom_boxplot() +
  labs(title = "Total Revolutions") + 
  theme_minimal() 
```

There are a few big outliers on Yuba. 

Let's narrow down the range a bit

```{r}
trap_operation |>
  filter(sample_period_revolutions < 25000) |># filters out high counter values on clear, battle, and yuba river - comments for these instances describe that counters were tampered with and not accurate on these days or that the were not reset the day before
  ggplot(aes(sample_period_revolutions, color = stream)) +
  geom_boxplot() +
  labs(title = "Total Revolutions") + 
  theme_minimal() 
```

#### Max Range for QA Recomendations

Total Revolutions max range should be >0 - 25000, if 0 something must be wrong with trap. 

Most revolutions we would expect < 12500 but can get bigger ones especially if it took longer than 24 hours to check the trap                                                             


### RPMs (start and end)

Numeric Summary: 
  - shows lots of NAs
- Max much higher than 3rd quartile (likely some outliers)

```{r}
summary(trap_operation$rpms_start)
summary(trap_operation$rpms_end)
```

Box Plot

```{r}
trap_operation |>
  filter(rpms_start < 3000) |> # filter out 2 values above 3000 on butte
  ggplot(aes(rpms_start, color = stream)) +
  geom_boxplot() +
  labs(title = "RPMs Start") + 
  theme_minimal() 
```

Still a couple of outliers on Butte Creek 

Let's look at RPM End - we are seeing similar distributions 

```{r}
trap_operation |>
  filter(rpms_end < 1000) |>  # filter out 3 values above 1000
  ggplot(aes(rpms_end, color = stream)) +
  geom_boxplot() +
  labs(title = "RPMs end") + 
  theme_minimal() 
```

#### Max Range for QA Recomendations

RPM Start: 0 - 30

RPM End: 0 - 30

### Debris Volumne  

Numeric Summary: 
- shows lots of NAs
- Max much higher than 3rd quartile (likely some outliers)

```{r}
summary(trap_operation$debris_volume)
```

Box Plot

```{r}
trap_operation |>
  ggplot(aes(debris_volume, color = stream)) +
  geom_boxplot() +
  labs(title = "Debris Volumne") + 
  theme_minimal() 
```

#### Max Range for QA Recomendations

Debris Volume 0 - 500 (gallons, will likely convert to L)


## Explore Catch Variables {.tabset}
* `fork_length` - fork length (mm)
* `weight`- observed fish weight

### Fork Length

Numeric Summary: 
- shows lots of NAs
- Max much higher than 3rd quartile (likely some outliers)

```{r}
summary(catch$fork_length)
```

Box Plot

```{r}
catch %>%
  ggplot(aes(fork_length, color = lifestage)) +
  geom_boxplot() +
  labs(title="Fork Lengths") + 
  theme_minimal()
```

It is clear that adult fish are much larger than Juveniles and can be over 1000 millimeter long. We will want to distinguish between lifestages when applying QA to data entry. 

Lets look at summaries of the data by lifestage: 

```{r}
catch |> 
  group_by(lifestage) |> 
  summarize(min = min(fork_length, na.rm = T),
            max = max(fork_length, na.rm = T),
            mean = mean(fork_length, na.rm = T),
            median = median(fork_length, na.rm = T))
```

#### Max Range for QA Recomendations

Adult: 80 - 1000
Juv: >0 - 100 (is this too restrictive or are all the outliers errors)

### Weight

Numeric Summary: 
- shows lots of NAs
- Max much higher than 3rd quartile (likely some outliers)

```{r}
summary(catch$weight)
```

Box Plot

```{r}
catch %>%
  filter(weight < 1000) |>  # Filter out 2 values above 1000
  ggplot(aes(weight, color = lifestage)) +
  geom_boxplot() +
  labs(title = "Weight") + 
  theme_minimal()
```


#### Max Range for QA Recomendations

Adult Range: 50 - 400 grams 
Juvenile: 0 - 50 grams
Most will fall below 50 grams but adult fish can be bigger

