---
title: "QC Site Names"
author: "Erin Cain"
date: "2022-08-18"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(lubridate)
library(googleCloudStorageR)
library(tidyverse)
root.dir <- rprojroot::find_rstudio_root_file()
knitr::opts_knit$set(root.dir)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval = F}
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

# QC Checks {.tabset}

This markdown goes through all standard format datasets to QC and review site names to ensure that all names are consistant across datasets. 

## Standard flow

Updated yub to yuba river. yub is a subsite of yuba river not a site.

```{r}
flow_raw <- read_csv(here::here("data","standard-format-data", "standard_flow.csv")) %>% glimpse

flow_raw  %>%  pull(site)  %>%  unique()
```


## Standard catch

Updated standard catch battle site to "ubc"

```{r}
catch_raw <- read_csv(here::here("data","standard-format-data", "standard_catch.csv")) %>% glimpse

catch_raw  %>%  pull(site)  %>%  unique()
```

Catch data is only reported at the site level not subsite. 

## Standard trap

Updated standard trap battle site to "ubc"
No tisdale site for standard trap  
All other sites are consistant with standard catch

```{r}
trap_raw <- read_csv(here::here("data","standard-format-data", "standard_trap.csv")) %>% glimpse

trap_raw  %>%  pull(site)  %>%  unique()
```

Trap data is collected at the sub site level: 
```{r}
trap_raw  %>%  pull(subsite)  %>%  unique()

```

## Standard environmental

No tisdale site for environmental. 

```{r}
environment_raw <- read_csv(here::here("data","standard-format-data", "standard_environmental.csv")) %>% glimpse

environment_raw  %>%  pull(site)  %>%  unique()
```

Environmental data is collected at the sub site level: 
```{r}
environment_raw  %>%  pull(subsite)  %>%  unique()
```

## Standard recapture

No site for recapture - must join with release 

```{r}
recapture_raw <- read_csv(here::here("data","standard-format-data", "standard_recapture.csv")) %>% glimpse

```
## Standard release

There are 2 NA values for site in the standard release table - otherwise sites match up with sites used in standard catch 

```{r}
release_raw <- read_csv(here::here("data","standard-format-data", "standard_release.csv")) %>% glimpse

release_raw  %>%  pull(site)  %>%  unique()

# 2 NA values for site in release raw
filter(release_raw, is.na(site))
```

