---
title: "RST QA/QC Ranges"
author: "Erin Cain"
date: "4/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(googleCloudStorageR)
```


## RST QA/QC

Add description of QA/QC methods

## Load in combined RST Datasets {.tabset}

### Import Data-sets
```{r}
# Access Data from Google Cloud (Download and Import)
# Link : <https://console.cloud.google.com/storage/browser/jpe-dev-bucket/standard-format-data;tab=[…]iew=project&prefix=&forceOnObjectsSortingFiltering=false>

# Import Data-sets
environmental <- read_csv("standard_RST_environmental.csv")
trap_operation <- read_csv("standard-format-data_standard_rst_trap.csv")
catch <- read_csv("standard-format-data_standard_rst_catch.csv")
mark_recapture <- read_csv("standard-format-data_standard_recapture.csv")
```


## Explore Catch Ranges {.tabset}

Description of what variables we are defining ranges for. 

* `fork_length` - 
* `count` - 
* `weight` - 

### Variables
```{r}
# Identify Numerical Columns
catch_variables %>% glimpse()

# Filter for numerical columns
catch_variables <- catch %>%
  select(c(fork_length,count,weight))

# HELPER FUNCTION: Get Ranges of a column
find_range <- function(data) {
  # Get the min and max value
  result <- range(data, na.rm = TRUE)
}

# Get Ranges
catch_ranges <- as.data.frame(apply(catch_variables, 2, find_range)) # get ranges
rownames(catch_ranges) <- c("min","max") # labels
catch_ranges

```

## Explore Environmental Ranges {.tabset}
Description of what variables we are defining ranges for. 

* `turbidity` - 
* `flow` -
* `secchi` - 

### Variables
```{r}
# Identify Numerical Columns
environmental %>%
  glimpse()

# Filter for numerical columns
environmental_variables <- environmental %>%
  select(c(turbidity, flow, secchi))

# Get Ranges
environmental_ranges <- as.data.frame(apply(environmental_variables, 2, find_range)) # get ranges
rownames(environmental_ranges) <- c("min","max") # labels
catch_ranges
```

## Explore Trap Operation Ranges {.tabset}
Description of what variables we are defining ranges for. 

* `sample_period_revolutions` - 
* `cone_setting,river_left_depth` -
* `river_center_depth` - 
* `river_right_depth` - 
* `depth_adjust` -
* `avg_time_per_rev` -
* `debris_volume` - 
* `rpms_start` -
* `rpms_end` - 
* `cone_rpm` -

### Variables
```{r}
# Identify Numerical Columns
trap_operation %>%
  glimpse()

# Filter for numerical columns
trap_operation_variables <- trap_operation %>%
  select(c(sample_period_revolutions, cone_setting,river_left_depth,
           river_center_depth, river_right_depth, depth_adjust, avg_time_per_rev,
           debris_volume, rpms_start, rpms_end, cone_rpm))

# Get Ranges
trap_operation_ranges <- as.data.frame(apply(trap_operation_variables, 2, find_range)) # get ranges
rownames(trap_operation_ranges) <- c("min","max") # labels
trap_operation_ranges
```

## Explore Mark Recapture Ranges {.tabset}
Description of what variables we are defining ranges for. 

* `number_recaptured` -
* `median_fork_length_recaptured` - 

### Variables
```{r}
# Identify Numerical Columns
mark_recapture %>%
  glimpse()

# Filter for numerical columns
mark_recapture_variables <- mark_recapture %>%
  select(c(number_recaptured, median_fork_length_recaptured))

# Get Ranges
mark_recapture_ranges <- as.data.frame(apply(mark_recapture_variables, 2, find_range)) # get ranges
rownames(mark_recapture_ranges) <- c("min","max") # labels
mark_recapture_ranges
```