---
title: "RST QA/QC Ranges"
author: "Erin Cain"
date: "4/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(googleCloudStorageR)
```


## RST QA/QC

## Load in combined RST Datasets {.tabset}

### Import Data-sets
```{r}
# Access Data from Google Cloud (Download and Import)
# Link : <https://console.cloud.google.com/storage/browser/jpe-dev-bucket/standard-format-data;tab=[…]iew=project&prefix=&forceOnObjectsSortingFiltering=false>

# Google Cloud Set up
# Sys.setenv("GCS_AUTH_FILE" = "config.json")
# Sys.setenv("GCS_DEFAULT_BUCKET" = "jpe-dev-bucket")
# gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
# 
# gcs_get_object(object_name = "standard-format-data/standard_RST_environmental.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "standard_RST_environmental.csv",
#                overwrite = TRUE)
# gcs_get_object(object_name = "standard-format-data/standard_rst_trap.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "standard_rst_trap.csv",
#                overwrite = TRUE)
# gcs_get_object(object_name = "standard-format-data/standard_rst_catch.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "standard_rst_catch.csv",
#                overwrite = TRUE)
# gcs_get_object(object_name = "standard-format-data/standard_recapture.csv",
#                bucket = gcs_get_global_bucket(),
#                saveToDisk = "standard_recapture.csv",
#                overwrite = TRUE)

# Import Data-sets
environmental <- read_csv("standard_RST_environmental.csv")
trap_operation <- read_csv("standard_rst_trap.csv")
catch <- read_csv("standard_rst_catch.csv")
mark_recapture <- read_csv("standard_recapture.csv")
```


## Explore Catch Ranges
* `fork_length` - fork length (mm)
* `count` - number of fish observed
* `weight`- observed fish weight

```{r}
# Identify Numerical Columns
catch %>% glimpse()

# Filter for numerical columns
catch_variables <- catch %>%
  select(c(fork_length,count,weight))

# HELPER FUNCTION: Get Ranges of a column
find_range <- function(data) {
  # Get the min and max value
  result <- range(data, na.rm = TRUE)
}

# Get Ranges
catch_ranges <- as.data.frame(apply(catch_variables, 2, find_range)) # get ranges
rownames(catch_ranges) <- c("min","max") # labels
catch_ranges
```

### Boxplots
```{r}
catch_variables %>%
  ggplot(aes(fork_length)) +
  geom_boxplot() +
  labs(title="Fork Lengths")

catch_variables %>%
  ggplot(aes(count)) +
  geom_boxplot() +
  labs(title="Count")

catch_variables %>%
  ggplot(aes(weight)) +
  geom_boxplot() +
  labs(title="weight")
```

### Defining Ranges
```{r}
col1 = summary(catch_variables$fork_length)
col2 = summary(catch_variables$count)
col3 = summary(catch_variables$weight)
catch_5NumSum = matrix(c(col1,col2,col3),ncol=3)
rownames(catch_5NumSum) = c("Min","1st Qu.","Median","Mean","3rd Qu.","Max","NA's")
colnames(catch_5NumSum) = c("fork_length","count","weight")
catch_5NumSum
```

```{r}
catch_variables %>%
  filter(fork_length > 10) %>%
  ggplot(aes(fork_length)) +
  geom_boxplot() +
  labs(x="Fork Length",y="Length",title="Fork Lengths")
```

## Explore Environmental Ranges {.tabset}
* `turbidity` - turbidity measure in NTU 
* `flow`-  flow in cfs  
* `secchi` - secchi measure in ft 

### Variables
```{r}
# Identify Numerical Columns
environmental %>%
  glimpse()

# Filter for numerical columns
environmental_variables <- environmental %>%
  select(c(turbidity, flow, secchi))

# Get Ranges
environmental_ranges <- as.data.frame(apply(environmental_variables, 2, find_range)) # get ranges
rownames(environmental_ranges) <- c("min","max") # labels
catch_ranges
```

## Explore Trap Operation Ranges
Capture juvenile fish during out migration. During efficiency trials, the sample of fish are caught and later released to determine how effective the trap is at capturing the fish. Traps are sometimes measured with rotational as a function of flow going through trap and its affect on velocity.

# TODO: revolution and flow. IF flow is 10,000 then revloution could be. Pair ranging.Reasonal numbers might be rotary screwshop and rotary number on google. Box plots will help us narrow.

* `sample_period_revolutions` - number of revolutions on cone at given point of time
* `river_left_depth`- 
* `river_center_depth` 
* `river_right_depth`
* `depth_adjust`
* `avg_time_per_rev` 
* `debris_volume` 
* `rpms_start`- revolutions per minute at start
* `rpms_end` - revolutions per minute at end
* `cone_rpm`- 

### Variables
```{r}
# Identify Numerical Columns
trap_operation %>%
  glimpse()

# Filter for numerical columns
trap_operation_variables <- trap_operation %>%
  select(c(sample_period_revolutions,river_left_depth,
           river_center_depth, river_right_depth, depth_adjust, avg_time_per_rev,
           debris_volume, rpms_start, rpms_end, cone_rpm))

# Get Ranges
trap_operation_ranges <- as.data.frame(apply(trap_operation_variables, 2, find_range)) # get ranges
rownames(trap_operation_ranges) <- c("min","max") # labels
trap_operation_ranges
```

## Explore Mark Recapture Ranges 
Description of what variables we are defining ranges for. 

* `number_recaptured` - count of observed fish 
* `median_fork_length_recaptured` - average fork length

### Variables
```{r}
# Identify Numerical Columns
mark_recapture %>%
  glimpse()

# Filter for numerical columns
mark_recapture_variables <- mark_recapture %>%
  select(c(number_recaptured, median_fork_length_recaptured))

# Get Ranges
mark_recapture_ranges <- as.data.frame(apply(mark_recapture_variables, 2, find_range)) # get ranges
rownames(mark_recapture_ranges) <- c("min","max") # labels
mark_recapture_ranges
```