---
title: "pull flow data for trap locations"
output: 
  html_document:
  theme: flatly
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(CDECRetrieve)
library(dataRetrieval)
library(hms)
```

# Flow gage lookup

```{r}
# We need flow data for each of the sites included in the trap_locations table
# for all of the years listed.
# We do not need to go to subsite scale of resolution.
# For now this table is for reference to find the years that data is needed.
# This table is also helpful for understanding years where we need flow data for each location:
# https://github.com/FlowWest/JPE-datasets/tree/main/data-raw/qc-markdowns/rst

# trap_locations_raw <- read_csv("data-raw/standard-format-data-prep/rst_sites.csv")
# trap_locations_raw <- read_csv("rst_sites.csv")
# 
# trap_locations <- trap_locations_raw %>%
#   select(tributary, site_name, river_location, latitude, longitude, year) %>%
#   distinct()
  # identify years

gage_locations_raw <- read_csv("data-raw/standard-format-data-prep/gage_sites.csv")

# Pull flow data from each of the gage sites here
gage_locations <- gage_locations_raw %>%
  filter(flow_gage == T) %>%
  select(tributary, site_name, agency, latitude, longitude, identifier) %>%
  distinct() %>%
  rename(gage_latitude = latitude,
         gage_longitude = longitude) %>%
  glimpse()
  # get codes
```

# Pull in flow data {.tabset}

## Battle Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("BAT") 

BAT_CDEC <- cdec_query(station = "BAT", dur_code = "H", sensor_num = "20", start_date = "2003-01-01")

# Format data
BAT_hourly_flows <- BAT_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
BAT_USGS <- readNWISdv(11376550, "00060")

# Format to make tidier
BAT_daily_flows <- BAT_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 2003) %>%
  as_tibble() %>% 
  rename(date = Date)
```
#### QA/QC

```{r}
# plots to check the quality of data. no data below zero.
# TODO from ashley - curious what the -9998 means. I am guessing this is a code for something, maybe missing data.
ggplot(BAT_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()

# Adds plots for USGS flows too
ggplot(BAT_daily_flows, aes(x = date, y = flow_cfs)) +
  geom_line()
```

```{r}
# Clean negative values
BAT_hourly_flows_clean <- BAT_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

BAT_daily_flows_clean <- BAT_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

# check to make sure the cleaning worked for hourly flows. there was no issue for daily
ggplot(BAT_hourly_flows_clean, aes(x = date, y = flow_cfs)) +
  geom_line()

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
BAT_combined = BAT_hourly_flows_clean %>% 
  full_join(BAT_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(BAT_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

# from ashley - the way you did it works! just wanted to mention that you don't 
# need to combine data before plotting. example code below
# ggplot(BAT_daily_flows, aes(x = date, y = flow_cfs)) +
#   geom_line(color = "steelblue") +
#   geom_line(data = BAT_hourly_flows_clean, aes(x = date, y = flow_cfs), color = "darkred", alpha = 0.5)

```

#### USGS vs CDEC Mean data

```{r}
BAT_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

### Recommendations

Battle Creek : recommend using USGS because CDEC had a lot of NA (-9998) values.

```{r}
saveRDS(BAT_daily_flows_clean, "data-raw/standard-format-data-prep/flow_data/battle_flow.rds")
```

## Butte Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("BCK") 

BCK_CDEC <- cdec_query(station = "BCK", dur_code = "H", sensor_num = "20", start_date = "1995-01-01")

# Format data
BCK_hourly_flows <- BCK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
BCK_USGS <- readNWISdv(11390000, "00060")

# Format to make tidier
BCK_daily_flows <- BCK_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}

ggplot(BCK_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()

# Adds plots for USGS flows too
ggplot(BCK_daily_flows, aes(x = date, y = flow_cfs)) +
  geom_line()

# Clean negative values
BCK_hourly_flows_clean <- BCK_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

BCK_daily_flows_clean <- BCK_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
BCK_combined = BCK_hourly_flows_clean %>% 
  full_join(BCK_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(BCK_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
BCK_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
### Recommendations

Buttle Creek : Recommended to use CDEC to not lose significant fluctuations

```{r}

BCK_hourly_flows_clean_mean <- BCK_hourly_flows_clean %>%
  group_by(date) %>%
  summarize(flow_cfs = mean(flow_cfs, na.rm = T))
saveRDS(BCK_hourly_flows_clean_mean, "data-raw/standard-format-data-prep/flow_data/butte_flow.rds")
```

## Clear Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)
 
# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("IGO") 

IGO_CDEC <- cdec_query(station = "IGO", dur_code = "H", sensor_num = "20", start_date = "2003-01-01")

# Format data
IGO_hourly_flows <- IGO_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
IGO_USGS <- readNWISdv(11372000, "00060")

# Format to make tidier
IGO_daily_flows <- IGO_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 2003) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}

ggplot(IGO_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()

# Clean negative values
IGO_hourly_flows_clean <- IGO_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

IGO_daily_flows_clean <- IGO_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
IGO_combined = IGO_hourly_flows_clean %>% 
  full_join(IGO_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(IGO_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
IGO_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
### Recommendations

Clear Creek : recommend using USGS because CDEC had a lot of NA (-9998) values.

```{r}
saveRDS(IGO_daily_flows_clean, "data-raw/standard-format-data-prep/flow_data/clear_flow.rds")
```

## Deer Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both

cdec_datasets("DCV") 

DCV_CDEC <- cdec_query(station = "DCV", dur_code = "H", sensor_num = "20", start_date = "1995-01-01")

# Format data
DCV_hourly_flows <- DCV_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
  
```

### USGS data

```{r}
DCV_USGS <- readNWISdv(11383500, "00060")

# Format to make tidier
DCV_daily_flows <- DCV_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
ggplot(DCV_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
# Clean negative values
DCV_hourly_flows_clean <- DCV_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

DCV_daily_flows_clean <- DCV_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
DCV_combined = DCV_hourly_flows_clean %>% 
  full_join(DCV_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(DCV_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

#### USGS vs CDEC Mean data

```{r}
DCV_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS, na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
### Recommendations

Deer Creek : recommend using USGS because CDEC had a lot of NA (-9998) values.

```{r}
saveRDS(DCV_daily_flows_clean, "data-raw/standard-format-data-prep/flow_data/deer_flow.rds")
```

## Feather River

### High flow channel sites

```{r}
# We already identified the code for the CDEC gage of interest here: GRL
# Use this code to see what data are available
cdec_datasets("GRL") 
# For reference - this mapping tool shows CDEC sites: https://cdec.water.ca.gov/webgis/?appid=cdecstation

# Use this code to pull the data from CDEC. 
gridley_CDEC <- cdec_query(station = "GRL", dur_code = "H", sensor_num = "20", start_date = "1996-01-01")

# Format data
gridley_hourly_flows <- gridley_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
  
```

#### QA/QC 

It looks like there are some negative values. We should make these NA.

It would be important to note any gaps in data collection.

```{r}
ggplot(gridley_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
```
#### Clean data

```{r}
gridley_hourly_flows_clean <- gridley_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

gridley_hourly_flows_clean_mean <- gridley_hourly_flows_clean %>%
  group_by(date) %>%
  summarize(flow_cfs = mean(flow_cfs, na.rm = T))
```

### Low flow channel sites

```{r}
# For reference - this mapping tool shows USGS sites: https://maps.waterdata.usgs.gov/mapper/index.html
# You can use the mapper to search for sites and find the right parameter code. I use 0060 below for daily discharge. 
# Hourly or instaneous would be better but it is not available.
# Documentation for dataRetrieval: https://github.com/USGS-R/dataRetrieval

oroville_USGS <- readNWISdv(11407000, "00060")

# Format to make tidier
oroville_daily_flows <- oroville_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1996) %>%
  as_tibble() %>% 
  rename(date = Date)
```


#### QA/QC

```{r}
ggplot(oroville_daily_flows, aes(x = date, y = flow_cfs)) +
  geom_line()
# just checking to make sure there aren't any zeros
filter(oroville_daily_flows, flow_cfs == 0)
filter(oroville_daily_flows, is.na(flow_cfs))
min(oroville_daily_flows$flow_cfs)
```

### Lower Feather site

```{r}
# Find CDEC gage nearby the Lower Feather RST site
# Use this code to see what data are available
# cdec_datasets("FSB") 
# For reference - this mapping tool shows CDEC sites: https://cdec.water.ca.gov/webgis/?appid=cdecstation

# Use this code to pull the data from CDEC. 
boyds_landing_CDEC <- cdec_query(station = "FSB", dur_code = "H", 
                                 sensor_num = "20", start_date = "2010-01-01")

# Format data
boyds_landing_hourly_flows <- boyds_landing_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
  
```


#### QA/QC

```{r}
ggplot(boyds_landing_hourly_flows, aes(x = date, y = flow_cfs)) +
  geom_line()
# just checking to make sure there aren't any zeros
boyds_landing_hourly_flows |> filter(flow_cfs == 0) |> nrow()
boyds_landing_hourly_flows |> filter(is.na(flow_cfs)) |> nrow()
boyds_landing_hourly_flows |> filter(flow_cfs < 0) |> nrow()
boyds_landing_hourly_flows |> nrow()
```

#### Clean data

```{r}
boyds_landing_mean_daily_flows <- boyds_landing_hourly_flows |> 
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs)) |> 
  filter(!is.na(flow_cfs), !is.na(date)) |> 
  group_by(date) |> 
  summarize(flow_cfs = mean(flow_cfs, na.rm = T)) |> 
  glimpse()
```

### Recommendations

There are three major sites on the Feather River - High Flow Channel, Low Flow Channel, and the Lower Feather RST site. The Lower Feather RST site does not collect discharge data. 

We will use the CDEC GRL gage for HFC, the USGS Oroville gage for LFC, and the CDEC FSB gage for the Lower Feather.

```{r}
saveRDS(gridley_hourly_flows_clean_mean, "data-raw/standard-format-data-prep/flow_data/feather_hfc_flow.rds")
saveRDS(oroville_daily_flows, "data-raw/standard-format-data-prep/flow_data/feather_lfc_flow.rds")
saveRDS(boyds_landing_mean_daily_flows, 
        "data-raw/standard-format-data-prep/flow_data_prep_files/flow_data/lower_feather_flow.rds")
```

## Mill Creek

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both

cdec_datasets("MLM") 

MLM_CDEC <- cdec_query(station = "MLM", dur_code = "H", sensor_num = "20", start_date = "1996-01-01")

# Format data
MLM_hourly_flows <- MLM_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
  
```

### USGS data

```{r}
MLM_USGS <- readNWISdv(11381500, "00060")

# Format to make tidier
MLM_daily_flows <- MLM_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1996) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
ggplot(MLM_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
# Clean negative values
MLM_hourly_flows_clean <- MLM_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

MLM_daily_flows_clean <- MLM_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
MLM_combined = MLM_hourly_flows_clean %>% 
  full_join(MLM_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(MLM_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
MLM_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

### Recommendations

Mill Creek: Recommend to use USGS because CDEC has no available data from 2008-2020.

```{r}
saveRDS(MLM_daily_flows_clean, "data-raw/standard-format-data-prep/flow_data/mill_flow.rds")
```

## Yuba River

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("MRY") 

MRY_CDEC <- cdec_query(station = "MRY", dur_code = "H", sensor_num = "20", start_date = "1999-01-01")

# Format data
MRY_hourly_flows <- MRY_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
MRY_USGS <- readNWISdv(11421000, "00060")

# Format to make tidier
MRY_daily_flows <- MRY_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1999) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
ggplot(MRY_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
# Clean negative values
MRY_hourly_flows_clean <- MRY_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

MRY_daily_flows_clean <- MRY_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
MRY_combined = MRY_hourly_flows_clean %>% 
  full_join(MRY_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(MRY_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
#### USGS vs CDEC Mean data

```{r}
MRY_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
### Recommendations

Yuba : Recommend to use USGS because CDEC has no available data from 2003-2020.

```{r}
saveRDS(MRY_daily_flows_clean, "data-raw/standard-format-data-prep/flow_data/yuba_flow.rds")
```

## Sacramento River - Knights Landing

### CDEC data

```{r}
# Determine what data are available for the CDEC gauge vs the USGS gauge and how that
# matches with the years needed (reference trap_locations table for years)

# If both gauges offer similar data, is one at a finer scale (hourly vs daily)?

# Choose the gauge with the most data at the finest scale. If years don't overlap, could use
# combination of both
cdec_datasets("WLK") 

WLK_CDEC <- cdec_query(station = "WLK", dur_code = "H", sensor_num = "20", start_date = "1995-01-01")

# Format data
WLK_hourly_flows <- WLK_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)
```

### USGS data

```{r}
WLK_USGS <- readNWISdv(11390500, "00060")

# Format to make tidier
WLK_daily_flows <- WLK_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)
```

#### QA/QC

```{r}
ggplot(WLK_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
# Clean negative values
WLK_hourly_flows_clean <- WLK_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

WLK_daily_flows_clean <- WLK_daily_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

```

#### USGS vs CDEC data

```{r}
# comparison of USGS vs CDEC
# Does one have more data?
# Does one have a finer scale?
WLK_combined = WLK_hourly_flows_clean %>% 
  full_join(WLK_daily_flows_clean, by ="date") %>% 
  rename("flow_cfs.CDEC"="flow_cfs.x", "flow_cfs.USGS"="flow_cfs.y")

ggplot(WLK_combined, aes(x=date)) +
  geom_line(aes(y = flow_cfs.CDEC), color ="darkred") +
  geom_line(aes(y = flow_cfs.USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```

#### USGS vs CDEC Mean data

```{r}
WLK_combined %>% 
  group_by(date = date(datetime)) %>% 
  group_by(date) %>% 
  summarise(
    flow_CDEC = mean(flow_cfs.CDEC, na.rm = TRUE),
    flow_USGS = max(flow_cfs.USGS,na.rm = TRUE)) %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y = flow_CDEC), color ="darkred") +
  geom_line(aes(y = flow_USGS), color ="steelblue", linetype ="twodash") +
  labs(y="flow_cfs")

```
### Recommendations

Sacramento River- Knights Landing :  Use USGS because there are a few gaps where there is not data for CDEC

```{r}
saveRDS(WLK_daily_flows_clean, "data-raw/standard-format-data-prep/flow_data/knights_flow.rds")
```


## Sacramento River - Tisdale

### CDEC data

```{r}
# Only CDEC data is available for Tisdale
cdec_datasets("TIS") 

TIS_CDEC <- cdec_query(station = "TIS", dur_code = "H", sensor_num = "20", start_date = "2010-01-01")

# Format data
TIS_hourly_flows <- TIS_CDEC %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)

```

#### QA/QC

```{r}
# plots to check the quality of data. no data below zero.
ggplot(TIS_hourly_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()
```

```{r}
TIS_hourly_flows_clean <- TIS_hourly_flows %>%
  mutate(flow_cfs = ifelse(flow_cfs < 0, NA_real_, flow_cfs))

ggplot(TIS_hourly_flows_clean, aes(x = datetime, y = flow_cfs)) +
  geom_line()

```

There are a lot of gaps in this data. Try pulling less fine data to see if any better.
The event data is not any better. I looked at [map](https://flowwest.shinyapps.io/jpe-eda-app/)
and the USGS gage for Knights Landing is not that far from Tisdale. It would better to
use that data.

```{r}
TIS_CDEC_event <- cdec_query(station = "TIS", dur_code = "E", sensor_num = "20", start_date = "2010-01-01")

# Format data
TIS_event_flows <- TIS_CDEC_event %>%
  mutate(date = as_date(datetime),
         time = as_hms(datetime),
         flow_cfs = parameter_value)

ggplot(TIS_event_flows, aes(x = datetime, y = flow_cfs)) +
  geom_line()

```

## Recommendations

The CDEC data for Tisdale is really sparse. Recommend using the USGS gage for Knights
because it is very close.

```{r}
saveRDS(WLK_daily_flows_clean, "data-raw/standard-format-data-prep/flow_data/tisdale_flow.rds")
```


## Red Bluff

```{r}
RB_USGS <- readNWISdv(11377100, "00060")
# Format to make tidier
RB_daily_flows <- RB_USGS %>%
  select(Date, flow_cfs =  X_00060_00003) %>%
  filter(lubridate::year(Date) >= 1995) %>%
  as_tibble() %>% 
  rename(date = Date)
```

### QA/QC

Data looks pretty good 

```{r}
ggplot(RB_daily_flows, aes(x = date, y = flow_cfs)) +
  geom_line()
```

## Recommendations

USGS data is complete and looks reasonable

```{r}
saveRDS(RB_daily_flows, "data-raw/standard-format-data-prep/flow_data_prep_files/flow_data/rb_flow.rds")
```