---
title: "Lower Sacramento (Knight's Landing) RST data QC - sampling effort"
author: "Ashley Vizek"
date: "10/19/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE, fig.width=15, fig.height=10}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
```

# Lower Sacramento (Knight's Landing) Rotary Screw Trap Data

## Description of Monitoring Data

**Timeframe:** 

10/04/2002 through 05/31/2021

**Completeness of Record throughout timeframe:** 

- The 2013 sampling period is not complete (Oct 2012 through Dec 2012)
- We are missing data for 2017 and 2018. 

**Sampling Location:**

Knight's Landing


**Data Contact:** 

[Jeanine Phillips](mailto::Jeanine.Phillips@wildlife.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()

get_data <- function(name) {
gcs_get_object(object_name = paste0("rst/lower-sac-river/data/knights-landing/knl_combine_", name,".csv"),
               bucket = gcs_get_global_bucket(),
               saveToDisk = paste0("knl_combine_", name, ".csv"),
               overwrite = TRUE)
}

get_data("sampling_effort")
```

Read in data from google cloud, glimpse raw data and domain description sheet.
Need to change date, start_date, stop_date into appropriate format. Make changes
in changes section at end.

```{r}
# read in data to clean 
knl_effort <- read.csv("knl_combine_sampling_effort.csv")
glimpse(knl_effort)
```

## Data transformations

This work was done in other rmd (1_knights_landing_initial_clean.Rmd, 2_knights_landing_combined_data.Rmd)

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables (this way we know we do not miss any)
knl_effort %>%
  select_if(is.numeric) %>%
  colnames()
```

### Variable: `number_traps`

**Plotting `number_traps` over Period of Record**

This variable exists for years 2002-2011 and is consistently 2; except in a few cases.

```{r}
filter(knl_effort, year(date) <= 2011) %>%
ggplot(aes(y = number_traps)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

**Numeric Summary of `number_traps` over Period of Record**

```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = mean(number_traps, na.rm = T),
            median = median(number_traps, na.rm = T),
            min = min(number_traps, na.rm = T),
            max = max(number_traps, na.rm = T),
            na = length(which(is.na(number_traps))))
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_effort$number_traps))/nrow(knl_effort), 3) * 100` % of values in the `number_traps` column are NA.

### Variable: `hrs_fished`

**Plotting `hrs_fished` over Period of Record**


```{r}
# knl_effort %>%
#   mutate(fake_year = 2000, 
#          fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  
  
  ggplot(knl_effort, aes(hrs_fished)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

**Numeric Summary of `hrs_fished` over Period of Record**

```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(hrs_fished, na.rm = T),2),
            median = round(median(hrs_fished, na.rm = T),2),
            min = round(min(hrs_fished, na.rm = T),2),
            max = round(max(hrs_fished, na.rm = T),2),
            na = length(which(is.na(hrs_fished))))
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_effort$hrs_fished))/nrow(knl_effort), 3) * 100` % of values in the `hrs_fished` column are NA.


### Variable: `sampling_period_hrs`

**Plotting `sampling_period_hrs` over Period of Record**

Data for this variable was collected beginning in 2011 (sampling year 2012).

```{r}
ggplot(filter(knl_effort, year(date) >= 2011), aes(sampling_period_hrs)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

**Numeric Summary of `sampling_period_hrs` over Period of Record**

```{r}
filter(knl_effort, year(date) >= 2011) %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(sampling_period_hrs, na.rm = T),2),
            median = round(median(sampling_period_hrs, na.rm = T),2),
            min = round(min(sampling_period_hrs, na.rm = T),2),
            max = round(max(sampling_period_hrs, na.rm = T),2),
            na = length(which(is.na(sampling_period_hrs))))
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_effort$sampling_period_hrs))/nrow(knl_effort), 3) * 100` % of values in the `sampling_period_hrs` column are NA.

### Variable: `flow_cfs`

**Plotting `flow_cfs` over Period of Record**

```{r}
ggplot(knl_effort, aes(y = flow_cfs)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

```{r}
knl_effort %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  ggplot(aes(x = fake_date, y = flow_cfs)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("") +
  facet_wrap(~wy)
```

**Numeric Summary of `flow_cfs` over Period of Record**

```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(flow_cfs, na.rm = T),2),
            median = round(median(flow_cfs, na.rm = T),2),
            min = round(min(flow_cfs, na.rm = T),2),
            max = round(max(flow_cfs, na.rm = T),2),
            na = length(which(is.na(flow_cfs))))
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_effort$flow_cfs))/nrow(knl_effort), 3) * 100` % of values in the `flow_cfs` column are NA.

### Variable: `secchi_ft`

**Plotting `secchi_ft` over Period of Record**

Data for this variable was not collected after 2014.

```{r}
ggplot(filter(knl_effort, year(date) <= 2014), aes(y = secchi_ft)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

```{r}
filter(knl_effort, year(date) <= 2014) %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  ggplot(aes(x = fake_date, y = secchi_ft)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("") +
  facet_wrap(~wy)
```

**Numeric Summary of `secchi_ft` over Period of Record**

```{r}
filter(knl_effort, year(date) <= 2014) %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(secchi_ft, na.rm = T),2),
            median = round(median(secchi_ft, na.rm = T),2),
            min = round(min(secchi_ft, na.rm = T),2),
            max = round(max(secchi_ft, na.rm = T),2),
            na = length(which(is.na(secchi_ft))))
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_effort$secchi_ft))/nrow(knl_effort), 3) * 100` % of values in the `secchi_ft` column are NA.


### Variable: `water_t_f`

**Plotting `water_t_f` over Period of Record**


```{r}
ggplot(knl_effort, aes(y = water_t_f)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

```{r}
knl_effort %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  ggplot(aes(x = fake_date, y = water_t_f)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("") +
  facet_wrap(~wy)
```

**Numeric Summary of `water_t_f` over Period of Record**


```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(water_t_f, na.rm = T),2),
            median = round(median(water_t_f, na.rm = T),2),
            min = round(min(water_t_f, na.rm = T),2),
            max = round(max(water_t_f, na.rm = T),2),
            na = length(which(is.na(water_t_f))))
```

Seems unlikely that water temp would be 0 in 2009. Check to make sure this isn't a typo.
I think this is a typo and should be changed to 70. Made that change to knl_effort_changes 
in change section below.

```{r}
filter(knl_effort, year(date) == 2009, water_t_f == 0)
filter(knl_effort, year(date) == 2009, month(date) == 6)
```

**NA and Unknown Values**

Percent of NA:

`r round(sum(is.na(knl_effort$water_t_f))/nrow(knl_effort), 3) * 100` % of values in the `water_t_f` column are NA.

### Variable: `klci`

**Plotting `klci` over Period of Record**
  
Data for this variable was only collected in 2020 and 2021.
Very few observations that are not 0. I may remove this variable from the combined dataset
because there is so little information.

```{r}
ggplot(knl_effort, aes(y = klci)) +
  geom_histogram()
```

**Numeric Summary of `klci` over Period of Record**
  
```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(klci, na.rm = T),2),
            median = round(median(klci, na.rm = T),2),
            min = round(min(klci, na.rm = T),2),
            max = round(max(klci, na.rm = T),2),
            na = length(which(is.na(klci))))
```

**NA and Unknown Values**
  
  Percent of NA:
  
  `r round(sum(is.na(knl_effort$klci))/nrow(knl_effort), 3) * 100` % of values in the `klci` column are NA.
  
### Variable: `turbidity`

**Plotting `turbidity` over Period of Record**
  
There are two turbidity units - NTU and FTU. FTU was only collected in 2014.
I didn't notice any big differences though between values for NTU v FTU. There
are some cases where turbidity gets to 1,000 which is in the realm of possibility.

```{r}
knl_effort %>%
  mutate(wy = factor(ifelse(month(date) %in% 10:12, year(date) + 1, year(date))),
         fake_year = 2000,
         fake_year = ifelse(month(date) %in% 10:12, fake_year - 1, fake_year),
         fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%
  ggplot(aes(x = fake_date, y = turbidity)) +
  scale_x_date(date_breaks = "3 month", date_labels = "%b") +
  geom_line(size = 0.5) +
  xlab("") +
  facet_wrap(~wy)
```

**Numeric Summary of `turbidity` over Period of Record**
  
```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(turbidity, na.rm = T),2),
            median = round(median(turbidity, na.rm = T),2),
            min = round(min(turbidity, na.rm = T),2),
            max = round(max(turbidity, na.rm = T),2),
            na = length(which(is.na(turbidity))))
```

**NA and Unknown Values**
  
  Percent of NA:
  
  `r round(sum(is.na(knl_effort$turbidity))/nrow(knl_effort), 3) * 100` % of values in the `turbidity` column are NA.
  
### Variable: `cone_id`

**Plotting `cone_id` over Period of Record**
  
Data exists for all years.
This variable should be re-categorized as a character.
Changes are made in the make_changes section below.

```{r}
# knl_effort %>%
#   mutate(fake_year = 2000, 
#          fake_date = ymd(paste(fake_year, month(date), day(date)))) %>%

unique(knl_effort$cone_id)
```


**Numeric Summary of `cone_id` over Period of Record**
  
```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(na = length(which(is.na(cone_id))))
```

**NA and Unknown Values**
  
  Percent of NA:
  
  `r round(sum(is.na(knl_effort$cone_id))/nrow(knl_effort), 3) * 100` % of values in the `cone_id` column are NA.

### Variable: `cone_rpm`

**Plotting `cone_rpm` over Period of Record**

Data collected beginning in 2011 (sampling year 2012).

```{r}
ggplot(filter(knl_effort, year(date) >= 2011), aes(y = cone_rpm)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

**Numeric Summary of `cone_rpm` over Period of Record**
  
```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(cone_rpm, na.rm = T),2),
            median = round(median(cone_rpm, na.rm = T),2),
            min = round(min(cone_rpm, na.rm = T),2),
            max = round(max(cone_rpm, na.rm = T),2),
            na = length(which(is.na(cone_rpm))))
```

**NA and Unknown Values**
  
  Percent of NA:
  
`r round(sum(is.na(knl_effort$cone_rpm))/nrow(knl_effort), 3) * 100` % of values in the `cone_rpm` column are NA.
  
### Variable: `total_cone_rev`

**Plotting `total_cone_rev` over Period of Record**

Data for this variable was collected beginning in 2011 (sampling year 2012).

```{r}
ggplot(filter(knl_effort, year(date) >= 2011), aes(y = total_cone_rev)) +
  geom_histogram() +
  facet_wrap(~year(date))
```

**Numeric Summary of `total_cone_rev` over Period of Record**
  
```{r}
knl_effort %>%
  group_by(year(date)) %>%
  summarize(mean = round(mean(total_cone_rev, na.rm = T),2),
            median = round(median(total_cone_rev, na.rm = T),2),
            min = round(min(total_cone_rev, na.rm = T),2),
            max = round(max(total_cone_rev, na.rm = T),2),
            na = length(which(is.na(total_cone_rev))))
```
It seems like the total_cone_rev at 20,495 is a typo but not enough information to be sure. Will leave as is.

```{r}
filter(knl_effort, total_cone_rev > 20000)
filter(knl_effort, total_cone_rev > 12000)
```

**NA and Unknown Values**
  
  Percent of NA:
  
`r round(sum(is.na(knl_effort$total_cone_rev))/nrow(knl_effort), 3) * 100` % of values in the `total_cone_rev` column are NA.
  
  
## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

Need to change dates to appropriate format!

```{r}
knl_effort %>%
  select_if(is.character) %>%
  colnames()
```


### Variable: `location`
```{r}
table(knl_effort$location) 
```

Some years do not have location variable filled in. Fill in with `KL` for these years.

```{r}
filter(knl_effort, is.na(location)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_effort$location))/nrow(knl_effort), 3) * 100` % of values in the `location` column are NA.

### Variable: `gear`

```{r}
table(knl_effort$gear) 
```

Some years do not have gear filled in. Leave blank because do not know what gear type.
Change to "2x 8 in" for the 2 x 8. I think these mean two 8 in cones.

```{r}
filter(knl_effort, is.na(gear)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_effort$gear))/nrow(knl_effort), 3) * 100` % of values in the `gear` column are NA.

### Variable: `cone_sampling_effort`

I don't think this variable is very informative. Missing values for every year.

```{r}
table(knl_effort$cone_sampling_effort) 
```


```{r}
filter(knl_effort, is.na(cone_sampling_effort)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_effort$cone_sampling_effort))/nrow(knl_effort), 3) * 100` % of values in the `cone_sampling_effort` column are NA.

### Variable: `turbidity_units`
```{r}
table(knl_effort$turbidity_units) 
```

Some years do not have location variable filled in. Fill in for these years.

```{r}
filter(knl_effort, is.na(turbidity_units)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_effort$turbidity_units))/nrow(knl_effort), 3) * 100` % of values in the `turbidity_units` column are NA.

### Variable: `comments`

```{r}
filter(knl_effort, is.na(comments)) %>%
  group_by(year(date)) %>%
  tally()
```

**NA and Unknown Values**

`r round(sum(is.na(knl_effort$comments))/nrow(knl_effort), 3) * 100` % of values in the `comments` column are NA.

## Make any changes identified

```{r make_changes}
knl_effort_changes <- knl_effort %>%
  # fix typo in water_t_f
  mutate(water_t_f = case_when(date == "2009-06-17" ~ 70,
                               T ~ water_t_f),
         # cone_id as character instead of numeric
         cone_id = as.character(cone_id),
         # fix date formatting
         date = as.Date(date),
         start_date = ymd_hms(start_date),
         stop_date = ymd_hms(stop_date),
         # fill in location for all
         location = "KL",
         # change name for gear
         gear = case_when(gear %in% c("2 x 8' Cone", "2X 8' Cone") ~ "2x 8 in cone",
                          # all rst data so just leave blank for now
                          gear == "RST" ~ NA_character_,
                          T ~ gear)) %>%
  # very little data, only for sampling year 2021 so remove.
  select(-klci)
```

### Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
f <- function(input, output) write_csv(input, file = output)

upload_data <- function(data, name) {
gcs_upload(data,
           object_function = f,
           type = "csv",
           name = paste0("rst/lower-sac-river/data/knights-landing/knl_combine_", name, ".csv"))
}

upload_data(knl_effort_changes, "sampling_effort_clean")
```
