---
title: "Battle Creek Carcass Survey QC"
author: "Erin Cain"
date: "9/29/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl) 
library(knitr)
```

# Battle Creek Carcass Survey QC

## Description of Monitoring Data

These data were acquired via snorkel and kayak surveys on Battle Creek from 1996 to 2019 and describe spring-run and unknown run Chinook Salmon carcasses found within and along Battle Creek.						

**Timeframe:** 1996 - 2019

**Survey Season:** March through October (most years not the entire timeframe)

**Completeness of Record throughout timeframe:** Values for ever year in timeframe

**Sampling Location:** Battle Creek 

**Data Contact:** [Natasha Wingerter](mailto:natasha_wingerter@fws.gov); [RJ Bottaro](mailto:rj_bottaro@fws.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))

gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "adult-holding-redd-and-carcass-surveys/battle-creek/data-raw/battle_creek_adult_spawn_hold_carcass.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "raw_adult_spawn_hold_carcass.xlsx",
               overwrite = TRUE)
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
sheets <- excel_sheets("raw_adult_spawn_hold_carcass.xlsx")
sheets 
raw_carcass_data <- read_excel("raw_adult_spawn_hold_carcass.xlsx", 
                               sheet = "Carcass",
                               col_types = c("numeric", "numeric", "numeric", "numeric", "text", "text", 
                                             "text", "text", "text", "text", "numeric", "text", 
                                             "text", "text", "text", "text", "text", "text")) %>% glimpse()

```

## Data transformations

```{r}
cleaner_carcass_data <- raw_carcass_data %>%
  janitor::clean_names() %>% 
  rename("river_mile" = rivermile,
         "run" = fws_run_call) %>%
  mutate(date1 = date, 
         date2 = as.Date.numeric(as.numeric(date), origin = "1899-12-30"),
         date = if_else(is.na(date2), as.Date(date, format = "%m/%d/%Y"), date2)) %>% # see if there is a better way to deal with mixed date format
  select(-species, -spawn_year, -date1, -date2) %>% # All either CHINOOK or Chinook
  glimpse()
```

## Data Dictionary

The following table describes the variables included in this dataset and the percent that do not include data.

```{r data_dictionary}
percent_na <- cleaner_carcass_data %>%
  summarise_all(list(name = ~sum(is.na(.))/length(.))) %>%
  pivot_longer(cols = everything())
  
data_dictionary <- tibble(variables = colnames(cleaner_carcass_data),
                          description = c("GPS X point",
                                          "GPS Y point",
                                          "River mile number",
                                          "Sample date",
                                          "Survey method (barrier weir trap, fish rescue, kayak survey, snorkel survey, other)",
                                          "Reach number (1-7); other location",
                                          "Male, female, unknown",
                                          "True if the fish was observed and not sampled (T/F)",
                                          "Fork length of fish in mm",
                                          "Adipose status (absent, present, unknown)",
                                          "Carcass status (bright, fresh, non-fresh, unknown)",
                                          "Spawning status (green, ripe, spawned, unknown, unspawned)",
                                          "Run of Chinook (spring, unknown)",
                                          "Coded wire tag number",
                                          "Any other tag or recapture marking",
                                          "General comments from survey crew"),
                          percent_na = round(percent_na$value*100)
                          
)
kable(data_dictionary)
```

## Explore Numeric Variables: {.tabset}

```{r}
cleaner_carcass_data %>% select_if(is.numeric) %>% colnames()
```

### Variable: `longitude`, `latitude`

```{r}
summary(cleaner_carcass_data$latitude)
summary(cleaner_carcass_data$longitude)
```

All values look within an expected range 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$latitude))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `latitude` column are NA.
* `r round(sum(is.na(cleaner_carcass_data$longitude))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `longitude` column are NA.

### Variable: `river_mile`

Data are filtered for spring run.

**Plotting river mile over Period of Record**

```{r}
filter(cleaner_carcass_data, run == "SCS") %>% 
  ggplot(aes(x = river_mile, y = year(date))) +
  geom_point(size = 1.4, alpha = .5, color = "blue") + 
  labs(x = "River Mile", 
       y = "Date") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

It looks like river miles 0 - 5 and 15 - 17 most commonly have carcasses. Very little data on river mile pre 2010. 

```{r}
filter(cleaner_carcass_data, run == "SCS") %>% 
  ggplot(aes(x = river_mile)) +
  geom_histogram(alpha = .75) + 
  labs(x = "River Mile") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

**Numeric Summary of river mile over Period of Record**

```{r}
summary(cleaner_carcass_data$river_mile)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$river_mile))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `river_mile` column are NA. 

### Variable: `fork_length`

Fork length in mm. Data are filtered for spring run.

**Plotting distribution of fork length**

```{r}
filter(cleaner_carcass_data, run == "SCS") %>% 
  ggplot(aes(x = fork_length)) +
  geom_histogram(alpha = .75) + 
  labs(x = "Fork Length") +
  theme_minimal() + 
  theme(text = element_text(size = 15)) 
```

```{r}
filter(cleaner_carcass_data, run == "SCS") %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = fork_length, y = year)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Fork length summarized by year") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

There is one observation where year is NA. This record should be removed from the dataset because it does not contain any data.

```{r}
cleaner_carcass_data %>% 
  mutate(year = as.factor(year(date))) %>%
  filter(is.na(year))
```

**Numeric Summary of fork length over Period of Record**

```{r}
summary(cleaner_carcass_data$fork_length)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$fork_length))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `fork_length` column are NA.

## Explore Categorical variables: {.tabset}


```{r}
cleaner_carcass_data %>% select_if(is.character) %>% colnames()
```


### Variable: `method`
```{r}
table(cleaner_carcass_data$method) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$method <- case_when(cleaner_carcass_data$method == "Barrier Weir Trap" | 
                                           cleaner_carcass_data$method == "Weir" ~ "barrier weir trap",
                                         cleaner_carcass_data$method == "Fish Rescue" ~ "fish rescue", 
                                         cleaner_carcass_data$method == "Kayak Survey" ~ "kayak survey",
                                         cleaner_carcass_data$method == "Snorkel Survey" ~ "snorkel survey",
                                         cleaner_carcass_data$method == "Other" |
                                           cleaner_carcass_data$method == "Other (CDFG)" ~ "other")

table(cleaner_carcass_data$method) 
```

It looks like most carcasses are from barrier weir trap and snorkel survey 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$method))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `method` column are NA.

### Variable: `location`

Reach number (1-7); other location	

* CNFH - Coleman National Fish Hatchery 

```{r}
table(cleaner_carcass_data$location) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$location))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `location` column are NA.


### Variable: `sex`
```{r}
table(cleaner_carcass_data$sex) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$sex <- tolower(cleaner_carcass_data$sex)

table(cleaner_carcass_data$sex) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$sex))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `sex` column are NA.
* `r round(sum(cleaner_carcass_data$sex == "unknown")/nrow(cleaner_carcass_data), 3) * 100` % of values in the `sex` column are `Unknown`.

### Variable: `observed_only`
```{r}
table(cleaner_carcass_data$observed_only) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$observed_only <- case_when(cleaner_carcass_data$observed_only == "Y" ~ TRUE, 
                                              cleaner_carcass_data$observed_only == "N" ~ FALSE)

table(cleaner_carcass_data$observed_only) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$observed_only))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `observed_only` column are NA.



### Variable: `adipose`
```{r}
table(cleaner_carcass_data$adipose) 
```

- Fix inconsistencies with spelling, capitalization, and abbreviations. 
- Converted to boolean variable.

```{r}
cleaner_carcass_data$adipose <- tolower(cleaner_carcass_data$adipose) 
cleaner_carcass_data$adipose <- case_when(cleaner_carcass_data$adipose == "absent" ~ FALSE,
                                          cleaner_carcass_data$adipose == "present" ~ TRUE,
                                          T ~ NA)
table(cleaner_carcass_data$adipose) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$adipose))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `adipose` column are NA.
* `r round(sum(cleaner_carcass_data$adipose == "unknown")/nrow(cleaner_carcass_data), 3) * 100` % of values in the `adipose` column are `Unknown`.


### Variable: `carcass_live_status`
```{r}
table(cleaner_carcass_data$carcass_live_status) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$carcass_live_status <- tolower(cleaner_carcass_data$carcass_live_status)

table(cleaner_carcass_data$carcass_live_status) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$carcass_live_status))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `carcass_live_status` column are NA.
* `r round(sum(cleaner_carcass_data$carcass_live_status == "unknown", na.rm = T)/nrow(cleaner_carcass_data), 3) * 100` % of values in the `carcass_live_status` column are `Unknown`.

### Variable: `spawn_condition`
```{r}
table(cleaner_carcass_data$spawn_condition) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$spawn_condition <- tolower(cleaner_carcass_data$spawn_condition)

table(cleaner_carcass_data$spawn_condition) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$spawn_condition))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `spawn_condition` column are NA.
* `r round(sum(cleaner_carcass_data$spawn_condition == "unknown", na.rm = T)/nrow(cleaner_carcass_data), 3) * 100` % of values in the `spawn_condition` column are `Unknown`.

### Variable: `run`
```{r}
table(cleaner_carcass_data$run) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$run <- ifelse(cleaner_carcass_data$run == "SCS", "spring", "unknown")
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$run))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `run` column are NA.
* `r round(sum(cleaner_carcass_data$run == "unknown", na.rm = T)/nrow(cleaner_carcass_data), 3) * 100` % of values in the `run` column are `Unknown`.

### Variable: `cwt_code`

TODO Figure out what NTD means

```{r}
table(cleaner_carcass_data$cwt_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$cwt_code <- tolower(cleaner_carcass_data$cwt_code)

table(cleaner_carcass_data$cwt_code) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$cwt_code))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `cwt_code` column are NA.


### Variable: `other_tag`

TODO Figure out what AFTC means

```{r}
table(cleaner_carcass_data$other_tag) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_carcass_data$other_tag <- ifelse(cleaner_carcass_data$other_tag == "AFTC", "AFTC", NA)
table(cleaner_carcass_data$other_tag) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$other_tag))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `other_tag` column are NA.


### Variable: `comments`
```{r}
unique(cleaner_carcass_data$comments)[1:5]
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_carcass_data$comments))/nrow(cleaner_carcass_data), 3) * 100` % of values in the `comments` column are NA.

## Summary of identified issues

* Looks like mostly samples after 2010 
* Clarify what NTD means in CWT code and location info that is not reach length
* Other tag column was dropped because it has only one entry other than "none"

## Next steps

### Columns to remove

- Suggest removing the `location` item or transforming to `longitude`, `latitude` or `river_mile`.
- Decided to remove `other_tag` because no data except for one data point is included.
- `comments` and `observed_only` may not be needed.

## Save cleaned data back to google cloud 

```{r}
battle_carcass <- cleaner_carcass_data %>% 
  select(-other_tag) %>%
  filter(!is.na(date)) %>%
  glimpse()
```


```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)

gcs_upload(battle_carcass,
           object_function = f,
           type = "csv",
           name = "adult-holding-redd-and-carcass-surveys/battle-creek/data/battle_carcass.csv")
```
