---
title: "Combine Feather River Snorkel Data"
author: "Liz Stebbins"
date: "5/2/2024"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(googleCloudStorageR)
library(knitr)
color_pal <- c("#9A8822",  "#F8AFA8", "#FDDDA0", "#74A089", "#899DA4", "#446455", "#DC863B", "#C93312")

gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
```

## Snorkel Data Standardization

FlowWest received snorkel data for Feather River in two separate Access databases: one for pre-2004 data and one for 2004-2020 data. These are processed in separate markdowns.
  
```{r, message = FALSE, warning = FALSE}
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/pre_2004_snorkel.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "snorkel_pre_2004.csv"),
               overwrite = TRUE)
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/post_2004_snorkel.csv",
               bucket = gcs_get_global_bucket(),
               saveToDisk = here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "snorkel_post_2004.csv"),
               overwrite = TRUE)

snorkel_pre_2004 <- read_csv(here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "snorkel_pre_2004.csv"))
snorkel_post_2004 <- read_csv(here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "snorkel_post_2004.csv"))

```

## Standard format for Adult All Snorkel Data

## Read in data {.tabset}
  
Below we read in the snorkel data from the different databases and rename or select columns so that we can join all the monitoring datasets together in the section below.

### Pre-2004

#### Columns Removed

- `flow` and `adj_flow` - sparsely recorded and values not on the same scale as `river_flow`
- `fish_depth` seemed redundant with `river_depth`
- `max_fork_length` seemed redundant with `fork_length`

columns created/renamed:
- `river_flow` renamed to `flow`
- `river_depth` renamed to `water_depth_m`
- `huc_unit` renamed to `unit_type`

```{r}
clean_snorkel_pre_2004 <- snorkel_pre_2004 |> 
  select(-c(flow, adj_flow, fish_depth, max_fork_length)) |> 
  rename(flow = river_flow,
         water_depth_m = river_depth,
         unit_type = huc_unit) |> 
  glimpse()
```

### Post-2004

#### Columns Removed

- `survey_comments` and `observation_comments` - not necessary for final dataset
- `time_of_temperature`, `observation_id`, `lwd_number` - not necessary for final dataset

columns created/renamed:
- `weather_code` renamed to `weather`
- `hydrology_code` renamed to `hydrology`
- 

```{r}
# get lookup tables from pre-2004 table for some codes
db_filepath <- here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "feather-river-db.mdb")
library(Hmisc)
lookup_HUC_cover <- mdb.get(db_filepath, "HUCcoverLU")
lookup_HUC_o_cover <- mdb.get(db_filepath, "HUCOcoverLU")
lookup_HUC_substrate <- mdb.get(db_filepath, "HUCsubstrateLU")
detach(package:Hmisc)

# first join the lookup tables
clean_snorkel_post_2004_lu <- snorkel_post_2004 |> 
  left_join(lookup_HUC_o_cover, by = c("overhead_cover" = "CoverCode")) |> 
    mutate(Cover = case_when(overhead_cover == 12 ~ "Overhead Object/Veg. 0 - 0.5m, Overhead Object/Veg. 0.5 - 2m",
                           overhead_cover == 13 ~ "Overhead Object/Veg. 0 - 0.5m, Submerged Aquatic Veg/Algae",
                           overhead_cover == 4 ~ "unknown (code 4)",
                           TRUE ~ Cover),
           Cover = str_to_lower(Cover)) |> 
  select(-c(overhead_cover)) |> 
  rename(overhead_cover = Cover) |> 
  glimpse()

clean_snorkel_post_2004 <- clean_snorkel_post_2004_lu |> 
  select(-c(survey_comments, observation_comments, time_of_temperature,
            observation_id, lwd_number)) |> 
  rename(weather = weather_code,
         hydrology = hydrology_code) |> 
  mutate(hydrology = str_to_lower(hydrology),
         substrate = as.character(substrate)) |> 
  glimpse()
```


## Combine data:

```{r}
all_snorkel_data <- bind_rows(clean_snorkel_post_2004, 
                           clean_snorkel_pre_2004)  |> glimpse()
```

## Standardization of combined data

### Location

This code was written by Maddee and pulled in on 3-13-2024. Mapped to standard section names in DMP map.

```{r}
all_snorkel_data_section_names <- all_snorkel_data |> 
  mutate(section_name = case_when(section_name == "Vance West" ~ "Vance Riffle",
                                  section_name == "Eye" ~ "Eye Riffle",
                                  section_name == "Hatchery Side Ditch" ~ "Hatchery Ditch",
                                  section_name == "Hatchery Side Channel" ~ "Hatchery Riffle", 
                                  section_name == "Gridley Side Channel" ~ "Gridley Riffle", 
                                  section_name %in% c("Robinson", "Lower Robinson") ~ "Robinson Riffle",
                                  section_name == "Goose" ~ "Goose Riffle", 
                                  section_name == "Auditorium" ~ "Auditorium Riffle",
                                  section_name %in% c("Matthews", "Mathews", "Mathews Riffle") ~ "Matthews Riffle",
                                  section_name %in% c("G95 Side Channel", "G95 West Side Channel", "G95 Side West", "G95 Side") ~ "G95", 
                                  section_name %in% c("Vance West Riffle", "Vance W Riffle", "Vance East") ~ "Vance Riffle",
                                  section_name %in% c("Moes", "Moes Ditch") ~ "Mo's Ditch",
                                  section_name == "Aleck" ~ "Aleck Riffle",
                                  section_name %in% c("Lower Mcfarland", "Upper Mcfarland") ~ "McFarland",
                                  section_name %in% c("Bed Rock Riffle", "Bedrock Riffle", "Bedrock", "Bedrock Park") ~ "Bedrock Park Riffle",
                                  section_name == "Steep" ~ "Steep Riffle",
                                  section_name %in% c("Keister", "Keister Riffle") ~ "Kiester Riffle",
                                  section_name == "Junkyard" ~ "Junkyard Riffle",
                                  section_name == "Gateway" ~ "Gateway Riffle",
                                  section_name == "Trailer Park" ~ "Trailer Park Riffle",
                                  section_name %in% c("Hatchery Ditch And Moes", "Hatchery Ditch Moes Ditch", 
                                                      "Hatchery Side Channel Moes Ditch", 
                                                      "Hatchery Ditch Lower Moes Ditch Upper",
                                                      "Hatchery Ditch And Moes Ditch", 
                                                      "Hatchery Side Channel And Moes Ditch", 
                                                      "Hatchery Ditch Moes") ~ "Hatchery Ditch and Mo's Ditch",
                                  section_name %in% c("Hatchery And Moes Side Channels", "Hatchery Side Ch Moes Side Ch", 
                                                      "Hatchery Side Channel And Moes") ~ "Hatchery and Mo's Riffles", 
                                  .default = as.character(section_name))) |> 
  glimpse()
```


```{r, message = FALSE}
# map to section_number from DMP for VA feather snorkel
# map the section name to a section number (where applicable)
all_snorkel_data_final <- all_snorkel_data_section_names |> 
  mutate(section_number = case_when(section_name == "Aleck Riffle" ~ 8,
                                    section_name == "Auditorium Riffle" ~ 4,
                                    section_name == "Bedrock Park Riffle" ~ 5,
                                    section_name == "Bedrock Riffle" ~ 10,
                                    section_name == "Big Riffle" ~ 17,
                                    section_name == "Eye Riffle" ~ 11,
                                    section_name == "G95" ~ 14,
                                    section_name == "Gateway Riffle" ~ 12,
                                    section_name == "Goose Riffle" ~ 16,
                                    section_name == "Gridley Riffle" ~ 19,
                                    section_name == "Hatchery Ditch" ~ 2,
                                    section_name == "Hatchery Riffle" ~ 1,
                                    section_name == "Junkyard Riffle" ~ 20,
                                    section_name == "Kiester Riffle" ~ 15,
                                    section_name == "Matthews Riffle" ~ 7,
                                    section_name == "McFarland" ~ 18,
                                    section_name == "Mo's Ditch" ~ 3,
                                    section_name == "Robinson Riffle" ~ 9,
                                    section_name == "Steep Riffle" ~ 10,
                                    section_name == "Trailer Park Riffle" ~ 6,
                                    section_name == "Vance Riffle" ~ 13,
                                    TRUE ~ NA))
```

### Temperature - conversion of F to C

It looks like temperature was recorded in either `F` or `C` over the whole time period. There seems to be a distinct cutoff between `F` temperatures (the higher points) and `C` temperatures. For all temperatures above a threshold of `40` (which, if that were in `F`, would be `104`), we convert to `C`.

There are some `0` values.

```{r}
all_snorkel_data_final |> 
  ggplot(aes(x = date, y = temperature)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Date", y = "Temperature", title = "Temperature sampled over time frame")
```

```{r}
all_snorkel_data_final <- all_snorkel_data_final |> 
  mutate(temperature_c = ifelse(temperature >= 40, 
                                weathermetrics::fahrenheit.to.celsius(temperature),
                                temperature))
all_snorkel_data_final |> 
  ggplot(aes(x = date, y = temperature_c)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Date", y = "Temperature (C)", title = "Temperature in C sampled over time frame")

all_snorkel_data_final <- all_snorkel_data_final |> 
  select(-temperature) |> 
  rename(temperature = temperature_c)
```
## Save Cleaned Data to Google Cloud

```{r}
knitr::kable(all_snorkel_data_final  |> head())
```

```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(all_snorkel_data_final,
           object_function = f,
           type = "csv",
           name = "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data/combined_feather_snorkel_data.csv",
           predefinedAcl = "bucketLevel")
```
