---
title: "Feather River Snorkel Pre-2004 QC"
author: "Liz Stebbins"
date: "3/6/2024"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River Snorkel Data 1999 - 2003

## Description of Monitoring Data

Feather River Snorkel Data from 1999 - 2003. These data were provided to flowwest in an access database `FR_S_and_S_Oroville.mdb`. We extracted the tables of interest (`SnorkObservations` and `SnorkelSurvey`, as well as some lookup tables). Snorkel survey data from 2004 - 2020 are available in a separate database `Snorkel_Revised.mdb` and processed in the markdown `feather_snorkel_qc.Rmd`.

**Timeframe:** 1999 - 2003

**Completeness of Record throughout timeframe:** 

Only SR collected are in 1999 - 2001

**Sampling Location:** Feather River

**Data Contact:** [Casey Campos](mailto:Casey.Campos@water.ca.gov)

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/db_from_feather_river/FR_S_and_S_Oroville.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/seine-snorkel-data/feather-river/feather-river-db.mdb",
               overwrite = TRUE)

```

Read in data from google cloud, glimpse raw data: 
```{r, warning=FALSE, message = FALSE}
db_filepath <- here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "feather-river-db.mdb")

library(Hmisc)
mdb.get(db_filepath, tables = TRUE) 
all_species <- mdb.get(db_filepath, "All Species Query Table") # what is this table?
snorkel_obsv <- mdb.get(db_filepath, "SnorkObservationsTBL") 
snorkel_survey_metadata <- mdb.get(db_filepath, "SnorkSurveyTBL")
lookup_HUC_cover <- mdb.get(db_filepath, "HUCcoverLU")
lookup_HUC_o_cover <- mdb.get(db_filepath, "HUCOcoverLU")
lookup_HUC_substrate <- mdb.get(db_filepath, "HUCsubstrateLU")
lookup_HUC_unit <- mdb.get(db_filepath, "HUCunitLU")
lookup_weather <- mdb.get(db_filepath, "WeatherLU")
detach(package:Hmisc)

# write to csvs
write_csv(snorkel_obsv, here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather.csv"))
write_csv(snorkel_survey_metadata, here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather_metadata.csv"))

# read in data
snorkel_raw <- read_csv(here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather.csv"))
snorkel_metadata_raw <- read_csv(here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather_metadata.csv"))

```

## Data transformations
`All Species Query Table` only spans `1998-1999`.

Read in individual observation files and survey metadata files. Update column names and column types. Filter to only show chinook salmon.


```{r}
cleaner_snorkel_data <- snorkel_raw |> 
  janitor::clean_names() |> 
  mutate(huc_ocover = as.numeric(huc_ocover)) |> 
  filter(species %in% c("CHN", "None", "C", "CHNF", "CHNS", "CHNL")) |>  # filter species to relevant values (none is helpful to show they snorkeled and did not see anything)
  left_join(lookup_HUC_unit |> 
              rename(huc_unit = Unit), by = c("hu_cunit" = "UnitCode")) |>
  left_join(lookup_HUC_cover |> 
            rename(huc_cover = Cover), by = c("huc_icover" = "CoverCode")) |>
  left_join(lookup_HUC_substrate |> 
            rename(huc_substrate = Substrate), by = c("hu_csubstrate" = "SubstrateCode")) |>
  left_join(lookup_HUC_o_cover |> 
            rename(huc_o_cover = Cover), by = c("huc_ocover" = "CoverCode")) |>
  select(-c(hu_csubstrate, hu_cunit, huc_icover, huc_ocover, snorkler, obs_id)) |> 
    mutate(run = case_when(species  == "CHNF" ~ "fall",
                         species == "CHNS" ~ "spring",
                         species == "CHNL" ~ "late fall",
                         species %in% c("CHN",  "None", "C") ~ "unknown"),
           huc_unit = str_to_lower(huc_unit),
           huc_cover = str_to_lower(huc_cover),
           huc_o_cover = str_to_lower(huc_o_cover),
           huc_substrate = str_to_lower(huc_substrate)) |> 
  rename(flow = velocity, adj_flow = adj_velocity) |> 
  glimpse()

cleaner_snorkel_metadata <- snorkel_metadata_raw |> 
  janitor::clean_names() |> 
  left_join(lookup_weather, by = c("weather" = "WeatherCode")) |> 
  select(-c(visibility_comments, x_of_divers, x_of_center_passes, pass_width, comments,
          temp_time, snorkel_time_start, snorkel_time_stop, weather,
          snorkel_crew, shore_crew, recorder)) |> # doesn't seem like time information is being read in from mdb.get - TODO 
  mutate(location = str_to_title(location),
         survey_type = str_to_lower(survey_type),
         section_type = str_to_lower(section_type),
         weather = str_to_lower(Weather)) |> 
  select(-c(Weather)) |> 
  glimpse()

snorkel <- left_join(cleaner_snorkel_data, cleaner_snorkel_metadata, by = "survey_id") |> 
  # remove units (repetitive) and comments, rename some variables for clarity
  select(survey_id, date, location, unit, survey_type, section_type, count = number, species, run,
         fork_length = fl, max_fork_length = max_fl, fish_depth, river_depth = depth,
         flow, adj_flow, river_flow, bank_distance, huc_unit, huc_cover,
         huc_o_cover, huc_substrate, visibility, temperature, weather) |>
  filter(!is.na(date)) |> 
  glimpse()

```

## Explore Numeric Variables: {.tabset}

```{r}
snorkel |> select_if(is.numeric) |> colnames()
```

### Variable: `flow`

**Plotting flow over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = date, y = flow)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  theme_minimal()
```

Very inconsistent and sparse flow measures. No flow measurements after 2000-04-04.

```{r}
snorkel |> 
  ggplot(aes(x = flow)) +
  geom_histogram() + 
  theme_minimal()
```

Flow is between 0 - 5.8. 0 values must be errors or `NA`.  

**Numeric Summary of flow over Period of Record**

```{r}
summary(snorkel$flow)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$flow))/nrow(snorkel), 3) * 100` % of values in the `flow` column are NA. 

### Variable: `adj_flow`

**Plotting adjusted flow over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = date, y = adj_flow)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  theme_minimal()
```

Very inconsistent and sparse adjusted flow measures. No adjusted flow measurements after 2000-03-23.

```{r}
snorkel |> 
  ggplot(aes(x = adj_flow)) +
  geom_histogram() + 
  theme_minimal()
```

Adjusted is between 0 - 3.97. 0 values must be errors or `NA`.  

**Numeric Summary of flow over Period of Record**

```{r}
summary(snorkel$adj_flow)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$adj_flow))/nrow(snorkel), 3) * 100` % of values in the `adj_flow` column are NA. 

### Variable: `river_flow`

**Plotting river flow over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = date, y = river_flow)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  theme_minimal()
```

River flow is collected more consistently than `flow` or `adj_flow`

```{r}
snorkel |> 
  ggplot(aes(x = river_flow)) +
  geom_histogram() + 
  theme_minimal()
```

River flow is between 1 - 9564. River flow seems to be the most consistent and reasonable flow variable in this dataset.  

**Numeric Summary of river flow over Period of Record**

```{r}
summary(snorkel$river_flow)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$river_flow))/nrow(snorkel), 3) * 100` % of values in the `flow` column are NA. 


### Variable: `visibility`

**Plotting visibility over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = date, y = visibility)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  theme_minimal()
```

Visibility is collected sporadically.

```{r}
snorkel |>
  ggplot(aes(x = visibility)) + 
  geom_histogram() + 
  theme_minimal()
```

All visibility measures fall between 0.5 and 9.0. 

**Numeric Summary of visibility over Period of Record**

```{r}
summary(snorkel$visibility)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$visibility))/nrow(snorkel), 3) * 100` % of values in the `visibility` column are NA.

### Variable: `temperature`

TODO: Check that they are in both F and C and then divide appropriately  

**Plotting temperature over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = date, y = temperature)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  theme_minimal()
```

Consistency of temperature measures varies throughout the years. 

```{r}
snorkel |> 
  ggplot(aes(x = temperature)) +
  geom_histogram() +
  theme_minimal()
```

Appears as if most values are collected in Fahrenheit and a few are collected in Celsius (the lowest ones)?

**Numeric Summary of temperature over Period of Record**

```{r}
summary(snorkel$temperature)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$temperature))/nrow(snorkel), 3) * 100` % of values in the `temperature` column are NA.


### Variable: `count`

**Plotting count over Period of Record**

```{r, include=FALSE}
sac_indices <- waterYearType::water_year_indices |> 
    filter(location == "Sacramento Valley") |> 
    transmute(water_year = WY, year_type = as.character(Yr_type))
```

```{r}
snorkel |> 
  filter(run == "spring") |>
  group_by(date) |>
  summarise(total_daily_catch = sum(count, na.rm = T)) |>
  # filter(year(date) > 2014, year(date) < 2021) |>
  mutate(water_year = ifelse(month(date) %in% 10:12, year(date) + 1, year(date))) |> 
  left_join(sac_indices) |>
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) |>
  ggplot(aes(x = fake_date, y = total_daily_catch, fill = year_type)) + 
  geom_col() + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1901-03-01"), as.Date("1901-10-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Total Daily Raw Fish Count",
       y = "Total daily catch",
       x = "Date")+ 
  facet_wrap(~water_year, scales = "free") + 
  scale_fill_manual(values = wesanderson::wes_palette("Moonrise2", 5, type = "continuous"))
```


Very few SR fish, only caught SR in 1999, 2000, and 2001.  
```{r}
snorkel  |>
  mutate(year = as.factor(year(date))) |>
  ggplot(aes(x = year, y = count, fill = run)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Fish Counted each Year",
       y = "Total fish count") + 
  scale_fill_manual(values = c("#E1BD6D", "#0B775E", "#F2300F", "#35274A")) +
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of count over Period of Record**

```{r}
summary(snorkel$count)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$count))/nrow(snorkel), 3) * 100` % of values in the `count` column are NA.

### Variable: `fork_length`

**Plotting est_size over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = fork_length)) + 
  geom_histogram() + 
  theme_minimal()
```


**Numeric Summary of fork_length over Period of Record**

```{r}
summary(snorkel$fork_length)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$fork_length))/nrow(snorkel), 3) * 100` % of values in the `fork_length` column are NA.

### Variable: `max_fork_length`

**Plotting max fork length over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = max_fork_length)) + 
  geom_histogram() + 
  theme_minimal()
```

**Numeric Summary of max fork length over Period of Record**

```{r}
summary(snorkel$max_fork_length)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$max_fork_length))/nrow(snorkel), 3) * 100` % of values in the `max_fork_length` column are NA.

### Variable: `fish_depth`

Depth at which fish was observed.

**Plotting fish depth over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = fish_depth)) + 
  geom_histogram() + 
  theme_minimal()
```
Most of the fish were observed at a depth between 0-2m.

**Numeric Summary of fish depth over Period of Record**

```{r}
summary(snorkel$fish_depth)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$fish_depth))/nrow(snorkel), 3) * 100` % of values in the `fish_depth` column are NA.

### Variable: `river_depth`

Depth of the river at sampling site.

**Plotting water depth over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = river_depth)) + 
  geom_histogram() + 
  theme_minimal()
```
River depth was mostly between 1-2 m.

**Numeric Summary of depth of water over Period of Record**

```{r}
summary(snorkel$river_depth)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$river_depth))/nrow(snorkel), 3) * 100` % of values in the `river_depth` column are NA.

### Variable: `bank_distance`

Distance from river bank.

**Plotting distance from river bank over Period of Record**

```{r}
snorkel |> 
  ggplot(aes(x = bank_distance)) + 
  geom_histogram() + 
  theme_minimal()
```

Distance from river bank was generally between 0-2.5 m. 

**Numeric Summary of distance from river bank over Period of Record**

```{r}
summary(snorkel$bank_distance)
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$bank_distance))/nrow(snorkel), 3) * 100` % of values in the `bank_distance` column are NA.

## Explore Categorical variables: {.tabset}


```{r}
snorkel |> select_if(is.character) |> colnames()
```


### Variable: `location`

Locations appear to be mapped to a survey reach and can be inconsistent (i.e. some are reach X - reach Y, some are just reach X). In later years, this is `section_name`.

```{r}
table(snorkel$location)

format_site_name <- function(string) {
  clean <- 
    str_replace_all(string, "'", "") |>
    str_replace_all("G-95", "G95") |> 
    str_replace_all("[^[:alnum:]]", " ") |> 
    trimws() |> 
    stringr::str_squish() |>
    stringr::str_to_title()
}

snorkel$location <- format_site_name(snorkel$location)
table(snorkel$location)
```

There are `r length(unique(snorkel$location))` unique locations.

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$location))/nrow(snorkel), 3) * 100` % of values in the `location` column are NA.

### Variable: `unit`

Not sure what `unit` corresponds to - should clarify this. In later years (see other .Rmd) this is `section_name`. 

```{r}
table(snorkel$unit) 
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$unit))/nrow(snorkel), 3) * 100` % of values in the `unit` column are NA.

### Variable: `survey_type`

Survey type, either `comp` or `unit`.

```{r}
table(snorkel$survey_type) 
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$survey_type))/nrow(snorkel), 3) * 100` % of values in the `survey_type` column are NA.

### Variable: `section_type`

Section type: either `permanent` or `random`

```{r}
table(snorkel$section_type)

snorkel$section_type <- ifelse(snorkel$section_type == "n/a", NA, snorkel$section_type)

table(snorkel$section_type)
```

There are `r length(unique(snorkel$section_type))` unique groups units covered. 

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$section_type))/nrow(snorkel), 3) * 100` % of values in the `section_type` column are NA.

### Variable: `species` 

Dataset were filtered to only Chinook.

```{r}
table(snorkel$species)

snorkel$species <- ifelse(snorkel$species == "None", "unknown", "chinook")

table(snorkel$species)
```


**NA and Unknown Values**

* `r round(sum(is.na(snorkel$species))/nrow(snorkel), 3) * 100` % of values in the `species` column are NA.

### Variable: `run``

Run type.

```{r}
table(snorkel$run)
```

There are `r length(unique(snorkel$run))` unique observation ids

* `r round(sum(is.na(snorkel$run))/nrow(snorkel), 3) * 100` % of values in the `run` column are NA.

### Variable: `huc_unit`

```{r}
table(snorkel$huc_unit) 
```

There are `r length(unique(snorkel$huc_unit))` unique units covered. 

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$huc_unit))/nrow(snorkel), 3) * 100` % of values in the `huc_unit` column are NA.

### Variable: `huc_cover`

This is instream cover. Lookup table from database was joined in cleaning stage.

```{r}
snorkel <- snorkel |> 
  rename(instream_cover = huc_cover)
table(snorkel$instream_cover) 
```


**NA and Unknown Values**

* `r round(sum(is.na(snorkel$instream_cover))/nrow(snorkel), 3) * 100` % of values in the `instream_cover` column are NA.

### Variable: `huc_o_cover`

This is overhead cover. Lookup table from database was joined at cleaning stage.


```{r}
snorkel <- snorkel |> 
  rename(overhead_cover = huc_o_cover)
table(snorkel$overhead_cover) 
```

**NA and Unknown Values**

* `r round(sum(is.na(snorkel$overhead_cover))/nrow(snorkel), 3) * 100` % of values in the `overhead_cover` column are NA.

### Variable: `huc_substrate`

Lookup table from database was joined at cleaning stage.


```{r}
snorkel <- snorkel |> 
  rename(substrate = huc_substrate)
table(snorkel$substrate) 
```


**NA and Unknown Values**

* `r round(sum(is.na(snorkel$substrate))/nrow(snorkel), 3) * 100` % of values in the `substrate` column are NA.


### Variable: `weather`

Lookup table from database was joined at cleaning stage.


```{r}
table(snorkel$weather) 
```


**NA and Unknown Values**

* `r round(sum(is.na(snorkel$weather))/nrow(snorkel), 3) * 100` % of values in the `weather` column are NA.



## Summary of identified issues

* No spring run observed in `2002` and `2003`
* Some survey IDs do not have associated dates: `472`, some entries from survey id `546`, and some `NA` survey dates.
* Several flow variables, though `river_flow` is the most consistently collected
* Still a few unknown columns that we need to define and understand 
* Detailed information on location and units split up; need to update standardized reach lookup table so that it accounts for the new entries represented in this dataset

## Save cleaned data back to google cloud 

```{r, eval=FALSE}
f <- function(input, output) write_csv(input, file = output)
gcs_upload(snorkel,
           object_function = f,
           type = "csv",
           name = "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/pre_2004_snorkel.csv",
           predefinedAcl = "bucketLevel")
```

