---
title: "Feather River Snorkel Pre-2004 QC"
author: "Liz Stebbins"
date: "3/6/2024"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library(readxl)
```

# Feather River Snorkel Data Pre-2004

## Description of Monitoring Data

TODO update

Feather River Snorkel Data from 2004 - 2020. This data was provided to flowwest in an access database  Snorkel_Revided.mdb. We queried it to have all year and important variables. Snorkel data from 1999 - 2003 is available in a separate access database FR S and S Oroville.mdb

**Timeframe:** 2004 - 2020TODO update


**Completeness of Record throughout timeframe:** 

Only SR collected are in 2011 and 2012

No environmental data collected pre 2010.


No data for: 

* 2006
* 2008
* 2009
* 2014


**Sampling Location:** Feather River

**Data Contact:** [Casey Campos](mailto:Casey.Campos@water.ca.gov)

Any additional info?

## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running 
# getwd() to see how to specify paths 
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = 
                 "juvenile-rearing-monitoring/seine-and-snorkel-data/feather-river/data-raw/db_from_feather_river/FR_S_and_S_Oroville.mdb",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "data-raw/qc-markdowns/seine-snorkel-data/feather-river/feather-river-db.mdb",
               overwrite = TRUE)

```

Read in data from google cloud, glimpse raw data: 
```{r, warning=FALSE}
db_filepath <- here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "feather-river-db.mdb")

library(Hmisc)
mdb.get(db_filepath, tables = TRUE) 
all_species <- mdb.get(db_filepath, "All Species Query Table") # what is this table?
snorkel_obsv <- mdb.get(db_filepath, "SnorkObservationsTBL") 
snorkel_survey_metadata <- mdb.get(db_filepath, "SnorkSurveyTBL")
lookup_HUC_cover <- mdb.get(db_filepath, "HUCcoverLU")
lookup_HUC_o_cover <- mdb.get(db_filepath, "HUCOcoverLU")
lookup_HUC_substrate <- mdb.get(db_filepath, "HUCsubstrateLU")
lookup_HUC_unit <- mdb.get(db_filepath, "HUCunitLU")
lookup_weather <- mdb.get(db_filepath, "WeatherLU")
detach(package:Hmisc)

# write to csvs
write_csv(snorkel_obsv, here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather.csv"))
write_csv(snorkel_survey_metadata, here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather_metadata.csv"))

# read in data
snorkel_raw <- read_csv(here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather.csv"))
snorkel_metadata_raw <- read_csv(here::here("data-raw", "qc-markdowns", "seine-snorkel-data", "feather-river", "raw_pre_2004_snorkel_data_feather_metadata.csv"))

```

## Data transformations
`All Species Query Table` only spans `1998-1999`.

Read in individual observation files and survey metadata files. Update column names and column types. Filter to only show chinook salmon.


```{r}
cleaner_snorkel_data <- snorkel_raw |> 
  janitor::clean_names() |> 
  mutate(huc_ocover = as.numeric(huc_ocover)) |> 
  filter(species %in% c("CHN", "None", "C", "CHNF", "CHNS", "CHNL")) |>  # filter species to relevant values (none is helpful to show they snorkeled and did not see anything)
  left_join(lookup_HUC_unit |> 
              rename(huc_unit = Unit), by = c("hu_cunit" = "UnitCode")) |>
  left_join(lookup_HUC_cover |> 
            rename(huc_cover = Cover), by = c("huc_icover" = "CoverCode")) |>
  left_join(lookup_HUC_substrate |> 
            rename(huc_substrate = Substrate), by = c("hu_csubstrate" = "SubstrateCode")) |>
  left_join(lookup_HUC_o_cover |> 
            rename(huc_o_cover = Cover), by = c("huc_ocover" = "CoverCode")) |>
  select(-c(hu_csubstrate, hu_cunit, huc_icover, huc_ocover, snorkler, obs_id)) |> 
    mutate(run = case_when(species  == "CHNF" ~ "fall",
                         species == "CHNS" ~ "spring",
                         species == "CHNL" ~ "late fall",
                         species %in% c("CHN",  "None", "C") ~ "unknown"),
           huc_unit = str_to_lower(huc_unit),
           huc_cover = str_to_lower(huc_cover),
           huc_o_cover = str_to_lower(huc_o_cover),
           huc_substrate = str_to_lower(huc_substrate)) |> 
  glimpse()

cleaner_snorkel_metadata <- snorkel_metadata_raw |> 
  janitor::clean_names() |> 
  left_join(lookup_weather, by = c("weather" = "WeatherCode")) |> 
  select(-c(visibility_comments, x_of_divers, x_of_center_passes, pass_width, comments,
          temp_time, snorkel_time_start, snorkel_time_stop, weather,
          snorkel_crew, shore_crew, recorder)) |> # doesn't seem like time information is being read in from mdb.get - TODO 
  mutate(location = str_to_title(location),
         survey_type = str_to_lower(survey_type),
         section_type = str_to_lower(section_type),
         weather = str_to_lower(Weather)) |> 
  select(-c(Weather)) |> 
  rename(river_velocity = river_flow) |> 
  glimpse()

snorkel <- left_join(cleaner_snorkel_data, cleaner_snorkel_metadata, by = "survey_id") |> 
  # remove units (repetitive) and comments, rename some variables for clarity
  select(survey_id, date, location, unit, survey_type, section_type, count = number, species, run,
         fork_length = fl, max_fork_length = max_fl, fish_depth, river_depth = depth,
         velocity, adj_velocity, river_velocity, bank_distance, huc_unit, huc_cover,
         huc_o_cover, huc_substrate, visibility, temperature, weather) |> 
  glimpse()

```

## Explore Numeric Variables: {.tabset}

```{r}
snorkel |> select_if(is.numeric) |> colnames()
```

### Variable: `flow`

**Plotting flow over Period of Record**

```{r}
cleaner_snorkel_data |> 
  ggplot(aes(x = date, y = flow)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  facet_wrap(~year(date), scales = "free_x") + 
  theme_minimal()
```

Very inconsistent and sparse flow measures. 

```{r}
cleaner_snorkel_data |> 
  ggplot(aes(x = flow)) +
  geom_histogram() + 
  theme_minimal()
```

Flow is between 0 - 8000. 0 values must be errors or NA.  

**Numeric Summary of flow over Period of Record**

```{r}
summary(cleaner_snorkel_data$flow)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$flow))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `flow` column are NA. 


### Variable: `turbidity`

**Plotting turbidity over Period of Record**

```{r}
cleaner_snorkel_data |> 
  ggplot(aes(x = date, y = turbidity)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  facet_wrap(~year(date), scales = "free") + 
  theme_minimal()
```

We only have non zero turbidity values for 2011, 2012, 2017, 2018, and 2019

```{r}
cleaner_snorkel_data |>
  ggplot(aes(x = turbidity)) + 
  geom_histogram() + 
  theme_minimal()
```

Most turbidity measures are 0. All fall between 0 and 4.5. 

**Numeric Summary of turbidity over Period of Record**

```{r}
summary(cleaner_snorkel_data$turbidity)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$turbidity))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `turbidity` column are NA.

### Variable: `temperature`

TODO: Check that they are in both F and C and then divide appropriately  

**Plotting temperature over Period of Record**

```{r}
cleaner_snorkel_data |> 
  ggplot(aes(x = date, y = temperature)) + 
  geom_line(linetype = "dashed") + 
  geom_point() +
  facet_wrap(~year(date), scales = "free") + 
  theme_minimal()
```

No temp measures before 2010, consistency of temperature measures varies throughout the years. 

```{r}
cleaner_snorkel_data |> 
  ggplot(aes(x = temperature)) +
  geom_histogram() +
  theme_minimal()
```

Interesting temperature distribution. I wonder if some measures are in C and some in F? 0 values are likely also a mistake or NA.

**Numeric Summary of temperature over Period of Record**

```{r}
summary(cleaner_snorkel_data$temperature)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$temperature))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `temperature` column are NA.



### Variable: `count`

**Plotting count over Period of Record**

```{r, include=FALSE}
sac_indices <- waterYearType::water_year_indices |> 
    filter(location == "Sacramento Valley") |> 
    transmute(water_year = WY, year_type = as.character(Yr_type))
```

```{r}
cleaner_snorkel_data |> 
  filter(run == "spring") |>
  group_by(date) |>
  summarise(total_daily_catch = sum(count, na.rm = T)) |>
  # filter(year(date) > 2014, year(date) < 2021) |>
  mutate(water_year = ifelse(month(date) %in% 10:12, year(date) + 1, year(date))) |> 
  left_join(sac_indices) |>
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) |>
  ggplot(aes(x = fake_date, y = total_daily_catch, fill = year_type)) + 
  geom_col() + 
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1901-03-01"), as.Date("1901-10-01")), date_breaks = "1 month") + 
  theme_minimal() + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom") + 
  labs(title = "Total Daily Raw Fish Count",
       y = "Total daily catch",
       x = "Date")+ 
  facet_wrap(~water_year, scales = "free") + 
  scale_fill_manual(values = wesanderson::wes_palette("Moonrise2", 5, type = "continuous"))
```


Very few SR fish, only caught SR in 2011 and 2012.  
```{r}
cleaner_snorkel_data  |>
  mutate(year = as.factor(year(date))) |>
  ggplot(aes(x = year, y = count)) + 
  geom_col() + 
  theme_minimal() +
  labs(title = "Total Fish Counted each Year",
       y = "Total fish count") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


**Numeric Summary of count over Period of Record**

```{r}
summary(cleaner_snorkel_data$count)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$count))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `count` column are NA.



### Variable: `est_size`

TODO figure out what size is describing. Is it length? or weight? 

**Plotting est_size over Period of Record**

```{r}
cleaner_snorkel_data |> 
  ggplot(aes(x = est_size)) + 
  geom_histogram() + 
  theme_minimal()
  
```

Estimated sizes fall between 0 and 1000. Again we need to figure out what type of size variable they are estimating 

**Numeric Summary of est_size over Period of Record**

```{r}
summary(cleaner_snorkel_data$est_size)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$est_size))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `est_size` column are NA.



### Variable: `water_depth_m`

Depth of water, units: m 

**Plotting water depth over Period of Record**

```{r}
cleaner_snorkel_data |> 
  ggplot(aes(x = water_depth_m)) + 
  geom_histogram() + 
  theme_minimal()
```
Depth of water is between 0 and 3 meters. 

**Numeric Summary of depth of water over Period of Record**

```{r}
summary(cleaner_snorkel_data$water_depth_m)
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$water_depth_m))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `water_depth_m` column are NA.




## Explore Categorical variables: {.tabset}


```{r}
cleaner_snorkel_data |> select_if(is.character) |> colnames()
```


### Variable: `survey_id`

Each row does not correspond to a unique survey. 

```{r}
length(unique(cleaner_snorkel_data$survey_id)) == length(cleaner_snorkel_data)
```

There are `r length(unique(cleaner_snorkel_data$survey_id))` unique survey ids

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$survey_id))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `survey_id` column are NA.

### Variable: `weather_code`

No lookup table for weather codes provided. However, they are fairly easy to interperate. Would be best if we could get weather codes standardized throughout and condense these down. 

```{r}
table(cleaner_snorkel_data$weather_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_snorkel_data$weather_code <- case_when(cleaner_snorkel_data$weather_code %in% c("CLD", "CLDY") ~ "cloudy", 
                                               cleaner_snorkel_data$weather_code %in% c("CLR (Hot)", "CLR/Hot", "Hot and CLR", "CLR Hot") ~ "clear and hot",
                                               cleaner_snorkel_data$weather_code %in% c("RAIN", "RAN", "CLD/RAIN", "LT RAIN", "CLD, Wind, Light Sprinkles") ~
                                                 "precipitation",
                                               cleaner_snorkel_data$weather_code %in% c("CLR 95", "CLR") ~ "clear",
                                               cleaner_snorkel_data$weather_code %in% c("PT. CLDY", "CLR/CLD") ~ "partly cloudy",
                                               cleaner_snorkel_data$weather_code %in% c("sun", "SUN") ~ "sunny",
                                               cleaner_snorkel_data$weather_code == c("CLR WINDY") ~ "clear and windy",
                                               cleaner_snorkel_data$weather_code == c("WND") ~ "windy",
                                               cleaner_snorkel_data$weather_code == c("LT CLD/HAZE") ~ "hazy")

table(cleaner_snorkel_data$weather_code) 
```

I would still like to simplify these down more but this is better for now. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$weather_code))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `weather_code` column are NA.

### Variable: `section_name`

Cleaned up section names, TODO need more information to decide if we can combine some of these section names. 

```{r}

format_site_name <- function(string) {
  clean <- 
    str_replace_all(string, "'", "") |>
    str_replace_all("G-95", "G95") |> 
    str_replace_all("[^[:alnum:]]", " ") |> 
    trimws() |> 
    stringr::str_squish() |>
    stringr::str_to_title()
}

cleaner_snorkel_data$section_name <- format_site_name(cleaner_snorkel_data$section_name)
table(cleaner_snorkel_data$section_name)


```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$section_name))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `section_name` column are NA.

### Variable: `units_covered`

A lot of units described here. If multiple units are covered they are listed. 

```{r}
unique(cleaner_snorkel_data$units_covered)[1:10]
```

There are `r length(unique(cleaner_snorkel_data$units_covered))` unique groups units covered. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$units_covered))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `units_covered` column are NA.

### Variable: `survey_comments` `observation_comments`
```{r}
unique(cleaner_snorkel_data$survey_comments)[1:5]
```

```{r}
unique(cleaner_snorkel_data$observation_comments)[1:5]
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$survey_comments))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `survey_comments` column are NA.

* `r round(sum(is.na(cleaner_snorkel_data$observation_comments))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `observation_comments` column are NA.

### Variable: `observation_id`

Each row does not correspond to a unique observation.  
```{r}
length(unique(cleaner_snorkel_data$observation_id)) == length(cleaner_snorkel_data)
```

There are `r length(unique(cleaner_snorkel_data$observation_id))` unique observation ids

* `r round(sum(is.na(cleaner_snorkel_data$observation_id))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `observation_id` column are NA.

### Variable: `unit`

```{r}
table(cleaner_snorkel_data$unit) 
```

There are `r length(unique(cleaner_snorkel_data$unit))` unique units covered. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$unit))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `unit` column are NA.

### Variable: `size_class`

| Size Class | Size Range |
| ---------- | ---------- |
| I |	0-50 |
| II |	51-75 |
| III |	76-100 |
| IV |	101-150 |
| V |	151-300 |
| VI |	301-499 |
| VII |	500+ |

```{r}
table(cleaner_snorkel_data$size_class) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$size_class))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `size_class` column are NA.

### Variable: `substrate`

| SubstrateCode |	Substrate |
| ------------- | --------- |
| 1 |	Organic Fines, Mud (0.05 mm) |
| 2 |	Sand (0.05 to 2 mm) |
| 3 |	Small Gravel (2 to 50 mm) |
| 4 |	Large Gravel (50 to 150 mm) |
| 5 |	Cobble (150 to 300 mm) |
| 6 |	Boulder (> 300 mm) |
| 0 | ? |

It appears that if more than one substrate is present they list all by creating multi-didget numbers. Not sure what 0 stands for? TODO 

```{r}
table(cleaner_snorkel_data$substrate) 
```

What is the best way to show multiple substrates at different sites? In the seine data they have many substrate columns but I do not think I like that better. 

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$substrate))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `substrate` column are NA.

### Variable: `instream_cover`

| ICoverCode |	ICover |
| ---------- | ------- |
| A |	No apparent cover |
| B |	Small instream objects/small-medium woody debris |
| C |	Large instream objects/large woody debris |
| D |	Overhead objects |
| E |	Submerged aquatic veg/filamentous algae |
| F |	Undercut bank |

Instream cover appears to use a similar method as substrate where they add letters on if multiple covers are present. 

```{r}
table(cleaner_snorkel_data$instream_cover) 
```

What is the best way to display/simplify these?

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$instream_cover))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `instream_cover` column are NA.


### Variable: `overhead_cover`

| OCoverCode |	OCover |
| --------- | -------- |
| 0 |	No Apparent Cover |
| 1 |	Overhanging veg/obj (< 0.5 m above surface) |
| 2 |	Overhanging veg/obj (0.5 to 2 m above surface |
| 3 |	Surface turbulence, bubble curtain |

Overhead cover appears to use a similar method as substrate and instream cover where they add letters on if multiple covers are present. 

```{r}
table(cleaner_snorkel_data$overhead_cover) 
```


**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$overhead_cover))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `overhead_cover` column are NA.



### Variable: `hydrology_code`

Looks like these are already defined for the most part, GM = Glide Edgewater so we can change that. 

```{r}
table(cleaner_snorkel_data$hydrology_code) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
cleaner_snorkel_data$hydrology_code <- ifelse(cleaner_snorkel_data$hydrology_code == "GM", "Glide Edgewater", cleaner_snorkel_data$hydrology_code)

table(cleaner_snorkel_data$hydrology_code) 
```

* `r round(sum(is.na(cleaner_snorkel_data$hydrology_code))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `hydrology_code` column are NA.



### Variable: `lwd_number`

I am not sure what this number is refering to, TODO

```{r}
table(cleaner_snorkel_data$lwd_number) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$lwd_number))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `lwd_number` column are NA.

### Variable: `run`

Chinook Run 

```{r}
table(cleaner_snorkel_data$run) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$run))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `run` column are NA.


Only 13 identified SR fish. This data may not be very helpful for us in the JPE


### Variable: `tagged`

If the fish is tagged or not

```{r}
table(cleaner_snorkel_data$tagged) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$tagged))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `tagged` column are NA.

### Variable: `clipped`

If the fish has a clipped fin or not

```{r}
table(cleaner_snorkel_data$clipped) 
```

**NA and Unknown Values**

* `r round(sum(is.na(cleaner_snorkel_data$clipped))/nrow(cleaner_snorkel_data), 3) * 100` % of values in the `clipped` column are NA.


## Summary of identified issues

* Snorkel surveys are not done every year in the timeframe 
* Environmental data very sparse, especially in early years
* Still a few unknown columns that we need to define and understand 
* Figure out the best way to display and standardize substrate and cover information
* Very few Spring Run Chinook (13)
* Detailed information on location split up into units and sections, TODO need to improve organization in these settings

## Save cleaned data back to google cloud 

```{r}
# Write to google cloud 
# Name file [watershed]_[data type].csv
```

