---
title: "butte-creek-rst-qc-checklist"
author: "Inigo Peng"
date: "10/19/2021"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
library (RColorBrewer)
```
# Butte Creek RST Data  

**Description of Monitoring Data**

This dataset contains data for all Chinook salmon that were captured in the Butte Creek rotary screw trap (RSTR) or diversion fyke trap (DSTR) between the 1995-96 and 2014-15 trapping seasons.

**Timeframe:** 

1995 - 2015

**Completeness of Record throughout timeframe:**

* Life stage information lacks after 2005
* Inconsistent completeness of physical data after 2008

**Sampling Location:** 

3 locations on Butte Creek

**Data Contact:** [Jessica Nichols](mailto:Jessica.Nichols@Wildlife.ca.gov)


## Access Cloud Data

```{r, eval=FALSE}
# Run Sys.setenv() to specify GCS_AUTH_FILE and GCS_DEFAULT_BUCKET before running
# Open object from google cloud storage
# Set your authentication using gcs_auth
gcs_auth(json_file = Sys.getenv("GCS_AUTH_FILE"))
# Set global bucket 
gcs_global_bucket(bucket = Sys.getenv("GCS_DEFAULT_BUCKET"))
gcs_list_objects()
# git data and save as xlsx
gcs_get_object(object_name = "rst/butte-creek/data-raw/CDFW_Butte_Creek_RST_Captures.xlsx",
               bucket = gcs_get_global_bucket(),
               saveToDisk = "butte_creek_rst_raw.xlsx",
               Overwrite = TRUE)
```

```{r}
raw_data = readxl::read_excel('butte_creek_rst_raw.xlsx',
                              col_types = c("date","text","text","text","text","text","numeric","numeric","numeric","text",
                                            "text","date","text","text","numeric","numeric","numeric","numeric","text","text",
                                            "numeric","numeric","text","numeric","numeric","text"
                                            ))
glimpse(raw_data)
```

## Data Transformations
```{r}
cleaner_data <- raw_data %>% 
  set_names(tolower(colnames(raw_data))) %>%
  select(-c("staffgauge")) %>% #no value associated with reading (e.g. cfs, ft) %>%  
  rename('date'= sampledate,
         'station' = stationcode,
         'method'= methodcode,
         'trap_status' = trapstatus,
         'species'= organismcode,
         'fork_length' = forklength,
         'lifestage' = stagecode,
         'time' = sampletime,
         'gear_id' = gearid,
         'temperature' = watertemperature,
         'velocity' = watervelocity,
         'trap_revolutions' = traprevolutions,
         'rpms_start' = rpmsstart,
         'rpms_end'= rpmsend) %>% 
  mutate(time = hms::as_hms(time),
         date = as.Date(date)) %>%
  glimpse()
```
## Explore `date`
```{r}
summary(cleaner_data$date)
```

**NA and Unknown Values**  

* `r round(sum(is.na(cleaner_data$date))/nrow(cleaner_data), 3)*100` % of values in the `date` column are NA.

## Explore Categorical Variables

```{r}
cleaner_data %>% select_if(is.character) %>% colnames()
```
### Variable `station`

**Description:** trap location

* BCADAMS - Adams Dam

* BCOKIE-1 - Okie Dam 1

* BCOKIE-2 - Okie Dam 2


```{r}
cleaner_data <- cleaner_data %>% 
  mutate(station = case_when(
    station == 'BCADAMS' ~ 'Adams Dam',
    station == 'BCOKIE-1'~ 'Okie Dam 1',
    station == 'BCOKIE-2' ~ 'Okie Dam 2',
    TRUE ~ as.character(station)
  ))

table(cleaner_data$station)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$station))/nrow(cleaner_data), 3)*100` % of values in the `station` column are NA.


### Variable `method`

**Description:** method of capture

* DSTR - Diversion fyke trap

* RSTR - Rotary screw trap

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(method = set_names(tolower(method)),
         method = case_when(
          method =='dstr'~ 'diversion fyke trap',
          method =='rstr'~ 'rotary screw trap'
  )
)
table(cleaner_data$method)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$method))/nrow(cleaner_data), 3)*100` % of values in the `method` column are NA.


### Variable `trap_status`

**Description:** 

* Check - trap was checked normally , continued fishing

* Pull - trap was pulled after trap check

* Set - trap was set upon arrival 
```{r}
cleaner_data <- cleaner_data %>% 
  mutate(trap_status = set_names(tolower(trap_status)))

table(cleaner_data$trap_status)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$trap_status))/nrow(cleaner_data), 3)*100` % of values in the `trap_status` column are NA.


### Variable `species`

Species contain Chinooks Only

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$species))/nrow(cleaner_data), 3)*100` % of values in the `species` column are NA.

### Variable `dead`

Fix capitalization. n/p is changed to NA.

```{r}
table(cleaner_data$dead)
```
```{r}
cleaner_data$dead <- case_when(tolower(cleaner_data$dead) == "yes" ~ TRUE, 
                                    tolower(cleaner_data$dead) == "no" ~ FALSE)
table(cleaner_data$dead)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$dead))/nrow(cleaner_data), 3)*100` % of values in the `dead` column are NA.

### Variable `markcode`

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(markcode = tolower(cleaner_data$markcode),
         markcode = ifelse(markcode == "n/p", NA, markcode))
table(cleaner_data$markcode)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$markcode))/nrow(cleaner_data), 3)*100` % of values in the `markcode` column are NA.

### Variable `lifestage`


* 1 - Fry with visible yolk sac
* 2 - Fry with no visible yolk sac
* 3 - Parr
* 4 - Fingerling
* 5 - Smolt
* AD - Adult
* n/p - not provided 
* UNK - unknown

```{r}
table(cleaner_data$lifestage)
```
n/p is converted to NA

```{r}

cleaner_data <- cleaner_data %>% 
  mutate(lifestage = case_when(
    lifestage == 1~ 'yolk sac fry',
    lifestage == 2~ 'fry',
    lifestage == 3~ 'parr',
    lifestage == 4~ 'fingerling',
    lifestage == 5~ 'smolt',
    lifestage == 'AD'~ 'adult',
    lifestage == 'UNK'~ 'unknown'
  ))
table(cleaner_data$lifestage)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$lifestage))/nrow(cleaner_data), 3)*100` % of values in the `lifestage` column are NA.

### Variable `gear_id`

**Description:**

* DSTR1 - Diversion Fyke Trap 1

* RSTR1 - Rotary Screw Trap 1

* RSTR2 - Rotary Screw Trap 2

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(gear_id = case_when(
    gear_id == 'DSTR1' ~ 'diversion fyke trap 1',
    gear_id == 'RSTR1' ~ 'rotary screw trap 1',
    gear_id == 'RSTR2' ~ 'rotary screw trap 2'
  ))

table(cleaner_data$gear_id)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$gear_id))/nrow(cleaner_data), 3)*100` % of values in the `gear_id` column are NA.

### Variable `weathercode`

**Description:**

* CLD – Cloudy 
* CLR – Clear 
* FOG – Foggy 
* RAN – Rainy 

```{r}
cleaner_data <- cleaner_data %>% 
  mutate(weathercode = case_when(
    weathercode == "CLD" ~ "cloudy",
    weathercode == "CLR" ~ "clear",
    weathercode == "FOG" ~ "foggy",
    weathercode == "RAN" ~ "rainy"
  ))
table(cleaner_data$weathercode)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$weathercode))/nrow(cleaner_data), 3)*100` % of values in the `weathercode` column are NA.

### Variable `northbrush`, `southbrush`

**Description:** for diversion fyke trap, there are diversion screens for clearing debris on the north and south of the trap. 

False - screen not working (not clearing debris)
True - screen work (clearing debris)

```{r}
#Not sure if we are interested in diversion fyke trap - keeping the variable for now
table(cleaner_data$northbrush)
```
```{r}
table(cleaner_data$southbrush)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$northbrush))/nrow(cleaner_data), 3)*100` % of values in the `northbrush` column are NA.


*  `r round(sum(is.na(cleaner_data$southbrush))/nrow(cleaner_data), 3)*100` % of values in the `southbrush` column are NA.

### Variable `debris`

**Description:** visual assessment of debris in trap

```{r}

cleaner_data <- cleaner_data %>% 
  mutate(debris = set_names(tolower(debris)))
table(cleaner_data$debris)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$debris))/nrow(cleaner_data), 3)*100` % of values in the `debris` column are NA.

### Variable `comments`
```{r}
unique(cleaner_data$comments)[1:5]
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$comments))/nrow(cleaner_data), 3)*100` % of values in the `comments` column are NA.


## Explore Numerical Variables

```{r}
cleaner_data %>% select_if(is.numeric) %>% colnames()
```
### Variable `count`

```{r fig.height = 8, fig.width = 8}
cleaner_data %>% 
  mutate(water_year = ifelse(month(date) %in% 10:12, year(date) + 1, year(date))) %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900, 1901),
         fake_date = as.Date(paste0(fake_year,"-", month(date), "-", day(date)))) %>%
  group_by(date) %>% 
  mutate(total_daily_catch = sum(count)) %>% 
  ungroup() %>% 
  ggplot(aes(x = fake_date, y = total_daily_catch)) +
  geom_col()+
  # scale_x_date(labels = date_format("%b"), limits = c(as.Date("1995-10-01"), as.Date("2016-06-01")), date_breaks = "1 month")+
  scale_x_date(labels = date_format("%b"), limits = c(as.Date("1900-10-01"), as.Date("1901-06-01")), date_breaks = "1 month") + 
  theme_minimal()+
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90))+
  labs(title = "Total Daily Raw Passage 1995 - 2015",
       y = "Total daily catch",
       x = "Date")+ 
  facet_wrap(~water_year, scales = "free")
```

```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = year, y = count))+
  geom_col()+
  theme_minimal()+
  labs(title = "Total Fish Count By Year")+
  theme(text = element_text(size = 10),
        axis.text.x = element_text(angle = 90,  vjust = 0.5, hjust=1))
```

**Numeric Summary of count over Period of Record**

```{r}
summary(cleaner_data$count)
```

These numbers are massive. Are they passage estimates? Noted in the bottom.

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$count))/nrow(cleaner_data), 3)*100` % of values in the `count` column are NA.

### Variable `fork_length`

**Description:** fork length in millimeters (mm)

Need to filter out 0s
```{r}
cleaner_data %>% filter(fork_length == 0)
```
```{r}
cleaner_data$fork_length <- ifelse(cleaner_data$fork_length == 0, NA, cleaner_data$fork_length)
```

```{r}
cleaner_data %>% 
  filter(fork_length < 250) %>% #filtered out 52 points to see more clear distribution
  ggplot(aes(x = fork_length))+
  geom_histogram(binwidth = 2)+
  theme_minimal()+
  scale_x_continuous(breaks = seq(0, 250, by=25))+
  labs(title = "Fork Length Distribution")+
  theme(text = element_text(size=15),
        axis.text.x = element_text(vjust =0.5, hjust = 1))
```
```{r}
cleaner_data %>% 
  ggplot(aes(x = fork_length, y = lifestage))+
  geom_boxplot()+
  theme_minimal()+
  labs(title = 'Fork length summarized by life stage')+
  theme(text = element_text(size = 12))
```

**Numeric Summary of fork_length over Period of Record**

```{r}
summary(cleaner_data$fork_length)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$fork_length))/nrow(cleaner_data), 3)*100` % of values in the `fork_length` column are NA.


### Variable `weight`

**Description:** wet weight in grams(g)

Need to filter out 0s
```{r}
cleaner_data %>% filter(weight == 0)
```
```{r}
cleaner_data$fork_length <- ifelse(cleaner_data$weight == 0, NA, cleaner_data$weight)
```

```{r}
cleaner_data %>% 
  filter(weight< 30) %>%  #filtered out 26 data points to see more clear distribution
  ggplot(aes(x = weight))+
  geom_histogram(binwidth = 1)+
  scale_x_continuous(breaks = seq(0, 30, by=2))+
  theme_minimal()+
  labs(title = "Weight Distribution")
```
```{r}
cleaner_data %>% 
  filter(weight < 50) %>% 
  ggplot(aes(x = weight, y= lifestage))+
  geom_boxplot()+
  labs(title = 'Weight summarized by life stage')+
  theme(text = element_text(size = 12))+
  theme_minimal()
```

**Numeric Summary of weight over Period of Record**

```{r}
summary(cleaner_data$weight)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$weight))/nrow(cleaner_data), 3)*100` % of values in the `weight` column are NA.

### Variable `temperature`

**Description:** temperature of water in degrees Celsius

```{r}
cleaner_data %>%
  filter(temperature < 100) %>% #filter out 36 points with water temperature > 100 degrees
  ggplot(aes(x= temperature, y = station))+
  geom_boxplot()+
  theme_minimal()+
  labs(title = "Water Temperature by Station")
```
```{r}
cleaner_data %>% 
  filter(temperature < 100) %>%  #filter out 36 points with water temperature > 100 degrees (entry error?)
  group_by(date) %>% 
  mutate(daily_avg_temp = mean(temperature)) %>% 
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900,1901),
         fake_date = as.Date(paste0(fake_year, "-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = daily_avg_temp, color = year))+
  geom_point()+
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal()+
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "none")+
  labs(title = "Daily Water Temperature (colored by year)",
       x = 'Date',
       y = 'Average Daily Temp')
```
```{r}
cleaner_data %>% 
  filter(temperature < 100) %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = temperature, y = year))+
  geom_boxplot()+
  theme_minimal()+
  labs(title = "Water Temperature summarized by year")+
  theme(text = element_text(size = 15),
        axis.text.x = element_text(vjust =0.5, hjust = 1))
```


**Numeric Summary of temperature over Period of Record**

```{r}
summary(cleaner_data$temperature)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$temperature))/nrow(cleaner_data), 3)*100` % of values in the `temperature` column are NA.

### Variable `turbidity`

**Description:** Turbidity of water in NTU
```{r}
cleaner_data %>% 
  group_by(date) %>% 
  mutate(daily_avg_turb = mean(turbidity)) %>% 
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900,1901),
         fake_date = as.Date(paste0(fake_year, "-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = daily_avg_turb, color = year))+
  geom_point()+
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal()+
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "none")+
  labs(title = "Daily Turbidity (colored by year)",
       x = 'Date',
       y = 'Average Daily Turbidity')
```
```{r}
cleaner_data %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = turbidity, y = year))+
  geom_boxplot()+
  theme_minimal()+
  labs(title = "Turbidity summarized by year")+
  theme(text = element_text(size = 15),
        axis.text.x = element_text(vjust =0.5, hjust = 1))
```

**Numeric Summary of turbidity over Period of Record**
```{r}
summary(cleaner_data$turbidity)
```
**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$turbidity))/nrow(cleaner_data), 3)*100` % of values in the `turbidity` column are NA.

### Variable `velocity`

**Description:** water velocity measured in ft/s

Data Transformation
```{r}
#Convert water velocity from ft/s to m/s
cleaner_data <- cleaner_data %>% 
  mutate(velocity = velocity/3.281)
```
```{r}
cleaner_data %>%
  filter(velocity < 8) %>% #filtered out 8 data points to show a more clear graph
  ggplot(aes(x= velocity, y = station))+
  geom_boxplot()+
  theme_minimal()+
  labs(title = "Water Velocity by Station",
       x= "Velocity (m/s)")
```
```{r}
cleaner_data %>% 
  filter(velocity < 5) %>% #filtered out one point to show a more clear graph
  group_by(date) %>% 
  mutate(daily_avg_velocity = mean(velocity)) %>% 
  ungroup() %>% 
  mutate(year = as.factor(year(date)),
         fake_year = if_else(month(date) %in% 10:12, 1900,1901),
         fake_date = as.Date(paste0(fake_year, "-", month(date), "-", day(date)))) %>% 
  ggplot(aes(x = fake_date, y = daily_avg_velocity, color = year))+
  geom_point()+
  scale_x_date(labels = date_format("%b"), date_breaks = "1 month")+
  theme_minimal()+
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "none")+
  labs(title = "Daily Water Velocity (colored by year)",
       x = 'Date',
       y = 'Average Daily Velocity')
```
```{r}
cleaner_data %>% 
  filter(velocity<6) %>% 
  mutate(year = as.factor(year(date))) %>% 
  ggplot(aes(x = velocity, y = year))+
  geom_boxplot()+
  theme_minimal()+
  labs(title = "Water Velocity summarized by year")+
  theme(text = element_text(size = 15),
        axis.text.x = element_text(vjust =0.5, hjust = 1))
```


**Numeric summary of velocity over Period of Record**
```{r}
summary(cleaner_data$velocity)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$velocity))/nrow(cleaner_data), 3)*100` % of values in the `velocity` column are NA.

### Variable `trap_revolutions`

**Description:** Number of revolutions the RST cone had made since last being checked
```{r}
cleaner_data %>% 
  ggplot(aes(x = trap_revolutions))+
  geom_histogram(binwidth = 500)+
  labs(title = "Distribution of Trap Revolutions")+
  theme_minimal()+
  theme(text = element_text(size = 12))
```
```{r}
cleaner_data %>% 
  filter(station != "Adams Dam") %>% 
  ggplot(aes(y= station, x = trap_revolutions))+
  geom_boxplot()+
  labs(title = "Trap Revolutions Summarized by Location")+
  theme_minimal()+
  theme(text = element_text(size = 12))
```


**Numeric summary of trap_revolutions over Period of Record**
```{r}
summary(cleaner_data$trap_revolutions)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$trap_revolutions))/nrow(cleaner_data), 3)*100` % of values in the `trap_revolutions` column are NA.

### Variable `rpms_start`

**Description:** rotations per minute of RST cone at start of trapping window
```{r}
cleaner_data %>% 
  filter(rpms_start < 10) %>% #filtered out 28 data points to show more clear distribution
  ggplot(aes(x = rpms_start))+
  geom_histogram(binwidth = 1)+
  labs(title = "Distribution of RPMs Start")+
  theme_minimal()+
  theme(text = element_text(size = 12))
```
```{r}
cleaner_data %>% 
  filter(rpms_start < 10) %>% 
  ggplot(aes(y= station, x = rpms_start))+
  geom_boxplot()+
  labs(title = "RPMs Start Summarized by Location")+
  theme_minimal()+
  theme(text = element_text(size = 12))
```

**Numeric summary of rpms_start over Period of Record**
```{r}
summary(cleaner_data$rpms_start)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$rpms_start))/nrow(cleaner_data), 3)*100` % of values in the `rpms_start`` column are NA.

### Variable `rpms_end`

**Description:** rotations per minute of RST cone at end of trapping window
```{r}
cleaner_data %>% 
  filter(rpms_end < 10) %>% #filtered out 28 data points to show more clear distribution
  ggplot(aes(x = rpms_end))+
  geom_histogram(binwidth = 1)+
  labs(title = "Distribution of RPMs End")+
  theme_minimal()+
  theme(text = element_text(size = 12))
```

```{r}
cleaner_data %>% 
  filter(rpms_end < 10) %>% 
  ggplot(aes(y= station, x = rpms_end))+
  geom_boxplot()+
  labs(title = "RPMs End Summarized by Location")+
  theme_minimal()+
  theme(text = element_text(size = 12))
```

**Numeric summary of rpms_end over Period of Record**
```{r}
summary(cleaner_data$rpms_end)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$rpms_end))/nrow(cleaner_data), 3)*100` % of values in the `rpms_end` column are NA.


### Variable `secchi`

**Description:** Secchi depth in feet

```{r}
cleaner_data %>% 
  ggplot(aes(x=secchi))+
  geom_histogram()+
  theme_minimal()+
  labs(title = "Distribution of Secchi")
```

**Numeric summary of secchi over Period of Record**
```{r}
summary(cleaner_data$secchi)
```

**NA and Unknown Values**  

*  `r round(sum(is.na(cleaner_data$secchi))/nrow(cleaner_data), 3)*100` % of values in the `secchi` column are NA.

**Issues Identified**

* 50 points in water temperature reaches over 50 degrees celsius

* Turbidity data lacks in some years

* Fish count number is extremely high (up to 220,000 fish in one day)? 

### Add cleaned data back into google cloud
```{r}
butte_creek_rst <- cleaner_data %>% glimpse()
```
```{r}
write_csv(butte_creek_rst, "butte_rst.csv")
```


```{r, eval= FALSE}

f <- function(input, output) write_csv(input, file = output)

gcs_upload(butte_creek_rst,
           object_function = f,
           type = "csv",
           name = "rst/butte-creek/data/butte-creek-rst.csv")
```